### 2013-01-23 tex/article.tex
Result tables and secondary bibliography

% ---------------------------------------------------------------------
% -------------- PREAMBLE ---------------------------------------------
% ---------------------------------------------------------------------

%\documentclass[12pt,a4paper,finnish,oneside]{article}
\documentclass[lnbip]{svmultln}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{calc}      % käytetään laskurien (counter) yhteydessä (tiedot.tex)
\usepackage{natbib}
\usepackage{url}
\usepackage{listings}
\usepackage{hyphenat}

\usepackage{supertabular,array}
\usepackage{lipsum}
\usepackage{booktabs}
\usepackage{graphicx}

% Strikethrough
\usepackage[normalem]{ulem}

% Punctuation for references
\bibpunct{[}{]}{;}{n}{,}{,}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{multibib}
\newcites{gen}{References}
\usepackage{bibentry}

% ---------------------------------------------------------------------
% -------------- DOCUMENT ---------------------------------------------
% ---------------------------------------------------------------------

\begin{document}

\selectlanguage{english}

\mainmatter

\title{Large scale agile transformations: A literature review}

\titlerunning{Large scale agile transformations}

\author{Kim Dikert \and Maria Paasivaara \and Casper Lassenius}
\authorrunning{Kim Dikert et al.}   % abbreviated author list (for running head)

% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Kim Dikert, Maria Paasivaara, Casper Lassenius}

\institute{Aalto University . . . Address . . .
\\ \email{email@xx},
\\ WWW home page: \texttt{http://www...} }

\maketitle

% ---------------------------------------------------------------------
% -------------- ABSTRACT AND TEXT ------------------------------------
% ---------------------------------------------------------------------

\begin{abstract}        % give a summary of your paper
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words. It will be set in 9-point
font size and be inset 1.0 cm from the right and left margins.
There will be two blank lines before and after the Abstract.
%                         please supply keywords within your abstract
\keywords {agile, transformation, large scale}
\end{abstract}

% ---------------------------------------------------------------------
\section{Introduction}

As the competition in software industry is growing companies are constantly
looking to improve their effectiveness. Agile methods are claimed to increase
productivity and quality \citegen{Livermore2008}, which makes them attractive for
companies pursuing better performance. However, large organizations may have
problems introducing agile methods \citegen{Dyba2009}. As agile methods are
initially designed for small teams their application has proven to be
problematic in a larger scale \citegen{Boehm2005}.

It is typical for large companies to function according to traditional software
engineering models [Cite?!] (also why?). These models strive for consistent quality and performance
by rigorous planning and definition of process. This kind of approach is however
badly suited for software development, as projects typically encounter
situations that are too hard or impossible to foretell \citegen{Schwaber2002}.
The main problems in plan-driven development are the high cost of changes and
late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
resopnding to change and distance from customers undermine the  competitiveness
of companies. Agile methods are believed to bring remedy to these problems.

The goal of this research is to explore how and why large organizations adopt
agile methods. Studies on large scale agile transformations have been published,
but no extensive and summarizing research is available on the subject. In this
research we will give an overview of reported case studies on adopting agile
methods in large organizations. We will further discuss the motivations for
adopting agile methods, the characteristics of agile transformations, and
factors relating to success or challenge in the transformation process. Our
research is focused on the process of organizational change. We leave out topics
related to comparing the initial and transformed states of organisations, such
as measuring the performance of organizations before and after the
transformation.

This paper is organized as follows. The next section gives an overview of large
scale agile development. Section \ref{sec:method} presents the literature review
method used for conducting the research. The findings of the review are
discussed in section \ref{sec:results}, and section \ref{sec:conclusions}
discusses the findings and gives proposals for future research.

% ---------------------------------------------------------------------
\section{Background}
\label{sec:background}

Introducing agile methods in large organizations is more difficult than it is in
small organizations \citegen{Dyba2008}. The difficulty is partly related to size
bringing higher organizational inertia which slows down organizational change
\citegen{Livermore2008}. Agile methods are not founded on the use of individual
tools or practices, but rather on a holistic change in the way of thinking.
Adopting agile requires often change of the entire organizational culture
\citegen{Misra2010}.

One significant difference between small and large adoptions is that large
organizations have more dependencies between projects and teams. This increases
the need for formal documentation and thus reduces agility \citegen{Lindvall2004}.
In addition to inter-team coordination large organizations require development
teams to interact with other organizational units, which are often non-agile of
their nature. For instance, HR may demand individuals to have strictly specified
roles in projects \citegen{Boehm2005}, or a change control board may inhibit use of
continuous integration or refactoring \citegen{Lindvall2004}. All units affected by
the agile transformation need to be informed and heard, and the agile process
must be adjusted according to their needs \citegen{Lindvall2004, Cohn2003,
Boehm2005}.

Agile methods will also affect management and business related functions. A key
challenge is that management must move away from life cycle models and towards
iterative and feature centric models \citegen{Nerur2005}. The focus must be shifted
from a large scale scope to shorter term project planning \citegen{Misra2010}, as
agile methods emphasize that meaningful planning can only be done for the near
future \citegen{Cohn2003}. However, the lack of planning can be a concern as
business and customer relationships often build on long term roadmapping.
Enabling operation with shorter term planning requires educating stakeholders
and reviewing contracting practices \citegen{Boehm2005}.

% ---------------------------------------------------------------------
\section{Research method}
\label{sec:method}

% Research questions:
%
% What are the motivations for starting an organizational transformation?
%
% What kind of agile transformations have been reported?
%
% What were challenges and success factors in the transformation process?

The literature review was conducted as an application of Kitchenham's
method for systematic literature review (SLR) \citegen{Kitchenham2007}. The method
was applied to be performed by only one researcher, and to serve as a
preliminary work evaluating the feasibility of a more in-depth review. The
required effort was limited by selecting only the following subset of SLR steps:
identification of primary sources, data extraction and synthesis. Steps left out
were the development of a systematic review protocol, strict study quality
assessment, and reviews in the study selection process.

\subsection{Data sources and search strategy}

The searches for primary sources was conducted in two steps. First a preliminary
search was made to sample the existing literature. Then a systematic search was
done based on the preliminary search. The following databases were included in
the searches: IEEExplore\footnote{http://ieeexplore.ieee.org/},
ACM\footnote{http://dl.acm.org/}, Scopus\footnote{http://www.scopus.com/},
ProQuest\footnote{http://search.proquest.com/}.
Additionally the archive of International XP
Conference\footnote{http://link.springer.com/} was searched.

The search phrases used in the preliminary search were \textit{agile
transformation} and \textit{large scale agile}. For the systematic search three
facets and appropriate keywords were chosen, as presented in Table
\ref{table:searchterms}. To search the databases a query phrase was constructed
by joining the keywords of each facet by \texttt{OR} operators and the facets by
\texttt{AND} operators.

As the preliminary search omitted some of the keywords in Table
\ref{table:searchterms} it yielded a small number of additional matches. The
reason for variance in matches was that many articles had creative titles or the
abstracts and keywords were not informative enough. Because of this, the search
with more restrictive keywords was unable to capture all possible matches.
We are however confident that the search covered the majority of material
relating to agile transformations in large organizations. If a wider search
should be performed, our recommendation is to only regard the facets
\textit{agile methods} and \textit{organizational change}. This would lead to
significantly more matches in the first stage of selecting studies, but the
additional effort is necessary due to the limitations of a keyword search.

\begin{table}[h]
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Facet                  & Keywords   \\ \midrule
        Agile methods          & agile, scrum, lean, xp \\ 
        Organizational change  & transformation, transition, change, migration \\
        Large organization     & enterprise, organization, (large \texttt{AND} scale) \\
        \bottomrule
    \end{tabular}
    \caption{Facets and related search terms used}
    \label{table:searchterms}
\end{table}

The selection process of the primary studies is summarized in Figure
\ref{fig:selection_process}. The database search yielded 1603 matches from which
100 sources were chosen by the title. 17 unmatched sources from the preliminary
search were also included. Finally, 31 sources were selected based on the
abstract to be included for data extraction. One source was not available, which
reduced the final number of primary studies to 30.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{researchprocess}
    \caption{Process of selecting primary studies}
    \label{fig:selection_process}
  \end{center}
\end{figure}

\subsection{Inclusion criteria}

The 30 primary sources were selected by four criteria: study type, size of
organization, use of agile methods in software development, and viewpoint on
transformation. Firstly, only case studies and experience reports on agile
software development were included in this study. Papers discussing agile
methods or organizational transformation on as a general topic were left out.
Secondly, the size of the organization described in the included primary studies
was to be large enough to show the challenges of large scale agile development
\citegen{Lindvall2004}. Further, only studies discussing agile software development
were included, leaving out for instance studies on agile management. Finally, to
be included each study had to provide a viewpoint on the organizational
transformation.

\subsection{Data extraction}

The content of the resutls was guided by the data extraction form. The form
included some quantitative fields including \textit{start time of
transformation} and \textit{organization size}, but most of the data was
collected in open ended fields. The reason for starting a transformation was
captured in fileds \textit{initial state of organization} and \textit{reasons
for change}. The description of the transformation process was captured in one
field. Recommendations for transformation processes were captured in fields
\textit{reported good practices}, \textit{reported challenges}, and
\textit{lessons learned}. Additional fields that were used but not sufficiently
answered by the primary studies were \textit{satisfaction after transformation},
\textit{effect on onrganization} and \textit{reported measurements before/after
transformation}.

% \begin{table}
%     \begin{tabular}{ p{0.36\textwidth}@{ \hskip 0.2cm } l }
%         \toprule
%         Transformation mentioned in \newline text (Y/N) &
%         Large scale mentioned in text (Y/N) \\
%         Is empirical case study (Y/N) &
%         Is industry experience report (Y/N) \\
%         Has listing of practices (Y/N) &
%         Used research method (Y/N) \\
%         Relevance to this review (1-5) & \\
%         Objective of research (or publication) &
%         Research method \\
%         Author bias &
%         Validity threats \\
%         Organization size &
%         Time of transformation \\
%         Initial state of organization &
%         Why was the change initiated \\
%         How was the change conducted &
%         What is agility? / Which agile paractices are used? \\
%         Findings / lessons learned &
%         Good practices validated or suggested by study \\
%         Reported challenges &
%         Satisfaction after transformation \\
%         Effect on organization &
%         Measurements as results (quantitative or other) \\
%         Other notes &
%         Notable references \\
%         \bottomrule
%     \end{tabular}
%     \caption{The fields of the data extraction form. Fields with a (Y/N) label
%     were answered as yes or no. The field with a (1-5) label was answered by a
%     grade from 1 to 5. All other fields were answered by free form text.}
%     \label{table:dataform}
% \end{table}

%\subsection{Data sources and search strategy}
%
%\subsection{Final selection}
%
%\subsection{Data extraction and synthesis}

% ---------------------------------------------------------------------
\section{Results}
\label{sec:results}

Of the 30 included primary studies 23 were experience reports and 7 were case
studies. From a validity viewpoint it is important to acknowledge that most of
the studies were written by employees in the studied organizations.

All of the included studies had been published within a decade. On average the
studies were published theree years after the transformation was reported to
have begun. Figure \ref{fig:publications} gives a timeline view of the
publications.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=1\textwidth]{publicationschart.pdf}
    \caption{Time of publications and transformations}
    \label{fig:publications}
  \end{center}
\end{figure}

\subsection{Overview of agile methods used}

\begin{table}[h]
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Agile method    & Reported in number of studies   \\ \midrule
        Scrum           & 18 \\ 
        XP              & 7 \\
        Lean            & 5 \\
        Tailoring and combining methods & 9  (indirect mentionings not counted) \\
        \bottomrule
    \end{tabular}
    \caption{Reported use of agile methods}
    \label{table:practices}
\end{table}

\subsection{Motivations for transformation}

\begin{table}[h]
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Reason                                    & In studies   \\ \midrule
        Need to improve performance               &  \\ 
        Need to eliminate known process problems  &  \\
        Need to improve time to market            &  \\
        \bottomrule
    \end{tabular}
    \caption{Reported resasons for initiating a transformation}
    \label{table:motivations}
\end{table}

\subsection{Typical characteristics of a transformation}

Text only. text. text.

\subsection{Reported challenges}

\begin{table}[h]
    \begin{tabular}{ p{0.36\textwidth} p{0.62\textwidth} }
        \toprule
        Class of challenge  & Examples   \\ \midrule

        \raggedright Aligning organization for change  &
             Lack of management support \newline
             Change resistance \newline
             Reverting to old model under pressure \\

        \raggedright\rule{0pt}{0.4cm}Misunderstanding the agile model  &
             Misinterpretation of roles and responsibilities \newline
             Misunderstanding management role in agile \\

        \raggedright\rule{0pt}{0.4cm}Applicability of agile in the organization  &
            Conflicts with business (long term planning) \newline
            Conflicts with other departments (HR) \\

        \raggedright\rule{0pt}{0.4cm}Lack of resources  &
            Lack or training, insufficient resources for coaching \\
        \bottomrule
    \end{tabular}
    \caption{Typical challenges encountered in studies}
    \label{table:challenge}
\end{table}

\subsection{Success factors}

\begin{table}[h]
    \begin{tabular}{ p{0.36\textwidth} p{0.62\textwidth} }
        \toprule
        Class of success factors  & Examples   \\ \midrule

        \raggedright Organization alignment and leadership  &
             Management support \newline
             Communicating vision for new process \newline
             Demand all organizational units to participate \\

        \raggedright\rule{0pt}{0.4cm}Coaching and community  &
             Create community to make the new practices stick \\

        \raggedright\rule{0pt}{0.4cm}Tailoring and conformity  &
            Invest in tailoring to create the most suitable model \newline
            Conformity clarifies use of the new model \\

        \raggedright\rule{0pt}{0.4cm}Investing in change  &
            Training, consulting, physical spaces, tools \\
        \bottomrule
    \end{tabular}
    \caption{Success factors reported in studies}
    \label{table:success}
\end{table}

% ---------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusions}

-- Research on large scale agile transformations has been published.

-- Papers presenting the effect of agile on specific areas in software
development exist, but the whole has not been studied as much.

\subsection{Limitations}

-- The papers are mostly experience reports with no research method.

-- Some papers were of low quality, but they were included in order to give a
view of the topic that is broader and inclined towards practice.

-- Effect of organizational change is hard (or maybe impossible?) to measure
(kuulemma Marjo ja Kristian tietävät asiasta).

\subsection{Future research}

-- Very little studies on quantitative comparison of before/after transformation
has been done. (how about quantitative?)

-- What thihgs are important to report in an agile transformation case study?

-- Do not focus exclusively on the software perspective, but find parallels from
general research on organizaional transformation and leadership. (Tähän pitäs
vissiin sitte olla joku konkreettinen ehdotus?)

% ---------------------------------------------------------------------
% -------------- BIBLIOGRAPHY -----------------------------------------
% ---------------------------------------------------------------------

\bibliographystyle{plain}
\bibliographystylegen{plain}

\bibliographygen{sources}

\nobibliography{sources_s}

\subsection*{Primary studies}

\begin{supertabular}{ l p{11.3cm} }
    {[}S1{]} & \bibentry{S1} \\ 
    {[}S2{]} & \bibentry{S2} \\ 
    {[}S3{]} & \bibentry{S3} \\ 
    {[}S4{]} & \bibentry{S4} \\ 
    {[}S5{]} & \bibentry{S5} \\ 
    {[}S6{]} & \bibentry{S6} \\ 
    {[}S7{]} & \bibentry{S7} \\ 
    {[}S8{]} & \bibentry{S8} \\ 
    {[}S9{]} & \bibentry{S9} \\ 
    {[}S10{]} & \bibentry{S10} \\ 
    {[}S11{]} & \bibentry{S11} \\ 
    {[}S12{]} & \bibentry{S12} \\ 
    {[}S13{]} & \bibentry{S13} \\ 
    {[}S14{]} & \bibentry{S14} \\ 
    {[}S15{]} & \bibentry{S15} \\ 
    {[}S16{]} & \bibentry{S16} \\ 
    {[}S17{]} & \bibentry{S17} \\ 
    {[}S18{]} & \bibentry{S18} \\ 
    {[}S19{]} & \bibentry{S19} \\ 
    {[}S20{]} & \bibentry{S20} \\ 
    {[}S21{]} & \bibentry{S21} \\ 
    {[}S22{]} & \bibentry{S22} \\ 
    {[}S23{]} & \bibentry{S23} \\ 
    {[}S24{]} & \bibentry{S24} \\ 
    {[}S25{]} & \bibentry{S25} \\ 
    {[}S26{]} & \bibentry{S26} \\ 
    {[}S27{]} & \bibentry{S27} \\ 
    {[}S28{]} & \bibentry{S28} \\ 
    {[}S29{]} & \bibentry{S29} \\ 
    {[}S30{]} & \bibentry{S30} \\ 
\end{supertabular}

\end{document}

### 2013-01-23 tex/article.tex
Reducing size of references
References to transformation reasons

@@
         Reason                                    & In studies   \\ \midrule
-        Need to improve performance               &  \\ 
-        Need to eliminate known process problems  &  \\
-        Need to improve time to market            &  \\
+        Need to improve performance               & S8, S11, S15, S16, S17, S24, S27 \\ 
+        Need to eliminate known process problems  & S5, S6, S10, S13, S19, S21 \\
+        Need to improve time to market            & S6, S9, S13, S19, S24 \\
         \bottomrule
@@
 
+\small
+
 \begin{supertabular}{ l p{11.3cm} }
-    {[}S1{]} & \bibentry{S1} \\ 
+    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.8cm}
     {[}S2{]} & \bibentry{S2} \\ 

### 2013-01-23 tex/article.tex
Results

@@
 
-\subsection{Overview of agile methods used}
+Most of the primary studies stated which agile methods were in use. The most
+commonly reported methods are listed in Table \ref{table:methods}, of which
+Scrum was the most popular single method. An important observation is that
+tailoring of methods was very common. The use of tailoring can be considered
+relating particularly to large scale, as the core of agile methods give very
+little advice on interfacing with various units present in a large organization.
 
@@
     \caption{Reported use of agile methods}
-    \label{table:practices}
+    \label{table:methods}
 \end{table}
 
-\subsection{Motivations for transformation}
+The reasons to initiate a change fell into three categories, as summarized in
+Table \ref{table:motivations}. The primary reasons for change was a strategical
+need to improve the performance of the organization, such as shorten the time to
+market for new features. In many cases the change was started when management
+acknowledged problems in the current way of working.
 
@@
     \caption{Typical challenges encountered in studies}
-    \label{table:challenge}
+    \label{table:challenges}
 \end{table}

### 2013-01-24 tex/article.tex
Results

@@
 
-\subsection{Typical characteristics of a transformation}
+The most usual way of implementing the change was a stepwise adoption of agile,
+which usually included a pilot project. A typical stepwise adoption was to first
+create agile teams and introduce basic practices. Having intra-team agile
+practices in place made it possible to see how the entire organization would be
+affected by the agile way of working.
 
-Text only. text. text.
+The central role of management in the change was evident in many studies. This
+was typically shown in that the change was initiated by the management and
+management took responsibility of the change process. Several studies identified
+a champion who was the driver of the change. Some studies reported the change to
+be driven by external consultants or a cross organizational work group.
 
@@
         \raggedright Aligning organization for change  &
-             Lack of management support \newline
-             Change resistance \newline
-             Reverting to old model under pressure \\
+             Lack of management support [S1, S5, S13]\newline
+             Change resistance [S14, S22, S26] \newline
+             Reverting to old model [S4, S11, S23] \\
 
         \raggedright\rule{0pt}{0.4cm}Misunderstanding the agile model  &
-             Misinterpretation of roles and responsibilities \newline
-             Misunderstanding management role in agile \\
+             Wrong application of roles and responsibilities \newline
+             [S6, S10, S11, S13, S17, S23, S24] \\
 
         \raggedright\rule{0pt}{0.4cm}Applicability of agile in the organization  &
-            Conflicts with business (long term planning) \newline
-            Conflicts with other departments (HR) \\
+            Conflicts with business (eg. planning) [S6, S11, S21] \newline
+            Suitability for non-development (testing, maintenance, etc.) [S2, S4, S13]\newline
+            Conflicts with other functions [S6, S10, S11, S13] \\
 
         \raggedright\rule{0pt}{0.4cm}Lack of resources  &
-            Lack or training, insufficient resources for coaching \\
+            Lack of training, insufficient resources for coaching \newline
+            [S5, S13, S22, S28]\\
         \bottomrule
@@
         \raggedright Organization alignment and leadership  &
-             Management support \newline
-             Communicating vision for new process \newline
-             Demand all organizational units to participate \\
+             Management support [S8, S9, S10, S13, S16, S19] \newline
+             Communicate vision for new process [S7, S19] \newline
+             Demand all organizational units to participate [S13, S15, S17, S18, S19, S21] \\
 
         \raggedright\rule{0pt}{0.4cm}Coaching and community  &
-             Create community to make the new practices stick \\
+             Create community to make the new practices stick [S4, S6, S7, S28] \\
 
         \raggedright\rule{0pt}{0.4cm}Tailoring and conformity  &
-            Invest in tailoring to create the most suitable model \newline
-            Conformity clarifies use of the new model \\
+            Tailor the most suitable model [S8, S11, S20] \newline
+            Conformity clarifies use of the new model [S19, S24, S29, S30]\\
 
         \raggedright\rule{0pt}{0.4cm}Investing in change  &
-            Training, consulting, physical spaces, tools \\
+            Training, consulting, physical spaces, tools [S6, S14, S19]\\
         \bottomrule

### 2013-01-24 tex/article.tex
Todos, shortening

@@
 
-It is typical for large companies to function according to traditional software
+It is typical for large companies to function according to plan-driven software
 engineering models [Cite?!] (also why?). These models strive for consistent quality and performance
@@
 
-\subsection{Data sources and search strategy}
+%\subsection{Data sources and search strategy}
 
@@
 
-\subsection{Inclusion criteria}
+%\subsection{Inclusion criteria}
 
@@
 
-\subsection{Data extraction}
+%\subsection{Data extraction}
 
@@
 
-All of the included studies had been published within a decade. On average the
-studies were published theree years after the transformation was reported to
-have begun. Figure \ref{fig:publications} gives a timeline view of the
-publications.
-
-\begin{figure}[htb]
+\begin{figure}[h]
   \begin{center}
@@
 
-Most of the primary studies stated which agile methods were in use. The most
-commonly reported methods are listed in Table \ref{table:methods}, of which
-Scrum was the most popular single method. An important observation is that
+All of the included studies had been published within a decade. On average the
+studies were published theree years after the transformation was reported to
+have begun. Figure \ref{fig:publications} gives a timeline view of the
+publications.
+
+The most commonly reported methods are listed in Table \ref{table:methods}, of
+which Scrum was the most popular single method. An important observation is that
 tailoring of methods was very common. The use of tailoring can be considered
@@
 Table \ref{table:motivations}. The primary reasons for change was a strategical
-need to improve the performance of the organization, such as shorten the time to
-market for new features. In many cases the change was started when management
-acknowledged problems in the current way of working.
+need to improve the performance of the organization. Especially shortening the
+time to market was seen as XXX. In many cases the change was started when
+management acknowledged problems in the current way of working.
 
@@
         Need to improve performance               & S8, S11, S15, S16, S17, S24, S27 \\ 
+        Need to improve time to market            & S6, S9, S13, S19, S24 \\
         Need to eliminate known process problems  & S5, S6, S10, S13, S19, S21 \\
-        Need to improve time to market            & S6, S9, S13, S19, S24 \\
         \bottomrule
@@
 be driven by external consultants or a cross organizational work group.
-
-\subsection{Reported challenges}
 
@@
             [S5, S13, S22, S28]\\
-        \bottomrule
-    \end{tabular}
-    \caption{Typical challenges encountered in studies}
-    \label{table:challenges}
-\end{table}
 
-\subsection{Success factors}
-
-\begin{table}[h]
-    \begin{tabular}{ p{0.36\textwidth} p{0.62\textwidth} }
-        \toprule
+        \midrule
         Class of success factors  & Examples   \\ \midrule
@@
     \end{tabular}
-    \caption{Success factors reported in studies}
+    \caption{Challenges and success factors reported in studies}
     \label{table:success}
@@
 
--- Research on large scale agile transformations has been published.
+* -- Research on large scale agile transformations has been published.
 
@@
 
--- The papers are mostly experience reports with no research method.
+* -- The papers are mostly experience reports with no research method.
 
--- Some papers were of low quality, but they were included in order to give a
+* -- Some papers were of low quality, but they were included in order to give a
 view of the topic that is broader and inclined towards practice.
@@
 
--- Very little studies on quantitative comparison of before/after transformation
+* -- Very little studies on quantitative comparison of before/after transformation
 has been done. (how about quantitative?)
@@
 
+\subsection{TODO}
+
+eg. --> e.g.
+
+Abstract: miksi on tärkeätä tutkia tätä?
+-- Mitkä on pääasialliset löydökset?
+
+Otsikko:
+literature review -> motivation, success and challenges
+
+Missä sanotaan mikä on "large"
+
+Search kuva on selitetty myös tekstissä.
+
+Inclusion: viewpoint on the organizational transformation [from plan-driven, tms.]
+
+Used methods: interfacing with various units present in a large organization
+--> Developmentin koordinointi
+
+"Applicability of agile problem" --> Paremmin "ongelmaa" kuvaava ilmaisu
+--> Myös keskimmäinen pointti
+--> "Other functions" --> tarkemmin
+
+"Coaching and community" --> selitä myös coaching
+
+Success factor: Stepwise adoption (piloting)
+
+Conclusions:
+
+--> Päätä siihen että "kiinnostavaa materiaalia löytyy" vaikka ei eka oletettu
+--> Olemme päättäneet toteuttaa laajemman ja systemaattisen tutkimuksen aiheesta
+
 % ---------------------------------------------------------------------
@@
 % ---------------------------------------------------------------------
+
+\small
 
@@
 
-\small
-
 \begin{supertabular}{ l p{11.3cm} }
-    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.8cm}
+    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.9cm}
     {[}S2{]} & \bibentry{S2} \\ 

### 2013-01-24 tex/article.tex
Fixed "applicability of agile" in challenges

@@
 
+Typical challenges and and success factors reported in the studies are listed in
+Table \ref{table:success}.
+
 \begin{table}[h]
@@
 
-        \raggedright\rule{0pt}{0.4cm}Applicability of agile in the organization  &
+        \raggedright\rule{0pt}{0.4cm}Effect of agile outside of development  &
             Conflicts with business (eg. planning) [S6, S11, S21] \newline
-            Suitability for non-development (testing, maintenance, etc.) [S2, S4, S13]\newline
-            Conflicts with other functions [S6, S10, S11, S13] \\
+            Making non-development functions agile (testing, maintenance, etc.) [S2, S4, S13]\newline
+            Conflicts with other functions (marketing, HR, etc.) \newline[S6, S10, S11, S13] \\
 
@@
 
-"Applicability of agile problem" --> Paremmin "ongelmaa" kuvaava ilmaisu
---> Myös keskimmäinen pointti
---> "Other functions" --> tarkemmin
-
 "Coaching and community" --> selitä myös coaching

### 2013-01-25 tex/article.tex
Text for challenges

@@
 Typical challenges and and success factors reported in the studies are listed in
-Table \ref{table:success}.
+Table \ref{table:success}. In both cases the ability to align the organization
+was a central factor in both cases. Challenges surface if not everyone is on
+board with the change. Good ground for the change is created by strongly
+communicating a clear vision and enabling everyone to participate. Resourcing
+for the change was also seen both as a challenge and a success factor.
+The studies show that allocating extra training and hiring of external consultants
+facilitate the change.
+
+A particular challenge for adopting agile was misunderstanding how agile should
+be applied. Some symptoms of this were talking agile but still working the old
+way, and managers making commitments without hearing development teams.
+As expected, many studies reported that development teams had problems
+interfacing with other organizational functions, such as management and
+marketing.
 
@@
              Wrong application of roles and responsibilities \newline
-             [S6, S10, S11, S13, S17, S23, S24] \\
+             [S11, S13, S16, S17, S23, S24] \newline
+             Agile terms merely mask the old model [S10, S22] \\
 
         \raggedright\rule{0pt}{0.4cm}Effect of agile outside of development  &
-            Conflicts with business (eg. planning) [S6, S11, S21] \newline
+            Conflicts with business (e.g. planning) [S6, S11, S21] \newline
             Making non-development functions agile (testing, maintenance, etc.) [S2, S4, S13]\newline

### 2013-01-25 tex/article.tex
Optimizing, removing text

@@
 It is typical for large companies to function according to plan-driven software
-engineering models [Cite?!] (also why?). These models strive for consistent quality and performance
+engineering models. These models strive for consistent quality and performance
 by rigorous planning and definition of process. This kind of approach is however
@@
 \citegen{Livermore2008}. Agile methods are not founded on the use of individual
-tools or practices, but rather on a holistic change in the way of thinking.
+tools or practices, but rather on a holistic way of thinking.
 Adopting agile requires often change of the entire organizational culture
@@
 the need for formal documentation and thus reduces agility \citegen{Lindvall2004}.
-In addition to inter-team coordination large organizations require development
-teams to interact with other organizational units, which are often non-agile of
-their nature. For instance, HR may demand individuals to have strictly specified
-roles in projects \citegen{Boehm2005}, or a change control board may inhibit use of
-continuous integration or refactoring \citegen{Lindvall2004}. All units affected by
-the agile transformation need to be informed and heard, and the agile process
+In addition to inter-team coordination development teams are required to
+interact with other organizational units, which are often non-agile of their
+nature. For instance, HR may demand individuals to have strictly specified roles
+in projects \citegen{Boehm2005}, or a change control board may inhibit use of
+continuous integration or refactoring \citegen{Lindvall2004}. All units affected
+by the agile transformation need to be informed and heard, and the agile process
 must be adjusted according to their needs \citegen{Lindvall2004, Cohn2003,
@@
 challenge is that management must move away from life cycle models and towards
-iterative and feature centric models \citegen{Nerur2005}. The focus must be shifted
-from a large scale scope to shorter term project planning \citegen{Misra2010}, as
-agile methods emphasize that meaningful planning can only be done for the near
-future \citegen{Cohn2003}. However, the lack of planning can be a concern as
-business and customer relationships often build on long term roadmapping.
-Enabling operation with shorter term planning requires educating stakeholders
-and reviewing contracting practices \citegen{Boehm2005}.
+iterative and feature centric models \citegen{Nerur2005}. The focus must be
+shifted from a large scale scope to shorter term project planning
+\citegen{Misra2010}, as agile methods emphasize that planning is only meaningful
+for the near future \citegen{Cohn2003}. However, the lack of planning can be a
+concern as business and customer relationships often build on long term
+roadmapping. Enabling operation with shorter term planning requires educating
+stakeholders and reviewing contracting practices \citegen{Boehm2005}.
 
@@
 was applied to be performed by only one researcher, and to serve as a
-preliminary work evaluating the feasibility of a more in-depth review. The
-required effort was limited by selecting only the following subset of SLR steps:
+preliminary work evaluating the feasibility of a more in-depth review.
+Effort was limited by selecting only the following subset of SLR steps:
 identification of primary sources, data extraction and synthesis. Steps left out
@@
 
+\begin{table}
+    \begin{tabular}{ l@{ \hskip 0.4cm } l }
+        \toprule
+        Facet                  & Keywords   \\ \midrule
+        Agile methods          & agile, scrum, lean, xp \\ 
+        Organizational change  & transformation, transition, change, migration \\
+        Large organization     & enterprise, organization, (large \texttt{AND} scale) \\
+        \bottomrule
+    \end{tabular}
+    \caption{Facets and related search terms used}
+    \label{table:searchterms}
+\end{table}
+
 As the preliminary search omitted some of the keywords in Table
@@
 
-\begin{table}[h]
-    \begin{tabular}{ l@{ \hskip 0.4cm } l }
-        \toprule
-        Facet                  & Keywords   \\ \midrule
-        Agile methods          & agile, scrum, lean, xp \\ 
-        Organizational change  & transformation, transition, change, migration \\
-        Large organization     & enterprise, organization, (large \texttt{AND} scale) \\
-        \bottomrule
-    \end{tabular}
-    \caption{Facets and related search terms used}
-    \label{table:searchterms}
-\end{table}
+%\subsection{Inclusion criteria}
 
 The selection process of the primary studies is summarized in Figure
-\ref{fig:selection_process}. The database search yielded 1603 matches from which
-100 sources were chosen by the title. 17 unmatched sources from the preliminary
-search were also included. Finally, 31 sources were selected based on the
-abstract to be included for data extraction. One source was not available, which
+\ref{fig:selection_process}. The database search yielded 1603 matches which were
+filtered down to 31 selected primary studies. One study was not available, which
 reduced the final number of primary studies to 30.
+The 30 primary sources were selected by three criteria: study type, size of
+organization, and use of agile methods in software development. Firstly, only
+case studies and experience reports on agile software development were included
+in this study. Papers discussing agile methods or organizational transformation
+on as a general topic were left out. Secondly, the size of the organization
+described in the included primary studies was to be large enough to show the
+challenges of large scale agile development \citegen{Lindvall2004}. Further,
+only studies discussing agile software development were included, leaving out
+for instance studies on agile management.
 
-\begin{figure}[htb]
+\begin{figure}[t]
   \begin{center}
-    \includegraphics[width=0.5\textwidth]{researchprocess}
+    \includegraphics[width=0.45\textwidth]{researchprocess}
     \caption{Process of selecting primary studies}
@@
 
-%\subsection{Inclusion criteria}
-
-The 30 primary sources were selected by four criteria: study type, size of
-organization, use of agile methods in software development, and viewpoint on
-transformation. Firstly, only case studies and experience reports on agile
-software development were included in this study. Papers discussing agile
-methods or organizational transformation on as a general topic were left out.
-Secondly, the size of the organization described in the included primary studies
-was to be large enough to show the challenges of large scale agile development
-\citegen{Lindvall2004}. Further, only studies discussing agile software development
-were included, leaving out for instance studies on agile management. Finally, to
-be included each study had to provide a viewpoint on the organizational
-transformation.
-
 %\subsection{Data extraction}
 
-The content of the resutls was guided by the data extraction form. The form
-included some quantitative fields including \textit{start time of
-transformation} and \textit{organization size}, but most of the data was
-collected in open ended fields. The reason for starting a transformation was
-captured in fileds \textit{initial state of organization} and \textit{reasons
-for change}. The description of the transformation process was captured in one
-field. Recommendations for transformation processes were captured in fields
+Analysis of the primary studies was guided by using a data extraction form with
+open ended fileds. The reason for starting a transformation was captured in
+fileds \textit{initial state of organization} and \textit{reasons for change}.
+The description of the transformation process was captured in one field.
+Recommendations for transformation processes were captured in fields
 \textit{reported good practices}, \textit{reported challenges}, and
-\textit{lessons learned}. Additional fields that were used but not sufficiently
-answered by the primary studies were \textit{satisfaction after transformation},
-\textit{effect on onrganization} and \textit{reported measurements before/after
-transformation}.
+\textit{lessons learned}.
+Additional fields that were used but not sufficiently answered by the primary
+studies were \textit{satisfaction after transformation} and \textit{reported
+measurements before/after transformation}.
 
@@
 
-\begin{figure}[h]
+\begin{figure}
   \begin{center}
@@
 
-\begin{table}[h]
+\begin{table}
     \begin{tabular}{ l@{ \hskip 0.4cm } l }
@@
 need to improve the performance of the organization. Especially shortening the
-time to market was seen as XXX. In many cases the change was started when
-management acknowledged problems in the current way of working.
+time to market was desired. In many cases the change was started when management
+acknowledged problems in the current way of working.
 
-\begin{table}[h]
+\begin{table}
     \begin{tabular}{ l@{ \hskip 0.4cm } l }
@@
 \begin{supertabular}{ l p{11.3cm} }
-    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.9cm}
+    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-2.1cm}
     {[}S2{]} & \bibentry{S2} \\ 

### 2013-01-25 tex/article.tex
Success factors

@@
 
-\begin{table}
+\begin{table}[t]
     \begin{tabular}{ l@{ \hskip 0.4cm } l }
@@
 
-\begin{figure}
+\begin{figure}[b]
   \begin{center}
@@
 
-\begin{table}[h]
-    \begin{tabular}{ p{0.36\textwidth} p{0.62\textwidth} }
+It is difficult to make a new way of wroking stick in a large organization.
+Use of agile can be rooted into the organization by creating an agile community.
+Founding of interest groups and arranging regular seminars are good ways for
+building community. As agility emphasizes people centricity and flexibility the
+new ways of working can not be mandated. Instead, coaching should be applied for
+leading the change. Several studies suggest conformity in the application of
+agile models. Conformity avoids disorder when several teams are simultaneously
+finding their agile ways.
+
+\begin{table}[t]
+    \begin{tabular}{ p{0.31\textwidth} p{0.67\textwidth} }
         \toprule
@@
         \raggedright\rule{0pt}{0.4cm}Coaching and community  &
-             Create community to make the new practices stick [S4, S6, S7, S28] \\
+             Create community to make change stick [S6, S7, S19, S28] \newline
+             Lead change by a coaching approach [S1, S4, S22, S28] \\
 
@@
             Tailor the most suitable model [S8, S11, S20] \newline
-            Conformity clarifies use of the new model [S19, S24, S29, S30]\\
+            Conformity clarifies use of new model [S19, S24, S29, S30]\\
 
@@
     {[}S30{]} & \bibentry{S30} \\ 
-\end{supertabular}
-
-\end{document}
+\end{supertabular}\end{document}

### 2013-01-25 tex/article.tex
Title, conclusions

@@
 
-\title{Large scale agile transformations: A literature review}
+\title{Adopting agile in the large: motivations, success factors and challenges}
 
-\titlerunning{Large scale agile transformations}
+\titlerunning{Adopting agile in the large}
 
@@
 \institute{Aalto University . . . Address . . .
-\\ \email{email@xx},
-\\ WWW home page: \texttt{http://www...} }
+\\ \email{kim.dikert@aalto.fi}, \email{maria.paasivaara@aalto.fi} }
 
@@
 %                         please supply keywords within your abstract
-\keywords {agile, transformation, large scale}
+\keywords {agile, transformation, large scale, adopting agile}
 \end{abstract}
@@
 method used for conducting the research. The findings of the review are
-discussed in section \ref{sec:results}, and section \ref{sec:conclusions}
-discusses the findings and gives proposals for future research.
+discussed in section \ref{sec:results}, and section \ref{sec:conclusion}
+gives proposals for future research.
 
@@
 % ---------------------------------------------------------------------
-\section{Conclusions}
-\label{sec:conclusions}
+\section{Future research}
+\label{sec:conclusion}
 
-* -- Research on large scale agile transformations has been published.
-
--- Papers presenting the effect of agile on specific areas in software
-development exist, but the whole has not been studied as much.
-
-\subsection{Limitations}
+Our results show that a body of research on large scale agile transformations
+has emerged. We aim to continue research on this topic by conducting a
+systematic literature review. Beyond descriptive reports on agile
+transformations we found very little studies with quantitative comparison of
+organizations before and after the transformations.
 
@@
 
-\subsection{Future research}
-
-* -- Very little studies on quantitative comparison of before/after transformation
-has been done. (how about quantitative?)
-
--- What thihgs are important to report in an agile transformation case study?
-
--- Do not focus exclusively on the software perspective, but find parallels from
-general research on organizaional transformation and leadership. (Tähän pitäs
-vissiin sitte olla joku konkreettinen ehdotus?)
-
 \subsection{TODO}
@@
 -- Mitkä on pääasialliset löydökset?
-
-Otsikko:
-literature review -> motivation, success and challenges
 
@@
 
-\begin{supertabular}{ l p{11.3cm} }
-    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-2.1cm}
+\begin{supertabular}{ l p{11.4cm} }
+    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.6cm}
     {[}S2{]} & \bibentry{S2} \\ 
@@
     {[}S8{]} & \bibentry{S8} \\ 
-    {[}S9{]} & \bibentry{S9} \\ 
+    {[}S9{]} & \bibentry{S9} \\ \shrinkheight{-1.6cm}
     {[}S10{]} & \bibentry{S10} \\ 

### 2013-01-25 tex/article.tex
Conclusions, sorting out todo

@@
 \begin{abstract}        % give a summary of your paper
+Abstract: miksi on tärkeätä tutkia tätä?
+-- Mitkä on pääasialliset löydökset?
 The abstract should summarize the contents of the paper
@@
 agile methods. Studies on large scale agile transformations have been published,
-but no extensive and summarizing research is available on the subject. In this
+but no extensive or summarizing research is available on the subject. In this
 research we will give an overview of reported case studies on adopting agile
@@
 scale agile development. Section \ref{sec:method} presents the literature review
-method used for conducting the research. The findings of the review are
+method used for conducting this research. The findings of the review are
 discussed in section \ref{sec:results}, and section \ref{sec:conclusion}
-gives proposals for future research.
+presents limitations and proposals for future research.
 
@@
 reduced the final number of primary studies to 30.
-The 30 primary sources were selected by three criteria: study type, size of
-organization, and use of agile methods in software development. Firstly, only
-case studies and experience reports on agile software development were included
-in this study. Papers discussing agile methods or organizational transformation
-on as a general topic were left out. Secondly, the size of the organization
-described in the included primary studies was to be large enough to show the
-challenges of large scale agile development \citegen{Lindvall2004}. Further,
-only studies discussing agile software development were included, leaving out
-for instance studies on agile management.
+The 30 primary sources were selected by four criteria: study type, size of
+organization, use of agile methods in software development, and a transformation
+viewpoint. Firstly, only case studies and experience reports on agile software
+development were included in this study. Papers discussing agile methods or
+organizational transformation on as a general topic were left out. Secondly, the
+size of the organization described in the included primary studies was to be
+large enough to show the challenges of large scale agile development
+\citegen{Lindvall2004}. Further, only studies discussing agile software
+development were included, leaving out for instance studies on agile management.
+Finally, the included studies had to provide a viewpoint on the adoption of
+agile methods.
 
@@
 measurements before/after transformation}.
-
-% \begin{table}
-%     \begin{tabular}{ p{0.36\textwidth}@{ \hskip 0.2cm } l }
-%         \toprule
-%         Transformation mentioned in \newline text (Y/N) &
-%         Large scale mentioned in text (Y/N) \\
-%         Is empirical case study (Y/N) &
-%         Is industry experience report (Y/N) \\
-%         Has listing of practices (Y/N) &
-%         Used research method (Y/N) \\
-%         Relevance to this review (1-5) & \\
-%         Objective of research (or publication) &
-%         Research method \\
-%         Author bias &
-%         Validity threats \\
-%         Organization size &
-%         Time of transformation \\
-%         Initial state of organization &
-%         Why was the change initiated \\
-%         How was the change conducted &
-%         What is agility? / Which agile paractices are used? \\
-%         Findings / lessons learned &
-%         Good practices validated or suggested by study \\
-%         Reported challenges &
-%         Satisfaction after transformation \\
-%         Effect on organization &
-%         Measurements as results (quantitative or other) \\
-%         Other notes &
-%         Notable references \\
-%         \bottomrule
-%     \end{tabular}
-%     \caption{The fields of the data extraction form. Fields with a (Y/N) label
-%     were answered as yes or no. The field with a (1-5) label was answered by a
-%     grade from 1 to 5. All other fields were answered by free form text.}
-%     \label{table:dataform}
-% \end{table}
-
-%\subsection{Data sources and search strategy}
-%
-%\subsection{Final selection}
-%
-%\subsection{Data extraction and synthesis}
 
@@
 relating particularly to large scale, as the core of agile methods give very
-little advice on interfacing with various units present in a large organization.
+little advice on interfacing with other units present in a large organization.
+The agile team must interface with units such as other development teams,
+management, and marketing.
 
@@
 way, and managers making commitments without hearing development teams.
-As expected, many studies reported that development teams had problems
-interfacing with other organizational functions, such as management and
-marketing.
+As expected, many studies reported that teams had problems interfacing with
+other organizational units, such as management and marketing.
 
@@
             Making non-development functions agile (testing, maintenance, etc.) [S2, S4, S13]\newline
-            Conflicts with other functions (marketing, HR, etc.) \newline[S6, S10, S11, S13] \\
+            Conflicts with other units (marketing, HR, etc.) \newline[S6, S10, S11, S13] \\
 
@@
 % ---------------------------------------------------------------------
-\section{Future research}
+\section{Limitations and future research}
 \label{sec:conclusion}
 
-Our results show that a body of research on large scale agile transformations
-has emerged. We aim to continue research on this topic by conducting a
-systematic literature review. Beyond descriptive reports on agile
-transformations we found very little studies with quantitative comparison of
-organizations before and after the transformations.
+The papers included in this study were mostly experience reports with no
+research method. Some papers were of low quality, but they were included as our
+aim was to give a view of the topic that is broader and inclined towards practice.
 
-* -- The papers are mostly experience reports with no research method.
+Differing from our initial beliefs, this research shows that a body of studies
+on large scale agile transformations has emerged. We aim to continue our
+research on this topic by conducting a systematic literature review. Beyond
+descriptive reports on agile transformations we found very little studies with
+quantitative comparison of organizations' performance before and after the
+transformations.
 
-* -- Some papers were of low quality, but they were included in order to give a
-view of the topic that is broader and inclined towards practice.
+% \subsection{TODO}
 
--- Effect of organizational change is hard (or maybe impossible?) to measure
-(kuulemma Marjo ja Kristian tietävät asiasta).
+% Vielä:
+% Missä sanotaan mikä on "large"
+% Success factor: Stepwise adoption (piloting)
 
-\subsection{TODO}
-
-eg. --> e.g.
-
-Abstract: miksi on tärkeätä tutkia tätä?
--- Mitkä on pääasialliset löydökset?
-
-Missä sanotaan mikä on "large"
-
-Search kuva on selitetty myös tekstissä.
-
-Inclusion: viewpoint on the organizational transformation [from plan-driven, tms.]
-
-Used methods: interfacing with various units present in a large organization
---> Developmentin koordinointi
-
-"Coaching and community" --> selitä myös coaching
-
-Success factor: Stepwise adoption (piloting)
-
-Conclusions:
-
---> Päätä siihen että "kiinnostavaa materiaalia löytyy" vaikka ei eka oletettu
---> Olemme päättäneet toteuttaa laajemman ja systemaattisen tutkimuksen aiheesta
+% Pitäs olla done:
+% Search kuva on selitetty myös tekstissä.
+% Inclusion: viewpoint on the organizational transformation [from plan-driven, tms.]
+% Used methods: interfacing with various units present in a large organization
+%    --> Developmentin koordinointi
+% "Coaching and community" --> selitä myös coaching
 

### 2013-01-25 tex/article.tex
Tweaking, removed footnotes

@@
 
-%\subsection{Data sources and search strategy}
-
 The searches for primary sources was conducted in two steps. First a preliminary
@@
 done based on the preliminary search. The following databases were included in
-the searches: IEEExplore\footnote{http://ieeexplore.ieee.org/},
-ACM\footnote{http://dl.acm.org/}, Scopus\footnote{http://www.scopus.com/},
-ProQuest\footnote{http://search.proquest.com/}.
+the searches: IEEExplore (http://ieeexplore.ieee.org/),
+ACM (http://dl.acm.org/), Scopus (http://www.scopus.com/),
+ProQuest \linebreak (http://search.proquest.com/).
 Additionally the archive of International XP
-Conference\footnote{http://link.springer.com/} was searched.
+Conference (http://link.springer.com/) was searched.
 
@@
 \texttt{AND} operators.
+
+As the preliminary search omitted some of the keywords in Table
+\ref{table:searchterms} it yielded a small number of additional matches. The
+reason for variance in matches was that many articles had creative titles or the
+abstracts and keywords were not informative enough. Because of this, the search
+with more restrictive keywords was unable to capture all possible matches.
+We are however confident that the search covered the majority of material
+relating to agile transformations in large organizations. If a wider search
+should be performed, our recommendation is to only regard the facets
+\textit{agile methods} and \textit{organizational change}. This would lead to
+significantly more matches in the first stage of selecting studies, but the
+additional effort is necessary due to the limitations of a keyword search.
 
@@
 
-As the preliminary search omitted some of the keywords in Table
-\ref{table:searchterms} it yielded a small number of additional matches. The
-reason for variance in matches was that many articles had creative titles or the
-abstracts and keywords were not informative enough. Because of this, the search
-with more restrictive keywords was unable to capture all possible matches.
-We are however confident that the search covered the majority of material
-relating to agile transformations in large organizations. If a wider search
-should be performed, our recommendation is to only regard the facets
-\textit{agile methods} and \textit{organizational change}. This would lead to
-significantly more matches in the first stage of selecting studies, but the
-additional effort is necessary due to the limitations of a keyword search.
-
-%\subsection{Inclusion criteria}
+\begin{figure}[b]
+  \begin{center}
+    \includegraphics[width=0.45\textwidth]{researchprocess}
+    \caption{Process of selecting primary studies}
+    \label{fig:selection_process}
+  \end{center}
+\end{figure}
 
@@
 agile methods.
-
-\begin{figure}[t]
-  \begin{center}
-    \includegraphics[width=0.45\textwidth]{researchprocess}
-    \caption{Process of selecting primary studies}
-    \label{fig:selection_process}
-  \end{center}
-\end{figure}
 
@@
 
-\small
+%\small
+\fontsize{9pt}{10pt}\selectfont
 
@@
 \begin{supertabular}{ l p{11.4cm} }
-    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.6cm}
+    {[}S1{]} & \bibentry{S1} \\  \shrinkheight{-1cm}
     {[}S2{]} & \bibentry{S2} \\ 
@@
     {[}S8{]} & \bibentry{S8} \\ 
-    {[}S9{]} & \bibentry{S9} \\ \shrinkheight{-1.6cm}
+    {[}S9{]} & \bibentry{S9} \\
     {[}S10{]} & \bibentry{S10} \\ 
     {[}S11{]} & \bibentry{S11} \\ 
-    {[}S12{]} & \bibentry{S12} \\ 
+    {[}S12{]} & \bibentry{S12} \\  \shrinkheight{-1.4cm}
     {[}S13{]} & \bibentry{S13} \\ 

### 2013-01-25 tex/article.tex
Spellcheck

@@
 late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
-resopnding to change and distance from customers undermine the  competitiveness
+responding to change and distance from customers undermine the  competitiveness
 of companies. Agile methods are believed to bring remedy to these problems.
@@
 research is focused on the process of organizational change. We leave out topics
-related to comparing the initial and transformed states of organisations, such
+related to comparing the initial and transformed states of organizations, such
 as measuring the performance of organizations before and after the
@@
 Analysis of the primary studies was guided by using a data extraction form with
-open ended fileds. The reason for starting a transformation was captured in
-fileds \textit{initial state of organization} and \textit{reasons for change}.
+open ended fields. The reason for starting a transformation was captured in
+fields \textit{initial state of organization} and \textit{reasons for change}.
 The description of the transformation process was captured in one field.
@@
 All of the included studies had been published within a decade. On average the
-studies were published theree years after the transformation was reported to
+studies were published three years after the transformation was reported to
 have begun. Figure \ref{fig:publications} gives a timeline view of the
@@
         Lean            & 5 \\
-        Tailoring and combining methods & 9  (indirect mentionings not counted) \\
+        Tailoring and combining methods & 9  (indirect mentioning not counted) \\
         \bottomrule
@@
     \end{tabular}
-    \caption{Reported resasons for initiating a transformation}
+    \caption{Reported reasons for initiating a transformation}
     \label{table:motivations}
@@
 
-It is difficult to make a new way of wroking stick in a large organization.
+It is difficult to make a new way of working stick in a large organization.
 Use of agile can be rooted into the organization by creating an agile community.

### 2013-01-25 tex/article.tex
Institute and contacts
Abstract

@@
 
-\institute{Aalto University . . . Address . . .
-\\ \email{kim.dikert@aalto.fi}, \email{maria.paasivaara@aalto.fi} }
+\institute{Aalto University, School of Science,
+Department of Computer Science and Engineering
+\\ \email{kim-karol.dikert@aalto.fi}, \email{maria.paasivaara@aalto.fi}, \email{casper.lassenius@aalto.fi} }
 
@@
 
-\begin{abstract}        % give a summary of your paper
-Abstract: miksi on tärkeätä tutkia tätä?
--- Mitkä on pääasialliset löydökset?
-The abstract should summarize the contents of the paper
-using at least 70 and at most 150 words. It will be set in 9-point
-font size and be inset 1.0 cm from the right and left margins.
-There will be two blank lines before and after the Abstract.
-%                         please supply keywords within your abstract
+\begin{abstract}
+
+Agile methods have become an appealing alternative for companies striving to
+improve their performance. Agile methods have been originally designed for small
+and individual teams. This poses challenges when introducing agile to large
+organizations where development teams can not be autonomous. We reviewed 30
+experience reports and case studies describing the process of taking agile
+methods into use in large organizations. The primary motive for agile adoption
+is the need to improve performance. For a successful transformation the entire
+organization must be aligned and all organizational units must participate.
+Creating an agile communityin the organization will help the change to stick.
+Typical problems are misunderstanding of the agile model, and difficulties
+coordinating the agile teams.
+
 \keywords {agile, transformation, large scale, adopting agile}
@@
 Typical challenges and and success factors reported in the studies are listed in
-Table \ref{table:success}. In both cases the ability to align the organization
-was a central factor in both cases. Challenges surface if not everyone is on
+Table \ref{table:success}. The ability to align the organization was a central
+factor for both success and challenges. Challenges surface if not everyone is on
 board with the change. Good ground for the change is created by strongly
@@
 for the change was also seen both as a challenge and a success factor.
-The studies show that allocating extra training and hiring of external consultants
-facilitate the change.
+The studies show that allocating extra training and hiring of external
+consultants facilitate the change.
 
@@
     {[}S11{]} & \bibentry{S11} \\ 
-    {[}S12{]} & \bibentry{S12} \\  \shrinkheight{-1.4cm}
+    {[}S12{]} & \bibentry{S12} \\  \shrinkheight{-3cm}
     {[}S13{]} & \bibentry{S13} \\ 

### 2013-01-25 tex/article.tex
Introduction: remove talk on plan-driven, and replace with problems in large scale.

@@
 looking to improve their effectiveness. Agile methods are claimed to increase
-productivity and quality \citegen{Livermore2008}, which makes them attractive for
-companies pursuing better performance. However, large organizations may have
-problems introducing agile methods \citegen{Dyba2009}. As agile methods are
-initially designed for small teams their application has proven to be
-problematic in a larger scale \citegen{Boehm2005}.
+productivity and quality \citegen{Livermore2008}, which makes them attractive
+for companies pursuing better performance. Agile methods give advantage by
+enabling shorter lead times and stronger focus on customer needs
+\citegen{Petersen2010}.
 
-It is typical for large companies to function according to plan-driven software
-engineering models. These models strive for consistent quality and performance
-by rigorous planning and definition of process. This kind of approach is however
-badly suited for software development, as projects typically encounter
-situations that are too hard or impossible to foretell \citegen{Schwaber2002}.
-The main problems in plan-driven development are the high cost of changes and
-late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
-responding to change and distance from customers undermine the  competitiveness
-of companies. Agile methods are believed to bring remedy to these problems.
+However, large organizations may have problems introducing agile methods
+\citegen{Dyba2009}. As agile methods are initially designed for small teams
+their application has proven to be problematic in a larger scale
+\citegen{Boehm2005}. In a large organization development teams will need to
+interface with other organizational units. As agile methods tend to focus on
+inter-team practices which leaves it up to the organization to figure out how to
+make agile work \citegen{Lindvall2004}.
+
+% --> There is demand for a summarizing study of large agile transformations as
+% the evidence is currently scattered.
+% 
+% These models strive for consistent quality and performance
+% by rigorous planning and definition of process. This kind of approach is however
+% badly suited for software development, as projects typically encounter
+% situations that are too hard or impossible to foretell \citegen{Schwaber2002}.
+% The main problems in plan-driven development are the high cost of changes and
+% late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
+% responding to change and distance from customers undermine the  competitiveness
+% of companies. Agile methods are believed to bring remedy to these problems.
 
@@
 Effort was limited by selecting only the following subset of SLR steps:
-identification of primary sources, data extraction and synthesis. Steps left out
+identification of primary studies, data extraction and synthesis. Steps left out
 were the development of a systematic review protocol, strict study quality
@@
 
-The searches for primary sources was conducted in two steps. First a preliminary
+The searches for primary studies was conducted in two steps. First a preliminary
 search was made to sample the existing literature. Then a systematic search was
@@
 reduced the final number of primary studies to 30.
-The 30 primary sources were selected by four criteria: study type, size of
+The primary studies were selected by four criteria: study type, size of
 organization, use of agile methods in software development, and a transformation

### 2013-01-25 tex/article.tex
Improve: conformity

@@
 new ways of working can not be mandated. Instead, coaching should be applied for
-leading the change. Several studies suggest conformity in the application of
-agile models. Conformity avoids disorder when several teams are simultaneously
-finding their agile ways.
+leading the change. Several studies suggest demanding conformity in the
+application of agile practices. Agreeing on a common way of working avoids
+disorder when several teams are simultaneously finding their agile ways.
 

### 2013-01-25 tex/article.tex
Corrections

@@
 Agile methods have become an appealing alternative for companies striving to
-improve their performance. Agile methods have been originally designed for small
-and individual teams. This poses challenges when introducing agile to large
+improve their performance, but they are originally designed for small and
+individual teams. This poses challenges when introducing agile to large
 organizations where development teams can not be autonomous. We reviewed 30
@@
 organization must be aligned and all organizational units must participate.
-Creating an agile communityin the organization will help the change to stick.
+Creating an agile community in the organization will help the change to stick.
 Typical problems are misunderstanding of the agile model, and difficulties
-coordinating the agile teams.
+coordinating multiple agile teams.
 
@@
 \citegen{Boehm2005}. In a large organization development teams will need to
-interface with other organizational units. As agile methods tend to focus on
-inter-team practices which leaves it up to the organization to figure out how to
-make agile work \citegen{Lindvall2004}.
+interface with other organizational units. Agile methods tend to focus on
+intra-team practices, which leaves it up to the organization to figure out how to
+tailor agile practices to suite a large organization \citegen{Lindvall2004}.
 
@@
 but no extensive or summarizing research is available on the subject. In this
-research we will give an overview of reported case studies on adopting agile
+paper we will give an overview of reported case studies on adopting agile
 methods in large organizations. We will further discuss the motivations for
@@
 tools or practices, but rather on a holistic way of thinking.
-Adopting agile requires often change of the entire organizational culture
+Adopting agile often requires change of the entire organizational culture
 \citegen{Misra2010}.
 
-One significant difference between small and large adoptions is that large
+One significant difference between small and large scale adoptions is that large
 organizations have more dependencies between projects and teams. This increases
@@
 In addition to inter-team coordination development teams are required to
-interact with other organizational units, which are often non-agile of their
+interact with other organizational units, which are often non-agile in
 nature. For instance, HR may demand individuals to have strictly specified roles
-in projects \citegen{Boehm2005}, or a change control board may inhibit use of
+in projects \citegen{Boehm2005}, or a change control board may inhibit the use of
 continuous integration or refactoring \citegen{Lindvall2004}. All units affected
-by the agile transformation need to be informed and heard, and the agile process
-must be adjusted according to their needs \citegen{Lindvall2004, Cohn2003,
-Boehm2005}.
+by the agile transformation need to be informed and consulted, and the agile
+process must be adjusted according to their needs \citegen{Lindvall2004,
+Cohn2003, Boehm2005}.
 
 Agile methods will also affect management and business related functions. A key
-challenge is that management must move away from life cycle models and towards
+challenge is that management must move away from life-cycle models and towards
 iterative and feature centric models \citegen{Nerur2005}. The focus must be
@@
 The literature review was conducted as an application of Kitchenham's
-method for systematic literature review (SLR) \citegen{Kitchenham2007}. The method
+method for systematic literature reviews (SLR) \citegen{Kitchenham2007}. The method
 was applied to be performed by only one researcher, and to serve as a
@@
 done based on the preliminary search. The following databases were included in
-the searches: IEEExplore (http://ieeexplore.ieee.org/),
-ACM (http://dl.acm.org/), Scopus (http://www.scopus.com/),
-ProQuest \linebreak (http://search.proquest.com/).
-Additionally the archive of International XP
-Conference (http://link.springer.com/) was searched.
+the searches: IEEExplore, ACM, Scopus, ProQuest. Additionally the archive of the
+International XP was searched.
 
@@
 management took responsibility of the change process. Several studies identified
-a champion who was the driver of the change. Some studies reported the change to
+a champion who was the driver for the change. Some studies reported the change to
 be driven by external consultants or a cross organizational work group.
@@
 
-The papers included in this study were mostly experience reports with no
+The papers included in this study were mostly experience reports with no reported
 research method. Some papers were of low quality, but they were included as our

### 2013-01-25 tex/article.tex
Improvement: abstract

@@
 Agile methods have become an appealing alternative for companies striving to
-improve their performance, but they are originally designed for small and
+improve their performance, but the methods are originally designed for small and
 individual teams. This poses challenges when introducing agile to large

### 2013-01-26 tex/article.tex
Spellcheck
Fixed: "confident that search matched majority of relevant material"

@@
 
-\usepackage{calc}      % käytetään laskurien (counter) yhteydessä (tiedot.tex)
+\usepackage{calc}
 \usepackage{natbib}
@@
 
-% Strikethrough
+% Strike through
 \usepackage[normalem]{ulem}
@@
 with more restrictive keywords was unable to capture all possible matches.
-We are however confident that the search covered the majority of material
-relating to agile transformations in large organizations. If a wider search
-should be performed, our recommendation is to only regard the facets
-\textit{agile methods} and \textit{organizational change}. This would lead to
-significantly more matches in the first stage of selecting studies, but the
-additional effort is necessary due to the limitations of a keyword search.
+If a wider search should be performed, our recommendation is to only regard the
+facets \textit{agile methods} and \textit{organizational change}. This would
+lead to significantly more matches in the first stage of selecting studies.
+In our experience a keyword search will not match the large scale facet, but a
+manual semantic analysis of the article meta data is required.
 
@@
 The reasons to initiate a change fell into three categories, as summarized in
-Table \ref{table:motivations}. The primary reasons for change was a strategical
+Table \ref{table:motivations}. The primary reasons for change was a strategic
 need to improve the performance of the organization. Especially shortening the
@@
 
+% Fixed:
+% We are however confident that the search covered the majority of material
+% relating to agile transformations in large organizations.
+%    --> Removed 
+%    --> In our experience a keyword search will not match the large scale
+%    facet, but a manual semantic analysis of the article meta data is required.
+
 % Vielä:
@@
 
-% Pitäs olla done:
-% Search kuva on selitetty myös tekstissä.
+% Should be covered
 % Inclusion: viewpoint on the organizational transformation [from plan-driven, tms.]
 % Used methods: interfacing with various units present in a large organization
-%    --> Developmentin koordinointi
-% "Coaching and community" --> selitä myös coaching
+%    --> Coordination of development
 

### 2013-01-26 tex/article.tex
Stepwise adoption as success factor
Author bias because author is in organization

@@
 
+        \raggedright\rule{0pt}{0.4cm}Stepwise adoption  &
+            Piloting [S4, S6, S10, S11, S26, S28, S30]\newline
+            Adoption spreads gradually [S15, S22, S27, S29]\\
+
         \raggedright\rule{0pt}{0.4cm}Investing in change  &
@@
 
-The papers included in this study were mostly experience reports with no reported
-research method. Some papers were of low quality, but they were included as our
-aim was to give a view of the topic that is broader and inclined towards practice.
+The papers included were mostly experience reports with no reported research
+method. Most authors were in the described organizations, so author bias should
+be expected. Some papers were of low quality, but were included as our aim was
+to give a view of the topic that is broader and inclined towards practice.
 

### 2013-01-26 tex/article.tex
Better articulating "what is large scale"

@@
 organizational transformation on as a general topic were left out. Secondly, the
-size of the organization described in the included primary studies was to be
-large enough to show the challenges of large scale agile development
-\citegen{Lindvall2004}. Further, only studies discussing agile software
-development were included, leaving out for instance studies on agile management.
-Finally, the included studies had to provide a viewpoint on the adoption of
-agile methods.
+papers had to indicate that the development organization is considered being
+large (smallest included reported 50 persons). The presence of several
+development teams also had to be indicated. Further, only studies discussing
+agile software development were included, leaving out for instance studies on
+agile management. Finally, the included studies had to provide a viewpoint on
+the adoption of agile methods.
 

### 2013-01-26 tex/article.tex
Motivation: encourage large organizations to adopt agile

@@
 agile methods. Studies on large scale agile transformations have been published,
-but no extensive or summarizing research is available on the subject. In this
-paper we will give an overview of reported case studies on adopting agile
-methods in large organizations. We will further discuss the motivations for
-adopting agile methods, the characteristics of agile transformations, and
+but no extensive or summarizing research is available on the subject.
+By this literature review we hope to provide evidence that encourages large
+organizations to consider adopting agile methods.
+In this paper we will give an overview of reported case studies on adopting
+agile methods in large organizations. We will further discuss motivations
+for adopting agile methods, characteristics of agile transformations, and
 factors relating to success or challenge in the transformation process. Our
 research is focused on the process of organizational change. We leave out topics
-related to comparing the initial and transformed states of organizations, such
-as measuring the performance of organizations before and after the
-transformation.
+related to comparing initial and transformed states of organizations, such
+as measuring performance before and after the transformation.
 
@@
 bringing higher organizational inertia which slows down organizational change
-\citegen{Livermore2008}. Agile methods are not founded on the use of individual
+\citegen{Livermore2008}. Agile development is not founded on the use of individual
 tools or practices, but rather on a holistic way of thinking.
@@
 the need for formal documentation and thus reduces agility \citegen{Lindvall2004}.
-In addition to inter-team coordination development teams are required to
+In addition to inter-team coordination development teams must
 interact with other organizational units, which are often non-agile in

### 2013-10-28 journal/article.tex
Stub for new article

--- 2013-01-26 tex/article.tex
@@
 
-%\documentclass[12pt,a4paper,finnish,oneside]{article}
-\documentclass[lnbip]{svmultln}
+\documentclass[12pt,a4paper,english,oneside]{article}
+%\documentclass[lnbip]{svmultln}
 
@@
 
-As the competition in software industry is growing companies are constantly
-looking to improve their effectiveness. Agile methods are claimed to increase
-productivity and quality \citegen{Livermore2008}, which makes them attractive
-for companies pursuing better performance. Agile methods give advantage by
-enabling shorter lead times and stronger focus on customer needs
-\citegen{Petersen2010}.
-
-However, large organizations may have problems introducing agile methods
-\citegen{Dyba2009}. As agile methods are initially designed for small teams
-their application has proven to be problematic in a larger scale
-\citegen{Boehm2005}. In a large organization development teams will need to
-interface with other organizational units. Agile methods tend to focus on
-intra-team practices, which leaves it up to the organization to figure out how to
-tailor agile practices to suite a large organization \citegen{Lindvall2004}.
-
-% --> There is demand for a summarizing study of large agile transformations as
-% the evidence is currently scattered.
-% 
-% These models strive for consistent quality and performance
-% by rigorous planning and definition of process. This kind of approach is however
-% badly suited for software development, as projects typically encounter
-% situations that are too hard or impossible to foretell \citegen{Schwaber2002}.
-% The main problems in plan-driven development are the high cost of changes and
-% late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
-% responding to change and distance from customers undermine the  competitiveness
-% of companies. Agile methods are believed to bring remedy to these problems.
-
-The goal of this research is to explore how and why large organizations adopt
-agile methods. Studies on large scale agile transformations have been published,
-but no extensive or summarizing research is available on the subject. In this
-paper we will give an overview of reported case studies on adopting agile
-methods in large organizations. We will further discuss the motivations for
-adopting agile methods, the characteristics of agile transformations, and
-factors relating to success or challenge in the transformation process. Our
-research is focused on the process of organizational change. We leave out topics
-related to comparing the initial and transformed states of organizations, such
-as measuring the performance of organizations before and after the
-transformation.
-
-This paper is organized as follows. The next section gives an overview of large
-scale agile development. Section \ref{sec:method} presents the literature review
-method used for conducting this research. The findings of the review are
-discussed in section \ref{sec:results}, and section \ref{sec:conclusion}
-presents limitations and proposals for future research.
-
 % ---------------------------------------------------------------------
@@
 \label{sec:background}
-
-Introducing agile methods in large organizations is more difficult than it is in
-small organizations \citegen{Dyba2008}. The difficulty is partly related to size
-bringing higher organizational inertia which slows down organizational change
-\citegen{Livermore2008}. Agile methods are not founded on the use of individual
-tools or practices, but rather on a holistic way of thinking.
-Adopting agile often requires change of the entire organizational culture
-\citegen{Misra2010}.
-
-One significant difference between small and large scale adoptions is that large
-organizations have more dependencies between projects and teams. This increases
-the need for formal documentation and thus reduces agility \citegen{Lindvall2004}.
-In addition to inter-team coordination development teams are required to
-interact with other organizational units, which are often non-agile in
-nature. For instance, HR may demand individuals to have strictly specified roles
-in projects \citegen{Boehm2005}, or a change control board may inhibit the use of
-continuous integration or refactoring \citegen{Lindvall2004}. All units affected
-by the agile transformation need to be informed and consulted, and the agile
-process must be adjusted according to their needs \citegen{Lindvall2004,
-Cohn2003, Boehm2005}.
-
-Agile methods will also affect management and business related functions. A key
-challenge is that management must move away from life-cycle models and towards
-iterative and feature centric models \citegen{Nerur2005}. The focus must be
-shifted from a large scale scope to shorter term project planning
-\citegen{Misra2010}, as agile methods emphasize that planning is only meaningful
-for the near future \citegen{Cohn2003}. However, the lack of planning can be a
-concern as business and customer relationships often build on long term
-roadmapping. Enabling operation with shorter term planning requires educating
-stakeholders and reviewing contracting practices \citegen{Boehm2005}.
 
@@
 
-% Research questions:
-%
-% What are the motivations for starting an organizational transformation?
-%
-% What kind of agile transformations have been reported?
-%
-% What were challenges and success factors in the transformation process?
-
-The literature review was conducted as an application of Kitchenham's
-method for systematic literature reviews (SLR) \citegen{Kitchenham2007}. The method
-was applied to be performed by only one researcher, and to serve as a
-preliminary work evaluating the feasibility of a more in-depth review.
-Effort was limited by selecting only the following subset of SLR steps:
-identification of primary studies, data extraction and synthesis. Steps left out
-were the development of a systematic review protocol, strict study quality
-assessment, and reviews in the study selection process.
-
-The searches for primary studies was conducted in two steps. First a preliminary
-search was made to sample the existing literature. Then a systematic search was
-done based on the preliminary search. The following databases were included in
-the searches: IEEExplore, ACM, Scopus, ProQuest. Additionally the archive of the
-International XP was searched.
-
-The search phrases used in the preliminary search were \textit{agile
-transformation} and \textit{large scale agile}. For the systematic search three
-facets and appropriate keywords were chosen, as presented in Table
-\ref{table:searchterms}. To search the databases a query phrase was constructed
-by joining the keywords of each facet by \texttt{OR} operators and the facets by
-\texttt{AND} operators.
-
-As the preliminary search omitted some of the keywords in Table
-\ref{table:searchterms} it yielded a small number of additional matches. The
-reason for variance in matches was that many articles had creative titles or the
-abstracts and keywords were not informative enough. Because of this, the search
-with more restrictive keywords was unable to capture all possible matches.
-If a wider search should be performed, our recommendation is to only regard the
-facets \textit{agile methods} and \textit{organizational change}. This would
-lead to significantly more matches in the first stage of selecting studies.
-In our experience a keyword search will not match the large scale facet, but a
-manual semantic analysis of the article meta data is required.
-
-\begin{table}[t]
-    \begin{tabular}{ l@{ \hskip 0.4cm } l }
-        \toprule
-        Facet                  & Keywords   \\ \midrule
-        Agile methods          & agile, scrum, lean, xp \\ 
-        Organizational change  & transformation, transition, change, migration \\
-        Large organization     & enterprise, organization, (large \texttt{AND} scale) \\
-        \bottomrule
-    \end{tabular}
-    \caption{Facets and related search terms used}
-    \label{table:searchterms}
-\end{table}
-
-\begin{figure}[b]
-  \begin{center}
-    \includegraphics[width=0.45\textwidth]{researchprocess}
-    \caption{Process of selecting primary studies}
-    \label{fig:selection_process}
-  \end{center}
-\end{figure}
-
-The selection process of the primary studies is summarized in Figure
-\ref{fig:selection_process}. The database search yielded 1603 matches which were
-filtered down to 31 selected primary studies. One study was not available, which
-reduced the final number of primary studies to 30.
-The primary studies were selected by four criteria: study type, size of
-organization, use of agile methods in software development, and a transformation
-viewpoint. Firstly, only case studies and experience reports on agile software
-development were included in this study. Papers discussing agile methods or
-organizational transformation on as a general topic were left out. Secondly, the
-size of the organization described in the included primary studies was to be
-large enough to show the challenges of large scale agile development
-\citegen{Lindvall2004}. Further, only studies discussing agile software
-development were included, leaving out for instance studies on agile management.
-Finally, the included studies had to provide a viewpoint on the adoption of
-agile methods.
-
-%\subsection{Data extraction}
-
-Analysis of the primary studies was guided by using a data extraction form with
-open ended fields. The reason for starting a transformation was captured in
-fields \textit{initial state of organization} and \textit{reasons for change}.
-The description of the transformation process was captured in one field.
-Recommendations for transformation processes were captured in fields
-\textit{reported good practices}, \textit{reported challenges}, and
-\textit{lessons learned}.
-Additional fields that were used but not sufficiently answered by the primary
-studies were \textit{satisfaction after transformation} and \textit{reported
-measurements before/after transformation}.
-
 % ---------------------------------------------------------------------
@@
 
-Of the 30 included primary studies 23 were experience reports and 7 were case
-studies. From a validity viewpoint it is important to acknowledge that most of
-the studies were written by employees in the studied organizations.
-
-\begin{figure}[b]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{publicationschart.pdf}
-    \caption{Time of publications and transformations}
-    \label{fig:publications}
-  \end{center}
-\end{figure}
-
-All of the included studies had been published within a decade. On average the
-studies were published three years after the transformation was reported to
-have begun. Figure \ref{fig:publications} gives a timeline view of the
-publications.
-
-The most commonly reported methods are listed in Table \ref{table:methods}, of
-which Scrum was the most popular single method. An important observation is that
-tailoring of methods was very common. The use of tailoring can be considered
-relating particularly to large scale, as the core of agile methods give very
-little advice on interfacing with other units present in a large organization.
-The agile team must interface with units such as other development teams,
-management, and marketing.
-
-\begin{table}[t]
-    \begin{tabular}{ l@{ \hskip 0.4cm } l }
-        \toprule
-        Agile method    & Reported in number of studies   \\ \midrule
-        Scrum           & 18 \\ 
-        XP              & 7 \\
-        Lean            & 5 \\
-        Tailoring and combining methods & 9  (indirect mentioning not counted) \\
-        \bottomrule
-    \end{tabular}
-    \caption{Reported use of agile methods}
-    \label{table:methods}
-\end{table}
-
-The reasons to initiate a change fell into three categories, as summarized in
-Table \ref{table:motivations}. The primary reasons for change was a strategic
-need to improve the performance of the organization. Especially shortening the
-time to market was desired. In many cases the change was started when management
-acknowledged problems in the current way of working.
-
-\begin{table}[t]
-    \begin{tabular}{ l@{ \hskip 0.4cm } l }
-        \toprule
-        Reason                                    & In studies   \\ \midrule
-        Need to improve performance               & S8, S11, S15, S16, S17, S24, S27 \\ 
-        Need to improve time to market            & S6, S9, S13, S19, S24 \\
-        Need to eliminate known process problems  & S5, S6, S10, S13, S19, S21 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Reported reasons for initiating a transformation}
-    \label{table:motivations}
-\end{table}
-
-The most usual way of implementing the change was a stepwise adoption of agile,
-which usually included a pilot project. A typical stepwise adoption was to first
-create agile teams and introduce basic practices. Having intra-team agile
-practices in place made it possible to see how the entire organization would be
-affected by the agile way of working.
-
-The central role of management in the change was evident in many studies. This
-was typically shown in that the change was initiated by the management and
-management took responsibility of the change process. Several studies identified
-a champion who was the driver for the change. Some studies reported the change to
-be driven by external consultants or a cross organizational work group.
-
-Typical challenges and and success factors reported in the studies are listed in
-Table \ref{table:success}. The ability to align the organization was a central
-factor for both success and challenges. Challenges surface if not everyone is on
-board with the change. Good ground for the change is created by strongly
-communicating a clear vision and enabling everyone to participate. Resourcing
-for the change was also seen both as a challenge and a success factor.
-The studies show that allocating extra training and hiring of external
-consultants facilitate the change.
-
-A particular challenge for adopting agile was misunderstanding how agile should
-be applied. Some symptoms of this were talking agile but still working the old
-way, and managers making commitments without hearing development teams.
-As expected, many studies reported that teams had problems interfacing with
-other organizational units, such as management and marketing.
-
-It is difficult to make a new way of working stick in a large organization.
-Use of agile can be rooted into the organization by creating an agile community.
-Founding of interest groups and arranging regular seminars are good ways for
-building community. As agility emphasizes people centricity and flexibility the
-new ways of working can not be mandated. Instead, coaching should be applied for
-leading the change. Several studies suggest demanding conformity in the
-application of agile practices. Agreeing on a common way of working avoids
-disorder when several teams are simultaneously finding their agile ways.
-
-\begin{table}[t]
-    \begin{tabular}{ p{0.31\textwidth} p{0.67\textwidth} }
-        \toprule
-        Class of challenge  & Examples   \\ \midrule
-
-        \raggedright Aligning organization for change  &
-             Lack of management support [S1, S5, S13]\newline
-             Change resistance [S14, S22, S26] \newline
-             Reverting to old model [S4, S11, S23] \\
-
-        \raggedright\rule{0pt}{0.4cm}Misunderstanding the agile model  &
-             Wrong application of roles and responsibilities \newline
-             [S11, S13, S16, S17, S23, S24] \newline
-             Agile terms merely mask the old model [S10, S22] \\
-
-        \raggedright\rule{0pt}{0.4cm}Effect of agile outside of development  &
-            Conflicts with business (e.g. planning) [S6, S11, S21] \newline
-            Making non-development functions agile (testing, maintenance, etc.) [S2, S4, S13]\newline
-            Conflicts with other units (marketing, HR, etc.) \newline[S6, S10, S11, S13] \\
-
-        \raggedright\rule{0pt}{0.4cm}Lack of resources  &
-            Lack of training, insufficient resources for coaching \newline
-            [S5, S13, S22, S28]\\
-
-        \midrule
-        Class of success factors  & Examples   \\ \midrule
-
-        \raggedright Organization alignment and leadership  &
-             Management support [S8, S9, S10, S13, S16, S19] \newline
-             Communicate vision for new process [S7, S19] \newline
-             Demand all organizational units to participate [S13, S15, S17, S18, S19, S21] \\
-
-        \raggedright\rule{0pt}{0.4cm}Coaching and community  &
-             Create community to make change stick [S6, S7, S19, S28] \newline
-             Lead change by a coaching approach [S1, S4, S22, S28] \\
-
-        \raggedright\rule{0pt}{0.4cm}Tailoring and conformity  &
-            Tailor the most suitable model [S8, S11, S20] \newline
-            Conformity clarifies use of new model [S19, S24, S29, S30]\\
-
-        \raggedright\rule{0pt}{0.4cm}Stepwise adoption  &
-            Piloting [S4, S6, S10, S11, S26, S28, S30]\newline
-            Adoption spreads gradually [S15, S22, S27, S29]\\
-
-        \raggedright\rule{0pt}{0.4cm}Investing in change  &
-            Training, consulting, physical spaces, tools [S6, S14, S19]\\
-        \bottomrule
-    \end{tabular}
-    \caption{Challenges and success factors reported in studies}
-    \label{table:success}
-\end{table}
-
 % ---------------------------------------------------------------------
@@
 \label{sec:conclusion}
-
-The papers included were mostly experience reports with no reported research
-method. Most authors were in the described organizations, so author bias should
-be expected. Some papers were of low quality, but were included as our aim was
-to give a view of the topic that is broader and inclined towards practice.
-
-Differing from our initial beliefs, this research shows that a body of studies
-on large scale agile transformations has emerged. We aim to continue our
-research on this topic by conducting a systematic literature review. Beyond
-descriptive reports on agile transformations we found very little studies with
-quantitative comparison of organizations' performance before and after the
-transformations.
-
-% \subsection{TODO}
-
-% Fixed:
-% We are however confident that the search covered the majority of material
-% relating to agile transformations in large organizations.
-%    --> Removed 
-%    --> In our experience a keyword search will not match the large scale
-%    facet, but a manual semantic analysis of the article meta data is required.
-
-% Vielä:
-% Missä sanotaan mikä on "large"
-% Success factor: Stepwise adoption (piloting)
-
-% Should be covered
-% Inclusion: viewpoint on the organizational transformation [from plan-driven, tms.]
-% Used methods: interfacing with various units present in a large organization
-%    --> Coordination of development
 

### 2013-10-28 journal/article.tex
Some text on research process

@@
 
-\documentclass[12pt,a4paper,english,oneside]{article}
-%\documentclass[lnbip]{svmultln}
+%\documentclass[12pt,a4paper,english,oneside]{article}
+\documentclass[lnbip]{svmultln}
 
@@
 
+--> Introductory paragraph
+
+--> Motivation for research
+-- There are no systematic reviews for this
+-- A collection of experience reports can be used for creating a theoretical
+   base.
+
+--> This literature review is intended for researchers as a theroretical
+   and proven foundation for the real problems that organizations face in
+   transformations.
+   This is also intended to be directly usable by practicioners, to serve as a
+   base of general advice with strong motivation (although every organization
+   has unique needs).
+
 % ---------------------------------------------------------------------
@@
 
+\subsection{Research questions}
+
+RQ1 Why are transformations initiated?
+
+RQ2 How do transformations proceed usually; do transformations exhibit patterns?
+
+RQ3 What are typical successes and challenges in an large scale agile
+    transformation process?
+
+\subsection{Research process}
+
+The research was conducted as an application of Kitchenham's
+\cite{Kitchenham2007} guidelines for systematic literature review. The selection
+of primary studies was done as a keyword based database search and manual
+filtering of included studies. The filtering process was executed independently
+by two researchers (Dikert, Paasivaara), and the process was audited by a third
+external researcher (Lassenius). We deviated from Kitchenham's guideline in the
+parts of study assessment and data extraction. These deviations are explained
+below.
+
+The reason for deviating stydy assessment
+--> Original research consisted almost solely of experience reports
+--> The fact that experience reports have a subjective standpoint makes it
+    unreliable to make quantitative conclusions based on the data
+--> We also decided to include separate studies describing the same organization.
+    This was useful as it deepened the understanding of each case, and several
+    large organizations have produced more than one description of their agile
+    transformation.
+
+The fact that the study is based on a quantitative basis
+--> Experie reports have a subjective bias
+--> The data synthesis must not rely on quantitative measures. This means that
+    we also considered topics raised by only a marginal number of reports.
+--> It may still be interesting to make some quantitative observations such as
+    a particular problem being reported in the majority of reports.
+
+Data extraction was implemented partially using data extraction forms, but also
+by a iterative coding approach. (The reason was? ...). Questions for which the
+answer was easy to trace (for example, organization size, time of
+transformation, and RQ1 why are transformations initiated) were answered by
+filling in a data extraction form. For the more complex research questions (RQ2
+how does transformation usually proceed, RQ3 which are typical success factors
+and challenges) the iterative coding approach was used. The iterative coding
+approach...
+--> Possibility to find many different viewpoinits for the research questions
+--> Lots of items, which would have lead to much diversity in data extraction form
+
 % ---------------------------------------------------------------------
@@
 \label{sec:results}
+
+% ---------------------------------------------------------------------
+\section{Discussion}
+\label{sec:discussion}
 

### 2013-10-28 journal/article.tex
Research method

@@
 by two researchers (Dikert, Paasivaara), and the process was audited by a third
-external researcher (Lassenius). We deviated from Kitchenham's guideline in the
-parts of study assessment and data extraction. These deviations are explained
-below.
+external researcher (Lassenius).
+
+We deviated from Kitchenham's guideline in the parts of study assessment and
+data extraction. These deviations are explained below.
+--> This is maybe not true! The study assessment can also be done qualitatively.
+
+Based on the research questions and intended focus of the research, we defined
+an inclusion criteria with three facets: agile software engineering, focus on
+transformation, and large organizations. The first facet covers primary studies
+focusing on software engineering organizations applying or striving to apply
+agile methods. Examples on topics excluded by the first facet are agile
+manufacturing, and applying Scrum practices in management boards. The second
+facet states that included studies must provide insights relevant to
+organizational transformation, and specifically to the research questions.
+The third facet underlines the particular contribution we aim to provide with
+this research, namely the exclusive focus on large organizations.
+
+--> Definition of large scale
+
+--> Case studies ??
+
+A database keyword search was chosen as the search strategy. Based on the
+research questions three facets for keywords were chosen: agile software
+development, organizational transformation, and large scale. Preliminary
+searches showed however that picking keywords with good precision was very
+difficult. As a result we excluded the third facet (large scale). Instead of
+engineering the search terms, which proved to result in excluding some
+relevant articles, we chose an approach of manually filtering the search
+results. Still having a substantial part of the matches in the non-relevant  
+areas such as agile manufacturing, we decided to limit searches to either be
+included in major software publications or include the term ``software''. The
+resulting facets and keywords are listed in Table \ref{table:searchterms}. The
+databases included in the search are listed in Table \ref{table:databases}. All
+databases supported searches using boolean search commands constructed from the
+facets and keywords. The search yielded NNNN unique matches.
+
+\begin{table}
+    \begin{tabular}{ p{0.22\textwidth} p{0.76\textwidth} }
+        \toprule
+        Facet                  & Keywords   \\
+        \midrule
+        Agile methods\newline (before \& after) &
+            agile, scrum, "extreme programming",\newline
+            waterfall, "plan-driven", RUP \\
+        Organizational transformation &
+            transform*, transiti*, migrat*, journey, adopt*, deploy, introduc*,
+            "roll-out", rollout \\
+        Limit to software\newline related atricles &
+            (software OR (conference="agile, xp, icgse, icse"))\newline
+            NOT title+abs="manufacturing", conference="agile manufacturing"
+        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Facets and related search terms used}
+    \label{table:searchterms}
+\end{table}
+
+\begin{table}
+    \begin{tabular}{ p{0.24\textwidth} l l }
+        \toprule
+        Database      & URL                     & N of matches   \\
+        \midrule
+        IEEExplore    & http://ieeexplore.ieee.org      & 745 \\ 
+        ACM           & http://dl.acm.org               & 168 \\
+        Scopus        & http://www.scopus.com/home.url  & 1596 \\
+        Web of Knowledge\newline(Thomson Reuters) &
+        http://apps.webofknowledge.com & 786 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Databases included in search}
+    \label{table:databases}
+\end{table}
 
@@
 --> Lots of items, which would have lead to much diversity in data extraction form
+
+The primary research for this literature review consists almost exclusively of
+industry experience reports. This may be seen as a factor lessening the
+evidence.
+--> Why did we do this anyway?
+--> Case studies with research method studying organizational transformations in
+    software industry are very scarce.
+--> Large organizational changes are hard to measure experimentally (controlled
+    experiments?). Also due to this any student research was excluded.
 

### 2013-10-29 journal/article.tex
Notes on inclusion criteria

@@
 --> This is maybe not true! The study assessment can also be done qualitatively.
+
+\subsection{Inclusion criteria}
+- pitää olla systemaattinen
+- pitää olla toistettava
+==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
+
+There must be more discussion on transformation than a couple of sentences.
+--> Esimerkkejä
+--> Tasot 1--5 ??
+
+Listing of not included topics. Discussion on these topics are considered not
+providing insight to the research questions. Our goal is to address the research
+questions, and specifically focus on analyzing the transformation process.
+Not included topics:
+--> Benefits of agile in general, problems with agile in general
+--> Comparing agile and method X
+--> Reporting on the current method, without consideration of how the method has
+    been evolved (or transformed)
+Tämän perusteella pitää siis kertoa itse prosessista: --> Kirjoita auki tämä
+kriteeri paremmin.
+
+% -----------------------
 

### 2013-10-29 journal/article.tex
Set document style to IST as an example

@@
 %\documentclass[12pt,a4paper,english,oneside]{article}
-\documentclass[lnbip]{svmultln}
+\documentclass[preprint,authoryear,12pt]{elsarticle}
 
@@
 \usepackage{url}
-\usepackage{listings}
+%\usepackage{listings}
 \usepackage{hyphenat}
@@
 
+\journal{Information and Software Technology}
+
 \begin{document}
 
+% Extra definitions?
 \selectlanguage{english}
 
-\mainmatter
+\begin{frontmatter}
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
 
@@
 
-\titlerunning{Adopting agile in the large}
+\author{Kim Dikert, Maria Paasivaara, Casper Lassenius}
 
-\author{Kim Dikert \and Maria Paasivaara \and Casper Lassenius}
-\authorrunning{Kim Dikert et al.}   % abbreviated author list (for running head)
+\address{Aalto University, School of Science, Department of Computer Science and Engineering}
 
-% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Kim Dikert, Maria Paasivaara, Casper Lassenius}
-
-\institute{Aalto University, School of Science,
-Department of Computer Science and Engineering
-\\ \email{kim-karol.dikert@aalto.fi}, \email{maria.paasivaara@aalto.fi}, \email{casper.lassenius@aalto.fi} }
-
-\maketitle
+\ead{kim-karol.dikert@aalto.fi, maria.paasivaara@aalto.fi, casper.lassenius@aalto.fi}
 
 % ---------------------------------------------------------------------
-% -------------- ABSTRACT AND TEXT ------------------------------------
+% ------------------------- ABSTRACT ----------------------------------
 % ---------------------------------------------------------------------
@@
 
-\keywords {agile, transformation, large scale, adopting agile}
 \end{abstract}
 
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+agile \sep transformation \sep large scale \sep adopting agile
+
+\end{keyword}
+
+\end{frontmatter}
+
 % ---------------------------------------------------------------------
+% --------------------------- TEXT ------------------------------------
+% ---------------------------------------------------------------------
+
 \section{Introduction}
@@
 
+Jos koko on epäselvä:
+--> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
+--> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
+    pois
+
 % -----------------------
@@
 \begin{table}
-    \begin{tabular}{ p{0.24\textwidth} l l }
+    \begin{tabular}{ p{0.28\textwidth} l l }
         \toprule
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{elsarticle-harv}
+%\bibliographystyle{plain}
 \bibliographystylegen{plain}

### 2013-10-29 journal/article.tex
Style update; added example bibliography
Some notes

@@
 
-\usepackage{multibib}
-\newcites{gen}{References}
-\usepackage{bibentry}
+% Multiple bibliographies
+%\usepackage{multibib}
+%\newcites{gen}{References}
+%\usepackage{bibentry}
 
@@
 
-\author{Kim Dikert, Maria Paasivaara, Casper Lassenius}
+\author{Kim Dikert}
+\ead{kim-karol.dikert@aalto.fi}
+
+\author{Maria Paasivaara}
+\ead{maria.paasivaara@aalto.fi}
+
+\author{Casper Lassenius}
+\ead{casper.lassenius@aalto.fi}
 
 \address{Aalto University, School of Science, Department of Computer Science and Engineering}
-
-\ead{kim-karol.dikert@aalto.fi, maria.paasivaara@aalto.fi, casper.lassenius@aalto.fi}
 
@@
    base.
+--> There is interest in large scale [ref XP13 workshop]
+
+--> Why is large scale software development different from smaller scale?
 
@@
 
-RQ1 Why are transformations initiated?
+\begin{itemize}
 
-RQ2 How do transformations proceed usually; do transformations exhibit patterns?
+\item
+RQ1: Why are transformations initiated?
 
-RQ3 What are typical successes and challenges in an large scale agile
+\item
+RQ2: How do transformations proceed usually; do transformations exhibit patterns?
+
+\item
+RQ3: What are typical successes and challenges in an large scale agile
     transformation process?
+
+\end{itemize}
 
@@
 
---> Definition of large scale
+A key question in the study was the definition of \emph{large scale}.
+--> Ref: the XP conference
 
@@
 
-%\small
-\fontsize{9pt}{10pt}\selectfont
+\bibliographystyle{elsarticle-harv}
+%\bibliographystylegen{plain}
 
-\bibliographystyle{elsarticle-harv}
-%\bibliographystyle{plain}
-\bibliographystylegen{plain}
+\bibliography{sources_background}
 
-\bibliographygen{sources}
-
-\nobibliography{sources_s}
-
-\subsection*{Primary studies}
-
-\begin{supertabular}{ l p{11.4cm} }
-    {[}S1{]} & \bibentry{S1} \\  \shrinkheight{-1cm}
-    {[}S2{]} & \bibentry{S2} \\ 
-    {[}S3{]} & \bibentry{S3} \\ 
-    {[}S4{]} & \bibentry{S4} \\ 
-    {[}S5{]} & \bibentry{S5} \\ 
-    {[}S6{]} & \bibentry{S6} \\ 
-    {[}S7{]} & \bibentry{S7} \\ 
-    {[}S8{]} & \bibentry{S8} \\ 
-    {[}S9{]} & \bibentry{S9} \\
-    {[}S10{]} & \bibentry{S10} \\ 
-    {[}S11{]} & \bibentry{S11} \\ 
-    {[}S12{]} & \bibentry{S12} \\  \shrinkheight{-3cm}
-    {[}S13{]} & \bibentry{S13} \\ 
-    {[}S14{]} & \bibentry{S14} \\ 
-    {[}S15{]} & \bibentry{S15} \\ 
-    {[}S16{]} & \bibentry{S16} \\ 
-    {[}S17{]} & \bibentry{S17} \\ 
-    {[}S18{]} & \bibentry{S18} \\ 
-    {[}S19{]} & \bibentry{S19} \\ 
-    {[}S20{]} & \bibentry{S20} \\ 
-    {[}S21{]} & \bibentry{S21} \\ 
-    {[}S22{]} & \bibentry{S22} \\ 
-    {[}S23{]} & \bibentry{S23} \\ 
-    {[}S24{]} & \bibentry{S24} \\ 
-    {[}S25{]} & \bibentry{S25} \\ 
-    {[}S26{]} & \bibentry{S26} \\ 
-    {[}S27{]} & \bibentry{S27} \\ 
-    {[}S28{]} & \bibentry{S28} \\ 
-    {[}S29{]} & \bibentry{S29} \\ 
-    {[}S30{]} & \bibentry{S30} \\ 
-\end{supertabular}\end{document}
+\end{document}

### 2013-10-30 journal/article.tex
Notes on large scale

@@
 
+--> Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
+    Practitioners," IEEE Software, pp. 8-9, 2010.
+
 --> Why is large scale software development different from smaller scale?
@@
 \label{sec:background}
+
+Lindvall, M., et al., "Agile Software Development in Large Organizations,"
+IEEE Computer, vol. 37, pp. 26-34, 2004.
+
+Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
+Perplexed: Addison-Wesley, 2003.
+Boehm, B., "Get ready for agile methods, with care," IEEE Computer, vol. 35,
+pp. 64 - 69, 2002.
+--> Any hints on size?? Maybe size leaning towards discipline?
 
@@
 A key question in the study was the definition of \emph{large scale}.
+--> Needs to have complexity in software design?? --> What is complexity in software?
+    Is ``complexity'' an issue in agile?
+    What will bring complexity into software? Not costs or the length of the
+    schedule. --> Having many people will bring the complexity in terms of the
+    SDM to use.
+--> Some people see cost as a definition for large scale, some see code base
+    size [Petersen], some see project duration [Bjarnson?]. These are all
+    related, but do not relate to any specific problem. You can spend X amount
+    of money, but it does not in itself make any difference if the money is
+    spent in an agile way.
+--> Formal definitions (EU, etc.) look at project costs or company turnover.
+    These definitions do not take a standpoint on complexity of the product, so
+    they do not matter in the context of SDMs.
+
 --> Ref: the XP conference
-
---> Case studies ??
+--> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
+--> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
+    Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
+    --> See Boehm also for list of
 

### 2013-11-05 journal/article.tex
Fulltext filtering
Filtering and notes on inclusion criteria

@@
 
+--> Case ``only piloting'': Piloting cases that clearly span only one limited
+    project are excluded. Especially if there is no discussion of the continuity
+    of agile. See Scott 2008.
+
+--> Cases must be empirical, i.e. present a real case. (Otherwise we would need
+    to include textbooks also...)
+
 % -----------------------
@@
     --> See Boehm also for list of
+
+--> The company must already be with N teams or N people, not growing up to the
+    required numbers. Small companies in growth are therefore excluded.
 

### 2013-11-07 journal/article.tex
Filtering; note on include/exclude
Notes on include criteria

@@
 
+--> Case of one agile team in an enterprise: This is unrelated if the single
+    team has no effect on the rest of the enterprise. There is no
+    ``organizational transformation'' if there is just one team soloing.
+
+--> Transformation and ``scaling up'' are very closely related. In some cases
+    they can not really be separated. For instance, if transformation begins
+    with a pilot and spreads gradually through the organization, it is very
+    likely that the process is seen as a ``scaling up'' journey.
+    However, in general, we do not include discussions purely on scaling up,
+    especially if no organizational change is considered.
+
 % -----------------------

### 2013-11-27 journal/article.tex
Coding process

@@
 Based on the research questions and intended focus of the research, we defined
-an inclusion criteria with three facets: agile software engineering, focus on
-transformation, and large organizations. The first facet covers primary studies
+three facets to guide any inclusion/exclusion decisions: agile software
+engineering, focus on transformation, and large organizations. The first facet covers primary studies
 focusing on software engineering organizations applying or striving to apply
@@
 
+\subsection{Coding and synthesis}
+
+An integrated deductive and inductive approach was chosen for coding the primary
+studies \cite{Cruzes2011a}. Initially a list of codes was established for the
+contextual information, as presented in Table \ref{table:contextualcodes}. The
+codes related to the research questions were chosen to be created by an
+inductive procedure. The reason for inducting codes rather than using an a
+priori approach was to avoid the researchers' previous assumptions of the
+research area to affect the choice of codes. To give direction for choosing the
+codes seven different code families were established (see Table
+\ref{table:codefamilies}), including a description defining the scope of the
+families and examples for codes. The example codes would not necessarily be used
+as actual codes, and would not represent the entire or final scope of the
+families.
+
+\begin{table}[h]
+    \begin{tabular}{ p{0.28\textwidth} p{0.7\textwidth} }
+        \toprule
+        Contextual code     & Explanation   \\
+        \midrule
+
+Agile practice & Agile practice or methodology used (eg. Scrum, XP) \\
+
+Business area & The business area in which the organization operates. \\
+
+Organization size & Mentioning the size of the case organization. \\
+
+Time of\newline transformation & When the transformation has been in
+progress, or how long the transformation has taken. Possibly relative to
+the time of the research. \\
+
+Research process & The paper describes a research process. \\
+
+Geographical\newline location & Where the organization is located. Includes
+headquarters and multisite locations. \\
+
+Large scale definition & A definition of "large scale" software development. \\
+
+Measuring results of\newline transformation & Mentioning of how
+transormation progress is measured, or how previous indicators have changed. \\
+
+Satisfaction in agile & Internal or external satisfaction/trust/credibility is
+mentioned. \\
+
+Multisite / GSD & Mentioning of a multisite organization. This does not imply
+that the multisite environment has brought special viewpoints on the agile
+implementation. \\
+
+        \bottomrule
+    \end{tabular}
+    \caption{Context information}
+    \label{table:contextualcodes}
+\end{table}
+
+\begin{table}[h]
+    \begin{tabular}{ p{0.32\textwidth} p{0.65\textwidth} }
+        \toprule
+        Code family          &  Description
+        \\
+        \midrule
+
+        RQ1: Reason to change &
+        Reasons to start the transformation (demand for shorter
+        TTM\footnotemark[1]). \\
+
+		RQ2: Transformation\newline process &
+		Statements that describe the transformation process (top-down, big bang,
+		stepwise). \\
+
+		RQ3: Challenges &
+		Statements that present challenges in the transformation (change resistance).
+		\\
+
+		RQ3: Success factors &
+		Statements that present success factors in the transformation (management
+		support).
+		\\
+
+		Investing in change  &
+		Factors that present how the organization is investing in the
+		transformation (training, consultants, tools). \\
+
+		Practices &
+		Distinguishable practices that are used or established during transformation
+		(coaching, piloting, continuous integration, communities of practice). Other
+		distinguishable factors that are presented as neutral in the primary studies \\
+
+		Other &
+		Contextual codes, other codes not relating to RQ's. \\
+
+        \bottomrule
+    \end{tabular}
+    \caption{Context information}
+    \label{table:codefamilies}
+\end{table}
+
+\footnotetext[1]{Time To Market}
+
+-- How to choose a code?
+
+-- If the primary source makes an indirect statement relating to a code?
+
+-- Considering that mentioning the inclusion of a factor as a success, then
+   would the absence of the same factor be a challenge?
+
+-- If a challenge is described also try to include the resolution in the
+   quotation, if one is available. This should broaded the understanding of the
+   challenges when looking at the quotations only.
+
+The total accumulated codes and quotes of the integrated deductive and inductive
+coding process are presented in Table \ref{table:codecount}.
+
+\begin{table}[h]
+    \begin{tabular}{ l r r }
+        \toprule
+        Code family    &  Codes  &  Quotes
+        \\
+        \midrule
+        RQ1: Reason to change &        10 &   70 \\
+		RQ2: Transformation process &  13 &  538 \\
+		RQ3: Challenges &              39 &  307 \\
+		RQ3: Success factors &         44 &  247 \\
+		Investing in change  &          5 &  130 \\
+		Practices &                    11 &  160 \\
+		Other &                        11 &  194 \\
+		Total &                       133 & 1440\footnotemark[2]
+		\\
+        \bottomrule
+    \end{tabular}
+    \caption{Context information}
+    \label{table:codecount}
+\end{table}
+
+\footnotetext[2]{Due to technical reasons a small number of ``free quotations''
+is included.}
+
 % ---------------------------------------------------------------------

### 2013-11-27 journal/article.tex
Updates on coding

@@
 
+-- Any statement that represents its own concept is coded. The statement becomes
+   a quote.
+
+-- The statements that are quoted must related to the research questions, or
+   contextual codes.
+
+-- The concept and quote is first considered as free standing, i.e. not
+   related to any existing code.
+
+-- If there is an existing code that is close to the new code it is considered
+   to use the existing code.
+
 -- If the primary source makes an indirect statement relating to a code?
+
+%% -----------
 

### 2013-11-27 journal/article.tex
Corrected table captions
Review of code families

@@
 
-Agile practice & Agile practice or methodology used (eg. Scrum, XP) \\
+Agile method & Agile methods used in the organization (eg. Scrum, XP) \\
 
@@
 
-Geographical\newline location & Where the organization is located. Includes
-headquarters and multisite locations. \\
+Geographical\newline location & Where the organization is located. \\
 
@@
 
-Measuring results of\newline transformation & Mentioning of how
-transormation progress is measured, or how previous indicators have changed. \\
-
-Satisfaction in agile & Internal or external satisfaction/trust/credibility is
-mentioned. \\
-
-Multisite / GSD & Mentioning of a multisite organization. This does not imply
-that the multisite environment has brought special viewpoints on the agile
-implementation. \\
+Multisite / GSD & Mentioning of a multisite organization. \\
 
@@
     \end{tabular}
-    \caption{Context information}
+    \caption{Code families}
     \label{table:codefamilies}
@@
 The total accumulated codes and quotes of the integrated deductive and inductive
-coding process are presented in Table \ref{table:codecount}.
+coding process are presented in Table \ref{table:codecount}. The total number of
+quotations is less than the sum of quotations in the categories because a single
+quotation may have multiple codes.
 
@@
         \toprule
-        Code family    &  Codes  &  Quotes
+        Code family    &  Codes  &  Quotations
         \\
@@
 		Practices &                    11 &  160 \\
-		Other &                        11 &  194 \\
-		Total &                       133 & 1440\footnotemark[2]
-		\\
+		Contextual &                    8 &  167 \\
+		Total &                       133 & 1440 \\
         \bottomrule
     \end{tabular}
-    \caption{Context information}
+    \caption{Number of codes and quotations per code family}
     \label{table:codecount}
 \end{table}
-
-\footnotetext[2]{Due to technical reasons a small number of ``free quotations''
-is included.}
 

### 2013-11-27 journal/article.tex
Coding description

@@
 An integrated deductive and inductive approach was chosen for coding the primary
-studies \cite{Cruzes2011a}. Initially a list of codes was established for the
-contextual information, as presented in Table \ref{table:contextualcodes}. The
-codes related to the research questions were chosen to be created by an
-inductive procedure. The reason for inducting codes rather than using an a
-priori approach was to avoid the researchers' previous assumptions of the
-research area to affect the choice of codes. To give direction for choosing the
-codes seven different code families were established (see Table
-\ref{table:codefamilies}), including a description defining the scope of the
-families and examples for codes. The example codes would not necessarily be used
-as actual codes, and would not represent the entire or final scope of the
+studies \cite{Cruzes2011a}. The coding was also designed to have a contextual
+part and a findings part as suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
+list of codes was established for contextual information, as presented in Table
+\ref{table:contextualcodes}. Codes related to the research questions were chosen
+to be created by an inductive process. The reason for inducting codes rather
+than using an a priori approach was to avoid the researchers' previous
+assumptions of the research area to affect the choice of codes. To give
+direction for code induction seven different code families were established (see
+Table \ref{table:codefamilies}). A description defining scope and examples for
+codes were defined for each family. The example codes would not necessarily be
+used as actual codes, and would not represent the entire or final scope of the
 families.
@@
 		Practices &
-		Distinguishable practices that are used or established during transformation
-		(coaching, piloting, continuous integration, communities of practice). Other
-		distinguishable factors that are presented as neutral in the primary studies \\
+		Distinguishable practices that are used or established during transformation.
+		Factors that are presented as neutral in terms of successes and challenges
+		in the primary studies.
+		(coaching, piloting, continuous integration, communities of practice) \\
 
-		Other &
-		Contextual codes, other codes not relating to RQ's. \\
+		Contextual &
+		Contextual codes defined in Table \ref{table:contextualcodes}. \\
 

### 2013-12-03 journal/article.tex
Inter-rater statistics
Inter-rater stats

@@
 
+Inter-rater statistics
+
+                    Maria
+           excl   ???   incl    TOT
+     excl  1578    18    15    1611
+Kimi ???     84    17     8     109
+     incl    51    42    62     155
+     TOT   1713    77    85    1875
+
+???'s =  169
+in/ex = 1706
+
+Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
+Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
+Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
+Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
+
+Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
+
+Pr(a) = (ex-ex + in-in) / tot = 0.8831
+
+k = 0.4063
+
 % ---------------------------------------------------------------------

### 2013-12-13 journal/article.tex
Reasons to change analysis

@@
 
+\subsection{Reasons to change}
+
+Problems
+xxxxxxxxxx     Overhead in process, the process is found to be ineffective
+xxx            Current processes are not taken seriously (followed)
+xxxxxxxxxxxx   Dysfunction in managing software development
+xx             Lack of flexibility
+xxxxxxxxx      Quality issues
+xx             Bad visibility and feedback of projects
+xxxxxxxxxxxxx  Difficulties to schedule or estimate software development, problems with deadlines
+x              Lack of knowledge transfer
+xx             Motivation and morale problems
+xxxx           Difficulties in customer relations and communication
+x              Problems meeting business needs
+
+x  Demand for speed (TTM)
+
 % ---------------------------------------------------------------------

### 2013-12-20 journal/article.tex
Coding test

@@
 
+\subsection{Validation of coding}
+
+Before proceeding with the synthesis we conducted an experiment with the goal to
+reveal the objectiveness of the coding by examining the amount of difference the
+codings of two independent researchers may have. In the experiment Dikert and
+Paasivaara coded independently a sample of five primary documents. The sample
+was hand picked to present a representative subset of the source material. The
+coding was done relying only on the above guidelines, and the researchers were
+careful not to discuss the choice of codes or interpretations relating to codes
+before the experiment was completed. Finally the independent codings of the two
+researchers were analyzed with two metrics: by the number of matching themes
+found, and by the amount of overlap in quotations similar codes have.
+
+The metric of matching themes measures how well the independent researchers are
+able to induct similar codes from the source material. As codes are inducted it
+is likely that two researchers label similar quotes differently. To overcome
+this problem we defined higher level themes under which the individual codes
+were organized. Through discussion we made decisions on whether differently
+labeled codes could be organized under the same theme or not. Table
+\ref{table:codingexperiment} presents the total number of themes the researchers
+found in each document, and how many of the themes were matching. We concluded
+that on average two of three themes that one researcher finds are similarly
+found by the other researcher, but one of three themes are not found by the
+other researcher.
+
+\begin{table}[h]
+    \begin{tabular}{ l r r r }
+        \toprule
+        Document    &  Matching themes  &  Total themes  &  Quotation overlap (0-3) \\
+        \midrule
+        P1          &  14  &  19   &  2.1  \\
+        P2          &   6  &  11   &  2.4  \\
+        P3          &   7  &  11   &  2.1  \\
+        P4          &   2  &   4   &  2.3  \\
+        P5          &  12  &  16   &  2.0  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Result of independent coding experiment}
+    \label{table:codingexperiment}
+\end{table}
+
 % ---------------------------------------------------------------------

### 2013-12-23 journal/article.tex
Explanation for coding test.

@@
 that on average two of three themes that one researcher finds are similarly
-found by the other researcher, but one of three themes are not found by the
+found by the other researcher, but one of three themes is not found by the
 other researcher.
+
+The metric of quotation overlap measures how well the quotations for similar
+codes match between two independent researchers. The intent is to measure how
+codes that are deemed similar by their names match in the actual quoted text. On
+a scale of 0 to 3 we evaluated the quoted text of codes in each theme that
+matched for both researchers, as described in the previous paragraph. A score of
+3 was given for matching codes whose quotation content was virtually the same
+for both researchers. A score of 2 was given if the quotations were mostly the
+same, but other researcher had included one text passage that the other had
+missed. A score of 1 was given if roughly half of the quotations were the same.
+If the quotations had less similarity than this a score of 0 was given. One code
+provided typically 1 to 3 quotations per document. The quotation overlap score
+was counted including only codes that had matching themes. We concluded that the
+average quotation overlap score was 2.2.
 
@@
 \end{table}
+
+Combining the quotation overlap result and the result that inductive coding
+yielded similar themes with a factor of two thirds, we conclude that two
+different researchers should agree completely on at least half of the quotations
+in the source material of our study. This agreement includes both that the
+researchers select the same text passage as a quotation, and that researchers
+agree with the meaning of the quotation. Further interpretation of the
+quotations and codes will still remain a subjective matter.
 

### 2014-01-07 journal/article.tex
Organizing research method

@@
 
+The goal of the research was to aggregate evidence on how large organizations
+transform into an agile way of working. The research was conducted as a
+systematic litterature review, based on empirical studies.
+
 \subsection{Research questions}
+
+The following research questions were set to guide the research process.
 
@@
 
+\subsection{Definition of large scale}
+
+A key question in the study was the definition of \emph{large scale}. Large
+scale requires additional coordination and makes communication difficult
+compared to software development in small teams.
+
+--> Needs to have complexity in software design?? --> What is complexity in software?
+    Is ``complexity'' an issue in agile?
+    What will bring complexity into software? Not costs or the length of the
+    schedule. --> Having many people will bring the complexity in terms of the
+    SDM to use.
+--> Some people see cost as a definition for large scale, some see code base
+    size [Petersen], some see project duration [Bjarnson?]. These are all
+    related, but do not relate to any specific problem. You can spend X amount
+    of money, but it does not in itself make any difference if the money is
+    spent in an agile way.
+--> Formal definitions (EU, etc.) look at project costs or company turnover.
+    These definitions do not take a standpoint on complexity of the product, so
+    they do not matter in the context of SDMs.
+
+--> Ref: the XP conference
+--> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
+--> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
+    Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
+    --> See Boehm also for list of
+
+--> The company must already be with N teams or N people, not growing up to the
+    required numbers. Small companies in growth are therefore excluded.
+
 \subsection{Research process}
@@
 filtering of included studies. The filtering process was executed independently
-by two researchers (Dikert, Paasivaara), and the process was audited by a third
+by two researchers (Dikert, Paasivaara). The selected primary studies were coded
+for data extraction and analysis. The entire process was audited by a third
 external researcher (Lassenius).
 
-We deviated from Kitchenham's guideline in the parts of study assessment and
-data extraction. These deviations are explained below.
---> This is maybe not true! The study assessment can also be done qualitatively.
+We deviated from Kitchenham's guideline on the part of data extraction. Instead
+of using data extraction forms we analyzed the primary studies by coding.
 
-\subsection{Inclusion criteria}
+\subsubsection{Inclusion criteria}
+
+Based on the research questions and intended focus of the research, we defined
+three facets to guide any inclusion/exclusion decisions: agile software
+engineering, focus on transformation, and large organizations. The first facet covers primary studies
+focusing on software engineering organizations applying or striving to apply
+agile methods. Examples on topics excluded by the first facet are agile
+manufacturing, and applying Scrum practices in management boards. The second
+facet states that included studies must provide insights relevant to
+organizational transformation, and specifically to the research questions.
+The third facet underlines the particular contribution we aim to provide with
+this research, namely the exclusive focus on large organizations.
+
+% -----------------------
+
 - pitää olla systemaattinen
@@
 
-% -----------------------
-
-Based on the research questions and intended focus of the research, we defined
-three facets to guide any inclusion/exclusion decisions: agile software
-engineering, focus on transformation, and large organizations. The first facet covers primary studies
-focusing on software engineering organizations applying or striving to apply
-agile methods. Examples on topics excluded by the first facet are agile
-manufacturing, and applying Scrum practices in management boards. The second
-facet states that included studies must provide insights relevant to
-organizational transformation, and specifically to the research questions.
-The third facet underlines the particular contribution we aim to provide with
-this research, namely the exclusive focus on large organizations.
-
-A key question in the study was the definition of \emph{large scale}.
---> Needs to have complexity in software design?? --> What is complexity in software?
-    Is ``complexity'' an issue in agile?
-    What will bring complexity into software? Not costs or the length of the
-    schedule. --> Having many people will bring the complexity in terms of the
-    SDM to use.
---> Some people see cost as a definition for large scale, some see code base
-    size [Petersen], some see project duration [Bjarnson?]. These are all
-    related, but do not relate to any specific problem. You can spend X amount
-    of money, but it does not in itself make any difference if the money is
-    spent in an agile way.
---> Formal definitions (EU, etc.) look at project costs or company turnover.
-    These definitions do not take a standpoint on complexity of the product, so
-    they do not matter in the context of SDMs.
-
---> Ref: the XP conference
---> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
---> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
-    Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
-    --> See Boehm also for list of
-
---> The company must already be with N teams or N people, not growing up to the
-    required numbers. Small companies in growth are therefore excluded.
+\subsubsection{Study selection}
 
@@
 
-\subsection{Coding and synthesis}
+\subsubsection{Coding of primary documents}
 
+The primary documents were coded using the Atlas TI qualitative data analysis software.
 An integrated deductive and inductive approach was chosen for coding the primary
@@
 
-The total accumulated codes and quotes of the integrated deductive and inductive
-coding process are presented in Table \ref{table:codecount}. The total number of
-quotations is less than the sum of quotations in the categories because a single
-quotation may have multiple codes.
+The total accumulated codes and quotations of the integrated deductive and
+inductive coding process are presented in Table \ref{table:codecount}. The total
+number of quotations is less than the sum of quotations in the categories
+because a single quotation may have multiple codes.
 
@@
 
-Inter-rater statistics
+% Inter-rater statistics
+%
+%                     Maria
+%            excl   ???   incl    TOT
+%      excl  1578    18    15    1611
+% Kimi ???     84    17     8     109
+%      incl    51    42    62     155
+%      TOT   1713    77    85    1875
+%
+% ???'s =  169
+% in/ex = 1706
+%
+% Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
+% Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
+% Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
+% Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
+%
+% Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
+%
+% Pr(a) = (ex-ex + in-in) / tot = 0.8831
+%
+% k = 0.4063
 
-                    Maria
-           excl   ???   incl    TOT
-     excl  1578    18    15    1611
-Kimi ???     84    17     8     109
-     incl    51    42    62     155
-     TOT   1713    77    85    1875
+\subsubsection{Validation of coding}
 
-???'s =  169
-in/ex = 1706
-
-Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
-Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
-Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
-Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
-
-Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
-
-Pr(a) = (ex-ex + in-in) / tot = 0.8831
-
-k = 0.4063
-
-\subsection{Validation of coding}
-
-Before proceeding with the synthesis we conducted an experiment with the goal to
-reveal the objectiveness of the coding by examining the amount of difference the
-codings of two independent researchers may have. In the experiment Dikert and
-Paasivaara coded independently a sample of five primary documents. The sample
-was hand picked to present a representative subset of the source material. The
-coding was done relying only on the above guidelines, and the researchers were
-careful not to discuss the choice of codes or interpretations relating to codes
-before the experiment was completed. Finally the independent codings of the two
-researchers were analyzed with two metrics: by the number of matching themes
-found, and by the amount of overlap in quotations similar codes have.
+Coding was done by only one researcher due to resource constraints and volume of
+source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
+an experiment with the goal to assess the consistency of coding that two
+independent researchers may achieve.
+In the experiment Dikert and Paasivaara coded independently a sample of five
+primary documents. The sample was hand picked as a representative subset of the
+source material. The coding was done relying solely on the guidelines presented
+in the previous chapter, and the researchers were careful not to discuss the
+choice of codes or interpretations relating to codes before the experiment was
+completed. Finally the independent codings of the two researchers were analyzed
+with two metrics: by the number of matching themes found, and by the amount of
+overlap in quotations similar codes have.
+% See Kitcenham, chapter 6.4
 

### 2014-01-08 journal/article.tex
Organizing research process chapter

@@
 transform into an agile way of working. The research was conducted as a
-systematic litterature review, based on empirical studies.
+systematic litterature review based on empirical studies.
 
@@
 \item
-RQ2: How do transformations proceed usually; do transformations exhibit patterns?
+RQ2: How do transformations usually proceed; do transformations exhibit patterns?
 
 \item
-RQ3: What are typical successes and challenges in an large scale agile
+RQ3: What are typical successes and challenges in a large scale agile
     transformation process?
@@
 \subsection{Definition of large scale}
+\label{sec:largescale}
 
@@
 We deviated from Kitchenham's guideline on the part of data extraction. Instead
-of using data extraction forms we analyzed the primary studies by coding.
+of using data extraction forms we analyzed the primary studies by coding. This
+deviation was made in order to minimize how possible preconceptions would affect
+the data extraction step, and to use a specialized software tool that allowed us
+to record evidence in as much detail as possible.
+% --> Possibility to find many different viewpoinits for the research questions
+% --> Lots of items, which would have lead to much diversity in data extraction form
 
 \subsubsection{Inclusion criteria}
+\label{sec:inclusioncriteria}
 
@@
 three facets to guide any inclusion/exclusion decisions: agile software
-engineering, focus on transformation, and large organizations. The first facet covers primary studies
-focusing on software engineering organizations applying or striving to apply
-agile methods. Examples on topics excluded by the first facet are agile
-manufacturing, and applying Scrum practices in management boards. The second
-facet states that included studies must provide insights relevant to
-organizational transformation, and specifically to the research questions.
-The third facet underlines the particular contribution we aim to provide with
-this research, namely the exclusive focus on large organizations.
+engineering, focus on transformation, and large scale. The first facet
+covers primary studies focusing on software engineering organizations applying
+or striving to apply agile methods. The second facet states that included
+studies must provide insights relevant to organizational transformation, and
+specifically to the research questions. The third facet underlines the
+particular contribution we aim to provide with this research, namely the
+exclusive focus on large organizations. These three facets were used for
+inclusion and exclusion in each study selection step, including search string
+design, filtering by abstracts, and full text filtering.
+
+Examples on topics excluded by the first facet are agile manufacturing, and
+applying Scrum practices in management boards. In addition, we required that the
+transformation was aimed to introduce agile methods into the studied
+organization.
+
+The facet of \emph{organizational transformation} was interpreted so that the
+primary study presents insights on the process of transformation. Excluded
+topics were: discussing a large agile effort but not presenting the how the
+organization had evolved [ref?], pilot cases that clearly span one limited
+project [Scott 2008?], and case studies that provide minimal information on the
+transformation process [ref?].
+--> Transformation and ``scaling up'' are very closely related. In some cases
+    they can not really be separated. For instance, if transformation begins
+    with a pilot and spreads gradually through the organization, it is very
+    likely that the process is seen as a ``scaling up'' journey.
+    However, in general, we do not include discussions purely on scaling up,
+    especially if no organizational change is considered.
+
+The facet of \emph{large scale} was interpreted as presented in section
+\ref{sec:largescale}.
+--> Case of one agile team in an enterprise: This is unrelated if the single
+    team has no effect on the rest of the enterprise. There is no
+    ``organizational transformation'' if there is just one team soloing.
+
+In addition to the criteria listed above we excluded studies that were not
+empirical. We excluded textbooks, studies presenting theories, and other studies
+that did not include any case organization. We also excluded student experiments
+as it is implausible to simulate the dynamics of a large organization in an
+experiment.
 
@@
 
-- pitää olla systemaattinen
-- pitää olla toistettava
-==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
+% - pitää olla systemaattinen
+% - pitää olla toistettava
+% ==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
+% 
+% Jos koko on epäselvä:
+% --> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
+% --> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
+%     pois
 
-There must be more discussion on transformation than a couple of sentences.
---> Esimerkkejä
---> Tasot 1--5 ??
-
+-- snip --
 Listing of not included topics. Discussion on these topics are considered not
@@
 
-Jos koko on epäselvä:
---> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
---> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
-    pois
+\subsubsection{Preliminary searches}
 
---> Case ``only piloting'': Piloting cases that clearly span only one limited
-    project are excluded. Especially if there is no discussion of the continuity
-    of agile. See Scott 2008.
+Before proceeding with the keyword search a series of preliminary searches were
+performed. Initial searches were made using keywords that were as general as
+possible, including ``agile transformation'' and ``large scale agile''.
+Additionally, a keyword search was done in the selected databases (listed in
+table \ref{table:databases}) using the following search string:
 
---> Cases must be empirical, i.e. present a real case. (Otherwise we would need
-    to include textbooks also...)
+\begin{verbatim}
+((agil* OR scrum OR xp OR lean) AND
+(transform* OR transit* OR change OR migrat*) AND
+((large AND (scale OR organization))) OR enterprise)
+\end{verbatim}
 
---> Case of one agile team in an enterprise: This is unrelated if the single
-    team has no effect on the rest of the enterprise. There is no
-    ``organizational transformation'' if there is just one team soloing.
-
---> Transformation and ``scaling up'' are very closely related. In some cases
-    they can not really be separated. For instance, if transformation begins
-    with a pilot and spreads gradually through the organization, it is very
-    likely that the process is seen as a ``scaling up'' journey.
-    However, in general, we do not include discussions purely on scaling up,
-    especially if no organizational change is considered.
+From these searches we compiled a list of 117 possibly relevant papers that
+should be matched by the keyword searches. The keyword search came to include 75
+of the preliminarily selected papers. The missed 32 preliminary papers were
+separately examined for inclusion, resulting in including 3 additional papers as
+primary studies.
 
@@
 
-A database keyword search was chosen as the search strategy. Based on the
-research questions three facets for keywords were chosen: agile software
-development, organizational transformation, and large scale. Preliminary
-searches showed however that picking keywords with good precision was very
-difficult. As a result we excluded the third facet (large scale). Instead of
-engineering the search terms, which proved to result in excluding some
-relevant articles, we chose an approach of manually filtering the search
-results. Still having a substantial part of the matches in the non-relevant  
-areas such as agile manufacturing, we decided to limit searches to either be
-included in major software publications or include the term ``software''. The
-resulting facets and keywords are listed in Table \ref{table:searchterms}. The
+A keyword search in online databases was chosen as the search strategy. The
 databases included in the search are listed in Table \ref{table:databases}. All
-databases supported searches using boolean search commands constructed from the
-facets and keywords. The search yielded NNNN unique matches.
-
-\begin{table}
-    \begin{tabular}{ p{0.22\textwidth} p{0.76\textwidth} }
-        \toprule
-        Facet                  & Keywords   \\
-        \midrule
-        Agile methods\newline (before \& after) &
-            agile, scrum, "extreme programming",\newline
-            waterfall, "plan-driven", RUP \\
-        Organizational transformation &
-            transform*, transiti*, migrat*, journey, adopt*, deploy, introduc*,
-            "roll-out", rollout \\
-        Limit to software\newline related atricles &
-            (software OR (conference="agile, xp, icgse, icse"))\newline
-            NOT title+abs="manufacturing", conference="agile manufacturing"
-        \\
-        \bottomrule
-    \end{tabular}
-    \caption{Facets and related search terms used}
-    \label{table:searchterms}
-\end{table}
+databases supported use of complex boolean logic in searches which allowed us
+flexibility when constructing the search strings.
 
@@
         Scopus           & http://www.scopus.com/home.url  & 1596 \\
-        Web of Knowledge\newline(Thomson Reuters) &
-        http://apps.webofknowledge.com & 786 \\
+        Web of Knowledge & http://apps.webofknowledge.com  & 786 \\
         \bottomrule
     \end{tabular}
-    \caption{Databases included in search}
+    \caption{Databases included in search, and number of matched articles}
     \label{table:databases}
 \end{table}
+
+We constructed a search string based on the facets \emph{agile software
+deveolpment} and \emph{organizational transformation}, as presented in section
+\ref{sec:inclusioncriteria}. Preliminary searches had showed that picking
+keywords with good precision was difficult. As a result we excluded the facet
+\emph{large scale} from the keyword search, and relied on manual filtering.
+Having a substantial part of the matches in non-relevant areas such as agile
+manufacturing, we considered only articles including the term ``software'' or
+articles published in relevant conferences. Instead of further engineering the
+search terms, which proved to result in excluding some relevant articles, we
+chose to manually filter the remaining search results. The resulting facets and
+keywords are listed in Table \ref{table:searchterms}. The search yielded 1874
+unique matches.
+
+\begin{table}
+    \begin{tabular}{ p{0.22\textwidth} p{0.76\textwidth} }
+        \toprule
+        Facet                  & Keywords   \\
+        \midrule
+        Agile methods\newline (before \& after) &
+            \texttt{agile, scrum, "extreme programming",
+            waterfall, "plan-driven", RUP}\\
+        Organizational transformation &
+            \texttt{transform*, transiti*, migrat*, journey, adopt*, deploy, introduc*,
+            "roll-out", rollout}\\
+        Only software\newline related atricles &
+            \texttt{(software OR (conference="agile, xp, icgse, icse"))
+            AND NOT (title+abs="manufacturing" OR conference="agile manufacturing")}
+        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Facets and related search terms used}
+    \label{table:searchterms}
+\end{table}
+
+\subsubsection{Study quality assessment}
+
+The primary research for this literature review consists almost exclusively of
+industry experience reports. This may be seen as a factor lessening the
+evidence.
+--> Why did we do this anyway?
+--> Case studies with research method studying organizational transformations in
+    software industry are very scarce.
+--> Large organizational changes are hard to measure experimentally (controlled
+    experiments?). Also due to this any student research was excluded.
 
@@
     a particular problem being reported in the majority of reports.
-
-Data extraction was implemented partially using data extraction forms, but also
-by a iterative coding approach. (The reason was? ...). Questions for which the
-answer was easy to trace (for example, organization size, time of
-transformation, and RQ1 why are transformations initiated) were answered by
-filling in a data extraction form. For the more complex research questions (RQ2
-how does transformation usually proceed, RQ3 which are typical success factors
-and challenges) the iterative coding approach was used. The iterative coding
-approach...
---> Possibility to find many different viewpoinits for the research questions
---> Lots of items, which would have lead to much diversity in data extraction form
-
-The primary research for this literature review consists almost exclusively of
-industry experience reports. This may be seen as a factor lessening the
-evidence.
---> Why did we do this anyway?
---> Case studies with research method studying organizational transformations in
-    software industry are very scarce.
---> Large organizational changes are hard to measure experimentally (controlled
-    experiments?). Also due to this any student research was excluded.
 
@@
 
+\subsection{Limitations of research method}
+
+-- A keyword search can not identify the facet of \emph{large scale}, and also
+   \emph{transformation} is difficult to capture with keywords. Because of this
+   we assume that it is likely that a some studies have been missed by the
+   keyword search. As an improvement we would suggest combining the entire
+   archive of \emph{Agile Conference} and \emph{Agile Processes in Software
+   Engineering and Extreme Programming} to the results of the keyword search.
+   These two conferences provided more than two thirds of the pimary sources.
+
+-- Few of the primary studies presented a research method. Most of the studies
+   were experience reports where the author was a member of the organization
+   discussed. Because of this we must assume that the primary studies are
+   biased. Due to this risk of bias we chose not to make statistical
+   interpretations of the results.
+
 % ---------------------------------------------------------------------

### 2014-01-09 journal/article.tex
Study selection process

@@
     However, in general, we do not include discussions purely on scaling up,
-    especially if no organizational change is considered.
+    especially if no organizational change is considered. [Maranzato 2012, Lyon 2008]
 
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/research_process.pdf}
+    \caption{Study selection process}
+    \label{fig:selectionprocess}
+  \end{center}
+\end{figure}
+
 A keyword search in online databases was chosen as the search strategy. The
@@
 chose to manually filter the remaining search results. The resulting facets and
-keywords are listed in Table \ref{table:searchterms}. The search yielded 1874
-unique matches.
+keywords are listed in Table \ref{table:searchterms}.
 
@@
 \end{table}
+
+The database keyword search matched 1874 unique papers. The abstracts of these
+papers were categorized independently by both Dikert and Paasivaara into three
+categories: include, excude, and uncertain. The result was combined by marking
+abstracts both researchers agreed on for inclusion or exclusion respecitvely.
+The inclusion of disagreed or uncertain abstracts was resolved through
+discussion. At this stage papers were excluded only if both reserachers deemed
+it clearly irrelevant, including any uncertain cases for full text filtering. As
+a result 170 papers were selected for full text filtering.
+
+% Inter-rater statistics
+%
+%                     Maria
+%            excl   ???   incl    TOT
+%      excl  1578    18    15    1611
+% Kimi ???     84    17     8     109
+%      incl    51    42    62     155
+%      TOT   1713    77    85    1875
+%
+% ???'s =  169
+% in/ex = 1706
+%
+% Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
+% Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
+% Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
+% Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
+%
+% Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
+%
+% Pr(a) = (ex-ex + in-in) / tot = 0.8831
+%
+% k = 0.4063
+
+Full text filtering was performed by evaluating each article against the three
+facets of the inclusion criteria. Filtering was done in two steps. In the first
+step Dikert extracted data relevant to the three facets of the inclusion
+criteria. Based on the extracted data 76 papers could be immediately deemed as
+included or excluded. The remaining 94 papers were evaluated against the
+inclusion criteria by both Dikert and Paasivaara, and a decision was made after
+discussing each paper separately. In difficult cases Lassenius was consulted to
+reach a decision. As a result 45 papers were selected for full text analysis. In
+addition 2 papers were included from the references of the evaluated papers, and
+3 papers were included from the results of the preliminary search. Finally, 50
+papers were selected as primary studies.
 
@@
 
-% Inter-rater statistics
-%
-%                     Maria
-%            excl   ???   incl    TOT
-%      excl  1578    18    15    1611
-% Kimi ???     84    17     8     109
-%      incl    51    42    62     155
-%      TOT   1713    77    85    1875
-%
-% ???'s =  169
-% in/ex = 1706
-%
-% Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
-% Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
-% Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
-% Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
-%
-% Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
-%
-% Pr(a) = (ex-ex + in-in) / tot = 0.8831
-%
-% k = 0.4063
-
 \subsubsection{Validation of coding}

### 2014-01-09 journal/article.tex
Research method

@@
 
-\begin{figure}
+The study selection process inlcuded three steps, each step refining the set of
+selected papers. First a database search was used to gather an initial set of
+papers. This set was then refined by filtering out abstracts and finally full
+text filtering. The study selection process is outlined in Figure
+\ref{fig:selectionprocess}.
+
+\begin{figure}[b]
   \begin{center}
@@
 
-A keyword search in online databases was chosen as the search strategy. The
-databases included in the search are listed in Table \ref{table:databases}. All
-databases supported use of complex boolean logic in searches which allowed us
-flexibility when constructing the search strings.
+A keyword search in online databases was chosen as the means to find potential
+primary studies. The databases included in the search are listed in Table
+\ref{table:databases}. All databases supported use of complex boolean logic in
+searches which allowed us flexibility when constructing the search strings.
 
@@
 An integrated deductive and inductive approach was chosen for coding the primary
-studies \cite{Cruzes2011a}. The coding was also designed to have a contextual
-part and a findings part as suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
+studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
+and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
 list of codes was established for contextual information, as presented in Table
@@
 than using an a priori approach was to avoid the researchers' previous
-assumptions of the research area to affect the choice of codes. To give
-direction for code induction seven different code families were established (see
-Table \ref{table:codefamilies}). A description defining scope and examples for
-codes were defined for each family. The example codes would not necessarily be
-used as actual codes, and would not represent the entire or final scope of the
-families.
+assumptions of the research area to affect the choice of codes
 
@@
 \end{table}
+
+To give direction for code induction seven different code families were
+established, as presented in Table \ref{table:codefamilies}. A description
+defining scope and examples for codes were defined for each family. The example
+codes would not necessarily be used as actual codes, and would not represent the
+entire or final scope of the families.
 
@@
 
-\begin{table}[h]
+\begin{table}
     \begin{tabular}{ l r r r }

### 2014-01-09 journal/article.tex
Inductive coding description

@@
 Based on the research questions and intended focus of the research, we defined
-three facets to guide any inclusion/exclusion decisions: agile software
-engineering, focus on transformation, and large scale. The first facet
-covers primary studies focusing on software engineering organizations applying
-or striving to apply agile methods. The second facet states that included
-studies must provide insights relevant to organizational transformation, and
-specifically to the research questions. The third facet underlines the
-particular contribution we aim to provide with this research, namely the
-exclusive focus on large organizations. These three facets were used for
-inclusion and exclusion in each study selection step, including search string
-design, filtering by abstracts, and full text filtering.
+three facets to guide inclusion/exclusion decisions: agile software engineering,
+focus on transformation, and large scale. The first facet covers primary studies
+focusing on software engineering organizations applying or striving to apply
+agile methods. The second facet states that included studies must provide
+insights relevant to organizational transformation, and specifically to the
+research questions. The third facet underlines the particular contribution we
+aim to provide with this research, namely the exclusive focus on large
+organizations. These three facets were used for inclusion and exclusion in each
+study selection step, including search string design, filtering by abstracts,
+and full text filtering.
 
@@
 \ref{sec:inclusioncriteria}. Preliminary searches had showed that picking
-keywords with good precision was difficult. As a result we excluded the facet
-\emph{large scale} from the keyword search, and relied on manual filtering.
+keywords with good precision was difficult. Therefore we decided to exclude the
+facet \emph{large scale} from the keyword search, with the consequence of
+increasing the manual filtering effort in the subsequent steps.
 Having a substantial part of the matches in non-relevant areas such as agile
-manufacturing, we considered only articles including the term ``software'' or
-articles published in relevant conferences. Instead of further engineering the
-search terms, which proved to result in excluding some relevant articles, we
-chose to manually filter the remaining search results. The resulting facets and
-keywords are listed in Table \ref{table:searchterms}.
+manufacturing, we included only articles including the term ``software'' or
+articles published in relevant conferences. Instead of engineering the search
+terms any further, which proved to result in excluding some relevant articles,
+we chose to rely more on manual filtering. The resulting facets and keywords are
+listed in Table \ref{table:searchterms}.
 
@@
 
--- How to choose a code?
+The inductive coding was done according to the guidelines described below.
 
--- Any statement that represents its own concept is coded. The statement becomes
-   a quote.
+A passage of text that presents any concept relating to the research questions
+or the contextual codes becomes a quotation. If the quotation corresponds
+closely to a concept that has already been coded the existing code is used.
+Otherwise a new code is created, and the code is assigned to one of the
+predefined code families. The labels of existing codes may need to be adjusted
+when new quotations are assigned. Many codes may be assigned to one quotation
+and quotations of different codes may be overlapping.
 
--- The statements that are quoted must related to the research questions, or
-   contextual codes.
+The length of the text passage that make up a quotation may vary. For the
+contextual codes the quoted passage should be of minimal length. For other
+quotes the length of the quotation should be long enough to highlight the
+context of the concept the quotation, varying from a few sentences to one
+paragraph, or even several paragraphs. As an example of context of the
+quotation, if a transformation challenge is described by a quotation it would be
+good if also the source or resolution of the challenge could be included in the
+quoted text passage. Having enough context should broaden the understanding
+of the coded concept when looking at the quotation in isolation..
 
--- The concept and quote is first considered as free standing, i.e. not
-   related to any existing code.
-
--- If there is an existing code that is close to the new code it is considered
-   to use the existing code.
-
--- If the primary source makes an indirect statement relating to a code?
-
-%% -----------
-
--- Considering that mentioning the inclusion of a factor as a success, then
-   would the absence of the same factor be a challenge?
-
--- If a challenge is described also try to include the resolution in the
-   quotation, if one is available. This should broaded the understanding of the
-   challenges when looking at the quotations only.
+% -- Considering that mentioning the inclusion of a factor as a success, then
+%    would the absence of the same factor be a challenge?
 

### 2014-01-12 journal/article.tex
Notes, small corrections

@@
 \item
-RQ1: Why are transformations initiated?
+RQ1: Why do large organizations initiate an agile transformation?
 
 \item
-RQ2: How do transformations usually proceed; do transformations exhibit patterns?
+RQ2: How do large scale agile transformations usually proceed; do
+     transformations exhibit patterns?
 
@@
 for data extraction and analysis. The entire process was audited by a third
-external researcher (Lassenius).
+researcher (Lassenius).
 
@@
 
+% --> Why are the preliminary searches done?
+
 Before proceeding with the keyword search a series of preliminary searches were
@@
 primary studies.
+% -->  Move this later.
 
@@
 discussing each paper separately. In difficult cases Lassenius was consulted to
-reach a decision. As a result 45 papers were selected for full text analysis. In
-addition 2 papers were included from the references of the evaluated papers, and
-3 papers were included from the results of the preliminary search. Finally, 50
-papers were selected as primary studies.
+reach a decision. As a result of the full text filtering 45 papers were selected
+to be included.
+In addition 2 papers were included from the references of the evaluated papers,
+and 3 papers were included from the results of the preliminary search. Finally,
+50 papers were selected as primary studies.
 

### 2014-01-13 journal/article.tex
Note on multiple papers for a single case

@@
 
+It was quite common that a single transformation case was described by many
+papers. For example, one paper would highlight the viewpoint of developers [ref.
+salesforce?], and another would consider the transformation from user experience
+designers' point of view [ref. salesforce?]. In stead of using the most complete
+paper as suggested by Kitchenham \cite{Kitchenham2007}, we combined the results
+presented in each paper and considered the case as a single unit. By doing this
+we avoided bias caused by duplicate publications, and were enabled to have a
+more detailed view of the case organization.
+
 \subsubsection{Coding of primary documents}

### 2014-01-13 journal/article.tex
Exclusion examples

--- 2014-01-12 journal/article.tex
@@
 Examples on topics excluded by the first facet are agile manufacturing, and
-applying Scrum practices in management boards. In addition, we required that the
-transformation was aimed to introduce agile methods into the studied
-organization.
+applying Scrum practices in management boards, as these do not relate to
+software engineering. In addition, we required that the organizational
+transformation was aimed to introduce agile methods, which excluded other
+development methods than agile (e.g. [Sagesser 2013]), and general talk on
+organizational transformation.
 
 The facet of \emph{organizational transformation} was interpreted so that the
-primary study presents insights on the process of transformation. Excluded
-topics were: discussing a large agile effort but not presenting the how the
-organization had evolved [ref?], pilot cases that clearly span one limited
-project [Scott 2008?], and case studies that provide minimal information on the
-transformation process [ref?].
---> Transformation and ``scaling up'' are very closely related. In some cases
-    they can not really be separated. For instance, if transformation begins
-    with a pilot and spreads gradually through the organization, it is very
-    likely that the process is seen as a ``scaling up'' journey.
-    However, in general, we do not include discussions purely on scaling up,
-    especially if no organizational change is considered. [Maranzato 2012, Lyon 2008]
+primary study must present insights on the process of transformation. Some
+topics that could clearly be excluded were discussing a large agile effort but
+
+-- not presenting the how the organization had evolved [Fitzgerald?],
+   Lyon [Lyon 2008] was excluded as it discusses technical challenges in scaling
+   up an agile process instead of providing organizational viewpoints.
+
+-- pilot cases that clearly span one limited project [Scott 2008?, Mc Cormic 2005],
+
+-- a single team in a large organization [Fulgham, Jensen].
+
+-- Some case studies that provide minimal information on the transformation process [Greening?].
+   [Mishra 2011] discusses agile practices in a large organization, but does not
+   provide viewpoints on the transformation process and introduction of agile methods.
+   Kim [Kim 2012] sounds promising at a start, but merely presents agile tools
+   lacking any transformation description.
+   Also [Hodgkins].
+
+Transformation and ``scaling up'' are very closely related and in some cases
+they are one and the same. For instance, if transformation begins with a pilot
+and spreads gradually through the organization, the process can very well be
+seen as a ``scaling up'' journey. However, we did not include discussions purely
+on scaling up, especially if no organizational change was considered. [Maranzato
+2012, Lyon 2008]
 
@@
     ``organizational transformation'' if there is just one team soloing.
+
+--> The case of difficult decision regarding size [Cloke 2007, Miller 2012].
+--> [Tudor 2006] Exclude, borderline. There is no evidence on size. Many teams
+    are mentioned, but there is still no direct indication of large size.
+
+--> ``The team'' referring to a large organization [Hodgkins]
 
@@
     a particular problem being reported in the majority of reports.
+
+It was quite common that a single transformation case was described by many
+papers. For example, one paper would highlight the viewpoint of developers
+\cite{Fry2007}, and another would consider the transformation from user
+experience designers' point of view \cite{Federoff2009}. Instead of using the
+most complete paper as suggested by Kitchenham \cite{Kitchenham2007}, we
+combined the results presented in each paper and considered the case as a single
+unit. By doing this we avoided bias caused by duplicate publications, and were
+enabled to have a more detailed view of the case organization.
 
@@
 
-\bibliography{sources_background}
+\bibliography{sources_background,sources_primary}
 

### 2014-01-14 journal/article.tex
Text and example references on inclusion criteria

@@
 transformation was aimed to introduce agile methods, which excluded other
-development methods than agile (e.g. [Sagesser 2013]), and general talk on
-organizational transformation.
+development methods than agile \cite{Sagesser2013}, and use of agile
+methods in other contexts than software development \cite{Hodgkins2007}.
 
 The facet of \emph{organizational transformation} was interpreted so that the
-primary study must present insights on the process of transformation. Some
-topics that could clearly be excluded were discussing a large agile effort but
-
--- not presenting the how the organization had evolved [Fitzgerald?],
-   Lyon [Lyon 2008] was excluded as it discusses technical challenges in scaling
-   up an agile process instead of providing organizational viewpoints.
-
--- pilot cases that clearly span one limited project [Scott 2008?, Mc Cormic 2005],
-
--- a single team in a large organization [Fulgham, Jensen].
-
--- Some case studies that provide minimal information on the transformation process [Greening?].
-   [Mishra 2011] discusses agile practices in a large organization, but does not
-   provide viewpoints on the transformation process and introduction of agile methods.
-   Kim [Kim 2012] sounds promising at a start, but merely presents agile tools
-   lacking any transformation description.
-   Also [Hodgkins].
+primary study must present insights on the process of transformation. This
+was the most important facet, as the pivot of this research is to examine
+how the transformations happen. Examples on excluded topics are comparing the
+original and agile development methods \cite{Petersen2010}, discussing use of
+agile in a large organization but not describing how the new methods were
+introduced \cite{Mishra2011}, and merely presenting agile tools in large scale
+use \cite{Kim2012}. These sources were excluded as they did not privide insights
+on organizational transformation.
+% Also [Fitzgerald 2005] on just presenting use of agile methods, no introducing
 
@@
 and spreads gradually through the organization, the process can very well be
-seen as a ``scaling up'' journey. However, we did not include discussions purely
-on scaling up, especially if no organizational change was considered. [Maranzato
-2012, Lyon 2008]
+seen as a ``scaling up'' journey. However, we excluded cases where a initially
+small organization was scaled up \cite{Maranzato2012}, and discussions with a
+process centered viewpoint without describing organizational change
+\cite{Lyon2008}.
+% Also [Greening 2010] on process centered viewpoint without transformation
 
 The facet of \emph{large scale} was interpreted as presented in section
-\ref{sec:largescale}.
---> Case of one agile team in an enterprise: This is unrelated if the single
-    team has no effect on the rest of the enterprise. There is no
-    ``organizational transformation'' if there is just one team soloing.
+\ref{sec:largescale}, having 50 development personel or 6 teams as as the limit
+for large scale. In some cases the source presented very vague indicators on
+size. For instance, we deemed the case presented by Cloke \cite{Cloke2007} to be
+included as there was indications of large scale considerations, although the
+size remained unclear. The case presented by Miller and Haddad \cite{Miller2012}
+was deemed to be excluded as there was no indications whatsoever on size. A
+borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
+no indication of large size, althoug there were many development teams.
+To complicate matters, some papers talked about ``the team'' in singular when
+referring to the organization \cite{Hodgkins2007}, making it clearly nontrivial
+to judge whether an organization was large based on the choice of words of the
+author.
 
---> The case of difficult decision regarding size [Cloke 2007, Miller 2012].
---> [Tudor 2006] Exclude, borderline. There is no evidence on size. Many teams
-    are mentioned, but there is still no direct indication of large size.
+A further example on exclusion by the \emph{large scale} facet were cases with
+large organizations but only a single team adopting agile \cite{Fulgham2011}.
+We considered cases of single teams (although in large organizations) unrelated
+to this research as our focus is on transformation of the entire organization.
+Also piloting cases that resulted only in single teams using agile were excluded
+\cite{Scott2008}.
 
---> ``The team'' referring to a large organization [Hodgkins]
+% Also [Jensen 2003] with only a single team
+% Possibly [Mc Cormic 2005] on single team piloting
 
 In addition to the criteria listed above we excluded studies that were not
-empirical. We excluded textbooks, studies presenting theories, and other studies
-that did not include any case organization. We also excluded student experiments
-as it is implausible to simulate the dynamics of a large organization in an
-experiment.
-
-% -----------------------
-
-% - pitää olla systemaattinen
-% - pitää olla toistettava
-% ==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
-% 
-% Jos koko on epäselvä:
-% --> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
-% --> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
-%     pois
-
--- snip --
-Listing of not included topics. Discussion on these topics are considered not
-providing insight to the research questions. Our goal is to address the research
-questions, and specifically focus on analyzing the transformation process.
-Not included topics:
---> Benefits of agile in general, problems with agile in general
---> Comparing agile and method X
---> Reporting on the current method, without consideration of how the method has
-    been evolved (or transformed)
-Tämän perusteella pitää siis kertoa itse prosessista: --> Kirjoita auki tämä
-kriteeri paremmin.
+empirical. We excluded textbooks, studies merely presenting theories, and other
+studies that did not include any case organization. We also excluded studies of
+benefits or limitations of agile in general. Lastly, we excluded student
+experiments as it is implausible to simulate the dynamics of a large
+organization in an experiment.
 
@@
 50 papers were selected as primary studies.
+
+% -----------------------
+
+% - pitää olla systemaattinen
+% - pitää olla toistettava
+% ==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
+% 
+% Jos koko on epäselvä:
+% --> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
+% --> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
+%     pois
 
@@
 
-\bibliography{sources_background,sources_primary}
+\bibliography{sources_background,sources_primary,sources_excluded}
 

### 2014-01-14 journal/article.tex
Better explanation on preliminary searches

@@
 
-Before proceeding with the keyword search a series of preliminary searches were
-performed. Initial searches were made using keywords that were as general as
-possible, including ``agile transformation'' and ``large scale agile''.
-Additionally, a keyword search was done in the selected databases (listed in
-table \ref{table:databases}) using the following search string:
+Before proceeding with the database search a number of preliminary searches were
+performed. The purpose of the preliminary searches was to create a benchmark of
+potentially relevant papers that should be matched by the actual database
+search. We also wanted to list top ranked hits by trivial keywords that the
+more complex final search string might miss. Initial searches were made using
+keywords that were as general as possible, including ``agile transformation''
+and ``large scale agile''. Additionally, a trial keyword search was done in the
+selected databases (listed in table \ref{table:databases}) using the following
+search string:
 
@@
 
-From these searches we compiled a list of 117 possibly relevant papers that
-should be matched by the keyword searches. The keyword search came to include 75
-of the preliminarily selected papers. The missed 32 preliminary papers were
-separately examined for inclusion, resulting in including 3 additional papers as
-primary studies.
-% -->  Move this later.
+From these preliminary searches we compiled a list of 117 possibly relevant
+papers that should be matched by the actual database search.
 
@@
 
+We also evaluated the keyword search against the benchmark created in the
+preliminary search step, concluding that 75 of the 117 preliminarily selected
+papers matched. The missed 32 preliminary papers were examined, resulting in
+including 3 additional papers as primary sources.
+
 % Inter-rater statistics
@@
 to be included.
-In addition 2 papers were included from the references of the evaluated papers,
-and 3 papers were included from the results of the preliminary search. Finally,
-50 papers were selected as primary studies.
+
+In the full text filtering step the references of all papers were also examined
+for relevance, including 2 referenced papers as primary sources. Most case study
+papers used references very scarcely, mostly referencing well known descriptions
+of agile methods.
+
+Finally, counting 45 papers from the full text filtering step, 3 papers from
+preliminary searches, and 2 papers from references, 50 papers were selected
+as primary studies.
 

### 2014-01-14 journal/article.tex
Choice of words

@@
 by two researchers (Dikert, Paasivaara). The selected primary studies were coded
-for data extraction and analysis. The entire process was audited by a third
-researcher (Lassenius).
+for data extraction and analysis. The entire process was audited and mentored by
+a third researcher (Lassenius).
 
@@
 borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
-no indication of large size, althoug there were many development teams.
+no indication of issues indicating large size, althoug there were several
+development teams.
 To complicate matters, some papers talked about ``the team'' in singular when

### 2014-04-07 journal/article.tex
Introduction paragraph

@@
 
+This section presents the findings of the literature review. In the first
+subsection a meta analysis of the analyzed papers is presented. In the
+subsequent sections findings are presented organized according to the research
+questions. The methods used for extracting the results were described in section
+\ref{sec:method}.
+
+\subsection{Summary of papers}
+
 \subsection{Reasons to change}

### 2014-04-08 journal/article.tex
Correction in the number of included studies

@@
 discussing each paper separately. In difficult cases Lassenius was consulted to
-reach a decision. As a result of the full text filtering 45 papers were selected
+reach a decision. As a result of the full text filtering 47 papers were selected
 to be included.
@@
 
-Finally, counting 45 papers from the full text filtering step, 3 papers from
-preliminary searches, and 2 papers from references, 50 papers were selected
+Finally, counting 47 papers from the full text filtering step, 3 papers from
+preliminary searches, and 2 papers from references, 52 papers were selected
 as primary studies.

### 2014-04-10 journal/article.tex
Explaining combining sources per case organization

@@
 
+In several cases the transformation of a single organization was described in
+more than one primary study. Even if the transformation of a single organization
+was described in many studies all sources that passed the inclusion criterial
+were included. The reason for this was that we wanted to gain as much insight
+from each organization as possible. As a result 40 unique organizations were
+identified in the primary studies.
+
 % -----------------------
@@
 subsequent sections findings are presented organized according to the research
-questions. The methods used for extracting the results were described in section
-\ref{sec:method}.
+questions. The methods used for extracting the results were described in
+section \ref{sec:method}.
 
-\subsection{Summary of papers}
+\subsection{Overview of primary studies}
+
+The results of this study include findings from 52 primary studies presenting
+how 40 different large organizations introduce agile software development.
+
+Time of transformation
+
+Organization size varied 50 -- 15000. Median.
+
+Agile methods used
 

### 2014-04-11 journal/article.tex
Meta-analysis
Journals vs. conferences

@@
 The results of this study include findings from 52 primary studies presenting
-how 40 different large organizations introduce agile software development.
+how 40 different large organizations introduce agile software development. Most
+of the included studies were experience reports (45 studies), and in many cases
+it was evident that the author had been personally involved in the transformation.
+Only 6 of the included studies had a research method explicitly stated. One of
+the studies was an interview article. The publication forums of the primary
+studies were distributed so that 57 sources were conference proceedings, 4
+sources were journal articles, and one source was a technical report.
 

### 2014-04-13 journal/article.tex
Organization size; multiple case studies numbered as separate organizations

@@
 
-In several cases the transformation of a single organization was described in
-more than one primary study. Even if the transformation of a single organization
-was described in many studies all sources that passed the inclusion criterial
-were included. The reason for this was that we wanted to gain as much insight
-from each organization as possible. As a result 40 unique organizations were
-identified in the primary studies.
+In several cases there were more than one primary study presenting the
+transformation of the same organization. Even if the transformation of a single
+organization was described in many studies all sources that passed the inclusion
+criterial were included. Studies presenting the same organization were treated
+as one unit so that we could gain as much insight from each organization as
+possible. There were also a few papers that presented multiple case studies, and
+in those cases we treated each studied organization individually. As a result 42
+unique organizations were identified in the primary studies.
 
@@
 The results of this study include findings from 52 primary studies presenting
-how 40 different large organizations introduce agile software development. Most
-of the included studies were experience reports (45 studies), and in many cases
-it was evident that the author had been personally involved in the transformation.
+how 42 different large software organizations introduce agile methods. Most of
+the included studies were experience reports (45 studies), and in many cases it
+was evident that the author had been personally involved in the transformation.
 Only 6 of the included studies had a research method explicitly stated. One of
@@
 
+The size of the organizations varied from the minimum included size of 50 to
+18,000 people. The median size was 300 people. In 7 studies the size was
+presented in terms of teams, ranging from the minimum of 6 teams to 150 teams.
+The median was 10 teams. In 8 studies there was no direct indication on size
+but the issues discussed revealed that the organization was of large size
+accoding to our definition.
+
 Time of transformation
-
-Organization size varied 50 -- 15000. Median.
 

### 2014-04-14 journal/article.tex
Choice of words in validatino of coding

@@
 found in each document, and how many of the themes were matching. We concluded
-that on average two of three themes that one researcher finds are similarly
-found by the other researcher, but one of three themes is not found by the
-other researcher.
+that on average two of three themes found by one researcher are similarly found
+by the other researcher, but one of three themes is not found by the other
+researcher.
 

### 2014-04-14 journal/article.tex
Transformation time and size graphs
Agile methods used

@@
 but the issues discussed revealed that the organization was of large size
-accoding to our definition.
+accoding to our definition. The distribution of organization sizes is presented
+in figure \ref{fig:organization_size}.
 
-Time of transformation
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/organization_size.pdf}
+    \caption{Distribution of organization sizes in primary studies}
+    \label{fig:organization_size}
+  \end{center}
+\end{figure}
 
-Agile methods used
+Figure \ref{fig:transformation_time} summarizes the publication years of the
+primary studies, and the years when the transformations were reported to have
+started. All studies and transformations were dated after year 2000. There were
+peaks in transformation studies in years 2008 and 2011. The studies were
+typically published two years after the start of the transformation.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/transformation_time.pdf}
+    \caption{Publication years and reported start years of transformations}
+    \label{fig:transformation_time}
+  \end{center}
+\end{figure}
+
+\subsection{Agile methods used}
+
+\begin{table}
+    \begin{tabular}{ l r r r }
+        \toprule
+        Method    &  Reported in studies   \\
+        \midrule
+        Scrum            &  28 \\
+        XP               &  5  \\
+        Scrum with XP    &  6  \\
+        Scrum with Lean  &  7  \\
+        Other            &  4  \\
+        Not specified    &  4  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Agile methods reported in the transformed organization}
+    \label{table:agile_methods}
+\end{table}
+
+The agile methods reported being used in the transformed organizations are
+summarized in table \ref{table:agile_methods}. The most prevalent method was
+Scrum, which was the sole agile method mentioned in 28 studies. The second most
+mentioned agile method was Extreme Programming (XP). Lean software development
+was mentioned in some studies, although combined to Scrum in all cases. Other
+agile methods mentioned were Unified Process, ADM, and Rapid Application
+Development.
+
+It was quite common that organizations sought to combine agile methods.
+Especially Scrum, XP, and Lean software development were used together. In
+addition, many cases mentinoed use of XP practices without explicitly stating XP
+as the process being used (such as test driven development and continuous
+integration). Combining and customizing agile practices was also evident in that
+many organizations viewed the agile method as continuously evolving.
+Organizations evolved the agile methods for instance through retrospectives and
+continuous improvement.
 

### 2014-04-16 journal/article.tex
Business areas; Initial state

@@
 in those cases we treated each studied organization individually. As a result 42
-unique organizations were identified in the primary studies.
+unique organizations were identified in the primary studies. We use the word
+\textit{case} to refer to the individual case organizations, and the word
+\textit{study} to refer to the primary studies.
 
@@
 
+The business areas of the case organizations are presented in table
+\ref{table:businessareas}. Different business areas were represented somewhat
+evenly. Online services was the largest group, including companies
+providing software as a service solutions for businesses, online media players
+for consumers, online services for consumers, and communication software for
+businesses. The second largest group was telecommunications, including companies
+such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
+largest group was enterprise management solution providers with products for
+business process management, project protfolio management, and facility
+management.
+
+\begin{table}
+    \begin{tabular}{ l r r r }
+        \toprule
+        Business area    &  Case organizations   \\
+        \midrule
+		Online services for consumers and business  &  9  \\
+		Telecommunications                          &  7  \\
+		Enterprise management                       &  5  \\
+		Banking and financial services              &  4  \\
+		Healthcare                                  &  3  \\
+		IT services                                 &  2  \\
+		Government                                  &  1  \\
+		Information security                        &  1  \\
+		Not specified                               & 10  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Business areas of case organizations in primary studies}
+    \label{table:businessareas}
+\end{table}
+
 \subsection{Agile methods used}
@@
         \midrule
-        Scrum            &  28 \\
-        XP               &  5  \\
-        Scrum with XP    &  6  \\
-        Scrum with Lean  &  7  \\
+        Scrum            &  25 \\
+        XP               &  4  \\
+        Scrum with XP    &  5  \\
+        Scrum with Lean  &  6  \\
         Other            &  4  \\
-        Not specified    &  4  \\
+        Not specified    &  1  \\
         \bottomrule
@@
     \caption{Agile methods reported in the transformed organization}
-    \label{table:agile_methods}
+    \label{table:agilemethods}
 \end{table}
@@
 The agile methods reported being used in the transformed organizations are
-summarized in table \ref{table:agile_methods}. The most prevalent method was
+summarized in table \ref{table:agilemethods}. The most prevalent method was
 Scrum, which was the sole agile method mentioned in 28 studies. The second most
@@
 agile methods mentioned were Unified Process, ADM, and Rapid Application
-Development.
+Development. In one case the agile method was not named.
 
@@
 
+\subsection{Transformation process}
+
+\subsubsection{The initial state of the organization}
+
+We investigated the initial state of the case organizations before the
+transformation. The typical initial process model was described as waterfall, as
+stated in 25 cases. In several cases the authors characterized the initial state
+of the organization as ``\quote{traditional}''. CMMI was presented as a driver
+for the original process in 4 cases. Other process models the studies named were
+RUP [P19], Prince2 [P51], and Macroscope [P14]. In 9 cases the initial state was
+not described.
+
+The old model was described in most cases briefly and in a neutral sentiment.
+Waterfall development in general was referred to as ``\quote{traditional,
+plan-based methods}'' [P25], ``\quote{traditional phase-based approach}'' [P36],
+or ``\quote{old fashioned process driven waterfall}'' [P23]. Korhonen [P25]
+referred to the model presented by Royce \cite{Royce1970} when describing the
+initial state.
+
+In some cases specifics of the old model was criticized. The characteristics of
+the old process models were described as having ``\quote{many gates and
+sign-offs required by upper levels of management}'' [P9], ``\quote{lot of
+bureaucracy in project management and a strict hierarchy of requirements
+communication}'' [P39], and ``\quote{project managers typically used a
+command-and-control philosophy}'' [P44].
+
+In a few cases the initial process model was less rigid, and closer to the agile
+model that was to be introduced. Examples of such loose process models were
+described: ``\quote{The choice of development methodology is one of any number
+of degrees of freedom granted to teams at Amazon. Development teams are expected
+to solve their problems in the best way possible}'' [P3], ``\quote{the R&D group
+leveraged a loose, waterfall-based process with an entrepreneurial culture}''
+[P16]. Some organizations had even signs of agile development on team level
+although the surrounding organization operated in a waterfall fashion. This was
+described as ``\quote{Agile and Scrum were increasingly popular and emerging at
+a grass-roots level, at a corporate level Yahoo! still embraced a more
+traditional product development process}'' [P10], and ``\quote{some teams had
+applied scrum on the team level, despite the surrounding organization that was
+working with a waterfall software engineering model}'' [P26].
+
+In a few cases the organization initailly lacked strict process. For instance
+Long described the initial state in the following way: ``\quote{Project scope
+was in constant flux. Well-meaning market and product owners routinely inserted
+new requirements during product development.}'' [P30].
+
+Some authors described the developement being originally organized in silos. For
+instance Ranganath describes that ``\quote{Teams operated in silos.
+Collaboration, shared purpose and ownership for outcomes were hard to find}''
+[P40].
+
 % ---------------------------------------------------------------------

### 2014-04-16 journal/article.tex
Fixes to layout

@@
 unique organizations were identified in the primary studies. We use the word
-\textit{case} to refer to the individual case organizations, and the word
-\textit{study} to refer to the primary studies.
+\emph{case} to refer to the individual case organizations, and the word
+\emph{study} to refer to the primary studies.
 
@@
 
+\begin{figure}[b]
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/organization_size.pdf}
+    \caption{Distribution of organization sizes in primary studies}
+    \label{fig:organization_size}
+  \end{center}
+\end{figure}
+
 The size of the organizations varied from the minimum included size of 50 to
@@
 
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/organization_size.pdf}
-    \caption{Distribution of organization sizes in primary studies}
-    \label{fig:organization_size}
-  \end{center}
-\end{figure}
-
 Figure \ref{fig:transformation_time} summarizes the publication years of the
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \begin{center}
@@
 
-The business areas of the case organizations are presented in table
-\ref{table:businessareas}. Different business areas were represented somewhat
-evenly. Online services was the largest group, including companies
-providing software as a service solutions for businesses, online media players
-for consumers, online services for consumers, and communication software for
-businesses. The second largest group was telecommunications, including companies
-such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
-largest group was enterprise management solution providers with products for
-business process management, project protfolio management, and facility
-management.
-
-\begin{table}
-    \begin{tabular}{ l r r r }
+\begin{table}[bh]
+    \begin{tabular}{ l r }
         \toprule
@@
 
+The business areas of the case organizations are presented in table
+\ref{table:businessareas}. Different business areas were represented somewhat
+evenly. Online services was the largest group, including companies
+providing software as a service solutions for businesses, online media players
+for consumers, online services for consumers, and communication software for
+businesses. The second largest group was telecommunications, including companies
+such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
+largest group was enterprise management solution providers with products for
+business process management, project protfolio management, and facility
+management.
+
 \subsection{Agile methods used}
-
-\begin{table}
-    \begin{tabular}{ l r r r }
-        \toprule
-        Method    &  Reported in studies   \\
-        \midrule
-        Scrum            &  25 \\
-        XP               &  4  \\
-        Scrum with XP    &  5  \\
-        Scrum with Lean  &  6  \\
-        Other            &  4  \\
-        Not specified    &  1  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Agile methods reported in the transformed organization}
-    \label{table:agilemethods}
-\end{table}
 
@@
 continuous improvement.
+
+\begin{table}[t]
+    \begin{tabular}{ l r }
+        \toprule
+        Method    &  Reported in studies   \\
+        \midrule
+        Scrum            &  25 \\
+        XP               &  4  \\
+        Scrum with XP    &  5  \\
+        Scrum with Lean  &  6  \\
+        Other            &  4  \\
+        Not specified    &  1  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Agile methods reported in the transformed organization}
+    \label{table:agilemethods}
+\end{table}
 
@@
 stated in 25 cases. In several cases the authors characterized the initial state
-of the organization as ``\quote{traditional}''. CMMI was presented as a driver
-for the original process in 4 cases. Other process models the studies named were
-RUP [P19], Prince2 [P51], and Macroscope [P14]. In 9 cases the initial state was
-not described.
+of the organization as ``traditional''. CMMI was presented as a driver for the
+original process in 4 cases. Other process models the studies named were RUP
+[P19], Prince2 [P51], and Macroscope [P14]. In 9 cases the initial state was not
+described.
 
 The old model was described in most cases briefly and in a neutral sentiment.
-Waterfall development in general was referred to as ``\quote{traditional,
-plan-based methods}'' [P25], ``\quote{traditional phase-based approach}'' [P36],
-or ``\quote{old fashioned process driven waterfall}'' [P23]. Korhonen [P25]
-referred to the model presented by Royce \cite{Royce1970} when describing the
-initial state.
+Waterfall development in general was referred to as ``traditional, plan-based
+methods'' [P25], ``traditional phase-based approach'' [P36], or ``old fashioned
+process driven waterfall'' [P23]. Korhonen [P25] referred to the model presented
+by Royce \cite{Royce1970} when describing the initial state.
 
 In some cases specifics of the old model was criticized. The characteristics of
-the old process models were described as having ``\quote{many gates and
-sign-offs required by upper levels of management}'' [P9], ``\quote{lot of
-bureaucracy in project management and a strict hierarchy of requirements
-communication}'' [P39], and ``\quote{project managers typically used a
-command-and-control philosophy}'' [P44].
+the old process models were described as having ``many gates and sign-offs
+required by upper levels of management'' [P9], ``lot of bureaucracy in project
+management and a strict hierarchy of requirements communication'' [P39], and
+``project managers typically used a command-and-control philosophy'' [P44].
 
@@
 model that was to be introduced. Examples of such loose process models were
-described: ``\quote{The choice of development methodology is one of any number
-of degrees of freedom granted to teams at Amazon. Development teams are expected
-to solve their problems in the best way possible}'' [P3], ``\quote{the R&D group
-leveraged a loose, waterfall-based process with an entrepreneurial culture}''
-[P16]. Some organizations had even signs of agile development on team level
-although the surrounding organization operated in a waterfall fashion. This was
-described as ``\quote{Agile and Scrum were increasingly popular and emerging at
-a grass-roots level, at a corporate level Yahoo! still embraced a more
-traditional product development process}'' [P10], and ``\quote{some teams had
-applied scrum on the team level, despite the surrounding organization that was
-working with a waterfall software engineering model}'' [P26].
+described: ``The choice of development methodology is one of any number of
+degrees of freedom granted to teams at Amazon. Development teams are expected to
+solve their problems in the best way possible'' [P3], ``the R\&D group leveraged
+a loose, waterfall-based process with an entrepreneurial culture'' [P16]. Some
+organizations had even signs of agile development on team level although the
+surrounding organization operated in a waterfall fashion. This was described as
+``Agile and Scrum were increasingly popular and emerging at a grass-roots level,
+at a corporate level Yahoo! still embraced a more traditional product
+development process'' [P10], and ``some teams had applied scrum on the team
+level, despite the surrounding organization that was working with a waterfall
+software engineering model'' [P26].
 
 In a few cases the organization initailly lacked strict process. For instance
-Long described the initial state in the following way: ``\quote{Project scope
-was in constant flux. Well-meaning market and product owners routinely inserted
-new requirements during product development.}'' [P30].
+Long describes the initial state in the following way: ``Project scope was in
+constant flux. Well-meaning market and product owners routinely inserted new
+requirements during product development.'' [P30].
 
-Some authors described the developement being originally organized in silos. For
-instance Ranganath describes that ``\quote{Teams operated in silos.
-Collaboration, shared purpose and ownership for outcomes were hard to find}''
-[P40].
+Some authors describe the developement being originally organized in silos. For
+instance Ranganath describes that ``Teams operated in silos. Collaboration,
+shared purpose and ownership for outcomes were hard to find'' [P40].
 

### 2014-04-21 journal/article.tex
Reasons to change

@@
 
+As stated by research question 1 we investigated the reasons why organizations
+start a transformation process towards agile software development. Most primary
+studies gave reasons for starting the transformation, and only 7 cases of 42 did
+not mention any reasons. In cases giving no specific reason to change it was
+mentioned simply that the management had decided to change the organization.
+
+Some cases reported that the organization had been successful in waterfall
+delivery for a long time before the agile transformation.
+
 Problems

### 2014-04-22 journal/article.tex
Reasons to change table

--- 2014-04-16 journal/article.tex
@@
 
+As stated by Research Question 1 we investigated the reasons organizations start
+a transformation process towards agile software development. In most primary
+reasons for starting the transformation were mentioned, and only 7 cases of 42
+did not mention any reasons. In the cases giving no specific reason to change it
+was simply mentioned that management had decided to change the organization.
+The reasons for organizations to change that were found are presented in table
+\ref{table:reasonstochange}. The reasons to change were mainly demands for
+improvement or identified problems in the current process.
+
+Some cases reported that the organization had been successful in waterfall
+delivery for a long time before the agile transformation.
+
+\begin{table}
+    \begin{tabular}{ p{1.5cm} l l }
+        \toprule
+        Reason    &  Examples  &  Sources   \\
+        \midrule
+
+        \multicolumn{3}{l}{Demand of improved delivery speed or frequency}  \\
+         &  Time To Market must lessen           & x \\
+         &  Demand for frequent deliveries       & x \\
+         &  Too long cycle time before feedback  & x \\
+
+        \multicolumn{3}{l}{General problems with management of software development} \\
+         &  General management dysfunction                & x \\
+         &  Problems with motivation and morale           & x \\
+         &  Problems coordinating a growing organization  & x \\
+
+        \multicolumn{3}{l}{Demand for flexibility or responsiveness}  \\
+         &  Responsiveness to business needs  & x \\
+         &  Adaptation as projects evolve     & x \\
+
+        \multicolumn{3}{l}{Problems with overhead and excess in process}  \\
+         &  Excess, slowness and costs caused by process  & x \\
+         &  The current process is being ignored          & x \\
+
+        \multicolumn{3}{l}{Demand to improve product quality}  \\
+
+        \multicolumn{3}{l}{Problems with planning or keeping schedules}  \\
+
+        \multicolumn{3}{l}{Demand of better customer collaboration}  \\
+         &  Customers push for improvement  & x \\
+
+        \multicolumn{3}{l}{Lack of focus or priorities}  \\
+         &  Business needs or priorities are not regarded  & x \\
+         &  Requirements are not understood                & x \\
+
+        \multicolumn{3}{l}{Communication and collaboration dificulties}  \\
+         &  Communication difficulties across and within projects  & x \\
+         &  Teams work in silos or isolation  & x \\
+
+        \bottomrule
+    \end{tabular}
+    \caption{Summarizing why large organizations initiate an agile transformation (RQ1)}
+    \label{table:reasonstochange}
+\end{table}
+
 Problems

### 2014-04-22 journal/article.tex
Corrections for tables

@@
         RQ1: Reason to change &
-        Reasons to start the transformation (demand for shorter
-        TTM\footnotemark[1]). \\
+        Reasons to start the transformation (demand for faster delivery). \\
 
@@
 \end{table}
-
-\footnotetext[1]{Time To Market}
 
@@
         \midrule
-        RQ1: Reason to change &        10 &   70 \\
-		RQ2: Transformation process &  13 &  538 \\
-		RQ3: Challenges &              39 &  307 \\
-		RQ3: Success factors &         44 &  247 \\
-		Investing in change  &          5 &  130 \\
-		Practices &                    11 &  160 \\
-		Contextual &                    8 &  167 \\
-		Total &                       133 & 1440 \\
+        RQ1: Reason to change &        31 &  124 \\
+		RQ2: Transformation process &  16 &  580 \\
+		RQ3: Challenges &              40 &  323 \\
+		RQ3: Success factors &         44 &  260 \\
+		Investing in change  &          5 &  137 \\
+		Practices &                    11 &  170 \\
+		Contextual &                    8 &  215 \\
+		Total &                       155 & 1576 \\
         \bottomrule
@@
 		Not specified                               & 10  \\
+        \midrule
+		Total                                       & 42  \\
         \bottomrule

### 2014-04-22 journal/article.tex
Reasons to change: Delivery speed, TTM

@@
 
-Some cases reported that the organization had been successful in waterfall
-delivery for a long time before the agile transformation.
-
 \begin{table}
@@
         \multicolumn{3}{l}{Communication and collaboration dificulties}  \\
-         &  Communication difficulties across and within projects  & x \\
+         &  Problems in communication across and within projects  & x \\
          &  Teams work in silos or isolation  & x \\
@@
 
-Problems
-xxxxxxxxxx     Overhead in process, the process is found to be ineffective
-xxx            Current processes are not taken seriously (followed)
-xxxxxxxxxxxx   Dysfunction in managing software development
-xx             Lack of flexibility
-xxxxxxxxx      Quality issues
-xx             Bad visibility and feedback of projects
-xxxxxxxxxxxxx  Difficulties to schedule or estimate software development, problems with deadlines
-x              Lack of knowledge transfer
-xx             Motivation and morale problems
-xxxx           Difficulties in customer relations and communication
-x              Problems meeting business needs
+The most prominent reason for introducing agile methods was the demand to
+deliver software more frequently. Companies saw it critical to reduce the Time
+To Market (TTM) of new features.
 
-x  Demand for speed (TTM)
+% Some cases reported that the organization had been successful in waterfall
+% delivery for a long time before the agile transformation.
 

### 2014-04-23 journal/article.tex
Refactored reasons to change

@@
         \midrule
-        RQ1: Reason to change &        31 &  124 \\
+        RQ1: Reason to change &        30 &  123 \\
 		RQ2: Transformation process &  16 &  580 \\
@@
 		Contextual &                    8 &  215 \\
-		Total &                       155 & 1576 \\
+		Total &                       154 & 1575 \\
         \bottomrule
@@
    biased. Due to this risk of bias we chose not to make statistical
-   interpretations of the results.
+   interpretations of the results. We only use relative measure when comparing
+   results.
 
@@
 \begin{table}
-    \begin{tabular}{ p{1.5cm} l l }
+    \begin{tabular}{ p{1.2cm} p{1.2cm} l }
         \toprule
-        Reason    &  Examples  &  Sources   \\
+        \multicolumn{3}{l}{Reasons to change}  \\
         \midrule
 
-        \multicolumn{3}{l}{Demand of improved delivery speed or frequency}  \\
-         &  Time To Market must lessen           & x \\
-         &  Demand for frequent deliveries       & x \\
-         &  Too long cycle time before feedback  & x \\
-
-        \multicolumn{3}{l}{General problems with management of software development} \\
-         &  General management dysfunction                & x \\
-         &  Problems with motivation and morale           & x \\
-         &  Problems coordinating a growing organization  & x \\
-
-        \multicolumn{3}{l}{Demand for flexibility or responsiveness}  \\
-         &  Responsiveness to business needs  & x \\
-         &  Adaptation as projects evolve     & x \\
-
-        \multicolumn{3}{l}{Problems with overhead and excess in process}  \\
-         &  Excess, slowness and costs caused by process  & x \\
-         &  The current process is being ignored          & x \\
-
-        \multicolumn{3}{l}{Demand to improve product quality}  \\
-
-        \multicolumn{3}{l}{Problems with planning or keeping schedules}  \\
-
-        \multicolumn{3}{l}{Demand of better customer collaboration}  \\
-         &  Customers push for improvement  & x \\
-
-        \multicolumn{3}{l}{Lack of focus or priorities}  \\
-         &  Business needs or priorities are not regarded  & x \\
-         &  Requirements are not understood                & x \\
-
-        \multicolumn{3}{l}{Communication and collaboration dificulties}  \\
-         &  Problems in communication across and within projects  & x \\
-         &  Teams work in silos or isolation  & x \\
+        \multicolumn{3}{l}{External issues: Need to respond to changes in business}  \\
+
+        &  \multicolumn{2}{l}{Demand for more frequent deliveries} \\
+        &  &  Time To Market must lessen  \\
+        &  &  Long cycle times are seen as problematic  \\
+
+        &  \multicolumn{2}{l}{Demand for flexibility or responsiveness} \\
+        &  &  Demand for responsiveness to changing business needs  \\
+        &  &  Adaptation as projects evolve  \\
+
+        &  \multicolumn{2}{l}{Customers demand improvement} \\
+
+        \multicolumn{3}{l}{\rule{0pt}{4ex} Internal issues: Problems in managing software development}  \\
+
+        &  \multicolumn{2}{l}{General problems with management of software development} \\
+        &  &  General management dysfunction  \\
+        &  &  Problems with motivation and morale  \\
+        &  &  Problems coordinating a growing organization  \\
+
+        &  \multicolumn{2}{l}{Problems with overhead and excess in process}  \\
+        &  &  Excess, slowness and costs caused by process  \\
+        &  &  The current process is being ignored  \\
+
+        &  \multicolumn{2}{l}{Demand to improve product quality}  \\
+
+        &  \multicolumn{2}{l}{Problems with planning or keeping schedules}  \\
+
+        &  \multicolumn{2}{l}{Too long cycle time before feedback}  \\
+
+        \multicolumn{3}{l}{\rule{0pt}{4ex} Communication issues}  \\
+
+        &  \multicolumn{2}{l}{Communication and collaboration dificulties}  \\
+        &  &  Problems in communication across and within projects  \\
+        &  &  Teams work in silos or isolation  \\
+
+        &  \multicolumn{2}{l}{Requirements engineering issues}  \\
+        &  &  Requirements are not understood  \\
+        &  &  Unawareness of what is important  \\
+
+        &  \multicolumn{2}{l}{Demand of better customer collaboration}  \\
 
@@
     \end{tabular}
-    \caption{Summarizing why large organizations initiate an agile transformation (RQ1)}
+    \caption{Summarizing why large organizations initiate an agile transformation (RQ1).
+             The table presents a hierarchial view of reasons for the organization to change.}
     \label{table:reasonstochange}
@@
 The most prominent reason for introducing agile methods was the demand to
-deliver software more frequently. Companies saw it critical to reduce the Time
-To Market (TTM) of new features.
+deliver software more frequently. A frequent theme was the desire to reduce Time To Market (TTM).
+
+* TTM, lead time
+
+[P17]
+Faster TTM
+"BMC decided to introduce a new generation of their own successful products before competitors could."
+
+[P17]
+Product needs to be improved quickly.
+The deadline is looming (TTM) "Target delivery of the first release was less than a year away"
+
+[P18]
+Reduce TTM "Ericsson set a corporate R\&D goal to reduce Time to Market"
+Faster cycles became necessary "This forced the PDU to deliver products in much short cycles."
+
+[P18]
+Reduction in lead time, increase in responsiveness, "both appealing and necessary" "in our ‘new’ multimedia market place"
+
+[P32]
+TTM must be reduced
+"BT is facing rapid marketplace changes that require it to radically increase its speed to market."
+
+[P38]
+TTM must be improved, to remain competitive
+"The market was moving faster than the development group was able to deliver."
+
+[P39]
+The TTM needed to be shortened
+The testing time needed to be shortened
+
+[P47]
+There was a desire to reduce TTM
+
+* Long cycle time is seen as a problem in itself
+
+[P2] They THINK that agile will make them deliver faster
+
+[P8]
+There are challenges in achieving faster and more frequent delivery.
+
+[P25]
+Long cycle time is seen as a general problem
+
+[P33]
+Long cycles is seen as a problem in itself
+"the number of releases had slowed to one major release per year"
+
+[P36]
+A "small system" needed to be put in place as fast as possible
+
+[P41]
+The goal was to "deliver better quality faster"
+
+[P16]
+Lack of feedback
+"many people in R\&D had not participated in a major release"
+Release cycles are too long, which is clearly problematic
+
+[P49]
+It takes ``too long to develop''
+It takes ``too long to deploy''
+Too little assurance that the result will be good in the changing marketplace
+"software was taking too long to develop, too long to deploy, and providing too
+little assurance that what was either developed or deployed was really going to
+ferry the corporation through the tumultuous waters of its marketplace"
+
+* Slowness combined with cost
+
+[P36]
+It is realized that development was slow, and needs to be improved
+"Business ... felt the all-too-common frustrations with the pace and cost of development."
+
+[P46]
+Development is "too slow and too expensive across the board"
+
+* Customers or ``business'' demand shorter cycles
+
+[P23]
+Speed of delivery
+"Speed of delivery is of the essence – customers are no longer prepared to wait for 18 months or more for solutions to be delivered."
+
+[P39]
+Project stakeholders demand "as early releases as possible"
+
+[P46]
+Business cared only about speed to value.
+
+[P5]
+New solutions must be delivered on a fast schedule
+"The engineering team had to deliver rapid incremental improvement."
+Software quality needs to be improved.
+"Business pressure for rapid improvement was growing."
+"The organization would be required to demonstrate significant improvement in a very short timeframe. It was clear traditional software development methodologies simply would not work."
+
+* Length delivery is a problem in the competition
+
+[P11]
+"we couldn't do business as usual" --> delivery speed must be increased, to remain competitive
+
+[P19]
+Fear of competitors, something must be done
+Long release cycles/duration are "an alarming state that could enable competitors to gain market share"
+
+* Faster is better ROI
+
+[P8]
+Provide within shorter cycles in order to focus on highest ROI
+
+* Timeliness, accuracy
+
+[P9]
+Product deliveries must be timely, to remain competitive
+
+[P15]
+The number of releases needed to be increased, for accuracy
 

### 2014-04-23 journal/article.tex
Clarifications for cases and studies

@@
 in those cases we treated each studied organization individually. As a result 42
-unique organizations were identified in the primary studies. We use the word
-\emph{case} to refer to the individual case organizations, and the word
-\emph{study} to refer to the primary studies.
+unique organizations were identified in the primary studies. We use the term
+\emph{study} to refer to the primary study publications, and the term
+\emph{case} to refer to an individual case organization that may be described in
+several different studies.
 
@@
 subsequent sections findings are presented organized according to the research
-questions. The methods used for extracting the results were described in
-section \ref{sec:method}.
+questions. The methods used for extracting the results were described in section
+\ref{sec:method}. Note that the term \emph{study} is used when refering to a
+primary study publication, and the term \emph{case} is used when refering to an
+organization which may be described in several studies.
 
@@
 summarized in table \ref{table:agilemethods}. The most prevalent method was
-Scrum, which was the sole agile method mentioned in 28 studies. The second most
+Scrum, which was the sole agile method mentioned in 25 cases. The second most
 mentioned agile method was Extreme Programming (XP). Lean software development
@@
         \toprule
-        Method    &  Reported in studies   \\
+        Method                             &  N  \\
         \midrule
@@
         XP                                 &  4  \\
-        Scrum with XP    &  5  \\
-        Scrum with Lean  &  6  \\
+        Scrum and XP                       &  5  \\
         Other                              &  4  \\
+        Combining Lean to an agile method  &  6  \\
         Not specified                      &  1  \\
@@
     \end{tabular}
-    \caption{Agile methods reported in the transformed organization}
+    \caption{Agile methods reported in number of cases}
     \label{table:agilemethods}

### 2014-04-23 journal/article.tex
Reasons to change

@@
 The reasons for organizations to change that were found are presented in table
-\ref{table:reasonstochange}. The reasons to change were mainly demands for
-improvement or identified problems in the current process.
+\ref{table:reasonstochange}.
 
@@
         &  \multicolumn{2}{l}{Customers demand improvement} \\
+
+        &  \multicolumn{2}{l}{Something must be done to remain competitive} \\
 
@@
 
+The reasons to change were are grouped in three top-level categories: external
+issues, internal issues, and communication issues. External issues are related
+to the business area which the organization is operating in. Internal issues are
+related to the way the organization is managed and software is being developed.
+Communication issues represent problems in communication and coordination in
+software projects.
+
 The most prominent reason for introducing agile methods was the demand to
-deliver software more frequently. A frequent theme was the desire to reduce Time To Market (TTM).
+deliver software more frequently. A frequent statement was that it was necessary
+to reduce the Time To Market (TTM) [P17, P18, P32, P39, P47]. As stated in one
+study ``The market was moving faster than the development group was able to
+deliver'' [P38].
 
@@
 
-[P2] They THINK that agile will make them deliver faster
+[P2]
+They THINK that agile will make them deliver faster
 

### 2014-04-23 journal/article.tex
Reasons to change: external issues

@@
 
+        &  \multicolumn{2}{l}{Something must be done in order to remain competitive} \\
+
+        &  \multicolumn{2}{l}{Customers demand improvement} \\
+
         &  \multicolumn{2}{l}{Demand for flexibility or responsiveness} \\
@@
 
-        &  \multicolumn{2}{l}{Customers demand improvement} \\
-
-        &  \multicolumn{2}{l}{Something must be done to remain competitive} \\
-
         \multicolumn{3}{l}{\rule{0pt}{4ex} Internal issues: Problems in managing software development}  \\
@@
         &  \multicolumn{2}{l}{General problems with management of software development} \\
-        &  &  General management dysfunction  \\
+        &  &  Management dysfunction  \\
         &  &  Problems with motivation and morale  \\
@@
 The most prominent reason for introducing agile methods was the demand to
-deliver software more frequently. A frequent statement was that it was necessary
-to reduce the Time To Market (TTM) [P17, P18, P32, P39, P47]. As stated in one
-study ``The market was moving faster than the development group was able to
-deliver'' [P38].
-
-* TTM, lead time
-
-[P17]
-Faster TTM
-"BMC decided to introduce a new generation of their own successful products before competitors could."
-
-[P17]
-Product needs to be improved quickly.
-The deadline is looming (TTM) "Target delivery of the first release was less than a year away"
-
-[P18]
-Reduce TTM "Ericsson set a corporate R\&D goal to reduce Time to Market"
-Faster cycles became necessary "This forced the PDU to deliver products in much short cycles."
-
-[P18]
-Reduction in lead time, increase in responsiveness, "both appealing and necessary" "in our ‘new’ multimedia market place"
-
-[P32]
-TTM must be reduced
-"BT is facing rapid marketplace changes that require it to radically increase its speed to market."
-
-[P38]
-TTM must be improved, to remain competitive
-"The market was moving faster than the development group was able to deliver."
-
-[P39]
-The TTM needed to be shortened
-The testing time needed to be shortened
-
-[P47]
-There was a desire to reduce TTM
-
-* Long cycle time is seen as a problem in itself
-
-[P2]
-They THINK that agile will make them deliver faster
-
-[P8]
-There are challenges in achieving faster and more frequent delivery.
-
-[P25]
-Long cycle time is seen as a general problem
-
-[P33]
-Long cycles is seen as a problem in itself
-"the number of releases had slowed to one major release per year"
-
-[P36]
-A "small system" needed to be put in place as fast as possible
-
-[P41]
-The goal was to "deliver better quality faster"
-
-[P16]
-Lack of feedback
-"many people in R\&D had not participated in a major release"
-Release cycles are too long, which is clearly problematic
-
-[P49]
-It takes ``too long to develop''
-It takes ``too long to deploy''
-Too little assurance that the result will be good in the changing marketplace
-"software was taking too long to develop, too long to deploy, and providing too
-little assurance that what was either developed or deployed was really going to
-ferry the corporation through the tumultuous waters of its marketplace"
-
-* Slowness combined with cost
-
-[P36]
-It is realized that development was slow, and needs to be improved
-"Business ... felt the all-too-common frustrations with the pace and cost of development."
-
-[P46]
-Development is "too slow and too expensive across the board"
-
-* Customers or ``business'' demand shorter cycles
-
-[P23]
-Speed of delivery
-"Speed of delivery is of the essence – customers are no longer prepared to wait for 18 months or more for solutions to be delivered."
-
-[P39]
-Project stakeholders demand "as early releases as possible"
-
-[P46]
-Business cared only about speed to value.
-
-[P5]
-New solutions must be delivered on a fast schedule
-"The engineering team had to deliver rapid incremental improvement."
-Software quality needs to be improved.
-"Business pressure for rapid improvement was growing."
-"The organization would be required to demonstrate significant improvement in a very short timeframe. It was clear traditional software development methodologies simply would not work."
-
-* Length delivery is a problem in the competition
-
-[P11]
-"we couldn't do business as usual" --> delivery speed must be increased, to remain competitive
-
-[P19]
-Fear of competitors, something must be done
-Long release cycles/duration are "an alarming state that could enable competitors to gain market share"
-
-* Faster is better ROI
-
-[P8]
-Provide within shorter cycles in order to focus on highest ROI
-
-* Timeliness, accuracy
-
-[P9]
-Product deliveries must be timely, to remain competitive
-
-[P15]
-The number of releases needed to be increased, for accuracy
+deliver software faster and more frequently. A common statement was that it was
+necessary to reduce the Time To Market (TTM) [P17, P18, P32, P39, P47]. As
+stated in one study: ``The market was moving faster than the development group
+was able to deliver'' [P38]. Long cycle time, i.e. time from start of project to
+delivery, was seen as a problem in itself in some studies [P25, P36]. Many
+studies indicated that the organizations set as one of the transformation goals
+to deliver software faster [P2, P8]. Other cases described that development was
+generally considered as slow [P36, P46, P49].
+
+Commercial competition and market situation was seen as a trigger for
+transformation. Especially speed of delivery was seen as a means to respond to
+the competition [P9, P38]. As Greening describes: ``The average release duration
+peaked at 41 and 35 months, an alarming state that could enable competitors to
+gain market share.'' [P19]. In many cases the transformation was seen as
+essential for the company to remain competitive [P11, P18, P42, P49]. For
+instance, British Telecom realized that ``From a strategic view, there is little
+alternative than to change the old ways of working'' [P32].
+
+Demands for change and improvement were also expressed by internal and external
+customers. Again, the requests of the customers reflected the need of faster
+development and release cycles [P5, P39]. As described in one study: ``Speed of
+delivery is of the essence – customers are no longer prepared to wait for 18
+months or more for solutions to be delivered'' [P23]. In some cases business
+partners expressed mistrust in the original way of conducting development [P40,
+P46].
+
+% * Timeliness, accuracy
+% 
+% [P15]
+% The number of releases needed to be increased, for accuracy
 

### 2014-04-23 journal/article.tex
Reasons to change: external issues

@@
 
+\subsubsection{External issues}
+
 The most prominent reason for introducing agile methods was the demand to
@@
 gain market share.'' [P19]. In many cases the transformation was seen as
-essential for the company to remain competitive [P11, P18, P42, P49]. For
+essential for the company to remain competitive [P10, P11, P18, P42, P49]. For
 instance, British Telecom realized that ``From a strategic view, there is little
@@
 P46].
+
+Yet another frequent external factor driving for change was the need to
+accomodate change during development. Organizations saw agile as a opportunity
+to improve responsiveness to change in general [P25, P26, P42].
+Some organizations faced difficulties when changes to plans were needed. For
+instance, the organizational struture made adapting to changes difficult [52],
+and making changes created management overhead [P51]. Other organizations saw an
+business advantage if priorities could be changed as required [P8, P39].
+Customers and business representatives were in many cases attributed as the
+sources of change requests, and the responsiveness to customer requests needed
+to be improved [P18, P33, P36, P49].
 

### 2014-04-24 journal/article.tex
Reasons to change: management problems

@@
    results.
+
+-- We avoid numbers in results, such as frequency of a phenomena, except in the
+   meta-analysis section.
 
@@
 
+% Some cases reported that the organization had been successful in waterfall
+% delivery for a long time before the agile transformation.
+
 \begin{table}
@@
         &  \multicolumn{2}{l}{Something must be done in order to remain competitive} \\
+        &  &  Change is required to stay in business  \\
 
@@
         &  \multicolumn{2}{l}{Demand for flexibility or responsiveness} \\
-        &  &  Demand for responsiveness to changing business needs  \\
-        &  &  Adaptation as projects evolve  \\
+        &  &  Demand to accomodate changes efficiently  \\
+        &  &  Better responsiveness to customer requests  \\
 
@@
 was able to deliver'' [P38]. Long cycle time, i.e. time from start of project to
-delivery, was seen as a problem in itself in some studies [P25, P36]. Many
-studies indicated that the organizations set as one of the transformation goals
-to deliver software faster [P2, P8]. Other cases described that development was
-generally considered as slow [P36, P46, P49].
+delivery, was seen as a problem in itself [P25, P36]. Some studies indicated
+that the organizations set as one of the transformation goals to deliver
+software faster [P2, P8]. Other cases described that development was generally
+considered as slow [P36, P46, P49].
 
@@
 
-% * Timeliness, accuracy
-% 
-% [P15]
-% The number of releases needed to be increased, for accuracy
-
-% Some cases reported that the organization had been successful in waterfall
-% delivery for a long time before the agile transformation.
+\subsubsection{Internal issues}
+
+The internal issues relate to how software development is managed in the
+organization. In many organizations the project management practice was in a
+state of dysfunction, and it needed to be corrected.
+Management problems were for instance constantly recurring crisis events [P30],
+managers ignoring prioritization of tasks [P38], and reluctance towards focusing
+on delivering business value [P49].
+Some organizations wanted to improve generic project management practice, such
+as risk management [P8], management of offshoring [P38], and improving the
+quality of life of team members [P44].
+
+Command and control style management where ``a few people, usually those
+farthest from the action, made the decisions without directly involving the
+developers doing the actual work'' [P44] was brought up as a problem. For
+instance, schedules could be dictated by management, and teams had to comply to
+them regardless of the workload [P9].
+
+--> Morale?
+
+--> Pressing team members?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+R: Problems with management of software development
+R: Desire to improve management of software development
+
+* Command-and-control, distance from power
+
+[P9]
+But, project scope and dates were still dictated by upper management.  If a team
+was able to adhere to the project dates, it was commonly met by exhaustive
+heroics from the team.
+
+[P9]
+The senior director saw solutions to the PDP issues in the possibilities of
+Agile too, especially in addressing the command and control brittleness,
+fundamental organization issues, and empowerment of individual contributors.
+
+[P44]
+Primavera’s project managers typically used a command-and-control philosophy—a
+few people, usually those farthest from the action, made the decisions without
+directly involving the developers doing the actual work. Furthermore,
+relationships between the development team and other departments began to
+deteriorate because expectations were seldom met.
+
+[P44]
+In particular, we had to improve the company’s perceptions of its own
+development team, and the team itself had to reduce its deep-rooted dependency
+on the organizational hierarchy and break down functional silos within the
+department.
+
+[P39]
+Last but not least, they wanted to improve collaboration between the product
+management team and developers, and to increase team motivation and commitment.
+
+* Projects simply ``fail''
+
+[P21]
+The software development teams under this division experience many issues and
+challenges during the software development cycle, the main issues are: Failure
+of some projects
+
+% * General dysfunction
+% 
+% [P6]
+% At the time, the most visible symptoms of dysfunction in Yahoo! product
+% development were at the project and team layer (centered around issues of
+% planning, project management, release management, and team interactions), rather
+% than at the technical practices or tools layer.
+% 
+% % [P30]
+% % A crisis process was implemented to address critical issues, enabling teams to
+% % stop other work to focus on critical issues. The result was no fewer than three
+% % “crisis” events per month, with constant change in project schedules.
+% 
+% [P38]
+% Project managers were not forced to communicate and prioritize across projects.
+% It was not clear what, if any, progress to expect on a given project. It was up
+% to the developer to prioritize to the best of his/her ability.
+% 
+% [P38]
+% project managers who were focused on a particular project and not on the number
+% of products actually being produced for the customers
+% 
+% [P49]
+% Long delivery cycles of a year or more were largely the rule, some ending with
+% nothing more to show than documentation. The excuse “they kept changing their
+% mind” became an easy shield to protect the executive or manager in charge from
+% meaningful sanction, despite delivering nothing of business value. Business case
+% projections were rarely (never is a strong word) followed up on to determine if
+% the originally projected ROI was achieved.
+
+% * General project management
+% 
+% [P8]
+% Managing project risks through earlier identification of problems and easier 
+% mitigation of risks;
+% 
+% [P8]
+% Expanding the sourcing options for projects to allow an increased use of offshore 
+% resourced.
+% 
+% [P44]
+% When we decided to improve how we build software and increase the quality of
+% life for everyone on the team, we found our answer in agile software
+% development.
 

### 2014-04-24 journal/article.tex
Reasons to change: management problems

--- 2014-04-23 journal/article.tex
@@
 
+-- We avoid numbers in results, such as frequency of a phenomena, except in the
+   meta-analysis section.
+
 % ---------------------------------------------------------------------
@@
 
-% * Timeliness, accuracy
-% 
-% [P15]
-% The number of releases needed to be increased, for accuracy
-
-% Some cases reported that the organization had been successful in waterfall
-% delivery for a long time before the agile transformation.
+\subsubsection{Internal issues}
+
+The internal issues relate to how software development is managed in the
+organization. In many organizations the project management practice was in a
+state of dysfunction, and it needed to be corrected. Management problems were
+for instance constantly recurring crisis events [P30], managers ignoring
+prioritization of tasks [P38], and reluctance towards focusing on delivering
+business value [P49]. Some cases stated that projects simply failed [P21]. Some
+organizations wanted to improve generic project management practice, such as
+risk management [P8], management of offshoring [P38], and improving the quality
+of life of team members [P44]. Several organizations were growing and
+experienced that the old development method did not scale up [P16, P50, P51].
+There was also a desire to improve team motivation and commitment, and
+collaboration between management and developers [P39].
+
+Command and control style management where ``a few people, usually those
+farthest from the action, made the decisions without directly involving the
+developers doing the actual work'' [P44] was brought up as a problem. For
+instance, schedules could be dictated by management, and teams had to comply to
+them regardless of the workload [P9]. As a result teams were overworked and
+needed stretch far to finish projects on time. Low morale and motivation were
+identified as problems that needed to be fixed, and agile was seen as a solution
+[P25, P30, P40].
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-04-24 journal/article.tex
Reasons to change: management problems

@@
 
+A common problem that was acknowledged in organizations was overhead in process,
+which showed as bureucracy and needless costs. Examples on overhead in process
+were unproductive meetings [P38], sign-offs and gates [P9], excessive focus on
+conformance to process [P40], and change management overhead [P51]. Creating
+excessive documentation instead of business value was seen as a problem in some
+cases [P24, P35].
+From business representatives point of view the processes were unreasonably
+costly, while at the same time insufficiently capable of delivering business
+value [P36, P46, P49]. In one case business realized that customers did not
+request the heavy CMMI based process [P24].
+A heavy process resulted often in ignoring the process, only paying ``lip
+service'' to the process, or applying its steps retroactively [P6, P9, P49].
+Seffernick described how applying the waterfall process had teams ``still
+developing the same old way but slower because of the additional process''
+[P46].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-04-24 journal/article.tex
Internal issues

@@
 As stated by Research Question 1 we investigated the reasons organizations start
-a transformation process towards agile software development. In most primary
-reasons for starting the transformation were mentioned, and only 7 cases of 42
-did not mention any reasons. In the cases giving no specific reason to change it
-was simply mentioned that management had decided to change the organization.
-The reasons for organizations to change that were found are presented in table
+a transformation process towards agile software development. Most primary
+studies mentioned reasons for starting the transformation, and only 7 cases of
+42 did not mention any reasons. In the cases giving no specific reason to change
+it was simply mentioned that management had decided to change the organization.
+The identified reasons for organizations to change are presented in table
 \ref{table:reasonstochange}.
@@
 
+        &  \multicolumn{2}{l}{Problems with planning or keeping schedules}  \\
+
+        &  \multicolumn{2}{l}{Too long cycle time before feedback}  \\
+
         &  \multicolumn{2}{l}{Demand to improve product quality}  \\
-
-        &  \multicolumn{2}{l}{Problems with planning or keeping schedules}  \\
-
-        &  \multicolumn{2}{l}{Too long cycle time before feedback}  \\
 
@@
 of life of team members [P44]. Several organizations were growing and
-experienced that the old development method did not scale up [P16, P50, P51].
+experienced that the old development method did not scale up [P16, P30, P50, P51].
 There was also a desire to improve team motivation and commitment, and
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A slow process with long cycle times leading to late feedback was a problem in
+the old model. A large investment was required before results were visible [P5],
+and there was variances in expectations and what was delivered [P40]. A related
+problem was late integration which caused extra work [P25]. As organizations
+grew and products became more complex test cycles began to lengthen, which
+hindered flexibility and reduced speed [P33, P50].
+
+Yet another problem in managing development was problems with schedules.
+Some cases described schedules as unpredictable and estimation being difficult
+[P16, P30, P40, P51]. In other cases it was reported that release deadlines were
+often missed [P19, P21, P33, P41].
+
+Finally, the original process model was seen as a cause for quality problems.
+Several studies mentioned a demand for better quality, and quality improvement
+was set as a goal for the agile transformation [P25, P31, P41, P42]. Improvement
+in external quality and customer satisfaction were also anticipated from agile
+development [P2, P4]. Problems with technical debt were attributed to the
+original process. Some cases reported on prioritizing new feature development
+over bugfixes or fixing only the easiest issues [P30, P38].
 

### 2014-04-25 journal/article.tex
Communication and collaboration issues

@@
 
-        \multicolumn{3}{l}{\rule{0pt}{4ex} Communication issues}  \\
+        \multicolumn{3}{l}{\rule{0pt}{4ex} Communication and collaboration issues}  \\
 
@@
 related to the way the organization is managed and software is being developed.
-Communication issues represent problems in communication and coordination in
-software projects.
+Communication and collaboration issues are related to internal and external
+collaboration. Change was motivated by known problems in these categories, such
+as known process dysfunctions or communication problems. There was also internal
+and external demand for improvement, for instance in speed of delivery and
+quality.
 
@@
 
-Yet another frequent external factor driving for change was the need to
+Yet another frequent external factor driving for improvement was the need to
 accomodate change during development. Organizations saw agile as a opportunity
@@
 farthest from the action, made the decisions without directly involving the
-developers doing the actual work'' [P44] was brought up as a problem. For
+developers doing the actual work'' was highlighted as a problem [P44]. For
 instance, schedules could be dictated by management, and teams had to comply to
@@
 
+\subsubsection{Communication and collaboration issues}
+
+Many organizations experienced problems in communication and collaboration
+between projects, teams, and even members of a single team [P38, P21].
+Teams were in many cases organized around components and were working in silos.
+Working in silos diluted focus and made adapting to change difficult, as teams
+could only focus on their limited area of expertise [P7]. Functional silos acted
+against creating a shared purpose in operating [P40], and resulted in a ``ball
+over the wall'' mentality [P46]. Distance between teams was also created by
+document driven communication, and was causing misunderstanding of requirements
+[P24].
+
+Other communication issues were related to requirements handling and difficulty
+to focus on the most important aspects of a delivery.
+Examples on problems with requirements was too much focus on the documenting
+aspect instead of providing understanding [P24, P49], and ignoring the fact that
+requirements and understanding are evolving during a project [P4].
+In some cases the organization of work and management was such that it hindered
+focusing the development [P7], and prioritization of what to work on was not
+done efficiently [P38, P39].
+
+--> Demand of better customer collaboration
+
+A few organizations were experiencing rise of agile methods at grass roots
+level, but early use of agile resluted in problems when the organization used a
+waterfall method on the top level. Tensions arose when teams using different
+development approaches attempted to collaborate on products [P10]. On a team
+level agile had enabled some flexibility, but that flexibility could not be
+leveraged on the scale of the entire organization [P26]. This kind of
+experiences drove the organization to deploy agile on a full scale.
+
+%%%%%%%%%%%%%%
+
+% ----------------------------------------------------------------
+
 \subsection{Transformation process}

### 2014-04-25 journal/article.tex
Communication and collaboration issues

@@
 
---> Demand of better customer collaboration
+Too little customer and user collaboration led to problems in some
+organizations. For instance, too late user involvment led to change requests
+late in the project [P40]. There was a wish to collaborate more frequently and
+in less formal ways [P4, P26].
 
 A few organizations were experiencing rise of agile methods at grass roots
-level, but early use of agile resluted in problems when the organization used a
-waterfall method on the top level. Tensions arose when teams using different
-development approaches attempted to collaborate on products [P10]. On a team
-level agile had enabled some flexibility, but that flexibility could not be
-leveraged on the scale of the entire organization [P26]. This kind of
-experiences drove the organization to deploy agile on a full scale.
-
-%%%%%%%%%%%%%%
-
-% ----------------------------------------------------------------
+level, but the agile teams could not collaborate efficiently with other parts of
+a waterfall organization. Tensions arose when teams using different development
+approaches attempted to work together on products [P10]. On a team level agile
+had enabled some flexibility, but that flexibility could not be leveraged on the
+scale of the entire organization [P26]. This kind of experiences drove the
+organization to deploy agile on a full scale.
 

### 2014-04-27 journal/article.tex
Challenges

@@
 
+\subsubsection{Modes of transformation}
+
+Top down
+
+Bottom up
+
+Big bang
+
+Transitional
+
+\subsection{Challenges in transformation}
+
+Any organizational transformation involving numerous individuals will face some
+challenges. As stated by Research Question 3, we investigated what are typical
+challenges that large organizations introducing agile methods face.
+
+Challenges can be organized on a high level into three categories. First we
+discuss challenges relating to managing the transformation. These challenges
+relate to managing people and resourcing. Secondly we present challenges in
+choosing and establishing a suitable agile model. Thirdly challenges related to
+large organization size are presented. Size will bring challenges such as
+coordinating work between many teams and adapting other functions than
+development to agile.
+
+\subsubsection{Challenges caused by management of the transformation}
+
+In the beginning of the transformation the organization has realized the need to
+change and is going towards the initial push.
+In many cases the initiative to change came from management, but when presented
+wrong the initiative becomes a mandate. Spayd summarizes this as:
+``Organizations do not change merely because the boss says so, at least not in
+the way that is intended'' [P49]. In that organization the introduction of CMM
+was carried out consistently by a mandate, but the collaborative nature of agile
+did not comply with such a mindset. A top down mandate may also dilute the
+understanding of the reasons for starting the transformation, and further for
+agile development overall [P2].
+
+Organizations had bad experiences with prescribing how to carry out Scrum
+practices. For instance, doing agile based on checklists would not work [P41].
+Benefield described that it was challenging to find a balance between
+independence for teams to choose their way and diluting the value and structure
+of Scrum [P6]. Giving room for team autonomy was seen
+
+% Lack of plans (a guess -- not included)?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
 % ---------------------------------------------------------------------

### 2014-04-28 journal/article.tex
Managing change

@@
 
-\subsubsection{Challenges caused by management of the transformation}
+\subsubsection{Leadership challenges in the transformation}
 
@@
 
+* Other problems with leading change
+
+[P43]  Failing to set goals
+
+Our greatest failures have occurred during instances where we have not set down
+widely-communicated expectations.  We attempted to compel test-driven
+development through the scrum teams during our third and fourth agile releases.
+We failed, though, to set down precisely what our expectations were for our
+development teams and to police and reward or correct developers based on
+results.  As such, we had very spotty results with TDD, with some teams
+succeeding in implementing it, but most not meeting with the level of success
+for which we had hoped.  We set up a practice where we asked teams to create
+User Acceptance Testing criteria, to guide development and UAT, but we stopped
+enforcing it after a few iterations. Product Owners started delivering feedback
+that UATs were becoming less focused and far less productive. We now strongly
+recommend them for all teams.
+
+In a couple of teams we let ill-defined user stories into an iteration, with
+disastrous results.  Small stories dragged on for weeks because we hadn’t
+defined their scope.  When we missed retrospectives, we found that small
+problems would start to spiral.  When we didn’t fully populate five iterations
+of a release with user stories at the Release Planning Meeting we found
+important pieces of functionality getting deprioritized and missing the release
+date. When we missed updating our metrics, the teams lost focus on the end goal.
+
+[P40]  People do not speak their minds
+
+at the outset most meetings seemed productive, since we had a lot of nodding 
+heads and everyone stepped out of meetings in agreement. However, I soon
+realized that any visible agreements on an approach or direction at meetings did
+not always mean that the participants  supported the decision. Therefore,
+drawing out an implementation  plan based on any key decisions emerging from
+meetings would be futile.
+
+Most 'real' discussions and communication  of how people really felt about a
+direction  or decision happened outside of meetings. The undercurrents  of
+resistance or genuine concerns that people had were never voiced at meetings.
+Direct conversation and sharing feedback was further constrained when the
+meetings included senior management representatives.
+
+Given  the  organization's   passive-aggressive culture there was no doubt in my
+mind that the transition to agile required a mind shift, but I also recognized
+that mind shifts couldn't happen overnight.
 % ---------------------------------------------------------------------

### 2014-04-28 journal/article.tex
Leadership challenges in change

--- 2014-04-27 journal/article.tex
@@
 Especially Scrum, XP, and Lean software development were used together. In
-addition, many cases mentinoed use of XP practices without explicitly stating XP
+addition, many cases mentioned use of XP practices without explicitly stating XP
 as the process being used (such as test driven development and continuous
@@
 
-\subsubsection{Challenges caused by management of the transformation}
+\subsubsection{Challenges relating to leading the change}
 
@@
 independence for teams to choose their way and diluting the value and structure
-of Scrum [P6]. Giving room for team autonomy was seen
+of Scrum [P6]. Giving room for team autonomy is one of the critical facets of
+agile, but it must be ensured that teams do not discard practices that add value
+and structure.
+
+Another issue that should be avoided is overenthusiasm towards agile methods.
+In several cases there was mentionings about change leaders or teams becoming
+agile zealots. For instance, Evans describes how some members of the agile
+community were becoming too evangelic, which made people take sides for or
+against agile development [12]. In another case team members started out with an
+overly eager attitude, but when agile did not deliver improvements quickly the
+interest faded and started to go back to the old process [P20]. Attempting to
+implement agile too strictly may cause conflicts, and especially in a large
+organization the change leaders need to maintain a collaborative attitude
+towards various groups in the surrounding organization [P6]. Introducing agile
+methods does not guarantee success, and therefore they should not be followed
+blindly [P4].
+
+In one case a passive-aggressive organization culture was described as a
+challenge. Ranganath describes how people did not speak their minds in meetings,
+thus creating a false sense of agreement when talking about how the organization
+should develop. Because of this resistance and concerns were left growing, and
+acted against achieving a successful transformation. To to avoid these issues
+change leaders worked with team members on an individual level to win confidence 
+and make their voices heard [P40].
+
+\subsubsection{Challenges in managing the changing organization}
 

### 2014-04-28 journal/article.tex
Lack of management support

@@
 
+Work in progress...
+
 Top down
@@
 
-\subsubsection{Challenges relating to leading the change}
-
-In the beginning of the transformation the organization has realized the need to
-change and is going towards the initial push.
+\subsubsection{Tensions between management and development}
+
+% In the beginning of the transformation the organization has realized the need
+% to change and is going towards the initial push.
 In many cases the initiative to change came from management, but when presented
@@
 against agile development [12]. In another case team members started out with an
-overly eager attitude, but when agile did not deliver improvements quickly the
+overly eager attitude, but when agile did not deliver improvements the
 interest faded and started to go back to the old process [P20]. Attempting to
@@
 
-\subsubsection{Challenges in managing the changing organization}
+The problem of tension between management and development may also be opposite.
+In cases where the change was emerging bottom up the reluctancy to change was
+not in the team members but management.
+This caused that significant organizational change could not be raised above the
+team level. For instance, Atlas describes how it would have required executive
+support to integrate the product management office and separate quality
+assurance groups in the agile process [P3]. In this case, when managers owning
+projects with many teams were not involved it retrained agile development to
+spread beyond team level. Further, if management does not speak out a goal
+towards using agile, developers feel that the agile methods may be replaced by
+something else at any time [P12].
+Lack middle management support was seen as the most serious shortcoming by
+Spayd [P49]. Middle management is in a position to undermine the entire
+transformation, and may do so if they do not participate in, and thus
+understand, the agile method [P49].
 

### 2014-04-28 journal/article.tex
Change resistance

@@
 
+The problem of tension between management and development may also be opposite.
+In cases where the change was emerging bottom up the reluctancy to change was
+not in the team members but management.
+This caused that significant organizational change could not be raised above the
+team level. For instance, Atlas describes how it would have required executive
+support to integrate the product management office and separate quality
+assurance groups in the agile process [P3]. In this case, when managers owning
+projects with many teams were not involved it retrained agile development to
+spread beyond team level. Further, if management does not speak out a goal
+towards using agile, developers feel that the agile methods may be replaced by
+something else at any time [P12].
+Lack middle management support was seen as the most serious shortcoming by Spayd
+[P49]. Middle management is in a position to undermine the entire
+transformation, and may do so if they do not participate in, and thus
+understand, the agile method. Agile brings changes to some management roles
+[P2], and lack of understanding of agile development will leave managers feeling
+left out [P6].
+
+\subsubsection{Other organizational issues}
+
 Another issue that should be avoided is overenthusiasm towards agile methods.
@@
 
-The problem of tension between management and development may also be opposite.
-In cases where the change was emerging bottom up the reluctancy to change was
-not in the team members but management.
-This caused that significant organizational change could not be raised above the
-team level. For instance, Atlas describes how it would have required executive
-support to integrate the product management office and separate quality
-assurance groups in the agile process [P3]. In this case, when managers owning
-projects with many teams were not involved it retrained agile development to
-spread beyond team level. Further, if management does not speak out a goal
-towards using agile, developers feel that the agile methods may be replaced by
-something else at any time [P12].
-Lack middle management support was seen as the most serious shortcoming by
-Spayd [P49]. Middle management is in a position to undermine the entire
-transformation, and may do so if they do not participate in, and thus
-understand, the agile method [P49].
-
 % Lack of plans (a guess -- not included)?
+
+\subsubsection{Change resistance}
+
+Change resistance is a central challenge in any organizational transformation.
+People will not be willing to change unless there are good reasons, and the
+change is preceived easy enough. It is difficult to attain a buy-in for a
+change, and even organizations that have a flexible culture will face resistance
+[P10]. It should be expected that not everyone will be willing to change, and
+some employees will never adapt to the change [P6]. In one case it was reported
+that objections from team members was a major reason for loss in time and
+productivity [P29].
+
+On a team level numoerous reasons for change resistance were given. For
+instance, Fecarotta described the organization of Boeing as risk-averse and
+cautious which slows down any change [P14]. The organization was also described
+having a deeply rooted status-quo because of high employee retention, which
+further hindered change. In some cases it was reported that people worried about
+new roles and responsibilities that agile would possibly bring [P44]. For
+instance, testers were worried that they would need to take on cross functional
+tasks, which would be outside their area of expertise [P18]. Yet another reason
+for change resistance was the move from own offices to team areas [P22]. People
+were also feeling being monitored because of the increased level of interaction
+within the team and between project stakeholders [P45].
+One study reported the cultural change required by middle management as the
+biggest cause of resistance [P2].
+
+Change resistance was mitigated for instance by education [P45], and one-to-one
+discussions [P22]. Change resistance was also softened by using first only teams
+willing to try agile, so that insights could be gathered to ease the continued
+transformation [P45].
+
+% General -->
+
+% For instance, teams may not be willing to try
+% new approaches unless they have been proven [P6].
+
+% Resistance is to be expected --> there are many different stakeholders [P41]
 

### 2014-04-28 journal/article.tex
Change resistance

@@
 did not comply with such a mindset. A top down mandate may also dilute the
-understanding of the reasons for starting the transformation, and further for
-agile development overall [P2].
-
-Organizations had bad experiences with prescribing how to carry out Scrum
-practices. For instance, doing agile based on checklists would not work [P41].
-Benefield described that it was challenging to find a balance between
-independence for teams to choose their way and diluting the value and structure
-of Scrum [P6]. Giving room for team autonomy is one of the critical facets of
-agile, but it must be ensured that teams do not discard practices that add value
-and structure.
+understanding of the reasons for starting the transformation, and the
+understanding of agile development overall [P2]. O'Connor describes how team
+members became sceptical towards agile when implementing mandaetd changes did
+not bring any visible benefits [P37].
+
+Organizations had also bad experiences if the implementation of the agile model
+was prescribed. For instance, doing Scrum based on checklists of practices would
+not work [P41]. Benefield described that it was challenging to find a balance
+between independence for teams to choose their way and diluting the value and
+structure of Scrum [P6]. Giving room for team autonomy is one of the critical
+facets of agile, but it must be ensured that teams do not discard practices that
+add value and structure.
 
@@
 
+% Agile zealots
+
 Another issue that should be avoided is overenthusiasm towards agile methods.
@@
 
+% People do not speak their minds
+
 In one case a passive-aggressive organization culture was described as a
@@
 were also feeling being monitored because of the increased level of interaction
-within the team and between project stakeholders [P45].
-One study reported the cultural change required by middle management as the
-biggest cause of resistance [P2].
-
-Change resistance was mitigated for instance by education [P45], and one-to-one
-discussions [P22]. Change resistance was also softened by using first only teams
-willing to try agile, so that insights could be gathered to ease the continued
-transformation [P45].
+within the team and between project stakeholders [P45]. One study reported the
+cultural change required for middle management as the biggest cause of
+resistance [P2].
+
+In these cases change resistance was mitigated for instance by education [P45],
+and one-to-one discussions [P22]. Change resistance was also softened by using
+first only teams willing to try agile, so that insights could be gathered to
+ease the continued transformation [P45].
+
+Another common form of change resistance that was reported was skepticism.
+Seffernick describes how management did on one hand acknowledge benefits of
+agile, but on the other explained away the possibilitiy to apply it due to
+contract reasons, the current matrix organization, and other organizational
+practices [P46]. In that case the director of development still pushed the agile
+transformation forward. Skepticism was often created by other similar
+misconceptions, including that agile does not work for complex products [P5,
+P36], agile needs to be implemented in a prescriptive by-the-book way [P9], that
+frequent meetings will cause overhead [P18], and that agile equals avoiding
+governance and working without a plan [P8].
 
@@
 
+% * No value seen by some groups, e.g. mainframe or maintenance [P49]
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-04-28 journal/article.tex
Old work and commitments

@@
 
+% Lack of plans (a guess -- not included)?
+
 % Agile zealots
@@
 
-% Lack of plans (a guess -- not included)?
+% Old work and commitments
+
+A condition hindering the transformation was that old commitments still needed
+to be tended to even though the organization was being restructured.
+Some organizations started their agile journey in a state where people were
+overcommitted, but it was later realized that overloaded people will not be able
+to learn new ways of working and changing their behavior [P37, P42]. Cowan
+describes how time was allocated for tending to old commitments [P11]. However,
+even when extra work was prepared for people with specific expertise
+became overloaded. After time it was realized that overcommitment must be
+reduced by allowing to shift excess work for later time [P11]. In one case
+people were forced to remain committed to firm deadlines which resulted in
+ignoring new practices introduced by agile, and eventually in the agile method
+breaking down.
 
@@
 
-% General -->
-
 % For instance, teams may not be willing to try

### 2014-04-29 journal/article.tex
Challenge: lack of physical spaces

@@
 
-% Old work and commitments
+% Old work and commitments, schedule pressure
 
@@
 breaking down.
+
+% Physical spaces, challenge of finding a seating
+
+In some cases the old way of working had people seated physically distributed,
+in opposite to what agile methods typically suggest. Office spaces needed to be
+arranged so that the entire team could work in a single room, but it took time
+and effort, and it was not always possible [P36, P6]. Lewis and Neher describe
+how dedicated rooms for teams could not be arranged, and team events such as
+daily standups required booking a conference room [P29]. In another case the
+estate organization shut down the teams' attempts to modify working spaces
+[P49].
 

### 2014-04-29 journal/article.tex
Lack of coaching/training

@@
 
+% ADD THIS QUOTATION:
+% [P21]
+% The development projects require continues maintenance and support. In addition,
+% the team was working on a new project with firm deadlines. The work environment
+% was very demanding and team worked under pressure to produce products according
+% to the planned schedule.
+% Scrum adoption process started while every member of the team was engaged in
+% his/her everyday tasks. With such work pressure the daily Scrum meetings added
+% extra pressure to the employees. They used to meet weekly and later twice a
+% month and then only when required and usually after working hours. As teams
+% started to skip daily meetings it also affected the learning process of scrum
+% between the team members. This eventually led to the failure of learn and
+% implement concept of the agile method.
+
 % Physical spaces, challenge of finding a seating
@@
 
+% Lack of resources
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+C: Lack of resoruces for training/coaching/piloting
+
+* There was not enough coaches to spread the transformation
+* New coaches from within the organization
+[P1]
+However the number of teams was large, and there were insufficient trained
+coaches to go around, so new coaches were being trained from within the
+organization, an important aspect of which was an ‘agile boot camp’.
+
+* Not enough coaches --> Funding constraint
+* ``the demand for training and coaching was simply outstripping supply''
+* Teams go ahead by them selves, and fail
+* The need for more coaches was proven by showing figures of increased productivity for coached teams
+[P6]
+We were spread so thin that we lost track of the team trainings in progress and
+were unable to keep up with demand. The whole program was in danger of
+imploding. We said “no” to teams who were requesting help thinking that might
+stem the tide. Many of those teams went ahead We frequently ran into one reason
+that many Scrum implementations fail;
+
+Unfortunately, our funding and headcount just weren’t sufficient to support the
+accelerating adoption.  We had funding from various parts of the organization,
+including the internal professional development organization, but the demand for
+training and coaching was simply outstripping supply. We needed to show that
+well-coached Scrum teams were worth the cost of coaching. For that, we would
+need metrics.
+
+* Damage due to lack of coaching --> The legacy of old model does damage if the team is not coached for agile
+* Lack of coaches impeded repetition of success
+* Hard to find good coaches
+[P9] <--> [P6]
+Without proper coaching however, the success of these informal teams met with
+mixed results over the longer term.  In retrospect, many of these teams did
+damage to themselves by improperly applying Agile techniques, particularly when
+command and control mechanisms were still in place.
+
+* Coaches were not hired --> Financial constraint
+* A proper piloting effort was not granted
+* Lack of training demotivated
+[P20]
+Entity S management recognized the need to hire a contractor as an agile master.
+However the position was not filled due to financial constraints.
+
+”why investing too much and wasting time and resources for a pilot project if
+the Scrum method promises success?”.
+Later after few weeks of direct adoption of Scrum, the team started to face
+tremendous difficulties and started gradually going back to their previous work
+method to catch up with late deadlines.
+
+The employees were unmotivated by lack of resources and lack of training.
+
+* There was no coach --> Financial constraints
+* Lack of training
+[P21]
+Missing Agile Master: Management recognized the need to have a coach. However
+the position was not filled due to financial constraints and senior software
+managers from within entity S took the role as the Scrum Master.
+
+The employees were unmotivated by lack of resources and lack ofrecognition of
+their efforts to improve the software development method.
+
+* Not enough coaches
+[P23]
+The Cookbook had done an excellent job of establishing a baseline understanding
+of Agile Delivery approaches but it had also stimulated a wave of demand from
+project teams asking for more information and guidance. BT’s “jam” – a team of
+just two coaches in Agile Delivery - was very thinly spread, especially given
+the scale of BT Exact’s workforce. Ironically, the success of these initiatives
+meant that BT again found itself with a shortage of “lumps” in its “jam”.
+
+* ``not enough people with agile experience to lead by example and/or mentor the non-agile developers''
+* Too few people with agile experience
+[P37]
+As I considered where to begin, I gained my first ally. Another coach was
+floating from team to team trying to provide support for the transformation. The
+two of us had many conversations about the problems the team was having and came
+up with two leading problems to tackle. One: we just did not have enough people
+with agile experience to lead by example and/or mentor the non-agile developers.
+
+* Shortened training led to motivation issues
+[P45]
+Many teams who had participated in the shortened training struggled later with
+communication and motivation issues. It turned out that in contrast to the pilot
+phase too few Scrum mentors were available to deal with the numerous team
+problems. To overcome these problems team training sessions were extended to one
+day per team and the “lean mentor” role was created. Lean mentors underwent
+in-depth training in Scrum, lean methodologies, and communication. Scrum master
+and product ownertraining sessions were extended to two days each.
+
+* Too few coaches
+* Coaches with too little experience may do harm to the agile model --> Now enforcing standards and certification
+[P47]
+The number of Agile projects and apprentice coaches is out pacing the number of
+mentor coaches. Many people being put in the apprentice coach position are more
+accustomed to traditional project practices. These factors contribute to
+apprentices potentially gaining the wrong type of experience to enable them to
+be an effective coach.
+
+* Too little coaches --> Coaches did not have time to follow up with the team
+* Too little training --> There are no trainers available
+[P49]
+As mentioned, there was no meaningful training in XP except for those projects
+that went to Offsites. Even though some fifty or more Offsites were held, this
+represented perhaps 30\% or less of all project teams. In order to empower
+employees, training is key; its lack prevented the majority of employees from
+truly understanding XP, let alone using it effectively. Training certainly could
+have been done “on-the-job” by experienced XP coaches, but there were virtually
+none at Qwest available to work with project teams. Offsites became the only
+form of such training and it was clearly inadequate to drive largescale change.
+
+Implementing XP without coaching will not work unless the team environment is
+already very favorable, such as having developers who have already used XP
+effectively and others who really want to try it.
 

### 2014-05-02 journal/article.tex
Lack of resources

--- 2014-04-29 journal/article.tex
@@
 
+\subsubsection{Lack of resources}
+
+Lack of resources allocated for supporting the transformation was an evident
+problem in many cases. Resources were most often short for training and
+coaching, but underbudgeting these activities may be a serious mistake.
+Hajjdiab [P20] describes how reluctance of management to invest in training
+caused teams to be ill prepared for the agile transformation effort, and
+eventually resulted in ending the use of agile methods. As a less dramatic
+outcome of shortened training was lower motivation [P45].
+
+Proper coaching was a particular problem for very large organizations. In cases
+where the number of teams in need of coaching was large the demand often
+exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
+was also attributed as a reason why the success of pilot teams could not be
+repeated when agile was adopted more widely [P9, P45].
+Silva and Doss [P47] described that when it was hard to fill the coaching
+positions people less seasoned in agile were appointed as coaches, which risked
+that agile practices would not be taught correctly.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-03 journal/article.tex
Falling back to old model

@@
 
+\subsubsection{Falling back to old model}
+
+In several cases challenges in the transformation resulted in people falling
+back to the old model of working.
+In some cases it was only a struggle to learn agile practices, but in other
+cases the old model displaced agile.
+Development work has to go on during the transformation but there will be new
+things to learn for the team. Stress caused by the combination of schedule
+pressure and much change at once will pull people back to the old way of working
+[P11, P18, P38].
+Schatz and Abdelshafi tell how even subtle trouble may risk the transformation,
+as people will always look for reasons to revert to familiar behavior [P44].
+Teams without adequate training were struggling with applying agile practices
+correctly, and the challenge the new practices posed made people to abandon them
+and return to the ways they know [P6, P49].
+
+In some cases agile was seen as an overhead, and was therefore abandoned. For
+instance, as new practices were being introduced there was a decrease
+performance, and when teams realized that the benefits are not immediate they
+started reverting to the old model [P21].
+
+When a release schedule with fixed deadlines that business required was combined
+to agile managers started to behave authoritatively and the self organization of
+agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)
+
+In one case new senior appointments were changed the management to be less
+favorable towards agile, and a waterfall development model started to return
+[P12].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-03 journal/article.tex
Restructuring: Change resistance, other organizational problems

@@
 
-\subsubsection{Tensions between management and development}
-
-% In the beginning of the transformation the organization has realized the need
-% to change and is going towards the initial push.
-In many cases the initiative to change came from management, but when presented
-wrong the initiative becomes a mandate. Spayd summarizes this as:
-``Organizations do not change merely because the boss says so, at least not in
-the way that is intended'' [P49]. In that organization the introduction of CMM
-was carried out consistently by a mandate, but the collaborative nature of agile
-did not comply with such a mindset. A top down mandate may also dilute the
-understanding of the reasons for starting the transformation, and the
-understanding of agile development overall [P2]. O'Connor describes how team
-members became sceptical towards agile when implementing mandaetd changes did
-not bring any visible benefits [P37].
-
-Organizations had also bad experiences if the implementation of the agile model
-was prescribed. For instance, doing Scrum based on checklists of practices would
-not work [P41]. Benefield described that it was challenging to find a balance
-between independence for teams to choose their way and diluting the value and
-structure of Scrum [P6]. Giving room for team autonomy is one of the critical
-facets of agile, but it must be ensured that teams do not discard practices that
-add value and structure.
-
-The problem of tension between management and development may also be opposite.
-In cases where the change was emerging bottom up the reluctancy to change was
-not in the team members but management.
-This caused that significant organizational change could not be raised above the
-team level. For instance, Atlas describes how it would have required executive
-support to integrate the product management office and separate quality
-assurance groups in the agile process [P3]. In this case, when managers owning
-projects with many teams were not involved it retrained agile development to
-spread beyond team level. Further, if management does not speak out a goal
-towards using agile, developers feel that the agile methods may be replaced by
-something else at any time [P12].
-Lack middle management support was seen as the most serious shortcoming by Spayd
-[P49]. Middle management is in a position to undermine the entire
-transformation, and may do so if they do not participate in, and thus
-understand, the agile method. Agile brings changes to some management roles
-[P2], and lack of understanding of agile development will leave managers feeling
-left out [P6].
-
-\subsubsection{Other organizational issues}
-
-% Lack of plans (a guess -- not included)?
-
-% Agile zealots
-
-Another issue that should be avoided is overenthusiasm towards agile methods.
-In several cases there was mentionings about change leaders or teams becoming
-agile zealots. For instance, Evans describes how some members of the agile
-community were becoming too evangelic, which made people take sides for or
-against agile development [12]. In another case team members started out with an
-overly eager attitude, but when agile did not deliver improvements the
-interest faded and started to go back to the old process [P20]. Attempting to
-implement agile too strictly may cause conflicts, and especially in a large
-organization the change leaders need to maintain a collaborative attitude
-towards various groups in the surrounding organization [P6]. Introducing agile
-methods does not guarantee success, and therefore they should not be followed
-blindly [P4].
-
-% People do not speak their minds
-
-In one case a passive-aggressive organization culture was described as a
-challenge. Ranganath describes how people did not speak their minds in meetings,
-thus creating a false sense of agreement when talking about how the organization
-should develop. Because of this resistance and concerns were left growing, and
-acted against achieving a successful transformation. To to avoid these issues
-change leaders worked with team members on an individual level to win confidence 
-and make their voices heard [P40].
-
-% Old work and commitments, schedule pressure
-
-A condition hindering the transformation was that old commitments still needed
-to be tended to even though the organization was being restructured.
-Some organizations started their agile journey in a state where people were
-overcommitted, but it was later realized that overloaded people will not be able
-to learn new ways of working and changing their behavior [P37, P42]. Cowan
-describes how time was allocated for tending to old commitments [P11]. However,
-even when extra work was prepared for people with specific expertise
-became overloaded. After time it was realized that overcommitment must be
-reduced by allowing to shift excess work for later time [P11]. In one case
-people were forced to remain committed to firm deadlines which resulted in
-ignoring new practices introduced by agile, and eventually in the agile method
-breaking down.
-
-% ADD THIS QUOTATION:
-% [P21]
-% The development projects require continues maintenance and support. In addition,
-% the team was working on a new project with firm deadlines. The work environment
-% was very demanding and team worked under pressure to produce products according
-% to the planned schedule.
-% Scrum adoption process started while every member of the team was engaged in
-% his/her everyday tasks. With such work pressure the daily Scrum meetings added
-% extra pressure to the employees. They used to meet weekly and later twice a
-% month and then only when required and usually after working hours. As teams
-% started to skip daily meetings it also affected the learning process of scrum
-% between the team members. This eventually led to the failure of learn and
-% implement concept of the agile method.
-
-% Physical spaces, challenge of finding a seating
-
-In some cases the old way of working had people seated physically distributed,
-in opposite to what agile methods typically suggest. Office spaces needed to be
-arranged so that the entire team could work in a single room, but it took time
-and effort, and it was not always possible [P36, P6]. Lewis and Neher describe
-how dedicated rooms for teams could not be arranged, and team events such as
-daily standups required booking a conference room [P29]. In another case the
-estate organization shut down the teams' attempts to modify working spaces
-[P49].
-
 \subsubsection{Change resistance}
 
-Change resistance is a central challenge in any organizational transformation.
+To start a transformation the organization needs an initial push. However,
+people are usually disincluned to change, and it is likely that resistance
+will emerge in any change. 
+% Change resistance is a central challenge in any organizational transformation.
 People will not be willing to change unless there are good reasons, and the
@@
 
-\subsubsection{Lack of resources}
-
-Lack of resources allocated for supporting the transformation was an evident
-problem in many cases. Resources were most often short for training and
-coaching, but underbudgeting these activities may be a serious mistake.
+The way the transformation is initiated will also affect how change resistance
+will show.
+In many cases the initiative to change came from management, but when presented
+wrong the initiative becomes a mandate that pepole are not receptive for. Spayd
+[P49] summarizes this as: ``Organizations do not change merely because the boss
+says so, at least not in the way that is intended''. In that organization the
+introduction of CMM had been carried out consistently by a mandate, but the
+collaborative nature of agile did not mix well with such a mindset. A top down
+mandate may also dilute the understanding of the reasons for starting the
+transformation, and the understanding of agile development overall [P2].
+O'Connor describes how team members became sceptical towards agile when
+implementing mandaetd changes did not bring any visible benefits [P37].
+
+Organizations had also bad experiences if the implementation of the agile model
+was prescribed. For instance, doing Scrum based on checklists of practices would
+not work [P41]. Benefield described that it was challenging to find a balance
+between independence for teams to choose their way and diluting the value and
+structure of Scrum [P6]. Giving room for team autonomy is one of the critical
+facets of agile, but it must be ensured that teams do not discard practices that
+add value and structure.
+
+The problem of tension between management and development may also be opposite.
+In cases where the change was emerging bottom up the reluctancy to change was
+not in the team members but management. This caused that significant
+organizational change could not be raised above the team level.
+For instance, Atlas [P3] describes how it would have required executive support
+to integrate the product management office and separate quality assurance groups
+in the agile process. In this case, when managers owning projects with many
+teams were not involved in the transformation, the agile model was restrained
+from spreading beyond team level. Lack of middle management support was also
+seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
+position to undermine the entire transformation, and may do so if they do not
+participate in, and thus understand, the agile method. Agile brings changes to
+some management roles [P2], and lack of understanding of agile development will
+leave managers feeling left out [P6]. Still one problem is that if management
+does not speak out a goal towards using agile, developers feel that the agile
+methods may be replaced by something else at any time [P12].
+
+\subsubsection{Other organizational issues}
+
+% Lack of plans (a guess -- not included)?
+
+% Agile zealots
+
+Another issue that should be avoided is overenthusiasm towards agile methods.
+In several cases there was mentionings about change leaders or teams becoming
+agile zealots. For instance, Evans describes how some members of the agile
+community were becoming too evangelic, which made people take sides for or
+against agile development [12]. In another case team members started out with an
+overly eager attitude, but when agile did not deliver improvements the
+interest faded and started to go back to the old process [P20]. Attempting to
+implement agile too strictly may cause conflicts, and especially in a large
+organization the change leaders need to maintain a collaborative attitude
+towards various groups in the surrounding organization [P6]. Introducing agile
+methods does not guarantee success, and therefore they should not be followed
+blindly [P4].
+
+% People do not speak their minds
+
+In one case a passive-aggressive organization culture was described as a
+challenge. Ranganath describes how people did not speak their minds in meetings,
+thus creating a false sense of agreement when talking about how the organization
+should develop. Because of this resistance and concerns were left growing, and
+acted against achieving a successful transformation. To to avoid these issues
+change leaders worked with team members on an individual level to win confidence 
+and make their voices heard [P40].
+
+% Old work and commitments, schedule pressure
+
+A condition hindering the transformation was that old commitments still needed
+to be tended to even though the organization was being restructured.
+Some organizations started their agile journey in a state where people were
+overcommitted, but it was later realized that overloaded people will not be able
+to learn new ways of working and changing their behavior [P37, P42]. Cowan [P11]
+describes how time was allocated for tending to old commitments. However, even
+when extra work was prepared for people with specific expertise became
+overloaded. After time it was realized that overcommitment must be reduced by
+allowing to shift excess work for later time [P11]. In another case management
+forced people to remain committed to firm deadlines, even midst in the
+organizational transformation. The engagement in old commitments and tasks
+resulted in ignoring new practices introduced by agile, and eventually the agile
+method was breaking down [P21].
+
+% Physical spaces, challenge of finding a seating
+
+In some cases the old way of working had people seated physically distributed,
+in opposite to what agile methods typically suggest. Office spaces needed to be
+arranged so that the entire team could work in a single room, but it took time
+and effort, and it was not always possible [P36, P6]. Lewis and Neher describe
+how dedicated rooms for teams could not be arranged, and team events such as
+daily standups required booking a conference room [P29]. In another case the
+estate organization shut down the teams' attempts to modify working spaces
+[P49].
+
+% Lack of resources
+
+Lack of resources allocated for supporting a transformation was an evident
+problem. Resources were most often short for training and coaching, but
+underbudgeting these activities may be a serious mistake.
 Hajjdiab [P20] describes how reluctance of management to invest in training

### 2014-05-03 journal/article.tex
Managers' roles

@@
 
+\subsubsection{Agile is found to be hard}
+
+A challenge many organizations came across was that an agile model was not
+straight forward to implement.
+
+A common problem in the agile implementations was the role of managers.
+Especially the role of middle management was seen as unclear in agile
+methods, but a cultural change would still be required on that level [P2].
+It was described that people in management positions needed to resist the
+tendency to take command and control to allow room for self-organization, but
+the change in role was difficult to achieve for the people involved [P11, P52].
+Nielsen and McMunn [P30] describe how the project management group had
+previously worked through big up front plans and competing of resources, but
+those ways would need to dramatically change.
+
+Several other problems related to management roles were also presented. For
+instance, Spayd [P49] describes how managers were left outside the roles offered
+by the new agile model. In another case when Scrum masters were appointed from
+the ranks of managers developers felt being micromanaged because of vague
+understanding of the agile method [P27]. It was also described how mixing the
+role of project manager and Scrum master created a conflict in interest, and
+turned the Scrum manager's role into one of policing the team instead of
+supporting the team [P30]. In some cases problems with manager positions were
+solved by phasing out roles relating to the old model, and replacing them with
+new positions more suitable for agile [P38, P52].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-04 journal/article.tex
Improvements

@@
 describes how time was allocated for tending to old commitments. However, even
-when extra work was prepared for people with specific expertise became
+when there was preparation for extra work people with specific expertise became
 overloaded. After time it was realized that overcommitment must be reduced by
@@
 resulted in ignoring new practices introduced by agile, and eventually the agile
-method was breaking down [P21].
+method was breaking down [P21]. Spayd [P49] also describes how senior management 
+pressed teams to deliver what the customer had been promised, regardless of what
+the current velocity of development predicted.
 
@@
 
-% Lack of resources
+% Lack of training and coaching
 
@@
 that agile practices would not be taught correctly.
+
+% Add quote?
+% * No coaching --> Applying techniques improperly --> Damage
+% [P9]
+% Without proper coaching however, the success of these informal teams met with
+% mixed results over the longer term.  In retrospect, many of these teams did
+% damage to themselves by improperly applying Agile techniques, particularly when
+% command and control mechanisms were still in place.
 

### 2014-05-04 journal/article.tex
Misunderstandings in the agile model

@@
 
-A challenge many organizations came across was that an agile model was not
-straight forward to implement.
+A challenge most organizations came across was that an agile model was difficult
+implement. An experienced software team may do well in training, but when the
+time comes to apply agile techniques in practice the team may get lost [P21].
+Agile was also seen as a solution to all problems [P9], success was declared
+prematurely [P49], and expectations created by successful early experiments were
+not fullfilled [P38]. Disapointment was granted when these hopes were not
+fullfilled and implementing an agile model was found to be hard and time
+consuming. The difficulty in implementing agile is evident in misconceptions of
+the agile model, when by-the-book implementations prove to be insufficient, and
+when the agile methods were applied badly to the organization.
+
+There were many examples on problems caused by misconceptions of what agile
+software development is. On a general level, Bang [P4] describes how the values
+of the agile manifesto were not understood, and agile practices were carried out
+without understanding their purpose. In one case management saw the purpose of
+agile simply being faster product delivery [P9].
+
+On a team level there were several examples of misunderstanding the agile model.
+Cases describe how agile was seen as freedom to hack without documentation
+[P26], that development could be done without design tasks [P37], and that agile
+meant that everyone should become a generalist [P42]. Schatz and Abdelshafi
+describe how teams presented unfinished work at reviews and ignored the
+principle of showing only completed increments, which resulted in a backlog of
+bugs [P44]. In one case team members had perceived the introduction of agile as
+a means to squeeze out more efficiency [P45]. In addition, the flatter
+organization was seen as fewer career opportunities, and pushed team members to
+internal competition for visibility [P45].
+Still one misunderstanding of agile was due to perceiving it only through the
+tools used, such as project management software [P48]. Superficial focusing only
+on tools themselves and not the reasons behind their use led to frustration
+[P42].
+
+% Misunderstandings in the agile model
+
+% By the book is not good
+
+% Customizing badly (mapping to old model)
+
+% Differing vocabularies
+
+% Roles of managers
 
@@
 those ways would need to dramatically change.
+
+% Roles of managers (continued)
 

### 2014-05-04 journal/article.tex
Possible addition
Misunderstanding

@@
 
+% Include quotation?
+% * Some teams mirrored waterfall approaches in their behavior --> Did not work
+% [P32]
+% Then comes execution and the learning really begins. Teams’ working habits
+% determine their success.  Often teams quickly ignore plans in favor of using
+% tactical approaches that mirror waterfall thinking. This contrasts nicely with
+% those more attuned to applying agile.
+
 \subsubsection{Agile is found to be hard}
@@
 
-% Misunderstandings in the agile model
+In some cases the misunderstanding of the agile model was described as ``doing
+mini waterfalls'' [P6]. Managers were using Scrum terminology, but at the same
+time committing unreasonable workloads on teams [P6]. In another case a
+waterfall nature was seen in giving task estimates as hours left, and task
+breakdowns disconnected from what was really being done [P27].
 

### 2014-05-04 journal/article.tex
By the book is not good

@@
 
+% Misconceptions of agile
+
 There were many examples on problems caused by misconceptions of what agile
@@
 
+Even if there was no misunderstandings the correct agile model was found hard to
+implement. Several cases describe how an agile model was hard to learn from
+literature [P21, P27], especially when it comes to implementing it on a large
+scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
+where we could find easy answers on how to do things''.
+Schnitter and Mackert [P45] report that theoretical considerations on how to
+scale up the agile methos were not good enough, and that product managers and
+architects struggled when several Scrum teams were working concurrently.
+Another case concluded that all practices prescribd in XP did not fit enterprise
+scale development, and some practices required customization [P49]. It would not
+be good to prescribe practices with a check list as it would impose a single
+model in an organization with several individual teams and projects [P41].
+
+Further indications were given that the advice given in literature will not
+ferry all the way. Benefield [P6] describes how it was dificult finding a
+balance between prescribing a by-the-book implementation which may put people
+off, and giving too much freedom in the agile methods which may weaken core
+practices.
+The reality where the practices needed to be applied was described as
+messy in comparison to the ideal circumstances presented in literature, which
+made it difficult to choose an initial model and gain buy-in for it [P10].
+Difficulties in choosing the model to start with were also due to variances
+perceived in the agile approaches suggested in literature [P30].
+
 % Customizing badly (mapping to old model)
@@
 % Differing vocabularies
+
+\subsubsection{Large organization issues}
 

### 2014-05-05 journal/article.tex
Customizing badly

@@
 
+The difficulty of agile and misunderstandings were also evident in cases where
+the methods were customized badly. As by-the-book implementation often was not
+feasible the agile method was tailored to suite the needs of the organization.
+However, in some cases tailoring meant skipping some practices, which lead into
+problems. In one case certain individuals were allowed to ignore some core
+elements of Scrum, which came to turn the teams decision making into a variant
+of command and control [P9].
+Lewis and Neher [P29] describe how there was temptation to strip some practices
+and enhance others. Previous attempts had proven that one of the reasons early
+attempts of agile had failed was deviations from the process, because of which
+the agile mindset did not take root [P29]. In a bad implementation customization
+may lead to the team taking only practices that reflect their current needs, and
+therefore failing to achieve a real change in process and mindset [P9, P46].
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
 % Differing vocabularies

### 2014-05-05 journal/article.tex
Work in progress

@@
 
+WORKING IN SILOS --> Certain expertise is required in many places
+
+* Teams were reorganized without notice --> Planning was hampered when people did not konw who would stay on a team
+* Frequent changes --> People managers had to agree to minimize team changes
+[P11]
+When we launched multiple Scrum teams, we tried to keep existing teams together
+and form new teams based on skills.  As we gained experience, we realized that
+there were skills shortages in certain types of expertise that wasn’t possible
+to develop in a sprint or two. There was pressure to move engineers with certain
+skills to the teams that needed them for work on upcoming sprints or to address
+critical customer requirements.  When moves occurred without much notice, teams
+couldn’t plan effectively if they didn’t know who would stay on a team.  After
+several months of frequent changes, product owners and people managers agreed to
+minimize team changes, and even talked of instituting a “trading deadline.”
+
 % ---------------------------------------------------------------------
@@
 
+* Some success factors and challenges relate
+
+* Success factors and challenges may be complements, and may be presented as
+  complements
+
+* What is special in large scale?
+
+* What is similar in organizations of all scales?
+
 % ---------------------------------------------------------------------

### 2014-05-06 journal/article.tex
Differing vocabularies

@@
 
+\subsubsection{Large organization issues}
+
+% Challenge: Differing vocabularies
+
+In a large environment many agile teams coexist and it is possible that their
+agile models and vocabularies start diverging. When many teams implement agile
+without consistent guidance friction and fragmentation may be generated [P9].
+The organization may require moving people between teams from time to time, and
+therefore it would be desirable that no big differences in teams' agile cultures
+exist. Divergence in process would create risk increasing costs when relocating
+people [P38]. Further, if processes of teams diverged management had
+difficulties in forecasting and comparing work of different teams [P38, P48].
+To overcome problems with divergence in agile models some organizations defined
+standards to follow [P8, P43].
+
+% Coordinating between many teams
+
+* Hard to create roadmap between teams --> Teams see only their own priorities
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ---------------------------------------------------------------------
-
-% Differing vocabularies
-
-\subsubsection{Large organization issues}
 

### 2014-05-06 journal/article.tex
Work in progress

@@
 
+"C: Coordinating/integrating work of many teams"
+
+* Nightly builds with work products from all teams --> Integration problems
+[P5]
+The nightly build process served as the point in time when the work from five or
+more Scrum teams all came together into a single installable image.  This
+integration effort occurred with the production of every installable image. 
+Integration across so many teams every night was the primary cause of the build
+problems we faced.
+
+* Coordination challenge --> Team goal or product goal?
+* Team centric focus created strong boundaries --> Bad for coordination
+[P7]
+A challenge we faced in team coordination was understanding whether to optimize
+the team goal or the product goal. Some teams focused too much on their own
+goals and created strong boundaries, impeding coordination.
+
+* How to manage dependencies between teams? --> No further mentionings
+[P9]
+Unresolved at the end of the pilot, however, was the issue of how to manage
+external dependencies with other teams.
+
+* ?
+[P10]
+CROSS-FUNCTIONAL DEPENDENCIES: However well we did within the team, we also
+frequently ran into dependencies on other groups often in other locations –
+groups such as UED (User Experience and Design), operations, legal and the
+Yahoo! security team (affectionately called the Paranoids) – which were
+identified as blocking issues but almost never neatly resolved within our sprint
+boundaries.
+
+* Well into the transformation there was trouble with dependencies between teams
+* Management burden to manage dependencies --> Creates problem with roadmapping
+* Hard to create roadmap between teams --> Teams see only their own priorities
+* Reduce dependencies --> Help roadmapping
+[P10]
+Now  teams were  comprised of tight-knit groups that – while collocated wherever
+possible – were  responsible to deliver  value regardless of location,
+technology or historical allegiance.
+However, within six months or so, and despite a pattern  of  success and
+improvement, it became very clear that the new forking between horizontal
+and vertical  teams  placed an enormous burden on the product and program
+management teams to manage dependencies in a way that supported  the  broader
+organizational roadmap.  Teams were structured to execute independently,  but
+the broader product roadmap called for features that  implied  specific platform
+improvements. Establishing a  reliable roadmap therefore required massive (and 
+often unreliable) coordination between teams that were more often  than not
+head-down in their own (sub-sprint) priorities.
+Theoretically,  the product owners (assisted by the Music PMO) would ensure that
+capabilities required by the  roadmap would appear on their own team’s backlog
+at an appropriate level of priority.  In practice, it was very difficult to
+identify  and track such dependencies, especially when the  feature  set  (and
+associated set of dependencies) was constantly in flux.
+Phase 2 of the transition (continuing today –  see Figure 4) therefore  attempts
+to further reduce such dependencies and improve how they are managed over time
+to improve release and roadmap planning.
+
+* Team centricity --> Power --> However, they operate with independent priorities --> No respect for larger context (release)
+* Release Framework was created
+[P13]
+Shortly after the initial rollout to 30+ product teams it was clear that we had
+unleashed the power of our individual teams through their new found
+self-organization and focus on short sprints.  However, since we aggregate the
+work of many sprints into a single release it was also becoming clear that our
+scrum teams were operating independently without respect for the context of the
+larger release.  Herding 30+ teams working in the same codeline to a predictable
+time-boxed release date is very challenging in any traditional development model
+let alone an agile development model where we value self-organization over
+centralized control.
+Our answer to this problem was a lightweight Release Framework and the
+introduction of a Release Manager for the overall product release.
+
+* Ease coordination by maintaining code quality as top priority, above new features
+[P13]
+We needed to ensure that the process or framework we put into place would
+promote visibility and optimization of the whole release, or of the suite of
+product functionality that we deliver to our customers and our partners.
+With such a large organization working on the same codeline, it was imperative
+that our global priorities were consistent and clear to all scrum teams in order
+to maintain quality of the release throughout the development cycle.  The
+priorities were as follows:
+1. Maintain existing functionality and quality of the overall codeline on a
+daily basis.  All automation regression tests must pass at a 99\% passing
+threshold on every checkin.
+2. Building new functionality is secondary to codeline stability.
+While each scrum team had to invest time to maintain the quality of the codeline
+on a daily basis, they also experienced benefits from it as well.
+
+* Individual component teams gained success --> Coordination challenges across teams
+* Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
+* Help: Tool support to visualize dependencies for people
+[P17]
+As individual component teams began to experience success, the challenge changed
+to coordinating across teams. Although iterations were humming along at the team
+level it was initially difficult for managers to get a complete picture of how
+the release was progressing. They were unable to tell when particular features
+were complete as their requirements crossed team boundaries.
+Using Rally’s application for managing the development lifecycle, BMC is now
+able to guide the overall release objectives with themes and high-level
+features.
+
+* Initial adopting good
+* Problems with inter-team coordination --> Architecture fragile, code dialects, distrust between teams
+* (Multisite, teams on different sites)
+* Help: Retrospectives across teams --> Created global teams --> Shared backlog
+[P24]
+To embrace and implement Scrum across the different teams and organizations, we
+focused on doing regular retrospectives on both team and company level.
+The first couple of company retrospectives were quite hectic and required some
+extra facilitation techniques to be constructive. But we did identify some of
+the main problems when crossing the team boundaries and working together with
+other departments.
+No matter how much communication we tried to establish between the teams, we
+were still far away from shared team focus on the complete product. As a result
+we could see the architecture becoming more fragile, multiple code dialects and
+distrust between teams were also starting to grow.
+By doing regular retrospectives with all the different teams and across the
+company we identified the main problems and decided it was time to create the
+global Scrum teams. We wanted to improve communication and trust by working on
+shared commitments between the different teams.
+
+We still needed to improve the focus on the complete product and not on "bad
+code” from other teams. The main barrier for building quality into the product
+was the distrust between the teams who worked on the same product. Therefore the
+two R\&D teams in Denmark and Bangladesh were merged into one global team.
+Instead of two distributed teams, working with different sprint backlogs, we
+established one team with one sprint backlog. The problem of building quality
+into the complete product was now a shared team problem, not a management
+problem.
+
+* Scrum on team level --> Software interfaces not properly synchronized
+* Problems with larger features
+* Scrum on team level good --> True benefit requires entire program to be flexible for changes
+[P26]
+Using scrum only on the team level had led to additional problems. Software
+interfaces were not properly synchronized with other development teams.
+Larger features were causing synchronization problems due to dependencies in
+code. Synchronization to the main waterfall program was considered problematic.
+Deploying scrum on the team level had brought flexibility in the team, but the
+ability to embrace the change was not truly benefiting the whole program as it
+was lacking capabilities to systematically embed the changes in all teams and
+levels.
+
+* Problems with cross team integration, lack of unified build environment
+[P27]
+1. Our department had a lack of program-wide build scripts, build environment,
+   data alignment and testing strategy.
+2. Cross team integration build took too long.
+
+* Different customers compete --> Customer centricity does not view the system as a whole
+[P29]
+each customer base views the system only via their process flow.
+We have multiple competing customers crossing multiple co-dependent
+applications, including dependencies on applications outside of our
+organization; multiple funding sources; and a mixture of Waterfall and Scrum
+projects.
+For our software teams to be able to execute efficiently, we will need to push
+the complexity and prioritization upstream, scaling teams to meet demand while
+maintaining clean, simple prioritized Product Backlogs for each product.
+
+* First cross team dependencies were reduced --> Problems with too much dependencies
+* Traditional dependency management limited flexibility, which was demanded --> Inflexibility, work intensive, adversarial relationships
+[P34]
+From the onset, we sought to reduce cross-team dependencies as much as possible,
+allowing each of the Scrum teams to “run” at their fastest pace.  However, as
+the number of concurrent work streams increased, so did the number of
+dependencies.  In particular, we found that adopting a SOA ran counter to our
+efforts.
+We quickly faced the prospect of implementing and integrating with a sizeable
+set of centralized services that would be consumed by multiple project teams.
+Within a few months, we had accumulated a significant number of integration
+points, preventing us from delivering fully production-ready software until they
+were resolved.
+Accepting the fact that these dependencies would be part of the project, we
+sought to create a lightweight and efficient process for resolving dependencies
+and integration points.  We recognized that traditional approaches to dependency
+management tend to place significant constraints on the sequence of work,
+reducing the teams’ ability to independently prioritize their backlogs. 
+Furthermore, highly contractual, document-centric approaches to defining
+integration points are work-intensive, prone to error and can create adversarial
+relationships between teams.  We wanted to establish an approach based on direct
+communication and collaboration.
+Our initial integration efforts involving five of the module teams of the Core
+Application were reasonably successful. After some initial scrambling, the teams
+self-organized and established a set of procedures for managing inter-module
+integration.
+
+* Worked decently with 5 teams and SoS --> Problems with global deployment
+[P34]
+However, our efforts to facilitate cross-team collaboration within the Core
+Application benefited heavily from the fact that the five module teams were all
+closely co-located, within the same organizational structure and employed an
+active “Scrum of Scrums” model to ensure communication and coordination between
+the teams.  As we pushed this approach out further across the program and the
+globe, we began to experience significant challenges.
+
+* Problem: Agile teams were thinking inside the team room walls
+* Success: Thinking of the teams that will integrate to your work
+* Failure: Thinking only inside the team walls
+[P34]
+We found that we had varying degrees of success in managing heavily-used
+services.  In those cases where we were successful, it was invariably due to a
+small group of people from both the consuming and providing teams taking
+ownership of the services and proactively seeking out future integration points
+and potential challenges.  Conversely, in those situations where such ownership
+did not exist and people remained within their team walls, we invariably
+experienced integration challenges: services consistently failing to meet the
+consumer’s requirements and integration points that were poorly tested and
+failed in system testing.
+
+* Global effort problems --> Communication regressed to contractual --> Narrow communication channels were used: documents, e-mail
+* Success: Individuals that saw effort in wide-channel communication, Individuals that thought outside team room walls
+[P34]
+As more and more integration occurred across the globe, we witnessed a tendency
+to revert back to contractual forms of communication. As always, distributing
+the development team significantly increases the challenge to directly
+communicate and collaborate.  The natural tendency to rely on e-mail due to
+differences in time zones frequently led to communication breakdowns and
+unnecessary blockages – forming barriers to feedback. In some instances,
+individuals broke through these barriers using instant messaging products, web
+cams and adjusted working hours to facilitate as much direct communication as
+possible.  At the same time other teams regressed, relying on documentation and
+establishing an almost adversarial relationship with their integration partners.
+
+* Large program problems --> Green builds hard to maintain
+* Tools insufficient for large
+[P34]
+Given the size and distribution of our development team and the fact that the
+program produced a single, integrated enterprise application, the complexity of
+the build process and the level of effort to maintain it was significant.   As
+the program grew in size, we recognized that the tool set with which we began
+the project was no longer sufficient.  We began experiencing nearly continuous
+broken builds due to conflation and timing issues on the build server.  At this
+point, we began a lengthy initiative to replace the toolsets and revamp the
+entire build process using AntHillPro and Subversion.  As a result, we created a
+build process that effectively managed our myriad of dependencies and supported
+a total build of the application in a reasonable time.
+
+* Problems: Integration and dependency management
+* Goal: Foster communication and collaboration instead of rigid processes
+* Problems: Team avoiding external distractions, balance between team membership and project membership --> CI and green builds did not work
+* Problems: Technical leads must work outside of team --> To coordinate
+* Solution: Balance between own goals and project goals --> Think outside team room
+* The team must think of others than them selves, too
+[P34]
+Over the lifetime of the project we encountered two key challenges due to the
+scope and scale of this effort: integration and dependency management and
+continuous integration. When addressing these challenges, we sought to adhere to
+the basic principals of the Agile manifesto as much as possible: making the
+delivery of working software our overriding goal and fostering communication and
+collaboration instead of dictating rigid processes. We sought to empower
+individuals and teams to make decisions and formulate their own processes, and
+to take calculated risks without fear of reprisals.
+
+However, we found that some of the qualities and practices that made the teams
+work effectively in isolation actually reduced their effectiveness in the
+large-scale environment.
+For example: in smaller projects, we work aggressively to avoid external
+distractions and encourage team members to focus exclusively on their individual
+team goals as expressed in the backlog.  In large-scale projects, team members
+must participate in cross-team activities and allocate time to project-wide
+activities.  In particular, team technical leads were required to dedicate
+significant portions of their time outside of the team room.  We found that many
+of our team leads struggled with this task.  While they recognized the need for
+cross-team communication, they resented context switching and the impact on
+their time dedicated to the team’s sprint goals.
+Furthermore, all of the teams struggled with balancing their identity as a team
+member against their membership in the larger project. This was reflected most
+keenly in our failure to sustain continuous integration and produce frequent
+green builds.
+
+Agile team must make trade-offs between stabilizing an application and
+completing more story cards, the size of this project required teams to put
+significantly more time into tracking down issues that were caused by other
+teams or dependencies.   Many team members found this frustrating and
+demoralizing.
+
+[P43] Ryan 2008 - The Price of Agile Is Eternal Vigilance.pdf - 43:11 [We implemented a scrum-of-scru..]  (4:48-4:200)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
+No memos
+
+We 
+implemented a scrum-of-scrums run by our release 
+management team to identify inter-team dependencies 
+and resolve merge issues before they occur.
+
+[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:18 [What could not be mitigated we..]  (4:2133-4:2292)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
+No memos
+
+What could not be mitigated were issues with teams dependent on “unreliable”
+Scrum teams, at least if the Scrum teams were isolated in a non-Scrum environment.
+
+[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:28 [Several approaches are known t..]  (8:522-8:1054)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
+No memos
+
+Several approaches are known to co-ordinate the work of several Scrum teams working
+on one product, e.. Scrum of Scrums [9] or MetaScrum [18]. The size and complexity
+of SAP’s products makes it necessary to consider these techniques. To co-ordinate the
+work of multiple Scrum teams and to communicate the requirements via team product
+backlogs, various approaches were combined. Special product teams were formed for
+this purpose. These teams can be regarded as both a permanent Scrum of Scrums and a
+supervisory product owner.
+
+[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:32 [Product teams were introduced ..]  (8:1076-8:1627)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
+No memos
+
+Product teams were introduced as a second organizational layer above the Scrum teams.
+A product team is responsible for the work of up to 7 development teams. It consists
+of the product owners of those teams and the same number of team members who
+are specialists in particular ﬁelds. Depending on the problem area other experts may
+be included. This allows for full engineering coverage of all problem areas expected,
+well-organized communication among the teams, and direct communication with the
+development teams to detect and mitigate risks.
+
+[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:33 [Product teams can rarely suppo..]  (9:841-9:1526)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
+No memos
+
+Product teams can rarely support more than seven Scrum teams. Ifthe products or components are be too big or too complex to be developed in this organizational setup, an
+intermediate organizational layer is inserted between the product team and the Scrum
+teams. These teams are called “area product teams”. Their responsibility is the same as
+the product teams’ responsibility with the exception of budget and ﬁnal product decisions.
+In this three-layer organization a productteam works similar to a project management ofﬁce.
+The complete budget and product responsibility is with the product team,
+while the operational day-to-day decisions are handled by the area product teams.
+
+[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:38 [Scaling Scrum to a multi-team ..]  (10:1223-10:1784)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [O: Large scale definition - Family: Contextual] 
+No memos
+
+Scaling Scrum to a multi-team development organization is not as easy. Scrum of
+Scrums and backlog preparation for many teams can be combined by product teams
+as outlined above. Looking at all known parameters of agile project management at
+SAP we tend to say that not much more than 130 development employees may be
+organizedin that fashion. This number sums up developers in 7 teams (max. 70 people),
+the product team (max. 16), developmentinfrastructure responsibles (about 10), quality
+assurance and testers (about 25), general management (about 10).
+
+[P48] Smits 2011 - Agile Experience Report - Transition And Complexity At Cisco Voice Technology Group.pdf - 48:30 [Tracking progress of hundreds ..]  (4:1657-4:2111)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
+No memos
+
+Tracking progress of hundreds of teams working on 
+pieces of major features proved to be a major 
+problem. Tools that we were using did not have roll 
+up capability without imposing strict rules on how 
+data was entered, labeled and updated. We found out 
+our tools were not really scalable for the size and 
+complexity of the programs we were working on. 
+This proved a serious handicap in providing 
+executive summary information about progress.
+
+[P50] Sutherland 2011 - Hitting The Wall - What To Do When High Performing Scrum Teams Overwhelm Operations And Infrastructure.pdf - 50:20 [A key impediment was the distr..]  (4:1737-4:2643)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
+No memos
+
+A key impediment was the distributed 
+development model in use by the Scrum teams.  
+Originally, before Scrum adoption, all teams worked
+together in a common environment.  There were 
+many integration issues as a result and no solid 
+process to rectify them.  So with the initial adoption 
+of Scrum in summer 2009 it was decided that each 
+individual Scrum team should have a dedicated and 
+private server with integration to take place after the 
+build.  In retrospect in the fall of 2009, the downside 
+of the decision to break up the environment was 
+being felt as integration issues were being discovered 
+very late,  as  the  build  process  was  the  first  point  of 
+integration.  To overcome this impediment the 
+decision was made to re-integrate all teams onto a 
+single  server  environment  again,  so  that  integration 
+issues could be rectified before they left 
+development.
+
+[P51] Vlaanderen 2012 - Growing Into Agility - Process Implementation Paths For Scrum.pdf - 51:12 [Initially, the sprints lasted ..]  (5:230-5:717)   (Super)
+Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
+No memos
+
+Initially, the sprints lasted two or three weeks, depending on
+the team, causing a discrepancy between the sprint durations of some of the
+teams. At the same time, interdependencies started to become more of an issue
+as the amount of teams grew. This caused problems when, for instance, frontend teams needed functionality in the back-end that was not yet developed.
+These problems, which consisted mainly of delays, were solved by ﬁxating sprint
+duration at two weeks for all teams.
+
 % ---------------------------------------------------------------------
@@
 
+* Transformation and scaling up are connected concepts. Although this research
+  focuses on transformation process only, it can not be completely separated
+  from discussion on scaling up, or on large scale agile implementation in
+  general.
+
 % ---------------------------------------------------------------------

### 2014-05-07 journal/article.tex
Work in progress

@@
 the methods were customized badly. As by-the-book implementation often was not
-feasible the agile method was tailored to suite the needs of the organization.
+feasible the agile method was tailored to suit the needs of the organization.
 However, in some cases tailoring meant skipping some practices, which lead into
@@
 
-% Challenge: Differing vocabularies
-
-In a large environment many agile teams coexist and it is possible that their
-agile models and vocabularies start diverging. When many teams implement agile
-without consistent guidance friction and fragmentation may be generated [P9].
-The organization may require moving people between teams from time to time, and
-therefore it would be desirable that no big differences in teams' agile cultures
-exist. Divergence in process would create risk increasing costs when relocating
-people [P38]. Further, if processes of teams diverged management had
-difficulties in forecasting and comparing work of different teams [P38, P48].
-To overcome problems with divergence in agile models some organizations defined
-standards to follow [P8, P43].
+% The software produced in a large organization may have size and complexity that
+% agile struggles to handle.
+
+The primary studies revealed challenges on two areas which relate particularly
+to large scale. First, a large software development endeavour must handle
+integrating the work products of many teams. Second, a large organization will
+have management challenges not seen on small scale.
 
@@
 
+% Hard to create roadmap between teams --> Teams see only their own priorities
+
+Some cases reported that introducing agile practices soon yielded improvements
+on the team level. However, the initial success was followed by challenges when
+the teams needed to work as parts of the larger surrounding orgaization [P13,
+P17, P24]. Introducing agile had created flexibility on a team level, but the
+flexibility could not be fully utilized as the surrounding organization was not
+responsive enough [P26]. Even being agile, teams had dependencies to each other
+and the dependencies were causing problems [].
+
+%% ?? -->
+Large organization put agile methods under the test of scaling up, and problems
+were encountered. One case reported use of agile across hundreds of teams, but
+management could not get a roll up of the progress without having strict rules
+on how status was reported [P48]. Schnitter and Mackert [P45] report that a
+maximum workable size in a project managed by Scrum was 130 people including 7
+development teams.
+
+Some organizations attempted to initially create a model where teams operate
+autonomously, which is well in line with agile principles. However, a number of
+issues arose from independence. For instance, teams had to balance between the
+team's own goals and broader goals of the entire organization, and some teams
+chose focus on their own goal [P7]. Coordination was obstructed by independent
+teams that did not respect the larger context they were working in [P13].
+Coordination was also difficult when different teams were working for different
+customers but the applications being built were interdependent [P29]. One case
+reported that allowing teams to have different sprint durations created delays
+in delivery [P51].
+
+One case reported that team centricity created also distrust between teams [P24].
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+DEPENDENCIES
+
+* Dependencies --> Because of a general roadmap
+* Dependencies --> Teams must work for a general higher level goal
+* Management burden to manage dependencies --> Creates problem with roadmapping
 * Hard to create roadmap between teams --> Teams see only their own priorities
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-"C: Coordinating/integrating work of many teams"
+* Reduce dependencies --> Help roadmapping
+[P10]
+Now teams were comprised of tight-knit groups that – while collocated wherever
+possible – were responsible to deliver value regardless of location, technology
+or historical allegiance.
+However, within six months or so, and despite a pattern of success and
+improvement, it became very clear that the new forking between horizontal and
+vertical teams placed an enormous burden on the product and program management
+teams to manage dependencies in a way that supported the broader organizational
+roadmap. Teams were structured to execute independently, but the broader product
+roadmap called for features that implied specific platform improvements.
+Establishing a reliable roadmap therefore required massive (and often
+unreliable) coordination between teams that were more often than not head-down
+in their own (sub-sprint) priorities.
+Theoretically, the product owners (assisted by the Music PMO) would ensure that
+capabilities required by the roadmap would appear on their own team’s backlog at
+an appropriate level of priority. In practice, it was very difficult to identify
+and track such dependencies, especially when the feature set (and associated set
+of dependencies) was constantly in flux.
+Phase 2 of the transition (continuing today – see Figure 4) therefore attempts
+to further reduce such dependencies and improve how they are managed over time
+to improve release and roadmap planning.
+
+CODE SYNCHRONIZATION
+
+* Added problems by agile --> Software interfaces not properly synchronized
+* Problems with larger features --> dependencies in code
+* Synchronization to main waterfall problematic
+[P26]
+Using scrum only on the team level had led to additional problems. Software
+interfaces were not properly synchronized with other development teams.
+Larger features were causing synchronization problems due to dependencies in
+code. Synchronization to the main waterfall program was considered problematic.
+Deploying scrum on the team level had brought flexibility in the team, but the
+ability to embrace the change was not truly benefiting the whole program as it
+was lacking capabilities to systematically embed the changes in all teams and
+levels.
 
@@
 
-* Coordination challenge --> Team goal or product goal?
-* Team centric focus created strong boundaries --> Bad for coordination
-[P7]
-A challenge we faced in team coordination was understanding whether to optimize
-the team goal or the product goal. Some teams focused too much on their own
-goals and created strong boundaries, impeding coordination.
+* Problems with inter-team coordination --> Architecture fragile, code dialects, distrust between teams
+* (Multisite, teams on different sites)
+[P24]
+No matter how much communication we tried to establish between the teams, we
+were still far away from shared team focus on the complete product. As a result
+we could see the architecture becoming more fragile, multiple code dialects and
+distrust between teams were also starting to grow.
+We still needed to improve the focus on the complete product and not on "bad
+code” from other teams. The main barrier for building quality into the product
+was the distrust between the teams who worked on the same product. Therefore the
+two R\&D teams in Denmark and Bangladesh were merged into one global team.
+Instead of two distributed teams, working with different sprint backlogs, we
+established one team with one sprint backlog. The problem of building quality
+into the complete product was now a shared team problem, not a management
+problem.
+
+* Problems with cross team integration, lack of unified build environment
+[P27]
+1. Our department had a lack of program-wide build scripts, build environment,
+   data alignment and testing strategy.
+2. Cross team integration build took too long.
+
+* With Scrum each team started using their own integration server --> Problem with late integration
+* The shared integration server was taken back into use
+[P50]
+A key impediment was the distributed development model in use by the Scrum
+teams. Originally, before Scrum adoption, all teams worked together in a common
+environment. There were many integration issues as a result and no solid process
+to rectify them. So with the initial adoption of Scrum in summer 2009 it was
+decided that each individual Scrum team should have a dedicated and private
+server with integration to take place after the build. In retrospect in the fall
+of 2009, the downside of the decision to break up the environment was being felt
+as integration issues were being discovered very late, as  the build process was
+the first point of integration. To overcome this impediment the decision was
+made to re-integrate all teams onto a single server environment again, so that
+integration issues could be rectified before they left development.
+
+OTHER ``DEPENDENCIES''
 
@@
 
-* ?
-[P10]
-CROSS-FUNCTIONAL DEPENDENCIES: However well we did within the team, we also
-frequently ran into dependencies on other groups often in other locations –
-groups such as UED (User Experience and Design), operations, legal and the
-Yahoo! security team (affectionately called the Paranoids) – which were
-identified as blocking issues but almost never neatly resolved within our sprint
-boundaries.
-
-* Well into the transformation there was trouble with dependencies between teams
-* Management burden to manage dependencies --> Creates problem with roadmapping
-* Hard to create roadmap between teams --> Teams see only their own priorities
-* Reduce dependencies --> Help roadmapping
-[P10]
-Now  teams were  comprised of tight-knit groups that – while collocated wherever
-possible – were  responsible to deliver  value regardless of location,
-technology or historical allegiance.
-However, within six months or so, and despite a pattern  of  success and
-improvement, it became very clear that the new forking between horizontal
-and vertical  teams  placed an enormous burden on the product and program
-management teams to manage dependencies in a way that supported  the  broader
-organizational roadmap.  Teams were structured to execute independently,  but
-the broader product roadmap called for features that  implied  specific platform
-improvements. Establishing a  reliable roadmap therefore required massive (and 
-often unreliable) coordination between teams that were more often  than not
-head-down in their own (sub-sprint) priorities.
-Theoretically,  the product owners (assisted by the Music PMO) would ensure that
-capabilities required by the  roadmap would appear on their own team’s backlog
-at an appropriate level of priority.  In practice, it was very difficult to
-identify  and track such dependencies, especially when the  feature  set  (and
-associated set of dependencies) was constantly in flux.
-Phase 2 of the transition (continuing today –  see Figure 4) therefore  attempts
-to further reduce such dependencies and improve how they are managed over time
-to improve release and roadmap planning.
-
-* Team centricity --> Power --> However, they operate with independent priorities --> No respect for larger context (release)
+COORDINATION
+
+* Managers --> Getting big picture, overview
+* Coordination challenges across teams
+* Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
+[P17]
+As individual component teams began to experience success, the challenge changed
+to coordinating across teams. Although iterations were humming along at the team
+level it was initially difficult for managers to get a complete picture of how
+the release was progressing. They were unable to tell when particular features
+were complete as their requirements crossed team boundaries.
+Using Rally’s application for managing the development lifecycle, BMC is now
+able to guide the overall release objectives with themes and high-level
+features.
+
+* Size and complexity --> Scrum-of-Scrums is mandatory 
+[P45]
+The size and complexity of SAP’s products makes it necessary to consider these
+techniques. To co-ordinate the work of multiple Scrum teams and to communicate
+the requirements via team product backlogs, various approaches were combined.
+Special product teams were formed for this purpose. These teams can be regarded
+as both a permanent Scrum of Scrums and a supervisory product owner.
+
+* Problem: Tracking the work of hundreds of teams --> Handicap in providing executive summary on progress
+[P48]
+Tracking progress of hundreds of teams working on pieces of major features
+proved to be a major problem. Tools that we were using did not have roll up
+capability without imposing strict rules on how data was entered, labeled and
+updated. We found out our tools were not really scalable for the size and
+complexity of the programs we were working on. This proved a serious handicap in
+providing executive summary information about progress.
+
+"C: Coordinating/integrating work of many teams"
+
+HELP / SOLUTIONS
+
+* Individual component teams gained success --> Coordination challenges across teams
+* Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
+* Help: Tool support to visualize dependencies for people
+[P17]
+As individual component teams began to experience success, the challenge changed
+to coordinating across teams. Although iterations were humming along at the team
+level it was initially difficult for managers to get a complete picture of how
+the release was progressing. They were unable to tell when particular features
+were complete as their requirements crossed team boundaries.
+Using Rally’s application for managing the development lifecycle, BMC is now
+able to guide the overall release objectives with themes and high-level
+features.
+
 * Release Framework was created
@@
 on a daily basis, they also experienced benefits from it as well.
-
-* Individual component teams gained success --> Coordination challenges across teams
-* Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
-* Help: Tool support to visualize dependencies for people
-[P17]
-As individual component teams began to experience success, the challenge changed
-to coordinating across teams. Although iterations were humming along at the team
-level it was initially difficult for managers to get a complete picture of how
-the release was progressing. They were unable to tell when particular features
-were complete as their requirements crossed team boundaries.
-Using Rally’s application for managing the development lifecycle, BMC is now
-able to guide the overall release objectives with themes and high-level
-features.
 
@@
 
-* Scrum on team level --> Software interfaces not properly synchronized
-* Problems with larger features
-* Scrum on team level good --> True benefit requires entire program to be flexible for changes
-[P26]
-Using scrum only on the team level had led to additional problems. Software
-interfaces were not properly synchronized with other development teams.
-Larger features were causing synchronization problems due to dependencies in
-code. Synchronization to the main waterfall program was considered problematic.
-Deploying scrum on the team level had brought flexibility in the team, but the
-ability to embrace the change was not truly benefiting the whole program as it
-was lacking capabilities to systematically embed the changes in all teams and
-levels.
-
-* Problems with cross team integration, lack of unified build environment
-[P27]
-1. Our department had a lack of program-wide build scripts, build environment,
-   data alignment and testing strategy.
-2. Cross team integration build took too long.
-
-* Different customers compete --> Customer centricity does not view the system as a whole
-[P29]
-each customer base views the system only via their process flow.
-We have multiple competing customers crossing multiple co-dependent
-applications, including dependencies on applications outside of our
-organization; multiple funding sources; and a mixture of Waterfall and Scrum
-projects.
-For our software teams to be able to execute efficiently, we will need to push
-the complexity and prioritization upstream, scaling teams to meet demand while
-maintaining clean, simple prioritized Product Backlogs for each product.
+* Scrum of scrums to resolve inter-team dependencies
+[P43]
+We implemented a scrum-of-scrums run by our release management team to identify
+inter-team dependencies and resolve merge issues before they occur.
+
+* Solution: Product teams
+[P45]
+Product teams were introduced as a second organizational layer above the Scrum
+teams. A product team is responsible for the work of up to 7 development teams.
+It consists of the product owners of those teams and the same number of team
+members who are specialists in particular fields. Depending on the problem area
+other experts may be included. This allows for full engineering coverage of all
+problem areas expected, well-organized communication among the teams, and direct
+communication with the development teams to detect and mitigate risks.
+
+% --
 
@@
 
-[P43] Ryan 2008 - The Price of Agile Is Eternal Vigilance.pdf - 43:11 [We implemented a scrum-of-scru..]  (4:48-4:200)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
-No memos
-
-We 
-implemented a scrum-of-scrums run by our release 
-management team to identify inter-team dependencies 
-and resolve merge issues before they occur.
-
-[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:18 [What could not be mitigated we..]  (4:2133-4:2292)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
-No memos
-
-What could not be mitigated were issues with teams dependent on “unreliable”
-Scrum teams, at least if the Scrum teams were isolated in a non-Scrum environment.
-
-[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:28 [Several approaches are known t..]  (8:522-8:1054)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
-No memos
-
-Several approaches are known to co-ordinate the work of several Scrum teams working
-on one product, e.. Scrum of Scrums [9] or MetaScrum [18]. The size and complexity
-of SAP’s products makes it necessary to consider these techniques. To co-ordinate the
-work of multiple Scrum teams and to communicate the requirements via team product
-backlogs, various approaches were combined. Special product teams were formed for
-this purpose. These teams can be regarded as both a permanent Scrum of Scrums and a
-supervisory product owner.
-
-[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:32 [Product teams were introduced ..]  (8:1076-8:1627)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
-No memos
-
-Product teams were introduced as a second organizational layer above the Scrum teams.
-A product team is responsible for the work of up to 7 development teams. It consists
-of the product owners of those teams and the same number of team members who
-are specialists in particular ﬁelds. Depending on the problem area other experts may
-be included. This allows for full engineering coverage of all problem areas expected,
-well-organized communication among the teams, and direct communication with the
-development teams to detect and mitigate risks.
-
-[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:33 [Product teams can rarely suppo..]  (9:841-9:1526)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
-No memos
-
-Product teams can rarely support more than seven Scrum teams. Ifthe products or components are be too big or too complex to be developed in this organizational setup, an
-intermediate organizational layer is inserted between the product team and the Scrum
-teams. These teams are called “area product teams”. Their responsibility is the same as
-the product teams’ responsibility with the exception of budget and ﬁnal product decisions.
-In this three-layer organization a productteam works similar to a project management ofﬁce.
-The complete budget and product responsibility is with the product team,
-while the operational day-to-day decisions are handled by the area product teams.
-
-[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:38 [Scaling Scrum to a multi-team ..]  (10:1223-10:1784)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [O: Large scale definition - Family: Contextual] 
-No memos
-
-Scaling Scrum to a multi-team development organization is not as easy. Scrum of
-Scrums and backlog preparation for many teams can be combined by product teams
-as outlined above. Looking at all known parameters of agile project management at
-SAP we tend to say that not much more than 130 development employees may be
-organizedin that fashion. This number sums up developers in 7 teams (max. 70 people),
-the product team (max. 16), developmentinfrastructure responsibles (about 10), quality
-assurance and testers (about 25), general management (about 10).
-
-[P48] Smits 2011 - Agile Experience Report - Transition And Complexity At Cisco Voice Technology Group.pdf - 48:30 [Tracking progress of hundreds ..]  (4:1657-4:2111)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
-No memos
-
-Tracking progress of hundreds of teams working on 
-pieces of major features proved to be a major 
-problem. Tools that we were using did not have roll 
-up capability without imposing strict rules on how 
-data was entered, labeled and updated. We found out 
-our tools were not really scalable for the size and 
-complexity of the programs we were working on. 
-This proved a serious handicap in providing 
-executive summary information about progress.
-
-[P50] Sutherland 2011 - Hitting The Wall - What To Do When High Performing Scrum Teams Overwhelm Operations And Infrastructure.pdf - 50:20 [A key impediment was the distr..]  (4:1737-4:2643)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
-No memos
-
-A key impediment was the distributed 
-development model in use by the Scrum teams.  
-Originally, before Scrum adoption, all teams worked
-together in a common environment.  There were 
-many integration issues as a result and no solid 
-process to rectify them.  So with the initial adoption 
-of Scrum in summer 2009 it was decided that each 
-individual Scrum team should have a dedicated and 
-private server with integration to take place after the 
-build.  In retrospect in the fall of 2009, the downside 
-of the decision to break up the environment was 
-being felt as integration issues were being discovered 
-very late,  as  the  build  process  was  the  first  point  of 
-integration.  To overcome this impediment the 
-decision was made to re-integrate all teams onto a 
-single  server  environment  again,  so  that  integration 
-issues could be rectified before they left 
-development.
-
-[P51] Vlaanderen 2012 - Growing Into Agility - Process Implementation Paths For Scrum.pdf - 51:12 [Initially, the sprints lasted ..]  (5:230-5:717)   (Super)
-Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
-No memos
-
-Initially, the sprints lasted two or three weeks, depending on
-the team, causing a discrepancy between the sprint durations of some of the
-teams. At the same time, interdependencies started to become more of an issue
-as the amount of teams grew. This caused problems when, for instance, frontend teams needed functionality in the back-end that was not yet developed.
-These problems, which consisted mainly of delays, were solved by ﬁxating sprint
-duration at two weeks for all teams.
+% INDIVIDUAL TEAMS GOOD --> PROBLEMS WHEN MANY TEAMS NEED TO WORK TOGETHER
+% 
+% * Individual component teams gained success --> Coordination challenges across teams
+% * Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
+% * Help: Tool support to visualize dependencies for people
+% [P17]
+% As individual component teams began to experience success, the challenge changed
+% to coordinating across teams. Although iterations were humming along at the team
+% level it was initially difficult for managers to get a complete picture of how
+% the release was progressing. They were unable to tell when particular features
+% were complete as their requirements crossed team boundaries.
+% Using Rally’s application for managing the development lifecycle, BMC is now
+% able to guide the overall release objectives with themes and high-level
+% features.
+% 
+% 
+% * Team centricity --> Power --> However, they operate with independent priorities --> No respect for larger context (release)
+% * Release Framework was created
+% [P13]
+% Shortly after the initial rollout to 30+ product teams it was clear that we had
+% unleashed the power of our individual teams through their new found
+% self-organization and focus on short sprints.  However, since we aggregate the
+% work of many sprints into a single release it was also becoming clear that our
+% scrum teams were operating independently without respect for the context of the
+% larger release.  Herding 30+ teams working in the same codeline to a predictable
+% time-boxed release date is very challenging in any traditional development model
+% let alone an agile development model where we value self-organization over
+% centralized control.
+% Our answer to this problem was a lightweight Release Framework and the
+% introduction of a Release Manager for the overall product release.
+% 
+% 
+% * Initial adopting good
+% * Problems with inter-team coordination --> Architecture fragile, code dialects, distrust between teams
+% * (Multisite, teams on different sites)
+% * Help: Retrospectives across teams --> Created global teams --> Shared backlog
+% [P24]
+% To embrace and implement Scrum across the different teams and organizations, we
+% focused on doing regular retrospectives on both team and company level.
+% The first couple of company retrospectives were quite hectic and required some
+% extra facilitation techniques to be constructive. But we did identify some of
+% the main problems when crossing the team boundaries and working together with
+% other departments.
+% No matter how much communication we tried to establish between the teams, we
+% were still far away from shared team focus on the complete product. As a result
+% we could see the architecture becoming more fragile, multiple code dialects and
+% distrust between teams were also starting to grow.
+% By doing regular retrospectives with all the different teams and across the
+% company we identified the main problems and decided it was time to create the
+% global Scrum teams. We wanted to improve communication and trust by working on
+% shared commitments between the different teams.
+% 
+% We still needed to improve the focus on the complete product and not on "bad
+% code” from other teams. The main barrier for building quality into the product
+% was the distrust between the teams who worked on the same product. Therefore the
+% two R\&D teams in Denmark and Bangladesh were merged into one global team.
+% Instead of two distributed teams, working with different sprint backlogs, we
+% established one team with one sprint backlog. The problem of building quality
+% into the complete product was now a shared team problem, not a management
+% problem.
+% 
+% 
+% * Scrum on team level --> Software interfaces not properly synchronized
+% * Problems with larger features
+% * Scrum on team level good --> True benefit requires entire program to be flexible for changes
+% [P26]
+% Using scrum only on the team level had led to additional problems. Software
+% interfaces were not properly synchronized with other development teams.
+% Larger features were causing synchronization problems due to dependencies in
+% code. Synchronization to the main waterfall program was considered problematic.
+% Deploying scrum on the team level had brought flexibility in the team, but the
+% ability to embrace the change was not truly benefiting the whole program as it
+% was lacking capabilities to systematically embed the changes in all teams and
+% levels.
+%
 
 % ---------------------------------------------------------------------
+
+% Challenge: Differing vocabularies
+
+In a large environment many agile teams coexist and it is possible that their
+agile models and vocabularies start diverging. When many teams implement agile
+without consistent guidance friction and fragmentation may be generated [P9].
+The organization may require moving people between teams from time to time, and
+therefore it would be desirable that no big differences in teams' agile cultures
+exist. Divergence in process would create risk increasing costs when relocating
+people [P38]. Further, if processes of teams diverged management had
+difficulties in forecasting and comparing work of different teams [P38, P48].
+To overcome problems with divergence in agile models some organizations defined
+standards to follow [P8, P43].
 
@@
 
+* Transformation of an organization to use agile can be seen as transforming
+  also the agile method to suit the organization.
+  --> A large scale agile model may require tailoring and trial over time
+
+* It can be argued that any development effort with hundreds of teams will be
+  difficult to coordinate
+
+* Even though most of the primary studies are experience reports and subject to
+  bias, this study can draw valuable conclusions. --> If a matter is being
+  talked about it must be of some significance.
+
 % ---------------------------------------------------------------------

### 2014-05-07 journal/article.tex
Work in progress

--- 2014-05-06 journal/article.tex
@@
 
+% The software produced in a large organization may have size and complexity that
+% agile struggles to handle.
+
+The primary studies revealed challenges on two areas which relate particularly
+to large scale. First, a large software development endeavour must handle
+integrating the work products of many teams. Second, a large organization will
+have management challenges not seen on small scale.
+
+% Coordinating between many teams
+
+%% ?? -->
+% Include this ???
+% Large organization put agile methods under the test of scaling up, and problems
+% were encountered. One case reported use of agile across hundreds of teams, but
+% management could not get a roll up of the progress without having strict rules
+% on how status was reported [P48]. Schnitter and Mackert [P45] report that a
+% maximum workable size in a project managed by Scrum was 130 people including 7
+% development teams.
+
+Many cases reported that introducing agile practices soon yielded improvements
+on the team level, but as a part of a large transformation teams needed to
+change their external behavior also.
+The initial success was followed by challenges when the teams needed to work as
+parts of the larger surrounding orgaization [P13, P17, P24]. Introducing agile
+had created flexibility on a team level, but it could not be fully utilized as
+the surrounding organization was not responsive enough [P26].
+There were dependencies between teams in the agile model as well, and the
+dependencies made managing development difficult [P9, P10].
+
+In the agile model it was hard for managers to get an overview of the progress
+of the release [P17]. The status of features was hard to tell because different
+aspects of features were sepatated across teams [P17]. Smits and Rilliet [P48]
+report that providing executive overview of the development programs was limited
+by the tools used to manage agile projects.
+
+Some organizations attempted to initially create a model where teams operate
+autonomously, which is well in line with agile principles. However, a number of
+issues arose from independence. For instance, teams had to balance between the
+team's own goals and broader goals of the entire organization, and some teams
+chose focus on their own goal [P7]. Coordination was obstructed by independent
+teams that did not respect the larger context they were working in [P13].
+Coordination was also difficult when different teams were working for different
+customers but the applications being built were interdependent [P29]. One case
+reported that allowing teams to have different sprint durations created delays
+in delivery [P51].
+
+Also technical problems relating to inter-team coordination were reported.
+Integrating the products of teams was seen as a problem in some cases [P5, P50],
+and lack of standardized build scripts was mentioned as a problem in one case
+[P27].
+There was also problems in synchronizing definitions of software interfaces
+between teams, and dependencies in code caused problems in larger features
+[P26].
+One case reported that team centricity in the agile model created a fragile
+architecture, divergence in coding style, and even distrust between teams [P24].
+
+Many studies presented solutions for the inter-team challenges, which also
+highlight the challenges themselves. The general model of solving inter-team
+problems was introducing an additional layer of coordination, such as a
+Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
+customized release framework [P13]. In one case integration problems were
+controlled by prioritizing code quality and maintaining existing functionality
+over new features [P13]. Finally, the choice of tools was reported to have an
+impact on the coordination effort [P17].
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+[P43] Transformation is described. Focus on “scaling up” part of adoption.
+
+% --
+
+* First cross team dependencies were reduced --> Problems with too much dependencies
+* Traditional dependency management limited flexibility, which was demanded --> Inflexibility, work intensive, adversarial relationships
+[P34]
+From the onset, we sought to reduce cross-team dependencies as much as possible,
+allowing each of the Scrum teams to “run” at their fastest pace.  However, as
+the number of concurrent work streams increased, so did the number of
+dependencies.  In particular, we found that adopting a SOA ran counter to our
+efforts.
+We quickly faced the prospect of implementing and integrating with a sizeable
+set of centralized services that would be consumed by multiple project teams.
+Within a few months, we had accumulated a significant number of integration
+points, preventing us from delivering fully production-ready software until they
+were resolved.
+Accepting the fact that these dependencies would be part of the project, we
+sought to create a lightweight and efficient process for resolving dependencies
+and integration points.  We recognized that traditional approaches to dependency
+management tend to place significant constraints on the sequence of work,
+reducing the teams’ ability to independently prioritize their backlogs. 
+Furthermore, highly contractual, document-centric approaches to defining
+integration points are work-intensive, prone to error and can create adversarial
+relationships between teams.  We wanted to establish an approach based on direct
+communication and collaboration.
+Our initial integration efforts involving five of the module teams of the Core
+Application were reasonably successful. After some initial scrambling, the teams
+self-organized and established a set of procedures for managing inter-module
+integration.
+
+* Worked decently with 5 teams and SoS --> Problems with global deployment
+[P34]
+However, our efforts to facilitate cross-team collaboration within the Core
+Application benefited heavily from the fact that the five module teams were all
+closely co-located, within the same organizational structure and employed an
+active “Scrum of Scrums” model to ensure communication and coordination between
+the teams.  As we pushed this approach out further across the program and the
+globe, we began to experience significant challenges.
+
+* Problem: Agile teams were thinking inside the team room walls
+* Success: Thinking of the teams that will integrate to your work
+* Failure: Thinking only inside the team walls
+[P34]
+We found that we had varying degrees of success in managing heavily-used
+services.  In those cases where we were successful, it was invariably due to a
+small group of people from both the consuming and providing teams taking
+ownership of the services and proactively seeking out future integration points
+and potential challenges.  Conversely, in those situations where such ownership
+did not exist and people remained within their team walls, we invariably
+experienced integration challenges: services consistently failing to meet the
+consumer’s requirements and integration points that were poorly tested and
+failed in system testing.
+
+* Global effort problems --> Communication regressed to contractual --> Narrow communication channels were used: documents, e-mail
+* Success: Individuals that saw effort in wide-channel communication, Individuals that thought outside team room walls
+[P34]
+As more and more integration occurred across the globe, we witnessed a tendency
+to revert back to contractual forms of communication. As always, distributing
+the development team significantly increases the challenge to directly
+communicate and collaborate.  The natural tendency to rely on e-mail due to
+differences in time zones frequently led to communication breakdowns and
+unnecessary blockages – forming barriers to feedback. In some instances,
+individuals broke through these barriers using instant messaging products, web
+cams and adjusted working hours to facilitate as much direct communication as
+possible.  At the same time other teams regressed, relying on documentation and
+establishing an almost adversarial relationship with their integration partners.
+
+* Large program problems --> Green builds hard to maintain
+* Tools insufficient for large
+[P34]
+Given the size and distribution of our development team and the fact that the
+program produced a single, integrated enterprise application, the complexity of
+the build process and the level of effort to maintain it was significant.   As
+the program grew in size, we recognized that the tool set with which we began
+the project was no longer sufficient.  We began experiencing nearly continuous
+broken builds due to conflation and timing issues on the build server.  At this
+point, we began a lengthy initiative to replace the toolsets and revamp the
+entire build process using AntHillPro and Subversion.  As a result, we created a
+build process that effectively managed our myriad of dependencies and supported
+a total build of the application in a reasonable time.
+
+* Problems: Integration and dependency management
+* Goal: Foster communication and collaboration instead of rigid processes
+* Problems: Team avoiding external distractions, balance between team membership and project membership --> CI and green builds did not work
+* Problems: Technical leads must work outside of team --> To coordinate
+* Solution: Balance between own goals and project goals --> Think outside team room
+* The team must think of others than them selves, too
+[P34]
+Over the lifetime of the project we encountered two key challenges due to the
+scope and scale of this effort: integration and dependency management and
+continuous integration. When addressing these challenges, we sought to adhere to
+the basic principals of the Agile manifesto as much as possible: making the
+delivery of working software our overriding goal and fostering communication and
+collaboration instead of dictating rigid processes. We sought to empower
+individuals and teams to make decisions and formulate their own processes, and
+to take calculated risks without fear of reprisals.
+
+However, we found that some of the qualities and practices that made the teams
+work effectively in isolation actually reduced their effectiveness in the
+large-scale environment.
+For example: in smaller projects, we work aggressively to avoid external
+distractions and encourage team members to focus exclusively on their individual
+team goals as expressed in the backlog.  In large-scale projects, team members
+must participate in cross-team activities and allocate time to project-wide
+activities.  In particular, team technical leads were required to dedicate
+significant portions of their time outside of the team room.  We found that many
+of our team leads struggled with this task.  While they recognized the need for
+cross-team communication, they resented context switching and the impact on
+their time dedicated to the team’s sprint goals.
+Furthermore, all of the teams struggled with balancing their identity as a team
+member against their membership in the larger project. This was reflected most
+keenly in our failure to sustain continuous integration and produce frequent
+green builds.
+
+Agile team must make trade-offs between stabilizing an application and
+completing more story cards, the size of this project required teams to put
+significantly more time into tracking down issues that were caused by other
+teams or dependencies.   Many team members found this frustrating and
+demoralizing.
+
+% ---------------------------------------------------------------------
+
 % Challenge: Differing vocabularies
@@
 standards to follow [P8, P43].
-
-% Coordinating between many teams
-
-* Hard to create roadmap between teams --> Teams see only their own priorities
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
 
@@
 
+* Transformation and scaling up are connected concepts. Although this research
+  focuses on transformation process only, it can not be completely separated
+  from discussion on scaling up, or on large scale agile implementation in
+  general.
+
+* Transformation of an organization to use agile can be seen as transforming
+  also the agile method to suit the organization.
+  --> A large scale agile model may require tailoring and trial over time
+
+* It can be argued that any development effort with hundreds of teams will be
+  difficult to coordinate
+
+* Even though most of the primary studies are experience reports and subject to
+  bias, this study can draw valuable conclusions. --> If a matter is being
+  talked about it must be of some significance.
+
 % ---------------------------------------------------------------------

### 2014-05-08 journal/article.tex
Coordinating work between many teams

@@
 
+Moore and Spens [P34] describe the progression of the above mentioned
+coordination problems. At first cross-team dependencies were attempted to be
+reduced, but it became evident that the dependencies were an inherent part of
+the project. A traditional approach to managing dependencies created excess
+work, reduced independence and flexibility of teams, and created contract based
+and adversarial relationships within the organization. There was some success
+when up to 5 teams were coordinated using Scrum-of-Scrums, but that model could
+not be scaled to a global level. The main problem was teams that were thinking
+too much inside the team walls. When the new practices were scaled up the
+communication channels narrowed and created communication breakdowns. Problems
+with collaboration, dependencies, and integration were solved with presnce of
+individuals with an proactive and open mindset towards working over team
+boundaries. It was concluded that a balance is needed between completing new
+stories from the team backlog and maintaining overall stability of the
+application. [P34]
+
 Many studies presented solutions for the inter-team challenges, which also
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-[P43] Transformation is described. Focus on “scaling up” part of adoption.
-
-% --
-
-* First cross team dependencies were reduced --> Problems with too much dependencies
-* Traditional dependency management limited flexibility, which was demanded --> Inflexibility, work intensive, adversarial relationships
-[P34]
-From the onset, we sought to reduce cross-team dependencies as much as possible,
-allowing each of the Scrum teams to “run” at their fastest pace.  However, as
-the number of concurrent work streams increased, so did the number of
-dependencies.  In particular, we found that adopting a SOA ran counter to our
-efforts.
-We quickly faced the prospect of implementing and integrating with a sizeable
-set of centralized services that would be consumed by multiple project teams.
-Within a few months, we had accumulated a significant number of integration
-points, preventing us from delivering fully production-ready software until they
-were resolved.
-Accepting the fact that these dependencies would be part of the project, we
-sought to create a lightweight and efficient process for resolving dependencies
-and integration points.  We recognized that traditional approaches to dependency
-management tend to place significant constraints on the sequence of work,
-reducing the teams’ ability to independently prioritize their backlogs. 
-Furthermore, highly contractual, document-centric approaches to defining
-integration points are work-intensive, prone to error and can create adversarial
-relationships between teams.  We wanted to establish an approach based on direct
-communication and collaboration.
-Our initial integration efforts involving five of the module teams of the Core
-Application were reasonably successful. After some initial scrambling, the teams
-self-organized and established a set of procedures for managing inter-module
-integration.
-
-* Worked decently with 5 teams and SoS --> Problems with global deployment
-[P34]
-However, our efforts to facilitate cross-team collaboration within the Core
-Application benefited heavily from the fact that the five module teams were all
-closely co-located, within the same organizational structure and employed an
-active “Scrum of Scrums” model to ensure communication and coordination between
-the teams.  As we pushed this approach out further across the program and the
-globe, we began to experience significant challenges.
-
-* Problem: Agile teams were thinking inside the team room walls
-* Success: Thinking of the teams that will integrate to your work
-* Failure: Thinking only inside the team walls
-[P34]
-We found that we had varying degrees of success in managing heavily-used
-services.  In those cases where we were successful, it was invariably due to a
-small group of people from both the consuming and providing teams taking
-ownership of the services and proactively seeking out future integration points
-and potential challenges.  Conversely, in those situations where such ownership
-did not exist and people remained within their team walls, we invariably
-experienced integration challenges: services consistently failing to meet the
-consumer’s requirements and integration points that were poorly tested and
-failed in system testing.
-
-* Global effort problems --> Communication regressed to contractual --> Narrow communication channels were used: documents, e-mail
-* Success: Individuals that saw effort in wide-channel communication, Individuals that thought outside team room walls
-[P34]
-As more and more integration occurred across the globe, we witnessed a tendency
-to revert back to contractual forms of communication. As always, distributing
-the development team significantly increases the challenge to directly
-communicate and collaborate.  The natural tendency to rely on e-mail due to
-differences in time zones frequently led to communication breakdowns and
-unnecessary blockages – forming barriers to feedback. In some instances,
-individuals broke through these barriers using instant messaging products, web
-cams and adjusted working hours to facilitate as much direct communication as
-possible.  At the same time other teams regressed, relying on documentation and
-establishing an almost adversarial relationship with their integration partners.
-
-* Large program problems --> Green builds hard to maintain
-* Tools insufficient for large
-[P34]
-Given the size and distribution of our development team and the fact that the
-program produced a single, integrated enterprise application, the complexity of
-the build process and the level of effort to maintain it was significant.   As
-the program grew in size, we recognized that the tool set with which we began
-the project was no longer sufficient.  We began experiencing nearly continuous
-broken builds due to conflation and timing issues on the build server.  At this
-point, we began a lengthy initiative to replace the toolsets and revamp the
-entire build process using AntHillPro and Subversion.  As a result, we created a
-build process that effectively managed our myriad of dependencies and supported
-a total build of the application in a reasonable time.
-
-* Problems: Integration and dependency management
-* Goal: Foster communication and collaboration instead of rigid processes
-* Problems: Team avoiding external distractions, balance between team membership and project membership --> CI and green builds did not work
-* Problems: Technical leads must work outside of team --> To coordinate
-* Solution: Balance between own goals and project goals --> Think outside team room
-* The team must think of others than them selves, too
-[P34]
-Over the lifetime of the project we encountered two key challenges due to the
-scope and scale of this effort: integration and dependency management and
-continuous integration. When addressing these challenges, we sought to adhere to
-the basic principals of the Agile manifesto as much as possible: making the
-delivery of working software our overriding goal and fostering communication and
-collaboration instead of dictating rigid processes. We sought to empower
-individuals and teams to make decisions and formulate their own processes, and
-to take calculated risks without fear of reprisals.
-
-However, we found that some of the qualities and practices that made the teams
-work effectively in isolation actually reduced their effectiveness in the
-large-scale environment.
-For example: in smaller projects, we work aggressively to avoid external
-distractions and encourage team members to focus exclusively on their individual
-team goals as expressed in the backlog.  In large-scale projects, team members
-must participate in cross-team activities and allocate time to project-wide
-activities.  In particular, team technical leads were required to dedicate
-significant portions of their time outside of the team room.  We found that many
-of our team leads struggled with this task.  While they recognized the need for
-cross-team communication, they resented context switching and the impact on
-their time dedicated to the team’s sprint goals.
-Furthermore, all of the teams struggled with balancing their identity as a team
-member against their membership in the larger project. This was reflected most
-keenly in our failure to sustain continuous integration and produce frequent
-green builds.
-
-Agile team must make trade-offs between stabilizing an application and
-completing more story cards, the size of this project required teams to put
-significantly more time into tracking down issues that were caused by other
-teams or dependencies.   Many team members found this frustrating and
-demoralizing.
 

### 2014-05-08 journal/article.tex
Multisite problems

@@
 
+Further problems were encountered when scaling up agile over many sites.
+Operating in a distributed environment makes communication more difficult, which
+causes problems with agile when frequent and effective communication needs to be
+achieved. Distribution had negative effects, such as missing kick-off meetings,
+reduced feeling of proximity when telecommunication is necessary, and difficulty
+in arranging frequent meetings due to time zone differences [P45].
+Agile project management was also problematic in a distributed setting. In a
+waterfall model separate parts of projects could be isolated to different sites,
+but the agile model does not allow such strict slicing of projects [P29].
+In one case it was simply admitted that a distributed organization will impose
+additional burden on communication and require additional care in the process,
+but distribution and agile would still be used together [P48].
+
 Moore and Spens [P34] describe the progression of the above mentioned

### 2014-05-08 journal/article.tex
Challenge: Previous organization in silos

@@
 
+% Challenge: Previous organization in silos is haunting new model
+% Is this good??
+
+In some cases problems were caused by internal boundaries or silos in the
+initial organization. Cloke [P10] describes how use of Scrum revealed an
+internal segmentation where teams operated with differing priorities and
+agendas. This segmentation also hampered the initial effort to use Scrum in a
+larger context [P10].
+Projects needed specialized skills that were scarce and people were often
+relocated to match the needs of the moment, but too much reorganization made
+it diffucult for teams to plan ahead [P11].
+A dispersed matrix organization causes also problems when people are on one hand
+kept assigned to many projects simultaneously, but on the other hand functions
+such as database administration, architecture boards, and project management are
+centralized [P14].
+
 % Challenge: Differing vocabularies

### 2014-05-08 journal/article.tex
Future research ideas

@@
 
+* Compare to change leadership research, and organizational research. Which
+  challenges are typical to any organization, which are specific to software
+  development organizations.
+
+* Compare how recommendations in large scale application models of agile relate
+  to the challenges and success factors reported in the transformaiton proecss.
+
 % ---------------------------------------------------------------------

### 2014-05-08 journal/article.tex
Editing

@@
 therefore failing to achieve a real change in process and mindset [P9, P46].
+Another example of bad customization that counteracts change is replacing agile
+vocabulary by favoring familiar terminology from the old model [P44].
 
@@
 change their external behavior also.
-The initial success was followed by challenges when the teams needed to work as
-parts of the larger surrounding orgaization [P13, P17, P24]. Introducing agile
-had created flexibility on a team level, but it could not be fully utilized as
-the surrounding organization was not responsive enough [P26].
+The initial success was followed by challenges when teams needed to work with
+other teams and as parts of the larger surrounding orgaization [P13, P17, P24].
+Introducing agile had created flexibility on a team level, but it could not be
+fully utilized as the surrounding organization was not responsive enough [P26].
 There were dependencies between teams in the agile model as well, and the

### 2014-05-08 journal/article.tex
Analysis; Work in progress

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
 % Challenge: Previous organization in silos is haunting new model
@@
 
+% ---------------------------------------------------------------------
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-WORKING IN SILOS --> Certain expertise is required in many places
-
-* Teams were reorganized without notice --> Planning was hampered when people did not konw who would stay on a team
-* Frequent changes --> People managers had to agree to minimize team changes
-[P11]
-When we launched multiple Scrum teams, we tried to keep existing teams together
-and form new teams based on skills.  As we gained experience, we realized that
-there were skills shortages in certain types of expertise that wasn’t possible
-to develop in a sprint or two. There was pressure to move engineers with certain
-skills to the teams that needed them for work on upcoming sprints or to address
-critical customer requirements.  When moves occurred without much notice, teams
-couldn’t plan effectively if they didn’t know who would stay on a team.  After
-several months of frequent changes, product owners and people managers agreed to
-minimize team changes, and even talked of instituting a “trading deadline.”
+("C: Part of organization is in old mindset" | "C: Using old and new models side-by-side")
+
+* The agile model had to made fit in the context of the entire enterprise
+* There were old projects using existing approaches
+* General challenge of organizing
+[P8]
+The final concern was to align the agile delivery processes with the
+pre-existing delivery model in use at ABC Bank. Not only were many projects
+already in-flight using existing development approaches, but also many new
+projects would be a blend of traditional and agile approaches. A clear
+relationship to the existing delivery approaches was essential.
+
+* Bottom up transformation
+* Corporate level --> Use of waterfall
+* There was tension built up between these approaches --> within teams --> in management
+* Problems --> Generic ``problems'' --> Particularly when collaborating with external teams
+* An inception sprint --> Some sees as necessary --> Others consider as failing in iterativeness --> Possible to avoid being agile
+[P10]
+at a corporate level Yahoo! still embraced a more traditional (top-down,
+heavy-weight, big design up front, waterfall) product development process (known
+rather unimaginatively as PDP). The tension between these two approaches – and a
+corresponding confusion – began to show up at all levels within the company –
+within the sprint teams themselves, within the leadership of organizations such
+as Music, and particularly when teams from different organizations were
+attempting to collaborate on products while using  fundamentally different
+approaches to develop them.
+
+The sometimes awkward co-existence of agile and traditional product lifecycles
+continues to be a challenge at Yahoo!  For instance, one hotly contested
+concession to the ‘big design up front’ mentality was to accommodate a ‘Sprint
+0’ of indeterminate length that would allow for some of the more traditional
+planning, specification and design activities to be completed.  In the minds of
+some, this was a necessary bridge between the sprint-focused world of scrum and
+the traditional domain of portfolio management and organizational planning.  In
+the minds of many, however, the ‘Sprint 0’ stipulation struck at the heart of
+what it meant to iteratively refine an emergent solution within the agile
+process, and  in some cases it allowed teams and individuals to avoid being
+agile at all by relegating all their work to this formation sprint.
+
+* People are pulled off to fight fires (old commitments?)
+[P17]
+Within the first two iterations the pilot teams saw every common impediment to
+producing fully-tested software in rapid iterations – team members were pulled
+off project to fight fires; architectural blocks arose due to multiplexed
+resources; there was inadequate access to product owners; and development teams
+were spread across as many as four cities.
+
+* Management requires increasing documentation --> Extra burden on agile teams
+[P20]
+The documentation requirements were driven basically from the upper management,
+ISO certificate requirements and the traditional development method that is
+currently used. Although agile development promises sufficient documentation of
+the projects, it didn’t seem very convincing to the upper management when they
+end up receiving few documents in comparison with the previous model of
+documentation.
+Many attempts were made to balance between the upper management requirement
+regarding documentation and between adopting Scrum method. Agile teams started
+increasing the number of documents required for documentation and started to
+customize Scrum as much as possible to conform with all the upper management
+requirements of documentation norms.
+This did not work very well and it created extra burden on the agile teams.
+
+* The Scrum team was forced to comply with approval procedures --> unnecessary work
+[P21]
+The traditional method currently in S was customized to comply with the
+governmental system of other department. The new agile method being introduced,
+Scrum, is developed in such highly bureaucratic environment. The agile team has
+to secure approvals and signatures before moving from one step to another. This
+was perceived by the team members as unnecessary and more time was taken into
+account to develop a new project.
+The scrum method requires much less correspondence, less time in communication
+between customer and the team and requires significantly less paper work and
+approvals as the customer is supposed to be involved in every step. Here, the
+governmental bureaucratic system came on the way of succeeding in the adoption
+of agile.
+
+* Generic problem: Synchronizing with the waterfall program
+[P26]
+Using scrum only on the team level had led to additional problems. Software
+interfaces were not properly synchronized with other development teams.
+Larger features were causing synchronization problems due to dependencies in
+code. Synchronization to the main waterfall program was considered problematic.
+Deploying scrum on the team level had brought flexibility in the team, but the
+ability to embrace the change was not truly benefiting the whole program as it
+was lacking capabilities to systematically embed the changes in all teams and
+levels.
+
+* Teams were forced to work for several programs in parallel
+* Working in this way revealed that overloading of tasks was problematic
+[P26]
+For practical reasons, some scrum teams needed to work for several programs.
+Sharing resources has not been seen as a good option by the programs, as it
+results in non-steady output level and makes estimation harder.
+Programs with less dedicated resources have not made very good progress. In
+fact, agile program model deployment has revealed the overload of tasks to
+programs and teams and the true development velocity and capability.
+
+Comment:
+"Programs" or "program model is seen as the "old model" here. I am not entirely
+sure whether that was the "old model" or was it just planned like this --
+problematic from the start.
+
+* A mix of processes --> WaterScrum
+* Problem --> The entire organization can impossibly change at once
+* Problem --> Additional coordination necessary
+[P29]
+Many groups manage 50+ applications and at any given time, over half are in some
+stage of development or modification.  With complex technical, scheduling, and
+requirements interdependencies, implementing Scrum beyond a limited pilot
+present other challenges.  Today, it is not feasible to stop all systems to
+retool our processes.  We are essentially forced to “change the tires on a
+moving car”.
+We are currently examining various models for dealing with what we call
+“WaterScrum” - simultaneously managing co-dependent waterfall and SCRUM projects
+together 1 .  WaterScrum is like trying to play two sports on the same field at
+the same time.  Extra coordination is necessary to make it work.
+
+* Mixing agile and waterfall --> Waterfall projects needed design aspects up front but agile teams created design incrementally throughout the project
+[P29]
+However, one key issue arose that was not fully anticipated: the waterfall
+project would need to complete its design phase before LPO had completed all of
+their sprints, and because the Scrum project was incremental, some design
+aspects for later sprints would not be complete until after the waterfall
+project design complete milestone.  This mixture caused some temporary grief.
+
+* Old mindset
+* Management in watterfall mindset --> Suspicion when percentage complete is missing
+* Management wanted also fixed delivery dates and cost analysis
+* Development messed up --> Deliverable product not available after each iteration
+[P35]
+Management were used to the waterfall model project plans and lost confidence as
+they could not determine the percentage ofwork completed and what stage gate
+they were at. Management was also used to a delivery date, costs analysis and
+statistics on progress being presented however this was also not available. From
+the development and project planning perspective a deliverable product after
+each iteration was not available. Had one been available it would have restored
+management’s confidence.
+
+* Generic problems --> Management is traditionally oriented
+* Management failed to change --> Phasing out
+* Problem: Big up front oriented management --> Managing to plans
+[P38]
+Sadly not all of the people and roles were able to adapt to the changing
+environment. The most notable was the project management group. The group had
+been very traditionally oriented. They focused on meetings and big up front
+project plans. They managed to the plan, often competing with each other for
+resources. When much of the responsibility shifted to the product owner and team
+to formulate the plan, the project managers struggled to adjust.
+Six months into the transformation it was becoming apparent that the project
+management was continuing to have trouble adding value to the team. A tough
+decision had to be made and the CTO met with the rest of management and they
+decided to phase out the four project manager positions and create a new role on
+the teams called delivery leads.
+
+* Using old language to make agile familiar --> People get confused --> Nothing will change
+* Mapping agile to the old model
+[P44]
+Use the established agile language. Many organizations, in their efforts to
+adopt new techniques, try to soften the blow by adapting the language to
+something more familiar in the culture. This could be a mistake when
+implementing agile. Its language differs from what we know and recognize in
+traditional development, but using the new language forces people to think in
+new ways and helps foster creativity. If you change the agile language to match
+what was previously known, people tend to slip back into their old ways.
+
+* General problems with working with other teams --> Scrum teams considered unrealiable
+[P45]
+What could not be mitigated were issues with teams dependent on “unreliable”
+Scrum teams, at least if the Scrum teams were isolated in a non-Scrum
+environment. Scrum teams working jointly on a particular software product rarely
+had complaints about unreliable delivery because they were familiar with the
+Scrum methodology and figured out how to get their requirements added to other
+teams’ backlogs and with the right priorities.
+
+* Problem: Teams had to fill out forms of two processes --> New process seen as overhead
+[P48]
+The new practices, language, and delivery of increments – these things just
+worked. Besides the good news, though, we also got a share of bad news. First,
+Scrum was perceived to produce lots of overhead and due to the phase
+implementation, sometimes both old and new templates and processes had to be
+used.  Second, many teams thought Scrum was all about managing the tool, and
+information was stored in multiple locations but not matching.
+
+There is a certain amount of truth in the first complaint. The mismatch between
+program management using the existing tools to manage and report on the program
+and the teams using the new agile toolset contributed greatly to the additional
+overhead.
 

### 2014-05-09 journal/article.tex
Using old and new models side by side: Work in progress

--- 2014-05-08 journal/article.tex
@@
 Another example of bad customization that counteracts change is replacing agile
-vocabulary by favoring familiar terminology from the old model [P44].
+vocabulary in favor of familiar terminology from the old model, because
+introducing new language will also create new ways of thinking [P44].
 
@@
 
+% ---------------------------------------------------------------------
+
+In most cases a large organizational transformation proceeded gradually, and
+during the process it was possible that the new agile model was used in parallel
+with the old methods.
+
+Ongoing projects were set up with previous development methods, and the agile
+methods had to be arranged to fit with them [P8, P29].
+
+Using the old and new models side by side was seen as problematic [P26], and
+causing tensions on all organizational levels [P10].
+
+As Scrum teams did not commit to fixed schedules teams working according to the
+original model labeled them unreliable [P45].
+
+The agile teams were required to comply with current procedures producing excess
+documentation and stepping through approval gates [P20, P21]. The bureucracy of
+the previous traditional model was still enforced, and management was not
+convinced to lighten the process. In one case it was described that after
+introducing the agile method teams were required to fill in templates of two
+processes [P48].
+
+Despite the agile introduction development was attempted to be managed in the
+old waterfall fashion. O'Connor [P38] described management as ``focused on
+meetings and big up front project plans''. In another case management was losing
+confidence in agile because familiar reports on cost analysis and percentage
+complete were not produced [P35].
+
+A particular problem in sychronizing work between projects using different
+models was that in Scrum the software design was fleshed out over time as
+sprints progressed, but the waterfall model would require detailed design to be
+locked early in the project [P29].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-WORKING IN SILOS --> Certain expertise is required in many places
-
-* Teams were reorganized without notice --> Planning was hampered when people did not konw who would stay on a team
-* Frequent changes --> People managers had to agree to minimize team changes
-[P11]
-When we launched multiple Scrum teams, we tried to keep existing teams together
-and form new teams based on skills.  As we gained experience, we realized that
-there were skills shortages in certain types of expertise that wasn’t possible
-to develop in a sprint or two. There was pressure to move engineers with certain
-skills to the teams that needed them for work on upcoming sprints or to address
-critical customer requirements.  When moves occurred without much notice, teams
-couldn’t plan effectively if they didn’t know who would stay on a team.  After
-several months of frequent changes, product owners and people managers agreed to
-minimize team changes, and even talked of instituting a “trading deadline.”
 

### 2014-05-09 journal/article.tex
Edited: Using two models side by side

@@
 
-% ---------------------------------------------------------------------
+% Using two models side by side
 
@@
 during the process it was possible that the new agile model was used in parallel
-with the old methods.
-
-Ongoing projects were set up with previous development methods, and the agile
-methods had to be arranged to fit with them [P8, P29].
-
-Using the old and new models side by side was seen as problematic [P26], and
-causing tensions on all organizational levels [P10].
-
-As Scrum teams did not commit to fixed schedules teams working according to the
-original model labeled them unreliable [P45].
-
-The agile teams were required to comply with current procedures producing excess
-documentation and stepping through approval gates [P20, P21]. The bureucracy of
-the previous traditional model was still enforced, and management was not
-convinced to lighten the process. In one case it was described that after
-introducing the agile method teams were required to fill in templates of two
-processes [P48].
-
-Despite the agile introduction development was attempted to be managed in the
-old waterfall fashion. O'Connor [P38] described management as ``focused on
-meetings and big up front project plans''. In another case management was losing
-confidence in agile because familiar reports on cost analysis and percentage
-complete were not produced [P35].
-
+with the old methods. The arrangement of using the old and new models side by
+side was seen as problematic [P26], and causing tensions on all organizational
+levels [P10]. Ongoing projects had been set up with previous development
+methods, and the agile methods had to be arranged to fit with them [P8, P29].
 A particular problem in sychronizing work between projects using different
@@
 
+Despite the agile introduction there were problems with management working
+according to the old waterfall model. O'Connor [P38] described management as
+``focused on meetings and big up front project plans'' despite agile being in
+use. In another case management was losing confidence in agile because familiar
+reports on cost analysis and percentage complete were not produced [P35]. As
+Scrum teams did not commit to fixed schedules they were considered unreliable
+according to the original model [P45].
+
+There were also problems with duplicating management when two models were in
+effect. For instance, agile teams were required to comply with current
+procedures producing excess documentation and stepping through approval gates
+[P20, P21]. The bureucracy of the previous traditional model was still enforced,
+and management was not convinced to lighten the process. Another case describes
+that after introducing the agile method teams were required to fill in templates
+of two processes [P48].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-09 journal/article.tex
Work in progress: Adapting non-development

@@
 with the old methods. The arrangement of using the old and new models side by
-side was seen as problematic [P26], and causing tensions on all organizational
+side was seen as problematic [P1, P26], and causing tensions on all organizational
 levels [P10]. Ongoing projects had been set up with previous development
@@
 
+% Add quote ?
+% * Especially higher organizational levels were not using agile --> Problem
+% [P3]
+% Scrum still exists mostly in a way that avoids conflict with the larger
+% corporate organization, and there has been little progress in addressing larger,
+% organizational impediments.
+
+% [P18]
+% Real Agile teams will experience resistance when they start working with
+% non-Agile interfaces. Lack of understanding in ‘why to help create a cadence?’,
+% ‘why involve the customer more?’ will cause resistance in the change process.
+
+\subsubsection{Integrating other functions than development to agile}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+[P17] Like many organizations, BMC’s decision to adopt Agile practices started
+in the development organization.
+
+"C: Adapting non-development parts of organization to agile"
+
+* General problem: Development had to comply with incompatible surrounding
+* Not agile: Production activities
+[P1]
+Comment: This group is the development group. Most of the elements of this
+group’s TF led to a definition and practice of their development methods very
+close to the agile manifesto. However, they have used this to differentiate
+themselves from the rest of the organisation and have employed  workarounds in
+order to comply with an incompatible sociotechnical network.  Although the
+organization has become more tolerant of agile practices most of the existing
+production activities remain difficult to integrate with Agile.
+
+* General problem: There was not agreement to introduce agile to all levels and functions in the organization
+* Tension was created in the organization
+[P1]
+Three of the groups saw Agile as having value for them, while ‘the business’
+apparently do not see the value of Agile at all. One of the challenges faced by
+this organization is how to extend an agile way of thinking beyond the
+developers, and both the Agile Advocates and Project Z believed that agile
+should cover the entire process, while the development team were content with
+focusing on implementation only. This shows a tension in the process of
+sociotechnical change: trying to translate principles created for the
+development of software into broader knowledge and processes to an audience with
+different roles, understandings and expectations, sharing a contrasting TF. One
+thing (the only thing) which all groups agreed upon was that increased
+collaboration would result in a better product.
+
+* General challenge: moving agile from development to the rest of business
+[P2]
+A group of Agile Advocates and Coaches is driving the agile ‘push’. Their
+mission is to disseminate knowledge of agile methods and facilitate their
+successful adoption across the organisation. From a TF perspective, this means
+that advocates must persuade other staff to adopt the same elements of
+interpretation and practice to frame agile adoption as their own. One of the
+biggest challenges they were facing was to move Agile from the development teams
+into the rest of the business.
+
+* Problem: Collaboration is needed between all stakeholders --> Stakeholders do not share viewpoints which creates conflicts in interests
+* All communities should engage in the process earlier
+[P2]
+Another element of interpretation was the need for greater collaboration and
+communication across all stakeholders and the  problems associated with it,
+especially between marketing managers and IT delivery managers. The IT delivery
+manager characterized the differences by saying: “I (marketing says) want that
+box and I want that now, whereas we (IT delivery) unpack the box”. The technical
+architect agreed that all communities need to engage earlier in the process so
+that decisions make sense from a customer, business and technical point of view.
+
+* Development was content with using agile only internally
+* General problem: Scaling agile practices into use in different roles
+[P2]
+One of the challenges faced by this organisation is how to extend an agile way
+of thinking beyond the developers, and both the Agile Advocates and Project Z
+believed that agile should cover the entire process, while the development team
+were content with focusing on implementation only. This shows a tension in the
+process of sociotechnical change: trying to translate principles created for the
+development of software into broader knowledge and processes to an audience with
+different roles, understandings and expectations, sharing a contrasting TF.
+
+* Not agile: MRDs --> Stories had to be extracted from detailed documentation --> Not good
+* MRDs not based on reality
+[P2]
+Extracting user stories from existing, very detailed ‘marketing requirements
+documents’ (MRDs) was another workaround. These MRDs were a prominent artifact
+found across the organisation. Leaders of Project Z claimed that it was very
+difficult to work with such detailed documents in a project whose scope could
+change rapidly; one agile developer described MRDs as “not based in reality at
+all”.
+
+Comment:
+The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
+
+* Problem: PMO is employing a waterfall paradigm
+[P3]
+the Amazon PMO may some day become interested in Scrum, but they are enmeshed in
+maintaining and thriving within a traditional waterfall paradigm. Another
+example is the existence of separate QA groups. Executive sponsorship would help
+to begin to break down these entrenched impediments.
+
+* Problem: There was a gap between marketing or product management and development --> Requirement elaboration as cause
+[P5]
+We were able to successfully bridge the gap between marketing or product
+management feature and the detailed requirements that are required for Agile.
+
+* Problem: user experience specialists need to keep the big picture within iterations
+[P9]
+the user experience specialists’ main difficulty was how to keep the entire user
+interface design in perspective, and still iterate on specific components.
+Comment: The UX people were considered part of the team.
+
+* Problem: Idea of iterative design was not welcomed by user experience experts, interaction designers and visual designers
+* These groups would have preferred designing up front
+[P6]
+Design is made up of user experience experts, interaction designers and visual
+designers. The designers’ initial reaction to Agile was similar to the way
+engineering architects react when faced with the idea that you don’t design
+everything up front, that you constantly refactor, and that requirements will
+change. They were quite taken aback.
+
+* Problem: Management attempts to use a ``resource pool'' of individuals rather than coordinating teams
+[P6]
+we face the constant threat that restructuring and the instinct from management
+to track and leverage resources at the individual level (for example
+organization-wide resource pools) will destroy the team dynamic and adversely
+affect productivity.
+
+* Challenge: PMO with entrenched processes
+* Challenge: Process group at odds with Scrum teams
+[P6]
+Align with Project Management: In any large organization like Yahoo! there
+inevitably exist Project Management Offices with entrenched process control
+practices. We had conflicts arise in our international offices as they had
+continued to roll out the original PDP during the U.S. transition to Scrum. The
+teams attempting to use Scrum were often at odds with the process group and, as
+a result, the two groups were not communicating well. We tackled this challenge
+by building relationships and identifying the common goals that both groups were
+focused on achieving.
+
+* Problems: Interaction with some groups caused blocking issues
+* User Experience and Design, operations, legal, and security team
+[P10]
+CROSS-FUNCTIONAL DEPENDENCIES: However well we did within the team, we also
+frequently ran into dependencies on other groups often in other locations –
+groups such as UED (User Experience and Design), operations, legal and the
+Yahoo! security team (affectionately called the Paranoids) – which were
+identified as blocking issues but almost never neatly resolved within our sprint
+boundaries.
+
+* DS team members dissatisfied after agile introduction
+[P28]
+A survey sent to the Technology organization in March 2007 revealed that 80\% of
+respondents felt that agile was making their scrum team more effective. However,
+when looking at the data by functional area, only 30\% of DS team members agreed
+with this statement.  When asked about several other impacts of agile, the same
+trend was seen (Table 1).
+
+* The User Experience team members were dissatisfied --> Radical difference in comparison to development
+* Problem: Creating holistic design when agile is done in iterations
+* Problem: Using UX resources across different teams --> UX as a service to teams, instead of part of the team?
+* Solution: Less teams assigned per UX designer
+[P15, P28]
+After the transition, it was clear that User Experience (UX) team members were
+dissatisfied. When asked 6 months into the rollout if agile was making their
+work life better, only 24\% agreed.
+
+While moving to agile was clearly a success for the R&D organization at
+salesforce.com, the initial transition period was far from smooth for the User
+Experience team.  In 2006, the vast majority of literature on the topic of agile
+did not include guidance on the inclusion of User-Centered Design (UCD)
+processes [4].  In agile literature, ‘design’ typically referred only to coding
+or system design [1].  The User Experience team was left with many unanswered
+questions; some included:
+• How can we properly identify the target users and ensure that their needs are
+met within agile’s just-in-time process?
+• How can we accomplish holistic design when agile teams are planning and
+building features in a piecemeal fashion?
+• How do UX professionals succeed across multiple teams when agile espouses one
+team per person?
+The unhappiness of the UX team was illustrated clearly in the results of a
+survey sent out to the R&D organization about 6 months after the rollout. 
+Eighty percent of total respondents felt that agile was making their Scrum team
+more effective. However, when looking at the data by functional area, only 30\%
+of User Experience team members agreed with this statement.  When asked about
+several other impacts of agile, the same trend was seen (Figure 3).
+
+Some of the reasons cited by User Experience for their dissatisfaction with agile 
+included: 
+• Assigned to too many teams 
+• Spending too much time in meetings 
+• Not enough time to complete work 
+• Lack of focus on the big picture
+
+Management made the decision to reduce the number of Scrum teams per User
+Experience team member from four or more to a maximum of two. The process of
+determining which teams would receive assigned resources involved the UX
+managers meeting with the VPs of the different product areas to determine the
+priorities of the new features.
+This solution to de-support approximately 35\% of teams was a situation that no
+one was satisfied with. Certainly, the UX team members were pleased with their
+new, realistic workload, but the decision to strand some Scrum teams was not
+accepted. Ease of use is a core value at salesforce.com. Placing the burden on
+the Scrum teams put that value at risk. In the spirit of agile, UX team members
+brainstormed an alternate solution, and the concept of Office Hours (OH) emerged
+[5]. The UX team members determined that by giving two hours of their time per
+week, they would be able to assist Scrum teams that did not have assigned UX
+resources without spreading themselves too thin.
+
+* Problem: User Experience design was a long process --> Timeboxing to sprints compromised quality of design
+[P15]
+Tighter timelines were also mentioned. In Scrum, teams select what they will
+work on at the beginning of a sprint and then build those items to completion
+over the course of that sprint; in the case of salesforce.com that is a one
+month period of time, start to finish. These aggressive timeframes can cause UCD
+to be compromised, since investigating and iterating designs often takes longer
+than weeks, particularly for complex enterprise applications.  In this initial
+phase, the User Experience team struggled to find ways to achieve the success
+that they had reached using waterfall. The R&D division made significant
+progress using agile, though, so the UX team had no choice but to adapt and
+evolve.
+
+* Problem: UX work in a large application is too large to fit in one sprint
+* Solution: Parallel design and development
+* UX work is phased some sprints ahead of development.
+[P15]
+In waterfall, the designer could focus on the design in its entirety. In agile,
+the UX team was forced to take the design and break it into pieces that could be
+tackled by development in one month segments.  This can be difficult for complex
+enterprise features such as analytics or workflow which are often impossible to
+design and evaluate as component parts.
+Parallel development and design (Figure 3) is beneficial in the enterprise
+applications space because features are typically too big to be designed,
+usability tested, iterated, validated, built, translated and documented in one
+sprint.
+
+* Solution: DS teams had standard timeboxed times to help development teams
+* System Testing and Product Metrics --> Copied good solution from DS
+* These groups did not need a dedicated member in each team
+[P28]
+DS: Development Services --> Documentation, Usability, and User Interface Design
+OH: Office Hours
+Another benefit and success of the OH program has been that the DS team members
+have been able to support teams that traditionally would not have received any
+user experience or product documentation support at all. Teams outside of the
+Technology organization are permitted to book OH slots if the slots are not
+already taken by Technology scrum teams.
+
+* Problem: Development started with agile --> Other departments needed to adapt to iterative delivery
+* Challenge: Product manager --> rethink definition of releases
+* Challenge: Marketing --> Adjust how products are taken to market
+* Challenge: Sales --> Manage customer expectations --> Frequency of new functionality
+* Challenge: How to make commitments to customers --> Requirement priorities
+[P17]
+Like many organizations, BMC’s decision to adopt Agile practices started in the
+development organization. These changes inevitably impact the rest of the
+company as all departments adapt to this new pace of incremental value delivery.
+Product managers must re-think how they define releases and work with
+development. Marketing must adjust how it takes products to market. Sales must
+reset customer expectations on how frequently they have access to new
+functionality.
+As they gain experience with agile, BMC is changing the way they make
+commitments to customers. Discussions with customers are centered more on
+requirement priorities and how those priorities can be delivered earlier and
+with greater satisfaction.
+
+* General challenge: Agile practices work worse in non-development
+[P18]
+Agile practices worked well in development teams and some other teams. Not
+surprisingly, the more distant from development, the less successful the pilot.
+
+“Why change? Agile practices will not suit us.” The further away from actual
+software development, the less understanding about what benefits Agile ways of
+working will bring. The idea that Agile works best if customer contact and
+customer contract are synergistic does not easily penetrate a large
+organization.
+
+* New development moved to agile first 
+* Problem: Moving the product from early access to GA --> Value chain has traditional parts
+[P31]
+The biggest bricks in our Two-Year Wall were being laid outside the development
+organization. The basic principles of Agile and lean are to reduce waste and
+deliver working software to the customer as often as possible. However, the
+development organization does not do this alone. When you look at the end-to-end
+process, there are many other functions involved:  sales, marketing, legal,
+finance. What we didn’t understand when we first embarked upon the Agile journey
+is what a disruptive force our approach would be on the entire company. As I
+mentioned earlier, the first teams to transition to Agile were those working on
+a new set of products. These teams had been working very closely with a
+strategic customer, and in the months prior to the public release of these
+products the situation was ideal for the collaboration, iteration and pace of
+releases. The reality is, in the first 18 months of our transformation the
+original Agile teams were operating as an “Agile microcosm” insulated from the
+rest of our traditional, multi-phased business processes. Once it became time to
+move the product from “early access” to general availability – to move from
+factory to inventory – things started to come apart.  Friction arose as the
+Agile teams began bumping up against the other, traditional, parts of the value
+chain.
+
+* Problem: Teams completed backlog --> Continued with stories direct from customer --> Signe customer bias
+* Problem: Sales and marketing organization had a different vision of the product
+* Problem: Product marketing was not collaborative, and in old mindset
+* Problem: Requirements --> user stories more granular and prioritized
+[P31]
+After several months of working on the new products, the Agile teams “drained”
+the backlog.   For a while, they kept busy working on new user stories that came
+directly from the strategic customer.  This began to lead to a disconnect
+between what the marketing and sales organizations believed to be the overall
+direction of the product, and the actual direction it was taking based on the
+needs of a single customer. It was clearly time for a regroup, and the
+engineering team pushed back on the MRD.  They were in full production mode
+using Agile and they needed more granular, prioritized user stories to fuel
+their efforts.  The teams needed product marketing to be more collaborative in
+the decomposition of the requirements. This was a totally foreign concept for
+the product marketing team. They had approached the relationship with
+engineering as a “black box” – high level requirements went in, and something
+eventually came out (usually late, and sometimes off the mark on what they
+expected).
+
+* Problem: MRD is not dynamic
+* Problem: MRD requirements scoped once a year --> Teams decide for themselves
+* Problem: Product marketing team was very disrupted of the requirement of collaboration!
+[P31]
+Further, the MRD needed to be more dynamic. Because the teams had achieved a
+predictable, efficient pace of delivery, scoping the requirements of a release
+once a year wasn’t enough to feed the engine.  Without the proper fuel the teams
+were left to decide for themselves what new things to work on, driving further
+disconnect between the groups. The concept of continuous collaboration
+throughout the development cycle was a very disruptive concept for the product
+marketing team.
+
+* Challenge: Accomodate time for export clearance
+* Challenge: Acquire third-party licensing authority --> prior to product release
+* Challenge: These processes will remain traditional
+* Challenge: Marketing needs some lead time
+* Enablement activities (sales materials, training)
+* These steps take time, but they are part of the product lifecycle
+* Challenge: Must know beforehand what will be in the product
+[P31]
+While changing our process in the front end of the value chain began to yield
+success, we discovered that there was also a traditional wall at the back end:
+our product release process.
+While the Agile approach enables teams to be flexible with regard to
+prioritization, and very adaptive to changing market conditions and customer
+needs, there are certain “laws of physics” required to release commercial
+product that cannot be altered. For example, the time required to acquire export
+clearance or to acquire third-party licensing authority prior to product
+release.  As a vendor that makes products that must be marketed, sold and
+implemented, there are several traditional processes that are involved in
+releasing our products to market.
+
+First there are the regulatory requirements such as export compliance relative
+to exporting code with certain levels of encryption,  then there are the
+marketing launch plans (demand generation campaigns, collateral, Web site,
+public relations), and finally, the field enablement activities (sales
+materials, training).  All these initiatives take time, and they cannot begin
+without knowledge of exactly what is going to be in the product.
+
+* Problem: Marketing had to know on beforehand what will be in the release
+* Problem: Business went with requiring a fixed delivery date and a fixed feature set
+[P31]
+Serious friction arose. The development teams became frustrated because they
+felt marketing and the NPI process was stalling their progress and throwing up
+barriers against the core principles they had adopted.  Marketing, who needed to
+begin prepping for the launch of the product about three months before the
+actual date, found themselves struggling to message and develop materials and
+campaigns without a clear understanding of the features and functionality that
+would actually be “ready” on the release date. And, the NPI group could not
+apply for export compliance nor could they request any third party license
+authorization without knowing what was actually going to be in the product.
+
+* Solution: Development and product marketing work together
+* Solution: Product marketing can require a release to be completed in 90 days
+[P31] Maples 2009 - Enterprise Agile Transformation - The Two-Year Wall.pdf - 31:18 [What we instituted was what we..]  (5:1983-5:2374)   (Super)
+Codes:  [C: Adapting non-development parts of organization to agile - Family: RQ3: Challenges] 
+No memos
+What we instituted was what we call an “Iron Triangle.” The concept is that at
+some number of sprints out before a release, we would commit to delivering a
+fixed set of features and functionality  at a certain date.  Of course, there
+was always the opportunity for us to deliver more than what we committed to, 
+but we absolutely had to deliver the set features by the set date.
+
+If  product marketing wants to release product for any reason they can “declare”
+a release date, and  we’re able to lock the current feature set and have  the
+code ready to release within two sprints of the date they declare the release
+necessary.
+
+* Development adopted first --> No plans for how to further adopt the rest of the stakeholders
+* Challenges: Product stakeholders and management, marketing, and rest of business
+* Problem: Marketing --> Lead time for marketing material
+* Solution: non-development groups learned how to work more efficiently
+[P38]
+Another challenge was teaching the rest of the business how to interact,
+participate, and collaborate. This is not limited to the product stakeholders
+and management but also HR, facilities, IT, and marketing. Marketing has
+interesting and important requirements. There are lead times for printing
+material such as fliers, brochures, and conference booth banners. Press releases
+need to have design down to best show off the product’s content. When suddenly
+the site that is the product in many of these cases could change literally on a
+week to week basis, you can see some of the issues that had to be addressed.
+In the agile world people often think “we can figure out how to interface with
+the rest of the business once we’ve gotten the team up and running” and while
+there are certainly teams that learn to interface with a non-agile business it
+is a risk and a barrier to change. In our case we were lucky that with the
+support of the CTO we were able to influence many elements of the business to be
+more agile themselves, which allowed them to work more effectively with the new
+agile development group.
+
+The non-development groups have also embraced some agile principles and are not
+only better set to interact with the development group but also are functioning
+at a higher level themselves.
+
+* Challenge: The entire organization must work in the same mode
+* Problem: Marketing
+[P42]
+In order to achieve real flow, the whole organization must work in the same Lean
+mode. For example, managing the product as a whole is difficult if the culture
+of  neighbouring areas such as Human Resources or Marketing is not aligned with
+Lean thinking. Transformation is expanding and the whole organization  is moving
+in the same direction.  A seamless organization is needed to  obtain end-toend
+benefits.
+
+* Challenge: Product teams need to communicate with business management
+* Challenge: Communicating with marketing
+* Challenge: Field services
+* The ways of working had to be changed for these groups
+[P45]
+Upscaling methods were mainly driven by theoretical considerations. Implementing
+them revealed that the skills and knowledge ofproduct managers and software
+architects were often not sufficient to manage a number of Scrum teams
+concurrently. Product teams had to communicate with business management,
+marketing, and field services who were not fully prepared for how Scrum and the
+product teams’ expectations would change their daily work. The process of
+software requirements engineering were found to require a major adjustment.
+
+* Challenge: PMO was a bottleneck inhibiting agility
+[P46]
+To make this journey a success, it was clear that the transformation had to
+start with our strong PMO. Here we had the hub of all things project related –
+coordination, reporting, process quality assurance, compliance, code performance
+quality assurance and the 100+ project managers themselves – in a central place.
+Nothing happened without the PMO’s intimate involvement. It was a bottleneck
+that just could not exist for the organization to be agile. First and foremost,
+all the project managers were returned to their positions in the development
+organization, reporting directly to development managers. There the project
+managers (PMs) became accountable for project success – no longer were they
+simply facilitators and enforcers of the process. While they had always been
+accountable for truth in reporting, now they were accountable for results –
+their bonus was on the line.
+
+* Problem: Technology operations --> Problems keeping up with development
+[P46]
+With the inception of agile development, it became apparent that Technology
+Operations was not able to keep pace with rapid deployment and dedicated
+development teams.  Senior leaders from both development and operations came
+together to openly share their thoughts and provide feedback. As an outcome of
+those meetings, the Infrastructure Solutions Management team was formed within
+the Technology Operations group.  The goal of the team is to provide
+infrastructure support for the rapid deployment of applications from inception
+to production. Teams are aligned to each development group, with the purpose of
+helping the highest priority projects within the group to ensure that the
+Technology Operations teams are not being an impediment to development. In
+addition, this group will partner with other Technology Operations groups to
+streamline processes and enhance the current service catalog or request for
+services.
+
+* Challenge: Surrounding organization, internal customers --> manufacturing, customer services, marketing, sales, field and partner enablement, training, demos, documentation, finance
+* Problem: They were not familiar with incremental planning --> Before they committed once
+[P48]
+We have found, though, that the most difficult part of our transition is getting
+everything else that must feed and support the product development teams in
+place. In the past, supporting and surrounding organizations counted on having a
+content commitment once. From this, plans were formed and timelines aligned.
+Incremental planning and commitment was something new for the critical internal
+customer teams  as well (manufacturing, customer services, marketing, sales,
+field and partner enablement, training, demos, documentation, finance, and
+senior leaders.)
+
+* Problem: Infrastructure and Operations Teams --> Need to adopt to rapid delivery
+[P50]
+With the bulk of Engineering forming into Scrum teams and becoming highly
+productive, the Infrastructure and Operations Teams remained operating as usual
+but continued to fall further behind as there was not enough time to focus on
+improving process efficiency.  A significant portion of activity was spent on
+short-term low-gain fire-fighting which  only reduced the resources required to
+maintain and significantly improve the infrastructure.
 

### 2014-05-10 journal/article.tex
Work in progress

@@
 % Using two models side by side
+% Part of the organization is in old mindset
 
@@
 
+% --> ``Part is in the old mindset''
+% * Problem: Management attempts to use a ``resource pool'' of individuals rather than coordinating teams
+% [P6]
+% we face the constant threat that restructuring and the instinct from management
+% to track and leverage resources at the individual level (for example
+% organization-wide resource pools) will destroy the team dynamic and adversely
+% affect productivity.
+
 \subsubsection{Integrating other functions than development to agile}
 
+Agile methods are typically understood through a set of activities closely
+related to software development. Cases describe how development teams were
+typically the first entities in the organization that were considered when
+introducing agile [P2, P17, P31, P38]. However, development operates in a larger
+context, and will need to interface with other organizational functions.
+Challenges were faced when exposing other parts of the organization to the agile
+mode, and functions that were further away from development were resistant to
+change [P2, P18, P31]. A tension was growing between development and broader
+roles in the organization [P1]. The full beenfits of the transformation would
+not be attained unless the entire organization was set to work along the same
+paradigm [P18, P42].
+
+Primary sources listed various functions beyond development that are involved in
+bringing products to the market. Marketing was a particularly frequently
+mentioned group [P5, P17, P28, P31, P38, P45, P48]. Other groups involved in the
+whole of product delivery were sales [P17, P31], infrastructure and operations
+[P10, P45, P46, P50], user experience and design [P9, P10, P28], documentation
+[P28], legal [P10], security [P10], customer services [P48], and finance [P48].
+All of these groups were reported to be affected on some level by the agile
+transformation.
+% separate QA [P3], --> Later human resources --> Later
+
+Agile affects timings in the software delivery lifecycle which caused two types
+of problems. Firstly, the iterativeness of agile models was problematic when
+interfacing with other functions. Iterativeness changed the pace of delivery
+[P17], and forced design to focus on a shorter scope [P15]. Various groups
+feeding and supporting development had earlier committed once to a long term
+plan, but in the new arrangement they needed to adjust to incremental delivery
+[P48].
+
+Federoff and Courage [P15] describe how the user experience team struggled with
+the
+
+--> Iterativeness and UED
+
+--> DS + UED [P6, P9, P10, P15+P28]
+
+The second type of problems was that some release activities would inevitably
+have a long lead time, such as acquiring export clearances and licensing
+authority. A particular function that required extended periods of stability was
+sales and marketing, as they were required to make customer commitments and
+communicate upcomming releases a relatively long time ahead. [P31]
+[P38] --> See marketing lead time?
+
+The increased delivery speed of development teams caused infrastructure and
+operations teams to fall behind, and they were forced to update their way of
+working [P46, P50].
+
+Some cases reported that development efforts were controlled on a top level by a
+project management office (PMO). The PMO was described as entrenched in a
+waterfall paradigm, and the top-level rigidity was causing friction in the
+agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
+all aspects of projects, and its central structure had to be dismantled for
+the organization to become agile [P46].
+
+Problems in adopting agile were even showing in functions outside the software
+delivery lifecycle. A support function that was particularly mentioned to
+have issues in agile adoption was human resouces (HR).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-[P17] Like many organizations, BMC’s decision to adopt Agile practices started
-in the development organization.
-
-"C: Adapting non-development parts of organization to agile"
-
-* General problem: Development had to comply with incompatible surrounding
-* Not agile: Production activities
-[P1]
-Comment: This group is the development group. Most of the elements of this
-group’s TF led to a definition and practice of their development methods very
-close to the agile manifesto. However, they have used this to differentiate
-themselves from the rest of the organisation and have employed  workarounds in
-order to comply with an incompatible sociotechnical network.  Although the
-organization has become more tolerant of agile practices most of the existing
-production activities remain difficult to integrate with Agile.
-
-* General problem: There was not agreement to introduce agile to all levels and functions in the organization
-* Tension was created in the organization
-[P1]
-Three of the groups saw Agile as having value for them, while ‘the business’
-apparently do not see the value of Agile at all. One of the challenges faced by
-this organization is how to extend an agile way of thinking beyond the
-developers, and both the Agile Advocates and Project Z believed that agile
-should cover the entire process, while the development team were content with
-focusing on implementation only. This shows a tension in the process of
-sociotechnical change: trying to translate principles created for the
-development of software into broader knowledge and processes to an audience with
-different roles, understandings and expectations, sharing a contrasting TF. One
-thing (the only thing) which all groups agreed upon was that increased
-collaboration would result in a better product.
-
-* General challenge: moving agile from development to the rest of business
-[P2]
-A group of Agile Advocates and Coaches is driving the agile ‘push’. Their
-mission is to disseminate knowledge of agile methods and facilitate their
-successful adoption across the organisation. From a TF perspective, this means
-that advocates must persuade other staff to adopt the same elements of
-interpretation and practice to frame agile adoption as their own. One of the
-biggest challenges they were facing was to move Agile from the development teams
-into the rest of the business.
-
-* Problem: Collaboration is needed between all stakeholders --> Stakeholders do not share viewpoints which creates conflicts in interests
-* All communities should engage in the process earlier
-[P2]
-Another element of interpretation was the need for greater collaboration and
-communication across all stakeholders and the  problems associated with it,
-especially between marketing managers and IT delivery managers. The IT delivery
-manager characterized the differences by saying: “I (marketing says) want that
-box and I want that now, whereas we (IT delivery) unpack the box”. The technical
-architect agreed that all communities need to engage earlier in the process so
-that decisions make sense from a customer, business and technical point of view.
-
-* Development was content with using agile only internally
-* General problem: Scaling agile practices into use in different roles
-[P2]
-One of the challenges faced by this organisation is how to extend an agile way
-of thinking beyond the developers, and both the Agile Advocates and Project Z
-believed that agile should cover the entire process, while the development team
-were content with focusing on implementation only. This shows a tension in the
-process of sociotechnical change: trying to translate principles created for the
-development of software into broader knowledge and processes to an audience with
-different roles, understandings and expectations, sharing a contrasting TF.
-
-* Not agile: MRDs --> Stories had to be extracted from detailed documentation --> Not good
-* MRDs not based on reality
-[P2]
-Extracting user stories from existing, very detailed ‘marketing requirements
-documents’ (MRDs) was another workaround. These MRDs were a prominent artifact
-found across the organisation. Leaders of Project Z claimed that it was very
-difficult to work with such detailed documents in a project whose scope could
-change rapidly; one agile developer described MRDs as “not based in reality at
-all”.
-
-Comment:
-The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
-
-* Problem: PMO is employing a waterfall paradigm
-[P3]
-the Amazon PMO may some day become interested in Scrum, but they are enmeshed in
-maintaining and thriving within a traditional waterfall paradigm. Another
-example is the existence of separate QA groups. Executive sponsorship would help
-to begin to break down these entrenched impediments.
-
-* Problem: There was a gap between marketing or product management and development --> Requirement elaboration as cause
-[P5]
-We were able to successfully bridge the gap between marketing or product
-management feature and the detailed requirements that are required for Agile.
-
-* Problem: user experience specialists need to keep the big picture within iterations
-[P9]
-the user experience specialists’ main difficulty was how to keep the entire user
-interface design in perspective, and still iterate on specific components.
-Comment: The UX people were considered part of the team.
+UED
 
@@
 
-* Problem: Management attempts to use a ``resource pool'' of individuals rather than coordinating teams
-[P6]
-we face the constant threat that restructuring and the instinct from management
-to track and leverage resources at the individual level (for example
-organization-wide resource pools) will destroy the team dynamic and adversely
-affect productivity.
-
-* Challenge: PMO with entrenched processes
-* Challenge: Process group at odds with Scrum teams
-[P6]
-Align with Project Management: In any large organization like Yahoo! there
-inevitably exist Project Management Offices with entrenched process control
-practices. We had conflicts arise in our international offices as they had
-continued to roll out the original PDP during the U.S. transition to Scrum. The
-teams attempting to use Scrum were often at odds with the process group and, as
-a result, the two groups were not communicating well. We tackled this challenge
-by building relationships and identifying the common goals that both groups were
-focused on achieving.
+* Problem: user experience specialists need to keep the big picture within iterations
+[P9]
+the user experience specialists’ main difficulty was how to keep the entire user
+interface design in perspective, and still iterate on specific components.
+Comment: The UX people were considered part of the team.
 
@@
 
-* DS team members dissatisfied after agile introduction
-[P28]
-A survey sent to the Technology organization in March 2007 revealed that 80\% of
-respondents felt that agile was making their scrum team more effective. However,
-when looking at the data by functional area, only 30\% of DS team members agreed
-with this statement.  When asked about several other impacts of agile, the same
-trend was seen (Table 1).
-
-* The User Experience team members were dissatisfied --> Radical difference in comparison to development
-* Problem: Creating holistic design when agile is done in iterations
-* Problem: Using UX resources across different teams --> UX as a service to teams, instead of part of the team?
-* Solution: Less teams assigned per UX designer
-[P15, P28]
-After the transition, it was clear that User Experience (UX) team members were
-dissatisfied. When asked 6 months into the rollout if agile was making their
-work life better, only 24\% agreed.
-
-While moving to agile was clearly a success for the R&D organization at
-salesforce.com, the initial transition period was far from smooth for the User
-Experience team.  In 2006, the vast majority of literature on the topic of agile
-did not include guidance on the inclusion of User-Centered Design (UCD)
-processes [4].  In agile literature, ‘design’ typically referred only to coding
-or system design [1].  The User Experience team was left with many unanswered
-questions; some included:
-• How can we properly identify the target users and ensure that their needs are
-met within agile’s just-in-time process?
-• How can we accomplish holistic design when agile teams are planning and
-building features in a piecemeal fashion?
-• How do UX professionals succeed across multiple teams when agile espouses one
-team per person?
-The unhappiness of the UX team was illustrated clearly in the results of a
-survey sent out to the R&D organization about 6 months after the rollout. 
-Eighty percent of total respondents felt that agile was making their Scrum team
-more effective. However, when looking at the data by functional area, only 30\%
-of User Experience team members agreed with this statement.  When asked about
-several other impacts of agile, the same trend was seen (Figure 3).
-
-Some of the reasons cited by User Experience for their dissatisfaction with agile 
-included: 
-• Assigned to too many teams 
-• Spending too much time in meetings 
-• Not enough time to complete work 
-• Lack of focus on the big picture
-
-Management made the decision to reduce the number of Scrum teams per User
-Experience team member from four or more to a maximum of two. The process of
-determining which teams would receive assigned resources involved the UX
-managers meeting with the VPs of the different product areas to determine the
-priorities of the new features.
-This solution to de-support approximately 35\% of teams was a situation that no
-one was satisfied with. Certainly, the UX team members were pleased with their
-new, realistic workload, but the decision to strand some Scrum teams was not
-accepted. Ease of use is a core value at salesforce.com. Placing the burden on
-the Scrum teams put that value at risk. In the spirit of agile, UX team members
-brainstormed an alternate solution, and the concept of Office Hours (OH) emerged
-[5]. The UX team members determined that by giving two hours of their time per
-week, they would be able to assist Scrum teams that did not have assigned UX
-resources without spreading themselves too thin.
-
-* Problem: User Experience design was a long process --> Timeboxing to sprints compromised quality of design
-[P15]
-Tighter timelines were also mentioned. In Scrum, teams select what they will
-work on at the beginning of a sprint and then build those items to completion
-over the course of that sprint; in the case of salesforce.com that is a one
-month period of time, start to finish. These aggressive timeframes can cause UCD
-to be compromised, since investigating and iterating designs often takes longer
-than weeks, particularly for complex enterprise applications.  In this initial
-phase, the User Experience team struggled to find ways to achieve the success
-that they had reached using waterfall. The R&D division made significant
-progress using agile, though, so the UX team had no choice but to adapt and
-evolve.
-
-* Problem: UX work in a large application is too large to fit in one sprint
-* Solution: Parallel design and development
-* UX work is phased some sprints ahead of development.
-[P15]
-In waterfall, the designer could focus on the design in its entirety. In agile,
-the UX team was forced to take the design and break it into pieces that could be
-tackled by development in one month segments.  This can be difficult for complex
-enterprise features such as analytics or workflow which are often impossible to
-design and evaluate as component parts.
-Parallel development and design (Figure 3) is beneficial in the enterprise
-applications space because features are typically too big to be designed,
-usability tested, iterated, validated, built, translated and documented in one
-sprint.
-
-* Solution: DS teams had standard timeboxed times to help development teams
-* System Testing and Product Metrics --> Copied good solution from DS
-* These groups did not need a dedicated member in each team
-[P28]
-DS: Development Services --> Documentation, Usability, and User Interface Design
-OH: Office Hours
-Another benefit and success of the OH program has been that the DS team members
-have been able to support teams that traditionally would not have received any
-user experience or product documentation support at all. Teams outside of the
-Technology organization are permitted to book OH slots if the slots are not
-already taken by Technology scrum teams.
+"C: Adapting non-development parts of organization to agile"
+
+SOME KIND OF GENERAL ADJUSTMENT?
+
+* Challenge: Product teams need to communicate with business management
+* Challenge: Communicating with marketing
+* Challenge: Field services
+* The ways of working had to be changed for these groups
+[P45]
+Upscaling methods were mainly driven by theoretical considerations. Implementing
+them revealed that the skills and knowledge of product managers and software
+architects were often not sufficient to manage a number of Scrum teams
+concurrently. Product teams had to communicate with business management,
+marketing, and field services who were not fully prepared for how Scrum and the
+product teams’ expectations would change their daily work. The process of
+software requirements engineering were found to require a major adjustment.
+
+ITERATIVENESS / SPEEDED UP INCREMENTAL DELIVERY
 
@@
 
-* General challenge: Agile practices work worse in non-development
-[P18]
-Agile practices worked well in development teams and some other teams. Not
-surprisingly, the more distant from development, the less successful the pilot.
-
-“Why change? Agile practices will not suit us.” The further away from actual
-software development, the less understanding about what benefits Agile ways of
-working will bring. The idea that Agile works best if customer contact and
-customer contract are synergistic does not easily penetrate a large
-organization.
-
-* New development moved to agile first 
-* Problem: Moving the product from early access to GA --> Value chain has traditional parts
-[P31]
-The biggest bricks in our Two-Year Wall were being laid outside the development
-organization. The basic principles of Agile and lean are to reduce waste and
-deliver working software to the customer as often as possible. However, the
-development organization does not do this alone. When you look at the end-to-end
-process, there are many other functions involved:  sales, marketing, legal,
-finance. What we didn’t understand when we first embarked upon the Agile journey
-is what a disruptive force our approach would be on the entire company. As I
-mentioned earlier, the first teams to transition to Agile were those working on
-a new set of products. These teams had been working very closely with a
-strategic customer, and in the months prior to the public release of these
-products the situation was ideal for the collaboration, iteration and pace of
-releases. The reality is, in the first 18 months of our transformation the
-original Agile teams were operating as an “Agile microcosm” insulated from the
-rest of our traditional, multi-phased business processes. Once it became time to
-move the product from “early access” to general availability – to move from
-factory to inventory – things started to come apart.  Friction arose as the
-Agile teams began bumping up against the other, traditional, parts of the value
-chain.
-
-* Problem: Teams completed backlog --> Continued with stories direct from customer --> Signe customer bias
-* Problem: Sales and marketing organization had a different vision of the product
-* Problem: Product marketing was not collaborative, and in old mindset
-* Problem: Requirements --> user stories more granular and prioritized
-[P31]
-After several months of working on the new products, the Agile teams “drained”
-the backlog.   For a while, they kept busy working on new user stories that came
-directly from the strategic customer.  This began to lead to a disconnect
-between what the marketing and sales organizations believed to be the overall
-direction of the product, and the actual direction it was taking based on the
-needs of a single customer. It was clearly time for a regroup, and the
-engineering team pushed back on the MRD.  They were in full production mode
-using Agile and they needed more granular, prioritized user stories to fuel
-their efforts.  The teams needed product marketing to be more collaborative in
-the decomposition of the requirements. This was a totally foreign concept for
-the product marketing team. They had approached the relationship with
-engineering as a “black box” – high level requirements went in, and something
-eventually came out (usually late, and sometimes off the mark on what they
-expected).
-
-* Problem: MRD is not dynamic
-* Problem: MRD requirements scoped once a year --> Teams decide for themselves
-* Problem: Product marketing team was very disrupted of the requirement of collaboration!
-[P31]
-Further, the MRD needed to be more dynamic. Because the teams had achieved a
-predictable, efficient pace of delivery, scoping the requirements of a release
-once a year wasn’t enough to feed the engine.  Without the proper fuel the teams
-were left to decide for themselves what new things to work on, driving further
-disconnect between the groups. The concept of continuous collaboration
-throughout the development cycle was a very disruptive concept for the product
-marketing team.
-
-* Challenge: Accomodate time for export clearance
-* Challenge: Acquire third-party licensing authority --> prior to product release
-* Challenge: These processes will remain traditional
-* Challenge: Marketing needs some lead time
-* Enablement activities (sales materials, training)
-* These steps take time, but they are part of the product lifecycle
-* Challenge: Must know beforehand what will be in the product
-[P31]
-While changing our process in the front end of the value chain began to yield
-success, we discovered that there was also a traditional wall at the back end:
-our product release process.
-While the Agile approach enables teams to be flexible with regard to
-prioritization, and very adaptive to changing market conditions and customer
-needs, there are certain “laws of physics” required to release commercial
-product that cannot be altered. For example, the time required to acquire export
-clearance or to acquire third-party licensing authority prior to product
-release.  As a vendor that makes products that must be marketed, sold and
-implemented, there are several traditional processes that are involved in
-releasing our products to market.
-
-First there are the regulatory requirements such as export compliance relative
-to exporting code with certain levels of encryption,  then there are the
-marketing launch plans (demand generation campaigns, collateral, Web site,
-public relations), and finally, the field enablement activities (sales
-materials, training).  All these initiatives take time, and they cannot begin
-without knowledge of exactly what is going to be in the product.
-
-* Problem: Marketing had to know on beforehand what will be in the release
-* Problem: Business went with requiring a fixed delivery date and a fixed feature set
-[P31]
-Serious friction arose. The development teams became frustrated because they
-felt marketing and the NPI process was stalling their progress and throwing up
-barriers against the core principles they had adopted.  Marketing, who needed to
-begin prepping for the launch of the product about three months before the
-actual date, found themselves struggling to message and develop materials and
-campaigns without a clear understanding of the features and functionality that
-would actually be “ready” on the release date. And, the NPI group could not
-apply for export compliance nor could they request any third party license
-authorization without knowing what was actually going to be in the product.
-
-* Solution: Development and product marketing work together
-* Solution: Product marketing can require a release to be completed in 90 days
-[P31] Maples 2009 - Enterprise Agile Transformation - The Two-Year Wall.pdf - 31:18 [What we instituted was what we..]  (5:1983-5:2374)   (Super)
-Codes:  [C: Adapting non-development parts of organization to agile - Family: RQ3: Challenges] 
-No memos
-What we instituted was what we call an “Iron Triangle.” The concept is that at
-some number of sprints out before a release, we would commit to delivering a
-fixed set of features and functionality  at a certain date.  Of course, there
-was always the opportunity for us to deliver more than what we committed to, 
-but we absolutely had to deliver the set features by the set date.
-
-If  product marketing wants to release product for any reason they can “declare”
-a release date, and  we’re able to lock the current feature set and have  the
-code ready to release within two sprints of the date they declare the release
-necessary.
-
+* Challenge: Surrounding organization adaption
+* Internal customers --> manufacturing, customer services, marketing, sales, field and partner enablement, training, demos, documentation, finance
+* Problem: They were not familiar with incremental planning --> Originally they committed once
+[P48]
+We have found, though, that the most difficult part of our transition is getting
+everything else that must feed and support the product development teams in
+place. In the past, supporting and surrounding organizations counted on having a
+content commitment once. From this, plans were formed and timelines aligned.
+Incremental planning and commitment was something new for the critical internal
+customer teams  as well (manufacturing, customer services, marketing, sales,
+field and partner enablement, training, demos, documentation, finance, and
+senior leaders.)
+
+LEAD TIME CONSIDERATION
+
+* Marketing causes trouble
 * Development adopted first --> No plans for how to further adopt the rest of the stakeholders
@@
 
-* Challenge: The entire organization must work in the same mode
-* Problem: Marketing
-[P42]
-In order to achieve real flow, the whole organization must work in the same Lean
-mode. For example, managing the product as a whole is difficult if the culture
-of  neighbouring areas such as Human Resources or Marketing is not aligned with
-Lean thinking. Transformation is expanding and the whole organization  is moving
-in the same direction.  A seamless organization is needed to  obtain end-toend
-benefits.
-
-* Challenge: Product teams need to communicate with business management
-* Challenge: Communicating with marketing
-* Challenge: Field services
-* The ways of working had to be changed for these groups
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+DEVELOPMENT SUPPORT TEAMS were not satisfied with agile
+* UX lacked big picture --> Iterations
+* Problem: UX work in a large application is too large to fit in one sprint
+
+* DS team members dissatisfied after agile introduction
+[P28]
+A survey sent to the Technology organization in March 2007 revealed that 80\% of
+respondents felt that agile was making their scrum team more effective. However,
+when looking at the data by functional area, only 30\% of DS team members agreed
+with this statement.  When asked about several other impacts of agile, the same
+trend was seen (Table 1).
+
+* The User Experience team members were dissatisfied --> Radical difference in comparison to development
+* Problem: Creating holistic design when agile is done in iterations
+* Problem: Using UX resources across different teams --> UX as a service to teams, instead of part of the team?
+* Solution: Less teams assigned per UX designer
+[P15, P28]
+After the transition, it was clear that User Experience (UX) team members were
+dissatisfied. When asked 6 months into the rollout if agile was making their
+work life better, only 24\% agreed.
+
+While moving to agile was clearly a success for the R\&D organization at
+salesforce.com, the initial transition period was far from smooth for the User
+Experience team.  In 2006, the vast majority of literature on the topic of agile
+did not include guidance on the inclusion of User-Centered Design (UCD)
+processes [4].  In agile literature, ‘design’ typically referred only to coding
+or system design [1].  The User Experience team was left with many unanswered
+questions; some included:
+• How can we properly identify the target users and ensure that their needs are
+met within agile’s just-in-time process?
+• How can we accomplish holistic design when agile teams are planning and
+building features in a piecemeal fashion?
+• How do UX professionals succeed across multiple teams when agile espouses one
+team per person?
+The unhappiness of the UX team was illustrated clearly in the results of a
+survey sent out to the R\&D organization about 6 months after the rollout. 
+Eighty percent of total respondents felt that agile was making their Scrum team
+more effective. However, when looking at the data by functional area, only 30\%
+of User Experience team members agreed with this statement.  When asked about
+several other impacts of agile, the same trend was seen (Figure 3).
+
+Some of the reasons cited by User Experience for their dissatisfaction with agile 
+included: 
+• Assigned to too many teams 
+• Spending too much time in meetings 
+• Not enough time to complete work 
+• Lack of focus on the big picture
+
+Management made the decision to reduce the number of Scrum teams per User
+Experience team member from four or more to a maximum of two. The process of
+determining which teams would receive assigned resources involved the UX
+managers meeting with the VPs of the different product areas to determine the
+priorities of the new features.
+This solution to de-support approximately 35\% of teams was a situation that no
+one was satisfied with. Certainly, the UX team members were pleased with their
+new, realistic workload, but the decision to strand some Scrum teams was not
+accepted. Ease of use is a core value at salesforce.com. Placing the burden on
+the Scrum teams put that value at risk. In the spirit of agile, UX team members
+brainstormed an alternate solution, and the concept of Office Hours (OH) emerged
+[5]. The UX team members determined that by giving two hours of their time per
+week, they would be able to assist Scrum teams that did not have assigned UX
+resources without spreading themselves too thin.
+
+* Problem: User Experience design was a long process --> Timeboxing to sprints compromised quality of design
+[P15]
+Tighter timelines were also mentioned. In Scrum, teams select what they will
+work on at the beginning of a sprint and then build those items to completion
+over the course of that sprint; in the case of salesforce.com that is a one
+month period of time, start to finish. These aggressive timeframes can cause UCD
+to be compromised, since investigating and iterating designs often takes longer
+than weeks, particularly for complex enterprise applications.  In this initial
+phase, the User Experience team struggled to find ways to achieve the success
+that they had reached using waterfall. The R\&D division made significant
+progress using agile, though, so the UX team had no choice but to adapt and
+evolve.
+
+* Problem: UX work in a large application is too large to fit in one sprint
+* Solution: Parallel design and development
+* UX work is phased some sprints ahead of development.
+[P15]
+In waterfall, the designer could focus on the design in its entirety. In agile,
+the UX team was forced to take the design and break it into pieces that could be
+tackled by development in one month segments.  This can be difficult for complex
+enterprise features such as analytics or workflow which are often impossible to
+design and evaluate as component parts.
+Parallel development and design (Figure 3) is beneficial in the enterprise
+applications space because features are typically too big to be designed,
+usability tested, iterated, validated, built, translated and documented in one
+sprint.
+
+* Solution: DS teams had standard timeboxed times to help development teams
+* System Testing and Product Metrics --> Copied good solution from DS
+* These groups did not need a dedicated member in each team
+[P28]
+DS: Development Services --> Documentation, Usability, and User Interface Design
+OH: Office Hours
+Another benefit and success of the OH program has been that the DS team members
+have been able to support teams that traditionally would not have received any
+user experience or product documentation support at all. Teams outside of the
+Technology organization are permitted to book OH slots if the slots are not
+already taken by Technology scrum teams.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+MARKETING NEEDS TO KNOW ON BEFOREHAND WHAT WILL BE RELEASED
+
+* Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
+* Problem: Sales and marketing organization had a different vision of the product
+* Problem: Product marketing was not collaborative, and in old mindset
+* Problem: Requirements --> user stories more granular and prioritized
+[P31]
+After several months of working on the new products, the Agile teams “drained”
+the backlog.   For a while, they kept busy working on new user stories that came
+directly from the strategic customer.  This began to lead to a disconnect
+between what the marketing and sales organizations believed to be the overall
+direction of the product, and the actual direction it was taking based on the
+needs of a single customer. It was clearly time for a regroup, and the
+engineering team pushed back on the MRD.  They were in full production mode
+using Agile and they needed more granular, prioritized user stories to fuel
+their efforts.  The teams needed product marketing to be more collaborative in
+the decomposition of the requirements. This was a totally foreign concept for
+the product marketing team. They had approached the relationship with
+engineering as a “black box” – high level requirements went in, and something
+eventually came out (usually late, and sometimes off the mark on what they
+expected).
+
+* Challenge: Accomodate time for export clearance
+* Challenge: Acquire third-party licensing authority --> prior to product release
+* Challenge: These processes will remain traditional
+* Challenge: Marketing needs some lead time
+* Enablement activities (sales materials, training)
+* These steps take time, but they are part of the product lifecycle
+* Challenge: Must know beforehand what will be in the product
+[P31]
+While changing our process in the front end of the value chain began to yield
+success, we discovered that there was also a traditional wall at the back end:
+our product release process.
+While the Agile approach enables teams to be flexible with regard to
+prioritization, and very adaptive to changing market conditions and customer
+needs, there are certain “laws of physics” required to release commercial
+product that cannot be altered. For example, the time required to acquire export
+clearance or to acquire third-party licensing authority prior to product
+release.  As a vendor that makes products that must be marketed, sold and
+implemented, there are several traditional processes that are involved in
+releasing our products to market.
+
+First there are the regulatory requirements such as export compliance relative
+to exporting code with certain levels of encryption,  then there are the
+marketing launch plans (demand generation campaigns, collateral, Web site,
+public relations), and finally, the field enablement activities (sales
+materials, training).  All these initiatives take time, and they cannot begin
+without knowledge of exactly what is going to be in the product.
+
+* Problem: Marketing had to know on beforehand what will be in the release
+* Problem: Business went with requiring a fixed delivery date and a fixed feature set
+[P31]
+Serious friction arose. The development teams became frustrated because they
+felt marketing and the NPI process was stalling their progress and throwing up
+barriers against the core principles they had adopted.  Marketing, who needed to
+begin prepping for the launch of the product about three months before the
+actual date, found themselves struggling to message and develop materials and
+campaigns without a clear understanding of the features and functionality that
+would actually be “ready” on the release date. And, the NPI group could not
+apply for export compliance nor could they request any third party license
+authorization without knowing what was actually going to be in the product.
+
+* Solution: Development and product marketing work together
+* Solution: Product marketing can require a release to be completed in 90 days
+[P31]
+What we instituted was what we call an “Iron Triangle.” The concept is that at
+some number of sprints out before a release, we would commit to delivering a
+fixed set of features and functionality  at a certain date.  Of course, there
+was always the opportunity for us to deliver more than what we committed to, 
+but we absolutely had to deliver the set features by the set date.
+
+If  product marketing wants to release product for any reason they can “declare”
+a release date, and  we’re able to lock the current feature set and have  the
+code ready to release within two sprints of the date they declare the release
+necessary.
+
+% TOP LAYER OF ORGANIZATION IS IN WATERFALL MODE
+% --> Generic ``PMO''
+% 
+% * Problem: PMO is employing a waterfall paradigm
+% [P3]
+% the Amazon PMO may some day become interested in Scrum, but they are enmeshed in
+% maintaining and thriving within a traditional waterfall paradigm. Another
+% example is the existence of separate QA groups. Executive sponsorship would help
+% to begin to break down these entrenched impediments.
+% 
+% 
+% * Problem: There was a gap between marketing or product management and development --> Requirement elaboration as cause
+% [P5]
+% We were able to successfully bridge the gap between marketing or product
+% management feature and the detailed requirements that are required for Agile.
+% 
+% 
+% * Challenge: PMO with entrenched processes
+% * Challenge: Process group at odds with Scrum teams
+% [P6]
+% Align with Project Management: In any large organization like Yahoo! there
+% inevitably exist Project Management Offices with entrenched process control
+% practices. We had conflicts arise in our international offices as they had
+% continued to roll out the original PDP during the U.S. transition to Scrum. The
+% teams attempting to use Scrum were often at odds with the process group and, as
+% a result, the two groups were not communicating well. We tackled this challenge
+% by building relationships and identifying the common goals that both groups were
+% focused on achieving.
+% 
+% 
+% * Challenge: PMO was a bottleneck inhibiting agility
+% [P46]
+% To make this journey a success, it was clear that the transformation had to
+% start with our strong PMO. Here we had the hub of all things project related –
+% coordination, reporting, process quality assurance, compliance, code performance
+% quality assurance and the 100+ project managers themselves – in a central place.
+% Nothing happened without the PMO’s intimate involvement. It was a bottleneck
+% that just could not exist for the organization to be agile. First and foremost,
+% all the project managers were returned to their positions in the development
+% organization, reporting directly to development managers. There the project
+% managers (PMs) became accountable for project success – no longer were they
+% simply facilitators and enforcers of the process. While they had always been
+% accountable for truth in reporting, now they were accountable for results –
+% their bonus was on the line.
+% 
+% * Problem: Development started with agile --> Other departments needed to adapt to iterative delivery
+% [P17]
+% Changes inevitably impact the rest of the company as all departments adapt to
+% this new pace of incremental value delivery. Product managers must re-think how
+% they define releases and work with development.
+
+%--
+
+% [P17] Like many organizations, BMC’s decision to adopt Agile practices started
+% in the development organization.
+
+% * Development was content with using agile only internally
+% * General problem: Scaling agile practices into use in different roles
+% [P2]
+% One of the challenges faced by this organisation is how to extend an agile way
+% of thinking beyond the developers, and both the Agile Advocates and Project Z
+% believed that agile should cover the entire process
+
+% * General challenge: Agile practices work worse in non-development
+% [P18]
+% Agile practices worked well in development teams and some other teams. Not
+% surprisingly, the more distant from development, the less successful the pilot.
+% 
+% “Why change? Agile practices will not suit us.” The further away from actual
+% software development, the less understanding about what benefits Agile ways of
+% working will bring. The idea that Agile works best if customer contact and
+% customer contract are synergistic does not easily penetrate a large
+% organization.
+
+% * Development of new products moved to agile first 
+% * Problem: Moving the product from early access to GA --> Value chain has traditional parts
+% [P31]
+% The biggest bricks in our Two-Year Wall were being laid outside the development
+% organization. The basic principles of Agile and lean are to reduce waste and
+% deliver working software to the customer as often as possible. However, the
+% development organization does not do this alone. When you look at the end-to-end
+% process, there are many other functions involved:  sales, marketing, legal,
+% finance. What we didn’t understand when we first embarked upon the Agile journey
+% is what a disruptive force our approach would be on the entire company. As I
+% mentioned earlier, the first teams to transition to Agile were those working on
+% a new set of products. These teams had been working very closely with a
+% strategic customer, and in the months prior to the public release of these
+% products the situation was ideal for the collaboration, iteration and pace of
+% releases. The reality is, in the first 18 months of our transformation the
+% original Agile teams were operating as an “Agile microcosm” insulated from the
+% rest of our traditional, multi-phased business processes. Once it became time to
+% move the product from “early access” to general availability – to move from
+% factory to inventory – things started to come apart.  Friction arose as the
+% Agile teams began bumping up against the other, traditional, parts of the value
+% chain.
+% 
+% * Development adopted first --> No plans for how to further adopt the rest of the stakeholders
+% [P38]
+% When suddenly the site that is the product in many of these cases could change
+% literally on a week to week basis, you can see some of the issues that had to be
+% addressed. In the agile world people often think “we can figure out how to
+% interface with the rest of the business once we’ve gotten the team up and
+% running” and while there are certainly teams that learn to interface with a
+% non-agile business it is a risk and a barrier to change. In our case we were
+% lucky that with the support of the CTO we were able to influence many elements
+% of the business to be more agile themselves, which allowed them to work more
+% effectively with the new agile development group.
+
+% * Challenge: The entire organization must work in the same mode
+% [P42]
+
+% ---------------------------------------------------------------------
+
+REQUIREMENT AND MRD COMMUNICATION PROBLEMS
+
+* Not agile: MRDs --> Stories had to be extracted from detailed documentation --> Not good
+* MRDs not based on reality
+[P2]
+Extracting user stories from existing, very detailed ‘marketing requirements
+documents’ (MRDs) was another workaround. These MRDs were a prominent artifact
+found across the organisation. Leaders of Project Z claimed that it was very
+difficult to work with such detailed documents in a project whose scope could
+change rapidly; one agile developer described MRDs as “not based in reality at
+all”.
+Comment:
+The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
+
+* Problem: MRD is not dynamic
+* Problem: MRD requirements scoped once a year --> Teams decide for themselves
+* Problem: Product marketing team was very disrupted of the requirement of collaboration!
+[P31]
+Further, the MRD needed to be more dynamic. Because the teams had achieved a
+predictable, efficient pace of delivery, scoping the requirements of a release
+once a year wasn’t enough to feed the engine.  Without the proper fuel the teams
+were left to decide for themselves what new things to work on, driving further
+disconnect between the groups. The concept of continuous collaboration
+throughout the development cycle was a very disruptive concept for the product
+marketing team.
+
+* Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
+* Problem: Sales and marketing organization had a different vision of the product
+* Problem: Product marketing was not collaborative, and in old mindset
+* Problem: Requirements --> user stories more granular and prioritized
+[P31]
+
+* Requirements engineering must be done differently
 [P45]
-Upscaling methods were mainly driven by theoretical considerations. Implementing
-them revealed that the skills and knowledge ofproduct managers and software
-architects were often not sufficient to manage a number of Scrum teams
-concurrently. Product teams had to communicate with business management,
-marketing, and field services who were not fully prepared for how Scrum and the
-product teams’ expectations would change their daily work. The process of
-software requirements engineering were found to require a major adjustment.
-
-* Challenge: PMO was a bottleneck inhibiting agility
-[P46]
-To make this journey a success, it was clear that the transformation had to
-start with our strong PMO. Here we had the hub of all things project related –
-coordination, reporting, process quality assurance, compliance, code performance
-quality assurance and the 100+ project managers themselves – in a central place.
-Nothing happened without the PMO’s intimate involvement. It was a bottleneck
-that just could not exist for the organization to be agile. First and foremost,
-all the project managers were returned to their positions in the development
-organization, reporting directly to development managers. There the project
-managers (PMs) became accountable for project success – no longer were they
-simply facilitators and enforcers of the process. While they had always been
-accountable for truth in reporting, now they were accountable for results –
-their bonus was on the line.
-
-* Problem: Technology operations --> Problems keeping up with development
-[P46]
-With the inception of agile development, it became apparent that Technology
-Operations was not able to keep pace with rapid deployment and dedicated
-development teams.  Senior leaders from both development and operations came
-together to openly share their thoughts and provide feedback. As an outcome of
-those meetings, the Infrastructure Solutions Management team was formed within
-the Technology Operations group.  The goal of the team is to provide
-infrastructure support for the rapid deployment of applications from inception
-to production. Teams are aligned to each development group, with the purpose of
-helping the highest priority projects within the group to ensure that the
-Technology Operations teams are not being an impediment to development. In
-addition, this group will partner with other Technology Operations groups to
-streamline processes and enhance the current service catalog or request for
-services.
-
-* Challenge: Surrounding organization, internal customers --> manufacturing, customer services, marketing, sales, field and partner enablement, training, demos, documentation, finance
-* Problem: They were not familiar with incremental planning --> Before they committed once
-[P48]
-We have found, though, that the most difficult part of our transition is getting
-everything else that must feed and support the product development teams in
-place. In the past, supporting and surrounding organizations counted on having a
-content commitment once. From this, plans were formed and timelines aligned.
-Incremental planning and commitment was something new for the critical internal
-customer teams  as well (manufacturing, customer services, marketing, sales,
-field and partner enablement, training, demos, documentation, finance, and
-senior leaders.)
-
-* Problem: Infrastructure and Operations Teams --> Need to adopt to rapid delivery
-[P50]
-With the bulk of Engineering forming into Scrum teams and becoming highly
-productive, the Infrastructure and Operations Teams remained operating as usual
-but continued to fall further behind as there was not enough time to focus on
-improving process efficiency.  A significant portion of activity was spent on
-short-term low-gain fire-fighting which  only reduced the resources required to
-maintain and significantly improve the infrastructure.
-
-% ---------------------------------------------------------------------
+Product teams had to communicate with business management, marketing, and field
+services who were not fully prepared for how Scrum and the product teams’
+expectations would change their daily work. The process of software requirements
+engineering were found to require a major adjustment.
 

### 2014-05-11 journal/article.tex
Work in progress

--- 2014-05-09 journal/article.tex
@@
 
+% Add quote ?
+% * Especially higher organizational levels were not using agile --> Problem
+% [P3]
+% Scrum still exists mostly in a way that avoids conflict with the larger
+% corporate organization, and there has been little progress in addressing larger,
+% organizational impediments.
+
+% [P18]
+% Real Agile teams will experience resistance when they start working with
+% non-Agile interfaces. Lack of understanding in ‘why to help create a cadence?’,
+% ‘why involve the customer more?’ will cause resistance in the change process.
+
+% --> ``Part is in the old mindset''
+% * Problem: Management attempts to use a ``resource pool'' of individuals rather than coordinating teams
+% [P6]
+% we face the constant threat that restructuring and the instinct from management
+% to track and leverage resources at the individual level (for example
+% organization-wide resource pools) will destroy the team dynamic and adversely
+% affect productivity.
+
+\subsubsection{Integrating other functions than development to agile}
+
+Agile methods are typically understood through a set of activities closely
+related to software development. Cases describe how development teams were
+typically the first entities in the organization that were considered when
+introducing agile [P2, P17, P31, P38]. However, development operates in a larger
+context, and will need to interface with other organizational functions.
+Challenges were faced when exposing other parts of the organization to the agile
+mode, and functions that were further away from development were resistant to
+change [P2, P18, P31]. A tension was growing between development and broader
+roles in the organization [P1]. The full beenfits of the transformation would
+not be attained unless the entire organization was set to work along the same
+paradigm [P18, P42].
+
+Primary sources listed various functions beyond development that are involved in
+bringing products to the market. Marketing was a particularly frequently
+mentioned group [P5, P17, P28, P31, P38, P45, P48]. Other groups involved in the
+whole of product delivery were sales [P17, P31], infrastructure and operations
+[P10, P45, P46, P50], user experience and design [P9, P10, P28], documentation
+[P28], legal [P10], security [P10], customer services [P48], and finance [P48].
+All of these groups were reported to be affected on some level by the agile
+transformation.
+% separate QA [P3], --> Later
+% human resources --> Later
+
+Agile affects timings in the software delivery lifecycle which caused two types
+of problems. Firstly, the iterativeness of agile models was problematic when
+interfacing with other functions. Iterativeness changed the pace of delivery
+[P17], and forced design to focus on a shorter scope [P15]. Various groups
+feeding and supporting development had earlier committed once to a long term
+plan, but in the new arrangement they needed to adjust to incremental delivery
+[P48].
+Especially user experience (UX) and interaction designers did not welcome an
+incremental model [P6], and were struggling with maintaining the big picture in
+design while working in iterations [P9, P10].
+The infrastructure and operations teams were falling behind due to the increased
+delivery speed of development teams, and they were forced to update their way of
+working [P46, P50].
+
+Federoff and Courage [P15] elaborate on the user experience team's problems
+with the short time horizons of Scrum. Previously the UX team had supplied
+development with designs within a long release cycle, but Scrum required
+completing features within weeks. The problem was that that the UX work required
+a holistic view instead of an incremental one, and the design process did not
+fit well into the time frame of a sprint. The UX team was overloaded and
+discontent, but the situation was fixed by refining the schedule for interations
+with development teams. [P15, P28]
+
+The second type of problems was that some release activities would inevitably
+have a long lead time but required committing to a set of functionalities in
+good time. The agile models' emphasis on short time horizon and flexibility in
+prioritization did not mix well with such activities.
+For instance, it would take time to prepare printed marketing material and press
+releases, but they needed to accurately present the upcoming features of the
+product [P38].
+
+Maples [P31] describes that marketing needed three months for preparing the
+product launch, but the agile process allowed development to keep changing the
+content of the product. As a result marketing was struggling with creating
+material and campaigns without knowing the exact product features. Further, the
+processes of acquiring export clearances and licensing authority required that
+the the feature set of the product was known well on beforehand [P31].
+As a result of these requests development felt that their process was being
+slowed down. The issues were solved by agreeing on that marketing could at any
+time request a release in 90 days, and development would be able to commit to a
+fixed feature set to be delivered [P31].
+
+% --
+
+Some cases reported that development efforts were controlled on a top level by a
+project management office (PMO). The PMO was described as entrenched in a
+waterfall paradigm, and the top-level rigidity was causing friction in the
+agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
+all aspects of projects, and its central structure had to be dismantled for
+the organization to become agile [P46].
+
+Problems in adopting agile were even showing in functions outside the software
+delivery lifecycle. A support function that was particularly mentioned to
+have issues in agile adoption was human resouces (HR).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+"C: Adapting non-development parts of organization to agile"
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% MARKETING NEEDS TO KNOW ON BEFOREHAND WHAT WILL BE RELEASED
+% 
+% 
+% * Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
+% * Problem: Sales and marketing organization had a different vision of the product
+% * Problem: Product marketing was not collaborative, and in old mindset
+% * Problem: Requirements --> user stories more granular and prioritized
+% [P31]
+% After several months of working on the new products, the Agile teams “drained”
+% the backlog.   For a while, they kept busy working on new user stories that came
+% directly from the strategic customer.  This began to lead to a disconnect
+% between what the marketing and sales organizations believed to be the overall
+% direction of the product, and the actual direction it was taking based on the
+% needs of a single customer. It was clearly time for a regroup, and the
+% engineering team pushed back on the MRD.  They were in full production mode
+% using Agile and they needed more granular, prioritized user stories to fuel
+% their efforts.  The teams needed product marketing to be more collaborative in
+% the decomposition of the requirements. This was a totally foreign concept for
+% the product marketing team. They had approached the relationship with
+% engineering as a “black box” – high level requirements went in, and something
+% eventually came out (usually late, and sometimes off the mark on what they
+% expected).
+% 
+% 
+% * Challenge: Accomodate time for export clearance
+% * Challenge: Acquire third-party licensing authority --> prior to product release
+% * Challenge: These processes will remain traditional
+% * Challenge: Marketing needs some lead time
+% * Enablement activities (sales materials, training)
+% * These steps take time, but they are part of the product lifecycle
+% * Challenge: Must know beforehand what will be in the product
+% [P31]
+% While changing our process in the front end of the value chain began to yield
+% success, we discovered that there was also a traditional wall at the back end:
+% our product release process.
+% While the Agile approach enables teams to be flexible with regard to
+% prioritization, and very adaptive to changing market conditions and customer
+% needs, there are certain “laws of physics” required to release commercial
+% product that cannot be altered. For example, the time required to acquire export
+% clearance or to acquire third-party licensing authority prior to product
+% release.  As a vendor that makes products that must be marketed, sold and
+% implemented, there are several traditional processes that are involved in
+% releasing our products to market.
+% 
+% First there are the regulatory requirements such as export compliance relative
+% to exporting code with certain levels of encryption,  then there are the
+% marketing launch plans (demand generation campaigns, collateral, Web site,
+% public relations), and finally, the field enablement activities (sales
+% materials, training).  All these initiatives take time, and they cannot begin
+% without knowledge of exactly what is going to be in the product.
+% 
+% 
+% * Problem: Marketing had to know on beforehand what will be in the release
+% * Problem: Business went with requiring a fixed delivery date and a fixed feature set
+% [P31]
+% Serious friction arose. The development teams became frustrated because they
+% felt marketing and the NPI process was stalling their progress and throwing up
+% barriers against the core principles they had adopted.  Marketing, who needed to
+% begin prepping for the launch of the product about three months before the
+% actual date, found themselves struggling to message and develop materials and
+% campaigns without a clear understanding of the features and functionality that
+% would actually be “ready” on the release date. And, the NPI group could not
+% apply for export compliance nor could they request any third party license
+% authorization without knowing what was actually going to be in the product.
+% 
+% 
+% * Solution: Development and product marketing work together
+% * Solution: Product marketing can require a release to be completed in 90 days
+% [P31]
+% What we instituted was what we call an “Iron Triangle.” The concept is that at
+% some number of sprints out before a release, we would commit to delivering a
+% fixed set of features and functionality  at a certain date.  Of course, there
+% was always the opportunity for us to deliver more than what we committed to, 
+% but we absolutely had to deliver the set features by the set date.
+% 
+% If  product marketing wants to release product for any reason they can “declare”
+% a release date, and  we’re able to lock the current feature set and have  the
+% code ready to release within two sprints of the date they declare the release
+% necessary.
+
 % ---------------------------------------------------------------------
+
+REQUIREMENT AND MRD COMMUNICATION PROBLEMS
+
+* Not agile: MRDs --> Stories had to be extracted from detailed documentation --> Not good
+* MRDs not based on reality
+[P2]
+Extracting user stories from existing, very detailed ‘marketing requirements
+documents’ (MRDs) was another workaround. These MRDs were a prominent artifact
+found across the organisation. Leaders of Project Z claimed that it was very
+difficult to work with such detailed documents in a project whose scope could
+change rapidly; one agile developer described MRDs as “not based in reality at
+all”.
+Comment:
+The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
+
+* Problem: MRD is not dynamic
+* Problem: MRD requirements scoped once a year --> Teams decide for themselves
+* Problem: Product marketing team was very disrupted of the requirement of collaboration!
+[P31]
+Further, the MRD needed to be more dynamic. Because the teams had achieved a
+predictable, efficient pace of delivery, scoping the requirements of a release
+once a year wasn’t enough to feed the engine.  Without the proper fuel the teams
+were left to decide for themselves what new things to work on, driving further
+disconnect between the groups. The concept of continuous collaboration
+throughout the development cycle was a very disruptive concept for the product
+marketing team.
+
+* Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
+* Problem: Sales and marketing organization had a different vision of the product
+* Problem: Product marketing was not collaborative, and in old mindset
+* Problem: Requirements --> user stories more granular and prioritized
+[P31]
+
+* Requirements engineering must be done differently
+[P45]
+Product teams had to communicate with business management, marketing, and field
+services who were not fully prepared for how Scrum and the product teams’
+expectations would change their daily work. The process of software requirements
+engineering were found to require a major adjustment.
 

### 2014-05-11 journal/article.tex
Work in progress

@@
 
-"C: Adapting non-development parts of organization to agile"
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MARKETING NEEDS TO KNOW ON BEFOREHAND WHAT WILL BE RELEASED
-% 
-% 
-% * Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
-% * Problem: Sales and marketing organization had a different vision of the product
-% * Problem: Product marketing was not collaborative, and in old mindset
-% * Problem: Requirements --> user stories more granular and prioritized
-% [P31]
-% After several months of working on the new products, the Agile teams “drained”
-% the backlog.   For a while, they kept busy working on new user stories that came
-% directly from the strategic customer.  This began to lead to a disconnect
-% between what the marketing and sales organizations believed to be the overall
-% direction of the product, and the actual direction it was taking based on the
-% needs of a single customer. It was clearly time for a regroup, and the
-% engineering team pushed back on the MRD.  They were in full production mode
-% using Agile and they needed more granular, prioritized user stories to fuel
-% their efforts.  The teams needed product marketing to be more collaborative in
-% the decomposition of the requirements. This was a totally foreign concept for
-% the product marketing team. They had approached the relationship with
-% engineering as a “black box” – high level requirements went in, and something
-% eventually came out (usually late, and sometimes off the mark on what they
-% expected).
-% 
-% 
-% * Challenge: Accomodate time for export clearance
-% * Challenge: Acquire third-party licensing authority --> prior to product release
-% * Challenge: These processes will remain traditional
-% * Challenge: Marketing needs some lead time
-% * Enablement activities (sales materials, training)
-% * These steps take time, but they are part of the product lifecycle
-% * Challenge: Must know beforehand what will be in the product
-% [P31]
-% While changing our process in the front end of the value chain began to yield
-% success, we discovered that there was also a traditional wall at the back end:
-% our product release process.
-% While the Agile approach enables teams to be flexible with regard to
-% prioritization, and very adaptive to changing market conditions and customer
-% needs, there are certain “laws of physics” required to release commercial
-% product that cannot be altered. For example, the time required to acquire export
-% clearance or to acquire third-party licensing authority prior to product
-% release.  As a vendor that makes products that must be marketed, sold and
-% implemented, there are several traditional processes that are involved in
-% releasing our products to market.
-% 
-% First there are the regulatory requirements such as export compliance relative
-% to exporting code with certain levels of encryption,  then there are the
-% marketing launch plans (demand generation campaigns, collateral, Web site,
-% public relations), and finally, the field enablement activities (sales
-% materials, training).  All these initiatives take time, and they cannot begin
-% without knowledge of exactly what is going to be in the product.
-% 
-% 
-% * Problem: Marketing had to know on beforehand what will be in the release
-% * Problem: Business went with requiring a fixed delivery date and a fixed feature set
-% [P31]
-% Serious friction arose. The development teams became frustrated because they
-% felt marketing and the NPI process was stalling their progress and throwing up
-% barriers against the core principles they had adopted.  Marketing, who needed to
-% begin prepping for the launch of the product about three months before the
-% actual date, found themselves struggling to message and develop materials and
-% campaigns without a clear understanding of the features and functionality that
-% would actually be “ready” on the release date. And, the NPI group could not
-% apply for export compliance nor could they request any third party license
-% authorization without knowing what was actually going to be in the product.
-% 
-% 
-% * Solution: Development and product marketing work together
-% * Solution: Product marketing can require a release to be completed in 90 days
-% [P31]
-% What we instituted was what we call an “Iron Triangle.” The concept is that at
-% some number of sprints out before a release, we would commit to delivering a
-% fixed set of features and functionality  at a certain date.  Of course, there
-% was always the opportunity for us to deliver more than what we committed to, 
-% but we absolutely had to deliver the set features by the set date.
-% 
-% If  product marketing wants to release product for any reason they can “declare”
-% a release date, and  we’re able to lock the current feature set and have  the
-% code ready to release within two sprints of the date they declare the release
-% necessary.
+"C: Difficulties with requirements engineering or backlog"
+______________________________________________________________________
+
+P 1: Abdelnour-Nocera 2007 - Adopting Agile in a large organization - balancing the old with the new.pdf - 1:13 [Extracting user stories from e..]  (10:1790-10:2192)   (Super)
+Codes:	[C: Adapting non-development parts of organization to agile - Family: RQ3: Challenges] [C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Extracting user stories from existing, very detailed ‘marketing requirements 
+documents’ (MRDs) was another workaround. These MRDs were a prominent artifact 
+found across the organization. Leaders of Project Z claimed that it was very difficult 
+to work with such detailed documents in a project whose scope could change rapidly; 
+one Agile developer described MRDs as “not based in reality at all”.
+
+P 2: Abdelnour-Nocera 2008 - Adopting Agile In A Large Organisation.pdf - 2:20 [Extracting user stories from e..]  (9:389-9:791)   (Super)
+Codes:	[C: Adapting non-development parts of organization to agile - Family: RQ3: Challenges] [C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Extracting user stories from existing, very detailed ‘marketing requirements 
+documents’ (MRDs) was another workaround. These MRDs were a prominent artifact 
+found across the organisation. Leaders of Project Z claimed that it was very difficult 
+to work with such detailed documents in a project whose scope could change rapidly; 
+one agile developer described MRDs as “not based in reality at all”.
+
+Comment:
+The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
+
+P 5: Beavers 2007 - Managing A Large 'Agile' Software Engineering Organization.pdf - 5:32 [There was constant debate over..]  (5:2150-5:2761)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+There was 
+constant debate over both the meaning of the 
+requirement and whether or not the requirement was 
+met.  Quality Assurance engineers were frustrated 
+trying to determine whether or not we should accept 
+this feature.  Simply put, we had not invested enough 
+of our velocity in either up front or ongoing 
+requirements elaboration.
+Going into the second release, we knew we had to 
+do a better job with managing, elaborating, and 
+prioritizing requirements.  This had strictly been an 
+issue regarding where we chose to invest our energy.  
+Properly managing requirements was critical.
+
+P11: Cowan 2011 - When The VP Is A Scrum Master, You Hit The Ground Running.pdf - 11:31 [Writing and grooming user stor..]  (4:974-4:2298)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Writing and 
+grooming user stories seems like a simple process during 
+Scrum training.  When many teams start sprinting at once, no 
+one starts off as an expert at writing stories.  In the mad rush to 
+convert requirements to stories, many stories had the correct 
+format, but were often too big, too ambiguous, or too 
+prescriptive.  Product owners and teams who held grooming 
+sessions once or twice a week over time refined their backlogs 
+so that planning meetings were short, and there was no wasted 
+time between sprints trying to figure out what to do next.  For 
+one untended product backlog, we instituted “speed grooming” 
+where we spent a maximum of five minutes discussing a story 
+to categorize it as “ready for a team,” “obsolete,” “more 
+grooming,” etc., so the stories could be readied for upcoming 
+sprints.  Most of our stories were for system software, not end 
+user features, so teams consistently underestimated how long a 
+story would take, and would complain that it couldn’t be 
+broken down further.  All of sudden they were being required 
+to break down work that used to take months into tasks of a 
+few hours.  In hindsight, we could have reduced stress on 
+everyone if we had devoted more time upfront to developing 
+everyone’s skills in writing stories and decomposing epics.
+
+P17: Gat 2006 - How BMC is Scaling Agile Development.pdf - 17:18 [As BMC went through their init..]  (4:45-4:850)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+As BMC went through their initial iterations, they
+realized there was often a gap between the level of
+detail in their system level features and the detail
+needed for iteration planning and execution. The use of
+simple “throw away stories” was insufficient to their
+challenge of scale. To bridge this gap, BMC created
+the role of Requirements Architect as part of each
+development team.
+The responsibility of each Requirements Architect
+is to take the high-level features defined by product
+management and decompose these on a just-in-time
+basis into the more detailed requirements and stories
+needed to drive iteration planning and development.
+This new role increased the accuracy of development
+estimates and improved the collaboration between the
+development team and product management.
+
+P17: Gat 2006 - How BMC is Scaling Agile Development.pdf - 17:29 [One of the challenges BMC face..]  (6:138-6:450)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+One of the challenges BMC faces is defining the right
+level of requirement detail at the right time in the life
+cycle. BMC is now using Rally Software to manage a
+prioritized feature backlog and to elaborate the
+requirements and stories for each component that will
+likely be needed to implement the feature.
+
+P24: Hansen 2009 - From Cmmi And Isolation To Scrum, Agile, Lean And Collaboration.pdf - 24:13 [Initially we started with the ..]  (3:3312-3:3922)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] [S: Pay attention to the product owner role - Family: RQ4: Success factors] 
+No memos
+
+Initially we started with the Product Owner (PO) 
+as a role for sales and account management. It worked 
+reasonably well with local teams in Denmark when 
+they could collaborate and meet when required. But it 
+did not work well with teams in Bangladesh. We 
+moved some of the very experienced senior developers 
+and consultants into global PO roles to establish a more 
+structured flow of tasks and collaboration with the 
+teams.  
+To establish a structure to improve the flow of 
+work between the Danish and the Bangladeshi teams 
+we introduced a “Proxy PO” for each team in 
+Bangladesh (figure 3).
+
+P24: Hansen 2009 - From Cmmi And Isolation To Scrum, Agile, Lean And Collaboration.pdf - 24:14 [Understanding the problem doma..]  (3:4124-4:62)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Understanding the problem domain and how the 
+different requirements were related was still hard. We 
+used many different techniques to move the domain 
+285knowledge and user understanding out to the different 
+teams.
+
+P24: Hansen 2009 - From Cmmi And Isolation To Scrum, Agile, Lean And Collaboration.pdf - 24:20 [Global planning Having one tea..]  (4:2827-4:3579)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Global planning 
+Having one team and therefore one sprint backlog 
+meant, that we had to do sprint planning together. Not 
+an easy task, but after some experiments, we came up 
+with a solution that worked for us: 
+• The whole global team was presented with the user 
+stories online 
+• The global team went offline and started to break 
+down the stories and meet to synchronize sprint 
+backlog items into one sprint backlog 
+• The global team informed the PO of the result 
+We spent a lot of time trying to make it work well, 
+and it required a high level of trust between all parties.  
+It also required a lot of knowledge of the business 
+domain, but made it easy to ask for the relevant 
+knowledge when doing the breakdown and planning.
+
+P24: Hansen 2009 - From Cmmi And Isolation To Scrum, Agile, Lean And Collaboration.pdf - 24:22 [The R&D team could build the b..]  (5:3659-5:4291)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+The R&D team could build the best quality code, 
+but we kept building elements of the wrong things. It 
+was almost like building a top of the line V12 engine, 
+for a solar cell car race. The trust between the team and 
+the PO were at stake. “I told you to build an engine, 
+and you should have known it was an electric engine I 
+needed”. So once again we had to improve the quality 
+and understanding of requirements between the PO and 
+the global team both before and during the actual 
+development in the sprint. After retrospectives and 
+more discussions, we decided to try moving the PO 
+into the Danish R&D team room.
+
+P27: Lee 2008 - Forming To Performing - Transitioning Large-Scale Project Into Agile.pdf - 27:19 [The team had difficulty estima..]  (3:567-3:858)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+The team had difficulty
+estimating tasks during the planning session: What
+should go into a story? How should a task card be
+written? Is the story points system better than the
+hour’s system? We also had a lack of well-defined
+acceptance criteria before and during the planning
+session.
+
+P27: Lee 2008 - Forming To Performing - Transitioning Large-Scale Project Into Agile.pdf - 27:21 [One solution we came up with w..]  (3:2624-3:3190)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+One solution we came up with was to improve the
+relationship between us and our internal customers.
+Planning sessions were dreaded - long and not well
+structured. Our team was driven based on the internal
+customer requirements, which went into a backlog
+that the program manager also known as Susha
+[3]
+maintained. The more efficient way to understand
+what a task entailed was to involve the customers in
+our planning meeting and pair up with them on
+writing the actual story. The customers partnered up
+with us for 15 minutes during the planning sessions.
+
+P30: Long 2008 - Agile Supports Improved Culture And Quality For Healthwise.pdf - 30:11 [The introduction of Agile rais..]  (3:35-3:514)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+The introduction of Agile raised visibility of the 
+requirements churn and lack of clarity in feature 
+definition, but writing good user stories remained 
+difficult, due in part to the separation between the 
+client and product development teams. 
+Engineering became known as the “story-eating 
+engine.” Stories were being completed, but did 
+not always add value for clients.  
+Agile successfully exposed the next areas needing 
+attention: defining value-added stories.
+
+P30: Long 2008 - Agile Supports Improved Culture And Quality For Healthwise.pdf - 30:38 [Managing an effective enterpri..]  (6:1506-6:1546)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Managing an effective enterprise backlog.
+
+P31: Maples 2009 - Enterprise Agile Transformation - The Two-Year Wall.pdf - 31:5 [we discovered some areas that ..]  (2:1586-2:1895)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] [C: Problems with separate non-agile QA - Family: RQ3: Challenges] 
+No memos
+
+we discovered some areas that 
+would require us to adapt our chosen Scrum 
+methodology to account for the realities of complex, 
+large-scale development.  Two key activities fell 
+outside the scope of our original Scrum framework 1. 
+Informing how things get into the backlog, and 2. 
+Enterprise testing.
+
+P31: Maples 2009 - Enterprise Agile Transformation - The Two-Year Wall.pdf - 31:6 [Once the organization began to..]  (2:2602-2:3650)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Once the organization began to understand the 
+increased productivity of the Agile teams, and they saw 
+that every three weeks the teams selected what they 
+would work on, they began to grasp the possibility that 
+any feature request could be delivered in mere weeks.  
+And, suddenly, we found the backlog was “up for 
+grabs” at the beginning of each sprint. 
+Requests came flooding in from the field, from 
+marketing, from our support organization. They would 
+be automatically added to the backlog, and each sprint 
+became an opportunity to evaluate and reprioritize. 
+And, since the Agile approach increases collaboration, 
+everyone had the opportunity to “lobby” for his or her 
+requests to make the backlog. This constant influx of 
+demand from several sources began to put a strain on 
+the system.  Questions about how to prioritize the 
+backlog items and a subversive process of “backlog 
+bartering” began. This negatively impacted the teams 
+and their performance, and it also generated distrust 
+among the various groups.
+
+P31: Maples 2009 - Enterprise Agile Transformation - The Two-Year Wall.pdf - 31:11 [After several months of workin..]  (4:1-4:1041)   (Super)
+Codes:	[C: Adapting non-development parts of organization to agile - Family: RQ3: Challenges] [C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+After several months of working on the new 
+products, the Agile teams “drained” the backlog.   For 
+a while, they kept busy working on new user stories 
+that came directly from the strategic customer.  This 
+began to lead to a disconnect between what the 
+marketing and sales organizations believed to be the 
+overall direction of the product, and the actual direction 
+it was taking based on the needs of a single customer.   
+It was clearly time for a regroup, and the 
+engineering team pushed back on the MRD.  They 
+were in full production mode using Agile and they 
+needed more granular, prioritized user stories to fuel 
+their efforts.  The teams needed product marketing to 
+be more collaborative in the decomposition of the 
+requirements. This was a totally foreign concept for the 
+product marketing team.  They had approached the 
+relationship with engineering as a “black box” – highlevel requirements went in, and something eventually 
+came out (usually late, and sometimes off the mark on 
+what they expected).
+
+P33: Mencke 2008 - A Product Manager'S Guide To Surviving The Big Bang Approach To Agile Transitions.pdf - 33:7 [In addition to the hottest ide..]  (3:634-3:1384)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+In addition to the hottest ideas coming off 
+the IdeaExchange, salesforce.com’s product 
+managers also have backlogs of important 
+features, and these product backlogs are an 
+integral part of ADM.
+The initial creation and prioritization of 
+product backlogs was an ominous challenge, and 
+nearly impossible to achieve in the two-month 
+timeframe the  product managers were allotted. 
+Breaking down requirements into both epic and 
+detailed user stories was arduous, and the way in 
+which the requirements were broken down was
+inconsistent between teams because each team 
+had limited insight into how other teams were 
+organizing their product backlogs. 
+To address this issue, salesforce.com built 
+an in-house app called  Scrumforce
+
+P33: Mencke 2008 - A Product Manager'S Guide To Surviving The Big Bang Approach To Agile Transitions.pdf - 33:10 [The transition from writing fu..]  (4:786-4:979)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+The transition from writing full 
+product or functional specs to writing user 
+stories is challenging. Learn how to break up 
+larger features into smaller user stories, and 
+prioritize them.
+
+P33: Mencke 2008 - A Product Manager'S Guide To Surviving The Big Bang Approach To Agile Transitions.pdf - 33:11 [With user stories come smaller..]  (4:1058-4:1661)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] [S: Requirements engineering is implemented correctly - Family: RQ4: Success factors] 
+No memos
+
+With user stories come smaller and more 
+vague requirements, and because of the agile 
+nature of the development process, the 
+product owner has a tendency to change 
+requirements during a sprint. Often, this is 
+due to the product owner not specifying the 
+full requirements before the sprint starts. 
+This can waste  time and result in work that 
+is thrown away. User stories should be 
+complete, have acceptance criteria, and be 
+usability tested before being added to a 
+sprint. Having clear guidelines prepared for 
+when the team is ready to tackle a user story 
+will help reduce waste.
+
+P37: O'Connor 2010 - Letters from the Edge of an Agile Transition.pdf - 37:22 [Our business analysts were run..]  (4:952-4:1713)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Our business analysts were running into
+the same issue one level up. They were not used to providing
+feature requirements in story form, much less breaking down
+these ”feature” level stories into stories that could be reasonably estimated by the development team. I realized almost
+immediately the stories were too big, but you can’t force immediate understanding. Like any skill, estimating takes practice. We went through many agonizing estimation and story
+breakdown sessions before we started to get some traction.
+As we began to build up some historical data this closed the
+feedback loop. The team was able to see that we were frequently underestimating and that this was directly connected
+to the trouble we were still having with story breakdown.
+
+P37: O'Connor 2010 - Letters from the Edge of an Agile Transition.pdf - 37:27 [We started ?nishing cards, but..]  (4:4347-4:4838)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+We started ﬁnishing cards,
+but found that estimates of hours often took weeks. Stories
+would spawn additional stories in order to reach completion. For example, a story that started with a two day estimate would generate a week’s worth of stories (sometimes
+even more than a week!). Many agile developers would say
+”Great! That’s part of agile.” Gaining clarity on a story and
+understanding that there is more to it is part of the process. Unfortunately this was happening far too often.
+
+P39: Prokhorenko 2012 - Skiing And Boxing - Coaching Product And Enterprise Teams.pdf - 39:19 [there were a number of difficu..]  (3:3675-3:4041)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+there were a number of difficulties in 
+requirements management. Actual stakeholders weren’t 
+available for the teams due to  many reasons and they used 
+another consulting company in the UK to gather their 
+requirements, design UI and communicate them back to teams. 
+Even more requirements were provided from risk management 
+subject matter experts in London.
+
+Comment:
+This sounds a bit like the initial state.
+
+P39: Prokhorenko 2012 - Skiing And Boxing - Coaching Product And Enterprise Teams.pdf - 39:31 [Product discovery workshops ar..]  (6:2221-6:2778)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Product discovery workshops are an extremely efficient 
+technique allowing for quick elaboration of a product roadmap 
+and creating mid-term requirements details. However, as most 
+efficient tools do, it has certain limitations. One of them is the 
+need to have the PO and the team co-located for a sprint which 
+might be a budget issue for a distributed project. Another one is 
+the nature of a PO role in enterprise projects where you simply 
+can’t bring all the subject matter experts in one place with the 
+team for any reasonable amount of time.
+
+P45: Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:20 [Around the same time some prob..]  (5:998-5:2037)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Around the same time some problems surfaced which were not connected to Scrum
+usage but needed to be resolved before moving to Scrum on a large scale:
+1. Communication among product management, software architects, development
+teams, and quality assurance was not strong enough to ensure a common understanding of development goals.
+Product management often was unable to communicate product requirements to software engineers.
+Languages and notations differed signiﬁcantly.
+2. The role of software architecture (high-level design) was not yet well understood.
+This led to increased dependencies among products and components, e. g. because
+componentization was weak and code reuse was highly valued. Another result was
+the lack of conceptual preparation for the development teams for whom there were
+too many obstacles to harmonize on interfaces and adhere to delivery timelines.
+To overcome both problems a team was formed to provide education on software architecture for developers, software architects, and product management.
+
+P45: Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:35 [Upscaling methods were mainly ..]  (9:1718-9:2252)   (Super)
+Codes:	[C: Adapting non-development parts of organization to agile - Family: RQ3: Challenges] [C: Agile by the book is not good enough (no exact recipes) - Family: RQ3: Challenges] [C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges]
+No memos
+
+Upscaling methods were mainly driven by theoretical considerations. Implementing
+themrevealedthatthe skills andknowledge ofproductmanagers and software architects
+were often not sufﬁcient to manage a number of Scrum teams concurrently. Product
+teams had to communicate with business management, marketing, and ﬁeld services
+who were not fully prepared for how Scrum and the product teams’ expectations would
+change their daily work. The process of software requirements engineering were found
+to require a major adjustment.
+
+P45: Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:39 [A three-level hierarchy ofteam..]  (10:1787-10:2327)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] [T: Impact on organization - Family: RQ2: Transformation process] 
+No memos
+
+A three-level hierarchy ofteams to manage requirements for product backlog generation has not yet proven to work as expected. Currently the high complexity of SAP’s
+products which is reﬂected in the organizational structure of SAP’s development areas
+makes it impossible to drop one hierarchy level of project management. This observation has triggered signiﬁcant effort to rigorously break up SAP’s product lines into
+separate components to ensure that no product component needs more resources than
+the 130 people as calculated above.
+
+P46: Seffernick 2007 - Enabling Agile In A Large Organization Our Journey Down The Yellow Brick Road.pdf - 46:34 [After month 16 of our journey,..]  (3:3987-3:4981)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] [C: Mapping agile to match the old model - Family: RQ3: Challenges] [I: Training - Family: Investing in change] [S: Emphasize principles instead of mechanics - Family: RQ4: Success factors] [S: Pay attention to the product owner role - Family: RQ4: Success factors]
+No memos
+
+After month 16 of our journey, focus shifted 
+toward connecting the product owners, ScrumMasters 
+and technology managers prior to the project start. Our 
+projects were starting without good backlogs and the 
+level of product owner involvement was inconsistent, 
+causing projects to start slowly. After some soul 
+searching we concluded that we were treating agile 
+development as a paradigm shift (mapping existing 
+roles to Scrum roles, yet not doing anything different). 
+What we needed was to educate it as a new paradigm. 
+The result was Agile Connections. This is a sequence 
+of training that starts with an overview to create 
+awareness and desire to use agile practices in 
+delivering the business value. Then, hands-on 
+workshops transfer knowledge, followed by jump 
+starts and coaching to help reinforce the ability to use 
+the practices. The overall goal is to get the product 
+owner, ScrumMaster and technology manager together 
+to enable projects to start better.
+
+P48: Smits 2011 - Agile Experience Report - Transition And Complexity At Cisco Voice Technology Group.pdf - 48:31 [An executive Portfolio Investm..]  (4:2117-4:2972)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+An executive Portfolio Investment Committee was 
+initiated with the goal of working across all of the 
+business units in our Technology Group to set 
+investment priorities. Two problems emerged – we 
+did not have the management information to support 
+the decisions of the Portfolio Committee, and 
+information about the strategic decisions and 
+direction were not effectively communicated to the 
+organization so that all participants in the system 
+release planning process were informed. The next 
+stage of “decomposition” of Themes, Epics and User 
+Stories didn’t really happen, so that the chunks of 
+work were so large that the Product Teams and 
+Scrum teams had to spend an inordinate amount of 
+time breaking things down, (and clarifying the 
+requirements at the lower levels) before any sprint 
+zero or sprint planning could occur.
+
+P51: Vlaanderen 2012 - Growing Into Agility - Process Implementation Paths For Scrum.pdf - 51:13 [The sprint planning meeting in..]  (5:720-5:1333)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+The sprint planning meeting initially took approximately one day, because
+all the user stories had to be discussed, estimated and planned at once. This
+was considered too long to be eﬀective. Therefore, story estimation is now done
+regularly throughout the sprints. This requires shorter bursts of attention, which
+eases the process. On the day of the sprint planning, the product owners present
+the user stories to their own team(s), followed by the selection and planning of
+user stories by the development teams. The process of selection and planning
+now takes approximately one hour, instead of one day.
+
+P51: Vlaanderen 2012 - Growing Into Agility - Process Implementation Paths For Scrum.pdf - 51:26 [Until 2007, agile product deve..]  (6:2886-7:730)   (Super)
+Codes:	[C: Difficulties with requirements engineering or backlog - Family: RQ3: Challenges] 
+No memos
+
+Until 2007, agile product development was accompanied by
+non-agile product management. Although several stages of elaboration were employed, no ﬁxed cycles were used. This caused that product managers did not
+manage to provide development with suﬃciently detailed requirements before122 K. Vlaanderen et al.
+the start of each sprint. To improve this, the product management team adjusted their process to the Scrum principles as well. This implied a continuous
+adaptation of the product backlog to a changing environment.
+Since Scrum itself does not provide guidelines for eﬀectively managing large amounts of requirements of diﬀerent granularity, a set of stages has been introduced, called
+the Agile Requirements Reﬁnery [26]. Within the agile requirements reﬁnery, an
+idea will generally move through the stages vision, theme, concept, and requirement deﬁnition. During these stages, requirements are reﬁned with details and
+speciﬁcations. This has resulted in a new categorization on the product backlog.
 
@@
 % ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
 \section{Discussion}

### 2014-05-12 journal/article.tex
Work in progress

--- 2014-05-11 journal/article.tex
@@
 transformation.
-% separate QA [P3], --> Later
-% human resources --> Later
 
@@
 
+% This where?
+
 Some cases reported that development efforts were controlled on a top level by a
@@
 
+% Requirements engineering is difficult
+
+Point --> Requirements management also needs to adopt the new methodology
+Product management was initially done in a traditional way, but it was not
+capable of iteratively supplying small enough requirements to the agile
+development process [P51]. The agile method had to be extended to include a
+process on how to get large requirements into team backlogs [P31].
+
+And --> The methodology needs to be adjusted
+Scrum had to be adjusted to fit for large scale development, as the methodology
+did not have guidelines on how to large and complex requirements [P31, P51].
+
+In agile methods requirements engineering is usually a task closely tied to the
+development team. However, large development projects will demand more
+requirements management effort.
+This is apparent in cases where complexity of the product will require
+additional layers of product management [P45], requirements are created by
+several stakehoder groups and development teams can not be in contact with all
+of them [P39], or teams do not have access to stakeholders due to a distributed
+seetting [P24].
+
+In the development life cycle requirements were in some cases first defined on a
+high level in marketing requirements documents (MRD) [P1, P31] or functional
+specifications [P33]. The high level requirements needed to be elaborated to be
+used by development teams. The requirements breakdown in it self was found to be
+difficult [P1, P33], and also when to do the breakdown to which detail level
+[P17].
+
+Several cases described that much learning was needed to master the new process
+of creating user stories, both on product management and development levels
+[P11, P27, P30, P33, P37]. Also the effort estimation for stories was hard to do
+well [P27, P37].
+
+Product managers and business analysts were struggling in creating high level
+requirements [P33, P37], and teams were struggling in breaking them down to an
+estimable size [P33]. Smits and Rilliet [P48] describe how high level product
+management delivered requirements in too large chunks resulting in development
+teams spending large amounts of time on defining suitaby sized stories.
+
+Some cases reported  Some organizations used a quality assurance group separate
+from development.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-"C: Adapting non-development parts of organization to agile"
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MARKETING NEEDS TO KNOW ON BEFOREHAND WHAT WILL BE RELEASED
-% 
-% 
-% * Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
-% * Problem: Sales and marketing organization had a different vision of the product
-% * Problem: Product marketing was not collaborative, and in old mindset
-% * Problem: Requirements --> user stories more granular and prioritized
-% [P31]
-% After several months of working on the new products, the Agile teams “drained”
-% the backlog.   For a while, they kept busy working on new user stories that came
-% directly from the strategic customer.  This began to lead to a disconnect
-% between what the marketing and sales organizations believed to be the overall
-% direction of the product, and the actual direction it was taking based on the
-% needs of a single customer. It was clearly time for a regroup, and the
-% engineering team pushed back on the MRD.  They were in full production mode
-% using Agile and they needed more granular, prioritized user stories to fuel
-% their efforts.  The teams needed product marketing to be more collaborative in
-% the decomposition of the requirements. This was a totally foreign concept for
-% the product marketing team. They had approached the relationship with
-% engineering as a “black box” – high level requirements went in, and something
-% eventually came out (usually late, and sometimes off the mark on what they
-% expected).
-% 
-% 
-% * Challenge: Accomodate time for export clearance
-% * Challenge: Acquire third-party licensing authority --> prior to product release
-% * Challenge: These processes will remain traditional
-% * Challenge: Marketing needs some lead time
-% * Enablement activities (sales materials, training)
-% * These steps take time, but they are part of the product lifecycle
-% * Challenge: Must know beforehand what will be in the product
-% [P31]
-% While changing our process in the front end of the value chain began to yield
-% success, we discovered that there was also a traditional wall at the back end:
-% our product release process.
-% While the Agile approach enables teams to be flexible with regard to
-% prioritization, and very adaptive to changing market conditions and customer
-% needs, there are certain “laws of physics” required to release commercial
-% product that cannot be altered. For example, the time required to acquire export
-% clearance or to acquire third-party licensing authority prior to product
-% release.  As a vendor that makes products that must be marketed, sold and
-% implemented, there are several traditional processes that are involved in
-% releasing our products to market.
-% 
-% First there are the regulatory requirements such as export compliance relative
-% to exporting code with certain levels of encryption,  then there are the
-% marketing launch plans (demand generation campaigns, collateral, Web site,
-% public relations), and finally, the field enablement activities (sales
-% materials, training).  All these initiatives take time, and they cannot begin
-% without knowledge of exactly what is going to be in the product.
-% 
-% 
-% * Problem: Marketing had to know on beforehand what will be in the release
-% * Problem: Business went with requiring a fixed delivery date and a fixed feature set
-% [P31]
-% Serious friction arose. The development teams became frustrated because they
-% felt marketing and the NPI process was stalling their progress and throwing up
-% barriers against the core principles they had adopted.  Marketing, who needed to
-% begin prepping for the launch of the product about three months before the
-% actual date, found themselves struggling to message and develop materials and
-% campaigns without a clear understanding of the features and functionality that
-% would actually be “ready” on the release date. And, the NPI group could not
-% apply for export compliance nor could they request any third party license
-% authorization without knowing what was actually going to be in the product.
-% 
-% 
-% * Solution: Development and product marketing work together
-% * Solution: Product marketing can require a release to be completed in 90 days
-% [P31]
-% What we instituted was what we call an “Iron Triangle.” The concept is that at
-% some number of sprints out before a release, we would commit to delivering a
-% fixed set of features and functionality  at a certain date.  Of course, there
-% was always the opportunity for us to deliver more than what we committed to, 
-% but we absolutely had to deliver the set features by the set date.
-% 
-% If  product marketing wants to release product for any reason they can “declare”
-% a release date, and  we’re able to lock the current feature set and have  the
-% code ready to release within two sprints of the date they declare the release
-% necessary.
-
-% ---------------------------------------------------------------------
-
-REQUIREMENT AND MRD COMMUNICATION PROBLEMS
+LARGE SCALE DEVELOPMENT EFFORT WILL DEMAND MORE FROM REQUIREMENTS MANAGEMENT
+[P17]
+% [P31] --> Scrum needs adjustment [P51b]
+% [P39] --> All subject matter experts can not be brought together
+[P46] --> High level stakeholders must be connected at project start
+[P45] --> Communicating requirements down
+[P51b] -> Non agile requirements management
+  |
+  v
+BREAKDOWN IS HARD IN LARGE SCALE
+[P1]  --> Extracting from MRDs hard (in general)
+[P17] --> Have correct granularity at each phase in the life cycle
+[P33] --> Breaking down backlogs was really hard
+[P37] --> Business analysts not used to breaking down
+[P48] --> Too large requirements to teams
+[P51b] ->
+
+THERE IS A GAP BETWEEN HIGH LEVEL AND DEVELOPMENT
+[P31] --> Product marketing team was not collaborative at start (product managers)
+[P39] --> Requirements are gathered in a separate function
+[P45]
+
+PROBLEMS IN REQUIREMENTS DEFINITION BEGINNING
+[P31]
+[P46]
+
+CUSTOMERS NOT AVAILABLE
+% [P39]
+
+GENERAL AGILE PROBLEM
+[P5]  --> Requirements defined ambiguously
+
+BAD BEHAVIORS IN REQUIREMENTS MANAGEMENT
+[P31] --> Backlog bartering
+[P31] --> Teams listened to single customer instead of big picture set by product marketing team
+
+CREATING USER STORIES WAS GENERALLY HARD TO LEARN
+
+* Writing user stories was difficult at start --> It was hard to learn the breakdown practice
+* Hard to do well in sprint planning session
+[P11]
+Writing and grooming user stories seems like a simple process during Scrum
+training.  When many teams start sprinting at once, no one starts off as an
+expert at writing stories.  In the mad rush to convert requirements to stories,
+many stories had the correct format, but were often too big, too ambiguous, or
+too prescriptive.  Product owners and teams who held grooming sessions once or
+twice a week over time refined their backlogs so that planning meetings were
+short, and there was no wasted time between sprints trying to figure out what to
+do next.  For one untended product backlog, we instituted “speed grooming” where
+we spent a maximum of five minutes discussing a story to categorize it as “ready
+for a team,” “obsolete,” “more grooming,” etc., so the stories could be readied
+for upcoming sprints.  Most of our stories were for system software, not end
+user features, so teams consistently underestimated how long a story would take,
+and would complain that it couldn’t be broken down further.  All of sudden they
+were being required to break down work that used to take months into tasks of a
+few hours.  In hindsight, we could have reduced stress on everyone if we had
+devoted more time upfront to developing everyone’s skills in writing stories and
+decomposing epics.
+
+* In the beginning the team had problems learning how to break down requirements into stories
+* Solution: Better customer and team collaboration --> More real customer, less internal customer
+[P27]
+The team had difficulty estimating tasks during the planning session: What
+should go into a story? How should a task card be written? Is the story points
+system better than the hour’s system? We also had a lack of well-defined
+acceptance criteria before and during the planning session.
+
+One solution we came up with was to improve the relationship between us and our
+internal customers. Planning sessions were dreaded - long and not well
+structured. Our team was driven based on the internal customer requirements,
+which went into a backlog that the program manager also known as Susha
+maintained. The more efficient way to understand what a task entailed was to
+involve the customers in our planning meeting and pair up with them on writing
+the actual story. The customers partnered up with us for 15 minutes during the
+planning sessions.
+
+* Agile raised awareness of requirements churn
+* Completed stories did not always add value for clients --> Stories defined badly
+* Problem: There was a separation between development teams and customers
+[P30]
+The introduction of Agile raised visibility of the requirements churn and lack
+of clarity in feature definition, but writing good user stories remained
+difficult, due in part to the separation between the client and product
+development teams. Engineering became known as the “story-eating engine.”
+Stories were being completed, but did not always add value for clients.
+Agile successfully exposed the next areas needing attention: defining
+value-added stories.
+
+* The shift in practice from functional specs to user stories is challenging
+[P33]
+The transition from writing full product or functional specs to writing user
+stories is challenging. Learn how to break up larger features into smaller user
+stories, and prioritize them.
+
+* Business analysts have problems with stories
+* No skills in requirements breakdown
+* Story breakdown was difficult practice --> Teams
+* Things got better with time
+[P37]
+Our business analysts were running into the same issue one level up. They were
+not used to providing feature requirements in story form, much less breaking
+down these ”feature” level stories into stories that could be reasonably
+estimated by the development team. I realized almost immediately the stories
+were too big, but you can’t force immediate understanding. Like any skill,
+estimating takes practice. We went through many agonizing estimation and story
+breakdown sessions before we started to get some traction. As we began to build
+up some historical data this closed the feedback loop. The team was able to see
+that we were frequently underestimating and that this was directly connected to
+the trouble we were still having with story breakdown.
+
+We started finishing cards, but found that estimates of hours often took weeks.
+Stories would spawn additional stories in order to reach completion. For
+example, a story that started with a two day estimate would generate a week’s
+worth of stories (sometimes even more than a week!). Many agile developers would
+say ”Great! That’s part of agile.” Gaining clarity on a story and understanding
+that there is more to it is part of the process. Unfortunately this was
+happening far too often.
+
+"C: Difficulties with requirements engineering or backlog"
 
@@
 * MRDs not based on reality
-[P2]
+* Extracting user stories from MRDs --> Breakdown hard because of too much detail in MRDs
+[P1, P2]
 Extracting user stories from existing, very detailed ‘marketing requirements
@@
 The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
+
+* The requirements were vaguely defined in the beginning --> Abmbiguity in requiements
+* There was no extended QA period when requirements ambiguities were resolved in the old model
+* Connection to QA
+* Solution: better requirements management
+* Solution: New role to bridge gap between high and detailed level
+* Solution: Requirements are elaborated JIT
+* Solution: Requirements runway [P17]
+[P5, P17]
+There was constant debate over both the meaning of the requirement and whether
+or not the requirement was met.  Quality Assurance engineers were frustrated
+trying to determine whether or not we should accept this feature.  Simply put,
+we had not invested enough of our velocity in either up front or ongoing
+requirements elaboration.
+Going into the second release, we knew we had to do a better job with managing,
+elaborating, and prioritizing requirements.  This had strictly been an issue
+regarding where we chose to invest our energy.
+Properly managing requirements was critical.
+
+* The scale of the enterprise did not allow for ``throw away stories''
+* Problem: Scale/size demands better requirements management 
+[P17]
+As BMC went through their initial iterations, they realized there was often a
+gap between the level of detail in their system level features and the detail
+needed for iteration planning and execution. The use of simple “throw away
+stories” was insufficient to their challenge of scale. To bridge this gap, BMC
+created the role of Requirements Architect as part of each development team.
+The responsibility of each Requirements Architect is to take the high-level
+features defined by product management and decompose these on a just-in-time
+basis into the more detailed requirements and stories needed to drive iteration
+planning and development.
+This new role increased the accuracy of development estimates and improved the
+collaboration between the development team and product management.
+
+* Problem: Correct granularity in requirements at each stage in the life cycle
+[P17]
+One of the challenges BMC faces is defining the right level of requirement
+detail at the right time in the life cycle. BMC is now using Rally Software to
+manage a prioritized feature backlog and to elaborate the requirements and
+stories for each component that will likely be needed to implement the feature.
+
+* High level requirements are problematic
+[P30]
+Continuing progress includes:
+* Managing an effective enterprise backlog.
+
+* Large scale requires adjustment to Scrum practices --> Requirements --> Getting things in backlog
+* Enterprise QA
+[P31]
+we discovered some areas that would require us to adapt our chosen Scrum
+methodology to account for the realities of complex, large-scale development. 
+Two key activities fell outside the scope of our original Scrum framework
+1. Informing how things get into the backlog, and 2. Enterprise testing.
+
+* The agile setup allowed for change --> Backlogs could be changed at beginning of each sprint
+* Backlog bartering --> Negative effects --> Distrust
+* Collaboration between product marketing and development to manage requirements
+[P31]
+Once the organization began to understand the increased productivity of the
+Agile teams, and they saw that every three weeks the teams selected what they
+would work on, they began to grasp the possibility that any feature request
+could be delivered in mere weeks.
+And, suddenly, we found the backlog was “up for grabs” at the beginning of each
+sprint.
+Requests came flooding in from the field, from marketing, from our support
+organization. They would be automatically added to the backlog, and each sprint
+became an opportunity to evaluate and reprioritize.
+And, since the Agile approach increases collaboration, everyone had the
+opportunity to “lobby” for his or her requests to make the backlog. This
+constant influx of demand from several sources began to put a strain on the
+system.  Questions about how to prioritize the backlog items and a subversive
+process of “backlog bartering” began. This negatively impacted the teams and
+their performance, and it also generated distrust among the various groups.
+
+* Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
+* Problem: Sales and marketing organization had a different vision of the product
+* Problem: Product marketing was not collaborative, and in old mindset
+* Problem: Requirements --> user stories more granular and prioritized
+[P31]
+After several months of working on the new products, the Agile teams “drained”
+the backlog.   For a while, they kept busy working on new user stories that came
+directly from the strategic customer.  This began to lead to a disconnect
+between what the marketing and sales organizations believed to be the overall
+direction of the product, and the actual direction it was taking based on the
+needs of a single customer.
+It was clearly time for a regroup, and the engineering team pushed back on the
+MRD.  They were in full production mode using Agile and they needed more
+granular, prioritized user stories to fuel their efforts.  The teams needed
+product marketing to be more collaborative in the decomposition of the
+requirements. This was a totally foreign concept for the product marketing team.
+ They had approached the relationship with engineering as a “black box” –
+highlevel requirements went in, and something eventually came out (usually late,
+and sometimes off the mark on what they expected).
 
@@
 
-* Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
-* Problem: Sales and marketing organization had a different vision of the product
-* Problem: Product marketing was not collaborative, and in old mindset
-* Problem: Requirements --> user stories more granular and prioritized
-[P31]
+* Backlog management was difficult
+* Requirements breakdown was very difficult
+* Teams did not have insight on how to organize backlogs across teams
+* Inconsistency of backlog management practice between teams
+[P33]
+In addition to the hottest ideas coming off the IdeaExchange, salesforce.com’s
+product managers also have backlogs of important features, and these product
+backlogs are an integral part of ADM.
+The initial creation and prioritization of product backlogs was an ominous
+challenge, and nearly impossible to achieve in the two-month timeframe the
+product managers were allotted. Breaking down requirements into both epic and
+detailed user stories was arduous, and the way in which the requirements were
+broken down was inconsistent between teams because each team had limited insight
+into how other teams were organizing their product backlogs.
+To address this issue, salesforce.com built an in-house app called  Scrumforce
+
+* Challenge: Writing user stories is hard
+* The shift in practice from functional specs to user stories is challenging
+* User stories should be written complete and accurately --> in the beginning of the sprint
+* (no further mentionings of large scale)
+[P33]
+The transition from writing full product or functional specs to writing user
+stories is challenging. Learn how to break up larger features into smaller user
+stories, and prioritize them.
+
+With user stories come smaller and more vague requirements, and because of the
+agile nature of the development process, the product owner has a tendency to
+change requirements during a sprint. Often, this is due to the product owner not
+specifying the full requirements before the sprint starts. This can waste time
+and result in work that is thrown away. User stories should be complete, have
+acceptance criteria, and be usability tested before being added to a sprint.
+Having clear guidelines prepared for when the team is ready to tackle a user
+story will help reduce waste.
+
+* Requirement management was difficult
+* Actual stakeholders were not available for teams 
+[P39]
+there were a number of difficulties in requirements management. Actual
+stakeholders weren’t available for the teams due to  many reasons and they used
+another consulting company in the UK to gather their requirements, design UI and
+communicate them back to teams. Even more requirements were provided from risk
+management subject matter experts in London.
+Comment: This sounds a bit like the initial state.
+
+* Large scale problem
+* It is impossible to bring all subject matters together in a large project
+[P39]
+Product discovery workshops are an extremely efficient technique allowing for
+quick elaboration of a product roadmap and creating mid-term requirements
+details. However, as most efficient tools do, it has certain limitations. One of
+them is the need to have the PO and the team co-located for a sprint which might
+be a budget issue for a distributed project. Another one is the nature of a PO
+role in enterprise projects where you simply can’t bring all the subject matter
+experts in one place with the team for any reasonable amount of time.
+
+* Product management unable to communicate requirements down to development
+* High level design functions had hard time communicating together and with development
+[P45]
+Around the same time some problems surfaced which were not connected to Scrum
+usage but needed to be resolved before moving to Scrum on a large scale:
+1. Communication among product management, software architects, development
+   teams, and quality assurance was not strong enough to ensure a common
+   understanding of development goals. Product management often was unable to
+   communicate product requirements to software engineers. Languages and
+   notations differed significantly.
+2. The role of software architecture (high-level design) was not yet well
+   understood. This led to increased dependencies among products and components,
+   e.g. because componentization was weak and code reuse was highly valued.
+   Another result was the lack of conceptual preparation for the development
+   teams for whom there were too many obstacles to harmonize on interfaces and
+   adhere to delivery timelines.
+To overcome both problems a team was formed to provide education on software
+architecture for developers, software architects, and product management.
 
 * Requirements engineering must be done differently
+* Product management has a hard time managing many agile teams simultaneously
 [P45]
-Product teams had to communicate with business management, marketing, and field
-services who were not fully prepared for how Scrum and the product teams’
-expectations would change their daily work. The process of software requirements
-engineering were found to require a major adjustment.
+Upscaling methods were mainly driven by theoretical considerations. Implementing
+them revealed that the skills and knowledge of product managers and software
+architects were often not sufficient to manage a number of Scrum teams
+concurrently. Product teams had to communicate with business management,
+marketing, and field services who were not fully prepared for how Scrum and the
+product teams’ expectations would change their daily work. The process of
+software requirements engineering were found to require a major adjustment.
+
+* Backlog could not be managed sensibly for teams over 130 people
+* Product management with a too large scope did not work
+* Managing requirements for product backlog is difficult in large scale (generally)
+* Hierarchial requirements management is troublesome
+* The organization's hierachy is reflected in requirements management
+[P45]
+A three-level hierarchy of teams to manage requirements for product backlog
+generation has not yet proven to work as expected. Currently the high complexity
+of SAP’s products which is reflected in the organizational structure of SAP’s
+development areas makes it impossible to drop one hierarchy level of project
+management. This observation has triggered significant effort to rigorously
+break up SAP’s product lines into separate components to ensure that no product
+component needs more resources than the 130 people as calculated above.
+
+* Problems with backlogs at project start --> inconsistency at project start
+* Product management (PO) tasks have to be done at project start
+* Technology and business leadership (high level) needs to be engaged strongly at project start
+[P46]
+After month 16 of our journey, focus shifted toward connecting the product
+owners, Scrum Masters and technology managers prior to the project start. Our
+projects were starting without good backlogs and the level of product owner
+involvement was inconsistent, causing projects to start slowly. After some soul
+searching we concluded that we were treating agile development as a paradigm
+shift (mapping existing roles to Scrum roles, yet not doing anything different).
+What we needed was to educate it as a new paradigm. The result was Agile
+Connections. This is a sequence of training that starts with an overview to
+create awareness and desire to use agile practices in delivering the business
+value. Then, hands-on workshops transfer knowledge, followed by jump starts and
+coaching to help reinforce the ability to use the practices. The overall goal is
+to get the product owner, ScrumMaster and technology manager together to enable
+projects to start better.
+
+* Too large requirements to teams --> Could not break down stories
+* Difficulties in breaking down requirements to stories
+* Business (high level) requirements management gave too large chunks of requirements to development
+[P48]
+An executive Portfolio Investment Committee was initiated with the goal of
+working across all of the business units in our Technology Group to set
+investment priorities. Two problems emerged – we did not have the management
+information to support the decisions of the Portfolio Committee, and information
+about the strategic decisions and direction were not effectively communicated to
+the organization so that all participants in the system release planning process
+were informed. The next stage of “decomposition” of Themes, Epics and User
+Stories didn’t really happen, so that the chunks of work were so large that the
+Product Teams and Scrum teams had to spend an inordinate amount of time breaking
+things down, (and clarifying the requirements at the lower levels) before any
+sprint zero or sprint planning could occur.
+
+* Sprint planning was taking too long --> User story breakdown was too difficult
+* Solution: Story estimation is done more iteratively
+[P51] (Case A)
+The sprint planning meeting initially took approximately one day, because all
+the user stories had to be discussed, estimated and planned at once. This was
+considered too long to be effective. Therefore, story estimation is now done
+regularly throughout the sprints. This requires shorter bursts of attention,
+which eases the process. On the day of the sprint planning, the product owners
+present the user stories to their own team(s), followed by the selection and
+planning of user stories by the development teams. The process of selection and
+planning now takes approximately one hour, instead of one day.
+
+* Requirements given from product management too large --> not enough detail for development
+* Product management adapted to Scrum --> Continuous attention to product backlog
+* Scrum does not tell how to manage large requirements on different levels 
+[P51] (Case B)
+Until 2007, agile product development was accompanied by non-agile product
+management. Although several stages of elaboration were employed, no fixed
+cycles were used. This caused that product managers did not manage to provide
+development with sufficiently detailed requirements before the start of each
+sprint. To improve this, the product management team adjusted their process to
+the Scrum principles as well. This implied a continuous adaptation of the
+product backlog to a changing environment. Since Scrum itself does not provide
+guidelines for effectively managing large amounts of requirements of different
+granularity, a set of stages has been introduced, called the Agile Requirements
+Refinery [26]. Within the agile requirements refinery, an idea will generally
+move through the stages vision, theme, concept, and requirement definition.
+During these stages, requirements are refined with details and specifications.
+This has resulted in a new categorization on the product backlog.
+
+MULTISITE CASE
+
+% * Requirements engineerinf was difficult in Multisite
+% [P24]
+% Initially we started with the Product Owner (PO) as a role for sales and account
+% management. It worked reasonably well with local teams in Denmark when they
+% could collaborate and meet when required. But it did not work well with teams in
+% Bangladesh. We moved some of the very experienced senior developers and
+% consultants into global PO roles to establish a more structured flow of tasks
+% and collaboration with the teams. To establish a structure to improve the flow
+% of work between the Danish and the Bangladeshi teams we introduced a “Proxy PO”
+% for each team in Bangladesh.
+% 
+% Understanding the problem domain and how the different requirements were related
+% was still hard. We used many different techniques to move the domain knowledge
+% and user understanding out to the different teams.
+% 
+% 
+% * Solution: Global planning is required for multisite
+% [P24]
+% Global planning:
+% Having one team and therefore one sprint backlog meant, that we had to do sprint
+% planning together. Not an easy task, but after some experiments, we came up with
+% a solution that worked for us:
+% 
+% • The whole global team was presented with the user stories online
+% • The global team went offline and started to break down the stories and meet to
+%   synchronize sprint backlog items into one sprint backlog
+% • The global team informed the PO of the result
+% 
+% We spent a lot of time trying to make it work well, and it required a high level
+% of trust between all parties. It also required a lot of knowledge of the
+% business domain, but made it easy to ask for the relevant knowledge when doing
+% the breakdown and planning.
+% 
+% 
+% * Colocation of requirements communication is required even in multisite  
+% [P24]
+% The R\&D team could build the best quality code, but we kept building elements
+% of the wrong things. It was almost like building a top of the line V12 engine, for
+% a solar cell car race. The trust between the team and the PO were at stake. “I
+% told you to build an engine, and you should have known it was an electric engine
+% I needed”. So once again we had to improve the quality and understanding of
+% requirements between the PO and the global team both before and during the
+% actual development in the sprint. After retrospectives and more discussions, we
+% decided to try moving the PO into the Danish R\&D team room.
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
 

### 2014-05-12 journal/article.tex
Requirements engineering challenges: Completed

--- 2014-05-11 journal/article.tex
@@
 processes of acquiring export clearances and licensing authority required that
-the the feature set of the product was known well on beforehand [P31].
+the the feature set of the product was known well on beforehand.
 As a result of these requests development felt that their process was being
@@
 
+% Requirements engineering is difficult
+
+In agile methods requirements engineering is usually closely tied to the
+development team. However, large development projects will demand additional
+high level requirements management.
+This is apparent in cases where complexity of the product will require
+additional layers of product management [P45], requirements are created by
+several stakehoder groups and development teams can not be in contact with all
+of them [P39], or teams do not have access to stakeholders due to a distributed
+seetting [P24].
+As an integral part of the software development lifecycle, requirements
+management will be affected by the new agile process. In large organizations
+product and requirements management was initially done in a traditional way, but
+it was contrasted by the agile process that required iteratively supplying small
+requirements to development [P51]. The agile method had to be extended to
+include a process on how to get large requirements into team backlogs [P31].
+
+Requirements were in some cases first defined on a high level in marketing
+requirements documents (MRD) [P1, P31] or functional specifications [P33]. These
+high level requirements needed to be elaborated to be used by agile development
+teams. The requirements breakdown in it self was found to be difficult [P1,
+P33], and also when to do the breakdown and to which detail level [P17].
+Product managers and business analysts were struggling in creating high level
+requirements [P33, P37], and teams were struggling in breaking them down to an
+estimable size [P33]. Smits and Rilliet [P48] describe how high level product
+management delivered requirements in too large chunks resulting in development
+teams spending large amounts of time on defining suitaby sized stories. In one
+case the teams were overwhelmed by reprioritization and requirements bartering
+when stakeholders learned that requirements could be rearranged flexibly [P31].
+
+Several cases highlighted that much learning was needed to master the new
+process of creating user stories, both on product management and development
+levels [P27, P30, P33, P37]. There were problems such as ambiguity in
+requirements [P5, P11], and effort estimation for stories was hard to do well
+[P27, P37].
+
 % --
+
+% This where?
 
@@
 
+% --
+
+Some cases reported  Some organizations used a quality assurance group separate
+from development.
+
 Problems in adopting agile were even showing in functions outside the software
@@
 
-"C: Adapting non-development parts of organization to agile"
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MARKETING NEEDS TO KNOW ON BEFOREHAND WHAT WILL BE RELEASED
-% 
-% 
-% * Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
-% * Problem: Sales and marketing organization had a different vision of the product
-% * Problem: Product marketing was not collaborative, and in old mindset
-% * Problem: Requirements --> user stories more granular and prioritized
-% [P31]
-% After several months of working on the new products, the Agile teams “drained”
-% the backlog.   For a while, they kept busy working on new user stories that came
-% directly from the strategic customer.  This began to lead to a disconnect
-% between what the marketing and sales organizations believed to be the overall
-% direction of the product, and the actual direction it was taking based on the
-% needs of a single customer. It was clearly time for a regroup, and the
-% engineering team pushed back on the MRD.  They were in full production mode
-% using Agile and they needed more granular, prioritized user stories to fuel
-% their efforts.  The teams needed product marketing to be more collaborative in
-% the decomposition of the requirements. This was a totally foreign concept for
-% the product marketing team. They had approached the relationship with
-% engineering as a “black box” – high level requirements went in, and something
-% eventually came out (usually late, and sometimes off the mark on what they
-% expected).
-% 
-% 
-% * Challenge: Accomodate time for export clearance
-% * Challenge: Acquire third-party licensing authority --> prior to product release
-% * Challenge: These processes will remain traditional
-% * Challenge: Marketing needs some lead time
-% * Enablement activities (sales materials, training)
-% * These steps take time, but they are part of the product lifecycle
-% * Challenge: Must know beforehand what will be in the product
-% [P31]
-% While changing our process in the front end of the value chain began to yield
-% success, we discovered that there was also a traditional wall at the back end:
-% our product release process.
-% While the Agile approach enables teams to be flexible with regard to
-% prioritization, and very adaptive to changing market conditions and customer
-% needs, there are certain “laws of physics” required to release commercial
-% product that cannot be altered. For example, the time required to acquire export
-% clearance or to acquire third-party licensing authority prior to product
-% release.  As a vendor that makes products that must be marketed, sold and
-% implemented, there are several traditional processes that are involved in
-% releasing our products to market.
-% 
-% First there are the regulatory requirements such as export compliance relative
-% to exporting code with certain levels of encryption,  then there are the
-% marketing launch plans (demand generation campaigns, collateral, Web site,
-% public relations), and finally, the field enablement activities (sales
-% materials, training).  All these initiatives take time, and they cannot begin
-% without knowledge of exactly what is going to be in the product.
-% 
-% 
-% * Problem: Marketing had to know on beforehand what will be in the release
-% * Problem: Business went with requiring a fixed delivery date and a fixed feature set
-% [P31]
-% Serious friction arose. The development teams became frustrated because they
-% felt marketing and the NPI process was stalling their progress and throwing up
-% barriers against the core principles they had adopted.  Marketing, who needed to
-% begin prepping for the launch of the product about three months before the
-% actual date, found themselves struggling to message and develop materials and
-% campaigns without a clear understanding of the features and functionality that
-% would actually be “ready” on the release date. And, the NPI group could not
-% apply for export compliance nor could they request any third party license
-% authorization without knowing what was actually going to be in the product.
-% 
-% 
-% * Solution: Development and product marketing work together
-% * Solution: Product marketing can require a release to be completed in 90 days
-% [P31]
-% What we instituted was what we call an “Iron Triangle.” The concept is that at
-% some number of sprints out before a release, we would commit to delivering a
-% fixed set of features and functionality  at a certain date.  Of course, there
-% was always the opportunity for us to deliver more than what we committed to, 
-% but we absolutely had to deliver the set features by the set date.
-% 
-% If  product marketing wants to release product for any reason they can “declare”
-% a release date, and  we’re able to lock the current feature set and have  the
-% code ready to release within two sprints of the date they declare the release
-% necessary.
-
 % ---------------------------------------------------------------------
 
-REQUIREMENT AND MRD COMMUNICATION PROBLEMS
-
-* Not agile: MRDs --> Stories had to be extracted from detailed documentation --> Not good
-* MRDs not based on reality
-[P2]
-Extracting user stories from existing, very detailed ‘marketing requirements
-documents’ (MRDs) was another workaround. These MRDs were a prominent artifact
-found across the organisation. Leaders of Project Z claimed that it was very
-difficult to work with such detailed documents in a project whose scope could
-change rapidly; one agile developer described MRDs as “not based in reality at
-all”.
-Comment:
-The problem with MRDs is a symptom of the challenge to integrate "the business" with the agile process.
-
-* Problem: MRD is not dynamic
-* Problem: MRD requirements scoped once a year --> Teams decide for themselves
-* Problem: Product marketing team was very disrupted of the requirement of collaboration!
-[P31]
-Further, the MRD needed to be more dynamic. Because the teams had achieved a
-predictable, efficient pace of delivery, scoping the requirements of a release
-once a year wasn’t enough to feed the engine.  Without the proper fuel the teams
-were left to decide for themselves what new things to work on, driving further
-disconnect between the groups. The concept of continuous collaboration
-throughout the development cycle was a very disruptive concept for the product
-marketing team.
-
-* Problem: Teams completed backlog --> Continued with stories direct from customer --> Single customer bias
-* Problem: Sales and marketing organization had a different vision of the product
-* Problem: Product marketing was not collaborative, and in old mindset
-* Problem: Requirements --> user stories more granular and prioritized
-[P31]
-
-* Requirements engineering must be done differently
-[P45]
-Product teams had to communicate with business management, marketing, and field
-services who were not fully prepared for how Scrum and the product teams’
-expectations would change their daily work. The process of software requirements
-engineering were found to require a major adjustment.
+% ---------------------------------------------------------------------
 

### 2014-05-12 journal/article.tex
Problems with HR

@@
 
+% Problems with HR
+
+Problems in adopting agile were even showing in functions outside the software
+delivery lifecycle. In several cases human resouces (HR) was a support function
+mentioned to work against the agile adoption. In order to gain full results of
+the transformation the entire organization should to be aligned, including HR
+[P38, P42, P49].
+Especially the rewarding practices set by HR were seen as problematic for the
+agile model. Rewarding was tied to personal performance, which was acting
+against team centric thinking and the agile model in general [P3, P6, P49].
+One case reported even on a practice of tracking failures down to individuals
+[P40].
+
 % --
@@
 
-Problems in adopting agile were even showing in functions outside the software
-delivery lifecycle. A support function that was particularly mentioned to
-have issues in agile adoption was human resouces (HR).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-13 journal/article.tex
Work in progress: External QA

@@
 
-Some cases reported  Some organizations used a quality assurance group separate
-from development.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+% When developing large software systems individual teams will not have resources
+% to verify on their own that the results of an iteration pass the QA of the
+% entire system. In these cases separate testing groups will perform heavy testing
+% tasks.
+%  
+% The activities of specialized QA teams still needs to be coordinated with
+% development teams
+
+In an enterprise setting quality assurance (QA) tasks will require effort beyond
+the general agile convention of doing testing within a crossfunctional team. In
+large development efforts the agile model may need to be extended to accomodate
+additional testing [P31]. Organizations may have existing separate QA teams that
+must be coordinated to work with development teams [P17, P30, P48]. Integrating
+enterprise scale QA will pose some additional challenges to agile methods.
+
+Full quality assurance of a system requires special testing such as performance,
+load, and memory testing, but agile methods lacks focus on them [P17]. These
+testing activities can not be done within boundaries of user stories, and
+require more resources than teams can spare [P31]. As testing tasks overlap team
+boundaries it is sensible having a separate teams for the tasks, but
+coordination practice between specialized QA teams and development teams needs
+to be defined [P28, P48].
+
+Some cases reported on problems with lack of automated testing. For a large
+system the lack of automated tests caused excess testing work and late discovery
+of defects [P10, P11, P17].
+
+Still one challenge in QA was the relationship to requirements engineering. The
+QA functions were struggling if there was problems such as ambiguity or
+ineffective breakdown of large requirements [P48]. In a waterfall model an
+extended time was reserved for QA at the end of projects, which allowed to
+retroactively resolve ambiguities in requirements [P5]. However, as agile
+required QA work to be done incrementally with development problems with
+ambiguity in requirements became an impediment to QA [P5].
+
 % ---------------------------------------------------------------------
+
+* There is a separate test team --> Not integrated with development teams
+* Problem: There is a 2 iteration lag for defects found by the test team
+[P30]
+
+* Separate Systems Teams --> System validation
+* Development teams did not know how to create increments that would be testable
+* Connection to requirements engineering
+[P48]
+
+* There was offshore QA resources --> They still needed to be leveraged
+* An arrangement was set up to leverage external QA
+[P17]
+
+* Separate QA groups --> General impediment (not elaborated)
+[P3] (Skip this? too vague)
+
+ENTERPRISE TESTING EXPLAINED
+
+* Enterprise testing --> Falls outside Scrum
+[P31]
+
+* In waterfall there are several test environments the release goes through
+* In agile it is hard to fit going through these environments
+* Hard to fit large scale testing in agile
+* The test environment model had to be revised
+[P29]
+
+% LACK OF AUTOMATION
+% 
+% * Testers were not engaged with the agile process --> For what reason? --> Generally
+% [P10]
+% * Solution: QA responses are handled in the second half of the sprint
+% [P10] (skip this?)
+% * There was a separate QA --> It was a bit difficult to integrate QA with the development teams
+% * QA tasks did not fit inside sprints --> There was lots of legacy code that needed to be regression tested
+% * Legacy code took long to test
+% [P10]
+% 
+% * Lack of automated tests --> Rush in testing
+% * Rushing to release --> Little time to test --> Automated tests skipped
+% [P11]
+% 
+% * Continuous build environment missing --> late discovery of bugs
+% * Late discovery of bugs --> hard to stay releasable
+% * There are iterations that are forced to be only testing and fixing
+% * Invest in test automation
+% [P17]
+
+% CONNECTION TO RE
+% 
+% * Separate Systems Teams --> System validation
+% * Development teams did not know how to create increments that would be testable
+% * Connection to requirements engineering
+% [P48]
+% 
+% * Requirements ambiguity --> Was originally resolved in an extensive QA period in the end of the development cycle
+% * QA was struggling --> Requirements were done badly --> Learning was needed  
+% [P5]
+
+% SPECIALIZED TESTING
+% 
+% * Some testing activities are not focused in agile --> Performance, load, memory
+% * These testing activities must not be forgotten
+% * A separate Integration Test Team --> Works also on these special testing activities
+% [P17]
+% 
+% * Performance, scalability testing --> Require a lot of resources
+% * This enterprise testing must be performed --> Teams can not do it
+% * These testing types need to be centralized --> Specialized environments, specialized tasks
+% * Development teams share the enterprise testing team
+% * Solution: A separate EQC was established offsite
+% [P31]
+% 
+% * External system testing team --> Their work is not needed all the time by teams
+% * They perform work that is necessary for DoD of teams
+% * Problem: How to organize system testing resources so that all teams can be served
+% [P28]
+
+% ---------------------------------------------------------------------
+
+("C: QA is difficult in agile" | "C: Problems with separate non-agile QA")
+
+* Separate QA groups --> General impediment (not elaborated)
+[P3]
+the Amazon PMO may some day become interested in Scrum, but they are enmeshed in
+maintaining and thriving within a traditional waterfall paradigm. Another
+example is the existence of separate QA groups. Executive sponsorship would help
+to begin to break down these entrenched impediments.
+
+* Requirements ambiguity --> Was originally resolved in an extensive QA period in the end of the development cycle
+* QA was struggling --> Requirements were done badly --> Learning was needed  
+[P5]
+There was constant debate over both the meaning of the requirement and whether
+or not the requirement was met.  Quality Assurance engineers were frustrated
+trying to determine whether or not we should accept this feature.  Simply put,
+we had not invested enough of our velocity in either up front or ongoing
+requirements elaboration.
+Going into the second release, we knew we had to do a better job with managing,
+elaborating, and prioritizing requirements.  This had strictly been an issue
+regarding where we chose to invest our energy.
+Properly managing requirements was critical.
+
+* Testers were not engaged with the agile process --> For what reason? --> Generally
+[P10]
+INTEGRATING  QA: We noticed that many testing actions were inevitably deferred
+until the end of the sprint or remained incomplete, and that the testers seemed
+less engaged with the process.
+Comment: This does not say that the testers are separate from the team...
+
+* Solution: QA responses are handled in the second half of the sprint
+[P10] (skip this?)
+How could we adapt Scrum – or otherwise change our own behaviors – in order to
+integrate QA, manage dependencies, focus our daily meetings and release software
+at the end of sprints of a reasonable duration? Inspired by ideas encountered at
+an industry conference (SD Best Practices), the team decided to extend its scrum
+approach in two simple ways.
+Four weeks still felt like the shortest sprint duration that made sense if our
+aim was to deliver potentially shippable releases at the end of each sprint. 
+However, the addition of two-week TASK CYCLES [1] provided a way to structure
+our time within a sprint to good effect (as shown in Figure 5).  During the
+first task-cycle within a sprint the developers would focus on implementation of
+backlog items that we expected to appear in the sprint release.  These changes
+would be checked into our MAIN  code branch. During the second task cycle within
+the sprint the developers would divide their time between resolving issues
+reported by QA while testing those backlog items (as a priority), and continuing
+to work on backlog items that we expected to appear in the following sprint
+release.
+
+* There was a separate QA --> It was a bit difficult to integrate QA with the development teams
+* QA tasks did not fit inside sprints --> There was lots of legacy code that needed to be regression tested
+* Legacy code took long to test
+[P10]
+Indeed, each step forward was accompanied by the discovery of additional
+challenges (the next layer  of the  onion).  Several new problems now came to
+the fore:
+(1) REGRESSION DEFECIT: despite tighter integration of the  QA  team,  it
+quickly became clear that our large legacy code base made necessary a  much
+larger amount of testing than  we  could  easily accommodate  within a sprint,
+or even a multisprint release, even when  developers  dropped everything to join
+the testing effort  as  a  release approached.
+
+* Lack of automated tests --> Rush in testing
+* Rushing to release --> Little time to test --> Automated tests skipped
+[P11]
+Scrum preaches test driven design and releasable product every two to four
+weeks.  For some of our newer products that had been rushed to market, there
+were few tests and, due to restructuring, no test teams.  Fortunately, engineers
+were used to testing each others’ code within the team.  But there were no
+automated builds, and no automated tests, and some test cycles took an entire
+week or more.  Since multiple teams were in the same situation, it was easier to
+get buy in to launch a Scrum project focused on creating an organization-wide
+structure for automating builds and testing that all Cloud Engineering teams
+could use.
+
+* There was offshore QA resources --> They still needed to be leveraged
+* An arrangement was set up to leverage external QA
+[P17]
+BMC needed to continue leveraging QA resources in India while adopting agile
+methodologies. Their solution was to create blended QA teams that matched one
+Lead QA engineer co-located with the US-based development organization with
+several QA engineers located in India. The co-located Lead QA engineer provides
+the necessary coordination and communication at daily US stand-ups, and then
+communicates back to the India-based QA resources on progress and priorities.
+The result is that BMC is able to gain the benefits of a round-the-clock QA
+cycle.
+
+* Continuous build environment missing --> late discovery of bugs
+* Late discovery of bugs --> hard to stay releasable
+* There are iterations that are forced to be only testing and fixing
+* Invest in test automation
+[P17]
+One of the most difficult challenges is staying releasable throughout the
+development cycle. There are several causes – including late defect discovery
+and the deferral of major bugs due in part to lack of a continuous build
+environment. In turn, these issues cause the need for multiple hardening
+iterations spent on pure test and fix activities. Going forward, BMC will
+address these issues with two initiatives:
+1. Maintaining and improving the continuous integration environment.
+2. A focus on pulling testing forward in the cycle using test-driven development
+   (TDD) practices at both the unit test and acceptance test levels, and
+   increasing the investment in test automation.
+
+* Some testing activities are not focused in agile --> Performance, load, memory
+* These testing activities must not be forgotten
+* A separate Integration Test Team --> Works also on these special testing activities
+[P17]
+It is easy to overlook standard engineering best practices when focusing on the
+new planning, development and testing behaviors of the agile paradigm. Sometimes
+standard BMC practices like performance testing, load testing and monitory
+memory usage weren’t executed until late in the development cycle. Pulling these
+activities forward in the release cycle and maintaining an integration test bed
+is the responsibility of a newly-formed Integration Test Team that insures that
+cross-team functionality is performing properly throughout the release cycle.
+
+* External system testing team --> Their work is not needed all the time by teams
+* They perform work that is necessary for DoD of teams
+* Problem: How to organize system testing resources so that all teams can be served
+[P28]
+Both the System Testing team and the Product Metrics team have implemented OH.
+Both of these teams perform work that is vital to the Technology team’s
+definition of “done” for a sprint; therefore, it is important that scrum teams
+have access to these resources. However, not all scrum teams need a dedicated
+system testing engineer, and no scrum teams need a dedicated product metrics
+analyst. Therefore, these two functional teams are ideal candidates for OH.
+
+The System Testing team has implemented OH for scrum teams to utilize the
+expertise of the system testing engineers as well as the system testing
+environments. Like the DS team, the System Testing team does not have enough
+staff to meet the demands for its services. In addition, many of the tasks that
+the system testing engineers do can be performed by other scrum team members.
+Therefore, the System Testing team has instituted OH similar to the UE OH.
+During the System Testing OH, system testing engineers work with scrum team
+members on:
+
+• Training scrum team members on how to use system testing environments and
+  tools.
+• Reviewing system testing results.
+• Validating features that have system testing requirements for scrum teams
+  without an assigned system testing engineer.
+
+The Product Metrics team has implemented OH in a similar way. During Product
+Metrics OH, the product metrics analysts help scrum team members with:
+
+• Defining what types of metrics are relevant to track the adoption of their features.
+• Using and interpreting the feature adoption metrics that are in the metrics database.
+
+* In waterfall there are several test environments the release goes through
+* In agile it is hard to fit going through these environments
+* Hard to fit large scale testing in agile
+* The test environment model had to be revised
+[P29]
+In a waterfall process, code is iterated and each build is deployed to a
+development unit test (DUT) environment.  This environment is primarily used by
+developers to test their own code.
+When the code has been determined as stabilized, a build is deployed into
+another environment called the development integration test (DIT) environment. 
+This is the environment where component testing often begins with build
+verification and unit tests being run.
+When a build it deemed “ready” it is iterated and deployed into a system
+integration test environment (SIT) where end-to-end and system testing such as
+performance is done.
+The last stop for a build is the user acceptance test environment (UAT) where
+user and other types of testing is done.
+Every major build from release candidates to the final build will eventually
+have gone through each environment.  This model while effective in a waterfall
+method does not adapt easily to the Scrum Model.
+
+* There is a separate test team --> Not integrated with development teams
+* Problem: There is a 2 iteration lag for defects found by the test team
+[P30]
+The software testing team tried many ways to coordinate with development teams
+over the first few sprints, but ultimately decided to run its own Scrum. They
+received work items from development teams when work items were “completed.” The
+result: a two-week iteration lag between the discovery of a defect and the time
+it could be fixed. Defects were queued to enter the development team’s next
+sprint. Another iteration lapsed before verification of the fix could occur.
+Because of the high cost of fixing defects, triage on defects was conducted by
+product owners and testers who chose not to fix some defects in exchange for new
+features.
+
+* Enterprise testing --> Falls outside Scrum
+[P31]
+we discovered some areas that would require us to adapt our chosen Scrum
+methodology to account for the realities of complex, large-scale development. 
+Two key activities fell outside the scope of our original Scrum framework
+1. Informing how things get into the backlog, and 2. Enterprise testing.
+
+* Performance, scalability testing --> Require a lot of resources
+* This enterprise testing must be performed --> Teams can not do it
+* These testing types need to be centralized --> Specialized environments, specialized tasks
+* Development teams share the enterprise testing team
+* Solution: A separate EQC was established offsite
+[P31]
+we realized that we were relying on our partnership with a very large customer
+to test the scalability of one of our products.  We would “drop” new code into
+their environment after each sprint, and they would let us know where we had
+issues. We knew that as we scaled Agile practices to more of the organization,
+using our teams and customers as test labs would lead to serious problems. So we
+decided to establish a formalized enterprise testing process. According to Scrum
+practices, all testing, including enterprise testing, should be done within the
+Scrum team, and a user story is only complete once it is fully tested. Certain
+types of enterprise testing pose challenges to this principle. Performance
+requirements, for instance, are often based on scenarios that transcend
+individual user stories. To incorporate performance, scalability and integration
+testing, we would have to adapt our Scrum practices.
+
+A few years ago, Borland established an Enterprise Quality Center (EQC) in
+Singapore.  This test lab is our central resource for all scalability,
+performance and integration testing.
+
+While our solution involves teams being dependent on an external organization
+and requires us to have two versions of “complete” for a user story, the benefit
+is that the EQC has become a virtual resource that we are able to provision as
+needed to the various Scrum teams. While this may be a controversial move, it is
+absolutely necessary.  We simply cannot release products that have not been
+fully enterprise tested.
+
+* Separate Systems Teams --> System validation
+* Development teams did not know how to create increments that would be testable
+* Connection to requirements engineering
+[P48]
+• We agreed to make the Systems Teams (system validation and trials) the
+pacemaker – the metronome - of the new agile framework, and it turned out that
+moving to time-boxed and phased execution of their activities was actually a
+welcome change for many. Instead of random delivery of products to the test and
+trial phases, the model these groups developed had a schedule of acceptance
+dates, which provided the Systems Teams much more flexibility. It also gave the
+teams opportunities for early exposure of defects and integration problems.
+
+• Unfortunately, the Product Teams were not able to deliver software components
+that could be tested or trialed early. So in effect, Systems Teams were “all
+dressed up with no place to go…”. It seems that the Scrum teams didn’t know how
+to break down their work into consumable pieces that could logically be
+delivered, tested and trialed incrementally.  Primarily this was because our
+portfolio and roadmap processes were at “too high a level”.
 

### 2014-05-13 journal/article.tex
Problems with enterprise or external QA

--- 2014-05-12 journal/article.tex
@@
 
+% Problems with enterprise or external QA
+
+In an enterprise setting quality assurance (QA) tasks will require effort beyond
+the general agile convention of doing testing within a crossfunctional team. In
+large development efforts the agile model may need to be extended to accomodate
+additional testing [P31]. Organizations may have existing separate QA teams that
+must be coordinated to work with development teams [P17, P30, P48]. Integrating
+enterprise scale QA will pose some additional challenges to agile methods.
+
+Full quality assurance of a system requires special testing such as performance,
+load, and memory testing, but agile methods lacks focus on them [P17]. These
+testing activities can not be done within boundaries of user stories, and
+require more resources than teams can spare [P31]. As testing tasks overlap team
+boundaries it is sensible having a separate teams for the tasks, but
+coordination practice between specialized QA teams and development teams needs
+to be defined [P28, P48].
+
+Some cases reported on problems with lack of automated testing. For a large
+system the lack of automated tests caused excess testing work and late discovery
+of defects [P10, P11, P17].
+
+Still one challenge in QA was the relationship to requirements engineering. The
+QA functions were struggling if there was problems such as ambiguity or
+ineffective breakdown of large requirements [P48]. In a waterfall model an
+extended time was reserved for QA at the end of projects, which allowed to
+retroactively resolve ambiguities in requirements [P5]. However, as agile
+required QA work to be done incrementally with development problems with
+ambiguity in requirements became an impediment to QA [P5].
+
 % Problems with HR
@@
 
-% --
-
-Some cases reported  Some organizations used a quality assurance group separate
-from development.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-13 journal/article.tex
Work in progress

@@
 
+"C: Difficulties in long term planning"
+
+[P5]
+Even though we believed we had built a reasonable release plan, the team lacked
+confidence in our ability to achieve the release objectives.
+As a result, we met with the business owners to inform them  they  needed to
+either remove the constraint of minimal credible requirements or change the
+target date.  In negotiating with the business owners, we were able to shift the
+date by adding two additional iterations to the release.
+
+Comment:
+There had been a demand for a particular feature set for the release, set by business owners
+
+[P6]
+removing governance gates where processes were overly bureaucratic, and changing
+the way we conducted resource planning and portfolio management.
+
+Comment:
+Talking about removing "broader systemic impediments" with pilot teams
+
+[P9]
+While the significance of this success was widely profiled within the company,
+there were still issues to resolve.  The team had increased their adeptness at
+short term execution, but there were big gaps in long term planning. This gap
+was later addressed by applying a more advanced Scrum technique of 116multilevel
+planning [7]. Focus group lunches were initiated for the business unit’s Scrum
+Masters to share stories and issues.
+
+Comment:
+The success of one particular team's transformation
+
+[P10]
+The  need  for  improved release planning was also behind the second innovation
+made at this time arising from ideas encountered at the conference.  The  team
+had been struggling to balance the need for longer term planning (multiple
+product release over a 3-18 month period) with the scrum discipline of sprint
+planning (a tight 2-4 week focus) – but without re-introducing  a static
+schedule-driven approach that would undermine the agile dynamic.
+
+[P10]
+VELOCITY  VACUUM:  Despite best intentions, establishing a measurable team
+velocity over time remained a huge challenge,  undermining  our efforts at
+release planning.  At the sprint level, the team struggled to embrace the idea
+of recording actual  effort  remaining rather than simply proportionally burning
+down from their original task estimates.  Shifting priorities, fractional  or
+virtual  participation on multiple sprint teams, quirky tools, and the 
+management of ‘off-sprint’ tasks all made it very difficult to perceive actual
+team velocity. When it came to release planning, the team was naturally hesitant
+to estimate against product backlog items when it had such a limited grasp of
+actual performance at the task level.
+
+[P13]
+Part of the release framework included the creation of a few key milestones.  
+In our self-organizing, agile environment we were very sensitive to imposing
+heavy milestones on our teams.  However, we still believed that there were key
+milestones that would align our teams to our on-time, high quality release
+objectives while enhancing the agility of our teams.
+
+[P26]
+One of the management concerns when transitioning to agile was the conflict
+between long-term and short-term planning. One of the threats seen was that
+backlogs and agile planning would give us only short-term visibility.
+We had many experienced program managers, who were able to estimate the size and
+duration of new programs based on their earlier experiences, and we saw no
+reason to leave that experience unutilized. Thus we decided to keep those both:
+long-term vision based on experience, and short-term “strategy” based on
+backlogs. This has resulted in more time spent on actual planning but also in
+improvement in the plans themselves.
+
+[P31]
+We needed to find a way to make the requirements definition and product
+roadmapping process more Agile.  To help facilitate a solution, we brought in
+Enthiosys, a management consulting firm with deep expertise in Agile project
+management. Enthiosys helped us create and implement a system that would bring
+together development, marketing, product management and even representatives
+from sales and support each quarter to scope out a living roadmap of features
+for a product.
+
+[P38]
+Another tension is trying to produce reasonable initial estimates. The estimates
+assist the business side with macrolevel prioritization of which products to
+invest in. One of the elements we worked hard to develop in the teams is
+learning as we build instead of trying to gather all of the requirements and do
+big design up front.
+This is often problematic because the time when the estimates are required is
+also when the teams know the least about the product. They have only begun to
+learn about the solution. Of course the teams do their best to leverage
+similarities with other products, especially as many of them have similar
+elements given the publishing domain, but this does not completely resolve the
+issue and it continues to be an area to improve.
+The transformation has gone a long way toward developing and increasing the
+trust between the business and technology, thus helping the business feel more
+confident in when products will be delivered. For example, one team previously
+went through three rounds of high/macro-level estimation and refinement before
+being given the green light. On the next project the team was able to learn from
+the previous project and what the business was looking for in terms of packaging
+and presentation of the high level estimates. The result was the next project
+required only one round of estimation. Another result was that the project was
+broken up into more phases.
+The first project had been done in two large phases. In the first phase the
+majority of the features were completed and the second was the more or less for
+overflow.
+In the second project, the project was broken up into four phases and the
+business only budgeted out the first two phases.
+The team also focused on a smaller set of requirements and had a clearer idea of
+what was to be built. The second two phases were unfunded and were less defined.
+This allowed the development team to do less guessing while allowing the
+business to risk less by investing. Then after the first release and into the
+second they could refine and plan the other two phases and allow them to
+leverage feedback from the user base to determine whether further investment was
+warranted.
+
 % ---------------------------------------------------------------------

### 2014-05-13 journal/article.tex
Work in progress: Long term planning

@@
 
+% ----
+
+% Long term planning
+
+Requirements work and estimation was problematic as there was a gap between
+short term and long term planning [P9, P10]. Typical agile backlogs would give
+only short term visibility, so organization specific practices for business to
+do long term planning had to be created [P26, P31]. Several cases described that
+care had to be taken to avoid long term planning becoming a prescriptive
+practice. In order to preserve the agility of development a schedule driven
+approach was avoided [P10], sensitivity was used when considering setting
+milestones [P13], and striving for more accurate estimates was not let to become
+an excuse for gathering requirements up front [P38].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+* Business built a release plan --> The team was consulted afterwards and did not believe it was credible
+* The release plan was adjusted
+[P5]
+
+* A reliable record of velocity hampered long term planning
+[P10]
+
+* Teams were better in short term activities
+* Long term planning had gaps
+[P9]
+
+* Conflict between short term and long term planning --> Management had this
+* Backlogs give only short term visibility
+* Experienced managers were capable of estimating size and duration of new programs
+* Both were kept -- backlog for strategical short term
+[P26]
+
+* Product roadmapping had to be made more agile --> Generally
+* Product marketing considered it development's task to figure out ``how''
+* Solution: Collaboration between development and product marketing
+* Iron triangle
+[P31]
+
+CAREFUL OF RE-INTRODUCING A PLAN-DRIVEN APPROACH
+
+* There was a gap between long term planning and working in short sprints
+* A long term planning practice should be established
+* The long term planning should not disrupt the agile dynamics
+* Were careful of not re-introducing schedule driven approach
+[P10]
+
+* Sensitivity in imposing milestones to teams --> Conserving agility
+* It had to be admitted that there are some milestones in projects --> Release objectives: kickoff, feature freeze, release freeze
+[P13]
+
+* Initial estimates for projects were hard
+* Avoid gathering requirements up-front --> Problematic from estimation point of view
+[P38]
+
+GAP BETWEEN SHORT AND LONG TERM
+
+* Teams were better in short term activities
+* Long term planning had gaps
+[P9]
+
+* There was a gap between long term planning and working in short sprints
+* A long term planning practice should be established
+* The long term planning should not disrupt the agile dynamics
+* Were careful of not re-introducing schedule driven approach
+[P10]
+
 % ---------------------------------------------------------------------
@@
 
+* Business built a release plan --> The team was consulted afterwards and did not believe it was credible
+* The release plan was adjusted
 [P5]
@@
 
-[P6]
-removing governance gates where processes were overly bureaucratic, and changing
-the way we conducted resource planning and portfolio management.
-
-Comment:
-Talking about removing "broader systemic impediments" with pilot teams
-
+* Teams were better in short term activities
+* Long term planning had gaps
 [P9]
@@
 short term execution, but there were big gaps in long term planning. This gap
-was later addressed by applying a more advanced Scrum technique of 116multilevel
+was later addressed by applying a more advanced Scrum technique of multilevel
 planning [7]. Focus group lunches were initiated for the business unit’s Scrum
@@
 
+* There was a gap between long term planning and working in short sprints
+* A long term planning practice should be established
+* The long term planning should not disrupt the agile dynamics
+* Were careful of not re-introducing schedule driven approach
 [P10]
@@
 
+* A reliable record of velocity hampered long term planning
 [P10]
@@
 
+* Sensitivity in imposing milestones to teams --> Conserving agility
+* It had to be admitted that there are some milestones in projects --> Release objectives: kickoff, feature freeze, release freeze
 [P13]
@@
 
+* Conflict between short term and long term planning --> Management had this
+* Backlogs give only short term visibility
+* Experienced managers were capable of estimating size and duration of new programs
+* Both were kept -- backlog for strategical short term
 [P26]
@@
 
+* Product roadmapping had to be made more agile --> Generally
+* Product marketing considered it development's task to figure out ``how''
+* Solution: Collaboration between development and product marketing
+* Iron triangle
 [P31]
@@
 
+* Initial estimates for projects were hard
+* Avoid gathering requirements up-front --> Problematic from estimation point of view
 [P38]

### 2014-05-13 journal/article.tex
Long term planning

--- 2014-05-13 journal/article.tex
@@
 
+% Long term planning
+
+High level requirements work and estimation was problematic as there was a gap
+between short and long term planning [P9, P10]. Typical agile backlogs would
+give only short term visibility, so organization specific practices for business
+to do long term planning had to be created [P26, P31]. Several cases described
+that care had to be taken to avoid long term planning becoming a prescriptive
+practice. In order to preserve the agility of development a schedule driven
+approach was avoided [P10], sensitivity was used when considering setting
+milestones [P13], and striving for more accurate estimates was not let to become
+an excuse for gathering requirements up front [P38].
+
 % Problems with enterprise or external QA

### 2014-05-14 journal/article.tex
Editing: Challenges

@@
 
+The challenges presented in this section are factors that primary studies
+described hindering the progress of the transformation.
+Some of the challenges presented here were not explicitly presented as
+transformation challenges, but rather as general difficulties in agile
+imlementation. These issues are still included as creating an agile
+implementation can be seen as a part of the transformation process.
+Some solutions for challenges are also presented, but they should be considered
+as ways of describing the challenges rather than direct success factors.
+
 Challenges can be organized on a high level into three categories. First we
@@
 
+% Change resistance
+% Lack of investment in transformation
+% Agile is found to be hard
+% Large organization issues
+% Coordinating work between many teams
+% Integrating other functions than development to agile
+% Other challenges in transformation (zealots, falling back)
+
 \subsubsection{Change resistance}
@@
 To start a transformation the organization needs an initial push. However,
-people are usually disincluned to change, and it is likely that resistance
-will emerge in any change. 
-% Change resistance is a central challenge in any organizational transformation.
-People will not be willing to change unless there are good reasons, and the
-change is preceived easy enough. It is difficult to attain a buy-in for a
-change, and even organizations that have a flexible culture will face resistance
-[P10]. It should be expected that not everyone will be willing to change, and
-some employees will never adapt to the change [P6]. In one case it was reported
-that objections from team members was a major reason for loss in time and
-productivity [P29].
+people are usually disinclined to change, and it any organizational change
+should anticipate some resistance. People will not be willing to change unless
+there are good reasons, and the change is preceived easy enough. It is difficult
+to attain a buy-in for a change, and even organizations that have a flexible
+culture will face resistance [P10]. It should be expected that not everyone will
+be willing to change, and some employees will never adapt to the change [P6].
+During a transformation period objections to change may become a major reason
+for loss in time and productivity [P29].
 
 On a team level numoerous reasons for change resistance were given. For
-instance, Fecarotta described the organization of Boeing as risk-averse and
-cautious which slows down any change [P14]. The organization was also described
+instance, Fecarotta [P14] described the organization of Boeing as risk-averse
+and cautious which slows down any change. The organization was also described
 having a deeply rooted status-quo because of high employee retention, which
-further hindered change. In some cases it was reported that people worried about
-new roles and responsibilities that agile would possibly bring [P44]. For
+further hindered change [P14]. In some cases it was reported that people worried
+about new roles and responsibilities that agile would possibly bring [P44]. For
 instance, testers were worried that they would need to take on cross functional
@@
 
-In these cases change resistance was mitigated for instance by education [P45],
-and one-to-one discussions [P22]. Change resistance was also softened by using
-first only teams willing to try agile, so that insights could be gathered to
+In these cases change resistance was mitigated for instance by education [P45]
+and one-to-one discussions [P22]. Change resistance was also softened by first
+engaging only teams willing to try agile, so that insights could be gathered to
 ease the continued transformation [P45].
@@
 The way the transformation is initiated will also affect how change resistance
-will show.
-In many cases the initiative to change came from management, but when presented
-wrong the initiative becomes a mandate that pepole are not receptive for. Spayd
-[P49] summarizes this as: ``Organizations do not change merely because the boss
-says so, at least not in the way that is intended''. In that organization the
-introduction of CMM had been carried out consistently by a mandate, but the
-collaborative nature of agile did not mix well with such a mindset. A top down
-mandate may also dilute the understanding of the reasons for starting the
-transformation, and the understanding of agile development overall [P2].
-O'Connor describes how team members became sceptical towards agile when
+will show. In many cases the initiative to change came from management, but when
+presented wrong the initiative becomes a mandate that pepole are not receptive
+for. Spayd [P49] summarizes this as: ``Organizations do not change merely
+because the boss says so, at least not in the way that is intended''. In that
+organization the introduction of CMM had been carried out consistently by a
+mandate, but the collaborative nature of agile did not mix well with such a
+mindset. A top down mandate may also dilute the understanding of the reasons for
+starting the transformation, and the understanding of agile development overall
+[P2]. O'Connor describes how team members became sceptical towards agile when
 implementing mandaetd changes did not bring any visible benefits [P37].
@@
 
-The problem of tension between management and development may also be opposite.
-In cases where the change was emerging bottom up the reluctancy to change was
-not in the team members but management. This caused that significant
-organizational change could not be raised above the team level.
-For instance, Atlas [P3] describes how it would have required executive support
-to integrate the product management office and separate quality assurance groups
-in the agile process. In this case, when managers owning projects with many
-teams were not involved in the transformation, the agile model was restrained
-from spreading beyond team level. Lack of middle management support was also
+The direction in which a tension between management and development is built may
+also be opposite. In cases where the change was emerging bottom up the
+reluctancy to change was not in the team members but management, causing that
+significant organizational change could not be raised above the team level. For
+instance, Atlas [P3] describes how it would have required executive support to
+integrate the product management office and separate quality assurance groups in
+the agile process. In this case, when managers owning projects with many teams
+were not involved in the transformation, the agile model was restrained from
+spreading beyond team level [P3]. Lack of middle management support was also
 seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
@@
 
-\subsubsection{Other organizational issues}
-
-% Lack of plans (a guess -- not included)?
-
-% Agile zealots
-
-Another issue that should be avoided is overenthusiasm towards agile methods.
-In several cases there was mentionings about change leaders or teams becoming
-agile zealots. For instance, Evans describes how some members of the agile
-community were becoming too evangelic, which made people take sides for or
-against agile development [12]. In another case team members started out with an
-overly eager attitude, but when agile did not deliver improvements the
-interest faded and started to go back to the old process [P20]. Attempting to
-implement agile too strictly may cause conflicts, and especially in a large
-organization the change leaders need to maintain a collaborative attitude
-towards various groups in the surrounding organization [P6]. Introducing agile
-methods does not guarantee success, and therefore they should not be followed
-blindly [P4].
-
-% People do not speak their minds
-
-In one case a passive-aggressive organization culture was described as a
-challenge. Ranganath describes how people did not speak their minds in meetings,
-thus creating a false sense of agreement when talking about how the organization
-should develop. Because of this resistance and concerns were left growing, and
-acted against achieving a successful transformation. To to avoid these issues
-change leaders worked with team members on an individual level to win confidence 
-and make their voices heard [P40].
+\subsubsection{Lack of investment in transformation}
+
+It is clear that changing the way of working will create some costs for the
+organization. If there is no intention to allow resources for the transformation
+problems will ensue. Typical examples on this was lack of investment in
+training, and that people were kept allocated to tasks from the original
+arrangement. In some cases the transformation was impeded when work spaces could
+not be updated.
+
+% Lack of training and coaching
+
+Lack of resources allocated for supporting a transformation was an evident
+problem. Resources were most often short for training and coaching, but
+underbudgeting these activities may be a serious mistake.
+Hajjdiab [P20] describes how reluctance of management to invest in training
+caused teams to be ill prepared for the agile transformation effort, and
+eventually resulted in ending the use of agile methods. As a less dramatic
+outcome of shortened training was lower motivation [P45].
+
+Proper coaching was a particular problem for very large organizations. In cases
+where the number of teams in need of coaching was large the demand often
+exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
+was also attributed as a reason why the success of pilot teams could not be
+repeated when agile was adopted more widely [P9, P45].
+Silva and Doss [P47] described that when it was hard to fill the coaching
+positions people less seasoned in agile were appointed as coaches, which risked
+that agile practices would not be taught correctly.
+
+% Add quote?
+% * No coaching --> Applying techniques improperly --> Damage
+% [P9]
+% Without proper coaching however, the success of these informal teams met with
+% mixed results over the longer term.  In retrospect, many of these teams did
+% damage to themselves by improperly applying Agile techniques, particularly when
+% command and control mechanisms were still in place.
 
@@
 
-% Lack of training and coaching
-
-Lack of resources allocated for supporting a transformation was an evident
-problem. Resources were most often short for training and coaching, but
-underbudgeting these activities may be a serious mistake.
-Hajjdiab [P20] describes how reluctance of management to invest in training
-caused teams to be ill prepared for the agile transformation effort, and
-eventually resulted in ending the use of agile methods. As a less dramatic
-outcome of shortened training was lower motivation [P45].
-
-Proper coaching was a particular problem for very large organizations. In cases
-where the number of teams in need of coaching was large the demand often
-exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
-was also attributed as a reason why the success of pilot teams could not be
-repeated when agile was adopted more widely [P9, P45].
-Silva and Doss [P47] described that when it was hard to fill the coaching
-positions people less seasoned in agile were appointed as coaches, which risked
-that agile practices would not be taught correctly.
-
-% Add quote?
-% * No coaching --> Applying techniques improperly --> Damage
-% [P9]
-% Without proper coaching however, the success of these informal teams met with
-% mixed results over the longer term.  In retrospect, many of these teams did
-% damage to themselves by improperly applying Agile techniques, particularly when
-% command and control mechanisms were still in place.
-
-\subsubsection{Falling back to old model}
-
-In several cases challenges in the transformation resulted in people falling
-back to the old model of working.
-In some cases it was only a struggle to learn agile practices, but in other
-cases the old model displaced agile.
-Development work has to go on during the transformation but there will be new
-things to learn for the team. Stress caused by the combination of schedule
-pressure and much change at once will pull people back to the old way of working
-[P11, P18, P38].
-Schatz and Abdelshafi tell how even subtle trouble may risk the transformation,
-as people will always look for reasons to revert to familiar behavior [P44].
-Teams without adequate training were struggling with applying agile practices
-correctly, and the challenge the new practices posed made people to abandon them
-and return to the ways they know [P6, P49].
-
-In some cases agile was seen as an overhead, and was therefore abandoned. For
-instance, as new practices were being introduced there was a decrease
-performance, and when teams realized that the benefits are not immediate they
-started reverting to the old model [P21].
-
-When a release schedule with fixed deadlines that business required was combined
-to agile managers started to behave authoritatively and the self organization of
-agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)
-
-In one case new senior appointments were changed the management to be less
-favorable towards agile, and a waterfall development model started to return
-[P12].
-
-% Include quotation?
-% * Some teams mirrored waterfall approaches in their behavior --> Did not work
-% [P32]
-% Then comes execution and the learning really begins. Teams’ working habits
-% determine their success.  Often teams quickly ignore plans in favor of using
-% tactical approaches that mirror waterfall thinking. This contrasts nicely with
-% those more attuned to applying agile.
-
 \subsubsection{Agile is found to be hard}
@@
 time comes to apply agile techniques in practice the team may get lost [P21].
-Agile was also seen as a solution to all problems [P9], success was declared
-prematurely [P49], and expectations created by successful early experiments were
-not fullfilled [P38]. Disapointment was granted when these hopes were not
-fullfilled and implementing an agile model was found to be hard and time
-consuming. The difficulty in implementing agile is evident in misconceptions of
-the agile model, when by-the-book implementations prove to be insufficient, and
-when the agile methods were applied badly to the organization.
+Agile was also thought being a solution to all problems [P9], success was
+declared prematurely [P49], and expectations created by successful early
+experiments were not fullfilled [P38]. Disapointment was granted when these
+hopes were not fullfilled, but instead implementing the agile model was found
+to be hard and time consuming. The difficulty in implementing agile is evident
+in misconceptions of the agile model, when by-the-book implementations prove to
+be insufficient, and when the agile methods were applied badly to the
+organization.
 
@@
 Another case concluded that all practices prescribd in XP did not fit enterprise
-scale development, and some practices required customization [P49]. It would not
-be good to prescribe practices with a check list as it would impose a single
+scale development, and some practices required customization [P49]. The model of
+agile should not be prescribed as a check list, as it would impose a single
 model in an organization with several individual teams and projects [P41].
 
+% Check this paragraph
 Further indications were given that the advice given in literature will not
@@
 Lewis and Neher [P29] describe how there was temptation to strip some practices
-and enhance others. Previous attempts had proven that one of the reasons early
-attempts of agile had failed was deviations from the process, because of which
-the agile mindset did not take root [P29]. In a bad implementation customization
-may lead to the team taking only practices that reflect their current needs, and
-therefore failing to achieve a real change in process and mindset [P9, P46].
+and enhance others. Previous attempts had proven that one of the reasons for
+agile implementations to fail was deviations from the process, because of which
+the agile mindset did not take root [P29]. A bad customization may lead to the
+team taking only practices that reflect their current needs, and therefore
+failing to achieve a real change in process and mindset [P9, P46].
 Another example of bad customization that counteracts change is replacing agile
-vocabulary in favor of familiar terminology from the old model, because
-introducing new language will also create new ways of thinking [P44].
-
-\subsubsection{Large organization issues}
-
-% The software produced in a large organization may have size and complexity that
-% agile struggles to handle.
-
-The primary studies revealed challenges on two areas which relate particularly
-to large scale. First, a large software development endeavour must handle
-integrating the work products of many teams. Second, a large organization will
-have management challenges not seen on small scale.
-
-% Coordinating between many teams
-
-%% ?? -->
-% Include this ???
-% Large organization put agile methods under the test of scaling up, and problems
-% were encountered. One case reported use of agile across hundreds of teams, but
-% management could not get a roll up of the progress without having strict rules
-% on how status was reported [P48]. Schnitter and Mackert [P45] report that a
-% maximum workable size in a project managed by Scrum was 130 people including 7
-% development teams.
+vocabulary in favor of familiar terminology from the old model [P44]. A benefit
+of introducing new vocabulary is that it may stimulate new ways of thinking
+[P44].
+
+\subsubsection{Coordinating work between many teams}
+
+In large software projects work must inevitably be distributed between many
+teams. One of the most prominent transformation challenges was the difficulty in
+creating a model to coordinate work between many teams.
+% --> Present the different aspects of this topic?
 
@@
 
-In the agile model it was hard for managers to get an overview of the progress
-of the release [P17]. The status of features was hard to tell because different
-aspects of features were sepatated across teams [P17]. Smits and Rilliet [P48]
-report that providing executive overview of the development programs was limited
-by the tools used to manage agile projects.
+% Independence of teams
 
@@
 autonomously, which is well in line with agile principles. However, a number of
-issues arose from independence. For instance, teams had to balance between the
-team's own goals and broader goals of the entire organization, and some teams
-chose focus on their own goal [P7]. Coordination was obstructed by independent
-teams that did not respect the larger context they were working in [P13].
-Coordination was also difficult when different teams were working for different
-customers but the applications being built were interdependent [P29]. One case
-reported that allowing teams to have different sprint durations created delays
-in delivery [P51].
+issues arose from independence. For instance, teams needed to find a balance
+between the their own goals and broader goals of the entire organization, but
+some teams chose to focus only on their own goal [P7]. Independent teams that
+did not respect the larger context they were working in obstructed coordination
+[P13]. Coordination was also difficult when teams were working independently for
+different customers but the applications being built were interdependent [P29].
+One case reported that allowing teams to have different sprint durations created
+delays in delivery [P51].
+
+Further coordination problems were encountered when scaling up agile over many
+sites. Operating in a distributed environment makes communication more
+difficult, and causes problems when frequent and effective communication needs
+to be achieved. Distribution had negative effects, such as missing kick-off
+meetings, reduced feeling of proximity when telecommunication is necessary, and
+difficulty in arranging frequent meetings due to time zone differences [P45].
+Agile project management was also problematic in a distributed setting. In a
+waterfall model separate parts of projects could be isolated to different sites,
+but the agile model does not allow such strict slicing of projects [P29]. In one
+case it was simply admitted that a distributed organization will impose
+additional burden on communication and require additional care in the process,
+but distribution and agile would still be used together [P48].
 
@@
 and lack of standardized build scripts was mentioned as a problem in one case
-[P27].
-There was also problems in synchronizing definitions of software interfaces
-between teams, and dependencies in code caused problems in larger features
-[P26].
-One case reported that team centricity in the agile model created a fragile
-architecture, divergence in coding style, and even distrust between teams [P24].
-
-Further problems were encountered when scaling up agile over many sites.
-Operating in a distributed environment makes communication more difficult, which
-causes problems with agile when frequent and effective communication needs to be
-achieved. Distribution had negative effects, such as missing kick-off meetings,
-reduced feeling of proximity when telecommunication is necessary, and difficulty
-in arranging frequent meetings due to time zone differences [P45].
-Agile project management was also problematic in a distributed setting. In a
-waterfall model separate parts of projects could be isolated to different sites,
-but the agile model does not allow such strict slicing of projects [P29].
-In one case it was simply admitted that a distributed organization will impose
-additional burden on communication and require additional care in the process,
-but distribution and agile would still be used together [P48].
+[P27]. There was also problems in synchronizing definitions of software
+interfaces between teams, and dependencies in code caused problems in larger
+features [P26]. One case reported that team centricity in the agile model
+created a fragile architecture, divergence in coding style, and even distrust
+between teams [P24].
 
@@
 communication channels narrowed and created communication breakdowns. Problems
-with collaboration, dependencies, and integration were solved with presnce of
-individuals with an proactive and open mindset towards working over team
+with collaboration, dependencies, and integration were solved in the presnce of
+individuals having a proactive and open mindset towards working over team
 boundaries. It was concluded that a balance is needed between completing new
@@
 Many studies presented solutions for the inter-team challenges, which also
-highlight the challenges themselves. The general model of solving inter-team
-problems was introducing an additional layer of coordination, such as a
-Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
+highlight the nature of the challenges themselves. The general model of solving
+inter-team problems was introducing an additional layer of coordination, such as
+a Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
 customized release framework [P13]. In one case integration problems were
@@
 impact on the coordination effort [P17].
+
+\subsubsection{Large organization issues}
+
+% The software produced in a large organization may have size and complexity that
+% agile struggles to handle.
+
+The primary studies revealed challenges on two areas which relate particularly
+to large scale. First, a large software development endeavour must handle
+integrating the work products of many teams. Second, a large organization will
+have management challenges not seen on small scale.
+
+%% ?? -->
+% Include this ???
+% Large organization put agile methods under the test of scaling up, and problems
+% were encountered.
+% In the agile model it was hard for managers to get an overview of the progress
+% of the release [P17]. The status of features was hard to tell because different
+% aspects of features were sepatated across teams [P17].
+% One case reported use of agile across hundreds of teams, but management could
+% not get a roll up of the progress without having strict rules on how status was
+% reported [P48]. Schnitter and Mackert [P45] report that a maximum workable size
+% in a project managed by Scrum was 130 people including 7 development teams.
+% % Smits and Rilliet [P48] report that providing executive overview of the
+% % development programs was limited by the tools used to manage agile projects.
 
@@
 levels [P10]. Ongoing projects had been set up with previous development
-methods, and the agile methods had to be arranged to fit with them [P8, P29].
+methods, and the agile methods needed to be arranged to fit with them [P8, P29].
 A particular problem in sychronizing work between projects using different
@@
 % affect productivity.
+
+% This where?
+Some cases reported that development efforts were controlled on a top level by a
+project management office (PMO). The PMO was described as entrenched in a
+waterfall paradigm, and the top-level rigidity was causing friction in the
+agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
+all aspects of projects, and its central structure had to be dismantled for
+the organization to become agile [P46].
 
@@
 
-% --
-
-% This where?
-
-Some cases reported that development efforts were controlled on a top level by a
-project management office (PMO). The PMO was described as entrenched in a
-waterfall paradigm, and the top-level rigidity was causing friction in the
-agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
-all aspects of projects, and its central structure had to be dismantled for
-the organization to become agile [P46].
+\subsubsection{Other challenges in transformation}
+
+% Lack of plans (a guess -- not included)?
+
+% Agile zealots
+
+Another issue that should be avoided is overenthusiasm towards agile methods.
+In several cases there was mentionings about change leaders or teams becoming
+agile zealots. For instance, Evans describes how some members of the agile
+community were becoming too evangelic, which made people take sides for or
+against agile development [12]. In another case team members started out with an
+overly eager attitude, but when agile did not deliver improvements the
+interest faded and started to go back to the old process [P20]. Attempting to
+implement agile too strictly may cause conflicts, and especially in a large
+organization the change leaders need to maintain a collaborative attitude
+towards various groups in the surrounding organization [P6]. Introducing agile
+methods does not guarantee success, and therefore they should not be followed
+blindly [P4].
+
+% People do not speak their minds
+
+In one case a passive-aggressive organization culture was described as a
+challenge. Ranganath describes how people did not speak their minds in meetings,
+thus creating a false sense of agreement when talking about how the organization
+should develop. Because of this resistance and concerns were left growing, and
+acted against achieving a successful transformation. To to avoid these issues
+change leaders worked with team members on an individual level to win confidence 
+and make their voices heard [P40].
+
+% Falling back to old model
+
+In several cases challenges in the transformation resulted in people falling
+back to the old model of working.
+In some cases it was only a struggle to learn agile practices, but in other
+cases the old model displaced agile.
+Development work has to go on during the transformation but there will be new
+things to learn for the team. Stress caused by the combination of schedule
+pressure and much change at once will pull people back to the old way of working
+[P11, P18, P38].
+Schatz and Abdelshafi tell how even subtle trouble may risk the transformation,
+as people will always look for reasons to revert to familiar behavior [P44].
+Teams without adequate training were struggling with applying agile practices
+correctly, and the challenge the new practices posed made people to abandon them
+and return to the ways they know [P6, P49].
+
+In some cases agile was seen as an overhead, and was therefore abandoned. For
+instance, as new practices were being introduced there was a decrease
+performance, and when teams realized that the benefits are not immediate they
+started reverting to the old model [P21].
+
+When a release schedule with fixed deadlines that business required was combined
+to agile managers started to behave authoritatively and the self organization of
+agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)
+
+In one case new senior appointments were changed the management to be less
+favorable towards agile, and a waterfall development model started to return
+[P12].
+
+% Include quotation?
+% * Some teams mirrored waterfall approaches in their behavior --> Did not work
+% [P32]
+% Then comes execution and the learning really begins. Teams’ working habits
+% determine their success.  Often teams quickly ignore plans in favor of using
+% tactical approaches that mirror waterfall thinking. This contrasts nicely with
+% those more attuned to applying agile.
 

### 2014-05-14 journal/article.tex
Large organization issues

@@
 --> Any hints on size?? Maybe size leaning towards discipline?
+
+\subsection{Things that may need to be explained}
+
+* Scrum
+
+* XP
+
+* Definition of sprint
+
+* Self organization and independence of teams -- This is a question relating how
+  to organize management.
 
@@
 
-% The software produced in a large organization may have size and complexity that
-% agile struggles to handle.
-
-The primary studies revealed challenges on two areas which relate particularly
-to large scale. First, a large software development endeavour must handle
-integrating the work products of many teams. Second, a large organization will
-have management challenges not seen on small scale.
-
-%% ?? -->
-% Include this ???
-% Large organization put agile methods under the test of scaling up, and problems
-% were encountered.
-% In the agile model it was hard for managers to get an overview of the progress
-% of the release [P17]. The status of features was hard to tell because different
-% aspects of features were sepatated across teams [P17].
-% One case reported use of agile across hundreds of teams, but management could
-% not get a roll up of the progress without having strict rules on how status was
-% reported [P48]. Schnitter and Mackert [P45] report that a maximum workable size
-% in a project managed by Scrum was 130 people including 7 development teams.
-% % Smits and Rilliet [P48] report that providing executive overview of the
-% % development programs was limited by the tools used to manage agile projects.
-
-% Challenge: Previous organization in silos is haunting new model
-% Is this good??
-
-In some cases problems were caused by internal boundaries or silos in the
-initial organization. Cloke [P10] describes how use of Scrum revealed an
-internal segmentation where teams operated with differing priorities and
-agendas. This segmentation also hampered the initial effort to use Scrum in a
-larger context [P10].
-Projects needed specialized skills that were scarce and people were often
-relocated to match the needs of the moment, but too much reorganization made
-it diffucult for teams to plan ahead [P11].
-A dispersed matrix organization causes also problems when people are on one hand
-kept assigned to many projects simultaneously, but on the other hand functions
-such as database administration, architecture boards, and project management are
-centralized [P14].
-
-% Challenge: Differing vocabularies
-
-In a large environment many agile teams coexist and it is possible that their
-agile models and vocabularies start diverging. When many teams implement agile
-without consistent guidance friction and fragmentation may be generated [P9].
-The organization may require moving people between teams from time to time, and
-therefore it would be desirable that no big differences in teams' agile cultures
-exist. Divergence in process would create risk increasing costs when relocating
-people [P38]. Further, if processes of teams diverged management had
-difficulties in forecasting and comparing work of different teams [P38, P48].
-To overcome problems with divergence in agile models some organizations defined
-standards to follow [P8, P43].
+The primary studies revealed challenges which relate particularly to large
+scale. Additional layers of management are usually necessary to steer large
+groups of people, but agile methods provide little guidance on how to organize
+management. Also, as large organizations accomodate many development teams
+problems with synchronizing the development models of teams may cause problems.
 
@@
 
-A common problem in the agile implementations was the role of managers.
-Especially the role of middle management was seen as unclear in agile
-methods, but a cultural change would still be required on that level [P2].
+The existence of additional management positions may be problematic with agile
+processes as they emphasize self organization.
+The role of middle management was seen as unclear in agile methods [P2]. This is
+problematic as an agile transformation requires cultural change especially on
+the middle management level [P2].
 It was described that people in management positions needed to resist the
-tendency to take command and control to allow room for self-organization, but
+tendency to take command and control to allow room for self organization, but
 the change in role was difficult to achieve for the people involved [P11, P52].
@@
 
-% Roles of managers (continued)
-
 Several other problems related to management roles were also presented. For
 instance, Spayd [P49] describes how managers were left outside the roles offered
-by the new agile model. In another case when Scrum masters were appointed from
-the ranks of managers developers felt being micromanaged because of vague
-understanding of the agile method [P27]. It was also described how mixing the
-role of project manager and Scrum master created a conflict in interest, and
+by the new agile model. In another case when managers were appointed as Scrum
+masters developers felt being micromanaged [P27]. This was partially because of
+vague understanding of the agile method [P27]. It was also described how mixing
+the role of project manager and Scrum master created a conflict in interest, and
 turned the Scrum manager's role into one of policing the team instead of
@@
 with the old methods. The arrangement of using the old and new models side by
-side was seen as problematic [P1, P26], and causing tensions on all organizational
-levels [P10]. Ongoing projects had been set up with previous development
-methods, and the agile methods needed to be arranged to fit with them [P8, P29].
-A particular problem in sychronizing work between projects using different
-models was that in Scrum the software design was fleshed out over time as
-sprints progressed, but the waterfall model would require detailed design to be
-locked early in the project [P29].
+side was seen as problematic [P1, P26], and causing tensions on all
+organizational levels [P10]. Ongoing projects had been set up with previous
+development methods, and the agile methods needed to be arranged to fit with
+them [P8, P29]. A particular problem in collaborating between projects using
+different models was that in Scrum the software design was fleshed out over time
+as sprints progressed, but the waterfall model would require detailed design to
+be locked early in the project [P29].
 
@@
 
+Some cases reported that development efforts were still controlled on a top
+level by a project management office (PMO). The PMO was described as entrenched
+in a waterfall paradigm, and the top-level rigidity was causing friction in the
+agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
+all aspects of projects, and its central structure had to be dismantled for the
+organization to become agile [P46].
+
 There were also problems with duplicating management when two models were in
@@
 
+% Challenge: Differing vocabularies
+
+In addition to teams using agile and waterfall methods side by side it is
+possible in a large organization that the agile interpretations of teams start
+diverging. When many teams implement agile without consistent guidance friction
+and fragmentation may be generated [P9]. The organization may require moving
+people between teams from time to time, and therefore it would be desirable that
+no big differences in teams' agile cultures exist. Divergence in process would
+create risk increasing costs when relocating people [P38].
+Further, if processes of teams diverged management had difficulties in
+forecasting and comparing work of different teams [P38, P48]. To overcome
+problems with divergence in agile models some organizations defined standards to
+follow [P8, P43].
+
+% Challenge: Previous organization in silos is haunting new model
+% Is this good??
+In some cases the initial organization had internal boundaries or specialized
+knowledge was stuck silos. Such settings caused problems in the agile
+implementation.
+Cloke [P10] describes how use of Scrum revealed an internal segmentation where
+teams operated with differing priorities and agendas. This segmentation also
+hampered the initial effort to use Scrum in a larger context [P10].
+Cowan [P11] describes how projects needed specialized skills that were scarce
+and people were often relocated to match the needs of the moment. Too much
+reorganization made it diffucult for teams to plan ahead [P11].
+
+%% This??
+% A dispersed matrix organization causes problems when people are on one hand
+% kept assigned to many projects simultaneously, but on the other hand functions
+% such as database administration, architecture boards, and project management are
+% centralized [P14].
+
+%% ?? -->
+% Large organization put agile methods under the test of scaling up, and problems
+% were encountered.
+% In the agile model it was hard for managers to get an overview of the progress
+% of the release [P17]. The status of features was hard to tell because different
+% aspects of features were sepatated across teams [P17].
+%
+% One case reported use of agile across hundreds of teams, but management could
+% not get a roll up of the progress without having strict rules on how status was
+% reported [P48]. Schnitter and Mackert [P45] report that a maximum workable size
+% in a project managed by Scrum was 130 people including 7 development teams.
+% Smits and Rilliet [P48] report that providing executive overview of the
+% development programs was limited by the tools used to manage agile projects.
+
+% --
+
 % Add quote ?
@@
 % affect productivity.
-
-% This where?
-Some cases reported that development efforts were controlled on a top level by a
-project management office (PMO). The PMO was described as entrenched in a
-waterfall paradigm, and the top-level rigidity was causing friction in the
-agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
-all aspects of projects, and its central structure had to be dismantled for
-the organization to become agile [P46].
 

### 2014-05-14 journal/article.tex
Editing

@@
 % Integrating other functions than development to agile
+% Integrating requirements management and quality assurance
 % Other challenges in transformation (zealots, falling back)
@@
 groups of people, but agile methods provide little guidance on how to organize
-management. Also, as large organizations accomodate many development teams
-problems with synchronizing the development models of teams may cause problems.
+management. Also, as large organizations accomodate many development teams there
+may be problems if the development models between teams are not synchronized.
+Internal fragmenttation is a further problem that may be encountered in large
+organizations.
 
@@
 
-The existence of additional management positions may be problematic with agile
-processes as they emphasize self organization.
+The existence of additional management positions may pose problems to agile
+processes that emphasize self organization.
 The role of middle management was seen as unclear in agile methods [P2]. This is
@@
 fixed feature set to be delivered [P31].
+
+% Problems with HR
+
+Problems in adopting agile were even showing in functions outside the software
+delivery lifecycle. In several cases human resouces (HR) was a support function
+mentioned to work against the agile adoption. In order to gain full results of
+the transformation the entire organization should to be aligned, including HR
+[P38, P42, P49].
+Especially the rewarding practices set by HR were seen as problematic for the
+agile model. Rewarding was tied to personal performance, which was acting
+against team centric thinking and the agile model in general [P3, P6, P49].
+One case reported even on a practice of tracking failures down to individuals
+[P40].
+
+\subsubsection{Integrating requirements management and quality assurance}
+
+Traditionally development is preceeded by requirements engineering and succeeded
+by quality assurance (QA) activities. Agile methods suggest that essential parts
+of these activities would be performed within the development team. The effort
+required for managing requirements and QA for large systems may be too large for
+individual teams to handle.
 
@@
 
-In an enterprise setting quality assurance (QA) tasks will require effort beyond
+In an enterprise setting quality assurance tasks will require effort beyond
 the general agile convention of doing testing within a crossfunctional team. In
@@
 
-% Problems with HR
-
-Problems in adopting agile were even showing in functions outside the software
-delivery lifecycle. In several cases human resouces (HR) was a support function
-mentioned to work against the agile adoption. In order to gain full results of
-the transformation the entire organization should to be aligned, including HR
-[P38, P42, P49].
-Especially the rewarding practices set by HR were seen as problematic for the
-agile model. Rewarding was tied to personal performance, which was acting
-against team centric thinking and the agile model in general [P3, P6, P49].
-One case reported even on a practice of tracking failures down to individuals
-[P40].
-
 \subsubsection{Other challenges in transformation}
 
-% Lack of plans (a guess -- not included)?
-
-% Agile zealots
-
-Another issue that should be avoided is overenthusiasm towards agile methods.
-In several cases there was mentionings about change leaders or teams becoming
-agile zealots. For instance, Evans describes how some members of the agile
-community were becoming too evangelic, which made people take sides for or
-against agile development [12]. In another case team members started out with an
-overly eager attitude, but when agile did not deliver improvements the
-interest faded and started to go back to the old process [P20]. Attempting to
-implement agile too strictly may cause conflicts, and especially in a large
-organization the change leaders need to maintain a collaborative attitude
-towards various groups in the surrounding organization [P6]. Introducing agile
-methods does not guarantee success, and therefore they should not be followed
-blindly [P4].
-
-% People do not speak their minds
-
-In one case a passive-aggressive organization culture was described as a
-challenge. Ranganath describes how people did not speak their minds in meetings,
-thus creating a false sense of agreement when talking about how the organization
-should develop. Because of this resistance and concerns were left growing, and
-acted against achieving a successful transformation. To to avoid these issues
-change leaders worked with team members on an individual level to win confidence 
-and make their voices heard [P40].
+There were still some other challenges in the transformation that do not fit in
+the categories presented above. Two particular problems are worth mentioning.
+Firstly, there were certain factors that pulled people back to the old ways of
+working, and secondly, excessive enthusiasm in agile had negative effects.
 
@@
 In several cases challenges in the transformation resulted in people falling
-back to the old model of working.
-In some cases it was only a struggle to learn agile practices, but in other
-cases the old model displaced agile.
+back to the old model of working. In some cases it was only a temporary struggle
+while learning agile practices, but in other cases the old model displaced
+agile.
 Development work has to go on during the transformation but there will be new
@@
 started reverting to the old model [P21].
-
-When a release schedule with fixed deadlines that business required was combined
-to agile managers started to behave authoritatively and the self organization of
-agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)
-
-In one case new senior appointments were changed the management to be less
-favorable towards agile, and a waterfall development model started to return
-[P12].
-
-% Include quotation?
-% * Some teams mirrored waterfall approaches in their behavior --> Did not work
-% [P32]
-% Then comes execution and the learning really begins. Teams’ working habits
-% determine their success.  Often teams quickly ignore plans in favor of using
-% tactical approaches that mirror waterfall thinking. This contrasts nicely with
-% those more attuned to applying agile.
+Evans [P12] describes that new senior appointments changed the management to be
+less favorable towards agile, and a waterfall development model started to
+return.
+
+% Agile zealots
+
+A phenomena that troubled some organizations was overenthusiasm towards agile
+methods. In several cases there was mentionings about change leaders or teams
+becoming agile zealots. For instance, Evans [12] describes how some members of
+the agile community were becoming too evangelic, which made people take sides
+for or against agile development. In another case team members started out with
+an overly eager attitude, but when agile did not deliver improvements the
+interest faded and started to go back to the old process [P20]. Attempting to
+implement agile too strictly may cause conflicts, and especially in a large
+organization the change leaders need to maintain a collaborative attitude
+towards various groups in the surrounding organization [P6]. Introducing agile
+methods does not guarantee success, and therefore they should not be followed
+blindly [P4].
+
+% When a release schedule with fixed deadlines that business required was combined
+% to agile managers started to behave authoritatively and the self organization of
+% agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)
+
+% This??
+% % People do not speak their minds
+% In one case a passive-aggressive organization culture was described as a
+% challenge. Ranganath describes how people did not speak their minds in meetings,
+% thus creating a false sense of agreement when talking about how the organization
+% should develop. Because of this resistance and concerns were left growing, and
+% acted against achieving a successful transformation. To to avoid these issues
+% change leaders worked with team members on an individual level to win confidence 
+% and make their voices heard [P40].
 

### 2014-05-15 journal/article.tex
Editing: Challenges in other functions

@@
 
-Agile methods are typically understood through a set of activities closely
-related to software development. Cases describe how development teams were
-typically the first entities in the organization that were considered when
-introducing agile [P2, P17, P31, P38]. However, development operates in a larger
-context, and will need to interface with other organizational functions.
+Agile methods are typically considered as a set of activities closely related to
+software development. Cases describe how development teams were the first
+entities in the organization that were considered when introducing agile [P2,
+P17, P31, P38]. However, development operates in a larger context and will need
+to interface with other organizational functions.
 Challenges were faced when exposing other parts of the organization to the agile
-mode, and functions that were further away from development were resistant to
+model, and functions that were further away from development were resistant to
 change [P2, P18, P31]. A tension was growing between development and broader
@@
 Primary sources listed various functions beyond development that are involved in
-bringing products to the market. Marketing was a particularly frequently
-mentioned group [P5, P17, P28, P31, P38, P45, P48]. Other groups involved in the
-whole of product delivery were sales [P17, P31], infrastructure and operations
-[P10, P45, P46, P50], user experience and design [P9, P10, P28], documentation
-[P28], legal [P10], security [P10], customer services [P48], and finance [P48].
-All of these groups were reported to be affected on some level by the agile
-transformation.
+bringing products to the market. Functions interfacing with development and
+affected by the agile transformation are presented in figure
+\ref{fig:challenges_other_functions}.
+Marketing was a particularly frequently mentioned group [P5, P17, P28, P31, P38,
+P45, P48]. Other groups involved in the whole of product delivery were sales
+[P17, P31], infrastructure and operations [P10, P45, P46, P50], user experience
+and design [P9, P10, P28], documentation [P28], legal [P10], security [P10],
+customer services [P48], and finance [P48].
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{graphics/challenges_other_functions.pdf}
+    \caption{Functions affected by the agile transformation}
+    \label{fig:challenges_other_functions}
+  \end{center}
+\end{figure}
 

### 2014-05-15 journal/article.tex
Editing

@@
   to organize management.
+
+* Scrum master must not command and control -- Scrum master as manager??
 
@@
 
+Figure \ref{fig:transformation_time} summarizes the publication years of the
+primary studies, and the years when the transformations were reported to have
+started. All studies and transformations were dated after year 2000. There were
+peaks in transformation studies in years 2008 and 2011. The studies were
+typically published two years after the start of the transformation.
+
+\begin{figure}[t]
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/transformation_time.pdf}
+    \caption{Publication years and reported start years of transformations}
+    \label{fig:transformation_time}
+  \end{center}
+\end{figure}
+
+The size of the organizations varied from the minimum included size of 50 to
+18,000 people. The median size was 300 people. In 7 studies the size was
+presented in terms of teams, ranging from the minimum of 6 teams to 150 teams.
+The median was 10 teams. In 8 studies there was no direct indication on size
+but the issues discussed revealed that the organization was of large size
+accoding to our definition. The distribution of organization sizes is presented
+in figure \ref{fig:organization_size}.
+
 \begin{figure}[b]
@@
 
-The size of the organizations varied from the minimum included size of 50 to
-18,000 people. The median size was 300 people. In 7 studies the size was
-presented in terms of teams, ranging from the minimum of 6 teams to 150 teams.
-The median was 10 teams. In 8 studies there was no direct indication on size
-but the issues discussed revealed that the organization was of large size
-accoding to our definition. The distribution of organization sizes is presented
-in figure \ref{fig:organization_size}.
-
-Figure \ref{fig:transformation_time} summarizes the publication years of the
-primary studies, and the years when the transformations were reported to have
-started. All studies and transformations were dated after year 2000. There were
-peaks in transformation studies in years 2008 and 2011. The studies were
-typically published two years after the start of the transformation.
-
-\begin{figure}[t]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/transformation_time.pdf}
-    \caption{Publication years and reported start years of transformations}
-    \label{fig:transformation_time}
-  \end{center}
-\end{figure}
-
-\begin{table}[bh]
+The business areas of the case organizations are presented in table
+\ref{table:businessareas}. Different business areas were represented somewhat
+evenly. Online services was the largest group, including companies
+providing software as a service solutions for businesses, online media players
+for consumers, online services for consumers, and communication software for
+businesses. The second largest group was telecommunications, including companies
+such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
+largest group was enterprise management solution providers with products for
+business process management, project protfolio management, and facility
+management.
+
+\begin{table}[!t]
     \begin{tabular}{ l r }
@@
 
-The business areas of the case organizations are presented in table
-\ref{table:businessareas}. Different business areas were represented somewhat
-evenly. Online services was the largest group, including companies
-providing software as a service solutions for businesses, online media players
-for consumers, online services for consumers, and communication software for
-businesses. The second largest group was telecommunications, including companies
-such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
-largest group was enterprise management solution providers with products for
-business process management, project protfolio management, and facility
-management.
-
 \subsection{Agile methods used}
@@
 
+\clearpage
+
 \subsection{Challenges in transformation}
@@
 Any organizational transformation involving numerous individuals will face some
-challenges. As stated by Research Question 3, we investigated what are typical
-challenges that large organizations introducing agile methods face.
+challenges. As stated by Research Question 3, we investigated what challenges
+large organizations introducing agile methods may face.
 
@@
 
+\begin{figure}[!b]
+  \begin{center}
+    \includegraphics{graphics/challenges_summary.pdf}
+    \caption{Challenges in agile transformation}
+    \label{fig:challenges_summary}
+  \end{center}
+\end{figure}
+
 Challenges can be organized on a high level into three categories. First we
@@
 
-To start a transformation the organization needs an initial push. However,
-people are usually disinclined to change, and it any organizational change
-should anticipate some resistance. People will not be willing to change unless
+In an organizational transformation working habits and roles need to change.
+However, people are usually disinclined to change and some resistance should be
+anticipated in any change process. People will not be willing to change unless
 there are good reasons, and the change is preceived easy enough. It is difficult
@@
 further hindered change [P14]. In some cases it was reported that people worried
-about new roles and responsibilities that agile would possibly bring [P44]. For
+about new roles and responsibilities that the agile model might bring [P44]. For
 instance, testers were worried that they would need to take on cross functional
@@
 
-The primary studies revealed challenges which relate particularly to large
+The primary studies revealed challenges that relate particularly to large
 scale. Additional layers of management are usually necessary to steer large
 groups of people, but agile methods provide little guidance on how to organize
-management. Also, as large organizations accomodate many development teams there
-may be problems if the development models between teams are not synchronized.
-Internal fragmenttation is a further problem that may be encountered in large
-organizations.
+management. Problems may arise in a large environment if the schedule of agile
+adoption varies from team to team. Internal fragmenttation is a further problem
+that may be encountered in large organizations.
 
@@
 processes that emphasize self organization.
-The role of middle management was seen as unclear in agile methods [P2]. This is
-problematic as an agile transformation requires cultural change especially on
-the middle management level [P2].
-It was described that people in management positions needed to resist the
-tendency to take command and control to allow room for self organization, but
-the change in role was difficult to achieve for the people involved [P11, P52].
+Especially the role of middle management was seen as unclear in agile methods
+[P2]. This is problematic as an agile transformation requires cultural change
+particularly on the middle management level [P2].
+It was described that managers needed to resist the tendency to take command and
+control and allow room for self organization, but the change in role was
+difficult to achieve for the people involved [P11, P52].
 Nielsen and McMunn [P30] describe how the project management group had
@@
 with the old methods. The arrangement of using the old and new models side by
-side was seen as problematic [P1, P26], and causing tensions on all
+side was seen as generally problematic [P1, P26], and causing tensions on all
 organizational levels [P10]. Ongoing projects had been set up with previous
-development methods, and the agile methods needed to be arranged to fit with
-them [P8, P29]. A particular problem in collaborating between projects using
-different models was that in Scrum the software design was fleshed out over time
-as sprints progressed, but the waterfall model would require detailed design to
-be locked early in the project [P29].
-
-Despite the agile introduction there were problems with management working
-according to the old waterfall model. O'Connor [P38] described management as
-``focused on meetings and big up front project plans'' despite agile being in
-use. In another case management was losing confidence in agile because familiar
-reports on cost analysis and percentage complete were not produced [P35]. As
-Scrum teams did not commit to fixed schedules they were considered unreliable
-according to the original model [P45].
+development methods, and the agile methods needed to be arranged to fit along
+with them [P8, P29]. A particular problem in collaborating between projects
+using different models was that in Scrum the software design was fleshed out
+over time as sprints progressed, but the waterfall model would require detailed
+design to be locked early in the project [P29].
+
+Even when agile had been taken into use in development there were problems with
+management working according to the old waterfall model. O'Connor [P38]
+described management as ``focused on meetings and big up front project plans''
+despite agile being in use. In another case management was losing confidence in
+agile because familiar reports on cost analysis and percentage complete were not
+produced [P35]. As Scrum teams did not commit to fixed schedules they were
+considered unreliable according to the original model [P45].
 
@@
 
-In addition to teams using agile and waterfall methods side by side it is
-possible in a large organization that the agile interpretations of teams start
-diverging. When many teams implement agile without consistent guidance friction
-and fragmentation may be generated [P9]. The organization may require moving
-people between teams from time to time, and therefore it would be desirable that
-no big differences in teams' agile cultures exist. Divergence in process would
-create risk increasing costs when relocating people [P38].
+In addition to teams using agile and waterfall models side by side a large
+organization makes it possible that the agile interpretations of teams diverge.
+When many teams implement agile without consistent guidance friction and
+fragmentation may emerge [P9]. The organization may require moving people
+between teams from time to time, and therefore it would be desirable that no big
+differences in teams' agile cultures exist. Divergence in process would create
+risk of increased costs when relocating people [P38].
 Further, if processes of teams diverged management had difficulties in

### 2014-05-16 journal/article.tex
Editing challenges

@@
 * Scrum master must not command and control -- Scrum master as manager??
+
+* People centricity -- agile is not for squeezing efficiency -- agile is not for
+  fewer career opportunities in a flatter organization.
+
+* Agile is a mindset -- it can not be studied from a book -- need of coaching
 
@@
     \includegraphics{graphics/challenges_summary.pdf}
-    \caption{Challenges in agile transformation}
+    \caption{Overview of challenges in agile transformation}
     \label{fig:challenges_summary}
@@
 It is clear that changing the way of working will create some costs for the
-organization. If there is no intention to allow resources for the transformation
-problems will ensue. Typical examples on this was lack of investment in
-training, and that people were kept allocated to tasks from the original
-arrangement. In some cases the transformation was impeded when work spaces could
-not be updated.
+organization. Problems will ensue if there is no intention to allow resources
+for the transformation. Examples on this was lack of investment in training, and
+keeping people allocated to tasks from the old arrangement. In some cases the
+transformation was impeded when work spaces could not be updated.
 
@@
 
-Lack of resources allocated for supporting a transformation was an evident
-problem. Resources were most often short for training and coaching, but
-underbudgeting these activities may be a serious mistake.
+Lack of activities supporting a transformation was an evident problem. Training
+and coaching were often short of resources, but underbudgeting these activities
+may seriously hamper the transformation.
 Hajjdiab [P20] describes how reluctance of management to invest in training
@@
 
-Proper coaching was a particular problem for very large organizations. In cases
-where the number of teams in need of coaching was large the demand often
-exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
-was also attributed as a reason why the success of pilot teams could not be
-repeated when agile was adopted more widely [P9, P45].
+Arranging proper coaching was a particular problem in many organizations. It is
+critical to coach teams in their real work environment as a mindset change is
+difficult to achieve by only attending training sessions. In large organizations
+where numerous teams needed to be coached the demand often exceeded the capacity
+of available coaches [P1, P6, P23, P49]. Lack of coaching was also attributed as
+a reason why the success of pilot teams could not be repeated when agile was
+adopted more widely [P9, P45].
 Silva and Doss [P47] described that when it was hard to fill the coaching
@@
 
-A condition hindering the transformation was that old commitments still needed
-to be tended to even though the organization was being restructured.
+Even though the organization was in a state of transformation the workload of
+teams was not always adjusted to facilitate the process.
 Some organizations started their agile journey in a state where people were
-overcommitted, but it was later realized that overloaded people will not be able
-to learn new ways of working and changing their behavior [P37, P42]. Cowan [P11]
-describes how time was allocated for tending to old commitments. However, even
-when there was preparation for extra work people with specific expertise became
-overloaded. After time it was realized that overcommitment must be reduced by
-allowing to shift excess work for later time [P11]. In another case management
-forced people to remain committed to firm deadlines, even midst in the
-organizational transformation. The engagement in old commitments and tasks
-resulted in ignoring new practices introduced by agile, and eventually the agile
-method was breaking down [P21]. Spayd [P49] also describes how senior management 
-pressed teams to deliver what the customer had been promised, regardless of what
-the current velocity of development predicted.
+overcommitted, but it was later realized that when people are overloaded they
+will not be able to changing their behavior and learn new ways of working [P37,
+P42]. Cowan [P11] describes how time was allocated for tending to old
+commitments. However, even when there was preparation for extra work people with
+specific expertise became overloaded. After time it was realized that
+overcommitment must be reduced by allowing to shift excess work for later time
+[P11]. In another case management forced people to remain committed to firm
+deadlines, even midst in the organizational transformation [P21]. The engagement
+in old commitments and tasks resulted in ignoring new practices introduced by
+agile, and eventually the agile method was breaking down [P21]. Spayd [P49] also
+describes how senior management pressed teams to deliver what the customer had
+been promised, regardless of what the current velocity of development predicted.
+For one part the pressure to deliver interfered with the transformation, but it
+was also forcing a change in the old way of working [P49].
 
@@
 
-A challenge most organizations came across was that an agile model was difficult
-implement. An experienced software team may do well in training, but when the
-time comes to apply agile techniques in practice the team may get lost [P21].
-Agile was also thought being a solution to all problems [P9], success was
-declared prematurely [P49], and expectations created by successful early
-experiments were not fullfilled [P38]. Disapointment was granted when these
-hopes were not fullfilled, but instead implementing the agile model was found
-to be hard and time consuming. The difficulty in implementing agile is evident
-in misconceptions of the agile model, when by-the-book implementations prove to
-be insufficient, and when the agile methods were applied badly to the
-organization.
+A challenge most organizations came across was that the agile model was
+difficult to implement. An experienced software team may do well in training,
+but when the time comes to apply agile techniques in practice the team may get
+lost [P21]. The difficulty in implementing agile is evident in misconceptions of
+the agile model, when by-the-book implementations prove to be insufficient, and
+when the agile methods were applied badly to the organization.
 
@@
 
-On a team level there were several examples of misunderstanding the agile model.
-Cases describe how agile was seen as freedom to hack without documentation
-[P26], that development could be done without design tasks [P37], and that agile
-meant that everyone should become a generalist [P42]. Schatz and Abdelshafi
-describe how teams presented unfinished work at reviews and ignored the
-principle of showing only completed increments, which resulted in a backlog of
-bugs [P44]. In one case team members had perceived the introduction of agile as
-a means to squeeze out more efficiency [P45]. In addition, the flatter
+On a team level there were several examples of misunderstanding agile
+development. Cases describe how agile was seen as freedom to hack without
+documentation [P26], that development could be done without design tasks [P37],
+and that agile meant that everyone should become a generalist [P42]. Schatz and
+Abdelshafi describe how teams presented unfinished work at reviews and ignored
+the principle of showing only completed increments, which resulted in a backlog
+of bugs [P44]. In one case team members had perceived the introduction of agile
+as a means to squeeze out more efficiency [P45]. In addition, the flatter
 organization was seen as fewer career opportunities, and pushed team members to
 internal competition for visibility [P45].
-Still one misunderstanding of agile was due to perceiving it only through the
-tools used, such as project management software [P48]. Superficial focusing only
-on tools themselves and not the reasons behind their use led to frustration
-[P42].
+Still one misunderstanding of agile was due to viewing it only through the tools
+used, such as project management software [P48]. Superficial focusing only on
+tools themselves and not the reasons behind their use led to frustration [P42].
 
@@
 
+As a final example of misconceptions, some organizations considered agile being
+a solution to all problems [P9]. Success was declared prematurely [P49] but
+expectations created by successful early experiments were not fullfilled [P38].
+Disapointment was granted when these hopes were not fullfilled, and instead
+implementing the agile model was found to be hard and time consuming.
+
 % By the book is not good
 
-Even if there was no misunderstandings the correct agile model was found hard to
+Even if there was no misunderstandings the correct agile model was hard to
 implement. Several cases describe how an agile model was hard to learn from
@@
 architects struggled when several Scrum teams were working concurrently.
-Another case concluded that all practices prescribd in XP did not fit enterprise
+Another case concluded that all practices suggested by XP did not fit enterprise
 scale development, and some practices required customization [P49]. The model of
-agile should not be prescribed as a check list, as it would impose a single
+agile should not be prescribed in a strict way, as it would impose a single
 model in an organization with several individual teams and projects [P41].
@@
 % Check this paragraph
-Further indications were given that the advice given in literature will not
-ferry all the way. Benefield [P6] describes how it was dificult finding a
-balance between prescribing a by-the-book implementation which may put people
-off, and giving too much freedom in the agile methods which may weaken core
-practices.
-The reality where the practices needed to be applied was described as
-messy in comparison to the ideal circumstances presented in literature, which
-made it difficult to choose an initial model and gain buy-in for it [P10].
-Difficulties in choosing the model to start with were also due to variances
-perceived in the agile approaches suggested in literature [P30].
+Further indications were given that the advice from literature will not ferry
+all the way. Benefield [P6] describes how it was dificult finding a balance
+between prescribing a by-the-book implementation which may put people off, and
+giving too much freedom in the agile methods which may weaken core practices.
+The reality where the practices needed to be applied was described as messy in
+comparison to the ideal circumstances presented in literature [P10]. Because of
+this it was difficult to choose a model to start with and gain buy-in for [P10].
+Difficulties in choosing a initial model were also due to variances perceived in
+the agile approaches suggested in literature [P30].
 
@@
 Many cases reported that introducing agile practices soon yielded improvements
-on the team level, but as a part of a large transformation teams needed to
-change their external behavior also.
+on the team level, but as a part of a large transformation teams also needed to
+change their external behavior.
 The initial success was followed by challenges when teams needed to work with
@@
 some teams chose to focus only on their own goal [P7]. Independent teams that
-did not respect the larger context they were working in obstructed coordination
-[P13]. Coordination was also difficult when teams were working independently for
-different customers but the applications being built were interdependent [P29].
-One case reported that allowing teams to have different sprint durations created
-delays in delivery [P51].
+did not respect the larger context obstructed coordination [P13]. Coordination
+was also difficult when teams were working independently for different customers
+but the applications being built were interdependent [P29]. One case reported
+that even allowing teams to have different sprint durations created delays in
+delivery [P51].
 
@@
 by quality assurance (QA) activities. Agile methods suggest that essential parts
-of these activities would be performed within the development team. The effort
-required for managing requirements and QA for large systems may be too large for
-individual teams to handle.
+of these activities would be performed within the development team. However,
+when developming large systems the effort required for managing requirements and
+QA may be too heavy for individual teams to handle.
 
@@
 development team. However, large development projects will demand additional
-high level requirements management.
-This is apparent in cases where complexity of the product will require
-additional layers of product management [P45], requirements are created by
-several stakehoder groups and development teams can not be in contact with all
-of them [P39], or teams do not have access to stakeholders due to a distributed
-seetting [P24].
-As an integral part of the software development lifecycle, requirements
-management will be affected by the new agile process. In large organizations
-product and requirements management was initially done in a traditional way, but
-it was contrasted by the agile process that required iteratively supplying small
-requirements to development [P51]. The agile method had to be extended to
-include a process on how to get large requirements into team backlogs [P31].
-
-Requirements were in some cases first defined on a high level in marketing
-requirements documents (MRD) [P1, P31] or functional specifications [P33]. These
-high level requirements needed to be elaborated to be used by agile development
-teams. The requirements breakdown in it self was found to be difficult [P1,
-P33], and also when to do the breakdown and to which detail level [P17].
-Product managers and business analysts were struggling in creating high level
-requirements [P33, P37], and teams were struggling in breaking them down to an
-estimable size [P33]. Smits and Rilliet [P48] describe how high level product
-management delivered requirements in too large chunks resulting in development
-teams spending large amounts of time on defining suitaby sized stories. In one
-case the teams were overwhelmed by reprioritization and requirements bartering
-when stakeholders learned that requirements could be rearranged flexibly [P31].
+high level requirements management. This is apparent in cases where complexity
+of the product will require additional layers of product management [P45],
+requirements are created by several stakehoder groups and development teams can
+not be in contact with all of them [P39], or teams do not have access to
+stakeholders due to a distributed seetting [P24]. Agile methods need to be
+extended to include a process on how to get large requirements into team
+backlogs [P31].
+
+% As an integral part of the software development lifecycle, requirements
+% management will be affected by the new agile process. In large organizations
+% product and requirements management was initially done in a traditional way, but
+% it was contrasted by the agile process that required iteratively supplying small
+% requirements to development [P51].
 
@@
 
+Requirements were in some cases first defined on a high level in marketing
+requirements documents (MRD) [P1, P31] or functional specifications [P33].
+These high level requirements needed to be elaborated to be used by agile
+development teams. The requirements breakdown in it self was found to be
+difficult [P1, P33], and also when to do the breakdown and to which detail level
+[P17].
+Product managers and business analysts were struggling in creating high level
+requirements [P33, P37], and teams were struggling in breaking them down to an
+estimable size [P33]. Smits and Rilliet [P48] describe how high level product
+management delivered requirements in too large chunks resulting in development
+teams spending large amounts of time on defining suitaby sized stories.
+% ?
+In one case the teams were overwhelmed by reprioritization and requirements
+bartering when stakeholders learned that requirements could be rearranged
+flexibly [P31].
+
 % Long term planning
@@
 between short and long term planning [P9, P10]. Typical agile backlogs would
-give only short term visibility, so organization specific practices for business
-to do long term planning had to be created [P26, P31]. Several cases described
-that care had to be taken to avoid long term planning becoming a prescriptive
-practice. In order to preserve the agility of development a schedule driven
-approach was avoided [P10], sensitivity was used when considering setting
-milestones [P13], and striving for more accurate estimates was not let to become
-an excuse for gathering requirements up front [P38].
+give only short term visibility, so organization specific practices for long
+term planning had to be created [P26, P31]. Several cases described that care
+had to be taken to avoid long term planning becoming a prescriptive practice. In
+order to preserve the agility of development a schedule driven approach was
+avoided [P10], sensitivity was used when considering setting milestones [P13],
+and striving for more accurate estimates was not let to become an excuse for
+gathering requirements up front [P38].
 
@@
 
-In an enterprise setting quality assurance tasks will require effort beyond
-the general agile convention of doing testing within a crossfunctional team. In
-large development efforts the agile model may need to be extended to accomodate
-additional testing [P31]. Organizations may have existing separate QA teams that
-must be coordinated to work with development teams [P17, P30, P48]. Integrating
-enterprise scale QA will pose some additional challenges to agile methods.
+In an enterprise setting quality assurance tasks will require effort beyond the
+general agile convention of doing testing within a crossfunctional team.
+Similarly as with requirements engineering, the agile model may need to be
+extended to accomodate additional testing activities [P31]. Organizations may
+have existing separate QA teams that must be coordinated to work with
+development teams [P17, P30, P48]. Integrating enterprise scale QA will pose
+some additional challenges to agile methods.
 
@@
 testing activities can not be done within boundaries of user stories, and
-require more resources than teams can spare [P31]. As testing tasks overlap team
-boundaries it is sensible having a separate teams for the tasks, but
+require more resources than teams can spare [P31]. When testing tasks overlap
+team boundaries it is sensible having a separate teams for the tasks, but
 coordination practice between specialized QA teams and development teams needs
@@
 extended time was reserved for QA at the end of projects, which allowed to
-retroactively resolve ambiguities in requirements [P5]. However, as agile
-required QA work to be done incrementally with development problems with
-ambiguity in requirements became an impediment to QA [P5].
+retroactively resolve ambiguities in requirements [P5]. However, ambiguity
+became an impediment to QA when development worked in short increments and
+there was no extended period for stabilization [P5].
 
@@
 
+* Have flexibility for different teams in the agile model (agile found
+  hard [P41]) vs. conformity to a single model.
+
 % ---------------------------------------------------------------------

### 2014-05-16 journal/article.tex
Challenges: introductory paragraphs

@@
 
+\subsubsection{Typical activities in the transformation}
+
+Training
+
+Use of external consultants
+
+Coaching
+
+Pilot program
+
+Seminars, workshops, other public events
+
+Agile community
+
 \clearpage
@@
 
-Any organizational transformation involving numerous individuals will face some
-challenges. As stated by Research Question 3, we investigated what challenges
-large organizations introducing agile methods may face.
+Any organizational transformation that involves numerous individuals will face
+some challenges. As stated by Research Question 3, we investigated what
+challenges large organizations introducing agile methods may face.
 
 The challenges presented in this section are factors that primary studies
-described hindering the progress of the transformation.
-Some of the challenges presented here were not explicitly presented as
-transformation challenges, but rather as general difficulties in agile
-imlementation. These issues are still included as creating an agile
-implementation can be seen as a part of the transformation process.
-Some solutions for challenges are also presented, but they should be considered
-as ways of describing the challenges rather than direct success factors.
+described hindering the progress of the transformation. Some of the challenges
+presented here do not explicitly relate to the transformation process, but are
+rather general difficulties in agile imlementation. These issues are still
+included as settling the agile implementation can be seen as a part of the
+transformation process. Some solutions for challenges are also presented, but
+they should be considered as ways of describing the challenges rather than
+direct success factors.
 
@@
 
-Challenges can be organized on a high level into three categories. First we
-discuss challenges relating to managing the transformation. These challenges
-relate to managing people and resourcing. Secondly we present challenges in
-choosing and establishing a suitable agile model. Thirdly challenges related to
-large organization size are presented. Size will bring challenges such as
-coordinating work between many teams and adapting other functions than
-development to agile.
-
-% Change resistance
-% Lack of investment in transformation
-% Agile is found to be hard
-% Large organization issues
-% Coordinating work between many teams
-% Integrating other functions than development to agile
-% Integrating requirements management and quality assurance
-% Other challenges in transformation (zealots, falling back)
+The challenges found in the primary studies are summarized in figure
+\ref{fig:challenges_summary}. The challenges are organized on a high level into
+seven categories within three themes. The categories are discussed subsequently
+in this section. First we present generic transformation issues that may apply
+to any organization. As the second theme challenges relating particularly to
+sizable organizations are discussed. The final theme studied is challenges
+encountered when agile development needs to interface with other functions
+involved in a system development life cycle.
+
+It is to be noted that we do not emphasize any of the listed categories or
+challenges above another. The challenge categories we present were found evenly
+in the set of primary studies. We included only challenges that had been
+mentioned in several studies, or were particularly emphasized in fewer studies.
+Challenges with only sporadic mentionings were considered less significant and
+were left out. There was no further evidence that any challenge would be more
+significant than another.
 
@@
 
+\clearpage
+
+\subsection{Success factors in transformation}
+
+Some success factors ...
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+* The factors that are listed are the ones to discuss -- there may be other
+  factors affecting transformation, but this study provides evidence that these
+  are the most important ones. We included only challenges that are discussed in
+  more than studies, which proves that some filtering has been done.
+
+* System development life cycle ?? -- There are inevitably other areas than
+  development of increments, even in an agile process
+
 % ---------------------------------------------------------------------

### 2014-05-16 journal/article.tex
Summary table: Change resistance

@@
 culture will face resistance [P10]. It should be expected that not everyone will
-be willing to change, and some employees will never adapt to the change [P6].
-During a transformation period objections to change may become a major reason
-for loss in time and productivity [P29].
-
-On a team level numoerous reasons for change resistance were given. For
+be willing to change, even so that some employees will never adapt to the new
+model [P6]. During a transformation period objections to change may become a
+major reason for loss in time and productivity [P29]. The main types of change
+resistance discovered in the primary studies are summarized in table
+\ref{table:challenges_changeresistance}.
+
+\begin{table}[b]
+    \begin{tabular}{ l l }
+        \toprule
+        Types of change resistance           &  References  \\
+        \midrule
+        General resistance to change         &  P14, P18, P22, P44, P45   \\
+        Skepticism towards a new model       &  P5, P8, P9, P18, P36, P46 \\
+        Top down mandate creates resistance  &  P2, P12, P37, P49 \\
+        Management not willing to change     &  P2, P3, P6, P49 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Types of change resistance reported in studies}
+    \label{table:challenges_changeresistance}
+\end{table}
+
+Numerous reasons were given for why change resistance emerged. For
 instance, Fecarotta [P14] described the organization of Boeing as risk-averse
-and cautious which slows down any change. The organization was also described
-having a deeply rooted status-quo because of high employee retention, which
-further hindered change [P14]. In some cases it was reported that people worried
-about new roles and responsibilities that the agile model might bring [P44]. For
-instance, testers were worried that they would need to take on cross functional
-tasks, which would be outside their area of expertise [P18]. Yet another reason
-for change resistance was the move from own offices to team areas [P22]. People
-were also feeling being monitored because of the increased level of interaction
-within the team and between project stakeholders [P45]. One study reported the
-cultural change required for middle management as the biggest cause of
-resistance [P2].
+and cautious, which slows down any change. Any change was further hindered by a
+deeply rooted status-quo and high employee retention [P14]. In some cases it was
+reported that people worried about new roles and responsibilities that the agile
+model might bring [P44]. For instance, testers were worried that they would need
+to take on cross functional tasks, which would be outside their area of
+expertise [P18]. Yet another reason for change resistance was the move from own
+offices to team areas [P22]. People were also feeling being monitored because of
+the increased level of interaction within the team and between project
+stakeholders [P45].
 
@@
 
-Another common form of change resistance that was reported was skepticism.
-Seffernick describes how management did on one hand acknowledge benefits of
-agile, but on the other explained away the possibilitiy to apply it due to
-contract reasons, the current matrix organization, and other organizational
-practices [P46]. In that case the director of development still pushed the agile
-transformation forward. Skepticism was often created by other similar
+Another common form of change resistance that was reported was skepticism and
+distrust in suitability of agile. Seffernick describes how management did on one hand
+acknowledge benefits of agile, but on the other explained away the possibilitiy
+to apply it due to contract reasons, the current matrix organization, and other
+organizational practices [P46]. Skepticism was often created by other similar
 misconceptions, including that agile does not work for complex products [P5,
@@
 
-The way the transformation is initiated will also affect how change resistance
-will show. In many cases the initiative to change came from management, but when
+The way the transformation is initiated will affect how change resistance will
+show. In several cases the change initiative came from management, but when
 presented wrong the initiative becomes a mandate that pepole are not receptive
@@
 implementing mandaetd changes did not bring any visible benefits [P37].
-
-Organizations had also bad experiences if the implementation of the agile model
-was prescribed. For instance, doing Scrum based on checklists of practices would
-not work [P41]. Benefield described that it was challenging to find a balance
-between independence for teams to choose their way and diluting the value and
-structure of Scrum [P6]. Giving room for team autonomy is one of the critical
-facets of agile, but it must be ensured that teams do not discard practices that
-add value and structure.
+A problem relating to top down mandate is that if management does not speak out
+a goal towards using agile, developers may feel that the agile methods may be
+replaced by something else at any time [P12].
 
@@
 significant organizational change could not be raised above the team level. For
-instance, Atlas [P3] describes how it would have required executive support to
-integrate the product management office and separate quality assurance groups in
-the agile process. In this case, when managers owning projects with many teams
+instance, Atlas [P3] describes how executive support would have been required to
+extend the agile process to involve the product management office and separate
+quality assurance groups. In this case, when managers %for multi team projects
 were not involved in the transformation, the agile model was restrained from
-spreading beyond team level [P3]. Lack of middle management support was also
-seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
-position to undermine the entire transformation, and may do so if they do not
-participate in, and thus understand, the agile method. Agile brings changes to
-some management roles [P2], and lack of understanding of agile development will
-leave managers feeling left out [P6]. Still one problem is that if management
-does not speak out a goal towards using agile, developers feel that the agile
-methods may be replaced by something else at any time [P12].
+spreading beyond the level of development teams [P3]. Lack of middle management
+support for change and disinclination to change management culture were seen as
+some of the most serious problems in the transformation [P2, P49]. Middle
+management is in a position to undermine the entire transformation, and may do
+so if they do not participate in, and thus understand, the agile method. Agile
+brings changes to some management roles [P2], and lack of understanding of agile
+development will leave managers feeling left out [P6].
 
@@
 
+% Where would this belong?
+% Organizations had also bad experiences if the implementation of the agile model
+% was being prescribed. For instance, implementing Scrum based on checklists of
+% practices would not work [P41]. Benefield described that it was challenging to
+% find a balance between independence for teams to choose their way and diluting
+% the value and structure of Scrum [P6]. Giving room for team autonomy is one of
+% the critical facets of agile, but it must be ensured that teams do not discard
+% practices that add value and structure.
+
 % Customizing badly (mapping to old model)

### 2014-05-16 journal/article.tex
Editing and summary box: Investing in transformation

@@
         \toprule
-        Types of change resistance           &  References  \\
+        Types of change resistance           &  Described in  \\
         \midrule
@@
 organization. Problems will ensue if there is no intention to allow resources
-for the transformation. Examples on this was lack of investment in training, and
-keeping people allocated to tasks from the old arrangement. In some cases the
-transformation was impeded when work spaces could not be updated.
+for the transformation. Table \ref{table:challenges_lackofinvestment} presents
+investments and resourcing arrangements that were witheld, and which caused
+problems in transformation.
+
+\begin{table}[b]
+    \begin{tabular}{ l l }
+        \toprule
+        Factors of investment and resourcing  &  Described in \\
+        \midrule
+        Lack of coaching                      &  P1, P6, P23, P45, P47, P49  \\
+        Lack of training                      &  P11, P20, P45  \\
+        Excessive commitment of people        &  P37, P42, P49  \\
+        Old commitments are kept              &  P11, P21  \\
+        Physical spaces can not be rearranged  &  P6, P29, P36, P49  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of investment and resourcing factors that affect
+             transformation negatively}
+    \label{table:challenges_lackofinvestment}
+\end{table}
 
@@
 
-Lack of activities supporting a transformation was an evident problem. Training
-and coaching were often short of resources, but underbudgeting these activities
-may seriously hamper the transformation.
+Training and coaching are direct investments in transformation and their lack is
+an evident problem. These activities were still underbudgeted, which
+created difficulties in the transformation [P11].
 Hajjdiab [P20] describes how reluctance of management to invest in training
-caused teams to be ill prepared for the agile transformation effort, and
+caused teams to be ill prepared for the agile transformation effort, which
 eventually resulted in ending the use of agile methods. As a less dramatic
-outcome of shortened training was lower motivation [P45].
+outcome of too little training was lower motivation [P45].
 
 Arranging proper coaching was a particular problem in many organizations. It is
-critical to coach teams in their real work environment as a mindset change is
-difficult to achieve by only attending training sessions. In large organizations
-where numerous teams needed to be coached the demand often exceeded the capacity
-of available coaches [P1, P6, P23, P49]. Lack of coaching was also attributed as
-a reason why the success of pilot teams could not be repeated when agile was
-adopted more widely [P9, P45].
+critical to coach teams in their real work environment as a proper change in
+mindset is difficult to achieve by only attending training sessions. [P9, P23]
+In large
+organizations where numerous teams needed to be coached the demand often
+exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
+was also attributed as a reason why the success of pilot teams could not be
+repeated when agile was adopted more widely [P9, P45].
 Silva and Doss [P47] described that when it was hard to fill the coaching
@@
 Some organizations started their agile journey in a state where people were
-overcommitted, but it was later realized that when people are overloaded they
-will not be able to changing their behavior and learn new ways of working [P37,
-P42]. Cowan [P11] describes how time was allocated for tending to old
-commitments. However, even when there was preparation for extra work people with
-specific expertise became overloaded. After time it was realized that
-overcommitment must be reduced by allowing to shift excess work for later time
-[P11]. In another case management forced people to remain committed to firm
-deadlines, even midst in the organizational transformation [P21]. The engagement
-in old commitments and tasks resulted in ignoring new practices introduced by
-agile, and eventually the agile method was breaking down [P21]. Spayd [P49] also
-describes how senior management pressed teams to deliver what the customer had
-been promised, regardless of what the current velocity of development predicted.
-For one part the pressure to deliver interfered with the transformation, but it
-was also forcing a change in the old way of working [P49].
+overcommitted. It was later realized that when people are overloaded they will
+not be able to changing their behavior and learn new ways of working [P37, P42].
+Spayd [P49] also describes how senior management pressed teams to deliver what
+the customer had been promised, regardless of what the current velocity of
+development predicted. For one part the pressure to deliver interfered with the
+transformation, but it was also forcing a change in the old way of working
+[P49].
+
+In one case management forced people to remain committed to firm deadlines, even
+midst in the organizational transformation [P21]. The engagement in old
+commitments and tasks resulted in ignoring new practices introduced by agile,
+and eventually the agile method was breaking down [P21].
+Cowan [P11] describes how time was allocated for tending to old commitments.
+However, even when there was preparation for extra work people with specific
+expertise became overloaded. After time it was realized that overcommitment must
+be reduced by allowing to shift excess work for later time [P11].
 

### 2014-05-17 journal/article.tex
SUmmary table: Agile is hard, coordinating between teams

@@
 
-A challenge most organizations came across was that the agile model was
-difficult to implement. An experienced software team may do well in training,
+A challenge most organizations came across was that implementing the agile model
+was found to be difficult. An experienced software team may do well in training,
 but when the time comes to apply agile techniques in practice the team may get
-lost [P21]. The difficulty in implementing agile is evident in misconceptions of
-the agile model, when by-the-book implementations prove to be insufficient, and
-when the agile methods were applied badly to the organization.
+lost [P21]. Viewpoints on the difficulty of correctly implementing an agile
+model are summarized in table \ref{table:challenges_difficulty}.
+
+% The difficulty in implementing agile is evident in misconceptions of the agile
+% model, when by-the-book implementations prove to be insufficient, and when the
+% agile methods were applied badly to the organization.
+
+\begin{table}
+    \begin{tabular}{ p{0.61\textwidth} p{0.31\textwidth} }
+        \toprule
+        Type of difficulty in implementation  &  Described in \\
+        \midrule
+        Misunderstanding agile development concepts  &
+        P6, P26, P37, P38, P42, \newline
+        P44, P45, P48, P49      \\
+        Literature does not provide enough guidance  &
+        P5, P6, P10, P13, P21, \newline
+        P27, P30, P45, P49     \\
+        Agile practices are customized badly   &
+        P9, P29, P44, P46  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of viewpoints why agile implementation is difficult}
+    \label{table:challenges_difficulty}
+\end{table}
 
@@
 time committing unreasonable workloads on teams [P6]. In another case a
-waterfall nature was seen in giving task estimates as hours left, and task
-breakdowns disconnected from what was really being done [P27].
+waterfall nature was evident when task estimates were given as hours left, and
+task breakdowns disconnected from what was really being done [P27].
 
@@
 Another case concluded that all practices suggested by XP did not fit enterprise
-scale development, and some practices required customization [P49]. The model of
-agile should not be prescribed in a strict way, as it would impose a single
-model in an organization with several individual teams and projects [P41].
+scale development, and some practices required customization [P49].
+
+% In a large organization with several different teams and projects it is clear
+% that one single model can not fit all needs [P41]. The model of
+% agile should not be prescribed in a strict way, as it would impose a single
+% model in an organization with several individual teams and projects [P41].
 
@@
 the methods were customized badly. As by-the-book implementation often was not
-feasible the agile method was tailored to suit the needs of the organization.
+feasible, the agile method was tailored to suit the needs of the organization.
 However, in some cases tailoring meant skipping some practices, which lead into
@@
 Another example of bad customization that counteracts change is replacing agile
-vocabulary in favor of familiar terminology from the old model [P44]. A benefit
-of introducing new vocabulary is that it may stimulate new ways of thinking
-[P44].
+vocabulary in favor of familiar terminology from the old model [P44]. An
+important benefit of introducing new vocabulary is that it stimulates new ways
+of thinking [P44].
 
@@
 teams. One of the most prominent transformation challenges was the difficulty in
-creating a model to coordinate work between many teams.
-% --> Present the different aspects of this topic?
+creating a model to coordinate work between many teams. Table
+\ref{table:challenges_coordinatingteams} presents challenges that arose when
+agile was implemented in environments where many teams need to collaborate.
+
+\begin{table}
+    \begin{tabular}{ p{0.61\textwidth} p{0.31\textwidth} }
+        \toprule
+        Problems in an environment with many teams  &  Described in \\
+        \midrule
+        Interfacing between teams is difficult  & 
+        P9, P10, P13, P17, \newline
+        P24, P26 \\
+        A model with autonomous teams is not good  &  P7, P13, P29, P51  \\
+        Distributed setting causes additional problems  &  P29, P45, P48  \\
+        Technical integration and consistency  &  P5, P24, P26, P27, P50  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of challenges in a setting with many agile teams}
+    \label{table:challenges_coordinatingteams}
+\end{table}
 
@@
 fully utilized as the surrounding organization was not responsive enough [P26].
-There were dependencies between teams in the agile model as well, and the
-dependencies made managing development difficult [P9, P10].
+The rollout of agile had not removed dependencies, and the dependencies
+made managing development difficult [P9, P10].
 
@@
 between the their own goals and broader goals of the entire organization, but
-some teams chose to focus only on their own goal [P7]. Independent teams that
-did not respect the larger context obstructed coordination [P13]. Coordination
-was also difficult when teams were working independently for different customers
-but the applications being built were interdependent [P29]. One case reported
-that even allowing teams to have different sprint durations created delays in
-delivery [P51].
+some teams chose to focus only on their own goal [P7]. Coordination was
+obstructed by independent teams that did not respect the larger context [P13].
+Coordination was also difficult when teams were working independently for
+different customers but the applications being built were interdependent [P29].
+One case reported that even allowing teams to have different sprint durations
+created delays in delivery [P51].
 

### 2014-05-17 journal/article.tex
Summary table: large scale issues, RE and QA

@@
 
-\subsubsection{Large organization issues}
-
-The primary studies revealed challenges that relate particularly to large
-scale. Additional layers of management are usually necessary to steer large
-groups of people, but agile methods provide little guidance on how to organize
-management. Problems may arise in a large environment if the schedule of agile
-adoption varies from team to team. Internal fragmenttation is a further problem
-that may be encountered in large organizations.
+\subsubsection{Other issues in large organizations}
+
+In addition to inter-team coordination there were also other challenges that
+should only emerge in a sizable organization.
+Additional layers of management are usually necessary to steer large groups of
+people, but agile methods provide little guidance on how to organize management.
+Problems may arise in a large environment if the schedule of agile adoption
+varies from team to team. Internal fragmenttation is a further problem that may
+be encountered in large organizations.
+Table \label{table:challenges_largescale} summarizes these challenges.
+
+\begin{table}[b]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.56\textwidth}
+                     >{\raggedright\arraybackslash}p{0.36\textwidth} }
+        \toprule
+        Challenges relating to large scale  &  Described in \\
+        \midrule
+        Management roles unclear in agile       &  P2, P11, P27, P30, P38, P49, P52 \\
+        Using old and new models side by side   &  P1, P8, P10, P26, P29 \\
+        Management uses waterfall model         &  P3, P6,P35,P38, P45, P46  \\
+        Duplicating management with two models  &  P20, P48 \\
+        Agile models diverge between teams      &  P8, P9, P38, P43, P48 \\
+        Internal boundaries in organization     &  P10, P11 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of challenges relating especially to large scale}
+    \label{table:challenges_largescale}
+\end{table}
 
@@
 processes that emphasize self organization.
-Especially the role of middle management was seen as unclear in agile methods
-[P2]. This is problematic as an agile transformation requires cultural change
-particularly on the middle management level [P2].
+Especially the role of middle management was unclear in agile methods [P2]. This
+is problematic as an agile transformation requires cultural change particularly
+on the middle management level [P2].
 It was described that managers needed to resist the tendency to take command and
@@
 software development. Cases describe how development teams were the first
-entities in the organization that were considered when introducing agile [P2,
-P17, P31, P38]. However, development operates in a larger context and will need
-to interface with other organizational functions.
+entities that were considered when introducing agile [P2, P17, P31, P38].
+However, development operates in a larger context and will need to interface
+with other organizational functions.
 Challenges were faced when exposing other parts of the organization to the agile
@@
 
-Agile affects timings in the software delivery lifecycle which caused two types
+Agile affects timings in the software delivery lifecycle, what caused two types
 of problems. Firstly, the iterativeness of agile models was problematic when
@@
 have a long lead time but required committing to a set of functionalities in
-good time. The agile models' emphasis on short time horizon and flexibility in
+good time. The agile model's emphasis on short time horizon and flexibility in
 prioritization did not mix well with such activities.
@@
 
-Traditionally development is preceeded by requirements engineering and succeeded
+Development is traditionally preceeded by requirements engineering and succeeded
 by quality assurance (QA) activities. Agile methods suggest that essential parts
@@
 when developming large systems the effort required for managing requirements and
-QA may be too heavy for individual teams to handle.
+QA may be too heavy for individual teams to handle. Table
+\ref{table:challenges_requirementsandqa} presents challenges in implementing
+additional requirements management and quality assurance required in large
+development projects.
+
+\begin{table}[t]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
+                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
+        \toprule
+        Challenges in requirements engineering and quality assurance  &  Described in \\
+        \midrule
+        How to accommodate high level requirements management  &  P24, P31, P39, P45  \\
+        Problems with breakdown of requirements        &  P1, P17, P33, P48  \\
+        Creating and estimating user stories is hard   &  P5, P11, P27, P30, P33, P37 \\
+        Handling coexisting long and short term planning  &  P9, P10, P13, P26, P31, P38 \\
+        Accommodating performance, load, and memory testing  &  P17, P28, P31, P48 \\
+        Lack of automated testing slows down           &  P10, P11, P17 \\
+        New requirements management practices affect QA  &  P5, P48 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of requirements engineering and quality assurance related
+             challenges relating in the agile model}
+    \label{table:challenges_requirementsandqa}
+\end{table}
 
@@
 
-In agile methods requirements engineering is usually closely tied to the
-development team. However, large development projects will demand additional
-high level requirements management. This is apparent in cases where complexity
-of the product will require additional layers of product management [P45],
-requirements are created by several stakehoder groups and development teams can
-not be in contact with all of them [P39], or teams do not have access to
-stakeholders due to a distributed seetting [P24]. Agile methods need to be
+In agile methods requirements engineering is usually informal and closely tied
+to the development team. However, large development projects will demand
+additional high level requirements management. This is apparent in cases where
+complexity of the product will require additional layers of product management
+[P45], requirements are created by several stakehoder groups and development
+teams can not be in contact with all of them [P39], or teams do not have access
+to stakeholders due to a distributed seetting [P24]. Agile methods need to be
 extended to include a process on how to get large requirements into team
 backlogs [P31].
-
-% As an integral part of the software development lifecycle, requirements
-% management will be affected by the new agile process. In large organizations
-% product and requirements management was initially done in a traditional way, but
-% it was contrasted by the agile process that required iteratively supplying small
-% requirements to development [P51].
-
-Several cases highlighted that much learning was needed to master the new
-process of creating user stories, both on product management and development
-levels [P27, P30, P33, P37]. There were problems such as ambiguity in
-requirements [P5, P11], and effort estimation for stories was hard to do well
-[P27, P37].
 
@@
 teams spending large amounts of time on defining suitaby sized stories.
-% ?
-In one case the teams were overwhelmed by reprioritization and requirements
-bartering when stakeholders learned that requirements could be rearranged
-flexibly [P31].
+
+Several cases highlighted that much learning was needed to master the new
+process of creating user stories, both on product management and development
+levels [P27, P30, P33, P37]. There were problems such as ambiguity in
+requirements [P5, P11], and effort estimation for stories was hard to do well
+[P27, P37].
+
+% As an integral part of the software development lifecycle, requirements
+% management will be affected by the new agile process. In large organizations
+% product and requirements management was initially done in a traditional way, but
+% it was contrasted by the agile process that required iteratively supplying small
+% requirements to development [P51].
+
+% In one case the teams were overwhelmed by reprioritization and requirements
+% bartering when stakeholders learned that requirements could be rearranged
+% flexibly [P31].
 

### 2014-05-17 journal/article.tex
Edits

@@
 
+* Team centricity -- agile focuses on team rather than individuals
+
+* Agile focuses on short time horizons ???
+
 * Agile is a mindset -- it can not be studied from a book -- need of coaching
@@
 \begin{table}
-    \begin{tabular}{ p{0.28\textwidth} l l }
+    \centering
+    \begin{tabular}{ l l l }
         \toprule
@@
 \begin{table}
-    \begin{tabular}{ p{0.22\textwidth} p{0.76\textwidth} }
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.22\textwidth}
+                     >{\raggedright\arraybackslash}p{0.70\textwidth} }
         \toprule
@@
         \midrule
-        Agile methods\newline (before \& after) &
+        Agile methods (before \& after) &
             \texttt{agile, scrum, "extreme programming",
@@
             "roll-out", rollout}\\
-        Only software\newline related atricles &
+        Only software related atricles &
             \texttt{(software OR (conference="agile, xp, icgse, icse"))
@@
 \begin{table}[h]
-    \begin{tabular}{ p{0.28\textwidth} p{0.7\textwidth} }
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.26\textwidth}
+                     >{\raggedright\arraybackslash}p{0.66\textwidth} }
         \toprule
@@
 
-Time of\newline transformation & When the transformation has been in
+Time of transformation & When the transformation has been in
 progress, or how long the transformation has taken. Possibly relative to
@@
 
-Geographical\newline location & Where the organization is located. \\
+Geographical location & Where the organization is located. \\
 
@@
 
-\begin{table}[h]
-    \begin{tabular}{ p{0.32\textwidth} p{0.65\textwidth} }
+\begin{table}[t]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
+                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
         \toprule
@@
 
-		RQ2: Transformation\newline process &
+		RQ2: Transformation process &
 		Statements that describe the transformation process (top-down, big bang,
@@
 \begin{table}[h]
+    \centering
     \begin{tabular}{ l r r }
@@
 \begin{table}
+    \centering
     \begin{tabular}{ l r r r }
@@
 % ---------------------------------------------------------------------
+\clearpage
+
 \section{Results}
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
   \begin{center}
@@
 
-\begin{figure}[b]
+\begin{figure}[!t]
   \begin{center}
@@
 
-The business areas of the case organizations are presented in table
-\ref{table:businessareas}. Different business areas were represented somewhat
-evenly. Online services was the largest group, including companies
-providing software as a service solutions for businesses, online media players
-for consumers, online services for consumers, and communication software for
-businesses. The second largest group was telecommunications, including companies
-such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
-largest group was enterprise management solution providers with products for
-business process management, project protfolio management, and facility
-management.
-
 \begin{table}[!t]
+    \centering
     \begin{tabular}{ l r }
@@
 
+The business areas of the case organizations are presented in table
+\ref{table:businessareas}. Different business areas were represented somewhat
+evenly. Online services was the largest group, including companies
+providing software as a service solutions for businesses, online media players
+for consumers, online services for consumers, and communication software for
+businesses. The second largest group was telecommunications, including companies
+such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
+largest group was enterprise management solution providers with products for
+business process management, project protfolio management, and facility
+management.
+
 \subsection{Agile methods used}
@@
 
-\begin{table}[t]
+\begin{table}[h]
+    \centering
     \begin{tabular}{ l r }
@@
 
+\clearpage
+
 \subsection{Reasons to change}
@@
 \begin{table}
+    \centering
     \begin{tabular}{ p{1.2cm} p{1.2cm} l }
@@
 \begin{table}[b]
+    \centering
     \begin{tabular}{ l l }
@@
 
-\begin{table}[b]
+\begin{table}[h]
+    \centering
     \begin{tabular}{ l l }
@@
 \begin{table}
-    \begin{tabular}{ p{0.61\textwidth} p{0.31\textwidth} }
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
         \toprule
@@
         Misunderstanding agile development concepts  &
-        P6, P26, P37, P38, P42, \newline
-        P44, P45, P48, P49      \\
+                P6, P26, P37, P38, P42, P44, P45, P48, P49      \\
         Literature does not provide enough guidance  &
-        P5, P6, P10, P13, P21, \newline
-        P27, P30, P45, P49     \\
+                P5, P6, P10, P13, P21, P27, P30, P45, P49     \\
         Agile practices are customized badly   &
@@
 
-\begin{table}
-    \begin{tabular}{ p{0.61\textwidth} p{0.31\textwidth} }
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
+                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
         \toprule
-        Problems in an environment with many teams  &  Described in \\
+        Problems in environments with many teams  &  Described in \\
         \midrule
-        Interfacing between teams is difficult  & 
-        P9, P10, P13, P17, \newline
-        P24, P26 \\
+        Interfacing between teams is difficult     &  P9, P10, P13, P17, P24, P26 \\
         A model with autonomous teams is not good  &  P7, P13, P29, P51  \\

### 2014-05-17 journal/article.tex
Figure for full study selection process

@@
 \item
-RQ2: How do large scale agile transformations usually proceed; do
-     transformations exhibit patterns?
+RQ2: How do large scale agile transformations usually proceed?
 
@@
 
-% --> Why are the preliminary searches done?
-
 Before proceeding with the database search a number of preliminary searches were
@@
 potentially relevant papers that should be matched by the actual database
-search. We also wanted to list top ranked hits by trivial keywords that the
+search. We started by examining top ranked hits by trivial keywords that the
 more complex final search string might miss. Initial searches were made using
 keywords that were as general as possible, including ``agile transformation''
-and ``large scale agile''. Additionally, a trial keyword search was done in the
+and ``large scale agile''. Secondly, a trial keyword search was done in the
 selected databases (listed in table \ref{table:databases}) using the following
@@
 
-The study selection process inlcuded three steps, each step refining the set of
-selected papers. First a database search was used to gather an initial set of
-papers. This set was then refined by filtering out abstracts and finally full
-text filtering. The study selection process is outlined in Figure
-\ref{fig:selectionprocess}.
+The selection of studies was primarily based on a database search. The initial
+step in the database search was constructing a search string which yielded a
+reasonably large result set. The search result was then refined in two steps, by
+first filtering out abstracts and finally full text filtering. The study
+selection process starting from the database search is outlined in figure
+\ref{fig:selectionprocess}. The result of the database search was still amended
+with some results from the preliminary search and relevant references from all
+examined papers.
 
@@
     \includegraphics[width=1\textwidth]{graphics/research_process.pdf}
-    \caption{Study selection process}
+    \caption{Database search as study selection process}
     \label{fig:selectionprocess}
@@
 
-A keyword search in online databases was chosen as the means to find potential
-primary studies. The databases included in the search are listed in Table
-\ref{table:databases}. All databases supported use of complex boolean logic in
-searches which allowed us flexibility when constructing the search strings.
+A keyword search in online databases was chosen as the primary means to find
+potential primary studies. The databases included in the search are listed in
+Table \ref{table:databases}. All databases supported use of complex boolean
+logic in searches, which allowed us flexibility in constructing the search
+strings.
 
@@
 preliminary searches, and 2 papers from references, 52 papers were selected
-as primary studies.
+as primary studies. The full selection process is illustrated in figure
+\ref{fig:selectionprocess_full}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
+    \caption{Full study selection process}
+    \label{fig:selectionprocess_full}
+  \end{center}
+\end{figure}
 

### 2014-05-17 journal/article.tex
Editing: Large-scale hyphen

@@
 
-agile \sep transformation \sep large scale \sep adopting agile
+agile \sep transformation \sep large-scale \sep adopting agile
 
@@
 
---> Why is large scale software development different from smaller scale?
+--> Why is large-scale software development different from smaller scale?
 
@@
 \item
-RQ1: Why do large organizations initiate an agile transformation?
+RQ1: Why do large software development organizations initiate an agile
+transformation?
 
 \item
-RQ2: How do large scale agile transformations usually proceed?
+RQ2: How do large-scale agile transformations usually proceed?
 
 \item
-RQ3: What are typical successes and challenges in a large scale agile
+RQ3: What are typical successes and challenges in a large-scale agile
      transformation process?
@@
 agile in a large organization but not describing how the new methods were
-introduced \cite{Mishra2011}, and merely presenting agile tools in large scale
+introduced \cite{Mishra2011}, and merely presenting agile tools in large-scale
 use \cite{Kim2012}. These sources were excluded as they did not privide insights
@@
 size. For instance, we deemed the case presented by Cloke \cite{Cloke2007} to be
-included as there was indications of large scale considerations, although the
+included as there was indications of large-scale considerations, although the
 size remained unclear. The case presented by Miller and Haddad \cite{Miller2012}
@@
 
-Large scale definition & A definition of "large scale" software development. \\
+Large scale definition & A definition of large-scale software development. \\
 
@@
   focuses on transformation process only, it can not be completely separated
-  from discussion on scaling up, or on large scale agile implementation in
+  from discussion on scaling up, or on large-scale agile implementation in
   general.
@@
   also the agile method to suit the organization.
-  --> A large scale agile model may require tailoring and trial over time
+  --> A large-scale agile model may require tailoring and trial over time
 
@@
 
-* Compare how recommendations in large scale application models of agile relate
+* Compare how recommendations in large-scale application models of agile relate
   to the challenges and success factors reported in the transformaiton proecss.

### 2014-05-18 journal/article.tex
Improved inclusion facets

@@
 Based on the research questions and intended focus of the research, we defined
-three facets to guide inclusion/exclusion decisions: agile software engineering,
-focus on transformation, and large scale. The first facet covers primary studies
-focusing on software engineering organizations applying or striving to apply
-agile methods. The second facet states that included studies must provide
-insights relevant to organizational transformation, and specifically to the
-research questions. The third facet underlines the particular contribution we
-aim to provide with this research, namely the exclusive focus on large
-organizations. These three facets were used for inclusion and exclusion in each
-study selection step, including search string design, filtering by abstracts,
-and full text filtering.
+four facets to guide inclusion/exclusion decisions: agile software engineering,
+focus on transformation, large scale, and empiricality. The first facet covers
+primary studies focusing on software engineering organizations applying or
+striving to apply agile methods. The second facet states that included studies
+must provide insights relevant to organizational transformation, and
+specifically to the research questions. The third facet underlines the
+particular contribution we aim to provide with this research, namely the
+exclusive focus on large organizations. The fourth facet limits inclusion to
+studies presenting real world cases. These four facets were used for inclusion
+and exclusion in each study selection step, including search string design,
+filtering by abstracts, and full text filtering.
+
+Table \ref{table:facets} lists the facets and gives examples on matching topics
+and not relevant topics. For a study to be included it needed to be relevant
+in all facets. Even if a study discussed some of the irrelevant topics it could
+be included as long as it had some content matcing all the facets.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.20\textwidth}
+                     >{\raggedright\arraybackslash}p{0.34\textwidth}
+                     >{\raggedright\arraybackslash}p{0.33\textwidth} }
+        \toprule
+        Facet  &  Relevant topics  &  Not relevant topics  \\
+        \midrule
+        Agile software development  & 
+                The organization develops software;
+                The development method introduced is agile  &
+
+                Agile manufacturing;
+                Scrum in management boards  \\
+
+        Organizational transformation  &
+                Presenting insights in the process of transformation  &
+
+                Comparison of before and after;
+                How agile is being used in large scale   \\
+
+        Large scale    &
+                At least 50 development personel or 6 teams uses agile &
+
+                Scaling up from small;
+                A single agile team in a large setting \\
+
+        Empiricality   &
+                Case studies, experience reports  &
+
+                Textbooks, student experiments, theory papers  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Facets of inclusion criteria with examples of relevant and
+             not relevant topics}
+    \label{table:facets}
+\end{table}
 
@@
 seen as a ``scaling up'' journey. However, we excluded cases where a initially
-small organization was scaled up \cite{Maranzato2012}, and discussions with a
-process centered viewpoint without describing organizational change
-\cite{Lyon2008}.
+small organization was scaled up \cite{Maranzato2012}, and discussions focusing
+on processes or tools without describing organizational change \cite{Lyon2008}.
 % Also [Greening 2010] on process centered viewpoint without transformation
@@
 
-In addition to the criteria listed above we excluded studies that were not
-empirical. We excluded textbooks, studies merely presenting theories, and other
-studies that did not include any case organization. We also excluded studies of
-benefits or limitations of agile in general. Lastly, we excluded student
-experiments as it is implausible to simulate the dynamics of a large
-organization in an experiment.
+According to the facet of \emph{empiricality} we excluded studies that did not
+discuss a distinguishable real world case. We excluded textbooks, studies merely
+presenting theories, and other studies that did not include any case
+organization. We also excluded studies of benefits or limitations of agile in
+general. Lastly, we excluded student experiments as it is implausible to
+simulate the dynamics of a large organization in an experiment.
 
@@
 facet \emph{large scale} from the keyword search, with the consequence of
-increasing the manual filtering effort in the subsequent steps.
-Having a substantial part of the matches in non-relevant areas such as agile
+increasing the manual filtering effort in the subsequent steps. Also the facet
+or \emph{empiricality} could not be represented by keywords. Having a
+substantial part of the matches in non-relevant areas such as agile
 manufacturing, we included only articles including the term ``software'' or
@@
 
-Full text filtering was performed by evaluating each article against the three
+Full text filtering was performed by evaluating each article against the four
 facets of the inclusion criteria. Filtering was done in two steps. In the first
-step Dikert extracted data relevant to the three facets of the inclusion
-criteria. Based on the extracted data 76 papers could be immediately deemed as
-included or excluded. The remaining 94 papers were evaluated against the
-inclusion criteria by both Dikert and Paasivaara, and a decision was made after
-discussing each paper separately. In difficult cases Lassenius was consulted to
-reach a decision. As a result of the full text filtering 47 papers were selected
-to be included.
+step Dikert extracted data relevant to the four facets. Based on the extracted
+data 76 papers could be immediately deemed as included or excluded. The
+remaining 94 papers were evaluated against the inclusion criteria by both Dikert
+and Paasivaara, and a decision was made after discussing each paper separately.
+In difficult cases Lassenius was consulted to reach a decision. As a result of
+the full text filtering 47 papers were selected to be included.
 

### 2014-05-18 journal/article.tex
Editing: Research method

@@
 potentially relevant papers that should be matched by the actual database
-search. We started by examining top ranked hits by trivial keywords that the
-more complex final search string might miss. Initial searches were made using
-keywords that were as general as possible, including ``agile transformation''
-and ``large scale agile''. Secondly, a trial keyword search was done in the
-selected databases (listed in table \ref{table:databases}) using the following
-search string:
+search, and to evaluate different search strings. We started by examining top
+ranked hits by trivial keywords that the more complex final search string might
+miss. Initial searches were made using keywords that were as general as
+possible, including ``agile transformation'' and ``large scale agile''.
+Secondly, a trial keyword search was done in the selected databases (listed in
+table \ref{table:databases}) using the following search string:
 
@@
 From these preliminary searches we compiled a list of 117 possibly relevant
-papers that should be matched by the actual database search.
+papers that the actual database search would be benchmarked against.
 
@@
 
-\begin{figure}[b]
+\begin{figure}
   \begin{center}
@@
 
-We constructed a search string based on the facets \emph{agile software
-deveolpment} and \emph{organizational transformation}, as presented in section
-\ref{sec:inclusioncriteria}. Preliminary searches had showed that picking
-keywords with good precision was difficult. Therefore we decided to exclude the
-facet \emph{large scale} from the keyword search, with the consequence of
-increasing the manual filtering effort in the subsequent steps. Also the facet
-or \emph{empiricality} could not be represented by keywords. Having a
-substantial part of the matches in non-relevant areas such as agile
+We proceeded with constructing a search string based on the facets presented in
+section \ref{sec:inclusioncriteria}. However, preliminary searches had showed
+that picking keywords with good precision was difficult. Especially the facet
+\emph{large scale} was difficult to represent with precise words. Also the facet
+of \emph{empiricality} could not sensibly be represented by keywords. We decided
+to include only the factets \emph{agile software deveolpment} and
+\emph{organizational transformation} in the keyword search, with the consequence
+of increasing the manual filtering effort in the subsequent steps.
+
+Having a substantial part of the matches in non-relevant areas such as agile
 manufacturing, we included only articles including the term ``software'' or
@@
 
-In the full text filtering step the references of all papers were also examined
-for relevance, including 2 referenced papers as primary sources. Most case study
-papers used references very scarcely, mostly referencing well known descriptions
-of agile methods.
+After the full text filtering step the references of all papers were also
+examined for relevance. Most case study papers used references very scarcely,
+mostly referencing well known descriptions of agile methods. We included 2
+referenced papers in full text analysis.
 
@@
 To give direction for code induction seven different code families were
-established, as presented in Table \ref{table:codefamilies}. A description
-defining scope and examples for codes were defined for each family. The example
-codes would not necessarily be used as actual codes, and would not represent the
-entire or final scope of the families.
+established on beforehand, as presented in Table \ref{table:codefamilies}. A
+description defining scope and examples for codes were defined for each family.
+The example codes would not necessarily be used as actual codes, and would not
+represent the entire or final scope of the families.
 
@@
 predefined code families. The labels of existing codes may need to be adjusted
-when new quotations are assigned. Many codes may be assigned to one quotation
-and quotations of different codes may be overlapping.
+when new quotations are assigned. A clarifying description is written for every
+code, so that the precise meaning of the codes can be reviewed. Many codes may
+be assigned to one quotation and quotations of different codes may be
+overlapping.
 
@@
 quoted text passage. Having enough context should broaden the understanding
-of the coded concept when looking at the quotation in isolation..
+of the coded concept when looking at the quotation in isolation.
 
@@
 
-The total accumulated codes and quotations of the integrated deductive and
-inductive coding process are presented in Table \ref{table:codecount}. The total
-number of quotations is less than the sum of quotations in the categories
-because a single quotation may have multiple codes.
+The total number of codes and quotations created in the coding process is
+presented in Table \ref{table:codecount}. The total number of quotations is less
+than the sum of quotations in the categories because a single quotation may have
+multiple codes.
 
@@
 
+\clearpage
+
 \subsubsection{Validation of coding}
@@
 found in each document, and how many of the themes were matching. We concluded
-that on average two of three themes found by one researcher are similarly found
-by the other researcher, but one of three themes is not found by the other
+that on average two of three themes found by one researcher were similarly found
+by the other researcher, but one of three themes was not found by the other
 researcher.
@@
 codes match between two independent researchers. The intent is to measure how
-codes that are deemed similar by their names match in the actual quoted text. On
-a scale of 0 to 3 we evaluated the quoted text of codes in each theme that
-matched for both researchers, as described in the previous paragraph. A score of
-3 was given for matching codes whose quotation content was virtually the same
-for both researchers. A score of 2 was given if the quotations were mostly the
-same, but other researcher had included one text passage that the other had
-missed. A score of 1 was given if roughly half of the quotations were the same.
-If the quotations had less similarity than this a score of 0 was given. One code
-provided typically 1 to 3 quotations per document. The quotation overlap score
-was counted including only codes that had matching themes. We concluded that the
-average quotation overlap score was 2.2.
+well codes that are deemed similar by their names match in the actual quoted
+text. We assigned a quotation overlap score for each code in themes that both
+researchers had found. A score of 3 was given for matching codes whose quotation
+content was virtually the same for both researchers. A score of 2 was given if
+the quotations were mostly the same, but other researcher had included one text
+passage that the other had missed. A score of 1 was given if roughly half of the
+quotation was the same. If the quotations had less similarity than this a score
+of 0 was given. One code provided typically 1 to 3 quotations per document. The
+quotation overlap score was counted including only codes that had matching
+themes. We concluded that the average quotation overlap score was 2.2.
 

### 2014-05-18 journal/article.tex
Reasons to change refactored

@@
 it was simply mentioned that management had decided to change the organization.
-The identified reasons for organizations to change are presented in table
-\ref{table:reasonstochange}.
+The reported reasons for starting an organizational change towards agile are
+summarized in figure \ref{fig:reasonstochange_summary}.
 
@@
 
+The most prominent reason to start the transformation was that business demanded
+shorter lead times and faster delivery. Other top reasons to start change were
+dysfunction in management, overhead in the current process, problems in
+scheduling, and demand to improve quality.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics{graphics/reasonstochange_summary.pdf}
+    \caption{Reasons why organizations started an agile transformation}
+    \label{fig:reasonstochange_summary}
+  \end{center}
+\end{figure}
+
+\subsubsection{Business demands speed and flexibility}
+
+Most cases reported that a major reason to change was that business demanded
+improvement. Most companies wanted to improve their competitiveness, or even
+feared that they are losing competition. The way to respond was to improve
+delivery speed and responsiveness to change. Table
+\ref{table:reasonstochange_business} presents business related reasons to
+introduce an agile model.
+
 \begin{table}
     \centering
-    \begin{tabular}{ p{1.2cm} p{1.2cm} l }
-        \toprule
-        \multicolumn{3}{l}{Reasons to change}  \\
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        Business reasons for starting a change   &  Described in  \\
         \midrule
-
-        \multicolumn{3}{l}{External issues: Need to respond to changes in business}  \\
-
-        &  \multicolumn{2}{l}{Demand for more frequent deliveries} \\
-        &  &  Time To Market must lessen  \\
-        &  &  Long cycle times are seen as problematic  \\
-
-        &  \multicolumn{2}{l}{Something must be done in order to remain competitive} \\
-        &  &  Change is required to stay in business  \\
-
-        &  \multicolumn{2}{l}{Customers demand improvement} \\
-
-        &  \multicolumn{2}{l}{Demand for flexibility or responsiveness} \\
-        &  &  Demand to accomodate changes efficiently  \\
-        &  &  Better responsiveness to customer requests  \\
-
-        \multicolumn{3}{l}{\rule{0pt}{4ex} Internal issues: Problems in managing software development}  \\
-
-        &  \multicolumn{2}{l}{General problems with management of software development} \\
-        &  &  Management dysfunction  \\
-        &  &  Problems with motivation and morale  \\
-        &  &  Problems coordinating a growing organization  \\
-
-        &  \multicolumn{2}{l}{Problems with overhead and excess in process}  \\
-        &  &  Excess, slowness and costs caused by process  \\
-        &  &  The current process is being ignored  \\
-
-        &  \multicolumn{2}{l}{Problems with planning or keeping schedules}  \\
-
-        &  \multicolumn{2}{l}{Too long cycle time before feedback}  \\
-
-        &  \multicolumn{2}{l}{Demand to improve product quality}  \\
-
-        \multicolumn{3}{l}{\rule{0pt}{4ex} Communication and collaboration issues}  \\
-
-        &  \multicolumn{2}{l}{Communication and collaboration dificulties}  \\
-        &  &  Problems in communication across and within projects  \\
-        &  &  Teams work in silos or isolation  \\
-
-        &  \multicolumn{2}{l}{Requirements engineering issues}  \\
-        &  &  Requirements are not understood  \\
-        &  &  Unawareness of what is important  \\
-
-        &  \multicolumn{2}{l}{Demand of better customer collaboration}  \\
-
+        Time to market must be reduced    &  P17, P18, P32, P38, P39, P47  \\
+        Delivery is considered slow       &  P2, P8, P25, P36, P46, P49  \\
+        Something must be done to remain competitive  &
+                P9, P10, P11, P18, P19, P32, P38, P42, P49  \\
+        Business partners demand faster delivery  &  P5, P23, P39, P40, P46  \\
+        Need to accomodate change during projects  &
+                P4, P8, P18, P25, P26, P33, P36, P39, P42, P49, P51, P52  \\
         \bottomrule
     \end{tabular}
-    \caption{Summarizing why large organizations initiate an agile transformation (RQ1).
-             The table presents a hierarchial view of reasons for the organization to change.}
-    \label{table:reasonstochange}
+    \caption{Business related reasons for starting an agile transformation}
+    \label{table:reasonstochange_business}
 \end{table}
 
-The reasons to change were are grouped in three top-level categories: external
-issues, internal issues, and communication issues. External issues are related
-to the business area which the organization is operating in. Internal issues are
-related to the way the organization is managed and software is being developed.
-Communication and collaboration issues are related to internal and external
-collaboration. Change was motivated by known problems in these categories, such
-as known process dysfunctions or communication problems. There was also internal
-and external demand for improvement, for instance in speed of delivery and
-quality.
-
-\subsubsection{External issues}
-
-The most prominent reason for introducing agile methods was the demand to
+The most frequently mentioned reason for introducing agile methods was the demand to
 deliver software faster and more frequently. A common statement was that it was
-necessary to reduce the Time To Market (TTM) [P17, P18, P32, P39, P47]. As
+necessary to reduce the time to market (TTM) [P17, P18, P32, P39, P47]. As
 stated in one study: ``The market was moving faster than the development group
@@
 transformation. Especially speed of delivery was seen as a means to respond to
-the competition [P9, P38]. As Greening describes: ``The average release duration
-peaked at 41 and 35 months, an alarming state that could enable competitors to
-gain market share.'' [P19]. In many cases the transformation was seen as
-essential for the company to remain competitive [P10, P11, P18, P42, P49]. For
-instance, British Telecom realized that ``From a strategic view, there is little
-alternative than to change the old ways of working'' [P32].
+the competition [P9, P38]. As Greening [P19] describes: ``The average release
+duration peaked at 41 and 35 months, an alarming state that could enable
+competitors to gain market share.''. In many cases the transformation was seen
+as essential for the company to remain competitive [P10, P11, P18, P42, P49].
+For instance, British Telecom realized that ``From a strategic view, there is
+little alternative than to change the old ways of working'' [P32].
 
@@
 development and release cycles [P5, P39]. As described in one study: ``Speed of
-delivery is of the essence – customers are no longer prepared to wait for 18
+delivery is of the essence -- customers are no longer prepared to wait for 18
 months or more for solutions to be delivered'' [P23]. In some cases business
@@
 
-Yet another frequent external factor driving for improvement was the need to
-accomodate change during development. Organizations saw agile as a opportunity
-to improve responsiveness to change in general [P25, P26, P42].
+Yet another factor pushing for improvement was the need to accomodate change
+during development. Organizations saw agile as an opportunity to improve
+responsiveness to change in general [P25, P26, P42].
 Some organizations faced difficulties when changes to plans were needed. For
@@
 and making changes created management overhead [P51]. Other organizations saw an
-business advantage if priorities could be changed as required [P8, P39].
+business advantage in enabling priorities to be changed as required [P8, P39].
+Domain knowledge will deepen during projects and environments will inevitably
+change, and therefore the requirements should not stay fixed [P4].
 Customers and business representatives were in many cases attributed as the
@@
 
-\subsubsection{Internal issues}
-
-The internal issues relate to how software development is managed in the
-organization. In many organizations the project management practice was in a
-state of dysfunction, and it needed to be corrected. Management problems were
-for instance constantly recurring crisis events [P30], managers ignoring
-prioritization of tasks [P38], and reluctance towards focusing on delivering
-business value [P49]. Some cases stated that projects simply failed [P21]. Some
-organizations wanted to improve generic project management practice, such as
-risk management [P8], management of offshoring [P38], and improving the quality
-of life of team members [P44]. Several organizations were growing and
-experienced that the old development method did not scale up [P16, P30, P50, P51].
-There was also a desire to improve team motivation and commitment, and
-collaboration between management and developers [P39].
+\subsubsection{Management issues}
+
+The way software development was managed was typically seen as problematic. The
+problems related to project management, people management, and managing
+schedules. Management related problems that the transformation was hoped to
+correct are listed in table \ref{table:reasonstochange_management}.
+
+\begin{table}[b]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        Management problems           &  Described in  \\
+        \midrule
+        Software development is managed badly   &  P8, P21, P30, P38, P44, P49  \\
+        Command and control style management    &  P9, P25, P30, P40, P44  \\
+        Scheduling and estimation is difficult  &  P16, P19, P21, P30, P33, P40, P41, P51  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Management problems stated as reasons for organizational change}
+    \label{table:reasonstochange_management}
+\end{table}
+
+In many organizations the project management practice was in a state of
+dysfunction, and it needed to be corrected. Project management problems included
+constantly recurring crisis events [P30], managers ignoring prioritization of
+tasks [P38], and reluctance towards focusing on delivering business value [P49].
+Some cases stated that projects simply failed [P21]. Some organizations wanted
+to improve generic project management practice, such as risk management [P8],
+management of offshoring [P38], and generally improving the quality of life of
+team members [P44].
 
@@
 
-A common problem that was acknowledged in organizations was overhead in process,
-which showed as bureucracy and needless costs. Examples on overhead in process
-were unproductive meetings [P38], sign-offs and gates [P9], excessive focus on
-conformance to process [P40], and change management overhead [P51]. Creating
-excessive documentation instead of business value was seen as a problem in some
-cases [P24, P35].
-From business representatives point of view the processes were unreasonably
-costly, while at the same time insufficiently capable of delivering business
-value [P36, P46, P49]. In one case business realized that customers did not
-request the heavy CMMI based process [P24].
+Yet another problem in managing development was problems with schedules.
+Some cases described schedules as unpredictable and estimation being difficult
+[P16, P30, P40, P51]. In other cases it was reported that release deadlines were
+often missed [P19, P21, P33, P41].
+
+\subsubsection{Process problems}
+
+The old way of working was acknowledged to have problems which needed to be
+corrected. A summary of weaknesses in the original process is presented in table
+\ref{table:reasonstochange_process}.
+
+\begin{table}[b]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        Problems in the old process           &  Described in  \\
+        \midrule
+        Excess and overhead in process     &  P9, P24, P35, P36, P38, P40, P46, P49, P51  \\
+        The process is being ignored       &  P6, P9, P46, P49  \\
+        Late integration, testing, and feedback  &  P5, P25, P33, P40, P50  \\
+        New practices can not be integrated in the old process  &  P10, P26  \\
+        The old process does not scale up  &  P16, P30, P50, P51  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Problems because of which the old process should be changed}
+    \label{table:reasonstochange_process}
+\end{table}
+
+A most typical known problem was overhead in process, which showed as bureucracy and
+needless costs. Examples on overhead in process were unproductive meetings
+[P38], sign-offs and gates [P9], excessive focus on conformance to process
+[P40], and change management overhead [P51]. Creating excessive documentation
+instead of business value was seen as a problem in some cases [P24, P35]. From
+business representatives point of view the processes were unreasonably costly,
+while at the same time insufficiently capable of delivering business value [P36,
+P46, P49]. In one case business realized that customers did not see any benefit
+in the heavy CMMI based process [P24].
+
 A heavy process resulted often in ignoring the process, only paying ``lip
 service'' to the process, or applying its steps retroactively [P6, P9, P49].
-Seffernick described how applying the waterfall process had teams ``still
-developing the same old way but slower because of the additional process''
-[P46].
+Seffernick [P46] describes how applying the waterfall process had teams ``still
+developing the same old way but slower because of the additional process''.
 
@@
 
-Yet another problem in managing development was problems with schedules.
-Some cases described schedules as unpredictable and estimation being difficult
-[P16, P30, P40, P51]. In other cases it was reported that release deadlines were
-often missed [P19, P21, P33, P41].
-
-Finally, the original process model was seen as a cause for quality problems.
-Several studies mentioned a demand for better quality, and quality improvement
-was set as a goal for the agile transformation [P25, P31, P41, P42]. Improvement
-in external quality and customer satisfaction were also anticipated from agile
-development [P2, P4]. Problems with technical debt were attributed to the
-original process. Some cases reported on prioritizing new feature development
-over bugfixes or fixing only the easiest issues [P30, P38].
-
-\subsubsection{Communication and collaboration issues}
+A few organizations were experiencing rise of agile methods at grass roots
+level, but the agile teams could not integrate efficiently with other parts of
+a waterfall organization. Tensions arose when teams using different development
+approaches attempted to work together on products [P10]. On a team level agile
+had enabled some flexibility, but that flexibility could not be leveraged on the
+scale of the entire organization [P26].
+This kind of experiences drove the organization to deploy agile on a full scale.
+
+Still one problem was that some growing organizations experienced that the old
+development method did not scale up [P16, P30, P50, P51].
+
+\subsubsection{Other reasons to change}
+
+There were still a number of problems that were given as reasons to change.
+In addition to the above mentioned problems, organizations were striving for
+better collaboration and quality. Table \ref{table:reasonstochange_other} lists
+other issues that drove organizations to change.
+
+\begin{table}[b]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
+                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
+        Reasons to change           &  Described in  \\
+        \midrule
+        Teams can not collaborate properly   &  P21, P38  \\
+        Silos and internal boundaries in organization   &  P27, P40, P46  \\
+        Lack of customer collaboration       &  P4, P26, P40  \\
+        Demand for better quality            &  P2, P4, P25, P31, P30, P38, P41, P42  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Other reasons why organizations wanted to change the way of working}
+    \label{table:reasonstochange_other}
+\end{table}
 
@@
 
-Other communication issues were related to requirements handling and difficulty
-to focus on the most important aspects of a delivery.
-Examples on problems with requirements was too much focus on the documenting
-aspect instead of providing understanding [P24, P49], and ignoring the fact that
-requirements and understanding are evolving during a project [P4].
-In some cases the organization of work and management was such that it hindered
-focusing the development [P7], and prioritization of what to work on was not
-done efficiently [P38, P39].
-
 Too little customer and user collaboration led to problems in some
@@
 
-A few organizations were experiencing rise of agile methods at grass roots
-level, but the agile teams could not collaborate efficiently with other parts of
-a waterfall organization. Tensions arose when teams using different development
-approaches attempted to work together on products [P10]. On a team level agile
-had enabled some flexibility, but that flexibility could not be leveraged on the
-scale of the entire organization [P26]. This kind of experiences drove the
-organization to deploy agile on a full scale.
+% There was also a desire to improve team motivation and commitment, and
+% collaboration between management and developers [P39].
+
+% Examples of problems with requirements was too much focus on the documenting
+% aspect instead of providing understanding [P24, P49], and ignoring the fact that
+% requirements and understanding are evolving during a project [P4].
+% 
+% In some cases the organization of work and management was such that it caused
+% lack of focus the development [P7], and prioritization of what to work on was
+% not done efficiently [P38, P39].
+
+Lastly, the original model was seen as a cause for quality problems.
+Several studies mentioned a demand for better quality, and quality improvement
+was set as a goal for the agile transformation [P25, P31, P41, P42]. New feature
+development was being prioritized over bugfixes, and technical debt was
+developing [P30, P38]. Improvement in external quality and customer satisfaction
+were anticipated from agile development [P2, P4].
+
+\clearpage
 
@@
 In a few cases the organization initailly lacked strict process. For instance
-Long describes the initial state in the following way: ``Project scope was in
-constant flux. Well-meaning market and product owners routinely inserted new
-requirements during product development.'' [P30].
+Long [P30] describes the initial state in the following way: ``Project scope was
+in constant flux. Well-meaning market and product owners routinely inserted new
+requirements during product development.''.
 
 Some authors describe the developement being originally organized in silos. For
-instance Ranganath describes that ``Teams operated in silos. Collaboration,
-shared purpose and ownership for outcomes were hard to find'' [P40].
+instance, Ranganath [P40] describes that ``Teams operated in silos.
+Collaboration, shared purpose and ownership for outcomes were hard to find''.
 

### 2014-05-19 journal/article.tex
Spellcheck

@@
 
---> This literature review is intended for researchers as a theroretical
+--> This literature review is intended for researchers as a theoretical
    and proven foundation for the real problems that organizations face in
    transformations.
-   This is also intended to be directly usable by practicioners, to serve as a
+   This is also intended to be directly usable by practitioners, to serve as a
    base of general advice with strong motivation (although every organization
@@
 transform into an agile way of working. The research was conducted as a
-systematic litterature review based on empirical studies.
+systematic literature review based on empirical studies.
 
@@
 in all facets. Even if a study discussed some of the irrelevant topics it could
-be included as long as it had some content matcing all the facets.
+be included as long as it had some content matching all the facets.
 
@@
         Large scale    &
-                At least 50 development personel or 6 teams uses agile &
+                At least 50 development personnel or 6 teams uses agile &
 
@@
 introduced \cite{Mishra2011}, and merely presenting agile tools in large-scale
-use \cite{Kim2012}. These sources were excluded as they did not privide insights
+use \cite{Kim2012}. These sources were excluded as they did not provide insights
 on organizational transformation.
@@
 The facet of \emph{large scale} was interpreted as presented in section
-\ref{sec:largescale}, having 50 development personel or 6 teams as as the limit
+\ref{sec:largescale}, having 50 development personnel or 6 teams as as the limit
 for large scale. In some cases the source presented very vague indicators on
@@
 borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
-no indication of issues indicating large size, althoug there were several
+no indication of issues indicating large size, although there were several
 development teams.
@@
 potential primary studies. The databases included in the search are listed in
-Table \ref{table:databases}. All databases supported use of complex boolean
+Table \ref{table:databases}. All databases supported use of complex Boolean
 logic in searches, which allowed us flexibility in constructing the search
@@
 of \emph{empiricality} could not sensibly be represented by keywords. We decided
-to include only the factets \emph{agile software deveolpment} and
+to include only the facets \emph{agile software development} and
 \emph{organizational transformation} in the keyword search, with the consequence
@@
             "roll-out", rollout}\\
-        Only software related atricles &
+        Only software related articles &
             \texttt{(software OR (conference="agile, xp, icgse, icse"))
@@
 papers were categorized independently by both Dikert and Paasivaara into three
-categories: include, excude, and uncertain. The result was combined by marking
-abstracts both researchers agreed on for inclusion or exclusion respecitvely.
+categories: include, exclude, and uncertain. The result was combined by marking
+abstracts both researchers agreed on for inclusion or exclusion respectively.
 The inclusion of disagreed or uncertain abstracts was resolved through
-discussion. At this stage papers were excluded only if both reserachers deemed
+discussion. At this stage papers were excluded only if both researchers deemed
 it clearly irrelevant, including any uncertain cases for full text filtering. As
@@
 organization was described in many studies all sources that passed the inclusion
-criterial were included. Studies presenting the same organization were treated
+criteria were included. Studies presenting the same organization were treated
 as one unit so that we could gain as much insight from each organization as
@@
 
-The reason for deviating stydy assessment
+The reason for deviating study assessment
 --> Original research consisted almost solely of experience reports
@@
 		Statements that describe the transformation process (top-down, big bang,
-		stepwise). \\
+		step wise). \\
 
@@
    Engineering and Extreme Programming} to the results of the keyword search.
-   These two conferences provided more than two thirds of the pimary sources.
+   These two conferences provided more than two thirds of the primary sources.
 
@@
 questions. The methods used for extracting the results were described in section
-\ref{sec:method}. Note that the term \emph{study} is used when refering to a
-primary study publication, and the term \emph{case} is used when refering to an
+\ref{sec:method}. Note that the term \emph{study} is used when referring to a
+primary study publication, and the term \emph{case} is used when referring to an
 organization which may be described in several studies.
@@
 but the issues discussed revealed that the organization was of large size
-accoding to our definition. The distribution of organization sizes is presented
+according to our definition. The distribution of organization sizes is presented
 in figure \ref{fig:organization_size}.
@@
 		Banking and financial services              &  4  \\
-		Healthcare                                  &  3  \\
+		Health care                                  &  3  \\
 		IT services                                 &  2  \\
@@
 largest group was enterprise management solution providers with products for
-business process management, project protfolio management, and facility
+business process management, project portfolio management, and facility
 management.
@@
         Business partners demand faster delivery  &  P5, P23, P39, P40, P46  \\
-        Need to accomodate change during projects  &
+        Need to accommodate change during projects  &
                 P4, P8, P18, P25, P26, P33, P36, P39, P42, P49, P51, P52  \\
@@
 
-Yet another factor pushing for improvement was the need to accomodate change
+Yet another factor pushing for improvement was the need to accommodate change
 during development. Organizations saw agile as an opportunity to improve
@@
 Some organizations faced difficulties when changes to plans were needed. For
-instance, the organizational struture made adapting to changes difficult [52],
+instance, the organizational structure made adapting to changes difficult [52],
 and making changes created management overhead [P51]. Other organizations saw an
@@
 
-A most typical known problem was overhead in process, which showed as bureucracy and
+A most typical known problem was overhead in process, which showed as bureaucracy and
 needless costs. Examples on overhead in process were unproductive meetings
@@
 Too little customer and user collaboration led to problems in some
-organizations. For instance, too late user involvment led to change requests
+organizations. For instance, too late user involvement led to change requests
 late in the project [P40]. There was a wish to collaborate more frequently and
@@
 
-In a few cases the organization initailly lacked strict process. For instance
+In a few cases the organization initially lacked strict process. For instance
 Long [P30] describes the initial state in the following way: ``Project scope was
@@
 
-Some authors describe the developement being originally organized in silos. For
+Some authors describe development being originally organized in silos. For
 instance, Ranganath [P40] describes that ``Teams operated in silos.
@@
 presented here do not explicitly relate to the transformation process, but are
-rather general difficulties in agile imlementation. These issues are still
+rather general difficulties in agile implementation. These issues are still
 included as settling the agile implementation can be seen as a part of the
@@
 anticipated in any change process. People will not be willing to change unless
-there are good reasons, and the change is preceived easy enough. It is difficult
+there are good reasons, and the change is perceived easy enough. It is difficult
 to attain a buy-in for a change, and even organizations that have a flexible
@@
 and cautious, which slows down any change. Any change was further hindered by a
-deeply rooted status-quo and high employee retention [P14]. In some cases it was
+deeply rooted status quo and high employee retention [P14]. In some cases it was
 reported that people worried about new roles and responsibilities that the agile
@@
 distrust in suitability of agile. Seffernick describes how management did on one hand
-acknowledge benefits of agile, but on the other explained away the possibilitiy
+acknowledge benefits of agile, but on the other explained away the possibility
 to apply it due to contract reasons, the current matrix organization, and other
@@
 show. In several cases the change initiative came from management, but when
-presented wrong the initiative becomes a mandate that pepole are not receptive
+presented wrong the initiative becomes a mandate that people are not receptive
 for. Spayd [P49] summarizes this as: ``Organizations do not change merely
@@
 starting the transformation, and the understanding of agile development overall
-[P2]. O'Connor describes how team members became sceptical towards agile when
-implementing mandaetd changes did not bring any visible benefits [P37].
+[P2]. O'Connor describes how team members became skeptical towards agile when
+implementing mandated changes did not bring any visible benefits [P37].
 A problem relating to top down mandate is that if management does not speak out
@@
 also be opposite. In cases where the change was emerging bottom up the
-reluctancy to change was not in the team members but management, causing that
+reluctance to change was not in the team members but management, causing that
 significant organizational change could not be raised above the team level. For
@@
 for the transformation. Table \ref{table:challenges_lackofinvestment} presents
-investments and resourcing arrangements that were witheld, and which caused
+investments and resourcing arrangements that were withheld, and which caused
 problems in transformation.
@@
 Training and coaching are direct investments in transformation and their lack is
-an evident problem. These activities were still underbudgeted, which
+an evident problem. These activities were still under budgeted, which
 created difficulties in the transformation [P11].
@@
 how dedicated rooms for teams could not be arranged, and team events such as
-daily standups required booking a conference room [P29]. In another case the
+daily stand-ups required booking a conference room [P29]. In another case the
 estate organization shut down the teams' attempts to modify working spaces
@@
 a solution to all problems [P9]. Success was declared prematurely [P49] but
-expectations created by successful early experiments were not fullfilled [P38].
-Disapointment was granted when these hopes were not fullfilled, and instead
+expectations created by successful early experiments were not fulfilled [P38].
+Disappointment was granted when these hopes were not fulfilled, and instead
 implementing the agile model was found to be hard and time consuming.
@@
 Schnitter and Mackert [P45] report that theoretical considerations on how to
-scale up the agile methos were not good enough, and that product managers and
+scale up the agile methods were not good enough, and that product managers and
 architects struggled when several Scrum teams were working concurrently.
@@
 Further indications were given that the advice from literature will not ferry
-all the way. Benefield [P6] describes how it was dificult finding a balance
+all the way. Benefield [P6] describes how it was difficult finding a balance
 between prescribing a by-the-book implementation which may put people off, and
@@
 The initial success was followed by challenges when teams needed to work with
-other teams and as parts of the larger surrounding orgaization [P13, P17, P24].
+other teams and as parts of the larger surrounding organization [P13, P17, P24].
 Introducing agile had created flexibility on a team level, but it could not be
 fully utilized as the surrounding organization was not responsive enough [P26].
-The rollout of agile had not removed dependencies, and the dependencies
+The roll-out of agile had not removed dependencies, and the dependencies
 made managing development difficult [P9, P10].
@@
 communication channels narrowed and created communication breakdowns. Problems
-with collaboration, dependencies, and integration were solved in the presnce of
+with collaboration, dependencies, and integration were solved in the presence of
 individuals having a proactive and open mindset towards working over team
@@
 Problems may arise in a large environment if the schedule of agile adoption
-varies from team to team. Internal fragmenttation is a further problem that may
+varies from team to team. Internal fragmentation is a further problem that may
 be encountered in large organizations.
@@
 procedures producing excess documentation and stepping through approval gates
-[P20, P21]. The bureucracy of the previous traditional model was still enforced,
+[P20, P21]. The bureaucracy of the previous traditional model was still enforced,
 and management was not convinced to lighten the process. Another case describes
@@
 and people were often relocated to match the needs of the moment. Too much
-reorganization made it diffucult for teams to plan ahead [P11].
+reorganization made it difficult for teams to plan ahead [P11].
 
@@
 change [P2, P18, P31]. A tension was growing between development and broader
-roles in the organization [P1]. The full beenfits of the transformation would
+roles in the organization [P1]. The full benefits of the transformation would
 not be attained unless the entire organization was set to work along the same
@@
 
-Agile affects timings in the software delivery lifecycle, what caused two types
+Agile affects timings in the software delivery life cycle, what caused two types
 of problems. Firstly, the iterativeness of agile models was problematic when
@@
 fit well into the time frame of a sprint. The UX team was overloaded and
-discontent, but the situation was fixed by refining the schedule for interations
+discontent, but the situation was fixed by refining the schedule for interactions
 with development teams. [P15, P28]
@@
 Problems in adopting agile were even showing in functions outside the software
-delivery lifecycle. In several cases human resouces (HR) was a support function
+delivery life cycle. In several cases human resources (HR) was a support function
 mentioned to work against the agile adoption. In order to gain full results of
@@
 
-Development is traditionally preceeded by requirements engineering and succeeded
+Development is traditionally preceded by requirements engineering and succeeded
 by quality assurance (QA) activities. Agile methods suggest that essential parts
 of these activities would be performed within the development team. However,
-when developming large systems the effort required for managing requirements and
+when developing large systems the effort required for managing requirements and
 QA may be too heavy for individual teams to handle. Table
@@
 complexity of the product will require additional layers of product management
-[P45], requirements are created by several stakehoder groups and development
+[P45], requirements are created by several stakeholder groups and development
 teams can not be in contact with all of them [P39], or teams do not have access
-to stakeholders due to a distributed seetting [P24]. Agile methods need to be
+to stakeholders due to a distributed setting [P24]. Agile methods need to be
 extended to include a process on how to get large requirements into team
@@
 management delivered requirements in too large chunks resulting in development
-teams spending large amounts of time on defining suitaby sized stories.
+teams spending large amounts of time on defining suitably sized stories.
 
@@
 In an enterprise setting quality assurance tasks will require effort beyond the
-general agile convention of doing testing within a crossfunctional team.
+general agile convention of doing testing within a cross-functional team.
 Similarly as with requirements engineering, the agile model may need to be
-extended to accomodate additional testing activities [P31]. Organizations may
+extended to accommodate additional testing activities [P31]. Organizations may
 have existing separate QA teams that must be coordinated to work with
@@
 
-A phenomena that troubled some organizations was overenthusiasm towards agile
+A phenomena that troubled some organizations was over-enthusiasm towards agile
 methods. In several cases there was mentionings about change leaders or teams
@@
 * Compare how recommendations in large-scale application models of agile relate
-  to the challenges and success factors reported in the transformaiton proecss.
+  to the challenges and success factors reported in the transformation process.
 

### 2014-05-19 journal/article.tex
Minor edit

@@
 
-Examples on topics excluded by the first facet are agile manufacturing, and
-applying Scrum practices in management boards, as these do not relate to
-software engineering. In addition, we required that the organizational
-transformation was aimed to introduce agile methods, which excluded other
-development methods than agile \cite{Sagesser2013}, and use of agile
-methods in other contexts than software development \cite{Hodgkins2007}.
+Examples on topics excluded by the facet of \emph{agile software development}
+are agile manufacturing and applying Scrum practices in management boards, as
+these do not relate to software engineering. In addition, we required that the
+organizational transformation was aimed to introduce agile methods, which
+excluded other development methods than agile \cite{Sagesser2013}, and use of
+agile methods in other contexts than software development \cite{Hodgkins2007}.
 

### 2014-05-19 journal/article.tex
New title

@@
 
-\title{Adopting agile in the large: motivations, success factors and challenges}
+\title{Adopting agile software development in the large organizations:
+       A systematic literature review}
 
@@
 
-Some success factors ...
+The analyzed primary studies brought up several factors that contributed in
+success in the transformation. As stated by Research Question 4, we studied what
+factors were reported to facilitate the transformation process. We present
+factors that the primary studies particularly highlighted as being important in
+advancing the transformation process.
+
+* What is our standpoint on recommendations that experience reports give, but
+  which were not necessarily tried in the organization?
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ l l }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Training is necessary to achieve change  &  P3, P6, P14, 16, P30  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of investment and resourcing factors that affect
+             transformation negatively}
+    \label{table:challenges_lackofinvestment}
+\end{table}
+
+\subsubsection{Training and coaching}
+
+Training was seen as a strong factor helping the organization to transform.
+Several studies stated that training improved the chance of succeeding in the
+transformation [P3, 16, P30]. It was even highlighted that the change would have
+failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
+and again how training could make the difference between success and failure for
+a team''. Training also helped people to become more positively inclined towards
+the new model [P16, P37], and at best training made people enhusiastic to change
+[P3, P46]. A few cases reported that there was some disinclination to change,
+and training would have helped to overcome [P11, P14].
 

### 2014-05-19 journal/article.tex
Training and coaching

@@
 
+\subsubsection{Training and coaching}
+
 \begin{table}[h]
     \centering
-    \begin{tabular}{ l l }
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
         \toprule
@@
         \midrule
-        Training is necessary to achieve change  &  P3, P6, P14, 16, P30  \\
+        Training is necessary to achieve change  &
+                P3, P6, P11, P14, P16, P30, P37, P46 \\
+        Coaching is essential in transformation  &
+                P3, P6, P8, P9 P16, P17, P18, P30, P42, P44, P47, P48  \\
         \bottomrule
     \end{tabular}
-    \caption{Summary of investment and resourcing factors that affect
-             transformation negatively}
-    \label{table:challenges_lackofinvestment}
+    \caption{XXXXXXXXXXXXXX}
+    \label{table:success_NN}
 \end{table}
-
-\subsubsection{Training and coaching}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Agile methods avoid prescribing exact ways of working but rather emphasize a
+mindset of adapting to the situation. It is difficult to explain by theory how
+such a mindset should be applied, and the agile practices are best learnt by
+doing. Coaching teams while they apply the agile methods in practice was seen as
+an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
+that coaching was essential for success in transformation [P8, P18]. Also,
+without coaching teams can do damage to the agile transformation if techniques
+are applied improperly [P9].
+
+There were a number of statements of the benefits of coaching. The coach
+position allowed to watch for and correct problems when they arose [P16, P42].
+Coaches also helped to draw attention away from focus on tools and towards
+understanding the principles of agile [P42, P48].
+There were also benefits in using both internal and external coaches.
+External coaches were able to have an objective view of the organization [P16,
+P44], but internal coaches were more accessible and knew the specifics of the
+organization [P3, P6].
+
+\subsubsection{Choosing the agile model}
+
+The choice of agile practices did affect the success of the transformation.
+Also how the practices were choosen affected.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-05-19 journal/article.tex
Choosing the agile model well

@@
 
-\subsubsection{Agile is found to be hard}
+\subsubsection{Agile is difficult to implement}
 
@@
 
+* Some success factors are just calculated decisions that were made because
+  the organization saw them as the best way to proceed. The question whether
+  the decision really were critical for success is not proven.
+
 \subsubsection{Training and coaching}
@@
 
-The choice of agile practices did affect the success of the transformation.
-Also how the practices were choosen affected.
+The choice of agile practices did affect the success of the transformation, and
+also how the practices were choosen affected.
+
+% Mapping agile to the old model
+
+Mapping the agile model to match the old model was seen as necessary in a few
+cases. Although a general mapping to the old model may not be good [P44, P46],
+some cases needed to preserve high level management practices [P14, P29, P48].
+By allowing management to remain unchanged it was possible to introduced agile
+stepwise on team level, and helped in getting management buy-in for the process
+[P14, P29, P48].
+In one case Scrum was considered as a wrapper for existing practices, and could
+therefore be easily fit in the organization [P44]. The agile methodology was
+considered complementing the existing culture, which eased management buy-in for
+the new method [P44].
+% Similarities and differences of agile were documented, and Scrum terminology was
+% explained in terms of the old model. Mapping of management information was done
+% so that for instance the information earlier found in Engineering Gantt charts
+% was found in the Sprint Backlog. [P29]
+
+% Conformity in the agile model
+
+Some studies reported that conformity was considered in implementing the new
+model. A consistence common vocabulary was seen to benefit the organization and
+support the change [P8, P16, P43]. Other benefits in using a single model were
+possibility to compare work between teams, it was easier for people to relocate,
+and progress was predictable for stakeholders [P43]. Consistency and common
+understanding was created in discipline meetings and public events, or by peer
+coaching [P11, P12, P32].
 

### 2014-05-19 journal/article.tex
Success: Choosing the agile model

@@
 
+% Keep things simple
+
+One advice given by few cases was to keep the organization simple. Beavers [P5]
+describes that attempting to operate in a complex and global setting complicated
+even simple agile practices. The organizational chart was simplified, which was
+welcomed by both employees and managers [P5]. Rather than focusing on detail in
+process, communication practices, and tools the focus should be on engaging the
+teams [P40].
+
+% Emphasize principles over mechanics
+
+In a number of studies it was described that during transformation the
+principles of agile should be emphasized over practices and simple mechanics
+[P46, P48]. When people understand the agile values they will also understand
+why the chage is being done and feel motivated [P1, P16, P42].
+The people centric nature of agile development should be communicated correctly.
+McDowell and Dourambeis [P32] describe how the organization launched a series of
+communications events especially designed to emphasize the reasons behind the
+agile practices.
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ---------------------------------------------------------------------
+
+% ??
+In order to draw the most out of agile teams should innovate and find on
+practices that really work for their case [P40, P41]
+
+% ??
+Especially inexperienced coaches and Scrum masters made mistakes in focusing too
+much on the implementation of practces [P23, P30].
 

### 2014-05-20 journal/article.tex
Success: Customizing

@@
 --> Introductory paragraph
+
+-- Who will be interested of this?
+
+-- What can be learned from this`?
 
@@
         \midrule
-        How to accommodate high level requirements management  &  P24, P31, P39, P45  \\
+        How to accommodate high level requirements management  &  P24, P31, P39, P45, P51  \\
         Problems with breakdown of requirements        &  P1, P17, P33, P48  \\
@@
 Product managers and business analysts were struggling in creating high level
-requirements [P33, P37], and teams were struggling in breaking them down to an
+requirements [P33, P37, P51], and teams were struggling in breaking them down to an
 estimable size [P33]. Smits and Rilliet [P48] describe how high level product
@@
 
-One advice given by few cases was to keep the organization simple. Beavers [P5]
-describes that attempting to operate in a complex and global setting complicated
-even simple agile practices. The organizational chart was simplified, which was
-welcomed by both employees and managers [P5]. Rather than focusing on detail in
-process, communication practices, and tools the focus should be on engaging the
-teams [P40].
+One advice given by few cases was to keep the organization and process simple
+[P10, P16]. Beavers [P5] describes that attempting to operate in a complex and
+global setting complicated even simple agile practices. The organizational chart
+was simplified, which was welcomed by both employees and managers [P5]. Rather
+than focusing on detail in process, communication practices, and tools the focus
+should be on engaging the teams [P40].
 
@@
 
+% Customizing the agile model
+
+Customizing the agile model and practices was reported as an necessary step in
+the agile implementation. As each organization will have its own challenges in
+adopting agile, it should be carefully considered which organization specific
+areas to focus on when choosing the agile practices to implement  [P8, P43].
+Lewis and Nehrer [P29] recommended to choose the agile model according to the
+current corporate model in order to avoid interference.
+
+Especially in a large organization it is not feasible to use the same process
+for all projects [P49]. An individual application of the agile process is needed
+for different types of development, depending on for instance the type of
+software being developed or the project size [P45].
+Applying agile at scale will require to deviate from some of the typical
+recommendations [P31]. However, it is essential to remember the agile principles
+when doing customizations, and watch out for customization that contradict with
+the principles [P31].
+
+Cases reported that customization was part of a successful agile implementation.
+For instance, teams were let to customize their practices individually, to fit
+the needs of each team [P41].
+Spayd [P49] writes that teams who modified agile practices to suite te large and
+distributed environment ended up doing better than teams that did not.
+% Some intermediate point?
+To allow teams to become innovative and perform well the agile model should be
+customized in a pragmatical way instead of following a strict textbook
+interpretation [P10]
+
+The customization of the agile model was also seen as an evolutionary process.
+Cowan [11] suggested that in a big-bang transformation it might be necessary to
+selectively compromise some parts of the agile implementation so that the core
+practices can be set in place.
+The agile transformation is a constantly ongoing process, and it is recommended
+to regularly challenge teams to refine the agile implementation to reflect the
+courrent needs of the organization [P41, P43].
+
+% Customize the agile model?
+% ??
+% [P16]
+% This cross-functional team redesigned and rebuilt the development process from
+% the ground up using key values from the company’s founding: KISS (Keep it Simple
+% Stupid), iterate quickly, and listen to our customers. These values are a
+% natural match for agile methodologies.
+% 
+% ??? There are changes done to the agile proces, but doing it is not that much emphasized
+% [P51]
+% One after the other, all products where then switched to the new Scrum process.
+% Along the way, minor adjustments were often made. The teams were constantly
+% searching for the right balance between sprint duration, size estimation,
+% requirement size, and time spent on backlog administration. Changes to the
+% process were initially often managed by the the CTO. However, they were
+% increasingly delegated to the appropriate ScrumMasters.
+% Comment: This is not directly seen as a success factor.
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+* Some success factors and challenges are ``symmetrical'', so that lack of some
+  factor is a challenge, but availability of it provides success
+
 * Success factors and challenges may be complements, and may be presented as
@@
 * What is special in large scale?
+
+* Customization of practices was seen as necessary, as large scale has demands
+  over the original agile ideas
 

### 2014-05-20 journal/article.tex
Success: Requirements handling

@@
 * Definition of sprint
+
+* Definition of product owner
 
@@
 
+\subsubsection{Considering other parts than development}
+
+% Improve requirements handling
+
+Some cases reported on challenges in requirements management, especially
+highlighting difficulty in managing the gap between high level requirements and
+stories handled by teams. It was stated that it is important that the
+requirements are consise and small enough for the teams to handle [P5, P11].
+It was recommended to invest in teaching skills in breaking down and writing
+user stories [P11, P33, P48]. Gat [P17] describes how the gap between
+development and requirements management was bridged by having a dedicated
+Requirements Architect to help with requirements breakdown.
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-20 journal/article.tex
Success: Role of POs

@@
 
-Some cases reported on challenges in requirements management, especially
+Some cases had reported on challenges in requirements management, especially
 highlighting difficulty in managing the gap between high level requirements and
@@
 
+% Emphasize role of Product Owners
+
+A particular role that seemed to make much difference in the agile introduction
+was the Product Owner. Many cases reported on problems if the Product Owner did
+not perform adequately, and it was seen as critical to have a dedicated person
+in that role [P12, 14, P30, P39].
+Success or problems emerged respectively depending on how well the Product
+Owners were performing. In one case a well implemented Product Owner role was
+understood to be a key success factor when using agile methods [16, P33].
+It was reported that when the Product Owner role was implemented correctly the
+team performed better and the work products were correct [P24]. Projects started
+off better when the Product Owners were engaged early on, and some Product
+Owners even became agile enthusiasts when they learned how business and
+technology can collaborate in the agile model [P46].
+
+Several cases endorsed training and coaching for the Product Owner role [P16,
+P39, P46]. Product Owners should receive training on agile principles, backlog
+management, user story breakdown, and agile planning [P16]. Also tool support
+for backlog management and two way communication between Product Owners and
+teams should be worked on [P39].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-21 journal/article.tex
Success: Commitment and urgency

@@
 
+% ??
+In order to draw the most out of agile teams should innovate and find on
+practices that really work for their case [P40, P41]
+
+% ??
+Especially inexperienced coaches and Scrum masters made mistakes in focusing too
+much on the implementation of practces [P23, P30].
+
 \subsubsection{Considering other parts than development}
@@
 
+\subsubsection{Urgency and commitment to change}
+
+% Make it understood that there is no return to the old model
+
+To make a large organization change the direction of the change needs to be
+communicated clearly. Atriculating urgency at start will help getting the change
+under way [P49]. Feedback of changes should certainly be listened to, but in the
+end it must be made clear that the old model can not be returned to [P22].
+For a case where the organization was changed all at once it was reported that
+the magnitude of change helped created the impression that the change was non
+negotiable [P11]. Firstly, as it vas evident that a large change was to happen
+people felt encouragement and permission to move on from the old way, and
+secondly, the urgency eliminated wasteful discussions on wheter Scrum was a good
+model or not [P11].
+When the organization culture is ingrown he change vision must constantly be
+reminded of, and each step towards the new model consider a victory [P48].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ---------------------------------------------------------------------
-
-% ??
-In order to draw the most out of agile teams should innovate and find on
-practices that really work for their case [P40, P41]
-
-% ??
-Especially inexperienced coaches and Scrum masters made mistakes in focusing too
-much on the implementation of practces [P23, P30].
 

### 2014-05-21 journal/article.tex
S: Commitment

@@
 
+% Commitment (?)
+
+A large change will inevitably require strong commitment. The agile model is
+introduced because of problems in the old model, and therefore there will be
+organizational issues uncovered during the transformation [P10]. People must not
+be demoralized when facing challenges, and the determination to change must be
+maintained [P8, P10]. It is not the agile model that has flaws, but it is
+merely exposing problems in the organization [P10].
+Ryan and Scudiere [P43] describe that management commitment was showing as
+strong focus on certain agile practices that had been chosen as non negotiable,
+and constant assessment that the practices work. The expectations were clearly
+communicated to the teams, and constant assessment made it clear that change is
+desired [P43].
+A strong commitment from management will assure teams that the change is the
+right thing [P11].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ---------------------------------------------------------------------
+
 \section{Discussion}

### 2014-05-21 journal/article.tex
S: Piloting

@@
 
+% Make change necessary
+
+Spayd [P49] describes that senior management pushed hard on a mandate to have
+deliveries every 90 days. The mandate made change necessary, and while the
+strong pressure did not promote agile practices in every case the drive for
+change was perceived good in general [P49].
+
+\subsubsection{Create a positive start by piloting}
+
+% Have a successful pilot first
+
+The agile model was usually tried out in the organization before rolling it out
+on a general level. Having a pilot project was reported as a significant success
+factor in several cases [P3, P9, P38].
+The pilot project gave proof that the agile model would be suitable for the
+organization and the general acceptance of agile was increased [P3, 17].
+The pilot also cleared disbeliefs of the suitability of agile for large
+organizations, and created acceptance for both development and management [P31,
+P36].
+
+Pilots were especially important for gaining management acceptance. Cases
+reported that management gave approval for agile methods only after seeing
+successful pilots [P14, P47]. Prokhorenko [P39] also describes that seeing the
+example of successful pilots will make managers from other departments eager to
+try the new method, and thus help to spread the use of it.
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
 

### 2014-05-21 journal/article.tex
S: Piloting

@@
 The pilot also cleared disbeliefs of the suitability of agile for large
-organizations, and created acceptance for both development and management [P31,
-P36].
+organizations, and created acceptance for both development and management [P9,
+P31, P36].
 
@@
 
+% Pilot provides information on how to do the rest of the transformation
+
+A particular benefit in piloting was that it provided knowledge on how to fit
+the agile method to the particular organization.
+A pilot project enabled the organization to get feedback on how teams and
+managers are best introduced to the new methods [P8, P22].
+Organizations met challenges in the pilot projects. However, the pilot projects
+served as valuable learning experiments providing insight in how to mitigate
+problems when the rest of the organization is transforming [P27, P40, P45].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-21 journal/article.tex
Work in progress

@@
 
+S: Positive experience helps spreading transformation
+
+* A positive experience created word-of-mouth that was spreading in the company
+* Other teams took interest  --> Experience from pilot program
+[P6]
+
+* A survey was created --> The survey showed that Scrum was preferred over the old model
+* This helped other teams to start with agile
+[P6]
+
+* The benefits of agile are learned
+* Persons who have experenced success will work for long term establishment of agile
+[P9]
+
+* Failure of the old model was contrasted with success of Scrum
+* The company decided to invest in agile
+[P19]
+
+* Positive feedback from pilot encouraged the entire organization to change
+[P22]
+
+* The teaching event was designed to be engaging
+* For the interest to be maintained it was an important factor that people had fun in training
+* The agile practices were chosen tactically --> Even sceptics were convinced at the end
+[P32]
+
+* The successful involvment of agile laid a foundation for further work
+[P36]
+
+* The organization saw at first hand that agile can benefit them (while the agile consulting company delivered)
+* Management became interested
+* There was some resistance --> Success mitigated it
+* Without seeing at first hand how agile was beneficial the organization would
+  never had started a transformation
+[P36]
+
+* Positive results will create interest --> The new model will spread
+[P39]
+
+* Positive word-of-mouth led to more people applying for agile project positions
+[P40]
+
+* Positivity overall is important in change
+* Leaders should especially show positivity
+* Celebrate even trivial successes
+* Encourage toexperiment and learn from failure
+[P40]
+
+* Make benefits visible, publish success stories --> The transformation will spread
+* The organization will better move towards the common goal
+[P42]
+
+* Showing good results made other teams to catch on agile
+[P46]
+
 % ---------------------------------------------------------------------
@@
 
+* The primary studies were quite openly pro-agile -- how large a factor of bias
+  is this?
+
 % ---------------------------------------------------------------------

### 2014-05-21 journal/article.tex
Having a positive experience

--- 2014-05-21 journal/article.tex
@@
 
+% Having a positive experience
+
+It is clear that any positive results will facilitate an organizational
+transformation. Although a recipie to fabricate positive experiences can not be
+given, many cases described how they had succeeded. Benefield [P6] describes how
+a survey was done to objectively measure satisfaction in agile, and the positive
+results helped teams make the decision to change their way of working. Another
+case describes how the corporate agile awareness and teaching event was designed
+to be fun and engaging, which made people take more ethusiasm in applying the
+agile practices [P32]. The move towards agile is assisted by making any benefits
+publicly visible and celebrating even the small victories [P40, P42].
+
+The dynamics of positive results were explained in further ways. Several cases
+highlighted that the agile transformation was spread effectively through
+positive word-of-mouth [P6, P40]. When good results were shown by a team it
+created interest in others, and enthusiasm to try the new model will spread
+[P39, P46]. Some companies were using agile and waterfall methods side by side.
+This setting made it comparison possible, bringing up the benefits of the agile
+model [P19, P36].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-21 journal/article.tex
S:  Having a positive experience

@@
 a survey was done to objectively measure satisfaction in agile, and the positive
-results helped teams make the decision to change their way of working. Another
-case describes how the corporate agile awareness and teaching event was designed
-to be fun and engaging, which made people take more ethusiasm in applying the
-agile practices [P32]. The move towards agile is assisted by making any benefits
+results helped teams make the decision to change their way of working. Benefield
+describes further how management was requesting proof on better performance as a
+response to requests to increase coaching and training budget. After conducting
+inquiries the coaching team was able to present an estimate of a 30\% increase
+in performance, which made management highly convinced [P6]. Another case
+describes how the corporate agile awareness and teaching event was designed to
+be fun and engaging, which made people take more ethusiasm in applying the agile
+practices [P32]. The move towards agile is assisted by making any benefits
 publicly visible and celebrating even the small victories [P40, P42].

### 2014-05-22 journal/article.tex
S: Communicating the change

@@
 
+* Some success factors are compiled from ``what we learned'' or ``advice for
+  others'' lists. We considered whether the advice was something the
+  organization had tried in practice.
+
 \subsubsection{Training and coaching}
@@
 
+\subsubsection{Communication and transparency in the change}
+
+Intensive communication was emphasized in a number of studies. It is important
+to reach a maximum number of people throughout the organization, as without
+communiction the a new model will not take root [P16, P43].
+It was recommended that use of the new model is made highly visible on many
+communication channels and even over-communicated [P16, P33, P41]. Mencke [P33]
+summarizes the viewpoint on over-communication: ``Even if you think that your
+teams understand a new method or process, repeat your communication to be
+sure.'' Workshops, coaching sessions, online discussions, and newletters are
+examples on different communication formats used [P41]. Another approach in
+communicating the change was that managers explained and encouraged change in an
+extensive series of one-to-one discussions [P22].
+
+Enabling transparency during the transformation was another communication
+related theme that was brought up, and even highlighted as a critical factor for
+success [P11, P16]. Fry and Greene [P16] underline the importance of
+transparency and sharing of information: ``This bias to sharing information with
+everyone was critical in our ability to adapt on a daily basis to ensure our
+success.'' Transparency was achieved by showing both successes and challenges
+[P6], actively reaching out for feedback [P6], using project management tools to
+display project status publicly [P11], by rearranging physical spaces [P42], and
+by holding the meetings of the rollout team in public [P16].
+By sharing experiences and status of the transformation the organization was
+aligned and everyone was moving in the same direction [P42, P43].
+
+% The inertia and fear of sharing information must be overcome [P33].
+
+% Sharing everything with everyone was key to success [P16].
+
+% Organizations made also an effort to communicate the status of the
+% transformation [P16, P43].
+
+% Because of the volatility of software development there must be continuous and
+% fluent communication between managers, POs, and teams [P42].
+
+Communicating the goals of the transformation was highlighted in a some studies.
+Having a clear message of expectations helped to remove confusion and make
+people understand the purpose of the transformation [P43, P48]. The motivation
+to change can be increased by having a simple proposal on how to reach the
+desired outcomes [P40].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-22 journal/article.tex
Work in progress

@@
 
+Many cases highlighted how people were engaged broadly in the organization.
+One of the lessons learned presented was that it is important to involve all
+stakeholders to gain acceptance in the transformation [P4, P31]. The transition
+team made an effort to engage people by both inviting to give feedback online
+and holding an extensive number of feedback meetings [P6, P22, P16]
+Being inclusive towards everyone was seen as one of the key success factors in
+the transformation [P16]. Inclusiveness will encourage people to participate and
+use the new model visibly [P16].
+
+\subsubsection{Have a change leading coalition}
+
+Organizing a large number of people requires some management and leadership.
+Similarly, transforming the way of working of a large group requires coordinaion
+and leadership [P42, P31]. In addition to the leadership provided by coaches,
+specific change leaders were also mentioned. Cases indicated the importance of
+having spokespersons for the change [P3, P4, P31]. Cowan [P11] describes how one
+person was strongly driving the transformation, and made an indisputable
+contribution for transforming the organization. Also Maples [P31] describes how
+the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
+described that the change was lead by a competent rollout team, which had
+representatives from all parts of the organization [P16, P33]. Finally, the
+responsibility of line and project managers to act as change leaders was
+highlighted [P18].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
 
+The formation and influence of agile communities had a significant impact on
+transformation.
+
 % ---------------------------------------------------------------------
+
+* A success factor: Agile communities emerged
+* Communities are commonly created in agile initiatives
+[P3]
+
+* There was an agile community which spawned eager followers
+* The community shared experiences and advice
+* The community emerged self-organized
+[P12]
+
+* Project managers were in a matrix structure
+* The matrix enabled sharing of learnings and confirmation 
+[P12]
+
+* The organization is planning on establishing an agile CoP
+[P30]
+
+* The coaches built up an agile community
+* There was online discussions, newsletters, and workshop events
+[P41]
+
+* The company has communities of practice, which are effective in raising awareness
+  of engineering practices
+[P42]
+
+* The agile coach community is a driving force --> They achieved becomming a driving force
+* The community created practices promoting the use of agile methods in the organization
+[P47]
+
+* A collective group is required to make chage in an organization
+* Impediments may be too large for individuals to move
+[P47]
+
+S: Have an agile community
 

### 2014-05-22 journal/article.tex
S: Communities, alignment

@@
 
+% * Create a sense of urgency
+% * Find a common cause to motivate the change
+% [P11]
+% Find a rallying point.  You might not have a recession as a spur to receptivity,
+% but a common, critical cause to motivate change will make it easier to explain
+% how Scrum can contribute to the success  of that cause.
+
 \subsubsection{Create a positive start by piloting}
@@
 
+The formation and influence of agile communities was reported to have a
+significant impact on transformation.
+The emergence of an agile community was reported as a success factor in the
+transformation [P3]. A community was also seen as indispansable as it has power
+to overcome impediments that would be too large for individuals to affect [P47].
+The agile communities achieved to raisie awareness of agile methods in the
+organization [P42, P41, P47], and spawned eager followers who spread the word
+even further [P12].
+
+A factor in achieving change was creating alignment towards the common goal of
+introducing a new development method. It was seen as an important factor that
+all levels in the organization speak the same language and accept the change
+[P18, P31]. Alignment was built by promoting success stories and gathering lists
+of problems to tackle [P42, P46]. In one case complete alignment between teams
+and within management was considered as necessary to use the agile model in a
+large context [P43]. Another case highlighted creating and applying a structured
+rollout process to uniformly introduce a large nuber of teams to agile [P8].
+
+\subsubsection{Management support}
+
+The most prominent and clear success factor was having support from management.
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-23 journal/article.tex
S: Management support

@@
 
+% Management support in general
+
 The most prominent and clear success factor was having support from management.
+Numerous cases made it clear that management support is an absolute necessity
+[P8, P18, P27, P31, P33, P36, P43, P47, P49].
+This was reflected in statemets such as ``Adopting agile, or implementing any
+significant change, requires an executive’s sincere support.'' [P44],
+``Executive commitment was crucial to implementing massive change.'' [P16], and
+``Having upper management engaged, supportive and visible is critical for
+wholesale organization involvement with Scrum.'' [P11].
+
+Managers were seen to be in a key role in making the change stick, as they had
+the authority and power to remove impediments [P18]. Seffernick [P46] describes
+how a number of people attempted to explain away the applicability of agile
+methods, but the objections were overruled by the director's commitment to make
+agile work. Management support was similarly needed when tight release schedules
+had to be flexed in order to give room for the adoption process [P16, P27].
+Favorable management decisions were also critical when additional resources were
+allocated to training and coaching [P47].
+
+Visible involvment of management was also reported to motivate and encourage
+employees to adopt the new model [P40]. For instance, the CTO organized training
+sessions personnally [P24], and the engineering VP frequently visited sprint
+demos [P11]. When the corporate level support for the agile initiative was
+showing teams adopted agile methods even spontaneously [P9].
+
+There were a few suggestions on how to involve management better. Foremost, a
+successful experience will likely rise the interest of management [P36].
+Mencke [P33] suggests giving executives tasks related to the agile process,
+so that they have better visibility of the new way of working.
+
+% Educate management
+
+In order to gain support for agile several cases highlighted the need to educate
+managment [P8, P47].
+Without education managers might make harmful interventions in the process, or
+just stick to the old model [P16, P35, P46].
+Managers not understanding the principles of agile were feeling left out with
+the introduction of self organizing teams, which some times resulted in
+backlashes [P6]. Providing proper training was reported to correct the
+situation, and even create strong agile supporters in management [P6, P9].
+Training also cleared misconceptions that management had [P29], and helped
+create a consistent implementation of the agile model across the organization
+[P11].
 
@@
 
-The formation and influence of agile communities had a significant impact on
-transformation.
-
 % ---------------------------------------------------------------------
-
-* A success factor: Agile communities emerged
-* Communities are commonly created in agile initiatives
-[P3]
-
-* There was an agile community which spawned eager followers
-* The community shared experiences and advice
-* The community emerged self-organized
-[P12]
-
-* Project managers were in a matrix structure
-* The matrix enabled sharing of learnings and confirmation 
-[P12]
-
-* The organization is planning on establishing an agile CoP
-[P30]
-
-* The coaches built up an agile community
-* There was online discussions, newsletters, and workshop events
-[P41]
-
-* The company has communities of practice, which are effective in raising awareness
-  of engineering practices
-[P42]
-
-* The agile coach community is a driving force --> They achieved becomming a driving force
-* The community created practices promoting the use of agile methods in the organization
-[P47]
-
-* A collective group is required to make chage in an organization
-* Impediments may be too large for individuals to move
-[P47]
-
-S: Have an agile community
 

### 2014-05-23 journal/article.tex
Work in progress

@@
 
+The agile practice of giving teams the power to decide was seen as an important
+factor in advancing the transformation.
+
 % ---------------------------------------------------------------------
+
+* Management could not solve the issue of how to implement daily builds
+* The teams themselves were let to figure out the technicalities
+* The teams solved the technical problem management could not dictate a solution to
+* It was a relevation for management that the teams were the best in solving the problem
+[P5]
+
+* Management thought initially that command and control is the way to go
+* Pressure were put in points where meters were showing weakness
+* A better approach was to provide the direction and let teams figure out
+[P5]
+
+* Allowing self-organization is critical in order to create commitment and engagement
+[P16]
+
+* Management drives the change
+* When teams are given new responsibilities they will be naturally motivated to continue using the practices
+[P18]
+
+* It was seen as good that teams are allowed to personalize their work area
+[P22]
+
+* Recommendation: There will be vast improvements if the team is empowered
+* Decision making emerges in the team.
+[P27]
+
+* Teams were given the authority to affect what was assigned to them based on
+  the effort estimate they gave themselves
+* Developer morale was increased and productivity was increased
+* Team collaboration improved
+[P30]
+
+* Teams were empowered according to the agile manifesto --> productivity and performance rose
+[P34]
+
+* Prescribing practices led only so-far
+* Effectiveness of teaching and communication was declining
+* They understood that prescribing does not go any longer, and teams need to self-organize
+* The teams were asked to tailor their practices instead
+* The teams were asked to showcase their application of practices
+* A checklist did not work --> The process needed to be tailored from team to team and between projects
+* The teams were asked to tell which practices they knew to work by first hand experience
+* Teams competed of showcasing best practices
+[P41]
+
+* Empowering teams successfully transfered the ownership of tailored practices to teams
+[P41]
+
+* When teams have the power to make decisions concerning development speed
+  and quality, the methodology is widely accepted
+[P45]
+
+S: Empowering teams (autonomy)
 

### 2014-05-23 journal/article.tex
S: Empower teams

--- 2014-05-23 journal/article.tex
@@
 * Agile is a mindset -- it can not be studied from a book -- need of coaching
+
+* What are the agile principles? What is being talked about when ``agile
+  principles'' is referred to?
 
@@
 
+\subsubsection{Enable autonomy}
+
+% Empowering teams
+
+The agile principle of giving teams the power to decide over themselves was seen
+as an important factor in advancing the transformation.
+In some cases management initially attempted to prescribe how the new practices
+should be implemented [P5, P18, P41]. However, it was learned that only when
+full control was given to teams the agile methods could be properly established
+[P5, P16, P41].
+Allowing self-organization will create commitment to the change and motivation
+to continued use [P16, P18]. It will allow teams to take ownership of the
+development model and voluntarily improve it even further [P41].
+
+The acceptance of agile methods was easy when teams gained authority to decide
+on development speed and quality [P45]. Favoring empowerment over prescribing
+the new practices was also reported to improve performance [P27]. Giving teams
+authority to decide over work items increased productivity and morale [P30,
+P34].
+
+Roche and Vasquez-McCall [P41] describe how prescribing the agile methods to use
+lead only so-far in the transformation. The effectiveness of teaching and
+communicating the transformation started to decline. To enable the
+transformation to still proceed, the an organization-wide challenge was created
+where teams would independently develop and showcase the best agile practices.
+As a result the ownership of the methods was transfered to teams, and the new
+model gained a secure foothold. [P41]
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-23 journal/article.tex
S: No top-down mandate

@@
 
+% No top-down mandate
+
+An interesting success factor was the absence of a top-down mandate. Atlas [P3]
+describes that the use of Scrum was spreading because teams were both allowed to
+use and enabled to learn the method. The transformation itself was agile when
+people felt empowerment in making the decision to adopt. The incentive to change
+was created when teams learned about agile methods, and perceived that a change
+would be beneficial [P3].
+The absence of mandate granted genuine support for the model on the grass roots
+level, which was a necessity for the process to work [P6].
+It was also thought that proceeding with a top-down mandate would have made
+teams use a uniform process [P41]. This would have been suboptimal, as it was
+important that the practices were tailored for each team and project [P41].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-23 journal/article.tex
S: Previous culture like agile

@@
 
+% Previous culture is like agile
+
+A few organizations had already some structures in place that were similar to
+agile, which made the transformation easier.
+For instance, a previous organizational model based on small and autonomous
+teams strongly aided the adoption [P3].
+Another case described the original on-demand delivery model a natural fit for
+agile, and the transformation was presented as a return to core values instead
+of a remodeling [P16].
+% The agile culture must also be maintained in each step of the transformation
+% [P31]
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-23 journal/article.tex
S: People factors

@@
 
+% Engage people broadly
+
 Many cases highlighted how people were engaged broadly in the organization.
@@
 \subsubsection{Have a change leading coalition}
+
+% Change leaders
 
@@
 
+% Have an agile community
+
 The formation and influence of agile communities was reported to have a
@@
 even further [P12].
+
+% Align the organization
 
@@
 
+\subsubsection{People factors}
+
+The transformation did not entirely rely on choices in technology and
+management, but also different people factors were given as basis of success.
+
+Choosing people with a disposition towards agile methods was seen as a
+requirement for the change to take the right track. Teams are staffed usually
+based on technical competences, but considering personality aspects alike was
+emphasized for agile teams [P29, P34]. The personality of people was seen as a
+key aspect for achieving change [P6]. There was a need for collaborative and
+understanding persons who are prepared to discard preconceptions and willing to
+try new untested approaches [P6, P29].
+% Also, having previous agile experience did not directly imply suitability to
+% work in a larga scale agile environment [P34].
+% In the beginning of the transformation team members who did not find themselves
+% comfortable with the new model were exchanged for better suited persons [P29].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-25 journal/article.tex
S: Have social events

@@
 
+% Have social events
+
+Social events were reported to benefit the transformation. A few cases even
+described a the transformation being driven by a series of events where people
+received information and had the possibility to participate in shaping the new
+model of working [P32, P41]. Various social activities were presented as
+valuable tools for improving bonding within teams [P11, P27, P40]. The importance of
+creating personal bonding was also highlighted for change leading entities such
+as management and coaching teams [P11, P23].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-25 journal/article.tex
S: People factors

@@
 % Also, having previous agile experience did not directly imply suitability to
-% work in a larga scale agile environment [P34].
+% work in a large scale agile environment [P34].
 % In the beginning of the transformation team members who did not find themselves
 % comfortable with the new model were exchanged for better suited persons [P29].
+
+% People with previous agile expertise
+
+A few cases mentioned the importance people with previous agile experience had
+at the start of the transformation [P6]. For instance, it helped the product
+management office to implement agile over the entire enterprise when the
+staffing of was updated to include people with agile experience [P10]. Staffing
+teams partially with developers familiar with agile helped the rest of the team
+get a good hold of agile development [P37].
+
+% People without baggage of the past
+
+A few cases discussed the benefit of having change leaders without baggage of
+the past. Cowan [P11] describes how the newly hired director was able to
+circumvent territorial battles that existed from before, and therefore focus
+fully on getting the agile model implemented. In another case external coaches
+were described to better spot places for correctino in the agile model, and
+their advice was received better because they were considered impartial when
+being external to the organization [P16].
+% Old employees were also willing to try something new, because the organization
+% had gone through a time of uncertainty and restructuring [P11].
 

### 2014-05-26 journal/article.tex
Some notes
Transformation description

@@
 \item
-RQ3: What are typical successes and challenges in a large-scale agile
-     transformation process?
+RQ3: What are typical challenges in a large-scale agile transformation?
+
+\item
+RQ4: What factors facilitate adopting of agile methods?
 
@@
 
+The second research question was aimed to discover how the process of
+transformation can be described.
+We provide an answer to Research Question 2 in several parts. First the initial
+states of organizations are described. Secondly, we present a two-axial model
+for categorizing transformation types. In the rest of this section we discuss
+typical arrangements that relate to the change process, such as how training is
+arranged, how change is lead, and what investments are done.
+
+The challenge of giving a generic description on how transformations proceed is
+that every organization is unique, and further, publications tended to provide
+more focused viewpoints rather than describing the transformation process
+itself. Therefore it was not sensible to attempt constructing a detailed model
+on how large organizations adopt agile methods.
+Table \ref{table:transformation_descriptiontypes} presents types of insights
+each primary source provided in relation to Research Question 2.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.52\textwidth}
+                     >{\raggedright\arraybackslash}p{0.40\textwidth} }
+        How transformation is described   &  Primary sources  \\
+        \midrule
+        Narrative description of transformation process &
+                P3, P5, P6, P9, P11, P12, P16, P18,
+                P24, P27, P30, P36, P40, P41, P46   \\
+        Application of agile methods seen from a specific viewpoint (e.g. coaching team) &
+                P10, P13, P15, P19, P23, P32, P34,
+                P39, P47, P48, P50, P52    \\
+        Viewpoints of challenges and success factors  &
+                P17, P20, P21, P29, P37, P38, P49   \\
+        Discussion on establising large scale agile practices  &
+                P7, P8, P31, P42, P43, P44, P45   \\
+        Only brief discussion on transformation process  &
+                P1, P2, P4, P14, P22, P25, P26, P28, P33, P35, P51  \\
+        \bottomrule
+    \end{tabular}
+    \caption{The main theme of each primary source, relating to providing
+             insights of the transformation process}
+    \label{table:transformation_descriptiontypes}
+\end{table}
+
 \subsubsection{The initial state of the organization}
@@
 
-\subsubsection{Typical activities in the transformation}
-
-Training
-
-Use of external consultants
-
-Coaching
-
-Pilot program
-
-Seminars, workshops, other public events
-
-Agile community
+\subsubsection{Transformation activities}
+
+During transformation periods organizations conducted activities aimed towards
+creating change. Examples on such activities are piloting new models, seminars
+and workshops, training, and coaching on new models.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
+                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
+        Activity           &  Described in  \\
+        \midrule
+        Pilot programs and other piloting   & 
+                P6, P7, P8, P9, P10, P11, P14, P16, P18, P22, P23, P24,
+                P27, P29, P36, P38, P40, P41, P43, P45, P46, P47  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Activities in the organization aiming to create change}
+    \label{table:transformation_activities}
+\end{table}
+
+% Pilot programs
+
+Pilot projects and programs were a common practice at the start of the
+transformation. Piloting was discussed in half of the cases.
+
+When?
+
+How many?
+
+What kind?
+
+Reasons for piloting
+
+Big bang + Piloting
+[P11, P16, P18, P41]
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
+
+\subsubsection{Investing in new structures in transformation}
+
+Even if any activities aiming for transformation can be considered as
+investments in themselves, the transformation process included investments in
+some tangible entities. These investments were focused on enabling agile
+software development. Typical investments were new management tools,
+improvements in testing environments, and updating of the physical working
+spaces.
 
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
+% S: Have social events (get-togethers, etc.)
+% 
+% * Management team event --> Investment
+% * The transformation management team met in person --> Personal bonding
+% * A common understanding of the principles --> Consistent implementation
+% * Excursion and training
+% [P11]
+% 
+% * It was seen beneficial that the coach team had an event
+% * The team had been dispersed
+% * Road trip
+% [P23]
+% 
+% * Various social activities are given as example of improve team bonding
+% * Ways to build trust --> Trust is needed
+% * Team outings
+% [P27]
+% 
+% * Learn the difference between functions and formal roles
+% * Teams participated in learning games
+% * Learning games created understanding for team dynamics in the agile environment
+% [P40]
+% 
+% * Team competition-based events
+% * Short events in the middle of the work week
+% * Teams got excited when they were able to compete and demonstrate their own best practices
+% * Teams got encouragement to tailor practices
+% [P41]
+
+\clearpage
 

### 2014-05-26 journal/article.tex
Work in progress

@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
-
 \subsubsection{Investing in new structures in transformation}
@@
 improvements in testing environments, and updating of the physical working
-spaces.
+spaces. Table \ref{table:transformation_investments} summarizes the
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        Investment for transformation      &  Described in  \\
+        \midrule
+        Proprietary agile material or documentation   & 
+                P1, P2, P8, P14, P16, P23, P26, P29, P31, P32, P41, P45  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Activities in the organization aiming to create change}
+    \label{table:transformation_investments}
+\end{table}
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+P: In-house agile material or documentation
+
+P 1: Abdelnour-Nocera 2007 - Adopting Agile in a large organization - balancing the old with the new.pdf - 1:9 [A key tool which formed part o..]  (7:2069-7:2717)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+A key tool which formed part of this group’s TF was a manual of agile adoption 
+which was generated by this group. However, it was not designed to carry agile 
+methods beyond software engineering, i.e. into the wider business environment. There 
+was one document we saw being circulated at mid-management level which contained 
+some principles to make the whole of the business agile. According to one of the 
+stakeholders, this was intended to provide guidance in “getting rid of old projects and 
+keeping new ones”. The logic for this is that “other parts of the new business need to 
+be agile also for the agile adoption to be successful”.
+
+Comment:
+This group is the agile advocates and coaches.
+
+P 2: Abdelnour-Nocera 2008 - Adopting Agile In A Large Organisation.pdf - 2:10 [A key tool was a manual of agi..]  (5:2063-5:2242)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+A key 
+tool was a manual of agile adoption generated by this group, but it was not designed 
+to carry agile methods beyond software engineering, i.e. to the business environment.
+
+P 8: Brown 2011 - A Case Study In Agile-At-Scale Delivery.pdf - 8:14 [The customization and packagin..]  (6:624-6:1113)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+The customization and packaging of the agile 
+delivery process for ABC Bank was given its own identity (including a logo, 
+graphics, posters, and so on). This encapsulated the specific choices of concepts, 
+processes, and terminology for agile delivery and both expressed them in a 
+vocabulary that was familiar, and also made them more readily identifiable to the 
+wide range of stakeholders around ABC Bank that would be involved with the 
+enterprise software delivery organization.
+
+P14: Fecarotta 2008 - Myboeingfleet And Agile Software Development.pdf - 14:12 [It has been about six months s..]  (3:3330-3:3548)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+It has been about six months since this approval.  
+Ryan’s team has created the Agile AD&S website 
+using a wiki, which encourages participation, and 
+emphasizes the openness and collaborative nature of 
+the effort.
+
+P16: Fry 2007 - Large Scale Agile Transformation In An On-Demand World.pdf - 16:8 [One team member wrote a docume..]  (2:1312-2:1852)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] [S: Communicating the change - Family: RQ4: Success factors] [S: Engage people broadly in the organization - Family: RQ4: Success factors]
+No memos
+
+One team member wrote a document describing the
+new process, its benefits and why we were
+transitioning from the old process. We led 45 one-hour
+meetings with key people from all levels in the
+organization. Feedback from these meetings was
+incorporated into the document after each meeting,
+molding the design of the new process and creating
+broad organizational buy-in for change. This open
+communication feedback loop allowed everyone to
+participate in the design ofthe new process and engage
+as an active voice in the solution.
+
+P16: Fry 2007 - Large Scale Agile Transformation In An On-Demand World.pdf - 16:14 [Three key members from the cro..]  (2:3195-2:3413)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+Three key members from the cross-functional
+team developed a consolidated presentation and
+training deck that included concepts from the current
+methodology, Scrum [1], XP (eXtreme Programming)
+and Lean methods [2].
+
+P16: Fry 2007 - Large Scale Agile Transformation In An On-Demand World.pdf - 16:16 [We also created an internal, w..]  (2:3602-2:3775)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+We
+also created an internal, wiki-based website as a
+repository for all our information and as a valuable
+reference for team members transitioning to the new
+methodology.
+
+P23: Hanly 2006 - Agile Coaching In British Telecom - Making Strawberry Jam.pdf - 23:11 [It was clear that people would..]  (2:3889-2:4172)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+It was clear that people would need a quick and 
+easy way of finding out about Agile Delivery and how 
+it was to be applied in BT. The BT Agile Cookbook – 
+an online collection of wisdom, guidance and advice 
+on Agile Delivery as applied to BT – was developed to 
+meet the need.
+
+P23: Hanly 2006 - Agile Coaching In British Telecom - Making Strawberry Jam.pdf - 23:13 [BT’s cookbook describes Agile ..]  (3:1-3:313)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+BT’s cookbook describes Agile Delivery in terms of 
+its traditional software development roles: the 
+customer, project manager, business analyst, designer, 
+developer and tester. It describes what's different and 
+why, and it sets out the agile core practices which are 
+seen as giving BT Exact most benefit.
+
+P26: Laanti 2008 - Implementing Program Model With Agile Principles In A Large Software Development Organization.pdf - 26:8 [All these new practices were i..]  (3:132-3:216)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+All these new practices were institutionalized with the 
+Agile Policy (Appendix A).
+
+P27: Lee 2008 - Forming To Performing - Transitioning Large-Scale Project Into Agile.pdf - 27:9 [In each phase, we encountered ..]  (2:78-2:330)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] [S: Pilot program provides knowledge of transformation - Family: RQ4: Success factors] 
+No memos
+
+In each phase, we encountered different sets of
+problems and we came up with creative tricks to solve
+them. The solutions evolved into “better or best
+practices”. These techniques shared within our team
+and were then adopted by the whole department.
+
+P29: Lewis 2007 - Over The Waterfall In A Barrel - Msit Adventures In Scrum.pdf - 29:6 [The first task became mapping ..]  (1:3088-1:3554)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+The first task became mapping waterfall terms 
+with Scrum terms.  Pre-Baseline became one of the 
+“Phase-end reviews” and Business Requirements 
+Document (BRD) became Product Backlog, captured 
+at the start of each Sprint.   
+Once terms were defined, an Agile SDLC Guide 
+was written.  This document describes Scrum in 
+terms common to the SDLC and included such topics 
+as “What is Agile?”, “What is Scrum?” and “How 
+does the MS IT SDLC differ from Scrum?”
+
+P31: Maples 2009 - Enterprise Agile Transformation - The Two-Year Wall.pdf - 31:22 [When we first embarked on our ..]  (6:918-6:1469)   (Super)
+Codes:	[I: Inhouse coaching resources - Family: Investing in change] [P: In-house agile material or documentation - Family: Practices] [S: Change leaders or champions - Family: RQ4: Success factors] 
+No memos
+
+When we first embarked on our Agile journey, we 
+made the decision to hire an Agile expert to serve as 
+our Agile evangelist and “guidance counselor.”  This 
+person has been with Borland since the very start of 
+our transformation, and he established “Borland Agile” 
+– the baseline process and framework for our entire 
+organization. His role is very much to guide, coach and 
+mentor all of the teams and to ensure that as we scale 
+and evolve, the adaptations we make to the Scrum 
+practices never take us too far away from the 
+principles.
+
+P32: McDowell 2007 - British Telecom Experience Report - Agile Intervention - Bt'S Joining The Dots Events For Organizational Change.pdf - 32:9 [The key content pieces address..]  (5:1963-5:2633)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] 
+No memos
+
+The key content pieces address highest impact opportunities for change.  This involves agile planning and estimation (which infers iterative delivery and user stories), 
+customer collaboration (a current gap with business partner involvement inside the 
+delivery cycle) and inter/intra-team communication (how to ensure good communication with distributed teams).  Short, nine minute videos have been created in each area 
+to provide a uniform delivery of information.  While these videos are run by agile 
+coaches, the video format both ensures that every participant receives the same content as well as lessening the burden of having highly experienced presenters.
+
+P41: Roche 2009 - The Amazing Team Race A Team Based Agile Adoption.pdf - 41:13 [Increase competencies to devel..]  (2:992-2:1272)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] [S: A successful agile (pilot) project at the beginning - Family: RQ4: Success factors] 
+No memos
+
+Increase competencies to develop high 
+performing teams.  Ultimately we knew 
+establishing well-run Scrum teams would 
+determine the success of the adoption. We wanted 
+to showcase these teams, identify and share their 
+practices so that their successes could be 
+duplicated.
+
+P41: Roche 2009 - The Amazing Team Race A Team Based Agile Adoption.pdf - 41:22 [The community used the newslet..]  (3:439-3:621)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] [S: Communicating the change - Family: RQ4: Success factors] [S: Have an agile community - Family: RQ4: Success factors] 
+No memos
+
+The community used the newsletters 
+to keep up to date on the transition effort; goals of 
+the transition team, upcoming Scrum events, 
+interesting articles and Scrum trivia games.
+
+P45: Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:13 [A wiki was used to manage basi..]  (4:55-4:667)   (Super)
+Codes:	[P: In-house agile material or documentation - Family: Practices] [P: SIG's, CoP's - Family: Practices] 
+No memos
+
+A wiki was used to manage basic information and observations. For each pilot team
+the following data were collected by the Scrum mentors: scope of the project, duration,
+product owner, scrum master, dependencies with other teams, locality, critical decisions
+and impediments, and other observations. These data were used to select and communicate best practices, create a network of Scrum masters and product owners, and set
+up a round table of experienced practicioners. In-depth interviews with Scrum masters,
+product owners, and team members provided valuable information and suggestions for
+improvement.
+
+% ---------------------------------------------------------------------
 

### 2014-05-26 journal/article.tex
T: Physical spaces, Agile material

--- 2014-05-26 journal/article.tex
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
+Mentioning big bang versus step wise [P16].
 
@@
 improvements in testing environments, and updating of the physical working
-spaces.
+spaces. Table \ref{table:transformation_investments} summarizes the
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        Investment for transformation      &  Described in  \\
+        \midrule
+        Proprietary agile material or documentation   & 
+                P1, P2, P8, P14, P16, P23, P26, P29, P32, P45  \\
+        Refurbishing work areas and other spaces   &
+                P6, P22, P35, P36, P38, P42, P46, P48 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Activities in the organization aiming to create change}
+    \label{table:transformation_investments}
+\end{table}
+
+% Organization specific agile material
+
+A typical practice was creating some internal material promoting the new agile
+model. The material served both a documenting role and promoted awareness for
+the new development model. The material was created to help explaining the agile
+model to stakeholders, and to serve as a reference for teams implemeting it.
+Typical in-house agile material was a manual which presented the organization
+specific implementations of agile practices [P1]. Many cases described using a
+wiki-based website to maintain and distribute information on agile practices
+[P14, P16, P45]. Hanly et al. [P23] describes how a n organization specific
+cookbook for agile implementation was created. The cookbook resolved
+misconceptions, presented differences between the old model and agile, and
+described agile development in traditional terms [P23]. Other agile material
+organizations produced were short videos [P32], and a branded theme for the new
+process [P8]. One primary source provided the organization specific Agile Policy
+as an appendix [P26].
+
+% Physical spaces
+
+In many cases it was mentioned that the work spaces needed to be rearranged to
+better allow the team centric model of agile [P35, P42, P48]. The possibilities
+to create a team area varied between cases. Some cases had trouble with physical
+space arrangements [P36], in some cases spaces could be arranged with some
+bargaining with the facilities department [P6, P38], but some organizations
+invested readily in good facilities [P46]. In addition to the work areas some
+organizations took the chance to improve common areas also [P22, P46].
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
 
@@
         Old commitments are kept              &  P11, P21  \\
-        Physical spaces can not be rearranged  &  P6, P29, P36, P49  \\
+        Physical spaces can not be rearranged  &  P6, P29, P36, P38, P49  \\
         \bottomrule
@@
 arranged so that the entire team could work in a single room, but it took time
-and effort, and it was not always possible [P36, P6]. Lewis and Neher describe
+and effort, and it was not always possible [P6, P36, P38]. Lewis and Neher describe
 how dedicated rooms for teams could not be arranged, and team events such as

### 2014-05-26 journal/article.tex
Wording
Editing

@@
     \end{tabular}
-    \caption{The main theme of each primary source, relating to providing
+    \caption{The theme of each primary source, in relation to providing
              insights of the transformation process}
@@
 Even if any activities aiming for transformation can be considered as
-investments in themselves, the transformation process included investments in
-some tangible entities. These investments were focused on enabling agile
-software development. Typical investments were new management tools,
-improvements in testing environments, and updating of the physical working
-spaces. Table \ref{table:transformation_investments} summarizes the
+investments in themselves, the transformation process included specific
+investments in some tangible entities. These investments were focused on better
+enabling agile software development. Typical investments were new management
+tools, improvements in testing environments, and updating of the physical
+working spaces. Table \ref{table:transformation_investments} summarizes common
+types of investments.
 

### 2014-05-27 journal/article.tex
Discussion

@@
     \includegraphics{graphics/reasonstochange_summary.pdf}
-    \caption{Reasons why organizations started an agile transformation}
+    \caption{Reasons why the studied organizations started an agile
+             transformation}
     \label{fig:reasonstochange_summary}
@@
 
+Grounded theory may not be able to provide a single generally applicable model
+for transformation, at least not in a literature review made based on the
+current agile transformation publications. This is because of two things.
+First, the organizations seem to differ too much from each other. Secondly, the
+published studies focus too much on isolated phenomena within the larger
+transformation.
+It would require too big of an interpretation to combine all the various cases
+into one theoretical model. We have chosen to present transformatoins using an
+elementary two-axial classification of transformations, and listings of
+individual phenomena that typically occur in a transformation.
+
 * Some success factors and challenges relate

### 2014-05-27 journal/article.tex
T: Supporting tools

@@
 investments in some tangible entities. These investments were focused on better
-enabling agile software development. Typical investments were new management
-tools, improvements in testing environments, and updating of the physical
-working spaces. Table \ref{table:transformation_investments} summarizes common
-types of investments.
+enabling agile software development, and highlight the concerns that
+organizations had when implementing an agile model. Typical investments were new
+management tools, improvements in testing environments, and updating of the
+physical working spaces. Table \ref{table:transformation_investments} summarizes
+common types of investments.
 
@@
     \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.42\textwidth}
+                     >{\raggedright\arraybackslash}p{0.50\textwidth} }
         Investment for transformation      &  Described in  \\
         \midrule
+        Agile project management tools  &
+                P10, P11, P14, P16, P17, P26, P33,
+                P41, P43, P44, P45, P48, P50 \\
         Proprietary agile material or documentation   & 
@@
     \end{tabular}
-    \caption{Activities in the organization aiming to create change}
+    \caption{Investments aiming to make agile methods perform better}
     \label{table:transformation_investments}
 \end{table}
+
+% Agile development tools
+
+Using and investing in software tools supporting agile development was mentioned
+in several studies. Especially project management tools were discussed.
+Some reasons for introducing new project management tools was that spreadsheets
+became unmanageable [P16], and the status overview of agile projects was lacking
+[P17]. When a suitable agile tool was implemented it was attributed for better
+visibility [P26, P33]. Using a project management tool can also help in teaching
+the particularities of the new process [P14].
+
+It was typical to use web-based tools for project manangement [P10, P11, P16].
+The use of web based tools such as wiki pages was also attributed to lessen the
+effort in documentation [P14]. Other tools that were taken into use while
+implementing agile methods were teleconferencing tools [P11], discussion forums
+for contacting coaches [P41], and information radiators to raise the awareness
+of the new method [P22, P41]. Also development and delivery infrastructure
+required investments to match the needs of agile development [P50].
+
+Several organizations created their own tools for managing agile development
+[P10, P11, P26]. Especially companies developing management software, such as
+SAP, Primavera, and Salesforce, customized their product to support the agile
+model being implemented [P16, P44, P45]. The development of the tools themselves
+also served as a showcase of agile delivery [P10, P45].
 
@@
         Old commitments are kept              &  P11, P21  \\
-        Physical spaces can not be rearranged  &  P6, P29, P36, P38, P49  \\
+        Physical spaces can not be rearranged  &  P6, P14, P29, P36, P38, P49  \\
         \bottomrule
@@
 arranged so that the entire team could work in a single room, but it took time
-and effort, and it was not always possible [P6, P36, P38]. Lewis and Neher describe
-how dedicated rooms for teams could not be arranged, and team events such as
-daily stand-ups required booking a conference room [P29]. In another case the
-estate organization shut down the teams' attempts to modify working spaces
-[P49].
+and effort, and it was not always possible [P6, P14, P36, P38]. Lewis and Neher
+describe how dedicated rooms for teams could not be arranged, and team events
+such as daily stand-ups required booking a conference room [P29]. In another
+case the estate organization shut down the teams' attempts to modify working
+spaces [P49].
 

### 2014-05-27 journal/article.tex
T: Continuous integration

@@
   principles'' is referred to?
+
+* Principle of build often? Principle of nightly builds or CI? Is this from XP?
+  Is this an agile principle?
 
@@
                 P41, P43, P44, P45, P48, P50 \\
+        Continuous builds and testing  &
+                P5, P11, P16, P17, P25, P27, P34, P39, P42, P50 \\
         Proprietary agile material or documentation   & 
@@
 also served as a showcase of agile delivery [P10, P45].
+
+% Investing in continuous builds
+
+Continuous builds and testing was reported to play an important role in scaling
+the agile model up to large scale [P17, P34]. Many organizations were
+experiencing problems when a daily build system was not in place. The main
+problem was that progress was slowed down because of lacks in the build and
+automated testing systems [P5, P27, P39, P50].
+Other reasons for implementing build automation was the need for having stable
+nightly builds [P5], builds were often broken when integration was attempted
+[P27], and there were problems with timeliness in test execution [P11]. One case
+reported that the entire infrastructure needed to be revamped to attain benefits
+from the accelerated work pace of Scrum teams [P50].
+
+When a general problem of frequent broken builds existed it was easy to acuire a
+buy in for creating an organization-wide standard for automating builds and
+tests [P11]. On the other hand, many cases emphasized that building the
+continuous integration environment required a significant investment [P17, P34,
+P42]. In order to have the focus and resources for creating a proper integration
+environment a dedicated team was Some organizations appointed a dedicated team
+to ensure the focus and resources for creating a proper integration environment
+were available [P5, P16, P17].
+
+Having daily builds was considered a significant advantage in the agile
+implementation [P5, P16]. Having the continuous integration environment in place
+cases reported that development time was reduced [P39], the build times were
+significantly improved [P5, P42], teams could know on a daily basis whether
+their code could be integrated [P17]. It was also reported that the
+communication and trust between testing and development was improved [P42].
 
@@
   organization had tried in practice.
+
+* For instance, implementing continuous integration and automated builds was not
+  explicitly stated as a factor contributing to the success of the
+  transformation, although it is unquestionable that such systems are critical
+  for maintaining quality and enabling frequent releases.
 

### 2014-05-28 journal/article.tex
T: Change leaders

@@
   Is this an agile principle?
+
+* Kotter: Having a ``guiding team'' for change
 
@@
 
+\subsubsection{How change was lead}
+
+A significant change effort requires entities in the organization who take
+leadership in the change.
+
+Change was lead in X ways.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        Type of leading change         &  Described in  \\
+        \midrule
+        Agile task group foor coordinating change   &
+                P4, P18, P21, P23, P25, P30, P40, P41,
+                P43, P45, P48, P49, P50  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Which entities were leading change in the agile transformation}
+    \label{table:transformation_leadership}
+\end{table}
+
+% Agile task force
+
+Many organizations formed an agile task force or similar entity for coordinating
+and leading the change. It was typical that in the beginning of the
+transformation, or when it had been acknowledged that the old model had
+problems, a work group was founded for coordinating the changes to be done. The
+change team usually consisted of management personnel [P43, P50], and was
+reinforced with external agile consultants [P18, P21, P25, P30]. Some cases
+highlighted that the team had a cross organizational staff who were familiar
+with the varying functions that would be affected by changes [P4, P21, P41,
+P48]. It was important that theere was a group of people who were strongly
+committed to making a change happen [P43].
+
+The tasks of the change team included first finding an alternative delivery
+method to solve problems in the old model [P21, P30]. During the transformation
+the team stimulated agile adoption in the organization [P18, P45, P49]. Examples
+on their tasks were informing people on the agile model [P4], educating and
+coaching [P45, P48, P49], creating an agile cookbook [P23], and organizing
+presentations and workshops [P25, P41]. The change team also assessed the
+success of transformed teams, and worked for repeating their success across the
+organization [P40, P41].
+% The team was also empowered to remove the impediments they discovered [P50].
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
+
 \subsubsection{Transformation activities}
@@
 organizations took the chance to improve common areas also [P22, P46].
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
 
@@
 
+* How do transformations fit within Kotter's 8 step process?
+
 % ---------------------------------------------------------------------

### 2014-05-28 journal/article.tex
Work in progress

@@
         \midrule
+        High up change leaders or sponsors  &
+                P3, P5, P24, P31, P36, P38, P46, P49  \\
         Agile task group foor coordinating change   &
-                P4, P18, P21, P23, P25, P30, P40, P41,
+                P4, P16, P18, P21, P23, P25, P30, P40, P41,
                 P43, P45, P48, P49, P50  \\
@@
 highlighted that the team had a cross organizational staff who were familiar
-with the varying functions that would be affected by changes [P4, P21, P41,
+with the varying functions that would be affected by changes [P4, P16, P21, P41,
 P48]. It was important that theere was a group of people who were strongly
@@
 on their tasks were informing people on the agile model [P4], educating and
-coaching [P45, P48, P49], creating an agile cookbook [P23], and organizing
+coaching [P16, P45, P48, P49], creating an agile cookbook [P23], and organizing
 presentations and workshops [P25, P41]. The change team also assessed the
 success of transformed teams, and worked for repeating their success across the
-organization [P40, P41].
-% The team was also empowered to remove the impediments they discovered [P50].
+organization [P16, P40, P41]. Still one characteristic was that the team was
+empowered to act on impediments they discovered [P16, P50].
 
@@
 
+% High up change leaders
+
+In many cases it was described that the change was initiated by a limited number
+of people in high-up positions. In these cases typical positions attributed for
+giving the initial push for change were CxO's, vice presidents, and directors.
+The pressence of a high up manager was described to underline the organization's
+commitment for change [P24].
+
+% Other change leaders
+
+Some organizations looked for recruiting champions to push with the
+transformation in local business units [P9].
+
+An important factor in the transformation was that organizations needed people
+to spread the word of the new methodology, and persuade for change [P9].
+
+% --
+
+Who?
+
+How were they selected?
+
+What is their task?
+
+What did they do?
+
+Do they have challenges in their tasks?
+
+What is particular?
+
 % ---------------------------------------------------------------------
+
+[P3] High up leader
+* A Development Manager
+* Who: A single person
+* What: The team expressed a need to formalize their development process
+* What: He was able to push a few other people to get interested in agile
+* What: The word spread. There was some interest already before, but he helped ignite the spark
+
+[P5] High up leader
+* A VP joined, who was pro-agile
+* Who: A vice president manager
+* What: He started introducing agile to the company
+* He kept firm and committed to the agile transformation --> Regardless of a few casualities
+
+[P24] High up leader
+* Who: The CTO
+* What: He lead training and awareness sessions
+* Particular: His presence made the management commitment to change clear
+
+[P31] High up leader
+* Who: A vice president
+* What: Make the change happen, commitment
+* What: Educating executive peers, be a guardian of the agile culture
+* Particular: He was committed, and pushed for the change to happen
+
+[P36] Consulting company
+* The consulting company wanted to work in an agile way
+* The contracting company got a view of agile practices, and a demonstration of effectivity
+[P36] High up leader
+* A pilot project was successful, and a high up manager became convinced
+* Who: The director of IT
+* What: He was the management champion. He would anticipate and deal with challenges
+
+[P38] High up leader
+* The company hired a new CTO, known for helping troubled companies
+* Who: The new CTO
+* What: He assessed the current problems
+* What: He took steps for making people get interested in agile
+* What: He had determination to make difficult management decisions for getting agile working
+* What: Important role of persuading the rest of the business to change their way of working
+
+[P46] High up leaders
+* A new CIO was hired. Later a director was hired
+* Who: The new CIO, another director
+* What: CIO took attention to break down internal barriers, and make the company perform better
+* What: CIO persuaded business partners
+* What: Director: a vision of makin things work better, and making everyone proud of their work
+* What: The director persuaded the development group
+* What: The director stayed firm and committed, when presented with impediments
+
+[P49] High up leader
+* A grand experiment was started wit the lead of the CIO
+* Who: the CIO
+* What: Persuaded to do a transformation
+
+% --
+
+[P18] Managers
+* Managers are key in making change stick
+* Who: Managers
+* What: Lead people, make the change stick
+* Particular: Managers needed to be coached for taking this leadership role
+
+[P14] One manager
+* There was one project manager who wanted to drive agile
+* Who: One project manager
+* What: He was devoted to use agile methods
+* What: He fought for getting use of agile accepted
+
+[P2] Group
+* A group of agile advocates is driving the push
+* Who: Advocates, coaches
+* Task: facilitate successful agile adoption, spread knowledge
+* Task: Persuade other staff to adopt
+* Challenge: move agile from development teams to rest of the business
+
+[P16] Group (more for ``change team'' ? )
+* Who: A cross functional rollout team
+* Who: People representing different roles
+* What: Created a schedule for rollout, provided expertise, coaching
+* What: Spreading the word
+* What: Removal of obstacles
+
+[P6] Champions in teams
+* There were agile champions in teams
+* Who: People in teams, from all disciplines
+* What: They knew the problems at hand, and had the knowledge to adjust the process accordingly
+* Particular: Hard to find people who understand the principles, and can help their teams
+
+[P9] Champions, coaching staff, individuals in teams
+* Champions were hired in business uints to help spread knowledge
+* Who: People across the organization
+* Who: Agile coaching staff was hiring people
+* Task: Help spread agile knowledge, teach techniques
+* Particular: The coaching staff was disbanded, as they could not produce results
+* There were agile people in the trenches afterwards, who drove the agile use further
+
+[P17] An evangelist
+* It was seen as necessary to have an evangelist position when the scale became larger
+* Who: someone who lead the Scrum of Scrums
+* What: Coordinate work when it was obvious that coordination was needed
+
+[P27] A single developer
+* Who: A developer took the role of leading the team
+* What: The team was empowered when there was an internal leader
 

### 2014-05-28 journal/article.tex
T: Change leaders

--- 2014-05-28 journal/article.tex
@@
 A significant change effort requires entities in the organization who take
-leadership in the change.
+leadership in the change. Getting a large group of people aligned towards a
+common goal, in many cases a completely new way of working, requires that
+someone takes charge for the undertaking.
 
 Change was lead in X ways.
+
+Leadership for change showed on all organizational levels.
+% The ways how change was lead was similar on each level.
 
@@
                      >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        Type of leading change         &  Described in  \\
+        Change leaders         &  Described in  \\
         \midrule
-        Agile task group foor coordinating change   &
-                P4, P18, P21, P23, P25, P30, P40, P41,
+        High-up change leaders or sponsors  &
+                P3, P5, P24, P31, P36, P38, P46, P49  \\
+        Change leaders on other organizational levels  &
+                P2, P6, P9, P14, P17, P18, P27  \\
+        Agile task group for coordinating change   &
+                P4, P16, P18, P21, P23, P25, P30, P40, P41,
                 P43, P45, P48, P49, P50  \\
@@
     \end{tabular}
-    \caption{Which entities were leading change in the agile transformation}
+    \caption{Summary of entities leading change in the agile transformation}
     \label{table:transformation_leadership}
 \end{table}
+
+% High-up change leaders
+
+In many cases it was described that the change was initiated by a limited number
+of people in high-up positions. In these cases typical positions attributed for
+giving the initial push for change were CxO's, vice presidents, and directors.
+In these cases it was highlighted that one or very few high-up persons affected
+significantly the start or expansion of the agile transformation. In some cases
+the change leader was a new hire [P5, P38, P46].
+
+The pressence of a high-up manager was described to underline the organization's
+commitment for change when dealing with employees or other stakeholders [P24].
+The firm commitment of the high-up leader was key in keeping the agile
+transformation on course in cases where objections or other problems arose [P5,
+P14, P36, P46]. Another important role of the change leader was to persuade and
+work for getting other people raise an interest in agile methods [P3, P31, P38,
+P46, P49].
 
@@
 on their tasks were informing people on the agile model [P4], educating and
-coaching [P45, P48, P49], creating an agile cookbook [P23], and organizing
+coaching [P16, P45, P48, P49], creating an agile cookbook [P23], and organizing
 presentations and workshops [P25, P41]. The change team also assessed the
 success of transformed teams, and worked for repeating their success across the
-organization [P40, P41].
-% The team was also empowered to remove the impediments they discovered [P50].
+organization [P16, P40, P41]. Still one characteristic was that the team was
+empowered to act on impediments they discovered [P16, P50].
+
+% Other change leaders
+
+Leadership for change emerged also on other organizational levels. In many cases
+these change leaders were still in management positions [P14, P18, P17]. There
+were also agile champions in teams who drove the change on a grassroots level
+[P6, P9, P27].
+As with higher-ups, these people worked for making agile methods familiar [P2,
+P6, P9, P17], making change stick [P2, P18], and persuaded others to use agile
+methods [P2, P9, P14].
+Benefield [P6] describes that people in teams were well able to adjust the
+application of the agile process as they were familiar with the problems at
+hands. It was howerver hard to find people who both understood the agile
+principles well, and knew how to coach within the team [P6].
+% Also coaches in [P2], and a team in [P16]
+
+Chung and Drummond [P9] give a further description of grassroots level
+leadership. Yahoo! initially used a team of coaches to help applying agile, but
+the coach team failed to make agile take root. The coaches were later disbanded
+due to reductions in budget. The agile practitioners in teams took over the
+leadership for the continued application of agile [P9].
 
@@
 
+* A qualitative interpretation requires much text. Therefore the results got
+  long. See ``Meta-ethnography''.
+
 % ---------------------------------------------------------------------

### 2014-05-28 journal/article.tex
T: Change leaders: Edit

@@
 common goal, in many cases a completely new way of working, requires that
-someone takes charge for the undertaking.
-
-Change was lead in X ways.
-
-Leadership for change showed on all organizational levels.
-% The ways how change was lead was similar on each level.
+someone takes charge for the undertaking. The primary sources showed evidence of
+leadership for change on all organizational levels. Table
+\ref{table:transformation_leadership} summarizes different types of change
+leaders that were discovered.
 
@@
                 P3, P5, P24, P31, P36, P38, P46, P49  \\
-        Change leaders on other organizational levels  &
-                P2, P6, P9, P14, P17, P18, P27  \\
         Agile task group for coordinating change   &
@@
                 P43, P45, P48, P49, P50  \\
+        Change leaders on other organizational levels  &
+                P2, P6, P9, P14, P17, P18, P27  \\
+        Informal spreading of agile awareness  &
+                P3, P6, P40  \\
         \bottomrule
@@
 leadership for the continued application of agile [P9].
+
+% Informal spreading of awareness
+
+Change leadership showed on the lowest level as informal spreading of awareness
+of agile methods.
+Benefield [P6] describes that the strongest force over other change leaders was
+letting agile knowledge spread virally. New Scrum teams were emerging as people
+heard of the new methodology [P6].
+It was similarly described that in an organization with an introverted culture
+the proclamations of managers could not match with the power of practitioners
+explaining to others how they had succeeded in with the agile method [P40].
+Even casual hallway talks could spread the word effectively, and it was
+also easier to encounter people from other parts of the organization in an
+informal context [P3].
 

### 2014-05-29 journal/article.tex
Work in progress

@@
 
-Work in progress...
-
-Top down
-
-Bottom up
-
-Big bang
-
-Transitional
+* Top down
+
+* Bottom up
+
+* Big bang
+
+* Step wise
+
+* Transformation is continuous, hard work, and takes time
+
+Mentioning big bang versus step wise [P16].
 
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
-
 \subsubsection{Transformation activities}
@@
     \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
-                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
+                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
         Activity           &  Described in  \\
         \midrule
+        Training    &
+                P3, P6, P8, P9, P11, P16, P18-P21, P23-P25, P27, P30,
+                P32, P35, P37, P39-P41, P43-P46, P48-P51  \\  % 29
         Pilot programs and other piloting   & 
-                P6, P7, P8, P9, P10, P11, P14, P16, P18, P22, P23, P24,
-                P27, P29, P36, P38, P40, P41, P43, P45, P46, P47  \\
+                P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
+                P36, P38, P40, P41, P43, P45, P46, P47  \\  %22
         \bottomrule
     \end{tabular}
-    \caption{Activities in the organization aiming to create change}
+    \caption{Summary of activities aiming to advance the transformation}
     \label{table:transformation_activities}
@@
 
-Reasons for piloting
+Reasons for piloting?
+
+Did it work?
 
@@
 
-Mentioning big bang versus step wise [P16].
+% Training
+
+Who?
+
+What kind?
+
+When?
+
+Why?
+
+Who teaches?
+
+Did it work?
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+As the organization is changing it is necessary that people know what to change
+into. Therefore the majority of cases described the use of training in the
+transformation.
+
+It was very common that the first persons involved with agile received
+training. In the start change leaders and pilot teams attended firmal training
+on agile development.
+
+Management was also among the first ones to receive agile training, and they
+received more in-depth training than the rest of the organization [P11].
+
+The type of training that was offered to teams was typically only a single
+introductory session [P16, P20], although some organizations invested in a
+multi-day course [P19].
+
+Training was reported to make a difference between success and failure of teams
+[P6].
+It became clear that training was necessary to retain an agile culture [P19].
+
+The training curriculum was evolving as trainers improved and learned the needs
+of the organization [P6]. In some cases training workshops was even initially
+tailored to suite the needs of the project [P23].
+
+Typically the training started with sending management and possibly a pilot team
+to in-depth training. After the pilot other teams started transitioning to the
+new development model. Teams about to transition typically attended a shorter
+agile introduction session, lasting for one day or less. The reason for reduced
+training for development teams was that there were many people to be trained.
+The lack of training for teams was compensated with coaching during ongoing
+projects.
+
+Training management served the purpose of creating a unified understanding of
+the methods at top level [P11, P24].
+
+It was common that the first introductory events were held by known industry
+experts [P25, P27].
+
+Some training included simulation workshops [P40, P41].
+Hands-on workshops [P46].
+
+% ---------------------------------------------------------------------
+
+[P3]
+* Who: The change leader
+* When: Before starting with the transformation
+* What: CSM course
+* Teacher: Ken Schwaber, later the change leader
+* Works: Education in itself increased interest in agile
+* Particular: The insight increased less than anticipated for anyone taking the course
+
+[P6]
+* When: initial thing
+* Who: The pilot team
+* What: Comprehensive training, CSM course
+--
+* When: Later
+* Who: Everyone
+* What: The training curriculum was evolving as trainers improved and learned the needs of the organization
+* Works: There was a difference between success and failure
+* Particular: They have invested and investigated what training suits best
+
+[P8]
+* When: Pilot teams reported trouble
+* Who: Pilot teams
+* Particular: Training was increased when there was problems getting started in pilots
+
+[P9]
+* Who: Pilot teams
+* Who: Also other interested people
+* What: Multi-day comprehensive training
+* Teacher: Experienced external agile coaches
+
+[P11]
+* Who: Management team
+* When: At start of the transformation
+* What: Common understanding of Scrum principles
+--
+* Who: Rest of the organization
+* Teacher: Scrum masters
+* What: Two hour overview -- enough understanding to get started
+* What: The Scrum masters assisted the transformation of teams
+* Particular: The Scrum masters were supposed to deepen the teams' understanding over time
+
+[P16]
+* When: First
+* Who: Program and functional managers
+* What: CSM course
+--
+* Who: Every team
+* What: a short agile introduction session
+* In depth training for PO and CSM roles
+
+[P18]
+* When: First
+* Who: The pilot team
+* Teacher: Agile consultant
+* What: One day course, coaching wiith kick-off and initial planning
+
+[P19]
+* Teacher: Ken Schwaber
+* When: Initially (?)
+* Who: 60 Scrum masters
+-- 
+* When: Later
+* Who: 240 employees
+* What: 2 day course
+* Why: It became clear that agile training was necessary to retain the culture.
+
+[P20, 21]
+* When: Initial
+* Who 30 team members from different roles
+* Works: There was too little training
+* Particular: Failure was attributed partially to the lack of training
+
+[P23]
+* What: A full day deep dive program event
+* Who: Projects could request this training
+* What: Project specific training
+
+[P24]
+* Who: CEO, CTO, managers, high-ups
+* What: Scrum master training, product owner training
+* When: Initially
+* What: Create a unified understanding
+--
+* Who: Both the main site and the offshore site
+* Teacher: The CTO, internal certified Scrum trainers
+
+[P25]
+* When: During the first year
+* When: Initially, by acclaimed experts
+* Who: The personnel
+* Teacher: Agile transformation team, managers, external consultants
+
+[P27]
+* Teachers: Renowed industry experts
+* What: Educating specific XP practices
+* Why: Help teams transition from waterfall to agile
+
+[P30]
+* Teacher: Jeff Sutherland
+* Who: Most of the engineering group, several managers, executives
+
+[P32]
+* When: A series of education events
+* Who: 100 persons at a time, 3000 in total
+* What: Practice agile in workshops
+* Particular: A productized training workshop
+
+[P35]
+* Who: The teams
+* What: Training
+* Particular: Product managers did not receive training, and they stuck in old ways
+
+[P37]
+* When: Four months after start
+* Who: All developers
+* Why: To build consensus, acceptance
+* What: Two day training workshop, outside the daily project environment
+
+[P39]
+* When: First step
+* Who: The development team
+* Why: Team members shared same values, and used the same methods
+
+[P40]
+* When: Pilot projects
+* Who: Several teams
+* Teacher: Consultants
+* What: Training, CSM training
+* What: Simulation workshops
+
+[P41]
+* Who: All project teams
+* Teachers: Transition team, external coaches
+* What: full day workshop
+* What: Simulation
+* Particular: Mixing attending teams to create new relationships
+
+[P43]
+* Who: The rollout team
+* What: Training, learning from consultants
+
+[P44]
+* Who: Entire development team
+* What: Workshop on agile development and Scrum
+* What: 15 CSM certifications
+* Teacher: Ken Schwaber
+
+[P45]
+* When: Initial
+* Who: Pilot teams
+* What: Overview, one day training
+* What: In-depth training for Scrum masters and POs
+--
+* When: After one year
+* What: Mentor training for experienced agilists
+--
+* What: A one day lean training workshop, in the fashion of the earlier ones
+
+[P46]
+* Who: Everyone, developers and managers
+* What: Scrum and agile training
+* Teacher: Internal
+--
+* When: Pilot projects
+* What: Key persons in pilot teams
+* Teacher: Ken Schwaber
+--
+* What: A series of education events
+* What: Introduction, a hands-on workshop (simulation?)
+
+[P48]
+* When: The first step
+* Who: Everyone, 30 teams, project managers, 800 people in total
+* Why: To understand where the organization is going, when shifting away from waterfall
+* Particular: It took time, as there were so many people, but it was straight forward in the end
+
+[P49]
+* Teacher: Internal coaching team, external consultant
+* What: A productized agile learning workshop
+* What: 4 day event
+* Who: Several teams
+* Who: 30\% of teams got the workshop, others got nothing
+* Works: The workshop was heavy, but it delivered results
+
+[P50]
+* Who: Product marketing, management
+* What: product owner certification
+
+[P51]
+* Teacher: External consultant
+* Why: Figure out how Scrum works
+--
+* Who: The company
+* What: in-house training of how Scrum works
+* Teacher: in-house
+
+% ---------------------------------------------------------------------
 
@@
     \end{tabular}
-    \caption{Investments aiming to make agile methods perform better}
+    \caption{Summary of investments aiming to make agile methods perform better}
     \label{table:transformation_investments}

### 2014-05-29 journal/article.tex
Work in progress

@@
 
-As the organization is changing it is necessary that people know what to change
-into. Therefore the majority of cases described the use of training in the
-transformation.
+As the organization is changing it is necessary to train people so they get some
+familiarity with the new model that is being pursued. The majority of cases
+described the use of training in the transformation.
 
@@
 
-Management was also among the first ones to receive agile training, and they
-received more in-depth training than the rest of the organization [P11].
+Management was also among the first ones to receive agile training. In some
+cases management received more in-depth training than the rest of the
+organization [P11].
 
@@
 
+The training curriculum was evolving as trainers improved and learned the needs
+of the organization [P6]. In some cases training workshops was even initially
+tailored to suite the needs of the project [P23].
+
 Training was reported to make a difference between success and failure of teams
@@
 It became clear that training was necessary to retain an agile culture [P19].
-
-The training curriculum was evolving as trainers improved and learned the needs
-of the organization [P6]. In some cases training workshops was even initially
-tailored to suite the needs of the project [P23].
 

### 2014-05-29 journal/article.tex
Training

--- 2014-05-28 journal/article.tex
@@
 
+% Training
+
+As the organization is changing it is necessary to train people so they get
+familiarity with the new model that is being pursued. The majority of cases
+described the use of training in the transformation.
+Most typically the change leaders and pilot teams were the first to attend
+formal training on agile development. Management was also usually among the
+first ones to receive agile training. In some cases particularly management
+received more in-depth training [P11, P16, P24]. In addition to providing an initial
+understanding of agile it was considered important that the key people get a
+common understanding of the method [P11, P24, P39].
+
+After key people had attained knowledge on agile, and pilot teams had possibly
+tested the methods, other teams received training on the new development model.
+In some cases the training of development teams was described to be brief [P11,
+P16, P20]. Other cases implemented a full day workshop [P18, P23, P32, P41,
+P45], and some organizations invested even in multi-day training [P9, P19, P37,
+P49].
+The reduced training for engineers was compensated with coaching during ongoing
+projects [P11].
+
+Several cases mentioned that the first introductory events were held by known
+industry experts [P3, P19, P25, P27, P30, P46]. External consultants were also
+often used as trainers [P18, P25, P40, P41, P49, P51]. In some cases a limited
+number of people with agile knowledge was first trained or acquired, and later
+the training was arranged with internal resources [P3, P11, P24, P25, P41, P46,
+P49, P51].
+The training curriculum was evolving as trainers improved and learned the needs
+of the organization [P6]. In some cases training workshops was initially
+tailored to suite the needs of the project [P23].
+
+Training was considered as important. Training was reported to make a difference
+between success and failure of teams [P6].
+It became clear that training was necessary to retain an agile culture [P19].
+Training was also described to increase the general interest towards using agile
+methods [P3].
+There were also some negative observations related to training. Atlas [P3]
+describes that the insights gained from training were less than anticipated
+[P3]. In another case product managers did not receive training and they stuck
+in old ways of managing, which jeopardized the transformation [P35].
+
+% Some training included simulation workshops [P40, P41].
+% Hands-on workshops [P46].
+
 % Pilot programs
@@
 
+Why is piloting started?
+
+Who?
+
 When?
 
-How many?
-
 What kind?
 
-Reasons for piloting
+What amount of piloting is applied?
+
+Did it work?
 
@@
 
-Mentioning big bang versus step wise [P16].
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
 

### 2014-05-29 journal/article.tex
Work in progress: Coaching

@@
                 P32, P35, P37, P39-P41, P43-P46, P48-P51  \\  % 29
+        Coaching    &
+                P1, P2, P3, P6, P8, P9, P16, P18, P23, P27, P29,
+                P33, P36, P39, P40, P41, P45, P46, P47, P50, P52  \\  % 21
         Pilot programs and other piloting   & 
                 P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
-                P36, P38, P40, P41, P43, P45, P46, P47  \\  %22
+                P36, P38, P40, P41, P43, P45, P46, P47  \\  % 22
         \bottomrule
@@
 % Hands-on workshops [P46]. 
+% Certified Scrum master training --> In several cases
 
@@
 
+% ---------------------------------------------------------------------
+
+% Coaching
+
+Why is coaching used?
+
+Coaches are who?
+
+Who receives coaching --> Teams, very few cases with some other roles
+
+When?
+
+How is coaching done?
+
+Did it work?
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
 
+Coaches were driving the agile push [P2].
+
+Coaching was considered very important at project start. Otherwise there may be
+severe long term consequences [P8].
+
+Coaching was also seen as an important success factor [P29, P33].
+
+The agile adoption erquired coaches [P47].
+
+Internal coaches were required for making the change grow [P47].
+
+To be able to coach efficiently it was required that the coach had detailed
+understanding of the company's processes and systems [P23]
+
+Coaches were also involved in training [P27].
+
+Scrum masters that had been trained earlier were coaching the teams in some
+cases.
+
 % ---------------------------------------------------------------------
+
+[P1, P2]
+* Why: best for introducing agile methods
+* Why: They drive the push
+* How: Coach in the team
+
+[P3]
+* Why: The change leader persuaded to introduce coaching
+* Coach: The change leader
+* Who: teams
+* How: Teams were coached
+
+[P6]
+* Why: It helps to have people with experience
+* Coach: An internal change leader, a team of coaches
+* Who: Teams
+* How: Start heavy a kick-off, and later draw back
+* Particular: There were little coaches at start
+* Particular: The coaches showed positive results, and management was convinced to invest in coaching
+
+[P8]
+* Why: Intensive coaching is necessary 4-6 months at start
+
+[P9]
+* Coach: A hired consultant specialist
+* How: The coach embodied the best agile principles
+
+[P16]
+* Why: To ensure trainsformation is successful
+* Coach: The cross-functional rollout team
+
+[P18]
+* Coach: An agile consultant
+* When: For over a year
+* How: Accompanied the introduction for over a year, trained on specific agile practices
+
+[P23]
+* Coaches: In-house coaching team, some external coaches
+* The coach community needed to be tended to specifically
+* If there were too few coaches things did not work
+* Pair coaching, apprentice coaching
+* Particular: The coaches needed to have specific knowledge of the company's systems and processes
+* Particular: Coaches in pairs --> training of coaches
+* Particular: Focus on coaching
+
+[P27]
+* Coaches: External coaches
+* How: Train on agile practices
+
+[P29]
+* Coaches: Agile consultants
+* Why: Having a impartial third party was a key success factor
+--
+* Coaches: The pilot team, matured teams
+* Who: Teams
+* How: Helping other teams in getting started
+* How: Taking a formal mentoring role when it is required
+
+[P33]
+* Who: Product owners
+* Why: Crucial to make it successful
+* Particular: Coaches were brought in after the start of the transformation, when trouble already existed
+
+[P36]
+* Coaches: Coaches from a subcontractor
+* How: The coached drive the project
+* How: Mentor in-house leads to take lead as soon as possible
+* How: Retrospect and adapt
+* When: The need for coaches disappeared gradually
+* The coach role was apparently seen as the lead of the project
+
+[P39]
+* Coach: External support, hired Chief Scrum masters
+* How: inspect and adapt
+
+[P40]
+* Who: Teams
+* How: Coaching breakout sessions
+
+[P41]
+* Coach: Internal and external
+* Who: Teams
+* How: Coaching office hours
+* How: Have a forum where teams can ask for help, and to share their implementation
+* Particular: It was disliked to have peers as coaches, as they were not considered more knowledgeable
+
+[P45]
+* Coach: Internal small expert team
+* Who: Teams
+* How: Supporting, advocating agile development
+* How: Educate and support teams interested in Scrum
+* Particular: Later offering possibility to teach more Scrum mentors
+
+[P46]
+* The pilot teams were coached
+* Who: Teams that are about to start agile projects
+* How: A productized training workshop, and a jump start afterwards
+* How: Key team roles are coached in the start, so that the start is successful
+
+[P47]
+* Why: The agile adoption required coaches
+* Coaches: Initially external consultants
+* Coaches: Internal people started taking coach positions
+* Coaches: Apprentice coaches, guided by external experts
+* Particular: A structure was created around coach certification
+* Particular: A coach community
+
+[P50]
+* Coaches: Internal key roles, these received initial coach and role specific training
+* When: When teams were starting
+* How: Coached to facilitate start up of teams
+
+[P52]
+* Who: Paired line manager and Scrum master
+* How: Pair coaching the pair itself
+* Coach: In some cases an experienced Scrum master was brought to help the pair
 

### 2014-05-29 journal/article.tex
Spell check

@@
 The fact that the study is based on a quantitative basis
---> Experie reports have a subjective bias
+--> Experience reports have a subjective bias
 --> The data synthesis must not rely on quantitative measures. This means that
@@
                 P17, P20, P21, P29, P37, P38, P49   \\
-        Discussion on establising large scale agile practices  &
+        Discussion on establishing large scale agile practices  &
                 P7, P8, P31, P42, P43, P44, P45   \\
@@
 
-The pressence of a high-up manager was described to underline the organization's
+The presence of a high-up manager was described to underline the organization's
 commitment for change when dealing with employees or other stakeholders [P24].
@@
 with the varying functions that would be affected by changes [P4, P16, P21, P41,
-P48]. It was important that theere was a group of people who were strongly
+P48]. It was important that there was a group of people who were strongly
 committed to making a change happen [P43].
@@
 application of the agile process as they were familiar with the problems at
-hands. It was howerver hard to find people who both understood the agile
+hands. It was however hard to find people who both understood the agile
 principles well, and knew how to coach within the team [P6].
@@
 
-The agile adoption erquired coaches [P47].
+The agile adoption required coaches [P47].
 
@@
 [P16]
-* Why: To ensure trainsformation is successful
+* Why: To ensure transformation is successful
 * Coach: The cross-functional rollout team
@@
 
-It was typical to use web-based tools for project manangement [P10, P11, P16].
+It was typical to use web-based tools for project management [P10, P11, P16].
 The use of web based tools such as wiki pages was also attributed to lessen the
@@
 
-When a general problem of frequent broken builds existed it was easy to acuire a
+When a general problem of frequent broken builds existed it was easy to acquire a
 buy in for creating an organization-wide standard for automating builds and
@@
 the new development model. The material was created to help explaining the agile
-model to stakeholders, and to serve as a reference for teams implemeting it.
+model to stakeholders, and to serve as a reference for teams implementing it.
 Typical in-house agile material was a manual which presented the organization
@@
 a team''. Training also helped people to become more positively inclined towards
-the new model [P16, P37], and at best training made people enhusiastic to change
+the new model [P16, P37], and at best training made people enthusiastic to change
 [P3, P46]. A few cases reported that there was some disinclination to change,
@@
 mindset of adapting to the situation. It is difficult to explain by theory how
-such a mindset should be applied, and the agile practices are best learnt by
+such a mindset should be applied, and the agile practices are best learned by
 doing. Coaching teams while they apply the agile methods in practice was seen as
@@
 The choice of agile practices did affect the success of the transformation, and
-also how the practices were choosen affected.
+also how the practices were chosen affected.
 
@@
 By allowing management to remain unchanged it was possible to introduced agile
-stepwise on team level, and helped in getting management buy-in for the process
+step-wise on team level, and helped in getting management buy-in for the process
 [P14, P29, P48].
@@
 [P46, P48]. When people understand the agile values they will also understand
-why the chage is being done and feel motivated [P1, P16, P42].
+why the change is being done and feel motivated [P1, P16, P42].
 The people centric nature of agile development should be communicated correctly.
@@
 the needs of each team [P41].
-Spayd [P49] writes that teams who modified agile practices to suite te large and
+Spayd [P49] writes that teams who modified agile practices to suite the large and
 distributed environment ended up doing better than teams that did not.
@@
 to regularly challenge teams to refine the agile implementation to reflect the
-courrent needs of the organization [P41, P43].
+current needs of the organization [P41, P43].
 
@@
 Especially inexperienced coaches and Scrum masters made mistakes in focusing too
-much on the implementation of practces [P23, P30].
+much on the implementation of practices [P23, P30].
 
@@
 stories handled by teams. It was stated that it is important that the
-requirements are consise and small enough for the teams to handle [P5, P11].
+requirements are concise and small enough for the teams to handle [P5, P11].
 It was recommended to invest in teaching skills in breaking down and writing
@@
 To make a large organization change the direction of the change needs to be
-communicated clearly. Atriculating urgency at start will help getting the change
+communicated clearly. Articulating urgency at start will help getting the change
 under way [P49]. Feedback of changes should certainly be listened to, but in the
@@
 the magnitude of change helped created the impression that the change was non
-negotiable [P11]. Firstly, as it vas evident that a large change was to happen
+negotiable [P11]. Firstly, as it was evident that a large change was to happen
 people felt encouragement and permission to move on from the old way, and
-secondly, the urgency eliminated wasteful discussions on wheter Scrum was a good
+secondly, the urgency eliminated wasteful discussions on whether Scrum was a good
 model or not [P11].
@@
 It is clear that any positive results will facilitate an organizational
-transformation. Although a recipie to fabricate positive experiences can not be
+transformation. Although a recipe to fabricate positive experiences can not be
 given, many cases described how they had succeeded. Benefield [P6] describes how
@@
 describes how the corporate agile awareness and teaching event was designed to
-be fun and engaging, which made people take more ethusiasm in applying the agile
+be fun and engaging, which made people take more enthusiasm in applying the agile
 practices [P32]. The move towards agile is assisted by making any benefits
@@
 to reach a maximum number of people throughout the organization, as without
-communiction the a new model will not take root [P16, P43].
+communication the a new model will not take root [P16, P43].
 It was recommended that use of the new model is made highly visible on many
@@
 teams understand a new method or process, repeat your communication to be
-sure.'' Workshops, coaching sessions, online discussions, and newletters are
+sure.'' Workshops, coaching sessions, online discussions, and newsletters are
 examples on different communication formats used [P41]. Another approach in
@@
 display project status publicly [P11], by rearranging physical spaces [P42], and
-by holding the meetings of the rollout team in public [P16].
+by holding the meetings of the roll-out team in public [P16].
 By sharing experiences and status of the transformation the organization was
@@
 Organizing a large number of people requires some management and leadership.
-Similarly, transforming the way of working of a large group requires coordinaion
+Similarly, transforming the way of working of a large group requires coordination
 and leadership [P42, P31]. In addition to the leadership provided by coaches,
@@
 the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
-described that the change was lead by a competent rollout team, which had
+described that the change was lead by a competent roll-out team, which had
 representatives from all parts of the organization [P16, P33]. Finally, the
@@
 The emergence of an agile community was reported as a success factor in the
-transformation [P3]. A community was also seen as indispansable as it has power
+transformation [P3]. A community was also seen as indispensable as it has power
 to overcome impediments that would be too large for individuals to affect [P47].
-The agile communities achieved to raisie awareness of agile methods in the
+The agile communities achieved to raise awareness of agile methods in the
 organization [P42, P41, P47], and spawned eager followers who spread the word
@@
 large context [P43]. Another case highlighted creating and applying a structured
-rollout process to uniformly introduce a large nuber of teams to agile [P8].
+roll-out process to uniformly introduce a large number of teams to agile [P8].
 
@@
 [P8, P18, P27, P31, P33, P36, P43, P47, P49].
-This was reflected in statemets such as ``Adopting agile, or implementing any
+This was reflected in statements such as ``Adopting agile, or implementing any
 significant change, requires an executive’s sincere support.'' [P44],
@@
 
-Visible involvment of management was also reported to motivate and encourage
+Visible involvement of management was also reported to motivate and encourage
 employees to adopt the new model [P40]. For instance, the CTO organized training
-sessions personnally [P24], and the engineering VP frequently visited sprint
+sessions personally [P24], and the engineering VP frequently visited sprint
 demos [P11]. When the corporate level support for the agile initiative was
@@
 In order to gain support for agile several cases highlighted the need to educate
-managment [P8, P47].
+management [P8, P47].
 Without education managers might make harmful interventions in the process, or
@@
 where teams would independently develop and showcase the best agile practices.
-As a result the ownership of the methods was transfered to teams, and the new
+As a result the ownership of the methods was transferred to teams, and the new
 model gained a secure foothold. [P41]
@@
 It was also thought that proceeding with a top-down mandate would have made
-teams use a uniform process [P41]. This would have been suboptimal, as it was
+teams use a uniform process [P41]. This would have been sub-optimal, as it was
 important that the practices were tailored for each team and project [P41].
@@
 fully on getting the agile model implemented. In another case external coaches
-were described to better spot places for correctino in the agile model, and
+were described to better spot places for correction in the agile model, and
 their advice was received better because they were considered impartial when
@@
 It would require too big of an interpretation to combine all the various cases
-into one theoretical model. We have chosen to present transformatoins using an
+into one theoretical model. We have chosen to present transformations using an
 elementary two-axial classification of transformations, and listings of

### 2014-05-30 journal/article.tex
Editing: T: Training, coaching

@@
   principles'' is referred to?
+
+* What is coaching? Why is it done?
 
@@
 
-\subsubsection{Transformation activities}
-
-During transformation periods organizations conducted activities aimed towards
-creating change. Examples on such activities are piloting new models, seminars
-and workshops, training, and coaching on new models.
+\subsubsection{Training, coaching, and awareness}
+
+During transformation periods organizations invested in activities aimed to
+incorporate the new model into the thinking of people. People were new to agile
+methods, and a common understanding needed to be created. Training, coaching,
+and public events were used for communicating the new model to a significant
+part of the organization. Describing the use of training and coaching was a very
+prominent facet in transformation cases.
+Table \ref{table:transformation_training} summarizes activities that
+organizations use for deepening the understanding and awareness of the new
+development process being applied.
 
@@
                 P33, P36, P39, P40, P41, P45, P46, P47, P50, P52  \\  % 21
-        Pilot programs and other piloting   & 
-                P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
-                P36, P38, P40, P41, P43, P45, P46, P47  \\  % 22
         \bottomrule
     \end{tabular}
-    \caption{Summary of activities aiming to advance the transformation}
-    \label{table:transformation_activities}
+    \caption{Summary of activities for building awareness and knowledgeability
+             in the agile way of working}
+    \label{table:transformation_training}
 \end{table}
@@
 
-% Pilot programs
-
-Pilot projects and programs were a common practice at the start of the
-transformation. Piloting was discussed in half of the cases.
-
-Why is piloting started?
-
-Who?
+% ---------------------------------------------------------------------
+
+% Coaching
+
+Why is coaching used?
+
+Coaches are who?
+
+Who receives coaching --> Teams, very few cases with some other roles
 
@@
 
-What kind?
-
-What amount of piloting is applied?
+How is coaching done?
 
@@
 
-Big bang + Piloting
-[P11, P16, P18, P41]
-
-% ---------------------------------------------------------------------
-
-% Coaching
-
-Why is coaching used?
-
-Coaches are who?
-
-Who receives coaching --> Teams, very few cases with some other roles
-
-When?
-
-How is coaching done?
-
-Did it work?
-
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Coaches were driving the agile push [P2].
-
-Coaching was considered very important at project start. Otherwise there may be
-severe long term consequences [P8].
+% Coaching was a typical way for steering the agile adoption.
+In addition to training, a coaching approach was usually used to teach how to
+work in an agile environment. As agile lacks rigorous prescriptions and builds
+on human interactions, it is best learned in practice while a coach is giving
+advice.
+
+Coaches were driving the agile push [P2, P36], and advocating agile methods [P45].
+
+Coaching was considered very important at project start [P8, P33, P46]. Otherwise
+there may be severe long term consequences [P8].
 
 Coaching was also seen as an important success factor [P29, P33].
-
-The agile adoption required coaches [P47].
+The coach ensured that transformation would be successful [P16].
+
+Having a impartial third party to rely on was a key success factor [P29].
+
+Having too little coaches [23].
+
+The agile adoption required coaches, especially at the start of agile projects
+[P8, P46, P47].
 
@@
 cases.
+
+As the agile way of working penetrated the organization the need for coaching
+was decreasing [P36].
+
+Some cases presented building of a coach community as an integral part of the
+agile adoption [P47, P23].
 
@@
 * Coaches: Coaches from a subcontractor
-* How: The coached drive the project
+* How: The coaches drive the project
 * How: Mentor in-house leads to take lead as soon as possible
@@
 % ---------------------------------------------------------------------
+
+\subsubsection{Piloting}
+
+It was typical that before taking the transformation to full scale the new model
+was tested in use in the organization specific environment.
+
+% Pilot programs
+
+Pilot projects and programs were a common practice at the start of the
+transformation. Piloting was discussed in half of the cases.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
+                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
+        Activity           &  Described in  \\
+        \midrule
+        Pilot programs and other piloting   & 
+                P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
+                P36, P38, P40, P41, P43, P45, P46, P47  \\  % 22
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of activities aiming to advance the transformation}
+    \label{table:transformation_activities}
+\end{table}
+
+Why is piloting started?
+
+Who?
+
+When?
+
+What kind?
+
+What amount of piloting is applied?
+
+Did it work?
+
+Big bang + Piloting
+[P11, P16, P18, P41]
 

### 2014-05-30 journal/article.tex
Work in progress: Coaching, internal coaches

@@
         Coaching    &
-                P1, P2, P3, P6, P8, P9, P16, P18, P23, P27, P29,
-                P33, P36, P39, P40, P41, P45, P46, P47, P50, P52  \\  % 21
+                P1-P3, P6, P8, P9, P11, P16, P18, P23, P27, P29,
+                P31-P33, P36, P39-P41, P45-P47, P49, P50, P52  \\  % 25
+        Internal personnel is coaching  &
+                P2, P3, P6, P11, P16, P23, P29, P31, P32,
+                P39, P41, P45, P47, P49, P50, P52  \\    % 16
         \bottomrule
@@
 % Coaching was a typical way for steering the agile adoption.
+% What
 In addition to training, a coaching approach was usually used to teach how to
 work in an agile environment. As agile lacks rigorous prescriptions and builds
-on human interactions, it is best learned in practice while a coach is giving
-advice.
-
-Coaches were driving the agile push [P2, P36], and advocating agile methods [P45].
-
-Coaching was considered very important at project start [P8, P33, P46]. Otherwise
-there may be severe long term consequences [P8].
-
-Coaching was also seen as an important success factor [P29, P33].
-The coach ensured that transformation would be successful [P16].
-
-Having a impartial third party to rely on was a key success factor [P29].
-
-Having too little coaches [23].
-
-The agile adoption required coaches, especially at the start of agile projects
-[P8, P46, P47].
-
-Internal coaches were required for making the change grow [P47].
+on human interactions, it is best learned in practical situations while a coach
+is giving direction.
+The main task of coaches was to guide development teams in their daily work.
+Some cases described that roles such as the product owner or scrum master
+received specialized coaching [P33, P52].
+In addition to helping teams, coaches were driving the agile push [P2, P36,
+P49], and advocated agile methods in general [P45].
+
+% Why
+Coaching was considered necessary, especially in the beginning of agile projects
+[P6, P8, P33, P46, 47]. Brown [P8] stated that without adequate coaching there
+would have been severe long term consequences [P8]. The application of coaching
+was seen as an important success factor [P29, P33], and the job of the coach was
+described to be to ensure that the transformation would be successful [P16].
+One viewpoint was that the coach was an impartial third party who could be
+relied on when resolving transformation impediments [P29].
+
+% Having too little coaches [23].
+% Coaches were also involved in training [P27].
+
+% How coaching evolves
+As the agile way of working penetrated the organization the need for coaching
+was decreasing, and coaching could be reduced [P18, P36]. However, some cases
+described that the demand for coaching became permanent in order to keep the
+development model healthy and to continuously pursue improvement [P6, P23].
+Silva and Doss [P47] describe that internal coaches were required for making the
+change grow. Over time the arrangement shifted from having external consultants
+as coaches to creating an internal coaching community.
+
+Some cases presented building of a coach community as an integral part of the
+agile adoption [P6, P23, P32, P47].
 
 To be able to coach efficiently it was required that the coach had detailed
-understanding of the company's processes and systems [P23]
-
-Coaches were also involved in training [P27].
-
-Scrum masters that had been trained earlier were coaching the teams in some
-cases.
-
-As the agile way of working penetrated the organization the need for coaching
-was decreasing [P36].
-
-Some cases presented building of a coach community as an integral part of the
-agile adoption [P47, P23].
+understanding of the company's processes and systems [P23, P39].
+
+The maturity of coaches grew over time [P32].
+
+% Scrum masters that had been trained earlier were coaching the teams in some
+% cases.
 
@@
 * How: inspect and adapt
+* Particular: Internal coaches had better understanding of the company's procedures
 
@@
 * Coach: In some cases an experienced Scrum master was brought to help the pair
+
+% ---------------------------------------------------------------------
+
+% Inhouse coaching
+
+What?
+
+Why?
+
+Benefits?
+
+Differences?
+
+[P2]
+* Why: There were too few coaches, so new ones were trained internally
+
+[P3]
+* Benefit: The internal coaches were accessible. You could easily acquire coaching for your team
+
+[P6]
+* What: There is a train the trainer program
+* Why: Coaches can be expensive, more coaches can scale up agile more quickly
+
+[P9]
+* There was a lack of qualified coaches
+* Why: hiring good external coaches was difficult
+* It was hard to make the internal coaching resources work as well
+
+[P11]
+* Coaches: The Scrum masters served as internal coaches
+
+[P16]
+* Coaches: The roll-out team coached
+
+[P23]
+* A coach community evolved
+* Why: The organization was so large
+* How: Paired coaching
+
+[P31]
+* A certain person was hiread as senior agile coach, and he stayed through out the transformation
+* Coach: A person was hired permanently
+
+[P32]
+* Particular: The maturity of coaches grew over time
+* Why: There were so many people to educate
+
+[P39]
+* Benefit: Internal coaches had understanding of procedures and communication policies
+
+[P45]
+* What: After 8 months the coaches started training future Scrum mentors
+
+[P47]
+* Why: There was an internal need. It just was necessary to have coaches.
+* How: First there were some external coaches, then an internal community was built
+* Coaches: People who were passionate to teach agile
+
+[P49]
+* Why: To scale up agile implementation
+* Why: The internal coaches create the change
 

### 2014-05-30 journal/article.tex
T: Coaching

@@
 
-% ---------------------------------------------------------------------
-
 % Coaching
 
-Why is coaching used?
-
-Coaches are who?
-
-Who receives coaching --> Teams, very few cases with some other roles
-
-When?
-
-How is coaching done?
-
-Did it work?
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Coaching was a typical way for steering the agile adoption.
 % What
@@
 was decreasing, and coaching could be reduced [P18, P36]. However, some cases
-described that the demand for coaching became permanent in order to keep the
+described that the demand for coaching had become permanent in order to keep the
 development model healthy and to continuously pursue improvement [P6, P23].
-Silva and Doss [P47] describe that internal coaches were required for making the
-change grow. Over time the arrangement shifted from having external consultants
-as coaches to creating an internal coaching community.
-
-Some cases presented building of a coach community as an integral part of the
-agile adoption [P6, P23, P32, P47].
-
-To be able to coach efficiently it was required that the coach had detailed
-understanding of the company's processes and systems [P23, P39].
-
-The maturity of coaches grew over time [P32].
+The maturity of coaches was described to grow over time, as the agile
+transformation progressed [P32].
+
+% Internal coaching
+The coaching positions were most often held by internal staff.
+Internal coaches were required for making the change grow, and to scale up the
+agile implementation [P47, P49]. Internal coaches had also the benefit of
+understanding of the company's processes and systems [P23, P39], and teams were
+able to access them easily [P3, P41]. Some cases presented building of a coach
+community as an integral part of the agile adoption [P6, P23, P32, P47].
 
@@
 
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % ---------------------------------------------------------------------
-
-[P1, P2]
-* Why: best for introducing agile methods
-* Why: They drive the push
-* How: Coach in the team
-
-[P3]
-* Why: The change leader persuaded to introduce coaching
-* Coach: The change leader
-* Who: teams
-* How: Teams were coached
-
-[P6]
-* Why: It helps to have people with experience
-* Coach: An internal change leader, a team of coaches
-* Who: Teams
-* How: Start heavy a kick-off, and later draw back
-* Particular: There were little coaches at start
-* Particular: The coaches showed positive results, and management was convinced to invest in coaching
-
-[P8]
-* Why: Intensive coaching is necessary 4-6 months at start
-
-[P9]
-* Coach: A hired consultant specialist
-* How: The coach embodied the best agile principles
-
-[P16]
-* Why: To ensure transformation is successful
-* Coach: The cross-functional rollout team
-
-[P18]
-* Coach: An agile consultant
-* When: For over a year
-* How: Accompanied the introduction for over a year, trained on specific agile practices
-
-[P23]
-* Coaches: In-house coaching team, some external coaches
-* The coach community needed to be tended to specifically
-* If there were too few coaches things did not work
-* Pair coaching, apprentice coaching
-* Particular: The coaches needed to have specific knowledge of the company's systems and processes
-* Particular: Coaches in pairs --> training of coaches
-* Particular: Focus on coaching
-
-[P27]
-* Coaches: External coaches
-* How: Train on agile practices
-
-[P29]
-* Coaches: Agile consultants
-* Why: Having a impartial third party was a key success factor
---
-* Coaches: The pilot team, matured teams
-* Who: Teams
-* How: Helping other teams in getting started
-* How: Taking a formal mentoring role when it is required
-
-[P33]
-* Who: Product owners
-* Why: Crucial to make it successful
-* Particular: Coaches were brought in after the start of the transformation, when trouble already existed
-
-[P36]
-* Coaches: Coaches from a subcontractor
-* How: The coaches drive the project
-* How: Mentor in-house leads to take lead as soon as possible
-* How: Retrospect and adapt
-* When: The need for coaches disappeared gradually
-* The coach role was apparently seen as the lead of the project
-
-[P39]
-* Coach: External support, hired Chief Scrum masters
-* How: inspect and adapt
-* Particular: Internal coaches had better understanding of the company's procedures
-
-[P40]
-* Who: Teams
-* How: Coaching breakout sessions
-
-[P41]
-* Coach: Internal and external
-* Who: Teams
-* How: Coaching office hours
-* How: Have a forum where teams can ask for help, and to share their implementation
-* Particular: It was disliked to have peers as coaches, as they were not considered more knowledgeable
-
-[P45]
-* Coach: Internal small expert team
-* Who: Teams
-* How: Supporting, advocating agile development
-* How: Educate and support teams interested in Scrum
-* Particular: Later offering possibility to teach more Scrum mentors
-
-[P46]
-* The pilot teams were coached
-* Who: Teams that are about to start agile projects
-* How: A productized training workshop, and a jump start afterwards
-* How: Key team roles are coached in the start, so that the start is successful
-
-[P47]
-* Why: The agile adoption required coaches
-* Coaches: Initially external consultants
-* Coaches: Internal people started taking coach positions
-* Coaches: Apprentice coaches, guided by external experts
-* Particular: A structure was created around coach certification
-* Particular: A coach community
-
-[P50]
-* Coaches: Internal key roles, these received initial coach and role specific training
-* When: When teams were starting
-* How: Coached to facilitate start up of teams
-
-[P52]
-* Who: Paired line manager and Scrum master
-* How: Pair coaching the pair itself
-* Coach: In some cases an experienced Scrum master was brought to help the pair
-
-% ---------------------------------------------------------------------
-
-% Inhouse coaching
-
-What?
-
-Why?
-
-Benefits?
-
-Differences?
-
-[P2]
-* Why: There were too few coaches, so new ones were trained internally
-
-[P3]
-* Benefit: The internal coaches were accessible. You could easily acquire coaching for your team
-
-[P6]
-* What: There is a train the trainer program
-* Why: Coaches can be expensive, more coaches can scale up agile more quickly
-
-[P9]
-* There was a lack of qualified coaches
-* Why: hiring good external coaches was difficult
-* It was hard to make the internal coaching resources work as well
-
-[P11]
-* Coaches: The Scrum masters served as internal coaches
-
-[P16]
-* Coaches: The roll-out team coached
-
-[P23]
-* A coach community evolved
-* Why: The organization was so large
-* How: Paired coaching
-
-[P31]
-* A certain person was hiread as senior agile coach, and he stayed through out the transformation
-* Coach: A person was hired permanently
-
-[P32]
-* Particular: The maturity of coaches grew over time
-* Why: There were so many people to educate
-
-[P39]
-* Benefit: Internal coaches had understanding of procedures and communication policies
-
-[P45]
-* What: After 8 months the coaches started training future Scrum mentors
-
-[P47]
-* Why: There was an internal need. It just was necessary to have coaches.
-* How: First there were some external coaches, then an internal community was built
-* Coaches: People who were passionate to teach agile
-
-[P49]
-* Why: To scale up agile implementation
-* Why: The internal coaches create the change
 

### 2014-05-31 journal/article.tex
T: External consultants: Work in progress

@@
 improve their performance, but the methods are originally designed for small and
-individual teams. This poses challenges when introducing agile to large
-organizations where development teams can not be autonomous. We reviewed 30
-experience reports and case studies describing the process of taking agile
-methods into use in large organizations. The primary motive for agile adoption
-is the need to improve performance. For a successful transformation the entire
-organization must be aligned and all organizational units must participate.
+individual teams. This creates unique challenges when introducing agile to large
+organizations where development teams can not be autonomous. We conducted a
+systematic literature review including publications on 40 industry cases,
+describing the process of taking agile methods into use in large organizations.
+
+The primary motive for agile adoption is the need to improve performance. For a
+successful transformation the entire organization must be aligned and all
+organizational units must participate.
 Creating an agile community in the organization will help the change to stick.
@@
 would have been severe long term consequences [P8]. The application of coaching
-was seen as an important success factor [P29, P33], and the job of the coach was
-described to be to ensure that the transformation would be successful [P16].
-One viewpoint was that the coach was an impartial third party who could be
-relied on when resolving transformation impediments [P29].
+was seen as an important success factor [P18, P29, P33], and the job of the
+coach was described to be to ensure that the transformation would be successful
+[P16]. One viewpoint was that the coach was an impartial third party who could
+be relied on when resolving transformation impediments [P16, P29, P44].
 
@@
 development model healthy and to continuously pursue improvement [P6, P23].
+External coaches moved out and internal coaches took their place [P40].
 The maturity of coaches was described to grow over time, as the agile
@@
 
+% External consultants
+
+Who are they?
+
+Why are they hired?
+
+When in the transformation?
+
+What did they do?
+
+Benefit was what?
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
 
+External consultants were employed especially at the start of the
+transformation. At the time it was realized that the organization should change
+the way of working a consultant was typically hired to give an introduction to
+agile.
+
+It was described that external consultants were invited every now and then to
+bring in fresh ideas to the organization [P6].
+
+External consultants were especially used in the beginning to ignite interest in
+agile methods [P9].
+
+External consultants assessed the current state of the organization, and
+recommended what kind of model to apply [P51].
+
+It was recommended to bring in extenal help as early as possible, before truoble
+starts rising in the transformation [P16].
+
+People were also more comfortable when the advice was given by impartial persons
+external to the organization [P16, P29].
+
+A particular wish in using consultants was that they would speed up the adoption
+[P34]. Consultants were already experts in the agile methods and could quickly
+tell solutions to problems.
+
 % ---------------------------------------------------------------------
+
+[P5]
+* Who: Dean Leffingwell
+* What: Advise, and tell to do what's next
+
+[P6]
+* When: Pilot teams
+* Who: Ken Schwaber, etc.
+* What: Early coaching
+--
+* What: Coach Scrum masters
+* When: Continue inviting agile experts
+* What: Bring fresh ideas
+--
+* Why: It can save money and trouble if internal skills are scarce
+* How: Apply them wisely
+
+[P7]
+* When: In the introduction
+* Who: Craig Larman
+* What: Introduce, train certified Scrum masters
+
+[P8]
+* When: Earlier
+* What: External consulting had been contracted
+--
+* When: In the transformation
+* What: Used a wide range of external consultants
+
+[P9]
+* When: In the beginning
+* Who: Ken Schwaber, Jeff Sutherland
+* What: To ignite interest
+--
+* When: In the adaption
+* Who: Experienced agile coaches
+* What: To train Scrum on a regular basis
+
+[P16]
+* When: In the beginning
+* Who: Industry experts
+* What: The cross functional team reached out to consultants
+--
+* When: In the transformation
+* Who: Professional trainers
+* What: 25\% of R\&D sent to professional training 
+* Who: Experienced consultants
+* What: assist teams POs and ScMs
+* Benefit: A key success factor
+* Benefit: Their experience helped drive the transformation
+* Benefit: People were more comfortable receiving advice from outsiders
+
+[P17]
+* Who: Rally software
+* What: Provide software and services to accelerate adoption
+* Benefit: Top notch consulting is critical for success
+
+[P18]
+* When: For the pilot team
+* Who: Agile consultants
+* What: A one day training, acting as a coach
+
+[P19]
+* When: In the beginning
+* Who: Ken Schwaber
+* What: Train 60 Scrum masters
+
+[P20, P21]
+* When: In the beginning
+* Who: Expert agile consultant
+* What: Assessment of the current process, educating management and agile task force
+* What: Recommended to adopt Scrum
+* What: Supervised the agile course held to initial team members
+
+[P23]
+* When: In the transformation, when looking for more coaches
+* Who: Exoftware Ltd
+* What: A partnership to promoting and coaching agile delivery
+
+[P25]
+* When: During the first year
+* Who: External consultants
+* What: Participate in the transformation team, support transformation of teams
+--
+* When: Later
+* Who: Acclaimed agile professionals, external speakers
+* What: Give presentations, lead workshops
+
+[P27]
+* When: In the beginning
+* Who: Renowed industry experts, ThoughtWorks
+* What: Educate employees on agile practices
+--
+* When: In the first couple of months
+* What: Coaching the team
+* Particular: The teams were sceptical of having external consultants embedded
+--
+* When: In the middle of the transformation
+* Who: Agile consultants who were familiar with all phases of the transformation
+* What: Workshops, coaching
+* What: Enabled forward progress, avoided time wasted in searching solutions
+
+[P29]
+* When: Introducing Scrum to the department, later internal people took over
+* What: coaching teams
+* Particular: Good having an impartial third party to fall back on
+
+[P30]
+* When: In beginning, and through out
+* Who: Jeff Sutherland
+* What: Scrum training for most of the engineering group
+* What: Sell the agile approach in the beginning
+* Benefit: An outside expert sells the agile approach in the beginning
+* It worked out with the external expert where there had been failures earlier
+
+[P31]
+* When: Teams had implemented agile, it was time to renew the requirements management
+* Who: Enthiosys, management consulting company, with deep agile experience
+* What: Help in creating a system to deliver end to end in an agile way
+
+[P34]
+* When: In the beginning
+* Who: ThoughtWorks consulting company
+* What: Support project management and development
+* Why: Accelerate use of agile practices
+
+[P35]
+* Who: Agile contractors
+* Why: To speed up the development
+
+[P37, P38]
+* When: In the beginning
+* Who: A local consulting firm
+* What: Assist with the transformation
+[P38]
+* Why: To give a shining example of agile
+* What: Mentoring and helping teams
+
+[P39]
+* When: In the beginning
+* Who: Two consulting companies
+* What: Coach implementing agile
+
+[P40]
+* When: In the beginning, they moved out after three months
+* Who: Consultants
+* Why: The key projects were not to be affected, so people with previous
+       knowledge were needed to create a roadmap
+* Benefit: They moved quickly, as they had previous experience of similar transformations
+
+[P41]
+* When: In the beginning
+* Who: known agile coaches
+* What: Create a vision
+* What: Help in developing and leading a transition workshop
+
+[P43]
+* When: In the beginning
+* Who: Consultant firms
+* What: Give initial insights, give insights during the transformation
+
+[P44]
+* When: In the beginning
+* Who: Ken Schwaber
+* What: Assess the current state
+* What: Hold workshops, certify Scrum masters
+* Benefit: Get impartial feedback from an outside source
+
+[P46]
+* When: In the beginning
+* Who: Ken Schwaber
+* What: Organizational coach
+* Coach pilot teams in the beginning
+
+[P47]
+* When: in the beginning
+* Who: External consultants
+* What: Coach the apprentice coaches
+* Particular: The need for the external coaches reduced over time, and gave way for internal
+* Benefit: Support the early growth of apprentice coaches
+
+[P50]
+* Who: Consultants from Scrum foundation
+* What: Coaching, training
+
+[P51]
+* When: In the beginning
+* Who: External consultant
+* What: Assess the current state, and recommend what to do
+* What: Provide training
 
@@
         Coaching is essential in transformation  &
-                P3, P6, P8, P9 P16, P17, P18, P30, P42, P44, P47, P48  \\
+                P3, P6, P8, P9 P16, P17, P18, P29, P30, P33, P42, P44, P47, P48  \\
         \bottomrule
@@
 an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
-that coaching was essential for success in transformation [P8, P18]. Also,
-without coaching teams can do damage to the agile transformation if techniques
-are applied improperly [P9].
+that coaching was essential for success in transformation [P8, P18, P29, P33].
+Also, without coaching teams can do damage to the agile transformation if
+techniques are applied improperly [P9].
 
@@
 
+* It is difficult to distinguish between team and organization. In most cases
+  when talking about teams it refers to a handfull of people familiar to each
+  other. However, in some cases it was evident that an entire department was
+  referred to as a team.
+
+* Projects and organizations. How to distinguish between a separate adoption in
+  the beginning of each project, and an organization wide adoption.
+
 % ---------------------------------------------------------------------

### 2014-05-31 journal/article.tex
T: External consultants

--- 2014-05-30 journal/article.tex
@@
 Even if any activities aiming for transformation can be considered as
-investments in themselves, the transformation process included specific
-investments in some tangible entities. These investments were focused on better
-enabling agile software development, and highlight the concerns that
-organizations had when implementing an agile model. Typical investments were new
-management tools, improvements in testing environments, and updating of the
+investments in themselves, the transformation process included some specific
+tangible investments. These investments were focused on better enabling agile
+software development, and highlight the concerns that organizations had when
+implementing an agile model. Typical investments were bringing in external help,
+new management tools, improvements in testing environments, and updating of the
 physical working spaces. Table \ref{table:transformation_investments} summarizes
@@
     \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.42\textwidth}
-                     >{\raggedright\arraybackslash}p{0.50\textwidth} }
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.35\textwidth}
+                     >{\raggedright\arraybackslash}p{0.57\textwidth} }
         Investment for transformation      &  Described in  \\
         \midrule
+        Using external consultants  &
+                P5-P9, P16-P21, P23 P25, P27, P29-P31,
+                P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
         Agile project management tools  &
@@
 \end{table}
+
+% External consultants
+
+It was very typical that external help was brought in to help with the
+transformation.
+External consultants were employed especially at the start of the
+transformation. At the time it was realized that the way of working should be
+changed a consultant was often hired to assess the current state and give a
+recommendation on how to proceed [P16, P20, P41, P43, P41, P44, P51].
+In the beginning external consultants were used to ignite interest in agile
+methods, and provide initial education [P7, P9, P27, P30, P43].
+It was recommended to bring in extenal help as early as possible, when the
+internal agile expertise is lacking and before transformation challenges appear
+[P6, P16, P30].
+
+When the transformation was getting on its way, agile consultants were typically
+leading training sessions, arranging agile workshops, and coaching teams.
+In some cases consultants helped specifically with scaling up the use of agile
+beyond single teams [P31, P34].
+Consultants were used mostly in the first stages of the transformation.
+The need for external consultants reduced over time, and their tasks were
+transfered to internal positions [P29, P47].
+It was described that as the agile journey progressed, external consultants were
+invited every now and then to bring in fresh ideas to the organization [P6].
+
+External consultants were regarded to bring in critical assistance in
+transforming the organization [P6, P16, P17]. Qualities in external consultants
+that organizations looked for was leveraging their experience to bring more
+speed to development and agile adoption [P17, P27, P34, P35, P40], and to bring
+in agile skill when only little existed in-house [P27, P47]. It was considered
+beneficial having experts with previous experience of transforming organizations
+[P16, P27].
+Still one important aspect in using consultants was that people were more
+comfortable when the advice was given by impartial persons external to the
+organization [P16, P29, P44].
+
+% In one case the teams were sceptical of having external consultants embedded
+% in temas [P27].
 

### 2014-05-31 journal/article.tex
T: Seminars and workshops: Work in progress

@@
                 P31-P33, P36, P39-P41, P45-P47, P49, P50, P52  \\  % 25
-        Internal personnel is coaching  &
-                P2, P3, P6, P11, P16, P23, P29, P31, P32,
-                P39, P41, P45, P47, P49, P50, P52  \\    % 16
+        In-house coaching   &
+                P1-P3, P6, P11, P16, P23, P29, P31, P32,
+                P39, P41, P45, P47, P49, P50, P52  \\    % 17
+        Seminars and workshops   &
+                P1-P4, P7, P9, P10, P11, P16, P23, P25,
+                P26, P27, P32, P41, P44, P46, P47   \\   % 18
         \bottomrule
@@
 
+% Seminars and public events
+
+When?
+
+What?
+
+Why?
+
+For whom?
+
+Benefit?
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
 
+Organizations typically strived to increase the agile awareness by events
+promoting agile, ranging from public seminars to workshops. These events were
+aimed at a general audience, consisting of all organizational groups. Events
+were arranged especially in the beginning of the agile journey. In some cases
+the demand and interest in public events was maintained, and the events were
+continued.
+
+Some cases describe continually arranged agile events [P47].
+
+Organizations took the opportunity to promote the new way of working.
+
+Some organizations productized the training events, so that they could be
+replicated for a large number of people [P46].
+
+The ``how things are done here'' attitude was being broken down by conducting
+workshops [P16].
+Workshops were used to get people better involved [P16].
+
+The purpose of these events was to evangelize and expand the agile model
+throughout the organization [P10].
+
+Hands-on workshops were used as a teaching method as they could be made more
+engaging than traditional lectures [P32, P41].
+
+Another benefit of open events was that the same message was delivered to many
+people at the same time []. As everyone could heard the same questions and
+answers the understanding was consistent [P41].
+
 % ---------------------------------------------------------------------
+
+[P1, P2]
+* Purpose: Translating agile principles
+* What: Special workshops, presentations, events of public recognition eg. agile awards
+* For: Rest of business
+
+[P3]
+* What: A series of talks
+* For: People in the organization
+* Benefit: People kept comming back, even if the benefits were not tangible
+--
+* When: When there was enough interest
+* What: Scrum gathering
+
+[P4]
+* Purpose: Coordinating activities for increasing knowledge and understanding
+* What: Competency day, weekly meetings, seminars and presentations for customers
+
+[P7]
+* When: in the beginning
+* Purpose: To identify initial problems, practical implementation planning
+* What: Workshops
+* For: Teams
+
+[P9]
+* When: In the beginning
+* Purpose: To ignite interest
+* What: Agile experts talks, public open sessions
+--
+* When: During pilots
+* What: Practical workshops, real-world examples
+* Benefit: They became popular, even outside of pilot teams
+* Benefit: Common understanding across the organization
+--
+* When: Further in the transformation
+* What: Agile talks, External agile community events
+
+[P10]
+* When: In the beginning, when the agile transformation was on
+* What: Lunch sessions, classes in different agile disciplines
+* Purpose: To evangelize the agile model
+
+[P11] (?)
+* When: In the beginning
+* Purpose: To train people
+* What: A short introductory presentation
+
+[P16]
+* When: During the transformation
+* Purpose: To involve individuals better
+* What: Workshops to go over the agile way of working, and to identify problems
+* Benefit: The workshops should have been used earlier to improve the transformation
+
+[P23]
+* When: In the transformation
+* What: Agile roadshow, Agile program days, Agile learning projects
+* Purpose: Promote awareness, give a one day deep dive, project level engagements
+--
+* When: Continued
+* What: Based on the roadshow, knowledge share events
+* Why: Because of the initial interest
+
+[P25]
+* When: During the first year
+* What: introductory presentation
+--
+* When: Later
+* What: Workshops lead by external speakers
+
+[P26]
+* Purpose: To correct people from sidetracks
+* What: Public lectures on agile, in all development centers
+
+[P27]
+* What: Lunch seminars, open space meetings
+* Purpose: To spread domain knowledge
+* Purpose: Brainstorm solutions to problems with conslutants
+
+[P32]
+* When: Continuous
+* What: Joining the dots events, 100 people at a time
+* What: Hands-on workshop
+* Purpose: Communicate the new ways of working, ensure understanding of agile for everyone
+* Benefit: A fun way of learning
+
+[P41]
+* When: Kick off for teams
+* What: Transition day programme
+* For: All teams
+* Benefit: All teams could participate, and get a common understanding
+* Benefit: Everyone hears the same questions
+
+[P44] (?)
+* When: In the beginning
+* What: An introductory workshop by an agile consultant
+
+[P46]
+* When: During the transformation
+* What: A sequence of agile training
+* What: Workshops
+* Purpose: To set the team straight
+
+[P47]
+* What: Two meetings a month
+* What: An open meeting for everyone who wants to participate in shaping the agile model
+* What: Various agile topics
 

### 2014-05-31 journal/article.tex
T: Public events

--- 2014-05-31 journal/article.tex
@@
 % cases.
+
+% Seminars and public events
+
+% What public events
+Organizations typically strived to increase the agile awareness by introducing
+series of events for promoting agile, ranging from public seminars to brown bag
+meetings and hands-on workshops. These events were aimed at a general audience,
+consisting of all organizational groups. Events were arranged especially in the
+beginning of the agile journey. In some cases the demand and interest in public
+events was maintained, and the events were continued.
+Some cases describe continually arranged agile events serving a central role in
+maintaining and advancing further the agile model [P23, P47].
+Training events were even productized so that they could be replicated for a
+large number of people [P32, P46].
+
+% Content of events
+In addition to general training and enabling awareness, the purpose of agile
+events was to evangelize and expand the agile model throughout the organization
+[P1, P10, P23, P32, P46].
+Workshops were used to get people better involved [P16], and public lectures
+were used to set straight sidetracks in the agile implementation[P26].
+Hands-on workshops were used as a more engaging teaching method compared to
+traditional lectures [P32, P41]. Open events were good for delivering the same
+message to many people at the same time. As everyone could heard the same
+questions and answers the understanding was consistent [P9, P41].
+Workshops were arranged for identifying problems and planning the new model [P7,
+P16, P27].
 

### 2014-05-31 journal/article.tex
T: Communities: Work in progress

@@
         Informal spreading of agile awareness  &
-                P3, P6, P40  \\
+                P3, P6, P14, P40  \\
         \bottomrule
@@
 letting agile knowledge spread virally. New Scrum teams were emerging as people
-heard of the new methodology [P6].
+heard of the new methodology [P6]. In some cases interest groups were used to
+stimulate viral spreading of agile [P14].
 It was similarly described that in an organization with an introverted culture
@@
 explaining to others how they had succeeded in with the agile method [P40].
-Even casual hallway talks could spread the word effectively, and it was
-also easier to encounter people from other parts of the organization in an
-informal context [P3].
+Even casual hallway talks could spread the word effectively, and it was also
+easier to encounter people from other parts of the organization in an informal
+context [P3].
 
@@
         Seminars and workshops   &
-                P1-P4, P7, P9, P10, P11, P16, P23, P25,
+                P1-P4, P7, P9, P10, P18, P11, P16, P23, P25,
                 P26, P27, P32, P41, P44, P46, P47   \\   % 18
+        Communities, special interest groups  &
+                P3, P6, P7, P9, P10, P11, P12, P14, P32, P41,
+                P42, P45, P47, P52  \\   % 14 
         \bottomrule
@@
 Workshops were used to get people better involved [P16], and public lectures
-were used to set straight sidetracks in the agile implementation[P26].
+were used to set straight sidetracks in the agile implementation [P26].
 Hands-on workshops were used as a more engaging teaching method compared to
@@
 
+% SIGs, CoPs, and community
+
+When?
+
+What?
+
+Why?
+
+For whom?
+
+Benefit?
+
+Events and coaching activity were coordinated forms of promoting the new model,
+but the agile transformation gave also birth to communities [P3]. Typically
+communities were emerging later in the transformation, when the rollout team
+was starting to step aside [P3, P41].
+In some cases the communities were created by passionate individuals without
+corporate intervention, and participation in communities was voluntary [P47].
+Communities were created when the mass of people involved became too large for
+formal control [P52].
+In other cases the forming of communities was stimulated by the organization
+[P12, P45].
+Companies strived to create communities of practice [P42] or special interest
+groups [P41] in order to reinforce continuous learning in the organization.
+Communities helped people to form a learning habit, which promoted continuous
+improvement in the organization [P52].
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ---------------------------------------------------------------------
+
+[P3]
+* Who: Employees
+* When: During transformation
+* When: The community stayed
+* What: There were e-mail lists
+* What: Discssion forums were used for direct conversation instead of using the change leader
+* Particular: It is claimed that communities are often used
+* Benefit: Reinforcing agile practices
+
+[P6]
+* Who: People
+* What: Mailing list for sharing information on agile
+* What: Pockets of committed communities
+* Purpose: Leveraging promoting force of people in the trenches
+
+[P7]
+* Who: Scrum masters
+* When: Beginning
+* What: Peer groups for Scrum masters
+* Benefit: The peer groups evolved into a Scrum master community
+* Benefit: The community was essential in growing managers to agile roles
+
+[P9]
+* When: In the transformation
+* What: People attended to external communities, there were internal champions
+* Benefit: The movement created interest in agile
+
+[P10]
+* Who: tems
+* What: Teams started communicating through e-mail groups
+* What: Use of wikis, informal gatherings
+* Particular: The interest was voluntary, the company started creating official programme
+
+[P11]
+* Who: Scrum masters, people managers
+* What: Scrum masters met regularly --> Helped coordinate work
+* Particular: Product owners should also have formed their own community
+
+[P12]
+* Who: Everyone
+* Why: The organization introduced a matrix structure
+* What: People met peers from other parts of the company, there were discipline specific meetings
+* Benefit: People started talking with each other
+* Particular: The agile culture was initiated
+
+[P14]
+* What: Agile support group
+* What: Attempted to support viral spreading of agile
+
+[P32]
+* What: A coaching community
+* Particular: The community took time to build, as coaching is a skill not learned easily
+
+[P41]
+* When: The transition team moved aside
+* Who: Scrum master, teams
+* What: Special interest groups were formed to reinforce the new practices
+* Purpose: Create learning organization
+
+[P42]
+* Why: Needed for creating a learning organization and evolve the organization
+* What: Communities of practice
+* What: Bring in as many ideas as possible
+* Purpose: Create learning organization
+
+[P45]
+* What: All data from teams was gathered
+* Purpose: Set up a network of information for Scrum masters and product owners
+* Purpose: Create a round table of experts
+
+[P47]
+* When: Broader adoption of agile
+* Why: Coaches saw a need to create an environment to support learning and influence agile
+* What: Coach community, supported by a particular experienced team
+* Particular: voluntarily participation
+* Particular: It was created by passionate individuals rather than management decision
+* What: The community organizes agile events
+
+[P52]
+* When: There was many scrum masters
+* What: Community of Scrum masters
+* Why: The team of Scrum masters became so big that it was hard to coordinate formally.
+       They started changing experiences less formally, in order to become better in the task.
+* Benefit: When people started contributing from the wider scope a positive feedback loop was created
+* Benefit: People developed a learning habit, continuous improvement in the organization
 

### 2014-05-31 journal/article.tex
T: Communities

--- 2014-05-31 journal/article.tex
@@
         Seminars and workshops   &
-                P1-P4, P7, P9, P10, P11, P16, P23, P25,
-                P26, P27, P32, P41, P44, P46, P47   \\   % 18
+                P1-P4, P7, P9, P10, P18, P11, P16, P23, P25,
+                P26, P27, P32, P41, P44, P46, P47   \\   % 19
+        Communities, special interest groups  &
+                P3, P6, P7, P9, P10, P11, P12, P14, P32, P41,
+                P42, P45, P47, P52  \\   % 14 
         \bottomrule
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
+% SIGs, CoPs, and community
+
+Events and coaching activity were coordinated forms of promoting the new model,
+but the agile transformation gave also birth to communities [P3]. Typically
+communities were emerging later in the transformation, when the rollout team
+was starting to step aside [P3, P41].
+In some cases the communities were created by passionate individuals without
+corporate intervention, and participation in communities was voluntary [P47].
+Communities were created when the mass of people involved became too large for
+formal control [P52].
+In other cases the forming of communities was stimulated by the organization
+[P12, P45].
+Companies strived to create communities of practice [P42] or special interest
+groups [P41] in order to reinforce continuous learning in the organization.
+Communities helped people to form a learning habit, which promoted continuous
+improvement in the organization [P52].
 
@@
 [P11, P16, P18, P41]
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
 

### 2014-06-01 journal/article.tex
T: Piloting: Work in progress

@@
 
-What kind?
+What kind?   Phony, real projects
 
@@
 
-Did it work?
-
-Big bang + Piloting
-[P11, P16, P18, P41]
+What was found?
+
+Benefits?
+
+What after?
 
@@
 
+% Piloting
+
+A logical first step in an agile transformation is to try out the new methods in
+a pilot project. A large amount of the primary studies described how piloting
+was used in the beginning of the transformation.
+
+A piloting stage was seen as important for any transformation [P8, P21].
+
+Organizations typically start wit a pilot, and go through a gradual
+transformation [P11].
+
+Even reports which presented a big bang adoption described that the main
+transformation was preceeded by piloting [P11, P18, P41].
+
+Pilots could refer to considerable size. In one case over 100 people
+participated in 8 pilot projects [P8].
+
+A structured initiation approach for agile projects was created based on the
+pilots [P8].
+
+As a result of a successful pilot management approved use of agile [P14].
+
 % ---------------------------------------------------------------------
+
+[P6]
+* Who: Broad set of services
+* Amount: 4 teams
+
+[P7]
+* Who: Identified teams
+* Amount: Many teams
+* Result: Promising changes were seen
+* Particular: Those who sere not involved were not impressed
+
+[P8]
+* Why: A pilot is important in any transformation
+* Who: Teams with varying characteristics, a particular selection process
+* Kind: Piloting with real projects were seen as important, multi-team pilot projects
+* Amount: 8 projects with 100 people
+* Found: The startup activities were difficult
+* Particular: There was a selection process to screen teams
+* Particular: Many teams in one pilot project
+* Benefit: Piloting revealed critical spots: Start up, coaching, product owner role
+* Benefit: A structured initiation approach for agile projects was created based on the pilots
+
+[P9]
+* Who: Volunteer teams
+* What: Optional pilot programs were recommended to d ifferent parts of the organization
+* Amount: 4 projects
+* Benefit: The pilot provided insights in deficiencies
+* Benefit: The pilot demonstrated benefits: Breaking down silos
+* Particular: Some teams were mandated to use agile based on the pilots
+
+[P10]
+* Why: Some people just decided to give the new method a go
+
+[P11, P33]
+* Why: Mangement wanted to see if a new method could be achieved and be maintained
+* Amount: One team
+
+[P14]
+* When: The agile process had been defined
+* Amount: 7 pilot projects, 4 months
+* Found: The pilots were successful
+* Benefit: Management approved agile
+
+[P16]
+* Particular: A specific pilot project was decided to be avoided, so that there
+              would not be dissonances in the organization
+* A large amount of people were sent to training, and then they just started using agile
+
+[P18]
+* Amount: 1 team
+* Result: Management decided to go on with agile. Other teams started copying the approach
+* Result: More pilots were initiated
+* Amount: 20 teams, 3 months
+* Result: Continue agile where it works, in other cases not
+* Result: The pilots surfaced problems that had been ignored
+
+[P21] (!)
+* A pilot was not conducted
+* It was a critical mistake to ignore piloting
+
+[P22]
+* Amount: 1 team
+* Result: The pilot gave the final look for the team area, a Team Area concept was created
+
+[P24]
+* What: There was experimenting with Scrum
+* Result: Good results, and it was decided to roll out the method in all loactions
+
+[P29] (?)
+* Why: The method was piloted first within the team in order to give it a soft landing
+* What: Try the method without altering it too much
+* What: Piloting seems to mean that the methodology is just tried non-full-heartedly
+* What: The pilot project lasted long enough to face many of the problems that
+        agile had in the environment in question
+
+[P36]
+* Amount: A project
+* What: Agile was tried within the organization
+* Result: The project was successful and management became convinced of benefits
+
+[P38]
+* Why: Management needed to show benefits of agile greatly
+* Who: A project was selected particularly for the pilot
+* Who: It was a project deemed to be plausible for a good end result
+* Result: The pilot was successful
+* Result: Business was pleased with the result
+* Particular: People got false expectations based on the success of the pilot
 

### 2014-06-01 journal/article.tex
Work in progress

@@
 * Particular: People got false expectations based on the success of the pilot
+
+[P40]
+* Why: The IT organization was to pilot new methods as parts of the company strategy
+* Who: Pilot projects were identified by consultants
+* Amount: 3 projects
+* What: Piloting in critical projects
+* Why: To identify what management tooling can be used in agile
+* Result: Data showed that agile was working
+* Result: There were problems in the distributed pilot project
+* Result: The consultants moved out, and continuing agile became an internal matter
+* Result: Pilots showed where method was not working
+* Benefit: Results of pilots were used to shape the agile model
+* Benefit: The pilots showcased agile in a good way, there was good word-of-mouth
+* Benefit: People got interested
 

### 2014-06-01 journal/article.tex
Work in progress

@@
 * Particular: People got false expectations based on the success of the pilot
+* Who: The pilot team was an all star team
 
@@
 * Benefit: People got interested
+
+[P40]
+* Amount: 5 projects
+* Result: The pilots were successful
+* Result: The organization proceeded with wider adopiton
+
+[P43]
+* Amount: 8 months of experimenting
+* Result: Clear success in projects that used agile
+* Result: The organization was ready for full roll-out
+
+[P45]
+* Who: Teams by their own interest
+* Amount: A number of pilot projects
+* Result: There was important findings in applicability and limitations
+* Result: Data was gathered, and adjustments were made thereby
+* Particular: The pilot gave unrealistic expectations, and the scaling up was more difficult than the pilot projects
+
+[P46]
+* Why: To get started
+* Amount: 2 pilot projects
+* Result: The pilots got done?
+
+[P47]
+* Amount: Many areas of the company
+* Result: Pilots showed great successs
+* Result: Teams showed a strong desire to continue with agile
+* Result: Executive management was supportive
 

### 2014-06-01 journal/article.tex
T: Piloting

--- 2014-06-01 journal/article.tex
@@
 Mentioning big bang versus step wise [P16].
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% ---------------------------------------------------------------------
+
+% ---------------------------------------------------------------------
 
@@
 
-It was typical that before taking the transformation to full scale the new model
-was tested in use in the organization specific environment.
-
-% Pilot programs
-
-Pilot projects and programs were a common practice at the start of the
-transformation. Piloting was discussed in half of the cases.
+% It was typical that before taking the transformation to full scale the new model
+% was tested in use in the organization specific environment.
+% Pilot projects and programs were a common practice at the start of the
+% transformation. Piloting was discussed in about half of the cases.
+
+A logical first step in an agile transformation is to try out the new methods in
+a pilot project. A large amount of the primary studies described how piloting
+was used in the beginning of the transformation. Primary sources discussing
+the use of piloting are listed in \ref{table:transformation_piloting}.
 
@@
                      >{\raggedright\arraybackslash}p{0.61\textwidth} }
-        Activity           &  Described in  \\
-        \midrule
-        Pilot programs and other piloting   & 
+        Sources discussing viewpoints on piloting   & 
                 P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
-                P36, P38, P40, P41, P43, P45, P46, P47  \\  % 22
+                P36, P38, P40, P41, P43, P45, P46, P47  \\
         \bottomrule
@@
     \caption{Summary of activities aiming to advance the transformation}
-    \label{table:transformation_activities}
+    \label{table:transformation_piloting}
 \end{table}
 
-Why is piloting started?
-
-Who?
-
-When?
-
-What kind?   Phony, real projects
-
-What amount of piloting is applied?
-
-What was found?
-
-Benefits?
-
-What after?
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Piloting
-
-A logical first step in an agile transformation is to try out the new methods in
-a pilot project. A large amount of the primary studies described how piloting
-was used in the beginning of the transformation.
-
+Piloting was typically described as the first step in transformation, and
+yielded typically success and positive results [P7, P9, P14, P38]. In many cases
+management needed proof of benefits of agile methods, and successful pilots
+granted approval for agile [P14, P18, P24, P36, P41, P43, P47].
 A piloting stage was seen as important for any transformation [P8, P21].
 
-Organizations typically start wit a pilot, and go through a gradual
-transformation [P11].
-
-Even reports which presented a big bang adoption described that the main
-transformation was preceeded by piloting [P11, P18, P41].
-
-Pilots could refer to considerable size. In one case over 100 people
-participated in 8 pilot projects [P8].
-
-A structured initiation approach for agile projects was created based on the
-pilots [P8].
-
-As a result of a successful pilot management approved use of agile [P14].
-
-% ---------------------------------------------------------------------
-
-[P6]
-* Who: Broad set of services
-* Amount: 4 teams
-
-[P7]
-* Who: Identified teams
-* Amount: Many teams
-* Result: Promising changes were seen
-* Particular: Those who sere not involved were not impressed
-
-[P8]
-* Why: A pilot is important in any transformation
-* Who: Teams with varying characteristics, a particular selection process
-* Kind: Piloting with real projects were seen as important, multi-team pilot projects
-* Amount: 8 projects with 100 people
-* Found: The startup activities were difficult
-* Particular: There was a selection process to screen teams
-* Particular: Many teams in one pilot project
-* Benefit: Piloting revealed critical spots: Start up, coaching, product owner role
-* Benefit: A structured initiation approach for agile projects was created based on the pilots
-
-[P9]
-* Who: Volunteer teams
-* What: Optional pilot programs were recommended to d ifferent parts of the organization
-* Amount: 4 projects
-* Benefit: The pilot provided insights in deficiencies
-* Benefit: The pilot demonstrated benefits: Breaking down silos
-* Particular: Some teams were mandated to use agile based on the pilots
-
-[P10]
-* Why: Some people just decided to give the new method a go
-
-[P11, P33]
-* Why: Mangement wanted to see if a new method could be achieved and be maintained
-* Amount: One team
-
-[P14]
-* When: The agile process had been defined
-* Amount: 7 pilot projects, 4 months
-* Found: The pilots were successful
-* Benefit: Management approved agile
-
-[P16]
-* Particular: A specific pilot project was decided to be avoided, so that there
-              would not be dissonances in the organization
-* A large amount of people were sent to training, and then they just started using agile
-
-[P18]
-* Amount: 1 team
-* Result: Management decided to go on with agile. Other teams started copying the approach
-* Result: More pilots were initiated
-* Amount: 20 teams, 3 months
-* Result: Continue agile where it works, in other cases not
-* Result: The pilots surfaced problems that had been ignored
-
-[P21] (!)
-* A pilot was not conducted
-* It was a critical mistake to ignore piloting
-
-[P22]
-* Amount: 1 team
-* Result: The pilot gave the final look for the team area, a Team Area concept was created
-
-[P24]
-* What: There was experimenting with Scrum
-* Result: Good results, and it was decided to roll out the method in all loactions
-
-[P29] (?)
-* Why: The method was piloted first within the team in order to give it a soft landing
-* What: Try the method without altering it too much
-* What: Piloting seems to mean that the methodology is just tried non-full-heartedly
-* What: The pilot project lasted long enough to face many of the problems that
-        agile had in the environment in question
-
-[P36]
-* Amount: A project
-* What: Agile was tried within the organization
-* Result: The project was successful and management became convinced of benefits
-
-[P38]
-* Why: Management needed to show benefits of agile greatly
-* Who: A project was selected particularly for the pilot
-* Who: It was a project deemed to be plausible for a good end result
-* Result: The pilot was successful
-* Result: Business was pleased with the result
-* Particular: People got false expectations based on the success of the pilot
-
-% ---------------------------------------------------------------------
+% Cases described starting the transformation with a pilot, and then proceeding
+% gradually [P11].
+
+There was some variance of approaches in choosing the pilot projects.
+In some cases the project or teams to pilot the agile methods were chosen
+carefully [P7, P8, P38, P40], but others chose the pilot teams from volunteers
+[P9, P10, P45].
+Some organizations considered it important to pilot agile with business critical
+projects, as the applicability had to be proven in a releant environment [P8,
+P40].
+
+Piloting could refer to considerable size or time. In one case over 100 people
+participated in 8 pilot projects [P8], and another case described an
+experimentation period of 8 months [P34].
+It was also described that in order to gain broad insights piloting was applied
+in many different projects [P6], and also in projects involving many teams [P8].
+
+Cases described that piloting gave valuable insights that helped the
+transformation further.
+Pilot projects revealed organization specific problems and critical factors for
+the use of agile [P8, P9, P29, P40, P45].
+Based on the experience gathered in pilots a structured initiation approach for
+agile projects was created [P8], and the requirements for well functioning team
+areas were discovered [P22].
+The visible success of the pilots also encouraged new teams to copy and start
+applying the agile methods [P18].
+
+% What other things did the pilot teach?
+There were still a few particular insights regarding use of piloting.
+Fry and Greene [P16] describe a particular big bang adoption where piloting was
+omitted. A pilot project was avioded because having two models side by side was
+deemed to create organizational dissonance. Instead a significant amout of
+people completed agile training and the new model was introduced overnight
+[P16].
+In some cases the success of the pilot gave unrealistic expectations [P38, P45].
+Organizations should beware if the team is set up of all-star developers, as
+the results will likely be much better than average [P38].
+
+% Even reports which presented a big bang adoption described that the main
+% transformation was preceeded by piloting [P11, P18, P41].
 

### 2014-06-02 journal/article.tex
T: Transformation process

@@
 subsequent sections findings are presented organized according to the research
-questions. The methods used for extracting the results were described in section
+questions.
+
+The methods used for extracting the results were described in section
 \ref{sec:method}. Note that the term \emph{study} is used when referring to a
@@
                      >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
         Business reasons for starting a change   &  Described in  \\
@@
                      >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
         Management problems           &  Described in  \\
@@
                      >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
         Problems in the old process           &  Described in  \\
@@
                      >{\raggedright\arraybackslash}p{0.31\textwidth} }
+        \toprule
         Reasons to change           &  Described in  \\
@@
 We provide an answer to Research Question 2 in several parts. First the initial
-states of organizations are described. Secondly, we present a two-axial model
-for categorizing transformation types. In the rest of this section we discuss
-typical arrangements that relate to the change process, such as how training is
-arranged, how change is lead, and what investments are done.
-
-The challenge of giving a generic description on how transformations proceed is
-that every organization is unique, and further, publications tended to provide
-more focused viewpoints rather than describing the transformation process
-itself. Therefore it was not sensible to attempt constructing a detailed model
-on how large organizations adopt agile methods.
+states of organizations are described. Secondly, we present a model for
+categorizing transformation types. In the rest of this section we discuss
+typical arrangements that relate to the change process, including how training
+and piloting is arranged, how change is lead, and what investments are done.
+
+The fact that every organization is unique creates challenges in giving generic
+descriptions on how transformations proceed. Further, publications tended to
+provide more focused viewpoints rather than describing the transformation
+process itself. Therefore it was not sensible to attempt constructing a detailed
+model on how large organizations adopt agile methods.
 Table \ref{table:transformation_descriptiontypes} presents types of insights
@@
                      >{\raggedright\arraybackslash}p{0.40\textwidth} }
+        \toprule
         How transformation is described   &  Primary sources  \\
@@
 
-* Top down
-
-* Bottom up
-
-* Big bang
-
-* Step wise
-
-* Transformation is continuous, hard work, and takes time
-
-Mentioning big bang versus step wise [P16].
+Based on the primary studies we identified two dimensions in the characteristics
+of transformation processes. The first dimension is the pivot point of
+leadership in change. This dimension has two extremes: top-down leadership where
+the change is mandated by management, and bottom-up leadership where the chage
+emerges from team level. The second dimension is the velocity of the change.
+Organizations were reported to perform transformations as a big bang or
+overnight fashion, or as a step wise process lasting for a long time. Figure
+\ref{fig:transformation_dimensions} summarizes the transformation process
+dimensions.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics{graphics/transformation_dimensions.pdf}
+    \caption{Dimensions for categorizing transformation types}
+    \label{fig:transformation_dimensions}
+  \end{center}
+\end{figure}
+
+We do not quantitively measure cases on the transformation process dimensions,
+because the criteria for creating a measurement would be vague. For instance, we
+do not think that presenting a single quantitative measurement of the weight
+point of leadership on an axis between management centric and team centric would
+describe reality accurately. The dimensions of pivot of leadership and
+transformation velocity nevertheless exist.
+
+As the primary sources describe, transformation cases combined the dimensions of
+leadership pivot and velocity. Four combinations of the characteristics could be
+observed. Figure \ref{fig:transformation_types} summarizes the combinations of
+transformation characteristics.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/transformation_process.pdf}
+    \caption{Categorization of transformation types}
+    \label{fig:transformation_types}
+  \end{center}
+\end{figure}
+
+The most extreme case of adoption was a bing bang transformation. To realize
+such a large change it is necessary that the change was centrally coordinated.
+For step wise transformations there were three variants depending on the amount
+of involvement of management and development teams. In some cases teams palyed a
+passive role, and in some cases management was described to be reluctant to
+change. In many cases both management and devlopment dealed with the
+transformation on their respective level of influence.
+
+As a concept separate from the dimensions of leadership pivot and velocity, the
+transformation was described as a continuous process. There were a few
+viewpoints illustrating the concept of continuous transformation. Several
+organizations strived to introduce a culture of continuous improvement, which
+would make organizational adjustment a permanent state. In many cases the agile
+transformation was not considered to have ended at any distinct point in time,
+and the impression that it was still ongoing was given. Finally, transformation
+was simply described as hard work that will inevitably take a long time.
+
+% Mentioning big bang versus step wise [P16].
 
@@
 
-\begin{table}
+\begin{table}[h]
     \centering
@@
                      >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
         Change leaders         &  Described in  \\
@@
                      >{\raggedright\arraybackslash}p{0.61\textwidth} }
+        \toprule
         Activity           &  Described in  \\
@@
 was used in the beginning of the transformation. Primary sources discussing
-the use of piloting are listed in \ref{table:transformation_piloting}.
-
-\begin{table}
+the use of piloting are listed in table \ref{table:transformation_piloting}.
+
+\begin{table}[h]
     \centering
@@
                      >{\raggedright\arraybackslash}p{0.61\textwidth} }
+        \toprule
         Sources discussing viewpoints on piloting   & 
@@
 tangible investments. These investments were focused on better enabling agile
-software development, and highlight the concerns that organizations had when
+software development. They also highlight the concerns that organizations had when
 implementing an agile model. Typical investments were bringing in external help,
@@
                      >{\raggedright\arraybackslash}p{0.57\textwidth} }
-        Investment for transformation      &  Described in  \\
+        \toprule
+        Investments during transformation      &  Described in  \\
         \midrule
@@
                 P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
-        Agile project management tools  &
+        New project management tools for agile  &
                 P10, P11, P14, P16, P17, P26, P33,
@@
     \end{tabular}
-    \caption{Summary of investments aiming to make agile methods perform better}
+    \caption{Summary of investments done during transformation, aiming to make
+             agile methods perform better}
     \label{table:transformation_investments}

### 2014-06-02 journal/article.tex
Spell check

@@
 leadership in change. This dimension has two extremes: top-down leadership where
-the change is mandated by management, and bottom-up leadership where the chage
+the change is mandated by management, and bottom-up leadership where the change
 emerges from team level. The second dimension is the velocity of the change.
@@
 
-We do not quantitively measure cases on the transformation process dimensions,
+We do not quantitatively measure cases on the transformation process dimensions,
 because the criteria for creating a measurement would be vague. For instance, we
@@
 
-The most extreme case of adoption was a bing bang transformation. To realize
+The most extreme case of adoption was a big bang transformation. To realize
 such a large change it is necessary that the change was centrally coordinated.
 For step wise transformations there were three variants depending on the amount
-of involvement of management and development teams. In some cases teams palyed a
+of involvement of management and development teams. In some cases teams played a
 passive role, and in some cases management was described to be reluctant to
-change. In many cases both management and devlopment dealed with the
+change. In many cases both management and development dealt with the
 transformation on their respective level of influence.
@@
 viewpoints illustrating the concept of continuous transformation. Several
-organizations strived to introduce a culture of continuous improvement, which
+organizations were striving to introduce a culture of continuous improvement, which
 would make organizational adjustment a permanent state. In many cases the agile
@@
 Some organizations considered it important to pilot agile with business critical
-projects, as the applicability had to be proven in a releant environment [P8,
+projects, as the applicability had to be proven in a relevant environment [P8,
 P40].
@@
 Fry and Greene [P16] describe a particular big bang adoption where piloting was
-omitted. A pilot project was avioded because having two models side by side was
-deemed to create organizational dissonance. Instead a significant amout of
+omitted. A pilot project was avoided because having two models side by side was
+deemed to create organizational dissonance. Instead a significant amount of
 people completed agile training and the new model was introduced overnight
@@
 methods, and provide initial education [P7, P9, P27, P30, P43].
-It was recommended to bring in extenal help as early as possible, when the
+It was recommended to bring in external help as early as possible, when the
 internal agile expertise is lacking and before transformation challenges appear
@@
 The need for external consultants reduced over time, and their tasks were
-transfered to internal positions [P29, P47].
+transferred to internal positions [P29, P47].
 It was described that as the agile journey progressed, external consultants were
@@
 * It is difficult to distinguish between team and organization. In most cases
-  when talking about teams it refers to a handfull of people familiar to each
+  when talking about teams it refers to a handful of people familiar to each
   other. However, in some cases it was evident that an entire department was

### 2014-06-03 journal/article.tex
List of improvements
TODOs

@@
 \end{figure}
+
+TODO:
+* Duplicates removed?
+* Ei voi ymmärtää miten preliminary search filtteröitiin
 
@@
 
+TODO:
+* Three variants?
+* Listaa paremmin
+
 As a concept separate from the dimensions of leadership pivot and velocity, the
@@
 \end{table}
+
+TODO:
+* Other organizational levels?
+* On development team level
 
@@
 
+TODO:
+* Mikä erityisyys on inhouse coachingissa?
+* Organizaation sijotus sisäisiin coach-positionihin?
+
 % Training
@@
 
+TODO:
+* Continuous builds and testing --> setting up the environment?
+* Hassusti otsikoitu?
+
 % External consultants
@@
 
+TODO:
+* Otsikko ei hyvä
+* --> Large organization issues
+
 In addition to inter-team coordination there were also other challenges that
@@
 \subsubsection{Integrating requirements management and quality assurance}
+
+TODO:
+* Muuta ylemmäksi järjestyksessä
+* Mieti rakenne
+* Mieti onko ``other than development'' hyvä otsikko
 
@@
 
+* Vertailu kirjallisuuteen
+
 * Some success factors and challenges relate

### 2014-06-03 journal/article.tex
Fixing todos

--- 2014-06-02 journal/article.tex
@@
 
-The most extreme case of adoption was a big bang transformation. To realize
-such a large change it is necessary that the change was centrally coordinated.
-For step wise transformations there were three variants depending on the amount
-of involvement of management and development teams. In some cases teams played a
-passive role, and in some cases management was described to be reluctant to
-change. In many cases both management and development dealt with the
-transformation on their respective level of influence.
+The most extreme case of adoption was a big bang transformation. To realize such
+a large change it is necessary that the change was centrally coordinated.
+For step wise transformations the the amount of management and development team
+involvement varied. At one extreme, the change was lead by management while
+teams played a passive role. At the other end management was described to be
+reluctant to change while teams pushed for change. Many cases took a middle
+course where both management and development dealt with the transformation on
+their respective level of influence.
 
@@
 viewpoints illustrating the concept of continuous transformation. Several
-organizations were striving to introduce a culture of continuous improvement, which
-would make organizational adjustment a permanent state. In many cases the agile
-transformation was not considered to have ended at any distinct point in time,
-and the impression that it was still ongoing was given. Finally, transformation
-was simply described as hard work that will inevitably take a long time.
+organizations were striving to introduce a culture of continuous improvement,
+which would make organizational adjustment a permanent state. In many cases the
+agile transformation was not considered to have ended at any distinct point in
+time, and the impression that it was still ongoing was given. Finally,
+transformation was simply described as hard work that will inevitably take a
+long time.
 
@@
                 P43, P45, P48, P49, P50  \\
-        Change leaders on other organizational levels  &
+        Change leaders on development team levels  &
                 P2, P6, P9, P14, P17, P18, P27  \\
@@
 
-The tasks of the change team included first finding an alternative delivery
-method to solve problems in the old model [P21, P30]. During the transformation
-the team stimulated agile adoption in the organization [P18, P45, P49]. Examples
-on their tasks were informing people on the agile model [P4], educating and
-coaching [P16, P45, P48, P49], creating an agile cookbook [P23], and organizing
-presentations and workshops [P25, P41]. The change team also assessed the
-success of transformed teams, and worked for repeating their success across the
-organization [P16, P40, P41]. Still one characteristic was that the team was
-empowered to act on impediments they discovered [P16, P50].
+The typical first task of the change team was to decide on an alternative
+delivery method to solve problems in the old model [P21, P30]. During the
+transformation the team stimulated agile adoption in the organization [P18, P45,
+P49]. Other examples on thechange team's tasks were informing people on the
+agile model [P4], educating and coaching [P16, P45, P48, P49], creating an agile
+cookbook [P23], and organizing presentations and workshops [P25, P41]. The
+change team also assessed the success of transformed teams, and worked for
+repeating their success across the organization [P16, P40, P41]. The change team
+was empowered to act on removing impediments they discovered [P16, P50].
 
@@
 
-Leadership for change emerged also on other organizational levels. In many cases
-these change leaders were still in management positions [P14, P18, P17]. There
-were also agile champions in teams who drove the change on a grassroots level
-[P6, P9, P27].
+Beyond executive positions, leadership for change emerged also on lower
+organizational levels. In many cases these change leaders were still in
+management positions [P14, P18, P17]. There were also agile champions in teams
+who drove the change on a grassroots level [P6, P9, P27].
 As with higher-ups, these people worked for making agile methods familiar [P2,
@@
                 P31-P33, P36, P39-P41, P45-P47, P49, P50, P52  \\  % 25
-        In-house coaching   &
-                P1-P3, P6, P11, P16, P23, P29, P31, P32,
-                P39, P41, P45, P47, P49, P50, P52  \\    % 17
+%         In-house coaching   &
+%                 P1-P3, P6, P11, P16, P23, P29, P31, P32,
+%                 P39, P41, P45, P47, P49, P50, P52  \\    % 17
         Seminars and workshops   &
@@
 
+% Internal coaching
+The coaching positions were most often held by internal staff.
+Internal coaches were required for making the change grow, and to scale up the
+agile implementation [P47, P49]. Internal coaches had also the benefit of
+understanding of the company's processes and systems [P23, P39], and teams were
+able to access them easily [P3, P41]. Some cases presented building of a coach
+community as an integral part of the agile adoption [P6, P23, P32, P47].
+
 % How coaching evolves
@@
 transformation progressed [P32].
-
-% Internal coaching
-The coaching positions were most often held by internal staff.
-Internal coaches were required for making the change grow, and to scale up the
-agile implementation [P47, P49]. Internal coaches had also the benefit of
-understanding of the company's processes and systems [P23, P39], and teams were
-able to access them easily [P3, P41]. Some cases presented building of a coach
-community as an integral part of the agile adoption [P6, P23, P32, P47].
 
@@
                 P41, P43, P44, P45, P48, P50 \\
-        Continuous builds and testing  &
+        Creating a continuous build and integration system  &
                 P5, P11, P16, P17, P25, P27, P34, P39, P42, P50 \\
@@
 
-When a general problem of frequent broken builds existed it was easy to acquire a
-buy in for creating an organization-wide standard for automating builds and
-tests [P11]. On the other hand, many cases emphasized that building the
-continuous integration environment required a significant investment [P17, P34,
-P42]. In order to have the focus and resources for creating a proper integration
-environment a dedicated team was Some organizations appointed a dedicated team
-to ensure the focus and resources for creating a proper integration environment
-were available [P5, P16, P17].
+When a general problem of frequent broken builds existed it was easy to acquire
+a buy in for creating an organization-wide standard for automating builds and
+tests [P11]. On the other hand, many cases emphasized that building a continuous
+integration environment required a significant investment [P17, P34, P42].
+In order to have the focus and resources for creating a proper integration
+environment a dedicated team was created. Having an integration team ensured
+that focus and resources were available for creating a proper integration
+environment [P5, P16, P17].
 
@@
 
-\subsubsection{Other issues in large organizations}
+\subsubsection{Issues specific to large organizations}
 
@@
 should only emerge in a sizable organization.
-Additional layers of management are usually necessary to steer large groups of
-people, but agile methods provide little guidance on how to organize management.
-Problems may arise in a large environment if the schedule of agile adoption
+Most importantly, additional layers of management are usually necessary to steer
+large groups of people. The problem is that agile methods provide little
+guidance on how to organize management.
+Problems may also arise in a large environment if the schedule of agile adoption
 varies from team to team. Internal fragmentation is a further problem that may
@@
 
-\subsubsection{Integrating other functions than development to agile}
-
-Agile methods are typically considered as a set of activities closely related to
-software development. Cases describe how development teams were the first
-entities that were considered when introducing agile [P2, P17, P31, P38].
-However, development operates in a larger context and will need to interface
-with other organizational functions.
-Challenges were faced when exposing other parts of the organization to the agile
-model, and functions that were further away from development were resistant to
-change [P2, P18, P31]. A tension was growing between development and broader
-roles in the organization [P1]. The full benefits of the transformation would
-not be attained unless the entire organization was set to work along the same
-paradigm [P18, P42].
-
-Primary sources listed various functions beyond development that are involved in
-bringing products to the market. Functions interfacing with development and
-affected by the agile transformation are presented in figure
-\ref{fig:challenges_other_functions}.
-Marketing was a particularly frequently mentioned group [P5, P17, P28, P31, P38,
-P45, P48]. Other groups involved in the whole of product delivery were sales
-[P17, P31], infrastructure and operations [P10, P45, P46, P50], user experience
-and design [P9, P10, P28], documentation [P28], legal [P10], security [P10],
-customer services [P48], and finance [P48].
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{graphics/challenges_other_functions.pdf}
-    \caption{Functions affected by the agile transformation}
-    \label{fig:challenges_other_functions}
-  \end{center}
-\end{figure}
-
-Agile affects timings in the software delivery life cycle, what caused two types
-of problems. Firstly, the iterativeness of agile models was problematic when
-interfacing with other functions. Iterativeness changed the pace of delivery
-[P17], and forced design to focus on a shorter scope [P15]. Various groups
-feeding and supporting development had earlier committed once to a long term
-plan, but in the new arrangement they needed to adjust to incremental delivery
-[P48].
-Especially user experience (UX) and interaction designers did not welcome an
-incremental model [P6], and were struggling with maintaining the big picture in
-design while working in iterations [P9, P10].
-The infrastructure and operations teams were falling behind due to the increased
-delivery speed of development teams, and they were forced to update their way of
-working [P46, P50].
-
-Federoff and Courage [P15] elaborate on the user experience team's problems
-with the short time horizons of Scrum. Previously the UX team had supplied
-development with designs within a long release cycle, but Scrum required
-completing features within weeks. The problem was that that the UX work required
-a holistic view instead of an incremental one, and the design process did not
-fit well into the time frame of a sprint. The UX team was overloaded and
-discontent, but the situation was fixed by refining the schedule for interactions
-with development teams. [P15, P28]
-
-The second type of problems was that some release activities would inevitably
-have a long lead time but required committing to a set of functionalities in
-good time. The agile model's emphasis on short time horizon and flexibility in
-prioritization did not mix well with such activities.
-For instance, it would take time to prepare printed marketing material and press
-releases, but they needed to accurately present the upcoming features of the
-product [P38].
-
-Maples [P31] describes that marketing needed three months for preparing the
-product launch, but the agile process allowed development to keep changing the
-content of the product. As a result marketing was struggling with creating
-material and campaigns without knowing the exact product features. Further, the
-processes of acquiring export clearances and licensing authority required that
-the the feature set of the product was known well on beforehand.
-As a result of these requests development felt that their process was being
-slowed down. The issues were solved by agreeing on that marketing could at any
-time request a release in 90 days, and development would be able to commit to a
-fixed feature set to be delivered [P31].
-
-% Problems with HR
-
-Problems in adopting agile were even showing in functions outside the software
-delivery life cycle. In several cases human resources (HR) was a support function
-mentioned to work against the agile adoption. In order to gain full results of
-the transformation the entire organization should to be aligned, including HR
-[P38, P42, P49].
-Especially the rewarding practices set by HR were seen as problematic for the
-agile model. Rewarding was tied to personal performance, which was acting
-against team centric thinking and the agile model in general [P3, P6, P49].
-One case reported even on a practice of tracking failures down to individuals
-[P40].
-
 \subsubsection{Integrating requirements management and quality assurance}
+
+TODO:
+* Mieti rakenne
+* Mieti onko seuraava ``other than development'' hyvä otsikko
 
@@
 there was no extended period for stabilization [P5].
+
+\subsubsection{Integrating other functions than development to agile}
+
+Agile methods are typically considered as a set of activities closely related to
+software development. Cases describe how development teams were the first
+entities that were considered when introducing agile [P2, P17, P31, P38].
+However, development operates in a larger context and will need to interface
+with other organizational functions.
+Challenges were faced when exposing other parts of the organization to the agile
+model, and functions that were further away from development were resistant to
+change [P2, P18, P31]. A tension was growing between development and broader
+roles in the organization [P1]. The full benefits of the transformation would
+not be attained unless the entire organization was set to work along the same
+paradigm [P18, P42].
+
+Primary sources listed various functions beyond development that are involved in
+bringing products to the market. Functions interfacing with development and
+affected by the agile transformation are presented in figure
+\ref{fig:challenges_other_functions}.
+Marketing was a particularly frequently mentioned group [P5, P17, P28, P31, P38,
+P45, P48]. Other groups involved in the whole of product delivery were sales
+[P17, P31], infrastructure and operations [P10, P45, P46, P50], user experience
+and design [P9, P10, P28], documentation [P28], legal [P10], security [P10],
+customer services [P48], and finance [P48].
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{graphics/challenges_other_functions.pdf}
+    \caption{Functions affected by the agile transformation}
+    \label{fig:challenges_other_functions}
+  \end{center}
+\end{figure}
+
+Agile affects timings in the software delivery life cycle, what caused two types
+of problems. Firstly, the iterativeness of agile models was problematic when
+interfacing with other functions. Iterativeness changed the pace of delivery
+[P17], and forced design to focus on a shorter scope [P15]. Various groups
+feeding and supporting development had earlier committed once to a long term
+plan, but in the new arrangement they needed to adjust to incremental delivery
+[P48].
+Especially user experience (UX) and interaction designers did not welcome an
+incremental model [P6], and were struggling with maintaining the big picture in
+design while working in iterations [P9, P10].
+The infrastructure and operations teams were falling behind due to the increased
+delivery speed of development teams, and they were forced to update their way of
+working [P46, P50].
+
+Federoff and Courage [P15] elaborate on the user experience team's problems
+with the short time horizons of Scrum. Previously the UX team had supplied
+development with designs within a long release cycle, but Scrum required
+completing features within weeks. The problem was that that the UX work required
+a holistic view instead of an incremental one, and the design process did not
+fit well into the time frame of a sprint. The UX team was overloaded and
+discontent, but the situation was fixed by refining the schedule for interactions
+with development teams. [P15, P28]
+
+The second type of problems was that some release activities would inevitably
+have a long lead time but required committing to a set of functionalities in
+good time. The agile model's emphasis on short time horizon and flexibility in
+prioritization did not mix well with such activities.
+For instance, it would take time to prepare printed marketing material and press
+releases, but they needed to accurately present the upcoming features of the
+product [P38].
+
+Maples [P31] describes that marketing needed three months for preparing the
+product launch, but the agile process allowed development to keep changing the
+content of the product. As a result marketing was struggling with creating
+material and campaigns without knowing the exact product features. Further, the
+processes of acquiring export clearances and licensing authority required that
+the the feature set of the product was known well on beforehand.
+As a result of these requests development felt that their process was being
+slowed down. The issues were solved by agreeing on that marketing could at any
+time request a release in 90 days, and development would be able to commit to a
+fixed feature set to be delivered [P31].
+
+% Problems with HR
+
+Problems in adopting agile were even showing in functions outside the software
+delivery life cycle. In several cases human resources (HR) was a support function
+mentioned to work against the agile adoption. In order to gain full results of
+the transformation the entire organization should to be aligned, including HR
+[P38, P42, P49].
+Especially the rewarding practices set by HR were seen as problematic for the
+agile model. Rewarding was tied to personal performance, which was acting
+against team centric thinking and the agile model in general [P3, P6, P49].
+One case reported even on a practice of tracking failures down to individuals
+[P40].
 

### 2014-06-04 journal/article.tex
Editing: Success factors

@@
 Some organizations faced difficulties when changes to plans were needed. For
-instance, the organizational structure made adapting to changes difficult [52],
+instance, the organizational structure made adapting to changes difficult [P52],
 and making changes created management overhead [P51]. Other organizations saw an
@@
 transformation the team stimulated agile adoption in the organization [P18, P45,
-P49]. Other examples on thechange team's tasks were informing people on the
+P49]. Other examples on the change team's tasks were informing people on the
 agile model [P4], educating and coaching [P16, P45, P48, P49], creating an agile
@@
 Coaching was considered necessary, especially in the beginning of agile projects
-[P6, P8, P33, P46, 47]. Brown [P8] stated that without adequate coaching there
+[P6, P8, P33, P46, P47]. Brown [P8] stated that without adequate coaching there
 would have been severe long term consequences [P8]. The application of coaching
@@
 
-% Having too little coaches [23].
+% Having too little coaches [P23].
 % Coaches were also involved in training [P27].
@@
 by quality assurance (QA) activities. Agile methods suggest that essential parts
-of these activities would be performed within the development team. However,
-when developing large systems the effort required for managing requirements and
-QA may be too heavy for individual teams to handle. Table
-\ref{table:challenges_requirementsandqa} presents challenges in implementing
-additional requirements management and quality assurance required in large
-development projects.
+of these activities would be performed within the development team. However, for
+large systems the effort required for managing requirements and QA may be too
+heavy for individual teams to handle. For instance, a single development team
+can manage and elaborate items in its own backlog, but a large feature may
+require work across many teams. The feature needs to be divided to team specific
+items on a higher level, and the process must be managed somehow.
+Table \ref{table:challenges_requirementsandqa} presents challenges in
+implementing additional requirements management and quality assurance required
+in large development projects.
 
@@
 methods. In several cases there was mentionings about change leaders or teams
-becoming agile zealots. For instance, Evans [12] describes how some members of
+becoming agile zealots. For instance, Evans [P12] describes how some members of
 the agile community were becoming too evangelic, which made people take sides
@@
 
-\begin{table}[h]
+Agile methods come as new ways of working to many in the organization. In order
+to get a good start in the transformation the ideas of the new methods need to
+be explained thoroughly. Training sessions are good starting points, but teams
+should be accompanied by coaching for a longer time to ensure that the new
+practices are learned and applied correctly in practice. Table
+\ref{table:success_training} lists references that highlight training and
+coaching as success factors.
+
+\begin{table}
     \centering
@@
         \midrule
-        Training is necessary to achieve change  &
+        Providing training on agile methods  &
                 P3, P6, P11, P14, P16, P30, P37, P46 \\
-        Coaching is essential in transformation  &
-                P3, P6, P8, P9 P16, P17, P18, P29, P30, P33, P42, P44, P47, P48  \\
+        Coaching teams as they learn by doing  &
+                P3, P6, P8, P9 P16, P17, P18, P29,
+                P30, P33, P42, P44, P47, P48  \\
         \bottomrule
     \end{tabular}
-    \caption{XXXXXXXXXXXXXX}
-    \label{table:success_NN}
+    \caption{Training and coaching presented as success factors}
+    \label{table:success_training}
 \end{table}
@@
 Several studies stated that training improved the chance of succeeding in the
-transformation [P3, 16, P30]. It was even highlighted that the change would have
+transformation [P3, P16, P30]. It was even highlighted that the change would have
 failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
@@
 
-\subsubsection{Choosing the agile model}
-
-The choice of agile practices did affect the success of the transformation, and
-also how the practices were chosen affected.
+\subsubsection{Use of piloting}
+
+The agile model was usually tried out in the organization before rolling it out
+on a general level. Beyond being a common practice, the use of pilot programs
+was also attributed as a success factor. Table \ref{table:success_pilots} lists
+sources discussing piloting as a success factor.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Have a pilot project at start  &
+                P3, P9, P14, P17, P31, P36, P38, P39, P47  \\
+        Gather insights from pilot projects  &
+                P8, P22, P27, P40, P45  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Piloting the agile model as a success factor}
+    \label{table:success_pilots}
+\end{table}
+
+% Have a successful pilot first
+
+Having a pilot project was reported as a significant success factor in several
+cases [P3, P9, P38].
+The pilot project gave proof that the agile model would be suitable for the
+organization and the general acceptance of agile was increased [P3, P17].
+The pilot also cleared disbeliefs of the suitability of agile for large
+organizations, and created acceptance for both development and management [P9,
+P31, P36].
+
+Pilots were especially important for gaining management acceptance. Cases
+reported that management gave approval for agile methods only after seeing
+successful pilots [P14, P47]. Prokhorenko [P39] also describes that seeing the
+example of successful pilots will make managers from other departments eager to
+try the new method, and thus help to spread the use of it.
+
+% Pilot provides information on how to do the rest of the transformation
+
+A particular benefit in piloting was that it provided knowledge on how to fit
+the agile method to the particular organization.
+A pilot project enabled the organization to get feedback on how teams and
+managers are best introduced to the new methods [P8, P22].
+Organizations met challenges in the pilot projects. However, the pilot projects
+served as valuable learning experiments providing insight in how to mitigate
+problems when the rest of the organization is transforming [P27, P40, P45].
+
+\subsubsection{Choosing the agile model to fit the organization}
+
+The choice of agile practices did affect the success of the transformation. It
+was common that cases described that it was beneficial to customize agile
+practices to fit the organization specific environment. As agile models do not
+prescribe practices strictly there was opportunities to make choices in the
+implementation. A few recommendations were given for choosing the agile model,
+such as using the same practices for all teams, mapping to the old model, and
+keeping practices simple. Table \ref{table:success_choosemodel} lists viewpoints
+relating to the choice of the agile model which were reported to provide
+benefits.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Customizing the agile model  &
+                P8, P10, P11, P29, P31, P40, P41, P43, P45, P49, P51  \\
+        Conformity to a single agile model  &
+                P8, P11, P12, P16, P32, P43  \\
+        Mapping to the old model eases adaptation  &
+                P14, P29, P44, P46, P48  \\
+        Keeping the new model simple  &
+                P5, P10, P16, P40  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Recommendations for choosing the agile model}
+    \label{table:success_choosemodel}
+\end{table}
+
+% Customizing the agile model
+
+Customizing the agile model and practices was often seen as an necessary step in
+the agile implementation. As each organization will have its own challenges in
+adopting agile, it should be carefully considered which organization specific
+areas to focus on when choosing the agile practices to implement [P8, P43].
+Lewis and Nehrer [P29] recommended to choose the agile model according to the
+current corporate model in order to avoid interference.
+
+Cases reported that customization was part of a successful agile implementation.
+For instance, teams were let to customize their practices individually, to fit
+the needs of each team [P41].
+Spayd [P49] writes that teams who modified agile practices to suite the large and
+distributed environment ended up doing better than teams that did not.
+% Some intermediate point?
+To allow teams to become innovative and perform well the agile model should be
+customized in a pragmatical way instead of following a strict textbook
+interpretation [P10]
+In order to draw the most out of agile teams should innovate and find on
+practices that really work for their case [P40, P41]
+
+Especially in a large organization it is not feasible to use the same process
+for all projects [P49]. An individual application of the agile process is needed
+for different types of development, depending on for instance the type of
+software being developed or the project size [P45].
+Applying agile at scale will require to deviate from some of the typical
+recommendations [P31]. However, it is essential to remember the agile principles
+when doing customizations, and watch out for customization that contradict with
+the principles [P31].
+
+The customization of the agile model was also seen as an evolutionary process.
+Cowan [P11] suggested that in a big-bang transformation it might be necessary to
+selectively compromise some parts of the agile implementation so that the core
+practices can be set in place.
+The agile transformation is a constantly ongoing process, and it is recommended
+to regularly challenge teams to refine the agile implementation to reflect the
+current needs of the organization [P41, P43].
+
+% Conformity in the agile model
+
+Some studies reported that conformity was considered in implementing the new
+model. A consistence common vocabulary was seen to benefit the organization and
+support the change [P8, P16, P43]. Other benefits in using a single model were
+possibility to compare work between teams, it was easier for people to relocate,
+and progress was predictable for stakeholders [P43]. Consistency and common
+understanding was created in discipline meetings and public events, or by peer
+coaching [P11, P12, P32].
 
@@
 
-% Conformity in the agile model
-
-Some studies reported that conformity was considered in implementing the new
-model. A consistence common vocabulary was seen to benefit the organization and
-support the change [P8, P16, P43]. Other benefits in using a single model were
-possibility to compare work between teams, it was easier for people to relocate,
-and progress was predictable for stakeholders [P43]. Consistency and common
-understanding was created in discipline meetings and public events, or by peer
-coaching [P11, P12, P32].
-
 % Keep things simple
@@
 
-% Emphasize principles over mechanics
-
-In a number of studies it was described that during transformation the
-principles of agile should be emphasized over practices and simple mechanics
-[P46, P48]. When people understand the agile values they will also understand
-why the change is being done and feel motivated [P1, P16, P42].
-The people centric nature of agile development should be communicated correctly.
-McDowell and Dourambeis [P32] describe how the organization launched a series of
-communications events especially designed to emphasize the reasons behind the
-agile practices.
-
-% Customizing the agile model
-
-Customizing the agile model and practices was reported as an necessary step in
-the agile implementation. As each organization will have its own challenges in
-adopting agile, it should be carefully considered which organization specific
-areas to focus on when choosing the agile practices to implement  [P8, P43].
-Lewis and Nehrer [P29] recommended to choose the agile model according to the
-current corporate model in order to avoid interference.
-
-Especially in a large organization it is not feasible to use the same process
-for all projects [P49]. An individual application of the agile process is needed
-for different types of development, depending on for instance the type of
-software being developed or the project size [P45].
-Applying agile at scale will require to deviate from some of the typical
-recommendations [P31]. However, it is essential to remember the agile principles
-when doing customizations, and watch out for customization that contradict with
-the principles [P31].
-
-Cases reported that customization was part of a successful agile implementation.
-For instance, teams were let to customize their practices individually, to fit
-the needs of each team [P41].
-Spayd [P49] writes that teams who modified agile practices to suite the large and
-distributed environment ended up doing better than teams that did not.
-% Some intermediate point?
-To allow teams to become innovative and perform well the agile model should be
-customized in a pragmatical way instead of following a strict textbook
-interpretation [P10]
-
-The customization of the agile model was also seen as an evolutionary process.
-Cowan [11] suggested that in a big-bang transformation it might be necessary to
-selectively compromise some parts of the agile implementation so that the core
-practices can be set in place.
-The agile transformation is a constantly ongoing process, and it is recommended
-to regularly challenge teams to refine the agile implementation to reflect the
-current needs of the organization [P41, P43].
-
-% Customize the agile model?
-% ??
-% [P16]
-% This cross-functional team redesigned and rebuilt the development process from
-% the ground up using key values from the company’s founding: KISS (Keep it Simple
-% Stupid), iterate quickly, and listen to our customers. These values are a
-% natural match for agile methodologies.
-% 
-% ??? There are changes done to the agile proces, but doing it is not that much emphasized
-% [P51]
-% One after the other, all products where then switched to the new Scrum process.
-% Along the way, minor adjustments were often made. The teams were constantly
-% searching for the right balance between sprint duration, size estimation,
-% requirement size, and time spent on backlog administration. Changes to the
-% process were initially often managed by the the CTO. However, they were
-% increasingly delegated to the appropriate ScrumMasters.
-% Comment: This is not directly seen as a success factor.
-
-% ??
-In order to draw the most out of agile teams should innovate and find on
-practices that really work for their case [P40, P41]
-
-% ??
-Especially inexperienced coaches and Scrum masters made mistakes in focusing too
-much on the implementation of practices [P23, P30].
-
-\subsubsection{Considering other parts than development}
+\subsubsection{Considering the Product Owner role and requirements management}
+
+A particularly difficult part to implement efficiently was requirements
+management work. The key role in handling requirements is the Product Owner.
+Many cases described that careful attention to the requirements management roles
+was necessary for a successful transformation.
+Table \ref{table:success_requirements} lists references that discuss careful
+implementation of the product owner role and requirements management as success
+factors.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Pay attention to the Product Owner role  &
+                P12, P14, P16, P24, P30, P33, P39, P46  \\
+        Pay attention to requirements management  &
+                P5, P11, P17, P33, P48  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Careful implementation of requirements management seen as a
+             success factor}
+    \label{table:success_requirements}
+\end{table}
+
+% Emphasize role of Product Owners
+
+A particular role that seemed to make much difference in the agile introduction
+was the Product Owner. Many cases reported on problems if the Product Owner did
+not perform adequately, and it was seen as critical to have a dedicated person
+in that role [P12, P14, P30, P39].
+Success or problems emerged respectively depending on how well the Product
+Owners were performing. In one case a well implemented Product Owner role was
+understood to be a key success factor when using agile methods [P16, P33].
+It was reported that when the Product Owner role was implemented correctly the
+team performed better and the work products were correct [P24]. Projects started
+off better when the Product Owners were engaged early on [P46]. Some Product
+Owners were even described to become agile enthusiasts when they learned how
+business and technology can collaborate in the agile model [P46].
+
+Several cases endorsed training and coaching for the Product Owner role [P16,
+P39, P46]. Product Owners should receive training on agile principles, backlog
+management, user story breakdown, and agile planning [P16]. Also tool support
+for backlog management and two way communication between Product Owners and
+teams should be worked on [P39].
 
@@
 
-% Emphasize role of Product Owners
-
-A particular role that seemed to make much difference in the agile introduction
-was the Product Owner. Many cases reported on problems if the Product Owner did
-not perform adequately, and it was seen as critical to have a dedicated person
-in that role [P12, 14, P30, P39].
-Success or problems emerged respectively depending on how well the Product
-Owners were performing. In one case a well implemented Product Owner role was
-understood to be a key success factor when using agile methods [16, P33].
-It was reported that when the Product Owner role was implemented correctly the
-team performed better and the work products were correct [P24]. Projects started
-off better when the Product Owners were engaged early on, and some Product
-Owners even became agile enthusiasts when they learned how business and
-technology can collaborate in the agile model [P46].
-
-Several cases endorsed training and coaching for the Product Owner role [P16,
-P39, P46]. Product Owners should receive training on agile principles, backlog
-management, user story breakdown, and agile planning [P16]. Also tool support
-for backlog management and two way communication between Product Owners and
-teams should be worked on [P39].
+\subsubsection{Management support}
+
+The most prominent and clear success factor was having support from management.
+When the leadership of the organization is committed and supportive change is
+much more likely to happen. Table \ref{table:success_management} lists
+success factors related to management support.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Have management support  &
+                P8, P11, P16, P18, P27, P31, P33, P36,
+                P43, P44, P46, P47, P49  \\
+        Visible management support encourages employees  &
+                P9, P11, P24, P40  \\
+        Educate management on agile methods  &
+                P6, P8, P9, P11, P16, P29, P35, P46, P47  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Success factors relating to management support for agile}
+    \label{table:success_management}
+\end{table}
+
+% Management support in general
+
+Numerous cases made it clear that management support is an absolute necessity
+[P8, P18, P27, P31, P33, P36, P43, P47, P49].
+This was reflected in statements such as ``Adopting agile, or implementing any
+significant change, requires an executive’s sincere support.'' [P44],
+``Executive commitment was crucial to implementing massive change.'' [P16], and
+``Having upper management engaged, supportive and visible is critical for
+wholesale organization involvement with Scrum.'' [P11].
+
+Managers were seen to be in a key role in making the change stick, as they had
+the authority and power to remove impediments [P18]. Seffernick [P46] describes
+how a number of people attempted to explain away the applicability of agile
+methods, but the objections were overruled by the director's commitment to make
+agile work. Management support was similarly needed when tight release schedules
+had to be flexed in order to give room for the adoption process [P16, P27].
+Favorable management decisions were also critical when additional resources were
+allocated to training and coaching [P47].
+
+Visible involvement of management was also reported to motivate and encourage
+employees to adopt the new model [P40]. For instance, the CTO organized training
+sessions personally [P24], and the engineering VP frequently visited sprint
+demos [P11]. When the corporate level support for the agile initiative was
+showing teams adopted agile methods even spontaneously [P9].
+
+% Educate management
+
+In order to gain support for agile several cases highlighted the need to educate
+management [P8, P47].
+Without education managers might make harmful interventions in the process, or
+just stick to the old model [P16, P35, P46].
+Managers not understanding the principles of agile were feeling left out with
+the introduction of self organizing teams, which some times resulted in
+backlashes [P6]. Providing proper training was reported to correct the
+situation, and even create strong agile supporters in management [P6, P9].
+Training also cleared misconceptions that management had [P29], and helped
+create a consistent implementation of the agile model across the organization
+[P11].
+
+There were a few suggestions on how to involve management better. Foremost, a
+successful experience will likely rise the interest of management [P36].
+Mencke [P33] suggests giving executives tasks related to the agile process,
+so that they have better visibility of the new way of working.
 
@@
 
-\subsubsection{Create a positive start by piloting}
-
-% Have a successful pilot first
-
-The agile model was usually tried out in the organization before rolling it out
-on a general level. Having a pilot project was reported as a significant success
-factor in several cases [P3, P9, P38].
-The pilot project gave proof that the agile model would be suitable for the
-organization and the general acceptance of agile was increased [P3, 17].
-The pilot also cleared disbeliefs of the suitability of agile for large
-organizations, and created acceptance for both development and management [P9,
-P31, P36].
-
-Pilots were especially important for gaining management acceptance. Cases
-reported that management gave approval for agile methods only after seeing
-successful pilots [P14, P47]. Prokhorenko [P39] also describes that seeing the
-example of successful pilots will make managers from other departments eager to
-try the new method, and thus help to spread the use of it.
-
-% Pilot provides information on how to do the rest of the transformation
-
-A particular benefit in piloting was that it provided knowledge on how to fit
-the agile method to the particular organization.
-A pilot project enabled the organization to get feedback on how teams and
-managers are best introduced to the new methods [P8, P22].
-Organizations met challenges in the pilot projects. However, the pilot projects
-served as valuable learning experiments providing insight in how to mitigate
-problems when the rest of the organization is transforming [P27, P40, P45].
-
-% Having a positive experience
-
-It is clear that any positive results will facilitate an organizational
-transformation. Although a recipe to fabricate positive experiences can not be
-given, many cases described how they had succeeded. Benefield [P6] describes how
-a survey was done to objectively measure satisfaction in agile, and the positive
-results helped teams make the decision to change their way of working. Benefield
-describes further how management was requesting proof on better performance as a
-response to requests to increase coaching and training budget. After conducting
-inquiries the coaching team was able to present an estimate of a 30\% increase
-in performance, which made management highly convinced [P6]. Another case
-describes how the corporate agile awareness and teaching event was designed to
-be fun and engaging, which made people take more enthusiasm in applying the agile
-practices [P32]. The move towards agile is assisted by making any benefits
-publicly visible and celebrating even the small victories [P40, P42].
-
-The dynamics of positive results were explained in further ways. Several cases
-highlighted that the agile transformation was spread effectively through
-positive word-of-mouth [P6, P40]. When good results were shown by a team it
-created interest in others, and enthusiasm to try the new model will spread
-[P39, P46]. Some companies were using agile and waterfall methods side by side.
-This setting made it comparison possible, bringing up the benefits of the agile
-model [P19, P36].
-
-\subsubsection{Communication and transparency in the change}
-
-Intensive communication was emphasized in a number of studies. It is important
-to reach a maximum number of people throughout the organization, as without
-communication the a new model will not take root [P16, P43].
-It was recommended that use of the new model is made highly visible on many
-communication channels and even over-communicated [P16, P33, P41]. Mencke [P33]
-summarizes the viewpoint on over-communication: ``Even if you think that your
-teams understand a new method or process, repeat your communication to be
-sure.'' Workshops, coaching sessions, online discussions, and newsletters are
-examples on different communication formats used [P41]. Another approach in
-communicating the change was that managers explained and encouraged change in an
-extensive series of one-to-one discussions [P22].
-
-Enabling transparency during the transformation was another communication
-related theme that was brought up, and even highlighted as a critical factor for
-success [P11, P16]. Fry and Greene [P16] underline the importance of
-transparency and sharing of information: ``This bias to sharing information with
-everyone was critical in our ability to adapt on a daily basis to ensure our
-success.'' Transparency was achieved by showing both successes and challenges
-[P6], actively reaching out for feedback [P6], using project management tools to
-display project status publicly [P11], by rearranging physical spaces [P42], and
-by holding the meetings of the roll-out team in public [P16].
-By sharing experiences and status of the transformation the organization was
-aligned and everyone was moving in the same direction [P42, P43].
-
-% The inertia and fear of sharing information must be overcome [P33].
-
-% Sharing everything with everyone was key to success [P16].
-
-% Organizations made also an effort to communicate the status of the
-% transformation [P16, P43].
-
-% Because of the volatility of software development there must be continuous and
-% fluent communication between managers, POs, and teams [P42].
-
-Communicating the goals of the transformation was highlighted in a some studies.
-Having a clear message of expectations helped to remove confusion and make
-people understand the purpose of the transformation [P43, P48]. The motivation
-to change can be increased by having a simple proposal on how to reach the
-desired outcomes [P40].
-
-% Engage people broadly
-
-Many cases highlighted how people were engaged broadly in the organization.
-One of the lessons learned presented was that it is important to involve all
-stakeholders to gain acceptance in the transformation [P4, P31]. The transition
-team made an effort to engage people by both inviting to give feedback online
-and holding an extensive number of feedback meetings [P6, P22, P16]
-Being inclusive towards everyone was seen as one of the key success factors in
-the transformation [P16]. Inclusiveness will encourage people to participate and
-use the new model visibly [P16].
-
 \subsubsection{Have a change leading coalition}
@@
 
-\subsubsection{Management support}
-
-% Management support in general
-
-The most prominent and clear success factor was having support from management.
-Numerous cases made it clear that management support is an absolute necessity
-[P8, P18, P27, P31, P33, P36, P43, P47, P49].
-This was reflected in statements such as ``Adopting agile, or implementing any
-significant change, requires an executive’s sincere support.'' [P44],
-``Executive commitment was crucial to implementing massive change.'' [P16], and
-``Having upper management engaged, supportive and visible is critical for
-wholesale organization involvement with Scrum.'' [P11].
-
-Managers were seen to be in a key role in making the change stick, as they had
-the authority and power to remove impediments [P18]. Seffernick [P46] describes
-how a number of people attempted to explain away the applicability of agile
-methods, but the objections were overruled by the director's commitment to make
-agile work. Management support was similarly needed when tight release schedules
-had to be flexed in order to give room for the adoption process [P16, P27].
-Favorable management decisions were also critical when additional resources were
-allocated to training and coaching [P47].
-
-Visible involvement of management was also reported to motivate and encourage
-employees to adopt the new model [P40]. For instance, the CTO organized training
-sessions personally [P24], and the engineering VP frequently visited sprint
-demos [P11]. When the corporate level support for the agile initiative was
-showing teams adopted agile methods even spontaneously [P9].
-
-There were a few suggestions on how to involve management better. Foremost, a
-successful experience will likely rise the interest of management [P36].
-Mencke [P33] suggests giving executives tasks related to the agile process,
-so that they have better visibility of the new way of working.
-
-% Educate management
-
-In order to gain support for agile several cases highlighted the need to educate
-management [P8, P47].
-Without education managers might make harmful interventions in the process, or
-just stick to the old model [P16, P35, P46].
-Managers not understanding the principles of agile were feeling left out with
-the introduction of self organizing teams, which some times resulted in
-backlashes [P6]. Providing proper training was reported to correct the
-situation, and even create strong agile supporters in management [P6, P9].
-Training also cleared misconceptions that management had [P29], and helped
-create a consistent implementation of the agile model across the organization
-[P11].
+\subsubsection{Communication and transparency in the change}
+
+Intensive communication was emphasized in a number of studies. It is important
+to reach a maximum number of people throughout the organization, as without
+communication the a new model will not take root [P16, P43].
+It was recommended that use of the new model is made highly visible on many
+communication channels and even over-communicated [P16, P33, P41]. Mencke [P33]
+summarizes the viewpoint on over-communication: ``Even if you think that your
+teams understand a new method or process, repeat your communication to be
+sure.'' Workshops, coaching sessions, online discussions, and newsletters are
+examples on different communication formats used [P41]. Another approach in
+communicating the change was that managers explained and encouraged change in an
+extensive series of one-to-one discussions [P22].
+
+Enabling transparency during the transformation was another communication
+related theme that was brought up, and even highlighted as a critical factor for
+success [P11, P16]. Fry and Greene [P16] underline the importance of
+transparency and sharing of information: ``This bias to sharing information with
+everyone was critical in our ability to adapt on a daily basis to ensure our
+success.'' Transparency was achieved by showing both successes and challenges
+[P6], actively reaching out for feedback [P6], using project management tools to
+display project status publicly [P11], by rearranging physical spaces [P42], and
+by holding the meetings of the roll-out team in public [P16].
+By sharing experiences and status of the transformation the organization was
+aligned and everyone was moving in the same direction [P42, P43].
+
+% The inertia and fear of sharing information must be overcome [P33].
+
+% Sharing everything with everyone was key to success [P16].
+
+% Organizations made also an effort to communicate the status of the
+% transformation [P16, P43].
+
+% Because of the volatility of software development there must be continuous and
+% fluent communication between managers, POs, and teams [P42].
+
+Communicating the goals of the transformation was highlighted in a some studies.
+Having a clear message of expectations helped to remove confusion and make
+people understand the purpose of the transformation [P43, P48]. The motivation
+to change can be increased by having a simple proposal on how to reach the
+desired outcomes [P40].
+
+% Emphasize principles over mechanics
+
+%         Emphasizing principles over mechanics  &
+%                 P1, P16, P32, P42, P46, P48  \\
+
+In a number of studies it was described that during transformation the
+principles of agile should be emphasized over practices and simple mechanics
+[P46, P48]. When people understand the agile values they will also understand
+why the change is being done and feel motivated [P1, P16, P42].
+The people centric nature of agile development should be communicated correctly.
+McDowell and Dourambeis [P32] describe how the organization launched a series of
+communications events especially designed to emphasize the reasons behind the
+agile practices.
+
+% ??
+Especially inexperienced coaches and Scrum masters made mistakes in focusing too
+much on the implementation of practices [P23, P30].
+
+% Engage people broadly
+
+Many cases highlighted how people were engaged broadly in the organization.
+One of the lessons learned presented was that it is important to involve all
+stakeholders to gain acceptance in the transformation [P4, P31]. The transition
+team made an effort to engage people by both inviting to give feedback online
+and holding an extensive number of feedback meetings [P6, P22, P16]
+Being inclusive towards everyone was seen as one of the key success factors in
+the transformation [P16]. Inclusiveness will encourage people to participate and
+use the new model visibly [P16].
 
@@
 
+% Having a positive experience
+
+It is clear that any positive results will facilitate an organizational
+transformation. Although a recipe to fabricate positive experiences can not be
+given, many cases described how they had succeeded. Benefield [P6] describes how
+a survey was done to objectively measure satisfaction in agile, and the positive
+results helped teams make the decision to change their way of working. Benefield
+describes further how management was requesting proof on better performance as a
+response to requests to increase coaching and training budget. After conducting
+inquiries the coaching team was able to present an estimate of a 30\% increase
+in performance, which made management highly convinced [P6]. Another case
+describes how the corporate agile awareness and teaching event was designed to
+be fun and engaging, which made people take more enthusiasm in applying the agile
+practices [P32]. The move towards agile is assisted by making any benefits
+publicly visible and celebrating even the small victories [P40, P42].
+
+The dynamics of positive results were explained in further ways. Several cases
+highlighted that the agile transformation was spread effectively through
+positive word-of-mouth [P6, P40]. When good results were shown by a team it
+created interest in others, and enthusiasm to try the new model will spread
+[P39, P46]. Some companies were using agile and waterfall methods side by side.
+This setting made it comparison possible, bringing up the benefits of the agile
+model [P19, P36].
+
 \clearpage

### 2014-06-05 journal/article.tex
Editing: Success factors

@@
     \includegraphics{graphics/challenges_summary.pdf}
-    \caption{Overview of challenges in agile transformation}
+    \caption{Overview of challenges in agile transformations}
     \label{fig:challenges_summary}
@@
 
-* What is our standpoint on recommendations that experience reports give, but
-  which were not necessarily tried in the organization?
-
-* Some success factors are just calculated decisions that were made because
-  the organization saw them as the best way to proceed. The question whether
-  the decision really were critical for success is not proven.
-
-* Some success factors are compiled from ``what we learned'' or ``advice for
-  others'' lists. We considered whether the advice was something the
-  organization had tried in practice.
-
-* For instance, implementing continuous integration and automated builds was not
-  explicitly stated as a factor contributing to the success of the
-  transformation, although it is unquestionable that such systems are critical
-  for maintaining quality and enabling frequent releases.
-
-\subsubsection{Training and coaching}
-
-Agile methods come as new ways of working to many in the organization. In order
-to get a good start in the transformation the ideas of the new methods need to
-be explained thoroughly. Training sessions are good starting points, but teams
-should be accompanied by coaching for a longer time to ensure that the new
-practices are learned and applied correctly in practice. Table
-\ref{table:success_training} lists references that highlight training and
-coaching as success factors.
-
-\begin{table}
+\begin{figure}[!b]
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/success_summary.pdf}
+    \caption{Overview of success factors in agile transformations}
+    \label{fig:success_summary}
+  \end{center}
+\end{figure}
+
+Figure \ref{fig:success_summary} summarizes success factors evident in the
+primary studies. The success factors are organized into ten categories within
+three themes, which are elaborated in this section. The first theme relates to
+management and leadership in the transformation. The second theme presents
+success factors that relate to finding a suitable agile model for the
+organization. The last theme includes success factors that are realized on the
+level of development teams.
+
+The discovered success factors did not exhibit strong differences in
+significance in respect to each other. Three categories of success factors can
+be considered of higher importance: management support, training and coaching,
+and choosing the agile model carefully. The remaining success factors should be
+considered having even significance.
+
+% * What is our standpoint on recommendations that experience reports give, but
+%   which were not necessarily tried in the organization?
+% 
+% * Some success factors are just calculated decisions that were made because
+%   the organization saw them as the best way to proceed. The question whether
+%   the decision really were critical for success is not proven. 
+% 
+% * Some success factors are compiled from ``what we learned'' or ``advice for
+%   others'' lists. We considered whether the advice was something the
+%   organization had tried in practice.
+% 
+% * For instance, implementing continuous integration and automated builds was not
+%   explicitly stated as a factor contributing to the success of the
+%   transformation, although it is unquestionable that such systems are critical
+%   for maintaining quality and enabling frequent releases.
+
+\subsubsection{Management support}
+
+The most prominent and clear success factor was having support from management.
+When the leadership of the organization is committed and supportive change is
+much more likely to happen. Table \ref{table:success_management} lists
+success factors related to management support.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.41\textwidth}
+                     >{\raggedright\arraybackslash}p{0.51\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Have management support  &
+                P8, P11, P16, P18, P27, P31, P33, P36,
+                P43, P44, P46, P47, P49  \\
+        Visible management support encourages employees  &
+                P9, P11, P24, P40  \\
+        Educate management on agile methods  &
+                P6, P8, P9, P11, P16, P29, P35, P46, P47  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Success factors relating to management support for agile}
+    \label{table:success_management}
+\end{table}
+
+% Management support in general
+
+Numerous cases made it clear that management support is an absolute necessity
+[P8, P18, P27, P31, P33, P36, P43, P47, P49].
+This was reflected in statements such as ``Adopting agile, or implementing any
+significant change, requires an executive’s sincere support.'' [P44],
+``Executive commitment was crucial to implementing massive change.'' [P16], and
+``Having upper management engaged, supportive and visible is critical for
+wholesale organization involvement with Scrum.'' [P11].
+
+Managers were seen to be in a key role in making the change stick, as they had
+the authority and power to remove impediments [P18]. Seffernick [P46] describes
+how a number of people attempted to explain away the applicability of agile
+methods, but the objections were overruled by the director's commitment to make
+agile work. Management support was similarly needed when tight release schedules
+had to be flexed in order to give room for the adoption process [P16, P27].
+Favorable management decisions were also critical when additional resources were
+allocated to training and coaching [P47].
+
+Visible involvement of management was also reported to motivate and encourage
+employees to adopt the new model [P40]. For instance, the CTO organized training
+sessions personally [P24], and the engineering VP frequently visited sprint
+demos [P11]. When the corporate level support for the agile initiative was
+showing teams adopted agile methods even spontaneously [P9].
+
+% Educate management
+
+In order to gain support for agile several cases highlighted the need to educate
+management [P8, P47].
+Without education managers might make harmful interventions in the process, or
+just stick to the old model [P16, P35, P46].
+Managers not understanding the principles of agile were feeling left out with
+the introduction of self organizing teams, which some times resulted in
+backlashes [P6]. Providing proper training was reported to correct the
+situation, and even create strong agile supporters in management [P6, P9].
+Training also cleared misconceptions that management had [P29], and helped
+create a consistent implementation of the agile model across the organization
+[P11].
+
+There were a few suggestions on how to involve management better. Foremost, a
+successful experience will likely rise the interest of management [P36].
+Mencke [P33] suggests giving executives tasks related to the agile process,
+so that they have better visibility of the new way of working.
+
+\subsubsection{Commitment to change}
+
+To make a large organization change the commitment to change must be made clear.
+People should be made to understand that there is no return to the old model.
+If and when the changes uncover problems they must not be allowed to set back
+the transformation. Table \ref{table:success_commitment} lists sources which
+present commitment to change as a success factor.
+
+\begin{table}[h]
     \centering
@@
         \midrule
-        Providing training on agile methods  &
-                P3, P6, P11, P14, P16, P30, P37, P46 \\
-        Coaching teams as they learn by doing  &
-                P3, P6, P8, P9 P16, P17, P18, P29,
-                P30, P33, P42, P44, P47, P48  \\
+        Make it understood that the change is non-negotiable  &
+                P11, P22, P48, P49  \\
+        Commitment to change  &
+                P8, P10, P43  \\
         \bottomrule
     \end{tabular}
-    \caption{Training and coaching presented as success factors}
-    \label{table:success_training}
+    \caption{Commitment to change as a success factor}
+    \label{table:success_commitment}
 \end{table}
 
-Training was seen as a strong factor helping the organization to transform.
-Several studies stated that training improved the chance of succeeding in the
-transformation [P3, P16, P30]. It was even highlighted that the change would have
-failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
-and again how training could make the difference between success and failure for
-a team''. Training also helped people to become more positively inclined towards
-the new model [P16, P37], and at best training made people enthusiastic to change
-[P3, P46]. A few cases reported that there was some disinclination to change,
-and training would have helped to overcome [P11, P14].
-
-Agile methods avoid prescribing exact ways of working but rather emphasize a
-mindset of adapting to the situation. It is difficult to explain by theory how
-such a mindset should be applied, and the agile practices are best learned by
-doing. Coaching teams while they apply the agile methods in practice was seen as
-an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
-that coaching was essential for success in transformation [P8, P18, P29, P33].
-Also, without coaching teams can do damage to the agile transformation if
-techniques are applied improperly [P9].
-
-There were a number of statements of the benefits of coaching. The coach
-position allowed to watch for and correct problems when they arose [P16, P42].
-Coaches also helped to draw attention away from focus on tools and towards
-understanding the principles of agile [P42, P48].
-There were also benefits in using both internal and external coaches.
-External coaches were able to have an objective view of the organization [P16,
-P44], but internal coaches were more accessible and knew the specifics of the
-organization [P3, P6].
+% Make it understood that there is no return to the old model
+
+Articulating urgency at start will help getting the change under way [P49].
+Feedback of changes should certainly be listened to, but in the end it must be
+made clear that the old model can not be returned to [P22].
+For a case where the organization was changed all at once it was reported that
+the magnitude of change helped created the impression that the change was non
+negotiable [P11]. Firstly, as it was evident that a large change was to happen
+people felt encouragement and permission to move on from the old way, and
+secondly, the urgency eliminated wasteful discussions on whether Scrum was a
+good model or not [P11].
+When the organization culture is ingrown he change vision must constantly be
+reminded of, and each step towards the new model consider a victory [P48].
+
+% Make change necessary
+
+Transformation can be facilitated by creating mechanisms that force change.
+Spayd [P49] describes that senior management pushed hard on a mandate to have
+deliveries every 90 days. The mandate made change necessary, and while the
+strong pressure did not promote agile practices in every case, the drive for
+change was perceived good in general [P49].
+
+% Commitment (?)
+
+A large change will inevitably require strong commitment, but problems in the
+transformation can put the commitment to test. The agile model is introduced
+because of problems in the old model, and therefore there will be organizational
+issues uncovered during the transformation [P10]. People must not be demoralized
+when facing challenges, and the determination to change must be maintained [P8,
+P10]. It is not the agile model that has flaws, but it is merely exposing
+problems in the organization [P10].
+
+Ryan and Scudiere [P43] describe that management commitment was showing as
+strong focus on certain agile practices that had been chosen as non negotiable,
+and constant assessment that the practices work. The expectations were clearly
+communicated to the teams, and constant assessment made it clear that change is
+desired [P43].
+A strong commitment from management will assure teams that the change is the
+right thing [P11].
+
+% * Create a sense of urgency
+% * Find a common cause to motivate the change
+% [P11]
+% Find a rallying point.  You might not have a recession as a spur to receptivity,
+% but a common, critical cause to motivate change will make it easier to explain
+% how Scrum can contribute to the success  of that cause.
+
+\subsubsection{Leadership and alignment}
+
+The importance of leadership in change was described in many sources. Individual
+persons as well as task forces and communities acted as change leaders. A factor
+related to leadership was creating organizational alignment.
+Table \ref{table:success_changeleaders} summarizes factors related to leadership
+in change which benefit the transformation.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Have change leaders  &
+                P3, P4, P11, P16, P18, P31, P33, P42  \\
+        Have an agile community  &
+                P3, P12, P41, P42, P47  \\
+        Align the organization  &
+                P8, P18, P31, P42, P43, P46  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Success factors relating to leadership in change}
+    \label{table:success_changeleaders}
+\end{table}
+
+% Change leaders
+
+Organizing a large number of people requires some management and leadership.
+Similarly, transforming the way of working of a large group requires coordination
+and leadership [P31, P42]. In addition to the leadership provided by coaches,
+specific change leaders were also mentioned. Cases indicated the importance of
+having spokespersons for the change [P3, P4, P31]. Cowan [P11] describes how one
+person was strongly driving the transformation, and made an indisputable
+contribution for transforming the organization. Also Maples [P31] describes how
+the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
+described that the change was lead by a competent roll-out team, which had
+representatives from all parts of the organization [P16, P33]. Finally, the
+responsibility of line and project managers to act as change leaders was
+highlighted [P18].
+
+% Have an agile community
+
+The formation and influence of agile communities was reported to have a
+significant impact on transformation.
+The emergence of an agile community was reported as a success factor in the
+transformation [P3]. A community was also seen as indispensable as it has power
+to overcome impediments that would be too large for individuals to affect [P47].
+The agile communities achieved to raise awareness of agile methods in the
+organization [P41, P42, P47], and spawned eager followers who spread the word
+even further [P12].
+
+% Align the organization
+
+A factor in achieving change was creating alignment towards the common goal of
+introducing new development methods. It was seen as an important factor that all
+levels in the organization speak the same language and accept the change [P18,
+P31]. Alignment was built by promoting success stories and gathering lists of
+problems to tackle [P42, P46]. In one case complete alignment between teams and
+within management was considered as necessary to use the agile model in a large
+context [P43]. Another case highlighted creating and applying a structured
+roll-out process to uniformly introduce agile to a large number of teams [P8].
 
@@
         \midrule
-        Customizing the agile model  &
+        Customize the agile model to the needs of the organization  &
                 P8, P10, P11, P29, P31, P40, P41, P43, P45, P49, P51  \\
@@
                 P14, P29, P44, P46, P48  \\
-        Keeping the new model simple  &
+        Keep the new model simple  &
                 P5, P10, P16, P40  \\
@@
 
-\begin{table}
+\begin{table}[h]
     \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.46\textwidth}
+                     >{\raggedright\arraybackslash}p{0.46\textwidth} }
         \toprule
@@
 
-\subsubsection{Management support}
-
-The most prominent and clear success factor was having support from management.
-When the leadership of the organization is committed and supportive change is
-much more likely to happen. Table \ref{table:success_management} lists
-success factors related to management support.
-
-\begin{table}
+\subsubsection{Training and coaching}
+
+Agile methods come as new ways of working to many. In order to get a good start
+in the transformation the ideas of the new methods need to be explained
+thoroughly. Training sessions are good starting points, but teams should be
+accompanied by coaching for a longer time to ensure that the new practices are
+learned and applied correctly in practice. Table \ref{table:success_training}
+lists references that highlight training and coaching as success factors.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.46\textwidth}
+                     >{\raggedright\arraybackslash}p{0.46\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Providing training on agile methods  &
+                P3, P6, P11, P14, P16, P30, P37, P46 \\
+        Coaching teams as they learn by doing  &
+                P3, P6, P8, P9 P16, P17, P18, P29,
+                P30, P33, P42, P44, P47, P48  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Training and coaching presented as success factors}
+    \label{table:success_training}
+\end{table}
+
+Training was seen as a strong factor helping the organization to transform.
+Several studies stated that training improved the chance of succeeding in the
+transformation [P3, P16, P30]. It was even highlighted that the change would have
+failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
+and again how training could make the difference between success and failure for
+a team''. Training also helped people to become more positively inclined towards
+the new model [P16, P37], and at best training made people enthusiastic to change
+[P3, P46]. A few cases reported that there was some disinclination to change,
+and training would have helped to overcome [P11, P14].
+
+Agile methods avoid prescribing exact ways of working but rather emphasize a
+mindset of adapting to the situation. It is difficult to explain by theory how
+such a mindset should be applied, and the agile practices are best learned by
+doing. Coaching teams while they apply the agile methods in practice was seen as
+an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
+that coaching was essential for success in transformation [P8, P18, P29, P33].
+Also, without coaching teams can do damage to the agile transformation if
+techniques are applied improperly [P9].
+
+There were a number of statements of the benefits of coaching. The coach
+position allowed to watch for and correct problems when they arose [P16, P42].
+Coaches also helped to draw attention away from focus on tools and towards
+understanding the principles of agile [P42, P48].
+There were also benefits in using both internal and external coaches.
+External coaches were able to have an objective view of the organization [P16,
+P44], but internal coaches were more accessible and knew the specifics of the
+organization [P3, P6].
+
+\subsubsection{Communication and transparency in the change}
+
+Communication and transparency was an important success factor brought up in the
+promary studies. The ongoing change should be communicated intensively to
+everyone. Transparency and inclusiveness were success factors related to
+communication.
+Table \ref{table:success_communication} presents success factors relating to
+communicating the change.
+
+\begin{table}[h]
     \centering
@@
         \midrule
-        Have management support  &
-                P8, P11, P16, P18, P27, P31, P33, P36,
-                P43, P44, P46, P47, P49  \\
-        Visible management support encourages employees  &
-                P9, P11, P24, P40  \\
-        Educate management on agile methods  &
-                P6, P8, P9, P11, P16, P29, P35, P46, P47  \\
+        Communicate the change intensively  &
+                P16, P22, P33, P40, P41, P43, P48  \\
+        Enable transparency in the change  &
+                P6, P11, P16, P42, P43  \\
+        Engage people broadly in the organization  &
+                P4, P6, P16, P22, P31 \\
         \bottomrule
     \end{tabular}
-    \caption{Success factors relating to management support for agile}
-    \label{table:success_management}
+    \caption{Succes factors relating to communicating the change}
+    \label{table:success_communication}
 \end{table}
-
-% Management support in general
-
-Numerous cases made it clear that management support is an absolute necessity
-[P8, P18, P27, P31, P33, P36, P43, P47, P49].
-This was reflected in statements such as ``Adopting agile, or implementing any
-significant change, requires an executive’s sincere support.'' [P44],
-``Executive commitment was crucial to implementing massive change.'' [P16], and
-``Having upper management engaged, supportive and visible is critical for
-wholesale organization involvement with Scrum.'' [P11].
-
-Managers were seen to be in a key role in making the change stick, as they had
-the authority and power to remove impediments [P18]. Seffernick [P46] describes
-how a number of people attempted to explain away the applicability of agile
-methods, but the objections were overruled by the director's commitment to make
-agile work. Management support was similarly needed when tight release schedules
-had to be flexed in order to give room for the adoption process [P16, P27].
-Favorable management decisions were also critical when additional resources were
-allocated to training and coaching [P47].
-
-Visible involvement of management was also reported to motivate and encourage
-employees to adopt the new model [P40]. For instance, the CTO organized training
-sessions personally [P24], and the engineering VP frequently visited sprint
-demos [P11]. When the corporate level support for the agile initiative was
-showing teams adopted agile methods even spontaneously [P9].
-
-% Educate management
-
-In order to gain support for agile several cases highlighted the need to educate
-management [P8, P47].
-Without education managers might make harmful interventions in the process, or
-just stick to the old model [P16, P35, P46].
-Managers not understanding the principles of agile were feeling left out with
-the introduction of self organizing teams, which some times resulted in
-backlashes [P6]. Providing proper training was reported to correct the
-situation, and even create strong agile supporters in management [P6, P9].
-Training also cleared misconceptions that management had [P29], and helped
-create a consistent implementation of the agile model across the organization
-[P11].
-
-There were a few suggestions on how to involve management better. Foremost, a
-successful experience will likely rise the interest of management [P36].
-Mencke [P33] suggests giving executives tasks related to the agile process,
-so that they have better visibility of the new way of working.
-
-\subsubsection{Urgency and commitment to change}
-
-% Make it understood that there is no return to the old model
-
-To make a large organization change the direction of the change needs to be
-communicated clearly. Articulating urgency at start will help getting the change
-under way [P49]. Feedback of changes should certainly be listened to, but in the
-end it must be made clear that the old model can not be returned to [P22].
-For a case where the organization was changed all at once it was reported that
-the magnitude of change helped created the impression that the change was non
-negotiable [P11]. Firstly, as it was evident that a large change was to happen
-people felt encouragement and permission to move on from the old way, and
-secondly, the urgency eliminated wasteful discussions on whether Scrum was a good
-model or not [P11].
-When the organization culture is ingrown he change vision must constantly be
-reminded of, and each step towards the new model consider a victory [P48].
-
-% Commitment (?)
-
-A large change will inevitably require strong commitment. The agile model is
-introduced because of problems in the old model, and therefore there will be
-organizational issues uncovered during the transformation [P10]. People must not
-be demoralized when facing challenges, and the determination to change must be
-maintained [P8, P10]. It is not the agile model that has flaws, but it is
-merely exposing problems in the organization [P10].
-Ryan and Scudiere [P43] describe that management commitment was showing as
-strong focus on certain agile practices that had been chosen as non negotiable,
-and constant assessment that the practices work. The expectations were clearly
-communicated to the teams, and constant assessment made it clear that change is
-desired [P43].
-A strong commitment from management will assure teams that the change is the
-right thing [P11].
-
-% Make change necessary
-
-Spayd [P49] describes that senior management pushed hard on a mandate to have
-deliveries every 90 days. The mandate made change necessary, and while the
-strong pressure did not promote agile practices in every case the drive for
-change was perceived good in general [P49].
-
-% * Create a sense of urgency
-% * Find a common cause to motivate the change
-% [P11]
-% Find a rallying point.  You might not have a recession as a spur to receptivity,
-% but a common, critical cause to motivate change will make it easier to explain
-% how Scrum can contribute to the success  of that cause.
-
-\subsubsection{Have a change leading coalition}
-
-% Change leaders
-
-Organizing a large number of people requires some management and leadership.
-Similarly, transforming the way of working of a large group requires coordination
-and leadership [P42, P31]. In addition to the leadership provided by coaches,
-specific change leaders were also mentioned. Cases indicated the importance of
-having spokespersons for the change [P3, P4, P31]. Cowan [P11] describes how one
-person was strongly driving the transformation, and made an indisputable
-contribution for transforming the organization. Also Maples [P31] describes how
-the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
-described that the change was lead by a competent roll-out team, which had
-representatives from all parts of the organization [P16, P33]. Finally, the
-responsibility of line and project managers to act as change leaders was
-highlighted [P18].
-
-% Have an agile community
-
-The formation and influence of agile communities was reported to have a
-significant impact on transformation.
-The emergence of an agile community was reported as a success factor in the
-transformation [P3]. A community was also seen as indispensable as it has power
-to overcome impediments that would be too large for individuals to affect [P47].
-The agile communities achieved to raise awareness of agile methods in the
-organization [P42, P41, P47], and spawned eager followers who spread the word
-even further [P12].
-
-% Align the organization
-
-A factor in achieving change was creating alignment towards the common goal of
-introducing a new development method. It was seen as an important factor that
-all levels in the organization speak the same language and accept the change
-[P18, P31]. Alignment was built by promoting success stories and gathering lists
-of problems to tackle [P42, P46]. In one case complete alignment between teams
-and within management was considered as necessary to use the agile model in a
-large context [P43]. Another case highlighted creating and applying a structured
-roll-out process to uniformly introduce a large number of teams to agile [P8].
-
-\subsubsection{Communication and transparency in the change}
 
@@
 
+Also communicating the goals of the transformation was highlighted in a some
+studies.
+Having a clear message of expectations helped to remove confusion and make
+people understand the purpose of the transformation [P43, P48]. The motivation
+to change can be increased by having a simple proposal on how to reach the
+desired outcomes [P40].
+McDowell and Dourambeis [P32] describe how the organization launched a series of
+communications events especially designed to emphasize the reasons behind the
+agile practices.
+
 Enabling transparency during the transformation was another communication
@@
 
-Communicating the goals of the transformation was highlighted in a some studies.
-Having a clear message of expectations helped to remove confusion and make
-people understand the purpose of the transformation [P43, P48]. The motivation
-to change can be increased by having a simple proposal on how to reach the
-desired outcomes [P40].
-
-% Emphasize principles over mechanics
-
-%         Emphasizing principles over mechanics  &
-%                 P1, P16, P32, P42, P46, P48  \\
-
-In a number of studies it was described that during transformation the
-principles of agile should be emphasized over practices and simple mechanics
-[P46, P48]. When people understand the agile values they will also understand
-why the change is being done and feel motivated [P1, P16, P42].
-The people centric nature of agile development should be communicated correctly.
-McDowell and Dourambeis [P32] describe how the organization launched a series of
-communications events especially designed to emphasize the reasons behind the
-agile practices.
-
-% ??
-Especially inexperienced coaches and Scrum masters made mistakes in focusing too
-much on the implementation of practices [P23, P30].
-
 % Engage people broadly
 
-Many cases highlighted how people were engaged broadly in the organization.
+Many cases highlighted the importance of engaging people broadly in the
+organization.
 One of the lessons learned presented was that it is important to involve all
@@
 team made an effort to engage people by both inviting to give feedback online
-and holding an extensive number of feedback meetings [P6, P22, P16]
-Being inclusive towards everyone was seen as one of the key success factors in
-the transformation [P16]. Inclusiveness will encourage people to participate and
-use the new model visibly [P16].
-
-\subsubsection{Enable autonomy}
+and holding an extensive number of feedback meetings [P6, P16, P22] Being
+inclusive towards everyone was seen as one of the key success factors in the
+transformation [P16]. Inclusiveness will encourage people to participate and use
+the new model visibly [P16].
+
+\subsubsection{Enable autonomy for development teams}
+
+A key principle in agile is to allow development teams decide for themselves how
+they execute their tasks. To make te agile model work well teams need to
+self-organize, and be allowed to make decisions on the pace of development.
+When teams can perform their tasks autonomically the transformation towards
+agile will progress.
+Table \ref{table:success_autonomy} lists success factors relating to autonomy of
+development teams.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
+                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Allow teams to self-organize  &
+                P5, P16, P18, P41  \\
+        Empower teams to decide over speed and quality  &
+                P27, P30, P34, P45  \\
+        No top-down mandate  &
+                P3, P6, P41  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Success factors relating to enabling autonomy for development teams}
+    \label{table:success_autonomy}
+\end{table}
 
@@
 communicating the transformation started to decline. To enable the
-transformation to still proceed, the an organization-wide challenge was created
+transformation to proceed further, an organization-wide challenge was created
 where teams would independently develop and showcase the best agile practices.
@@
 It was also thought that proceeding with a top-down mandate would have made
-teams use a uniform process [P41]. This would have been sub-optimal, as it was
-important that the practices were tailored for each team and project [P41].
-
-% Previous culture is like agile
-
-A few organizations had already some structures in place that were similar to
-agile, which made the transformation easier.
-For instance, a previous organizational model based on small and autonomous
-teams strongly aided the adoption [P3].
-Another case described the original on-demand delivery model a natural fit for
-agile, and the transformation was presented as a return to core values instead
-of a remodeling [P16].
-% The agile culture must also be maintained in each step of the transformation
-% [P31]
-
-\subsubsection{People factors}
+teams conform to a single process [P41]. This would have been sub-optimal, as it
+was important that each individual team and project tailored their practices
+[P41].
+
+\subsubsection{Choosing the right people}
 
@@
 management, but also different people factors were given as basis of success.
+The choice of people to initially involve in the transformation was reported to
+make a difference. Interest or previous experience of agile methods were
+valuable characteristics to look for.
+Table \ref{table:success_people} lists factors choosing people that can benefit
+the transformation.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ ll }
+        \toprule
+        Success factors  &  Described in \\
+        \midrule
+        Start with people who are receptive for agile methods  &
+                P6, P29, P34  \\
+        Use people with previous agile experience  &
+                P6, P10, P37  \\
+        Have change leaders without baggage of the past  &
+                P11, P16  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Success factors relating to choosing the right people}
+    \label{table:success_people}
+\end{table}
 
@@
 % had gone through a time of uncertainty and restructuring [P11].
-
-% Have social events
-
-Social events were reported to benefit the transformation. A few cases even
-described a the transformation being driven by a series of events where people
-received information and had the possibility to participate in shaping the new
-model of working [P32, P41]. Various social activities were presented as
-valuable tools for improving bonding within teams [P11, P27, P40]. The importance of
-creating personal bonding was also highlighted for change leading entities such
-as management and coaching teams [P11, P23].
 
@@
 
+\subsubsection{Other success factors}
+
+A few succes factors which did not fit in the above categories stood out. These
+success factors are creating a positive experience in the beginning, arranging
+socaial events, emphasizing principles over mechanics, and the previous
+organizational culture being like agile.
+
 % Having a positive experience
@@
 
+% Have social events
+
+Social events were reported to benefit the transformation. A few cases even
+described a the transformation being driven by a series of events where people
+received information and had the possibility to participate in shaping the new
+model of working [P32, P41]. Various social activities were presented as
+valuable tools for improving bonding within teams [P11, P27, P40]. The importance of
+creating personal bonding was also highlighted for change leading entities such
+as management and coaching teams [P11, P23].
+
+% Emphasize principles over mechanics
+
+In a number of studies it was described that the principles of agile should be
+emphasized over practices and simple mechanics [P46, P48]. When people
+understand the agile values they will also understand why the change is being
+done and feel motivated [P1, P16, P42].
+Some cases described that inexperienced coaches and Scrum masters made mistakes
+in focusing too much on the implementation of practices [P23, P30].
+
+% Previous culture is like agile
+
+A few organizations had already some structures in place that were similar to
+agile, which made the transformation easier.
+For instance, a previous organizational model based on small and autonomous
+teams strongly aided the adoption [P3].
+Another case described the original on-demand delivery model a natural fit for
+agile, and the transformation was presented as a return to core values instead
+of a remodeling [P16].
+
 \clearpage

### 2014-06-05 journal/article.tex
Spell check

@@
 Communication and transparency was an important success factor brought up in the
-promary studies. The ongoing change should be communicated intensively to
+primary studies. The ongoing change should be communicated intensively to
 everyone. Transparency and inclusiveness were success factors related to
@@
     \end{tabular}
-    \caption{Succes factors relating to communicating the change}
+    \caption{Success factors relating to communicating the change}
     \label{table:success_communication}
@@
 A key principle in agile is to allow development teams decide for themselves how
-they execute their tasks. To make te agile model work well teams need to
+they execute their tasks. To make the agile model work well teams need to
 self-organize, and be allowed to make decisions on the pace of development.
@@
 
-A few succes factors which did not fit in the above categories stood out. These
+A few success factors which did not fit in the above categories stood out. These
 success factors are creating a positive experience in the beginning, arranging
-socaial events, emphasizing principles over mechanics, and the previous
+social events, emphasizing principles over mechanics, and the previous
 organizational culture being like agile.

### 2014-06-06 journal/article.tex
Research method: editing, additions

@@
 
+The research process is presented starting from the research questions. After
+that we present the definition of large scale which this research builds upon.
+The bulk of this section concentrates on presenting the steps of the research
+process, starting from the definition of inclusion criteria, continuing with
+describing how the primary studies were selected, and finally presenting the
+method for synthesizing the primary studies. The section concludes with a
+discussion on limitations of the method.
+
 \subsection{Research questions}
@@
 
-A key question in the study was the definition of \emph{large scale}. Large
+A key question in this study was the definition of \emph{large scale}. Large
 scale requires additional coordination and makes communication difficult
-compared to software development in small teams.
+compared to software development in small teams. One of the motivations for this
+research was to identify phenomena that appear particularly in large-scale
+transformations, and which may not be present in small organizations.
+
+In this section we first present the definition of large scale.
 
@@
 The research was conducted as an application of Kitchenham's
-\cite{Kitchenham2007} guidelines for systematic literature review. The selection
-of primary studies was done as a keyword based database search and manual
-filtering of included studies. The filtering process was executed independently
-by two researchers (Dikert, Paasivaara). The selected primary studies were coded
-for data extraction and analysis. The entire process was audited and mentored by
-a third researcher (Lassenius).
+\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
+primary studies was done as a keyword based database search and manual filtering
+of included studies. The filtering process was executed independently by two
+researchers (Dikert, Paasivaara). The selected primary studies were coded for
+qualitative data extraction and analysis. The entire process was audited and
+mentored by a third researcher (Lassenius).
 
@@
 
-Before proceeding with the database search a number of preliminary searches were
-performed. The purpose of the preliminary searches was to create a benchmark of
-potentially relevant papers that should be matched by the actual database
+Before proceeding with identifying the primary studies a number of preliminary
+searches were performed. The purpose of the preliminary searches was to create a
+benchmark of potentially relevant papers that should be matched by the actual
 search, and to evaluate different search strings. We started by examining top
@@
 
-\subsubsection{Study selection}
-
-The selection of studies was primarily based on a database search. The initial
-step in the database search was constructing a search string which yielded a
-reasonably large result set. The search result was then refined in two steps, by
-first filtering out abstracts and finally full text filtering. The study
-selection process starting from the database search is outlined in figure
-\ref{fig:selectionprocess}. The result of the database search was still amended
-with some results from the preliminary search and relevant references from all
-examined papers.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/research_process.pdf}
-    \caption{Database search as study selection process}
-    \label{fig:selectionprocess}
-  \end{center}
-\end{figure}
-
-A keyword search in online databases was chosen as the primary means to find
-potential primary studies. The databases included in the search are listed in
-Table \ref{table:databases}. All databases supported use of complex Boolean
-logic in searches, which allowed us flexibility in constructing the search
-strings.
+\subsubsection{Identification of primary studies}
+
+The gathering of potential primary studies was based on a search in online
+databases. The databases searched are listed in Table \ref{table:databases}. All
+databases supported use of complex Boolean logic in searches, which allowed us
+flexibility in constructing the search strings. The main task in the database
+search was to construct a search string which yielded a reasonably large result
+set. The set of matched primary studies was later ammended with a small number
+of studies found in references of the matched papers.
 
@@
 
-The database keyword search matched 1874 unique papers. The abstracts of these
-papers were categorized independently by both Dikert and Paasivaara into three
-categories: include, exclude, and uncertain. The result was combined by marking
-abstracts both researchers agreed on for inclusion or exclusion respectively.
-The inclusion of disagreed or uncertain abstracts was resolved through
+The database keyword search matched 1875 unique papers. The results of the
+search were gathered for the selection step using reference management software.
+
+\subsubsection{Study selection}
+
+The set of potential primary studies identified by the database search was
+refined in two steps, by first filtering out abstracts and finally full text
+filtering. The study selection process starting from the database search is
+outlined in figure \ref{fig:selectionprocess_full}. The result of the database
+search was amended with some results from the preliminary search and relevant
+references from all examined papers.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
+    \caption{Full study selection process}
+    \label{fig:selectionprocess_full}
+  \end{center}
+\end{figure}
+
+TODO:
+* Duplicates removed?
+* Ei voi ymmärtää miten preliminary search filtteröitiin
+
+The database keyword search had matched 1875 unique papers. The abstracts of
+these papers were categorized independently by both Dikert and Paasivaara into
+three categories: include, exclude, and uncertain. There were 1578 exclusions
+and 62 inclusions that both researchers agreed on. The inclusion decisions
+for the 235 abstracts with disagreement or uncertainty were resolved through
 discussion. At this stage papers were excluded only if both researchers deemed
@@
 a result 170 papers were selected for full text filtering.
-
-We also evaluated the keyword search against the benchmark created in the
-preliminary search step, concluding that 75 of the 117 preliminarily selected
-papers matched. The missed 32 preliminary papers were examined, resulting in
-including 3 additional papers as primary sources.
 
@@
 
-After the full text filtering step the references of all papers were also
-examined for relevance. Most case study papers used references very scarcely,
-mostly referencing well known descriptions of agile methods. We included 2
-referenced papers in full text analysis.
+We evaluated the result of the database keyword search against the benchmark
+created in the preliminary search step. We concluded that 75 of the 117
+preliminarily selected papers were matched by the database keyword search.
+The missed 32 preliminary papers were examined, resulting in including 3
+additional papers as primary sources.
+
+In parallell with the full text filtering step the references of all papers were
+also examined for relevance. Most papers used references very scarcely, typicaly
+referencing well known descriptions of agile methods. We included 2 referenced
+papers in full text analysis.
 
@@
 preliminary searches, and 2 papers from references, 52 papers were selected
-as primary studies. The full selection process is illustrated in figure
-\ref{fig:selectionprocess_full}.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
-    \caption{Full study selection process}
-    \label{fig:selectionprocess_full}
-  \end{center}
-\end{figure}
-
-TODO:
-* Duplicates removed?
-* Ei voi ymmärtää miten preliminary search filtteröitiin
+as primary studies.
+
+\subsubsection{Handling of duplicate reports on a single case}
 
 In several cases there were more than one primary study presenting the
-transformation of the same organization. Even if the transformation of a single
-organization was described in many studies all sources that passed the inclusion
-criteria were included. Studies presenting the same organization were treated
-as one unit so that we could gain as much insight from each organization as
-possible. There were also a few papers that presented multiple case studies, and
-in those cases we treated each studied organization individually. As a result 42
-unique organizations were identified in the primary studies. We use the term
-\emph{study} to refer to the primary study publications, and the term
-\emph{case} to refer to an individual case organization that may be described in
-several different studies.
-
-% -----------------------
-
-% - pitää olla systemaattinen
-% - pitää olla toistettava
-% ==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
+transformation of the same organization. Duplicate descriptions of the same
+organization focused typically on different aspects. For example, one paper
+would highlight the viewpoint of developers \cite{Fry2007}, and another would
+consider the transformation from user experience designers' point of view
+\cite{Federoff2009}.
+
+Even if the transformation of a single organization was described in many
+studies, all sources that passed the inclusion criteria were included. Studies
+presenting the same organization were treated as one unit so that we could gain
+as much insight on each organization as possible. Conversely, there were also a
+few papers that presented multiple case studies, and in those cases we treated
+each studied organization individually.
+
+As a result 42 unique organizations were identified in the primary studies. We
+use the term \emph{study} to refer to the primary study publications, and the
+term \emph{case} to refer to an individual case organization that may be
+described in several different studies.
+
+\subsubsection{Study quality assessment}
+
+The primary research for this literature review consists almost exclusively of
+industry experience reports. There were only 6 case studies with a research
+method, and observations on transformation were presented only as a minor part
+in these studies. Based on this finding we concluded that case studies
+presenting viewpoints on organizational transformations are very scarce in
+software industry. We deemed that the results would be distorted heavily and
+many valuable studies would be left out if a strict quality assessment would be
+part of the inclusion criteria. As a result we decided to include all experience
+reports, regardless of the perceived objectivity.
+
+The absence of research methods and subjective viewpoints of experience reports
+can be seen as a factor lessening the evidence. Because of this we were
+refrained from making quantitative interpretations of the data, but based the
+results around describing qualitative observations. Regardless of the bias of
+the primary studies, we think that the concepts that emerged from the source
+material should not be questioned on a fundamental level. The overall existence
+of organizational phenomena should not be questioned based on possible author
+bias in experience reports.
+
+% --> Large organizational changes are hard to measure experimentally (controlled
+%     experiments?). Also due to this any student research was excluded.
+
+% --> It may still be interesting to make some quantitative observations such as
+%     a particular problem being reported in the majority of reports.
+
+% * The primary studies did not make distinction between method, results, and
+%   discussion
 % 
-% Jos koko on epäselvä:
-% --> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
-% --> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
-%     pois
-
-\subsubsection{Study quality assessment}
-
-The primary research for this literature review consists almost exclusively of
-industry experience reports. This may be seen as a factor lessening the
-evidence.
---> Why did we do this anyway?
---> Case studies with research method studying organizational transformations in
-    software industry are very scarce.
---> Large organizational changes are hard to measure experimentally (controlled
-    experiments?). Also due to this any student research was excluded.
-
-The reason for deviating study assessment
---> Original research consisted almost solely of experience reports
---> The fact that experience reports have a subjective standpoint makes it
-    unreliable to make quantitative conclusions based on the data
---> We also decided to include separate studies describing the same organization.
-    This was useful as it deepened the understanding of each case, and several
-    large organizations have produced more than one description of their agile
-    transformation.
-
-The fact that the study is based on a quantitative basis
---> Experience reports have a subjective bias
---> The data synthesis must not rely on quantitative measures. This means that
-    we also considered topics raised by only a marginal number of reports.
---> It may still be interesting to make some quantitative observations such as
-    a particular problem being reported in the majority of reports.
-
-It was quite common that a single transformation case was described by many
-papers. For example, one paper would highlight the viewpoint of developers
-\cite{Fry2007}, and another would consider the transformation from user
-experience designers' point of view \cite{Federoff2009}. Instead of using the
-most complete paper as suggested by Kitchenham \cite{Kitchenham2007}, we
-combined the results presented in each paper and considered the case as a single
-unit. By doing this we avoided bias caused by duplicate publications, and were
-enabled to have a more detailed view of the case organization.
+% * The studies neither did any implications for generalizability, but rather
+%   presented narratives or experiences. Some cases presented lists of
+% %   recommendations.
+
+As presented in the previous section, we allowed several descriptions of the
+same case to be included. Instead of using the most complete paper as suggested
+by Kitchenham \cite{Kitchenham2007}, we combined the results presented in each
+paper and considered the case as a single unit. Keeping in mind the potential
+bias caused by duplicate publications, including all papers enabled us to have a
+more detailed view of the case organization.
 
 \subsubsection{Coding of primary documents}
+\label{sec:coding}
 
@@
 
+\subsubsection{Synthesis of primary studies based on coding}
+
+The synthesis of findings was created based on the result of coding presented in
+section \ref{sec:coding}. An initial organization of codes into categories was
+made based on the labels of the codes. Each codes was linked to a single
+research question, so the categorization of codes was made separately within
+the topic of each research question. The categories of codes were labeled
+appropriately to give a high level description of the linked codes. Some codes
+having minor importance and few quotations were left out.
+
+After assigning each code to a unique category, the content of the categories
+was studied. Each category was studied by reading through each quotation of each
+code included. Typically the quotations were displayed and examined in their
+original context, considering also surrounding paragraphs of the quotation.
+Notes were made on each quotation presenting noteworthy observations. Based on
+the notes, an accurate description emerged for each code.
+
+As an accurate description had been created for each code, the initially defined
+categories could be refined. Some codes were recategorized, and the definitions
+of a few categories were revisited, until a final ordering was reached. The
+results are presented according to the final categorization.
+
+% * If a code had more quotations there were more observations made (longer
+%   result text)
+
 \subsection{Limitations of research method}
@@
    meta-analysis section.
+
+Bias  --  systematic error
+
+Internal quality  --  validity
+
+External quality  --  generalizability
 

### 2014-06-07 journal/article.tex
Threats to validity; research process figures

@@
 describing how the primary studies were selected, and finally presenting the
-method for synthesizing the primary studies. The section concludes with a
-discussion on limitations of the method.
+method for analyzing and synthesizing the primary studies. The section concludes
+with a discussion on limitations of the method.
 
@@
 \cite{Kitchenham2007} guidelines for systematic literature review. Selection of
-primary studies was done as a keyword based database search and manual filtering
-of included studies. The filtering process was executed independently by two
-researchers (Dikert, Paasivaara). The selected primary studies were coded for
-qualitative data extraction and analysis. The entire process was audited and
-mentored by a third researcher (Lassenius).
+primary studies was done using first a keyword based database search to identify
+potentially relevant sources, and then manually filtering the search result. The
+filtering process was executed independently by two researchers (Dikert,
+Paasivaara). The selected primary studies were coded for qualitative data
+extraction by one researcher (Dikert). The results of the study were finally
+elicitated by aggregating and analyzing the coding of the primary documents.
+Figure \ref{fig:researchprocess} presents an outline of the research process.
+The entire process was audited and mentored by a third researcher (Lassenius).
+
+\begin{figure}[b]
+  \begin{center}
+    \includegraphics{graphics/research_process.pdf}
+    \caption{Outline of the research process}
+    \label{fig:researchprocess}
+  \end{center}
+\end{figure}
 
@@
     \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
-    \caption{Full study selection process}
+    \caption{The study selection process}
     \label{fig:selectionprocess_full}
@@
 \end{figure}
-
-TODO:
-* Duplicates removed?
-* Ei voi ymmärtää miten preliminary search filtteröitiin
 
@@
 
-\subsubsection{Coding of primary documents}
+\subsubsection{Coding of primary studies}
 \label{sec:coding}
 
-The primary documents were coded using the Atlas TI qualitative data analysis software.
+The primary studies were coded using the Atlas TI qualitative data analysis software.
 An integrated deductive and inductive approach was chosen for coding the primary
@@
 
-\begin{table}[h]
+\begin{table}[t]
     \centering
@@
 
-\clearpage
-
 \subsubsection{Validation of coding}
@@
 In the experiment Dikert and Paasivaara coded independently a sample of five
-primary documents. The sample was hand picked as a representative subset of the
+primary studies. The sample was hand picked as a representative subset of the
 source material. The coding was done relying solely on the guidelines presented
@@
 
-\subsection{Limitations of research method}
-
--- A keyword search can not identify the facet of \emph{large scale}, and also
-   \emph{transformation} is difficult to capture with keywords. Because of this
-   we assume that it is likely that a some studies have been missed by the
-   keyword search. As an improvement we would suggest combining the entire
-   archive of \emph{Agile Conference} and \emph{Agile Processes in Software
-   Engineering and Extreme Programming} to the results of the keyword search.
-   These two conferences provided more than two thirds of the primary sources.
-
--- Few of the primary studies presented a research method. Most of the studies
-   were experience reports where the author was a member of the organization
-   discussed. Because of this we must assume that the primary studies are
-   biased. Due to this risk of bias we chose not to make statistical
-   interpretations of the results. We only use relative measure when comparing
-   results.
-
--- We avoid numbers in results, such as frequency of a phenomena, except in the
-   meta-analysis section.
-
-Bias  --  systematic error
-
-Internal quality  --  validity
-
-External quality  --  generalizability
+\subsection{Threats to validity and limitations}
+
+To assess the validity of this research internal validity and external validity
+needs to be considered. In this section we present the identified threats to
+validity and the strategies used to mitigate them. We highlight two specific
+limitations in the research: weakness in keyword search, and the strength of
+evidence of primary sources.
+
+The threat to internal validity is subjective bias, which affects two parts of
+the research. First, the selection of the primary studies may have been
+distorted by interpreting the inclusion criteria falsely. This risk was
+mitigated at start by using three researchers in designing the inclusion
+criteria. When the inclusion criteria was subsequently applied the initial
+filtering was performed independently by two researchers, and unclear cases were
+resolved by case-by-case discussions between two or three researchers.
+
+The second part of the research that may have been affected by subjective bias
+was the elicitation of results by coding and analysis. Our tools for mitigating
+this threat were limited, as the work stages in question are particularly
+laborous and for resource constraints the process could not be duplicated. We
+assessed the bias of the coding step using an experiment, and prevented
+subjective bias in analysis by making the results as tracable as possible. The
+coding experiment revealed that two researchers can find roughly 2 of 3 themes
+in an experience report identically. The traceability of results is achieved by
+supplying references to each claim presented in the results.
+
+This study had only a minor threat to external validity. The systematic
+literature review approach includes steps to identify all studies that may be
+relevant in the study area. As long as the systematic approach is adhered to
+threats against external validity are minimized. External validity was also
+confirmed by saturation of the primary studies. Saturation was evident, as when
+the majority of the primary studies had been coded almost all quotations for the
+remaining studies could be assigned to existing codes.
+
+A particular problem in this systematic literature review was the limitations of
+boolean keyword searches in online databases. A keyword search can not identify
+the facet of \emph{large scale}, and also \emph{transformation} is difficult to
+capture with keywords. Because of this we assume that it is likely that a some
+studies have been missed by the keyword search, potentially causing bias. To
+eliminate the potential bias we would suggest combining the entire archive of
+\emph{Agile Conference} and \emph{Agile Processes in Software Engineering and
+Extreme Programming} to the results of the keyword search. These two conferences
+provided more than two thirds of the primary sources.
+
+Another particular problem was that few of the primary studies presented a
+research method. Most of the studies were experience reports where the author
+was a member of the organization discussed. Because of this we must assume that
+the primary studies are biased to some extent. Due to this risk of bias we chose
+not to make quantitative interpretations in the results, and use only
+qualitative analysis. We conclude that in the current state of research it is
+necessary to include studies with varying strength of evidence in order to
+relevantly aggregate evidence on large scale transformations.
 

### 2014-06-09 journal/article.tex
Small edits

@@
 
-A key question in this study was the definition of \emph{large scale}. Large
+A key question in this study was the definition of \emph{large scale}. Our goal
+was to identify phenomena that appear particularly in large-scale
+transformations, and which may not be present in small organizations. Large
 scale requires additional coordination and makes communication difficult
-compared to software development in small teams. One of the motivations for this
-research was to identify phenomena that appear particularly in large-scale
-transformations, and which may not be present in small organizations.
+compared to software development in small teams.
 
@@
 needs to be considered. In this section we present the identified threats to
-validity and the strategies used to mitigate them. We highlight two specific
-limitations in the research: weakness in keyword search, and the strength of
-evidence of primary sources.
+validity and the strategies used to mitigate them. We also highlight two
+specific limitations in the research: weakness in keyword search, and the
+strength of evidence of primary sources.
 
@@
 this threat were limited, as the work stages in question are particularly
-laborous and for resource constraints the process could not be duplicated. We
+laborous. For resource constraints the process could not be duplicated. We
 assessed the bias of the coding step using an experiment, and prevented
@@
 coding experiment revealed that two researchers can find roughly 2 of 3 themes
-in an experience report identically. The traceability of results is achieved by
+in an experience report identically. Traceability of results was achieved by
 supplying references to each claim presented in the results.
 
-This study had only a minor threat to external validity. The systematic
-literature review approach includes steps to identify all studies that may be
-relevant in the study area. As long as the systematic approach is adhered to
-threats against external validity are minimized. External validity was also
-confirmed by saturation of the primary studies. Saturation was evident, as when
-the majority of the primary studies had been coded almost all quotations for the
-remaining studies could be assigned to existing codes.
+This study had a minor threat to external validity. The systematic literature
+review approach includes steps to identify all studies that may be relevant in
+the study area. As long as the systematic approach is adhered to, including
+systematic identification of research, threats against external validity are
+minimized. External validity was also confirmed by saturation of the primary
+studies. Saturation was evident, as when the majority of the primary studies had
+been coded almost all quotations for the remaining studies could be assigned to
+existing codes.
 
@@
 studies have been missed by the keyword search, potentially causing bias. To
-eliminate the potential bias we would suggest combining the entire archive of
-\emph{Agile Conference} and \emph{Agile Processes in Software Engineering and
-Extreme Programming} to the results of the keyword search. These two conferences
+eliminate the potential bias we would suggest manually searching the entire
+archive of \emph{Agile Conference} and \emph{Agile Processes in Software
+Engineering and Extreme Programming} for relevant studies. These two conferences
 provided more than two thirds of the primary sources.

### 2014-06-18 journal/article.tex
Large scale

@@
 
+Dingsoyr:
+The nature of software development, being innovative work that is only partially
+compatible with programmed coordination, requires a strong emphasis on personal
+communication [Kraut, R. E. and Streeter, L. A., "Coordination in software
+development," Communications of the ACM, vol. 38, pp. 69-81, 1995]
+
 % ---------------------------------------------------------------------
@@
 scale requires additional coordination and makes communication difficult
-compared to software development in small teams.
-
-In this section we first present the definition of large scale.
+compared to software development in small teams. In this section we present the
+definition of large scale that is used in this study.
+
+Proceeding a workshop at XP2013 Dingsøyr et al. \cite{Dingsoyr2013} compiled a
+description of the definition of large scale. Dingsøyr et al. describe that
+at the XP2013 workshop the most common view on large scale was the involvment of
+many people. Dingsøyr et al. also conducted a brief literature search to
+identify what definitions on large scale are used. The results showed that
+large scale had been regadred in terms of size in persons or teams, project
+budget, code base size, and project duration. Paasivaara et al.
+\cite{Paasivaara2008} referred to an organization of 40 people and 7 teams.
+Berger and Beynon-Davies \cite{Berger2009} present the project cost of over £10
+million as a factor of large scale, but also highlight a team size of over 50
+people. A product with a code base size of over 5 million lines of code was
+considered to represent a large scale development effort by Petersen and Wohlin
+\cite{Petersen2010}. A project time of 2 years was considered as a factor of
+large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a project scope
+of 60-80 features.
 
@@
     SDM to use.
---> Some people see cost as a definition for large scale, some see code base
-    size [Petersen], some see project duration [Bjarnson?]. These are all
-    related, but do not relate to any specific problem. You can spend X amount
-    of money, but it does not in itself make any difference if the money is
-    spent in an agile way.
 --> Formal definitions (EU, etc.) look at project costs or company turnover.
@@
 
---> Ref: the XP conference
 --> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
@@
 
-Transformation and ``scaling up'' are very closely related and in some cases
-they are one and the same. For instance, if transformation begins with a pilot
-and spreads gradually through the organization, the process can very well be
-seen as a ``scaling up'' journey. However, we excluded cases where a initially
-small organization was scaled up \cite{Maranzato2012}, and discussions focusing
-on processes or tools without describing organizational change \cite{Lyon2008}.
+Transformation and ``scaling up'' of agile practices in use are very closely
+related concepts, and in some cases they are one and the same. For instance, if
+transformation begins with a pilot and spreads gradually through the
+organization, the process can very well be seen as a ``scaling up'' journey.
+However, we excluded cases where an initially small organization was scaled up
+\cite{Maranzato2012}, and discussions focusing on processes or tools without
+describing organizational change \cite{Lyon2008}.
 % Also [Greening 2010] on process centered viewpoint without transformation
@@
 
-A further example on exclusion by the \emph{large scale} facet were cases with
+Further examples on exclusion by the \emph{large scale} facet were cases with
 large organizations but only a single team adopting agile \cite{Fulgham2011}.
@@
 Also piloting cases that resulted only in single teams using agile were excluded
-\cite{Scott2008}.
+\cite{Scott2008}. Finally, cases where the organization was growing to large
+scale, but did not meet the size criteria at the start of the transformation,
+were excluded.
 
@@
 be encountered in large organizations.
-Table \label{table:challenges_largescale} summarizes these challenges.
+Table \ref{table:challenges_largescale} summarizes these challenges.
 

### 2014-06-18 journal/article.tex
Large scale

@@
 
-Proceeding a workshop at XP2013 Dingsøyr et al. \cite{Dingsoyr2013} compiled a
-description of the definition of large scale. Dingsøyr et al. describe that
-at the XP2013 workshop the most common view on large scale was the involvment of
-many people. Dingsøyr et al. also conducted a brief literature search to
-identify what definitions on large scale are used. The results showed that
-large scale had been regadred in terms of size in persons or teams, project
-budget, code base size, and project duration. Paasivaara et al.
-\cite{Paasivaara2008} referred to an organization of 40 people and 7 teams.
-Berger and Beynon-Davies \cite{Berger2009} present the project cost of over £10
-million as a factor of large scale, but also highlight a team size of over 50
-people. A product with a code base size of over 5 million lines of code was
-considered to represent a large scale development effort by Petersen and Wohlin
-\cite{Petersen2010}. A project time of 2 years was considered as a factor of
-large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a project scope
-of 60-80 features.
+Proceeding a workshop at XP2013 \cite{Dingsoyr2013a} Dingsøyr et al.
+compiled a description of the definition of large scale \cite{Dingsoyr2013b}.
+Dingsøyr et al. describe that at the XP2013 workshop the most common view on
+large scale was the involvment of many people \cite{Dingsoyr2013b}. Dingsøyr et
+al. conducted a brief literature search to identify what definitions on large
+scale are used. The results showed that large scale had been regadred in terms
+of size in persons or teams, project budget, code base size, and project
+duration.
+Paasivaara et al. \cite{Paasivaara2008} referred to an organization of 40 people
+and 7 teams. Berger and Beynon-Davies \cite{Berger2009} present the project cost
+of over £10 million as a factor of large scale, but also highlight a team size
+of over 50 people. A product with a code base size of over 5 million lines of
+code was considered to represent a large scale development effort by Petersen
+and Wohlin \cite{Petersen2010}. A project time of 2 years was considered as a
+factor of large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a
+project scope of 60-80 features.
 

### 2014-06-19 journal/article.tex
Large scale

@@
 
---> Needs to have complexity in software design?? --> What is complexity in software?
-    Is ``complexity'' an issue in agile?
-    What will bring complexity into software? Not costs or the length of the
+--> What will bring complexity into software? Not costs or the length of the
     schedule. --> Having many people will bring the complexity in terms of the
     SDM to use.
---> Formal definitions (EU, etc.) look at project costs or company turnover.
-    These definitions do not take a standpoint on complexity of the product, so
-    they do not matter in the context of SDMs.
 

### 2014-06-21 journal/article.tex
References from primary docs

@@
 
+What is agile?
+[6] Conboy K., Fitzgerald B., “Toward a conceptual framework for agile methods:
+a study of agility in different disciplines,”
+Proc. ACM workshop on Interdisciplinary software engineering research (WISER), 2004, pp. 37-44.
+
+\subsection{Large scale general references from primary studies}
+
+Laanti 2008:
+[2] Kähkönen Tuomo, “Agile Methods for large organizations – building communities of practice”.
+Proc. Agile Development Conf. (ADC), 2004, pp. 2-10.
+
+[7] Eckstein Jutta, “Agile Software Development in the large: Diving into the deep”,
+Dorset House Publishing, 2004. ISBN 0-932633-57-9.
+[8] Leffingwell Dean, “Scaling Software Agility: Best Practices for Large Enterprises”,
+Addison-Wesley Professional, 2007. ISBN-10: 0321458192.
+[12] Highsmith J., Agile for the Enterprise, “From Agile Teams to Agile Organizations”,
+Cutter, 2005. http://www.cutter.com/program/ fulltext/reports/2005/01/index.html .
+[13] Schwaber Ken, “Enterprise and Scrum”, Microsoft Press 2007, ISBN-10: 0735623376.
+
 \subsection{Things that may need to be explained}
@@
 project scope of 60-80 features.
+
+----
+
+Moore and Spens \cite{Moore2008} define large scale global development in the
+scope of their research as 300 people across 3 sites.
+
+----
 

### 2014-06-23 journal/article.tex
Large scale references

@@
 
+Vlaanderen 2012:
+22. Sidky, A.S.: A Structured Approach to Adopting Agile Practices: The Agile Adop-
+tion Framework (June 2007)
+
+Cockburn 2003:
+R. Crocker, Large-Scale Agile Software Development, Addison-Wesley, 2003.
+
 \subsection{Things that may need to be explained}
@@
 
+Elshamy \cite{Elshamy2006}: Roughly speaking, we consider a development project
+large if the development team is anywhere between 50 and 100 people (includes
+developers, testers, business analysts, and managers).
+
+\cite{Fowler2000}: Cockburn's Crystal methodology has been used by teams of up
+to 50 people, but there is little evidence as to how you can use an adaptive
+approach, or even if such approaches work at all, beyond that size.
+
+\cite{Williams2003} Agile is best suitable for teams around 50 people or fewer.
+A few projects of several hundred people have used the agile framework
+succesfully. 9 However, during 2002, agile practitioners and other researchers
+concluded that the agile value set and practices best suit colocated teams of
+about 50 people or fewer who have easy access to user and business experts and
+are developing projects that are not life-critical. Although the agile value set
+might be adopted (and may be adapted) under other circumstances, the boundary
+conditions for truly agile behavior seem to be fairly set.
+
 ----
@@
 
---> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
 --> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
@@
 
---> The company must already be with N teams or N people, not growing up to the
-    required numbers. Small companies in growth are therefore excluded.
+% --> The company must already be with N teams or N people, not growing up to the
+%     required numbers. Small companies in growth are therefore excluded.
 

### 2014-06-24 journal/article.tex
Large scale

--- 2014-06-21 journal/article.tex
@@
 Proc. Agile Development Conf. (ADC), 2004, pp. 2-10.
-
 [7] Eckstein Jutta, “Agile Software Development in the large: Diving into the deep”,
@@
 
-The following research questions were set to guide the research process.
+The following research questions were set to guide the research.
 
@@
 
-Proceeding a workshop at XP2013 \cite{Dingsoyr2013a} Dingsøyr et al.
+The definition of large scale was discussed in a workshop at XP2013
+\cite{Dingsoyr2013a}. Proceeding the workshop, Dingsøyr et al.
 compiled a description of the definition of large scale \cite{Dingsoyr2013b}.
-Dingsøyr et al. describe that at the XP2013 workshop the most common view on
-large scale was the involvment of many people \cite{Dingsoyr2013b}. Dingsøyr et
-al. conducted a brief literature search to identify what definitions on large
-scale are used. The results showed that large scale had been regadred in terms
-of size in persons or teams, project budget, code base size, and project
-duration.
-Paasivaara et al. \cite{Paasivaara2008} referred to an organization of 40 people
-and 7 teams. Berger and Beynon-Davies \cite{Berger2009} present the project cost
-of over £10 million as a factor of large scale, but also highlight a team size
-of over 50 people. A product with a code base size of over 5 million lines of
-code was considered to represent a large scale development effort by Petersen
-and Wohlin \cite{Petersen2010}. A project time of 2 years was considered as a
-factor of large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a
-project scope of 60-80 features.
-
-----
-
-Moore and Spens \cite{Moore2008} define large scale global development in the
-scope of their research as 300 people across 3 sites.
-
-----
-
---> What will bring complexity into software? Not costs or the length of the
-    schedule. --> Having many people will bring the complexity in terms of the
-    SDM to use.
-
---> [Cockburn 2003] Agile is best suitable for teams around 50 people or fewer.
---> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
-    Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
-    --> See Boehm also for list of
-
---> The company must already be with N teams or N people, not growing up to the
-    required numbers. Small companies in growth are therefore excluded.
+The most common view in the XP2013 workshop was that the involvment of many
+people is the main factor in large scale \cite{Dingsoyr2013b}.
+
+Dingsøyr et al. conducted a brief literature search to identify what definitions
+on large scale are used \cite{Dingsoyr2013b}. The results showed that large
+scale had been regadred in terms of size in persons or teams, project budget,
+code base size, and project duration. As examples from the results, Paasivaara
+et al. \cite{Paasivaara2008} referred to an organization of 40 people and 7
+teams. Berger and Beynon-Davies \cite{Berger2009} present the project cost of
+over £10 million as a factor of large scale, but also highlight a team size of
+over 50 people. A product with a code base size of over 5 million lines of code
+was considered to represent a large scale development effort by Petersen and
+Wohlin \cite{Petersen2010}. A project time of 2 years was considered as a factor
+of large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a project
+scope of 60-80 features.
+
+As a part of our research, we identified further definitions of large scale. All
+of these definitions referred to the number of people involved.
+In early work on agile Fowler \cite{Fowler2000} regards the Crystal methodology
+being suitable for up to 50 people, thus setting a boundary beyond which agile
+methods would be expected to need special considerations. Also Williams and
+Cockburn \cite{Williams2003} present 50 people as a limit which practicioners
+and researchers has seen as the largest organization suitable for agile.
+Elshamy and Elssamadisy \cite{Elshamy2006} considered a development project
+large if it had a staff between 50 and 100 people, including all project
+personnel. Finally, the largest numbers were presented by Moore and Spens
+\cite{Moore2008} who define large scale in the scope of their research as 300
+people across 3 sites.
+
+Based on these findings we defined large scale in the scope of this research to
+denote organizations with 50 or more people. These 50 people would not all need
+to be developers, but they should be part of the development organization. For
+instance, Scrum masters and software architects would be counted in. As some
+studies presented the number of teams rather than number of people, we
+correspondignly defined large scale to denote development efforts involving at
+least 6 teams. Having 6 teams with an average teams size of 6 people, plus a
+number of supporting staff, can reasonably be considered to form an organization
+of 50 people.
+
+% --> What will bring complexity into software? Not costs or the length of the
+%     schedule. --> Having many people will bring the complexity in terms of the
+%     SDM to use.
+% 
+% --> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
+%     Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
+%     --> See Boehm also for list of
+
+% --> The company must already be with N teams or N people, not growing up to the
+%     required numbers. Small companies in growth are therefore excluded.
 
@@
         Using old and new models side by side   &  P1, P8, P10, P26, P29 \\
-        Management uses waterfall model         &  P3, P6,P35,P38, P45, P46  \\
+        Management uses waterfall model         &  P3, P6, P35, P38, P45, P46 \\
         Duplicating management with two models  &  P20, P48 \\

### 2014-06-24 journal/article.tex
Spellcheck

@@
 compiled a description of the definition of large scale \cite{Dingsoyr2013b}.
-The most common view in the XP2013 workshop was that the involvment of many
+The most common view in the XP2013 workshop was that the involvement of many
 people is the main factor in large scale \cite{Dingsoyr2013b}.
@@
 on large scale are used \cite{Dingsoyr2013b}. The results showed that large
-scale had been regadred in terms of size in persons or teams, project budget,
+scale had been regarded in terms of size in persons or teams, project budget,
 code base size, and project duration. As examples from the results, Paasivaara
@@
 methods would be expected to need special considerations. Also Williams and
-Cockburn \cite{Williams2003} present 50 people as a limit which practicioners
+Cockburn \cite{Williams2003} present 50 people as a limit which practitioners
 and researchers has seen as the largest organization suitable for agile.
@@
 studies presented the number of teams rather than number of people, we
-correspondignly defined large scale to denote development efforts involving at
+correspondingly defined large scale to denote development efforts involving at
 least 6 teams. Having 6 teams with an average teams size of 6 people, plus a
@@
 extraction by one researcher (Dikert). The results of the study were finally
-elicitated by aggregating and analyzing the coding of the primary documents.
+elicited by aggregating and analyzing the coding of the primary documents.
 Figure \ref{fig:researchprocess} presents an outline of the research process.
@@
 search was to construct a search string which yielded a reasonably large result
-set. The set of matched primary studies was later ammended with a small number
+set. The set of matched primary studies was later amended with a small number
 of studies found in references of the matched papers.
@@
 
-In parallell with the full text filtering step the references of all papers were
-also examined for relevance. Most papers used references very scarcely, typicaly
+In parallel with the full text filtering step the references of all papers were
+also examined for relevance. Most papers used references very scarcely, typically
 referencing well known descriptions of agile methods. We included 2 referenced
@@
 this threat were limited, as the work stages in question are particularly
-laborous. For resource constraints the process could not be duplicated. We
+laborious. For resource constraints the process could not be duplicated. We
 assessed the bias of the coding step using an experiment, and prevented
-subjective bias in analysis by making the results as tracable as possible. The
+subjective bias in analysis by making the results as traceable as possible. The
 coding experiment revealed that two researchers can find roughly 2 of 3 themes
@@
 A particular problem in this systematic literature review was the limitations of
-boolean keyword searches in online databases. A keyword search can not identify
+Boolean keyword searches in online databases. A keyword search can not identify
 the facet of \emph{large scale}, and also \emph{transformation} is difficult to

### 2014-06-24 journal/article.tex
Clarifications

@@
 
-The definition of large scale was discussed in a workshop at XP2013
+The definition of large scale was recently discussed in a workshop at XP2013
 \cite{Dingsoyr2013a}. Proceeding the workshop, Dingsøyr et al.
@@
 denote organizations with 50 or more people. These 50 people would not all need
-to be developers, but they should be part of the development organization. For
-instance, Scrum masters and software architects would be counted in. As some
-studies presented the number of teams rather than number of people, we
-correspondingly defined large scale to denote development efforts involving at
-least 6 teams. Having 6 teams with an average teams size of 6 people, plus a
-number of supporting staff, can reasonably be considered to form an organization
-of 50 people.
+to be developers, but they should be part of a single software development
+organization. For instance, Scrum masters and software architects would be
+counted in. As some studies presented the number of teams rather than number of
+people, we correspondingly defined large scale to denote development efforts
+involving at least 6 teams. Having 6 teams with an average teams size of 6
+people, plus a number of supporting staff, can reasonably be considered to form
+an organization of 50 people.
 

### 2014-08-09 journal/content.tex
Renaming content file

--- 2014-06-24 journal/article.tex
@@
 
--- What can be learned from this`?
+-- What can be learned from this?
 
@@
    base.
---> There is interest in large scale [ref XP13 workshop]
+--> There is interest in large scale [ref XP13 workshop] \cite{Dingsoyr2013a}
 
@@
 
-% ---------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
-
-Lindvall, M., et al., "Agile Software Development in Large Organizations,"
-IEEE Computer, vol. 37, pp. 26-34, 2004.
-
 Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
@@
 pp. 64 - 69, 2002.
---> Any hints on size?? Maybe size leaning towards discipline?
-
-What is agile?
-[6] Conboy K., Fitzgerald B., “Toward a conceptual framework for agile methods:
-a study of agility in different disciplines,”
-Proc. ACM workshop on Interdisciplinary software engineering research (WISER), 2004, pp. 37-44.
-
-\subsection{Large scale general references from primary studies}
+--> Size requires more discipline?
+
+% ---------------------------------------------------------------------
+\clearpage
+
+\section{Background}
+\label{sec:background}
+
+\subsection{Agile software development}
+
+Agile software development is a collection of methods developed as an alternative
+to so called traditional development methods. From the point of view of agile
+methods traditional methods strive to minimize change during the development
+process and attempt to fully define requirements before development. However,
+adaptability during the development life cycle and the ability to respond to
+conditions unforeseen at start are critical factors for the success of a
+software development project. Embracing change during development, people
+centricity are, and pursuing high quality from the start are fundamental themes
+in agile development.
+\cite{Highsmith2001,Cockburn2001}
+
+Agile methods have been both criticized and advocated, and research has shown
+that accommodating change may be a factor in both success and failure
+\cite{Boehm2002}. It has been shown that agile methods have improved
+satisfaction of both customers and developers, but on the other hand there is
+evidence that agile methods may not suite well for large undertakings
+\cite{Dyba2009}. As a solution \citet{Boehm2002} has suggested that each
+organization should find its own balance of agile and plan driven methods.
+
+Two of the most popular agile methods at the time being are Extreme Programming
+(XP) and Scrum \cite{Hamed2013}. Scrum is a method which focuses on the project
+management viewpoint of agile development \cite{Schwaber2002}. The Scrum method
+focuses on time-boxing, continuous tracking of project progress, and customer
+centricity. A prominent feature of Scrum is dividing work into sprints, which
+are time boxes of 1 to 4 weeks, aimed for producing a finished increment of
+working software \cite{Schwaber2002}. The XP development method is a collection
+of practices for enabling efficient incremental development \cite{Beck1999}.
+Some of the key practices of XP are continuous feedback on development, test
+driven development, close customer involvement, pair programming, and continuous
+integration (CI) \cite{Beck1999}.
+
+A methodology relating to and complementing agile development is lean software
+development. Similarly to Scrum and XP, lean software development aims to
+concentrate focus on activities that provide the most value to the customer.
+Core principles of lean software development are optimizing the whole, removal
+of waste, building quality in, and continuous learning. \cite{Poppendieck2012}
+
+% What is agile?
+% [6] Conboy K., Fitzgerald B., “Toward a conceptual framework for agile methods:
+% a study of agility in different disciplines,”
+% Proc. ACM workshop on Interdisciplinary software engineering research (WISER), 2004, pp. 37-44.
+
+\subsection{Large scale agile development}
 
@@
 
-\subsection{Things that may need to be explained}
+\subsection{Adopting agile methods in large organizations}
+
+Introducing agile methods in large organizations is more difficult than it is in
+small organizations \cite{Dyba2008}. The difficulty is partly related to size
+bringing higher organizational inertia which slows down organizational change
+\cite{Livermore2008}. Agile development is not founded on the use of individual
+tools or practices, but rather on a holistic way of thinking. Adopting agile
+often requires change of the entire organizational culture \cite{Misra2010}.
+
+One significant difference between small and large scale adoptions is that large
+organizations have more dependencies between projects and teams. This increases
+the need for formal documentation and thus reduces agility \cite{Lindvall2004}.
+In addition to inter-team coordination development teams must interact with
+other organizational units, which are often non-agile in nature. For instance,
+HR may demand individuals to have strictly specified roles in projects
+\cite{Boehm2005}, or a change control board may inhibit the use of continuous
+integration or refactoring \cite{Lindvall2004}. All units affected by the agile
+transformation need to be informed and consulted, and the agile process must be
+adjusted according to their needs \cite{Lindvall2004, Cohn2003, Boehm2005}.
+
+Agile methods will also affect management and business related functions. A key
+challenge is that management must move away from life-cycle models and towards
+iterative and feature centric models \cite{Nerur2005}. The focus must be shifted
+from a large scale scope to shorter term project planning \cite{Misra2010}, as
+agile methods emphasize that planning is only meaningful for the near future
+\cite{Cohn2003}. However, the lack of planning can be a concern as business and
+customer relationships often build on long term roadmapping. Enabling operation
+with shorter term planning requires educating stakeholders and reviewing
+contracting practices \cite{Boehm2005}.
+
+% \subsection{Challenges in adopting agile methods}
+%
+
+% \subsection{Success factors in adopting agile methods}
+%
+
+\subsection{Summary of things that may need to be explained}
 
@@
 % ---------------------------------------------------------------------
+\clearpage
+
 \section{Research method}
@@
         Large scale    &
-                At least 50 development personnel or 6 teams uses agile &
+                At least 50 development personnel or 6 teams use agile &
 
@@
 
-TODO:
-* Mieti rakenne
-* Mieti onko seuraava ``other than development'' hyvä otsikko
+% TODO:
+% * Mieti rakenne
+% * Mieti onko seuraava ``other than development'' hyvä otsikko
 

### 2014-08-09 journal/article.tex
Commented out TODO
Fixes
Preambles for article

--- 2014-06-24 journal/article.tex
@@
 
-\section{Introduction}
-
---> Introductory paragraph
-
--- Who will be interested of this?
-
--- What can be learned from this`?
-
---> Motivation for research
--- There are no systematic reviews for this
--- A collection of experience reports can be used for creating a theoretical
-   base.
---> There is interest in large scale [ref XP13 workshop]
-
---> Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
-    Practitioners," IEEE Software, pp. 8-9, 2010.
-
---> Why is large-scale software development different from smaller scale?
-
---> This literature review is intended for researchers as a theoretical
-   and proven foundation for the real problems that organizations face in
-   transformations.
-   This is also intended to be directly usable by practitioners, to serve as a
-   base of general advice with strong motivation (although every organization
-   has unique needs).
-
-Dingsoyr:
-The nature of software development, being innovative work that is only partially
-compatible with programmed coordination, requires a strong emphasis on personal
-communication [Kraut, R. E. and Streeter, L. A., "Coordination in software
-development," Communications of the ACM, vol. 38, pp. 69-81, 1995]
-
-% ---------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
-
-Lindvall, M., et al., "Agile Software Development in Large Organizations,"
-IEEE Computer, vol. 37, pp. 26-34, 2004.
-
-Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
-Perplexed: Addison-Wesley, 2003.
-Boehm, B., "Get ready for agile methods, with care," IEEE Computer, vol. 35,
-pp. 64 - 69, 2002.
---> Any hints on size?? Maybe size leaning towards discipline?
-
-What is agile?
-[6] Conboy K., Fitzgerald B., “Toward a conceptual framework for agile methods:
-a study of agility in different disciplines,”
-Proc. ACM workshop on Interdisciplinary software engineering research (WISER), 2004, pp. 37-44.
-
-\subsection{Large scale general references from primary studies}
-
-Laanti 2008:
-[2] Kähkönen Tuomo, “Agile Methods for large organizations – building communities of practice”.
-Proc. Agile Development Conf. (ADC), 2004, pp. 2-10.
-[7] Eckstein Jutta, “Agile Software Development in the large: Diving into the deep”,
-Dorset House Publishing, 2004. ISBN 0-932633-57-9.
-[8] Leffingwell Dean, “Scaling Software Agility: Best Practices for Large Enterprises”,
-Addison-Wesley Professional, 2007. ISBN-10: 0321458192.
-[12] Highsmith J., Agile for the Enterprise, “From Agile Teams to Agile Organizations”,
-Cutter, 2005. http://www.cutter.com/program/ fulltext/reports/2005/01/index.html .
-[13] Schwaber Ken, “Enterprise and Scrum”, Microsoft Press 2007, ISBN-10: 0735623376.
-
-Vlaanderen 2012:
-22. Sidky, A.S.: A Structured Approach to Adopting Agile Practices: The Agile Adop-
-tion Framework (June 2007)
-
-Cockburn 2003:
-R. Crocker, Large-Scale Agile Software Development, Addison-Wesley, 2003.
-
-\subsection{Things that may need to be explained}
-
-* Scrum
-
-* XP
-
-* Definition of sprint
-
-* Definition of product owner
-
-* Self organization and independence of teams -- This is a question relating how
-  to organize management.
-
-* Scrum master must not command and control -- Scrum master as manager??
-
-* People centricity -- agile is not for squeezing efficiency -- agile is not for
-  fewer career opportunities in a flatter organization.
-
-* Team centricity -- agile focuses on team rather than individuals
-
-* Agile focuses on short time horizons ???
-
-* Agile is a mindset -- it can not be studied from a book -- need of coaching
-
-* What are the agile principles? What is being talked about when ``agile
-  principles'' is referred to?
-
-* What is coaching? Why is it done?
-
-* Principle of build often? Principle of nightly builds or CI? Is this from XP?
-  Is this an agile principle?
-
-* Kotter: Having a ``guiding team'' for change
-
-% ---------------------------------------------------------------------
-\section{Research method}
-\label{sec:method}
-
-The goal of the research was to aggregate evidence on how large organizations
-transform into an agile way of working. The research was conducted as a
-systematic literature review based on empirical studies.
-
-The research process is presented starting from the research questions. After
-that we present the definition of large scale which this research builds upon.
-The bulk of this section concentrates on presenting the steps of the research
-process, starting from the definition of inclusion criteria, continuing with
-describing how the primary studies were selected, and finally presenting the
-method for analyzing and synthesizing the primary studies. The section concludes
-with a discussion on limitations of the method.
-
-\subsection{Research questions}
-
-The following research questions were set to guide the research.
-
-\begin{itemize}
-
-\item
-RQ1: Why do large software development organizations initiate an agile
-transformation?
-
-\item
-RQ2: How do large-scale agile transformations usually proceed?
-
-\item
-RQ3: What are typical challenges in a large-scale agile transformation?
-
-\item
-RQ4: What factors facilitate adopting of agile methods?
-
-\end{itemize}
-
-\subsection{Definition of large scale}
-\label{sec:largescale}
-
-A key question in this study was the definition of \emph{large scale}. Our goal
-was to identify phenomena that appear particularly in large-scale
-transformations, and which may not be present in small organizations. Large
-scale requires additional coordination and makes communication difficult
-compared to software development in small teams. In this section we present the
-definition of large scale that is used in this study.
-
-The definition of large scale was discussed in a workshop at XP2013
-\cite{Dingsoyr2013a}. Proceeding the workshop, Dingsøyr et al.
-compiled a description of the definition of large scale \cite{Dingsoyr2013b}.
-The most common view in the XP2013 workshop was that the involvement of many
-people is the main factor in large scale \cite{Dingsoyr2013b}.
-
-Dingsøyr et al. conducted a brief literature search to identify what definitions
-on large scale are used \cite{Dingsoyr2013b}. The results showed that large
-scale had been regarded in terms of size in persons or teams, project budget,
-code base size, and project duration. As examples from the results, Paasivaara
-et al. \cite{Paasivaara2008} referred to an organization of 40 people and 7
-teams. Berger and Beynon-Davies \cite{Berger2009} present the project cost of
-over £10 million as a factor of large scale, but also highlight a team size of
-over 50 people. A product with a code base size of over 5 million lines of code
-was considered to represent a large scale development effort by Petersen and
-Wohlin \cite{Petersen2010}. A project time of 2 years was considered as a factor
-of large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a project
-scope of 60-80 features.
-
-As a part of our research, we identified further definitions of large scale. All
-of these definitions referred to the number of people involved.
-In early work on agile Fowler \cite{Fowler2000} regards the Crystal methodology
-being suitable for up to 50 people, thus setting a boundary beyond which agile
-methods would be expected to need special considerations. Also Williams and
-Cockburn \cite{Williams2003} present 50 people as a limit which practitioners
-and researchers has seen as the largest organization suitable for agile.
-Elshamy and Elssamadisy \cite{Elshamy2006} considered a development project
-large if it had a staff between 50 and 100 people, including all project
-personnel. Finally, the largest numbers were presented by Moore and Spens
-\cite{Moore2008} who define large scale in the scope of their research as 300
-people across 3 sites.
-
-Based on these findings we defined large scale in the scope of this research to
-denote organizations with 50 or more people. These 50 people would not all need
-to be developers, but they should be part of the development organization. For
-instance, Scrum masters and software architects would be counted in. As some
-studies presented the number of teams rather than number of people, we
-correspondingly defined large scale to denote development efforts involving at
-least 6 teams. Having 6 teams with an average teams size of 6 people, plus a
-number of supporting staff, can reasonably be considered to form an organization
-of 50 people.
-
-% --> What will bring complexity into software? Not costs or the length of the
-%     schedule. --> Having many people will bring the complexity in terms of the
-%     SDM to use.
-% 
-% --> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
-%     Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
-%     --> See Boehm also for list of
-
-% --> The company must already be with N teams or N people, not growing up to the
-%     required numbers. Small companies in growth are therefore excluded.
-
-\subsection{Research process}
-
-The research was conducted as an application of Kitchenham's
-\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
-primary studies was done using first a keyword based database search to identify
-potentially relevant sources, and then manually filtering the search result. The
-filtering process was executed independently by two researchers (Dikert,
-Paasivaara). The selected primary studies were coded for qualitative data
-extraction by one researcher (Dikert). The results of the study were finally
-elicited by aggregating and analyzing the coding of the primary documents.
-Figure \ref{fig:researchprocess} presents an outline of the research process.
-The entire process was audited and mentored by a third researcher (Lassenius).
-
-\begin{figure}[b]
-  \begin{center}
-    \includegraphics{graphics/research_process.pdf}
-    \caption{Outline of the research process}
-    \label{fig:researchprocess}
-  \end{center}
-\end{figure}
-
-We deviated from Kitchenham's guideline on the part of data extraction. Instead
-of using data extraction forms we analyzed the primary studies by coding. This
-deviation was made in order to minimize how possible preconceptions would affect
-the data extraction step, and to use a specialized software tool that allowed us
-to record evidence in as much detail as possible.
-% --> Possibility to find many different viewpoinits for the research questions
-% --> Lots of items, which would have lead to much diversity in data extraction form
-
-\subsubsection{Inclusion criteria}
-\label{sec:inclusioncriteria}
-
-Based on the research questions and intended focus of the research, we defined
-four facets to guide inclusion/exclusion decisions: agile software engineering,
-focus on transformation, large scale, and empiricality. The first facet covers
-primary studies focusing on software engineering organizations applying or
-striving to apply agile methods. The second facet states that included studies
-must provide insights relevant to organizational transformation, and
-specifically to the research questions. The third facet underlines the
-particular contribution we aim to provide with this research, namely the
-exclusive focus on large organizations. The fourth facet limits inclusion to
-studies presenting real world cases. These four facets were used for inclusion
-and exclusion in each study selection step, including search string design,
-filtering by abstracts, and full text filtering.
-
-Table \ref{table:facets} lists the facets and gives examples on matching topics
-and not relevant topics. For a study to be included it needed to be relevant
-in all facets. Even if a study discussed some of the irrelevant topics it could
-be included as long as it had some content matching all the facets.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.20\textwidth}
-                     >{\raggedright\arraybackslash}p{0.34\textwidth}
-                     >{\raggedright\arraybackslash}p{0.33\textwidth} }
-        \toprule
-        Facet  &  Relevant topics  &  Not relevant topics  \\
-        \midrule
-        Agile software development  & 
-                The organization develops software;
-                The development method introduced is agile  &
-
-                Agile manufacturing;
-                Scrum in management boards  \\
-
-        Organizational transformation  &
-                Presenting insights in the process of transformation  &
-
-                Comparison of before and after;
-                How agile is being used in large scale   \\
-
-        Large scale    &
-                At least 50 development personnel or 6 teams uses agile &
-
-                Scaling up from small;
-                A single agile team in a large setting \\
-
-        Empiricality   &
-                Case studies, experience reports  &
-
-                Textbooks, student experiments, theory papers  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Facets of inclusion criteria with examples of relevant and
-             not relevant topics}
-    \label{table:facets}
-\end{table}
-
-Examples on topics excluded by the facet of \emph{agile software development}
-are agile manufacturing and applying Scrum practices in management boards, as
-these do not relate to software engineering. In addition, we required that the
-organizational transformation was aimed to introduce agile methods, which
-excluded other development methods than agile \cite{Sagesser2013}, and use of
-agile methods in other contexts than software development \cite{Hodgkins2007}.
-
-The facet of \emph{organizational transformation} was interpreted so that the
-primary study must present insights on the process of transformation. This
-was the most important facet, as the pivot of this research is to examine
-how the transformations happen. Examples on excluded topics are comparing the
-original and agile development methods \cite{Petersen2010}, discussing use of
-agile in a large organization but not describing how the new methods were
-introduced \cite{Mishra2011}, and merely presenting agile tools in large-scale
-use \cite{Kim2012}. These sources were excluded as they did not provide insights
-on organizational transformation.
-% Also [Fitzgerald 2005] on just presenting use of agile methods, no introducing
-
-Transformation and ``scaling up'' of agile practices in use are very closely
-related concepts, and in some cases they are one and the same. For instance, if
-transformation begins with a pilot and spreads gradually through the
-organization, the process can very well be seen as a ``scaling up'' journey.
-However, we excluded cases where an initially small organization was scaled up
-\cite{Maranzato2012}, and discussions focusing on processes or tools without
-describing organizational change \cite{Lyon2008}.
-% Also [Greening 2010] on process centered viewpoint without transformation
-
-The facet of \emph{large scale} was interpreted as presented in section
-\ref{sec:largescale}, having 50 development personnel or 6 teams as as the limit
-for large scale. In some cases the source presented very vague indicators on
-size. For instance, we deemed the case presented by Cloke \cite{Cloke2007} to be
-included as there was indications of large-scale considerations, although the
-size remained unclear. The case presented by Miller and Haddad \cite{Miller2012}
-was deemed to be excluded as there was no indications whatsoever on size. A
-borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
-no indication of issues indicating large size, although there were several
-development teams.
-To complicate matters, some papers talked about ``the team'' in singular when
-referring to the organization \cite{Hodgkins2007}, making it clearly nontrivial
-to judge whether an organization was large based on the choice of words of the
-author.
-
-Further examples on exclusion by the \emph{large scale} facet were cases with
-large organizations but only a single team adopting agile \cite{Fulgham2011}.
-We considered cases of single teams (although in large organizations) unrelated
-to this research as our focus is on transformation of the entire organization.
-Also piloting cases that resulted only in single teams using agile were excluded
-\cite{Scott2008}. Finally, cases where the organization was growing to large
-scale, but did not meet the size criteria at the start of the transformation,
-were excluded.
-
-% Also [Jensen 2003] with only a single team
-% Possibly [Mc Cormic 2005] on single team piloting
-
-According to the facet of \emph{empiricality} we excluded studies that did not
-discuss a distinguishable real world case. We excluded textbooks, studies merely
-presenting theories, and other studies that did not include any case
-organization. We also excluded studies of benefits or limitations of agile in
-general. Lastly, we excluded student experiments as it is implausible to
-simulate the dynamics of a large organization in an experiment.
-
-\subsubsection{Preliminary searches}
-
-Before proceeding with identifying the primary studies a number of preliminary
-searches were performed. The purpose of the preliminary searches was to create a
-benchmark of potentially relevant papers that should be matched by the actual
-search, and to evaluate different search strings. We started by examining top
-ranked hits by trivial keywords that the more complex final search string might
-miss. Initial searches were made using keywords that were as general as
-possible, including ``agile transformation'' and ``large scale agile''.
-Secondly, a trial keyword search was done in the selected databases (listed in
-table \ref{table:databases}) using the following search string:
-
-\begin{verbatim}
-((agil* OR scrum OR xp OR lean) AND
-(transform* OR transit* OR change OR migrat*) AND
-((large AND (scale OR organization))) OR enterprise)
-\end{verbatim}
-
-From these preliminary searches we compiled a list of 117 possibly relevant
-papers that the actual database search would be benchmarked against.
-
-\subsubsection{Identification of primary studies}
-
-The gathering of potential primary studies was based on a search in online
-databases. The databases searched are listed in Table \ref{table:databases}. All
-databases supported use of complex Boolean logic in searches, which allowed us
-flexibility in constructing the search strings. The main task in the database
-search was to construct a search string which yielded a reasonably large result
-set. The set of matched primary studies was later amended with a small number
-of studies found in references of the matched papers.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ l l l }
-        \toprule
-        Database         & URL                         & N of matches   \\
-        \midrule
-        IEEExplore       & http://ieeexplore.ieee.org      & 745 \\ 
-        ACM              & http://dl.acm.org               & 168 \\
-        Scopus           & http://www.scopus.com/home.url  & 1596 \\
-        Web of Knowledge & http://apps.webofknowledge.com  & 786 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Databases included in search, and number of matched articles}
-    \label{table:databases}
-\end{table}
-
-We proceeded with constructing a search string based on the facets presented in
-section \ref{sec:inclusioncriteria}. However, preliminary searches had showed
-that picking keywords with good precision was difficult. Especially the facet
-\emph{large scale} was difficult to represent with precise words. Also the facet
-of \emph{empiricality} could not sensibly be represented by keywords. We decided
-to include only the facets \emph{agile software development} and
-\emph{organizational transformation} in the keyword search, with the consequence
-of increasing the manual filtering effort in the subsequent steps.
-
-Having a substantial part of the matches in non-relevant areas such as agile
-manufacturing, we included only articles including the term ``software'' or
-articles published in relevant conferences. Instead of engineering the search
-terms any further, which proved to result in excluding some relevant articles,
-we chose to rely more on manual filtering. The resulting facets and keywords are
-listed in Table \ref{table:searchterms}.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.22\textwidth}
-                     >{\raggedright\arraybackslash}p{0.70\textwidth} }
-        \toprule
-        Facet                  & Keywords   \\
-        \midrule
-        Agile methods (before \& after) &
-            \texttt{agile, scrum, "extreme programming",
-            waterfall, "plan-driven", RUP}\\
-        Organizational transformation &
-            \texttt{transform*, transiti*, migrat*, journey, adopt*, deploy, introduc*,
-            "roll-out", rollout}\\
-        Only software related articles &
-            \texttt{(software OR (conference="agile, xp, icgse, icse"))
-            AND NOT (title+abs="manufacturing" OR conference="agile manufacturing")}
-        \\
-        \bottomrule
-    \end{tabular}
-    \caption{Facets and related search terms used}
-    \label{table:searchterms}
-\end{table}
-
-The database keyword search matched 1875 unique papers. The results of the
-search were gathered for the selection step using reference management software.
-
-\subsubsection{Study selection}
-
-The set of potential primary studies identified by the database search was
-refined in two steps, by first filtering out abstracts and finally full text
-filtering. The study selection process starting from the database search is
-outlined in figure \ref{fig:selectionprocess_full}. The result of the database
-search was amended with some results from the preliminary search and relevant
-references from all examined papers.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
-    \caption{The study selection process}
-    \label{fig:selectionprocess_full}
-  \end{center}
-\end{figure}
-
-The database keyword search had matched 1875 unique papers. The abstracts of
-these papers were categorized independently by both Dikert and Paasivaara into
-three categories: include, exclude, and uncertain. There were 1578 exclusions
-and 62 inclusions that both researchers agreed on. The inclusion decisions
-for the 235 abstracts with disagreement or uncertainty were resolved through
-discussion. At this stage papers were excluded only if both researchers deemed
-it clearly irrelevant, including any uncertain cases for full text filtering. As
-a result 170 papers were selected for full text filtering.
-
-% Inter-rater statistics
-%
-%                     Maria
-%            excl   ???   incl    TOT
-%      excl  1578    18    15    1611
-% Kimi ???     84    17     8     109
-%      incl    51    42    62     155
-%      TOT   1713    77    85    1875
-%
-% ???'s =  169
-% in/ex = 1706
-%
-% Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
-% Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
-% Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
-% Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
-%
-% Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
-%
-% Pr(a) = (ex-ex + in-in) / tot = 0.8831
-%
-% k = 0.4063
-
-Full text filtering was performed by evaluating each article against the four
-facets of the inclusion criteria. Filtering was done in two steps. In the first
-step Dikert extracted data relevant to the four facets. Based on the extracted
-data 76 papers could be immediately deemed as included or excluded. The
-remaining 94 papers were evaluated against the inclusion criteria by both Dikert
-and Paasivaara, and a decision was made after discussing each paper separately.
-In difficult cases Lassenius was consulted to reach a decision. As a result of
-the full text filtering 47 papers were selected to be included.
-
-We evaluated the result of the database keyword search against the benchmark
-created in the preliminary search step. We concluded that 75 of the 117
-preliminarily selected papers were matched by the database keyword search.
-The missed 32 preliminary papers were examined, resulting in including 3
-additional papers as primary sources.
-
-In parallel with the full text filtering step the references of all papers were
-also examined for relevance. Most papers used references very scarcely, typically
-referencing well known descriptions of agile methods. We included 2 referenced
-papers in full text analysis.
-
-Finally, counting 47 papers from the full text filtering step, 3 papers from
-preliminary searches, and 2 papers from references, 52 papers were selected
-as primary studies.
-
-\subsubsection{Handling of duplicate reports on a single case}
-
-In several cases there were more than one primary study presenting the
-transformation of the same organization. Duplicate descriptions of the same
-organization focused typically on different aspects. For example, one paper
-would highlight the viewpoint of developers \cite{Fry2007}, and another would
-consider the transformation from user experience designers' point of view
-\cite{Federoff2009}.
-
-Even if the transformation of a single organization was described in many
-studies, all sources that passed the inclusion criteria were included. Studies
-presenting the same organization were treated as one unit so that we could gain
-as much insight on each organization as possible. Conversely, there were also a
-few papers that presented multiple case studies, and in those cases we treated
-each studied organization individually.
-
-As a result 42 unique organizations were identified in the primary studies. We
-use the term \emph{study} to refer to the primary study publications, and the
-term \emph{case} to refer to an individual case organization that may be
-described in several different studies.
-
-\subsubsection{Study quality assessment}
-
-The primary research for this literature review consists almost exclusively of
-industry experience reports. There were only 6 case studies with a research
-method, and observations on transformation were presented only as a minor part
-in these studies. Based on this finding we concluded that case studies
-presenting viewpoints on organizational transformations are very scarce in
-software industry. We deemed that the results would be distorted heavily and
-many valuable studies would be left out if a strict quality assessment would be
-part of the inclusion criteria. As a result we decided to include all experience
-reports, regardless of the perceived objectivity.
-
-The absence of research methods and subjective viewpoints of experience reports
-can be seen as a factor lessening the evidence. Because of this we were
-refrained from making quantitative interpretations of the data, but based the
-results around describing qualitative observations. Regardless of the bias of
-the primary studies, we think that the concepts that emerged from the source
-material should not be questioned on a fundamental level. The overall existence
-of organizational phenomena should not be questioned based on possible author
-bias in experience reports.
-
-% --> Large organizational changes are hard to measure experimentally (controlled
-%     experiments?). Also due to this any student research was excluded.
-
-% --> It may still be interesting to make some quantitative observations such as
-%     a particular problem being reported in the majority of reports.
-
-% * The primary studies did not make distinction between method, results, and
-%   discussion
-% 
-% * The studies neither did any implications for generalizability, but rather
-%   presented narratives or experiences. Some cases presented lists of
-% %   recommendations.
-
-As presented in the previous section, we allowed several descriptions of the
-same case to be included. Instead of using the most complete paper as suggested
-by Kitchenham \cite{Kitchenham2007}, we combined the results presented in each
-paper and considered the case as a single unit. Keeping in mind the potential
-bias caused by duplicate publications, including all papers enabled us to have a
-more detailed view of the case organization.
-
-\subsubsection{Coding of primary studies}
-\label{sec:coding}
-
-The primary studies were coded using the Atlas TI qualitative data analysis software.
-An integrated deductive and inductive approach was chosen for coding the primary
-studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
-and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
-list of codes was established for contextual information, as presented in Table
-\ref{table:contextualcodes}. Codes related to the research questions were chosen
-to be created by an inductive process. The reason for inducting codes rather
-than using an a priori approach was to avoid the researchers' previous
-assumptions of the research area to affect the choice of codes
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.26\textwidth}
-                     >{\raggedright\arraybackslash}p{0.66\textwidth} }
-        \toprule
-        Contextual code     & Explanation   \\
-        \midrule
-
-Agile method & Agile methods used in the organization (eg. Scrum, XP) \\
-
-Business area & The business area in which the organization operates. \\
-
-Organization size & Mentioning the size of the case organization. \\
-
-Time of transformation & When the transformation has been in
-progress, or how long the transformation has taken. Possibly relative to
-the time of the research. \\
-
-Research process & The paper describes a research process. \\
-
-Geographical location & Where the organization is located. \\
-
-Large scale definition & A definition of large-scale software development. \\
-
-Multisite / GSD & Mentioning of a multisite organization. \\
-
-        \bottomrule
-    \end{tabular}
-    \caption{Context information}
-    \label{table:contextualcodes}
-\end{table}
-
-To give direction for code induction seven different code families were
-established on beforehand, as presented in Table \ref{table:codefamilies}. A
-description defining scope and examples for codes were defined for each family.
-The example codes would not necessarily be used as actual codes, and would not
-represent the entire or final scope of the families.
-
-\begin{table}[t]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
-                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
-        \toprule
-        Code family          &  Description
-        \\
-        \midrule
-
-        RQ1: Reason to change &
-        Reasons to start the transformation (demand for faster delivery). \\
-
-		RQ2: Transformation process &
-		Statements that describe the transformation process (top-down, big bang,
-		step wise). \\
-
-		RQ3: Challenges &
-		Statements that present challenges in the transformation (change resistance).
-		\\
-
-		RQ3: Success factors &
-		Statements that present success factors in the transformation (management
-		support).
-		\\
-
-		Investing in change  &
-		Factors that present how the organization is investing in the
-		transformation (training, consultants, tools). \\
-
-		Practices &
-		Distinguishable practices that are used or established during transformation.
-		Factors that are presented as neutral in terms of successes and challenges
-		in the primary studies.
-		(coaching, piloting, continuous integration, communities of practice) \\
-
-		Contextual &
-		Contextual codes defined in Table \ref{table:contextualcodes}. \\
-
-        \bottomrule
-    \end{tabular}
-    \caption{Code families}
-    \label{table:codefamilies}
-\end{table}
-
-The inductive coding was done according to the guidelines described below.
-
-A passage of text that presents any concept relating to the research questions
-or the contextual codes becomes a quotation. If the quotation corresponds
-closely to a concept that has already been coded the existing code is used.
-Otherwise a new code is created, and the code is assigned to one of the
-predefined code families. The labels of existing codes may need to be adjusted
-when new quotations are assigned. A clarifying description is written for every
-code, so that the precise meaning of the codes can be reviewed. Many codes may
-be assigned to one quotation and quotations of different codes may be
-overlapping.
-
-The length of the text passage that make up a quotation may vary. For the
-contextual codes the quoted passage should be of minimal length. For other
-quotes the length of the quotation should be long enough to highlight the
-context of the concept the quotation, varying from a few sentences to one
-paragraph, or even several paragraphs. As an example of context of the
-quotation, if a transformation challenge is described by a quotation it would be
-good if also the source or resolution of the challenge could be included in the
-quoted text passage. Having enough context should broaden the understanding
-of the coded concept when looking at the quotation in isolation.
-
-% -- Considering that mentioning the inclusion of a factor as a success, then
-%    would the absence of the same factor be a challenge?
-
-The total number of codes and quotations created in the coding process is
-presented in Table \ref{table:codecount}. The total number of quotations is less
-than the sum of quotations in the categories because a single quotation may have
-multiple codes.
-
-\begin{table}[t]
-    \centering
-    \begin{tabular}{ l r r }
-        \toprule
-        Code family    &  Codes  &  Quotations
-        \\
-        \midrule
-        RQ1: Reason to change &        30 &  123 \\
-		RQ2: Transformation process &  16 &  580 \\
-		RQ3: Challenges &              40 &  323 \\
-		RQ3: Success factors &         44 &  260 \\
-		Investing in change  &          5 &  137 \\
-		Practices &                    11 &  170 \\
-		Contextual &                    8 &  215 \\
-		Total &                       154 & 1575 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Number of codes and quotations per code family}
-    \label{table:codecount}
-\end{table}
-
-\subsubsection{Validation of coding}
-
-Coding was done by only one researcher due to resource constraints and volume of
-source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
-an experiment with the goal to assess the consistency of coding that two
-independent researchers may achieve.
-In the experiment Dikert and Paasivaara coded independently a sample of five
-primary studies. The sample was hand picked as a representative subset of the
-source material. The coding was done relying solely on the guidelines presented
-in the previous chapter, and the researchers were careful not to discuss the
-choice of codes or interpretations relating to codes before the experiment was
-completed. Finally the independent codings of the two researchers were analyzed
-with two metrics: by the number of matching themes found, and by the amount of
-overlap in quotations similar codes have.
-% See Kitcenham, chapter 6.4
-
-The metric of matching themes measures how well the independent researchers are
-able to induct similar codes from the source material. As codes are inducted it
-is likely that two researchers label similar quotes differently. To overcome
-this problem we defined higher level themes under which the individual codes
-were organized. Through discussion we made decisions on whether differently
-labeled codes could be organized under the same theme or not. Table
-\ref{table:codingexperiment} presents the total number of themes the researchers
-found in each document, and how many of the themes were matching. We concluded
-that on average two of three themes found by one researcher were similarly found
-by the other researcher, but one of three themes was not found by the other
-researcher.
-
-The metric of quotation overlap measures how well the quotations for similar
-codes match between two independent researchers. The intent is to measure how
-well codes that are deemed similar by their names match in the actual quoted
-text. We assigned a quotation overlap score for each code in themes that both
-researchers had found. A score of 3 was given for matching codes whose quotation
-content was virtually the same for both researchers. A score of 2 was given if
-the quotations were mostly the same, but other researcher had included one text
-passage that the other had missed. A score of 1 was given if roughly half of the
-quotation was the same. If the quotations had less similarity than this a score
-of 0 was given. One code provided typically 1 to 3 quotations per document. The
-quotation overlap score was counted including only codes that had matching
-themes. We concluded that the average quotation overlap score was 2.2.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ l r r r }
-        \toprule
-        Document    &  Matching themes  &  Total themes  &  Quotation overlap (0-3) \\
-        \midrule
-        P1          &  14  &  19   &  2.1  \\
-        P2          &   6  &  11   &  2.4  \\
-        P3          &   7  &  11   &  2.1  \\
-        P4          &   2  &   4   &  2.3  \\
-        P5          &  12  &  16   &  2.0  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Result of independent coding experiment}
-    \label{table:codingexperiment}
-\end{table}
-
-Combining the quotation overlap result and the result that inductive coding
-yielded similar themes with a factor of two thirds, we conclude that two
-different researchers should agree completely on at least half of the quotations
-in the source material of our study. This agreement includes both that the
-researchers select the same text passage as a quotation, and that researchers
-agree with the meaning of the quotation. Further interpretation of the
-quotations and codes will still remain a subjective matter.
-
-\subsubsection{Synthesis of primary studies based on coding}
-
-The synthesis of findings was created based on the result of coding presented in
-section \ref{sec:coding}. An initial organization of codes into categories was
-made based on the labels of the codes. Each codes was linked to a single
-research question, so the categorization of codes was made separately within
-the topic of each research question. The categories of codes were labeled
-appropriately to give a high level description of the linked codes. Some codes
-having minor importance and few quotations were left out.
-
-After assigning each code to a unique category, the content of the categories
-was studied. Each category was studied by reading through each quotation of each
-code included. Typically the quotations were displayed and examined in their
-original context, considering also surrounding paragraphs of the quotation.
-Notes were made on each quotation presenting noteworthy observations. Based on
-the notes, an accurate description emerged for each code.
-
-As an accurate description had been created for each code, the initially defined
-categories could be refined. Some codes were recategorized, and the definitions
-of a few categories were revisited, until a final ordering was reached. The
-results are presented according to the final categorization.
-
-% * If a code had more quotations there were more observations made (longer
-%   result text)
-
-\subsection{Threats to validity and limitations}
-
-To assess the validity of this research internal validity and external validity
-needs to be considered. In this section we present the identified threats to
-validity and the strategies used to mitigate them. We also highlight two
-specific limitations in the research: weakness in keyword search, and the
-strength of evidence of primary sources.
-
-The threat to internal validity is subjective bias, which affects two parts of
-the research. First, the selection of the primary studies may have been
-distorted by interpreting the inclusion criteria falsely. This risk was
-mitigated at start by using three researchers in designing the inclusion
-criteria. When the inclusion criteria was subsequently applied the initial
-filtering was performed independently by two researchers, and unclear cases were
-resolved by case-by-case discussions between two or three researchers.
-
-The second part of the research that may have been affected by subjective bias
-was the elicitation of results by coding and analysis. Our tools for mitigating
-this threat were limited, as the work stages in question are particularly
-laborious. For resource constraints the process could not be duplicated. We
-assessed the bias of the coding step using an experiment, and prevented
-subjective bias in analysis by making the results as traceable as possible. The
-coding experiment revealed that two researchers can find roughly 2 of 3 themes
-in an experience report identically. Traceability of results was achieved by
-supplying references to each claim presented in the results.
-
-This study had a minor threat to external validity. The systematic literature
-review approach includes steps to identify all studies that may be relevant in
-the study area. As long as the systematic approach is adhered to, including
-systematic identification of research, threats against external validity are
-minimized. External validity was also confirmed by saturation of the primary
-studies. Saturation was evident, as when the majority of the primary studies had
-been coded almost all quotations for the remaining studies could be assigned to
-existing codes.
-
-A particular problem in this systematic literature review was the limitations of
-Boolean keyword searches in online databases. A keyword search can not identify
-the facet of \emph{large scale}, and also \emph{transformation} is difficult to
-capture with keywords. Because of this we assume that it is likely that a some
-studies have been missed by the keyword search, potentially causing bias. To
-eliminate the potential bias we would suggest manually searching the entire
-archive of \emph{Agile Conference} and \emph{Agile Processes in Software
-Engineering and Extreme Programming} for relevant studies. These two conferences
-provided more than two thirds of the primary sources.
-
-Another particular problem was that few of the primary studies presented a
-research method. Most of the studies were experience reports where the author
-was a member of the organization discussed. Because of this we must assume that
-the primary studies are biased to some extent. Due to this risk of bias we chose
-not to make quantitative interpretations in the results, and use only
-qualitative analysis. We conclude that in the current state of research it is
-necessary to include studies with varying strength of evidence in order to
-relevantly aggregate evidence on large scale transformations.
-
-% ---------------------------------------------------------------------
-\clearpage
-
-\section{Results}
-\label{sec:results}
-
-This section presents the findings of the literature review. In the first
-subsection a meta analysis of the analyzed papers is presented. In the
-subsequent sections findings are presented organized according to the research
-questions.
-
-The methods used for extracting the results were described in section
-\ref{sec:method}. Note that the term \emph{study} is used when referring to a
-primary study publication, and the term \emph{case} is used when referring to an
-organization which may be described in several studies.
-
-\subsection{Overview of primary studies}
-
-The results of this study include findings from 52 primary studies presenting
-how 42 different large software organizations introduce agile methods. Most of
-the included studies were experience reports (45 studies), and in many cases it
-was evident that the author had been personally involved in the transformation.
-Only 6 of the included studies had a research method explicitly stated. One of
-the studies was an interview article. The publication forums of the primary
-studies were distributed so that 57 sources were conference proceedings, 4
-sources were journal articles, and one source was a technical report.
-
-Figure \ref{fig:transformation_time} summarizes the publication years of the
-primary studies, and the years when the transformations were reported to have
-started. All studies and transformations were dated after year 2000. There were
-peaks in transformation studies in years 2008 and 2011. The studies were
-typically published two years after the start of the transformation.
-
-\begin{figure}[b]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/transformation_time.pdf}
-    \caption{Publication years and reported start years of transformations}
-    \label{fig:transformation_time}
-  \end{center}
-\end{figure}
-
-The size of the organizations varied from the minimum included size of 50 to
-18,000 people. The median size was 300 people. In 7 studies the size was
-presented in terms of teams, ranging from the minimum of 6 teams to 150 teams.
-The median was 10 teams. In 8 studies there was no direct indication on size
-but the issues discussed revealed that the organization was of large size
-according to our definition. The distribution of organization sizes is presented
-in figure \ref{fig:organization_size}.
-
-\begin{figure}[!t]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/organization_size.pdf}
-    \caption{Distribution of organization sizes in primary studies}
-    \label{fig:organization_size}
-  \end{center}
-\end{figure}
-
-\begin{table}[!t]
-    \centering
-    \begin{tabular}{ l r }
-        \toprule
-        Business area    &  Case organizations   \\
-        \midrule
-		Online services for consumers and business  &  9  \\
-		Telecommunications                          &  7  \\
-		Enterprise management                       &  5  \\
-		Banking and financial services              &  4  \\
-		Health care                                  &  3  \\
-		IT services                                 &  2  \\
-		Government                                  &  1  \\
-		Information security                        &  1  \\
-		Not specified                               & 10  \\
-        \midrule
-		Total                                       & 42  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Business areas of case organizations in primary studies}
-    \label{table:businessareas}
-\end{table}
-
-The business areas of the case organizations are presented in table
-\ref{table:businessareas}. Different business areas were represented somewhat
-evenly. Online services was the largest group, including companies
-providing software as a service solutions for businesses, online media players
-for consumers, online services for consumers, and communication software for
-businesses. The second largest group was telecommunications, including companies
-such as British Telecom, Cisco, Ericsson, and Nokia Siemens Networks. The third
-largest group was enterprise management solution providers with products for
-business process management, project portfolio management, and facility
-management.
-
-\subsection{Agile methods used}
-
-The agile methods reported being used in the transformed organizations are
-summarized in table \ref{table:agilemethods}. The most prevalent method was
-Scrum, which was the sole agile method mentioned in 25 cases. The second most
-mentioned agile method was Extreme Programming (XP). Lean software development
-was mentioned in some studies, although combined to Scrum in all cases. Other
-agile methods mentioned were Unified Process, ADM, and Rapid Application
-Development. In one case the agile method was not named.
-
-It was quite common that organizations sought to combine agile methods.
-Especially Scrum, XP, and Lean software development were used together. In
-addition, many cases mentioned use of XP practices without explicitly stating XP
-as the process being used (such as test driven development and continuous
-integration). Combining and customizing agile practices was also evident in that
-many organizations viewed the agile method as continuously evolving.
-Organizations evolved the agile methods for instance through retrospectives and
-continuous improvement.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ l r }
-        \toprule
-        Method                             &  N  \\
-        \midrule
-        Scrum                              &  25 \\
-        XP                                 &  4  \\
-        Scrum and XP                       &  5  \\
-        Other                              &  4  \\
-        Combining Lean to an agile method  &  6  \\
-        Not specified                      &  1  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Agile methods reported in number of cases}
-    \label{table:agilemethods}
-\end{table}
-
-\clearpage
-
-\subsection{Reasons to change}
-
-As stated by Research Question 1 we investigated the reasons organizations start
-a transformation process towards agile software development. Most primary
-studies mentioned reasons for starting the transformation, and only 7 cases of
-42 did not mention any reasons. In the cases giving no specific reason to change
-it was simply mentioned that management had decided to change the organization.
-The reported reasons for starting an organizational change towards agile are
-summarized in figure \ref{fig:reasonstochange_summary}.
-
-% Some cases reported that the organization had been successful in waterfall
-% delivery for a long time before the agile transformation.
-
-The most prominent reason to start the transformation was that business demanded
-shorter lead times and faster delivery. Other top reasons to start change were
-dysfunction in management, overhead in the current process, problems in
-scheduling, and demand to improve quality.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics{graphics/reasonstochange_summary.pdf}
-    \caption{Reasons why the studied organizations started an agile
-             transformation}
-    \label{fig:reasonstochange_summary}
-  \end{center}
-\end{figure}
-
-\subsubsection{Business demands speed and flexibility}
-
-Most cases reported that a major reason to change was that business demanded
-improvement. Most companies wanted to improve their competitiveness, or even
-feared that they are losing competition. The way to respond was to improve
-delivery speed and responsiveness to change. Table
-\ref{table:reasonstochange_business} presents business related reasons to
-introduce an agile model.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Business reasons for starting a change   &  Described in  \\
-        \midrule
-        Time to market must be reduced    &  P17, P18, P32, P38, P39, P47  \\
-        Delivery is considered slow       &  P2, P8, P25, P36, P46, P49  \\
-        Something must be done to remain competitive  &
-                P9, P10, P11, P18, P19, P32, P38, P42, P49  \\
-        Business partners demand faster delivery  &  P5, P23, P39, P40, P46  \\
-        Need to accommodate change during projects  &
-                P4, P8, P18, P25, P26, P33, P36, P39, P42, P49, P51, P52  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Business related reasons for starting an agile transformation}
-    \label{table:reasonstochange_business}
-\end{table}
-
-The most frequently mentioned reason for introducing agile methods was the demand to
-deliver software faster and more frequently. A common statement was that it was
-necessary to reduce the time to market (TTM) [P17, P18, P32, P39, P47]. As
-stated in one study: ``The market was moving faster than the development group
-was able to deliver'' [P38]. Long cycle time, i.e. time from start of project to
-delivery, was seen as a problem in itself [P25, P36]. Some studies indicated
-that the organizations set as one of the transformation goals to deliver
-software faster [P2, P8]. Other cases described that development was generally
-considered as slow [P36, P46, P49].
-
-Commercial competition and market situation was seen as a trigger for
-transformation. Especially speed of delivery was seen as a means to respond to
-the competition [P9, P38]. As Greening [P19] describes: ``The average release
-duration peaked at 41 and 35 months, an alarming state that could enable
-competitors to gain market share.''. In many cases the transformation was seen
-as essential for the company to remain competitive [P10, P11, P18, P42, P49].
-For instance, British Telecom realized that ``From a strategic view, there is
-little alternative than to change the old ways of working'' [P32].
-
-Demands for change and improvement were also expressed by internal and external
-customers. Again, the requests of the customers reflected the need of faster
-development and release cycles [P5, P39]. As described in one study: ``Speed of
-delivery is of the essence -- customers are no longer prepared to wait for 18
-months or more for solutions to be delivered'' [P23]. In some cases business
-partners expressed mistrust in the original way of conducting development [P40,
-P46].
-
-Yet another factor pushing for improvement was the need to accommodate change
-during development. Organizations saw agile as an opportunity to improve
-responsiveness to change in general [P25, P26, P42].
-Some organizations faced difficulties when changes to plans were needed. For
-instance, the organizational structure made adapting to changes difficult [P52],
-and making changes created management overhead [P51]. Other organizations saw an
-business advantage in enabling priorities to be changed as required [P8, P39].
-Domain knowledge will deepen during projects and environments will inevitably
-change, and therefore the requirements should not stay fixed [P4].
-Customers and business representatives were in many cases attributed as the
-sources of change requests, and the responsiveness to customer requests needed
-to be improved [P18, P33, P36, P49].
-
-\subsubsection{Management issues}
-
-The way software development was managed was typically seen as problematic. The
-problems related to project management, people management, and managing
-schedules. Management related problems that the transformation was hoped to
-correct are listed in table \ref{table:reasonstochange_management}.
-
-\begin{table}[b]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Management problems           &  Described in  \\
-        \midrule
-        Software development is managed badly   &  P8, P21, P30, P38, P44, P49  \\
-        Command and control style management    &  P9, P25, P30, P40, P44  \\
-        Scheduling and estimation is difficult  &  P16, P19, P21, P30, P33, P40, P41, P51  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Management problems stated as reasons for organizational change}
-    \label{table:reasonstochange_management}
-\end{table}
-
-In many organizations the project management practice was in a state of
-dysfunction, and it needed to be corrected. Project management problems included
-constantly recurring crisis events [P30], managers ignoring prioritization of
-tasks [P38], and reluctance towards focusing on delivering business value [P49].
-Some cases stated that projects simply failed [P21]. Some organizations wanted
-to improve generic project management practice, such as risk management [P8],
-management of offshoring [P38], and generally improving the quality of life of
-team members [P44].
-
-Command and control style management where ``a few people, usually those
-farthest from the action, made the decisions without directly involving the
-developers doing the actual work'' was highlighted as a problem [P44]. For
-instance, schedules could be dictated by management, and teams had to comply to
-them regardless of the workload [P9]. As a result teams were overworked and
-needed stretch far to finish projects on time. Low morale and motivation were
-identified as problems that needed to be fixed, and agile was seen as a solution
-[P25, P30, P40].
-
-Yet another problem in managing development was problems with schedules.
-Some cases described schedules as unpredictable and estimation being difficult
-[P16, P30, P40, P51]. In other cases it was reported that release deadlines were
-often missed [P19, P21, P33, P41].
-
-\subsubsection{Process problems}
-
-The old way of working was acknowledged to have problems which needed to be
-corrected. A summary of weaknesses in the original process is presented in table
-\ref{table:reasonstochange_process}.
-
-\begin{table}[b]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Problems in the old process           &  Described in  \\
-        \midrule
-        Excess and overhead in process     &  P9, P24, P35, P36, P38, P40, P46, P49, P51  \\
-        The process is being ignored       &  P6, P9, P46, P49  \\
-        Late integration, testing, and feedback  &  P5, P25, P33, P40, P50  \\
-        New practices can not be integrated in the old process  &  P10, P26  \\
-        The old process does not scale up  &  P16, P30, P50, P51  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Problems because of which the old process should be changed}
-    \label{table:reasonstochange_process}
-\end{table}
-
-A most typical known problem was overhead in process, which showed as bureaucracy and
-needless costs. Examples on overhead in process were unproductive meetings
-[P38], sign-offs and gates [P9], excessive focus on conformance to process
-[P40], and change management overhead [P51]. Creating excessive documentation
-instead of business value was seen as a problem in some cases [P24, P35]. From
-business representatives point of view the processes were unreasonably costly,
-while at the same time insufficiently capable of delivering business value [P36,
-P46, P49]. In one case business realized that customers did not see any benefit
-in the heavy CMMI based process [P24].
-
-A heavy process resulted often in ignoring the process, only paying ``lip
-service'' to the process, or applying its steps retroactively [P6, P9, P49].
-Seffernick [P46] describes how applying the waterfall process had teams ``still
-developing the same old way but slower because of the additional process''.
-
-A slow process with long cycle times leading to late feedback was a problem in
-the old model. A large investment was required before results were visible [P5],
-and there was variances in expectations and what was delivered [P40]. A related
-problem was late integration which caused extra work [P25]. As organizations
-grew and products became more complex test cycles began to lengthen, which
-hindered flexibility and reduced speed [P33, P50].
-
-A few organizations were experiencing rise of agile methods at grass roots
-level, but the agile teams could not integrate efficiently with other parts of
-a waterfall organization. Tensions arose when teams using different development
-approaches attempted to work together on products [P10]. On a team level agile
-had enabled some flexibility, but that flexibility could not be leveraged on the
-scale of the entire organization [P26].
-This kind of experiences drove the organization to deploy agile on a full scale.
-
-Still one problem was that some growing organizations experienced that the old
-development method did not scale up [P16, P30, P50, P51].
-
-\subsubsection{Other reasons to change}
-
-There were still a number of problems that were given as reasons to change.
-In addition to the above mentioned problems, organizations were striving for
-better collaboration and quality. Table \ref{table:reasonstochange_other} lists
-other issues that drove organizations to change.
-
-\begin{table}[b]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
-                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
-        \toprule
-        Reasons to change           &  Described in  \\
-        \midrule
-        Teams can not collaborate properly   &  P21, P38  \\
-        Silos and internal boundaries in organization   &  P27, P40, P46  \\
-        Lack of customer collaboration       &  P4, P26, P40  \\
-        Demand for better quality            &  P2, P4, P25, P31, P30, P38, P41, P42  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Other reasons why organizations wanted to change the way of working}
-    \label{table:reasonstochange_other}
-\end{table}
-
-Many organizations experienced problems in communication and collaboration
-between projects, teams, and even members of a single team [P38, P21].
-Teams were in many cases organized around components and were working in silos.
-Working in silos diluted focus and made adapting to change difficult, as teams
-could only focus on their limited area of expertise [P7]. Functional silos acted
-against creating a shared purpose in operating [P40], and resulted in a ``ball
-over the wall'' mentality [P46]. Distance between teams was also created by
-document driven communication, and was causing misunderstanding of requirements
-[P24].
-
-Too little customer and user collaboration led to problems in some
-organizations. For instance, too late user involvement led to change requests
-late in the project [P40]. There was a wish to collaborate more frequently and
-in less formal ways [P4, P26].
-
-% There was also a desire to improve team motivation and commitment, and
-% collaboration between management and developers [P39].
-
-% Examples of problems with requirements was too much focus on the documenting
-% aspect instead of providing understanding [P24, P49], and ignoring the fact that
-% requirements and understanding are evolving during a project [P4].
-% 
-% In some cases the organization of work and management was such that it caused
-% lack of focus the development [P7], and prioritization of what to work on was
-% not done efficiently [P38, P39].
-
-Lastly, the original model was seen as a cause for quality problems.
-Several studies mentioned a demand for better quality, and quality improvement
-was set as a goal for the agile transformation [P25, P31, P41, P42]. New feature
-development was being prioritized over bugfixes, and technical debt was
-developing [P30, P38]. Improvement in external quality and customer satisfaction
-were anticipated from agile development [P2, P4].
-
-\clearpage
-
-\subsection{Transformation process}
-
-The second research question was aimed to discover how the process of
-transformation can be described.
-We provide an answer to Research Question 2 in several parts. First the initial
-states of organizations are described. Secondly, we present a model for
-categorizing transformation types. In the rest of this section we discuss
-typical arrangements that relate to the change process, including how training
-and piloting is arranged, how change is lead, and what investments are done.
-
-The fact that every organization is unique creates challenges in giving generic
-descriptions on how transformations proceed. Further, publications tended to
-provide more focused viewpoints rather than describing the transformation
-process itself. Therefore it was not sensible to attempt constructing a detailed
-model on how large organizations adopt agile methods.
-Table \ref{table:transformation_descriptiontypes} presents types of insights
-each primary source provided in relation to Research Question 2.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.52\textwidth}
-                     >{\raggedright\arraybackslash}p{0.40\textwidth} }
-        \toprule
-        How transformation is described   &  Primary sources  \\
-        \midrule
-        Narrative description of transformation process &
-                P3, P5, P6, P9, P11, P12, P16, P18,
-                P24, P27, P30, P36, P40, P41, P46   \\
-        Application of agile methods seen from a specific viewpoint (e.g. coaching team) &
-                P10, P13, P15, P19, P23, P32, P34,
-                P39, P47, P48, P50, P52    \\
-        Viewpoints of challenges and success factors  &
-                P17, P20, P21, P29, P37, P38, P49   \\
-        Discussion on establishing large scale agile practices  &
-                P7, P8, P31, P42, P43, P44, P45   \\
-        Only brief discussion on transformation process  &
-                P1, P2, P4, P14, P22, P25, P26, P28, P33, P35, P51  \\
-        \bottomrule
-    \end{tabular}
-    \caption{The theme of each primary source, in relation to providing
-             insights of the transformation process}
-    \label{table:transformation_descriptiontypes}
-\end{table}
-
-\subsubsection{The initial state of the organization}
-
-We investigated the initial state of the case organizations before the
-transformation. The typical initial process model was described as waterfall, as
-stated in 25 cases. In several cases the authors characterized the initial state
-of the organization as ``traditional''. CMMI was presented as a driver for the
-original process in 4 cases. Other process models the studies named were RUP
-[P19], Prince2 [P51], and Macroscope [P14]. In 9 cases the initial state was not
-described.
-
-The old model was described in most cases briefly and in a neutral sentiment.
-Waterfall development in general was referred to as ``traditional, plan-based
-methods'' [P25], ``traditional phase-based approach'' [P36], or ``old fashioned
-process driven waterfall'' [P23]. Korhonen [P25] referred to the model presented
-by Royce \cite{Royce1970} when describing the initial state.
-
-In some cases specifics of the old model was criticized. The characteristics of
-the old process models were described as having ``many gates and sign-offs
-required by upper levels of management'' [P9], ``lot of bureaucracy in project
-management and a strict hierarchy of requirements communication'' [P39], and
-``project managers typically used a command-and-control philosophy'' [P44].
-
-In a few cases the initial process model was less rigid, and closer to the agile
-model that was to be introduced. Examples of such loose process models were
-described: ``The choice of development methodology is one of any number of
-degrees of freedom granted to teams at Amazon. Development teams are expected to
-solve their problems in the best way possible'' [P3], ``the R\&D group leveraged
-a loose, waterfall-based process with an entrepreneurial culture'' [P16]. Some
-organizations had even signs of agile development on team level although the
-surrounding organization operated in a waterfall fashion. This was described as
-``Agile and Scrum were increasingly popular and emerging at a grass-roots level,
-at a corporate level Yahoo! still embraced a more traditional product
-development process'' [P10], and ``some teams had applied scrum on the team
-level, despite the surrounding organization that was working with a waterfall
-software engineering model'' [P26].
-
-In a few cases the organization initially lacked strict process. For instance
-Long [P30] describes the initial state in the following way: ``Project scope was
-in constant flux. Well-meaning market and product owners routinely inserted new
-requirements during product development.''.
-
-Some authors describe development being originally organized in silos. For
-instance, Ranganath [P40] describes that ``Teams operated in silos.
-Collaboration, shared purpose and ownership for outcomes were hard to find''.
-
-\subsubsection{Modes of transformation}
-
-Based on the primary studies we identified two dimensions in the characteristics
-of transformation processes. The first dimension is the pivot point of
-leadership in change. This dimension has two extremes: top-down leadership where
-the change is mandated by management, and bottom-up leadership where the change
-emerges from team level. The second dimension is the velocity of the change.
-Organizations were reported to perform transformations as a big bang or
-overnight fashion, or as a step wise process lasting for a long time. Figure
-\ref{fig:transformation_dimensions} summarizes the transformation process
-dimensions.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics{graphics/transformation_dimensions.pdf}
-    \caption{Dimensions for categorizing transformation types}
-    \label{fig:transformation_dimensions}
-  \end{center}
-\end{figure}
-
-We do not quantitatively measure cases on the transformation process dimensions,
-because the criteria for creating a measurement would be vague. For instance, we
-do not think that presenting a single quantitative measurement of the weight
-point of leadership on an axis between management centric and team centric would
-describe reality accurately. The dimensions of pivot of leadership and
-transformation velocity nevertheless exist.
-
-As the primary sources describe, transformation cases combined the dimensions of
-leadership pivot and velocity. Four combinations of the characteristics could be
-observed. Figure \ref{fig:transformation_types} summarizes the combinations of
-transformation characteristics.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/transformation_process.pdf}
-    \caption{Categorization of transformation types}
-    \label{fig:transformation_types}
-  \end{center}
-\end{figure}
-
-The most extreme case of adoption was a big bang transformation. To realize such
-a large change it is necessary that the change was centrally coordinated.
-For step wise transformations the the amount of management and development team
-involvement varied. At one extreme, the change was lead by management while
-teams played a passive role. At the other end management was described to be
-reluctant to change while teams pushed for change. Many cases took a middle
-course where both management and development dealt with the transformation on
-their respective level of influence.
-
-As a concept separate from the dimensions of leadership pivot and velocity, the
-transformation was described as a continuous process. There were a few
-viewpoints illustrating the concept of continuous transformation. Several
-organizations were striving to introduce a culture of continuous improvement,
-which would make organizational adjustment a permanent state. In many cases the
-agile transformation was not considered to have ended at any distinct point in
-time, and the impression that it was still ongoing was given. Finally,
-transformation was simply described as hard work that will inevitably take a
-long time.
-
-% Mentioning big bang versus step wise [P16].
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% ---------------------------------------------------------------------
-
-% ---------------------------------------------------------------------
-
-\subsubsection{How change was lead}
-
-A significant change effort requires entities in the organization who take
-leadership in the change. Getting a large group of people aligned towards a
-common goal, in many cases a completely new way of working, requires that
-someone takes charge for the undertaking. The primary sources showed evidence of
-leadership for change on all organizational levels. Table
-\ref{table:transformation_leadership} summarizes different types of change
-leaders that were discovered.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Change leaders         &  Described in  \\
-        \midrule
-        High-up change leaders or sponsors  &
-                P3, P5, P24, P31, P36, P38, P46, P49  \\
-        Agile task group for coordinating change   &
-                P4, P16, P18, P21, P23, P25, P30, P40, P41,
-                P43, P45, P48, P49, P50  \\
-        Change leaders on development team levels  &
-                P2, P6, P9, P14, P17, P18, P27  \\
-        Informal spreading of agile awareness  &
-                P3, P6, P14, P40  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of entities leading change in the agile transformation}
-    \label{table:transformation_leadership}
-\end{table}
-
-% High-up change leaders
-
-In many cases it was described that the change was initiated by a limited number
-of people in high-up positions. In these cases typical positions attributed for
-giving the initial push for change were CxO's, vice presidents, and directors.
-In these cases it was highlighted that one or very few high-up persons affected
-significantly the start or expansion of the agile transformation. In some cases
-the change leader was a new hire [P5, P38, P46].
-
-The presence of a high-up manager was described to underline the organization's
-commitment for change when dealing with employees or other stakeholders [P24].
-The firm commitment of the high-up leader was key in keeping the agile
-transformation on course in cases where objections or other problems arose [P5,
-P14, P36, P46]. Another important role of the change leader was to persuade and
-work for getting other people raise an interest in agile methods [P3, P31, P38,
-P46, P49].
-
-% Agile task force
-
-Many organizations formed an agile task force or similar entity for coordinating
-and leading the change. It was typical that in the beginning of the
-transformation, or when it had been acknowledged that the old model had
-problems, a work group was founded for coordinating the changes to be done. The
-change team usually consisted of management personnel [P43, P50], and was
-reinforced with external agile consultants [P18, P21, P25, P30]. Some cases
-highlighted that the team had a cross organizational staff who were familiar
-with the varying functions that would be affected by changes [P4, P16, P21, P41,
-P48]. It was important that there was a group of people who were strongly
-committed to making a change happen [P43].
-
-The typical first task of the change team was to decide on an alternative
-delivery method to solve problems in the old model [P21, P30]. During the
-transformation the team stimulated agile adoption in the organization [P18, P45,
-P49]. Other examples on the change team's tasks were informing people on the
-agile model [P4], educating and coaching [P16, P45, P48, P49], creating an agile
-cookbook [P23], and organizing presentations and workshops [P25, P41]. The
-change team also assessed the success of transformed teams, and worked for
-repeating their success across the organization [P16, P40, P41]. The change team
-was empowered to act on removing impediments they discovered [P16, P50].
-
-% Other change leaders
-
-Beyond executive positions, leadership for change emerged also on lower
-organizational levels. In many cases these change leaders were still in
-management positions [P14, P18, P17]. There were also agile champions in teams
-who drove the change on a grassroots level [P6, P9, P27].
-As with higher-ups, these people worked for making agile methods familiar [P2,
-P6, P9, P17], making change stick [P2, P18], and persuaded others to use agile
-methods [P2, P9, P14].
-Benefield [P6] describes that people in teams were well able to adjust the
-application of the agile process as they were familiar with the problems at
-hands. It was however hard to find people who both understood the agile
-principles well, and knew how to coach within the team [P6].
-% Also coaches in [P2], and a team in [P16]
-
-Chung and Drummond [P9] give a further description of grassroots level
-leadership. Yahoo! initially used a team of coaches to help applying agile, but
-the coach team failed to make agile take root. The coaches were later disbanded
-due to reductions in budget. The agile practitioners in teams took over the
-leadership for the continued application of agile [P9].
-
-% Informal spreading of awareness
-
-Change leadership showed on the lowest level as informal spreading of awareness
-of agile methods.
-Benefield [P6] describes that the strongest force over other change leaders was
-letting agile knowledge spread virally. New Scrum teams were emerging as people
-heard of the new methodology [P6]. In some cases interest groups were used to
-stimulate viral spreading of agile [P14].
-It was similarly described that in an organization with an introverted culture
-the proclamations of managers could not match with the power of practitioners
-explaining to others how they had succeeded in with the agile method [P40].
-Even casual hallway talks could spread the word effectively, and it was also
-easier to encounter people from other parts of the organization in an informal
-context [P3].
-
-\subsubsection{Training, coaching, and awareness}
-
-During transformation periods organizations invested in activities aimed to
-incorporate the new model into the thinking of people. People were new to agile
-methods, and a common understanding needed to be created. Training, coaching,
-and public events were used for communicating the new model to a significant
-part of the organization. Describing the use of training and coaching was a very
-prominent facet in transformation cases.
-Table \ref{table:transformation_training} summarizes activities that
-organizations use for deepening the understanding and awareness of the new
-development process being applied.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
-                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
-        \toprule
-        Activity           &  Described in  \\
-        \midrule
-        Training    &
-                P3, P6, P8, P9, P11, P16, P18-P21, P23-P25, P27, P30,
-                P32, P35, P37, P39-P41, P43-P46, P48-P51  \\  % 29
-        Coaching    &
-                P1-P3, P6, P8, P9, P11, P16, P18, P23, P27, P29,
-                P31-P33, P36, P39-P41, P45-P47, P49, P50, P52  \\  % 25
-%         In-house coaching   &
-%                 P1-P3, P6, P11, P16, P23, P29, P31, P32,
-%                 P39, P41, P45, P47, P49, P50, P52  \\    % 17
-        Seminars and workshops   &
-                P1-P4, P7, P9, P10, P18, P11, P16, P23, P25,
-                P26, P27, P32, P41, P44, P46, P47   \\   % 19
-        Communities, special interest groups  &
-                P3, P6, P7, P9, P10, P11, P12, P14, P32, P41,
-                P42, P45, P47, P52  \\   % 14 
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of activities for building awareness and knowledgeability
-             in the agile way of working}
-    \label{table:transformation_training}
-\end{table}
-
-% Training
-
-As the organization is changing it is necessary to train people so they get
-familiarity with the new model that is being pursued. The majority of cases
-described the use of training in the transformation.
-Most typically the change leaders and pilot teams were the first to attend
-formal training on agile development. Management was also usually among the
-first ones to receive agile training. In some cases particularly management
-received more in-depth training [P11, P16, P24]. In addition to providing an initial
-understanding of agile it was considered important that the key people get a
-common understanding of the method [P11, P24, P39].
-
-After key people had attained knowledge on agile, and pilot teams had possibly
-tested the methods, other teams received training on the new development model.
-In some cases the training of development teams was described to be brief [P11,
-P16, P20]. Other cases implemented a full day workshop [P18, P23, P32, P41,
-P45], and some organizations invested even in multi-day training [P9, P19, P37,
-P49].
-The reduced training for engineers was compensated with coaching during ongoing
-projects [P11].
-
-Several cases mentioned that the first introductory events were held by known
-industry experts [P3, P19, P25, P27, P30, P46]. External consultants were also
-often used as trainers [P18, P25, P40, P41, P49, P51]. In some cases a limited
-number of people with agile knowledge was first trained or acquired, and later
-the training was arranged with internal resources [P3, P11, P24, P25, P41, P46,
-P49, P51].
-The training curriculum was evolving as trainers improved and learned the needs
-of the organization [P6]. In some cases training workshops was initially
-tailored to suite the needs of the project [P23].
-
-Training was considered as important. Training was reported to make a difference
-between success and failure of teams [P6].
-It became clear that training was necessary to retain an agile culture [P19].
-Training was also described to increase the general interest towards using agile
-methods [P3].
-There were also some negative observations related to training. Atlas [P3]
-describes that the insights gained from training were less than anticipated
-[P3]. In another case product managers did not receive training and they stuck
-in old ways of managing, which jeopardized the transformation [P35].
-
-% Some training included simulation workshops [P40, P41].
-% Hands-on workshops [P46]. 
-% Certified Scrum master training --> In several cases
-
-% Coaching
-
-% What
-In addition to training, a coaching approach was usually used to teach how to
-work in an agile environment. As agile lacks rigorous prescriptions and builds
-on human interactions, it is best learned in practical situations while a coach
-is giving direction.
-The main task of coaches was to guide development teams in their daily work.
-Some cases described that roles such as the product owner or scrum master
-received specialized coaching [P33, P52].
-In addition to helping teams, coaches were driving the agile push [P2, P36,
-P49], and advocated agile methods in general [P45].
-
-% Why
-Coaching was considered necessary, especially in the beginning of agile projects
-[P6, P8, P33, P46, P47]. Brown [P8] stated that without adequate coaching there
-would have been severe long term consequences [P8]. The application of coaching
-was seen as an important success factor [P18, P29, P33], and the job of the
-coach was described to be to ensure that the transformation would be successful
-[P16]. One viewpoint was that the coach was an impartial third party who could
-be relied on when resolving transformation impediments [P16, P29, P44].
-
-% Having too little coaches [P23].
-% Coaches were also involved in training [P27].
-
-% Internal coaching
-The coaching positions were most often held by internal staff.
-Internal coaches were required for making the change grow, and to scale up the
-agile implementation [P47, P49]. Internal coaches had also the benefit of
-understanding of the company's processes and systems [P23, P39], and teams were
-able to access them easily [P3, P41]. Some cases presented building of a coach
-community as an integral part of the agile adoption [P6, P23, P32, P47].
-
-% How coaching evolves
-As the agile way of working penetrated the organization the need for coaching
-was decreasing, and coaching could be reduced [P18, P36]. However, some cases
-described that the demand for coaching had become permanent in order to keep the
-development model healthy and to continuously pursue improvement [P6, P23].
-External coaches moved out and internal coaches took their place [P40].
-The maturity of coaches was described to grow over time, as the agile
-transformation progressed [P32].
-
-% Scrum masters that had been trained earlier were coaching the teams in some
-% cases.
-
-% Seminars and public events
-
-% What public events
-Organizations typically strived to increase the agile awareness by introducing
-series of events for promoting agile, ranging from public seminars to brown bag
-meetings and hands-on workshops. These events were aimed at a general audience,
-consisting of all organizational groups. Events were arranged especially in the
-beginning of the agile journey. In some cases the demand and interest in public
-events was maintained, and the events were continued.
-Some cases describe continually arranged agile events serving a central role in
-maintaining and advancing further the agile model [P23, P47].
-Training events were even productized so that they could be replicated for a
-large number of people [P32, P46].
-
-% Content of events
-In addition to general training and enabling awareness, the purpose of agile
-events was to evangelize and expand the agile model throughout the organization
-[P1, P10, P23, P32, P46].
-Workshops were used to get people better involved [P16], and public lectures
-were used to set straight sidetracks in the agile implementation [P26].
-Hands-on workshops were used as a more engaging teaching method compared to
-traditional lectures [P32, P41]. Open events were good for delivering the same
-message to many people at the same time. As everyone could heard the same
-questions and answers the understanding was consistent [P9, P41].
-Workshops were arranged for identifying problems and planning the new model [P7,
-P16, P27].
-
-% SIGs, CoPs, and community
-
-Events and coaching activity were coordinated forms of promoting the new model,
-but the agile transformation gave also birth to communities [P3]. Typically
-communities were emerging later in the transformation, when the rollout team
-was starting to step aside [P3, P41].
-In some cases the communities were created by passionate individuals without
-corporate intervention, and participation in communities was voluntary [P47].
-Communities were created when the mass of people involved became too large for
-formal control [P52].
-In other cases the forming of communities was stimulated by the organization
-[P12, P45].
-Companies strived to create communities of practice [P42] or special interest
-groups [P41] in order to reinforce continuous learning in the organization.
-Communities helped people to form a learning habit, which promoted continuous
-improvement in the organization [P52].
-
-\subsubsection{Piloting}
-
-% It was typical that before taking the transformation to full scale the new model
-% was tested in use in the organization specific environment.
-% Pilot projects and programs were a common practice at the start of the
-% transformation. Piloting was discussed in about half of the cases.
-
-A logical first step in an agile transformation is to try out the new methods in
-a pilot project. A large amount of the primary studies described how piloting
-was used in the beginning of the transformation. Primary sources discussing
-the use of piloting are listed in table \ref{table:transformation_piloting}.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
-                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
-        \toprule
-        Sources discussing viewpoints on piloting   & 
-                P6-P11, P14, P16, P18, P22, P23, P24, P27, P29,
-                P36, P38, P40, P41, P43, P45, P46, P47  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of activities aiming to advance the transformation}
-    \label{table:transformation_piloting}
-\end{table}
-
-Piloting was typically described as the first step in transformation, and
-yielded typically success and positive results [P7, P9, P14, P38]. In many cases
-management needed proof of benefits of agile methods, and successful pilots
-granted approval for agile [P14, P18, P24, P36, P41, P43, P47].
-A piloting stage was seen as important for any transformation [P8, P21].
-
-% Cases described starting the transformation with a pilot, and then proceeding
-% gradually [P11].
-
-There was some variance of approaches in choosing the pilot projects.
-In some cases the project or teams to pilot the agile methods were chosen
-carefully [P7, P8, P38, P40], but others chose the pilot teams from volunteers
-[P9, P10, P45].
-Some organizations considered it important to pilot agile with business critical
-projects, as the applicability had to be proven in a relevant environment [P8,
-P40].
-
-Piloting could refer to considerable size or time. In one case over 100 people
-participated in 8 pilot projects [P8], and another case described an
-experimentation period of 8 months [P34].
-It was also described that in order to gain broad insights piloting was applied
-in many different projects [P6], and also in projects involving many teams [P8].
-
-Cases described that piloting gave valuable insights that helped the
-transformation further.
-Pilot projects revealed organization specific problems and critical factors for
-the use of agile [P8, P9, P29, P40, P45].
-Based on the experience gathered in pilots a structured initiation approach for
-agile projects was created [P8], and the requirements for well functioning team
-areas were discovered [P22].
-The visible success of the pilots also encouraged new teams to copy and start
-applying the agile methods [P18].
-
-% What other things did the pilot teach?
-There were still a few particular insights regarding use of piloting.
-Fry and Greene [P16] describe a particular big bang adoption where piloting was
-omitted. A pilot project was avoided because having two models side by side was
-deemed to create organizational dissonance. Instead a significant amount of
-people completed agile training and the new model was introduced overnight
-[P16].
-In some cases the success of the pilot gave unrealistic expectations [P38, P45].
-Organizations should beware if the team is set up of all-star developers, as
-the results will likely be much better than average [P38].
-
-% Even reports which presented a big bang adoption described that the main
-% transformation was preceeded by piloting [P11, P18, P41].
-
-\subsubsection{Investing in new structures in transformation}
-
-Even if any activities aiming for transformation can be considered as
-investments in themselves, the transformation process included some specific
-tangible investments. These investments were focused on better enabling agile
-software development. They also highlight the concerns that organizations had when
-implementing an agile model. Typical investments were bringing in external help,
-new management tools, improvements in testing environments, and updating of the
-physical working spaces. Table \ref{table:transformation_investments} summarizes
-common types of investments.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.35\textwidth}
-                     >{\raggedright\arraybackslash}p{0.57\textwidth} }
-        \toprule
-        Investments during transformation      &  Described in  \\
-        \midrule
-        Using external consultants  &
-                P5-P9, P16-P21, P23 P25, P27, P29-P31,
-                P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
-        New project management tools for agile  &
-                P10, P11, P14, P16, P17, P26, P33,
-                P41, P43, P44, P45, P48, P50 \\
-        Creating a continuous build and integration system  &
-                P5, P11, P16, P17, P25, P27, P34, P39, P42, P50 \\
-        Proprietary agile material or documentation   & 
-                P1, P2, P8, P14, P16, P23, P26, P29, P32, P45  \\
-        Refurbishing work areas and other spaces   &
-                P6, P22, P35, P36, P38, P42, P46, P48 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of investments done during transformation, aiming to make
-             agile methods perform better}
-    \label{table:transformation_investments}
-\end{table}
-
-% External consultants
-
-It was very typical that external help was brought in to help with the
-transformation.
-External consultants were employed especially at the start of the
-transformation. At the time it was realized that the way of working should be
-changed a consultant was often hired to assess the current state and give a
-recommendation on how to proceed [P16, P20, P41, P43, P41, P44, P51].
-In the beginning external consultants were used to ignite interest in agile
-methods, and provide initial education [P7, P9, P27, P30, P43].
-It was recommended to bring in external help as early as possible, when the
-internal agile expertise is lacking and before transformation challenges appear
-[P6, P16, P30].
-
-When the transformation was getting on its way, agile consultants were typically
-leading training sessions, arranging agile workshops, and coaching teams.
-In some cases consultants helped specifically with scaling up the use of agile
-beyond single teams [P31, P34].
-Consultants were used mostly in the first stages of the transformation.
-The need for external consultants reduced over time, and their tasks were
-transferred to internal positions [P29, P47].
-It was described that as the agile journey progressed, external consultants were
-invited every now and then to bring in fresh ideas to the organization [P6].
-
-External consultants were regarded to bring in critical assistance in
-transforming the organization [P6, P16, P17]. Qualities in external consultants
-that organizations looked for was leveraging their experience to bring more
-speed to development and agile adoption [P17, P27, P34, P35, P40], and to bring
-in agile skill when only little existed in-house [P27, P47]. It was considered
-beneficial having experts with previous experience of transforming organizations
-[P16, P27].
-Still one important aspect in using consultants was that people were more
-comfortable when the advice was given by impartial persons external to the
-organization [P16, P29, P44].
-
-% In one case the teams were sceptical of having external consultants embedded
-% in temas [P27].
-
-% Agile development tools
-
-Using and investing in software tools supporting agile development was mentioned
-in several studies. Especially project management tools were discussed.
-Some reasons for introducing new project management tools was that spreadsheets
-became unmanageable [P16], and the status overview of agile projects was lacking
-[P17]. When a suitable agile tool was implemented it was attributed for better
-visibility [P26, P33]. Using a project management tool can also help in teaching
-the particularities of the new process [P14].
-
-It was typical to use web-based tools for project management [P10, P11, P16].
-The use of web based tools such as wiki pages was also attributed to lessen the
-effort in documentation [P14]. Other tools that were taken into use while
-implementing agile methods were teleconferencing tools [P11], discussion forums
-for contacting coaches [P41], and information radiators to raise the awareness
-of the new method [P22, P41]. Also development and delivery infrastructure
-required investments to match the needs of agile development [P50].
-
-Several organizations created their own tools for managing agile development
-[P10, P11, P26]. Especially companies developing management software, such as
-SAP, Primavera, and Salesforce, customized their product to support the agile
-model being implemented [P16, P44, P45]. The development of the tools themselves
-also served as a showcase of agile delivery [P10, P45].
-
-% Investing in continuous builds
-
-Continuous builds and testing was reported to play an important role in scaling
-the agile model up to large scale [P17, P34]. Many organizations were
-experiencing problems when a daily build system was not in place. The main
-problem was that progress was slowed down because of lacks in the build and
-automated testing systems [P5, P27, P39, P50].
-Other reasons for implementing build automation was the need for having stable
-nightly builds [P5], builds were often broken when integration was attempted
-[P27], and there were problems with timeliness in test execution [P11]. One case
-reported that the entire infrastructure needed to be revamped to attain benefits
-from the accelerated work pace of Scrum teams [P50].
-
-When a general problem of frequent broken builds existed it was easy to acquire
-a buy in for creating an organization-wide standard for automating builds and
-tests [P11]. On the other hand, many cases emphasized that building a continuous
-integration environment required a significant investment [P17, P34, P42].
-In order to have the focus and resources for creating a proper integration
-environment a dedicated team was created. Having an integration team ensured
-that focus and resources were available for creating a proper integration
-environment [P5, P16, P17].
-
-Having daily builds was considered a significant advantage in the agile
-implementation [P5, P16]. Having the continuous integration environment in place
-cases reported that development time was reduced [P39], the build times were
-significantly improved [P5, P42], teams could know on a daily basis whether
-their code could be integrated [P17]. It was also reported that the
-communication and trust between testing and development was improved [P42].
-
-% Organization specific agile material
-
-A typical practice was creating some internal material promoting the new agile
-model. The material served both a documenting role and promoted awareness for
-the new development model. The material was created to help explaining the agile
-model to stakeholders, and to serve as a reference for teams implementing it.
-Typical in-house agile material was a manual which presented the organization
-specific implementations of agile practices [P1]. Many cases described using a
-wiki-based website to maintain and distribute information on agile practices
-[P14, P16, P45]. Hanly et al. [P23] describes how a n organization specific
-cookbook for agile implementation was created. The cookbook resolved
-misconceptions, presented differences between the old model and agile, and
-described agile development in traditional terms [P23]. Other agile material
-organizations produced were short videos [P32], and a branded theme for the new
-process [P8]. One primary source provided the organization specific Agile Policy
-as an appendix [P26].
-
-% Physical spaces
-
-In many cases it was mentioned that the work spaces needed to be rearranged to
-better allow the team centric model of agile [P35, P42, P48]. The possibilities
-to create a team area varied between cases. Some cases had trouble with physical
-space arrangements [P36], in some cases spaces could be arranged with some
-bargaining with the facilities department [P6, P38], but some organizations
-invested readily in good facilities [P46]. In addition to the work areas some
-organizations took the chance to improve common areas also [P22, P46].
-
-\clearpage
-
-\subsection{Challenges in transformation}
-
-Any organizational transformation that involves numerous individuals will face
-some challenges. As stated by Research Question 3, we investigated what
-challenges large organizations introducing agile methods may face.
-
-The challenges presented in this section are factors that primary studies
-described hindering the progress of the transformation. Some of the challenges
-presented here do not explicitly relate to the transformation process, but are
-rather general difficulties in agile implementation. These issues are still
-included as settling the agile implementation can be seen as a part of the
-transformation process. Some solutions for challenges are also presented, but
-they should be considered as ways of describing the challenges rather than
-direct success factors.
-
-\begin{figure}[!b]
-  \begin{center}
-    \includegraphics{graphics/challenges_summary.pdf}
-    \caption{Overview of challenges in agile transformations}
-    \label{fig:challenges_summary}
-  \end{center}
-\end{figure}
-
-The challenges found in the primary studies are summarized in figure
-\ref{fig:challenges_summary}. The challenges are organized on a high level into
-seven categories within three themes. The categories are discussed subsequently
-in this section. First we present generic transformation issues that may apply
-to any organization. As the second theme challenges relating particularly to
-sizable organizations are discussed. The final theme studied is challenges
-encountered when agile development needs to interface with other functions
-involved in a system development life cycle.
-
-It is to be noted that we do not emphasize any of the listed categories or
-challenges above another. The challenge categories we present were found evenly
-in the set of primary studies. We included only challenges that had been
-mentioned in several studies, or were particularly emphasized in fewer studies.
-Challenges with only sporadic mentionings were considered less significant and
-were left out. There was no further evidence that any challenge would be more
-significant than another.
-
-\subsubsection{Change resistance}
-
-In an organizational transformation working habits and roles need to change.
-However, people are usually disinclined to change and some resistance should be
-anticipated in any change process. People will not be willing to change unless
-there are good reasons, and the change is perceived easy enough. It is difficult
-to attain a buy-in for a change, and even organizations that have a flexible
-culture will face resistance [P10]. It should be expected that not everyone will
-be willing to change, even so that some employees will never adapt to the new
-model [P6]. During a transformation period objections to change may become a
-major reason for loss in time and productivity [P29]. The main types of change
-resistance discovered in the primary studies are summarized in table
-\ref{table:challenges_changeresistance}.
-
-\begin{table}[b]
-    \centering
-    \begin{tabular}{ l l }
-        \toprule
-        Types of change resistance           &  Described in  \\
-        \midrule
-        General resistance to change         &  P14, P18, P22, P44, P45   \\
-        Skepticism towards a new model       &  P5, P8, P9, P18, P36, P46 \\
-        Top down mandate creates resistance  &  P2, P12, P37, P49 \\
-        Management not willing to change     &  P2, P3, P6, P49 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Types of change resistance reported in studies}
-    \label{table:challenges_changeresistance}
-\end{table}
-
-Numerous reasons were given for why change resistance emerged. For
-instance, Fecarotta [P14] described the organization of Boeing as risk-averse
-and cautious, which slows down any change. Any change was further hindered by a
-deeply rooted status quo and high employee retention [P14]. In some cases it was
-reported that people worried about new roles and responsibilities that the agile
-model might bring [P44]. For instance, testers were worried that they would need
-to take on cross functional tasks, which would be outside their area of
-expertise [P18]. Yet another reason for change resistance was the move from own
-offices to team areas [P22]. People were also feeling being monitored because of
-the increased level of interaction within the team and between project
-stakeholders [P45].
-
-In these cases change resistance was mitigated for instance by education [P45]
-and one-to-one discussions [P22]. Change resistance was also softened by first
-engaging only teams willing to try agile, so that insights could be gathered to
-ease the continued transformation [P45].
-
-Another common form of change resistance that was reported was skepticism and
-distrust in suitability of agile. Seffernick describes how management did on one hand
-acknowledge benefits of agile, but on the other explained away the possibility
-to apply it due to contract reasons, the current matrix organization, and other
-organizational practices [P46]. Skepticism was often created by other similar
-misconceptions, including that agile does not work for complex products [P5,
-P36], agile needs to be implemented in a prescriptive by-the-book way [P9], that
-frequent meetings will cause overhead [P18], and that agile equals avoiding
-governance and working without a plan [P8].
-
-% For instance, teams may not be willing to try
-% new approaches unless they have been proven [P6].
-
-% Resistance is to be expected --> there are many different stakeholders [P41]
-
-% * No value seen by some groups, e.g. mainframe or maintenance [P49]
-
-The way the transformation is initiated will affect how change resistance will
-show. In several cases the change initiative came from management, but when
-presented wrong the initiative becomes a mandate that people are not receptive
-for. Spayd [P49] summarizes this as: ``Organizations do not change merely
-because the boss says so, at least not in the way that is intended''. In that
-organization the introduction of CMM had been carried out consistently by a
-mandate, but the collaborative nature of agile did not mix well with such a
-mindset. A top down mandate may also dilute the understanding of the reasons for
-starting the transformation, and the understanding of agile development overall
-[P2]. O'Connor describes how team members became skeptical towards agile when
-implementing mandated changes did not bring any visible benefits [P37].
-A problem relating to top down mandate is that if management does not speak out
-a goal towards using agile, developers may feel that the agile methods may be
-replaced by something else at any time [P12].
-
-The direction in which a tension between management and development is built may
-also be opposite. In cases where the change was emerging bottom up the
-reluctance to change was not in the team members but management, causing that
-significant organizational change could not be raised above the team level. For
-instance, Atlas [P3] describes how executive support would have been required to
-extend the agile process to involve the product management office and separate
-quality assurance groups. In this case, when managers %for multi team projects
-were not involved in the transformation, the agile model was restrained from
-spreading beyond the level of development teams [P3]. Lack of middle management
-support for change and disinclination to change management culture were seen as
-some of the most serious problems in the transformation [P2, P49]. Middle
-management is in a position to undermine the entire transformation, and may do
-so if they do not participate in, and thus understand, the agile method. Agile
-brings changes to some management roles [P2], and lack of understanding of agile
-development will leave managers feeling left out [P6].
-
-\subsubsection{Lack of investment in transformation}
-
-It is clear that changing the way of working will create some costs for the
-organization. Problems will ensue if there is no intention to allow resources
-for the transformation. Table \ref{table:challenges_lackofinvestment} presents
-investments and resourcing arrangements that were withheld, and which caused
-problems in transformation.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ l l }
-        \toprule
-        Factors of investment and resourcing  &  Described in \\
-        \midrule
-        Lack of coaching                      &  P1, P6, P23, P45, P47, P49  \\
-        Lack of training                      &  P11, P20, P45  \\
-        Excessive commitment of people        &  P37, P42, P49  \\
-        Old commitments are kept              &  P11, P21  \\
-        Physical spaces can not be rearranged  &  P6, P14, P29, P36, P38, P49  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of investment and resourcing factors that affect
-             transformation negatively}
-    \label{table:challenges_lackofinvestment}
-\end{table}
-
-% Lack of training and coaching
-
-Training and coaching are direct investments in transformation and their lack is
-an evident problem. These activities were still under budgeted, which
-created difficulties in the transformation [P11].
-Hajjdiab [P20] describes how reluctance of management to invest in training
-caused teams to be ill prepared for the agile transformation effort, which
-eventually resulted in ending the use of agile methods. As a less dramatic
-outcome of too little training was lower motivation [P45].
-
-Arranging proper coaching was a particular problem in many organizations. It is
-critical to coach teams in their real work environment as a proper change in
-mindset is difficult to achieve by only attending training sessions. [P9, P23]
-In large
-organizations where numerous teams needed to be coached the demand often
-exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
-was also attributed as a reason why the success of pilot teams could not be
-repeated when agile was adopted more widely [P9, P45].
-Silva and Doss [P47] described that when it was hard to fill the coaching
-positions people less seasoned in agile were appointed as coaches, which risked
-that agile practices would not be taught correctly.
-
-% Add quote?
-% * No coaching --> Applying techniques improperly --> Damage
-% [P9]
-% Without proper coaching however, the success of these informal teams met with
-% mixed results over the longer term.  In retrospect, many of these teams did
-% damage to themselves by improperly applying Agile techniques, particularly when
-% command and control mechanisms were still in place.
-
-% Old work and commitments, schedule pressure
-
-Even though the organization was in a state of transformation the workload of
-teams was not always adjusted to facilitate the process.
-Some organizations started their agile journey in a state where people were
-overcommitted. It was later realized that when people are overloaded they will
-not be able to changing their behavior and learn new ways of working [P37, P42].
-Spayd [P49] also describes how senior management pressed teams to deliver what
-the customer had been promised, regardless of what the current velocity of
-development predicted. For one part the pressure to deliver interfered with the
-transformation, but it was also forcing a change in the old way of working
-[P49].
-
-In one case management forced people to remain committed to firm deadlines, even
-midst in the organizational transformation [P21]. The engagement in old
-commitments and tasks resulted in ignoring new practices introduced by agile,
-and eventually the agile method was breaking down [P21].
-Cowan [P11] describes how time was allocated for tending to old commitments.
-However, even when there was preparation for extra work people with specific
-expertise became overloaded. After time it was realized that overcommitment must
-be reduced by allowing to shift excess work for later time [P11].
-
-% Physical spaces, challenge of finding a seating
-
-In some cases the old way of working had people seated physically distributed,
-in opposite to what agile methods typically suggest. Office spaces needed to be
-arranged so that the entire team could work in a single room, but it took time
-and effort, and it was not always possible [P6, P14, P36, P38]. Lewis and Neher
-describe how dedicated rooms for teams could not be arranged, and team events
-such as daily stand-ups required booking a conference room [P29]. In another
-case the estate organization shut down the teams' attempts to modify working
-spaces [P49].
-
-\subsubsection{Agile is difficult to implement}
-
-A challenge most organizations came across was that implementing the agile model
-was found to be difficult. An experienced software team may do well in training,
-but when the time comes to apply agile techniques in practice the team may get
-lost [P21]. Viewpoints on the difficulty of correctly implementing an agile
-model are summarized in table \ref{table:challenges_difficulty}.
-
-% The difficulty in implementing agile is evident in misconceptions of the agile
-% model, when by-the-book implementations prove to be insufficient, and when the
-% agile methods were applied badly to the organization.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Type of difficulty in implementation  &  Described in \\
-        \midrule
-        Misunderstanding agile development concepts  &
-                P6, P26, P37, P38, P42, P44, P45, P48, P49      \\
-        Literature does not provide enough guidance  &
-                P5, P6, P10, P13, P21, P27, P30, P45, P49     \\
-        Agile practices are customized badly   &
-                P9, P29, P44, P46  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of viewpoints why agile implementation is difficult}
-    \label{table:challenges_difficulty}
-\end{table}
-
-% Misconceptions of agile
-
-There were many examples on problems caused by misconceptions of what agile
-software development is. On a general level, Bang [P4] describes how the values
-of the agile manifesto were not understood, and agile practices were carried out
-without understanding their purpose. In one case management saw the purpose of
-agile simply being faster product delivery [P9].
-
-On a team level there were several examples of misunderstanding agile
-development. Cases describe how agile was seen as freedom to hack without
-documentation [P26], that development could be done without design tasks [P37],
-and that agile meant that everyone should become a generalist [P42]. Schatz and
-Abdelshafi describe how teams presented unfinished work at reviews and ignored
-the principle of showing only completed increments, which resulted in a backlog
-of bugs [P44]. In one case team members had perceived the introduction of agile
-as a means to squeeze out more efficiency [P45]. In addition, the flatter
-organization was seen as fewer career opportunities, and pushed team members to
-internal competition for visibility [P45].
-Still one misunderstanding of agile was due to viewing it only through the tools
-used, such as project management software [P48]. Superficial focusing only on
-tools themselves and not the reasons behind their use led to frustration [P42].
-
-In some cases the misunderstanding of the agile model was described as ``doing
-mini waterfalls'' [P6]. Managers were using Scrum terminology, but at the same
-time committing unreasonable workloads on teams [P6]. In another case a
-waterfall nature was evident when task estimates were given as hours left, and
-task breakdowns disconnected from what was really being done [P27].
-
-As a final example of misconceptions, some organizations considered agile being
-a solution to all problems [P9]. Success was declared prematurely [P49] but
-expectations created by successful early experiments were not fulfilled [P38].
-Disappointment was granted when these hopes were not fulfilled, and instead
-implementing the agile model was found to be hard and time consuming.
-
-% By the book is not good
-
-Even if there was no misunderstandings the correct agile model was hard to
-implement. Several cases describe how an agile model was hard to learn from
-literature [P21, P27], especially when it comes to implementing it on a large
-scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
-where we could find easy answers on how to do things''.
-Schnitter and Mackert [P45] report that theoretical considerations on how to
-scale up the agile methods were not good enough, and that product managers and
-architects struggled when several Scrum teams were working concurrently.
-Another case concluded that all practices suggested by XP did not fit enterprise
-scale development, and some practices required customization [P49].
-
-% In a large organization with several different teams and projects it is clear
-% that one single model can not fit all needs [P41]. The model of
-% agile should not be prescribed in a strict way, as it would impose a single
-% model in an organization with several individual teams and projects [P41].
-
-% Check this paragraph
-Further indications were given that the advice from literature will not ferry
-all the way. Benefield [P6] describes how it was difficult finding a balance
-between prescribing a by-the-book implementation which may put people off, and
-giving too much freedom in the agile methods which may weaken core practices.
-The reality where the practices needed to be applied was described as messy in
-comparison to the ideal circumstances presented in literature [P10]. Because of
-this it was difficult to choose a model to start with and gain buy-in for [P10].
-Difficulties in choosing a initial model were also due to variances perceived in
-the agile approaches suggested in literature [P30].
-
-% Where would this belong?
-% Organizations had also bad experiences if the implementation of the agile model
-% was being prescribed. For instance, implementing Scrum based on checklists of
-% practices would not work [P41]. Benefield described that it was challenging to
-% find a balance between independence for teams to choose their way and diluting
-% the value and structure of Scrum [P6]. Giving room for team autonomy is one of
-% the critical facets of agile, but it must be ensured that teams do not discard
-% practices that add value and structure.
-
-% Customizing badly (mapping to old model)
-
-The difficulty of agile and misunderstandings were also evident in cases where
-the methods were customized badly. As by-the-book implementation often was not
-feasible, the agile method was tailored to suit the needs of the organization.
-However, in some cases tailoring meant skipping some practices, which lead into
-problems. In one case certain individuals were allowed to ignore some core
-elements of Scrum, which came to turn the teams decision making into a variant
-of command and control [P9].
-Lewis and Neher [P29] describe how there was temptation to strip some practices
-and enhance others. Previous attempts had proven that one of the reasons for
-agile implementations to fail was deviations from the process, because of which
-the agile mindset did not take root [P29]. A bad customization may lead to the
-team taking only practices that reflect their current needs, and therefore
-failing to achieve a real change in process and mindset [P9, P46].
-Another example of bad customization that counteracts change is replacing agile
-vocabulary in favor of familiar terminology from the old model [P44]. An
-important benefit of introducing new vocabulary is that it stimulates new ways
-of thinking [P44].
-
-\subsubsection{Coordinating work between many teams}
-
-In large software projects work must inevitably be distributed between many
-teams. One of the most prominent transformation challenges was the difficulty in
-creating a model to coordinate work between many teams. Table
-\ref{table:challenges_coordinatingteams} presents challenges that arose when
-agile was implemented in environments where many teams need to collaborate.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
-                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
-        \toprule
-        Problems in environments with many teams  &  Described in \\
-        \midrule
-        Interfacing between teams is difficult     &  P9, P10, P13, P17, P24, P26 \\
-        A model with autonomous teams is not good  &  P7, P13, P29, P51  \\
-        Distributed setting causes additional problems  &  P29, P45, P48  \\
-        Technical integration and consistency      &  P5, P24, P26, P27, P50  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of challenges in a setting with many agile teams}
-    \label{table:challenges_coordinatingteams}
-\end{table}
-
-Many cases reported that introducing agile practices soon yielded improvements
-on the team level, but as a part of a large transformation teams also needed to
-change their external behavior.
-The initial success was followed by challenges when teams needed to work with
-other teams and as parts of the larger surrounding organization [P13, P17, P24].
-Introducing agile had created flexibility on a team level, but it could not be
-fully utilized as the surrounding organization was not responsive enough [P26].
-The roll-out of agile had not removed dependencies, and the dependencies
-made managing development difficult [P9, P10].
-
-% Independence of teams
-
-Some organizations attempted to initially create a model where teams operate
-autonomously, which is well in line with agile principles. However, a number of
-issues arose from independence. For instance, teams needed to find a balance
-between the their own goals and broader goals of the entire organization, but
-some teams chose to focus only on their own goal [P7]. Coordination was
-obstructed by independent teams that did not respect the larger context [P13].
-Coordination was also difficult when teams were working independently for
-different customers but the applications being built were interdependent [P29].
-One case reported that even allowing teams to have different sprint durations
-created delays in delivery [P51].
-
-Further coordination problems were encountered when scaling up agile over many
-sites. Operating in a distributed environment makes communication more
-difficult, and causes problems when frequent and effective communication needs
-to be achieved. Distribution had negative effects, such as missing kick-off
-meetings, reduced feeling of proximity when telecommunication is necessary, and
-difficulty in arranging frequent meetings due to time zone differences [P45].
-Agile project management was also problematic in a distributed setting. In a
-waterfall model separate parts of projects could be isolated to different sites,
-but the agile model does not allow such strict slicing of projects [P29]. In one
-case it was simply admitted that a distributed organization will impose
-additional burden on communication and require additional care in the process,
-but distribution and agile would still be used together [P48].
-
-Also technical problems relating to inter-team coordination were reported.
-Integrating the products of teams was seen as a problem in some cases [P5, P50],
-and lack of standardized build scripts was mentioned as a problem in one case
-[P27]. There was also problems in synchronizing definitions of software
-interfaces between teams, and dependencies in code caused problems in larger
-features [P26]. One case reported that team centricity in the agile model
-created a fragile architecture, divergence in coding style, and even distrust
-between teams [P24].
-
-Moore and Spens [P34] describe the progression of the above mentioned
-coordination problems. At first cross-team dependencies were attempted to be
-reduced, but it became evident that the dependencies were an inherent part of
-the project. A traditional approach to managing dependencies created excess
-work, reduced independence and flexibility of teams, and created contract based
-and adversarial relationships within the organization. There was some success
-when up to 5 teams were coordinated using Scrum-of-Scrums, but that model could
-not be scaled to a global level. The main problem was teams that were thinking
-too much inside the team walls. When the new practices were scaled up the
-communication channels narrowed and created communication breakdowns. Problems
-with collaboration, dependencies, and integration were solved in the presence of
-individuals having a proactive and open mindset towards working over team
-boundaries. It was concluded that a balance is needed between completing new
-stories from the team backlog and maintaining overall stability of the
-application. [P34]
-
-Many studies presented solutions for the inter-team challenges, which also
-highlight the nature of the challenges themselves. The general model of solving
-inter-team problems was introducing an additional layer of coordination, such as
-a Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
-customized release framework [P13]. In one case integration problems were
-controlled by prioritizing code quality and maintaining existing functionality
-over new features [P13]. Finally, the choice of tools was reported to have an
-impact on the coordination effort [P17].
-
-\subsubsection{Issues specific to large organizations}
-
-In addition to inter-team coordination there were also other challenges that
-should only emerge in a sizable organization.
-Most importantly, additional layers of management are usually necessary to steer
-large groups of people. The problem is that agile methods provide little
-guidance on how to organize management.
-Problems may also arise in a large environment if the schedule of agile adoption
-varies from team to team. Internal fragmentation is a further problem that may
-be encountered in large organizations.
-Table \ref{table:challenges_largescale} summarizes these challenges.
-
-\begin{table}[b]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.56\textwidth}
-                     >{\raggedright\arraybackslash}p{0.36\textwidth} }
-        \toprule
-        Challenges relating to large scale  &  Described in \\
-        \midrule
-        Management roles unclear in agile       &  P2, P11, P27, P30, P38, P49, P52 \\
-        Using old and new models side by side   &  P1, P8, P10, P26, P29 \\
-        Management uses waterfall model         &  P3, P6, P35, P38, P45, P46 \\
-        Duplicating management with two models  &  P20, P48 \\
-        Agile models diverge between teams      &  P8, P9, P38, P43, P48 \\
-        Internal boundaries in organization     &  P10, P11 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of challenges relating especially to large scale}
-    \label{table:challenges_largescale}
-\end{table}
-
-% Roles of managers
-
-The existence of additional management positions may pose problems to agile
-processes that emphasize self organization.
-Especially the role of middle management was unclear in agile methods [P2]. This
-is problematic as an agile transformation requires cultural change particularly
-on the middle management level [P2].
-It was described that managers needed to resist the tendency to take command and
-control and allow room for self organization, but the change in role was
-difficult to achieve for the people involved [P11, P52].
-Nielsen and McMunn [P30] describe how the project management group had
-previously worked through big up front plans and competing of resources, but
-those ways would need to dramatically change.
-
-Several other problems related to management roles were also presented. For
-instance, Spayd [P49] describes how managers were left outside the roles offered
-by the new agile model. In another case when managers were appointed as Scrum
-masters developers felt being micromanaged [P27]. This was partially because of
-vague understanding of the agile method [P27]. It was also described how mixing
-the role of project manager and Scrum master created a conflict in interest, and
-turned the Scrum manager's role into one of policing the team instead of
-supporting the team [P30]. In some cases problems with manager positions were
-solved by phasing out roles relating to the old model, and replacing them with
-new positions more suitable for agile [P38, P52].
-
-% Using two models side by side
-% Part of the organization is in old mindset
-
-In most cases a large organizational transformation proceeded gradually, and
-during the process it was possible that the new agile model was used in parallel
-with the old methods. The arrangement of using the old and new models side by
-side was seen as generally problematic [P1, P26], and causing tensions on all
-organizational levels [P10]. Ongoing projects had been set up with previous
-development methods, and the agile methods needed to be arranged to fit along
-with them [P8, P29]. A particular problem in collaborating between projects
-using different models was that in Scrum the software design was fleshed out
-over time as sprints progressed, but the waterfall model would require detailed
-design to be locked early in the project [P29].
-
-Even when agile had been taken into use in development there were problems with
-management working according to the old waterfall model. O'Connor [P38]
-described management as ``focused on meetings and big up front project plans''
-despite agile being in use. In another case management was losing confidence in
-agile because familiar reports on cost analysis and percentage complete were not
-produced [P35]. As Scrum teams did not commit to fixed schedules they were
-considered unreliable according to the original model [P45].
-
-Some cases reported that development efforts were still controlled on a top
-level by a project management office (PMO). The PMO was described as entrenched
-in a waterfall paradigm, and the top-level rigidity was causing friction in the
-agile adoption [P3, P6]. The PMO was seen as a hub and bottleneck controlling
-all aspects of projects, and its central structure had to be dismantled for the
-organization to become agile [P46].
-
-There were also problems with duplicating management when two models were in
-effect. For instance, agile teams were required to comply with current
-procedures producing excess documentation and stepping through approval gates
-[P20, P21]. The bureaucracy of the previous traditional model was still enforced,
-and management was not convinced to lighten the process. Another case describes
-that after introducing the agile method teams were required to fill in templates
-of two processes [P48].
-
-% Challenge: Differing vocabularies
-
-In addition to teams using agile and waterfall models side by side a large
-organization makes it possible that the agile interpretations of teams diverge.
-When many teams implement agile without consistent guidance friction and
-fragmentation may emerge [P9]. The organization may require moving people
-between teams from time to time, and therefore it would be desirable that no big
-differences in teams' agile cultures exist. Divergence in process would create
-risk of increased costs when relocating people [P38].
-Further, if processes of teams diverged management had difficulties in
-forecasting and comparing work of different teams [P38, P48]. To overcome
-problems with divergence in agile models some organizations defined standards to
-follow [P8, P43].
-
-% Challenge: Previous organization in silos is haunting new model
-% Is this good??
-In some cases the initial organization had internal boundaries or specialized
-knowledge was stuck silos. Such settings caused problems in the agile
-implementation.
-Cloke [P10] describes how use of Scrum revealed an internal segmentation where
-teams operated with differing priorities and agendas. This segmentation also
-hampered the initial effort to use Scrum in a larger context [P10].
-Cowan [P11] describes how projects needed specialized skills that were scarce
-and people were often relocated to match the needs of the moment. Too much
-reorganization made it difficult for teams to plan ahead [P11].
-
-%% This??
-% A dispersed matrix organization causes problems when people are on one hand
-% kept assigned to many projects simultaneously, but on the other hand functions
-% such as database administration, architecture boards, and project management are
-% centralized [P14].
-
-%% ?? -->
-% Large organization put agile methods under the test of scaling up, and problems
-% were encountered.
-% In the agile model it was hard for managers to get an overview of the progress
-% of the release [P17]. The status of features was hard to tell because different
-% aspects of features were sepatated across teams [P17].
-%
-% One case reported use of agile across hundreds of teams, but management could
-% not get a roll up of the progress without having strict rules on how status was
-% reported [P48]. Schnitter and Mackert [P45] report that a maximum workable size
-% in a project managed by Scrum was 130 people including 7 development teams.
-% Smits and Rilliet [P48] report that providing executive overview of the
-% development programs was limited by the tools used to manage agile projects.
-
-% --
-
-% Add quote ?
-% * Especially higher organizational levels were not using agile --> Problem
-% [P3]
-% Scrum still exists mostly in a way that avoids conflict with the larger
-% corporate organization, and there has been little progress in addressing larger,
-% organizational impediments.
-
-% [P18]
-% Real Agile teams will experience resistance when they start working with
-% non-Agile interfaces. Lack of understanding in ‘why to help create a cadence?’,
-% ‘why involve the customer more?’ will cause resistance in the change process.
-
-% --> ``Part is in the old mindset''
-% * Problem: Management attempts to use a ``resource pool'' of individuals rather than coordinating teams
-% [P6]
-% we face the constant threat that restructuring and the instinct from management
-% to track and leverage resources at the individual level (for example
-% organization-wide resource pools) will destroy the team dynamic and adversely
-% affect productivity.
-
-\subsubsection{Integrating requirements management and quality assurance}
-
-TODO:
-* Mieti rakenne
-* Mieti onko seuraava ``other than development'' hyvä otsikko
-
-Development is traditionally preceded by requirements engineering and succeeded
-by quality assurance (QA) activities. Agile methods suggest that essential parts
-of these activities would be performed within the development team. However, for
-large systems the effort required for managing requirements and QA may be too
-heavy for individual teams to handle. For instance, a single development team
-can manage and elaborate items in its own backlog, but a large feature may
-require work across many teams. The feature needs to be divided to team specific
-items on a higher level, and the process must be managed somehow.
-Table \ref{table:challenges_requirementsandqa} presents challenges in
-implementing additional requirements management and quality assurance required
-in large development projects.
-
-\begin{table}[t]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
-                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
-        \toprule
-        Challenges in requirements engineering and quality assurance  &  Described in \\
-        \midrule
-        How to accommodate high level requirements management  &  P24, P31, P39, P45, P51  \\
-        Problems with breakdown of requirements        &  P1, P17, P33, P48  \\
-        Creating and estimating user stories is hard   &  P5, P11, P27, P30, P33, P37 \\
-        Handling coexisting long and short term planning  &  P9, P10, P13, P26, P31, P38 \\
-        Accommodating performance, load, and memory testing  &  P17, P28, P31, P48 \\
-        Lack of automated testing slows down           &  P10, P11, P17 \\
-        New requirements management practices affect QA  &  P5, P48 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of requirements engineering and quality assurance related
-             challenges relating in the agile model}
-    \label{table:challenges_requirementsandqa}
-\end{table}
-
-% Requirements engineering is difficult
-
-In agile methods requirements engineering is usually informal and closely tied
-to the development team. However, large development projects will demand
-additional high level requirements management. This is apparent in cases where
-complexity of the product will require additional layers of product management
-[P45], requirements are created by several stakeholder groups and development
-teams can not be in contact with all of them [P39], or teams do not have access
-to stakeholders due to a distributed setting [P24]. Agile methods need to be
-extended to include a process on how to get large requirements into team
-backlogs [P31].
-
-Requirements were in some cases first defined on a high level in marketing
-requirements documents (MRD) [P1, P31] or functional specifications [P33].
-These high level requirements needed to be elaborated to be used by agile
-development teams. The requirements breakdown in it self was found to be
-difficult [P1, P33], and also when to do the breakdown and to which detail level
-[P17].
-Product managers and business analysts were struggling in creating high level
-requirements [P33, P37, P51], and teams were struggling in breaking them down to an
-estimable size [P33]. Smits and Rilliet [P48] describe how high level product
-management delivered requirements in too large chunks resulting in development
-teams spending large amounts of time on defining suitably sized stories.
-
-Several cases highlighted that much learning was needed to master the new
-process of creating user stories, both on product management and development
-levels [P27, P30, P33, P37]. There were problems such as ambiguity in
-requirements [P5, P11], and effort estimation for stories was hard to do well
-[P27, P37].
-
-% As an integral part of the software development lifecycle, requirements
-% management will be affected by the new agile process. In large organizations
-% product and requirements management was initially done in a traditional way, but
-% it was contrasted by the agile process that required iteratively supplying small
-% requirements to development [P51].
-
-% In one case the teams were overwhelmed by reprioritization and requirements
-% bartering when stakeholders learned that requirements could be rearranged
-% flexibly [P31].
-
-% Long term planning
-
-High level requirements work and estimation was problematic as there was a gap
-between short and long term planning [P9, P10]. Typical agile backlogs would
-give only short term visibility, so organization specific practices for long
-term planning had to be created [P26, P31]. Several cases described that care
-had to be taken to avoid long term planning becoming a prescriptive practice. In
-order to preserve the agility of development a schedule driven approach was
-avoided [P10], sensitivity was used when considering setting milestones [P13],
-and striving for more accurate estimates was not let to become an excuse for
-gathering requirements up front [P38].
-
-% Problems with enterprise or external QA
-
-In an enterprise setting quality assurance tasks will require effort beyond the
-general agile convention of doing testing within a cross-functional team.
-Similarly as with requirements engineering, the agile model may need to be
-extended to accommodate additional testing activities [P31]. Organizations may
-have existing separate QA teams that must be coordinated to work with
-development teams [P17, P30, P48]. Integrating enterprise scale QA will pose
-some additional challenges to agile methods.
-
-Full quality assurance of a system requires special testing such as performance,
-load, and memory testing, but agile methods lacks focus on them [P17]. These
-testing activities can not be done within boundaries of user stories, and
-require more resources than teams can spare [P31]. When testing tasks overlap
-team boundaries it is sensible having a separate teams for the tasks, but
-coordination practice between specialized QA teams and development teams needs
-to be defined [P28, P48].
-
-Some cases reported on problems with lack of automated testing. For a large
-system the lack of automated tests caused excess testing work and late discovery
-of defects [P10, P11, P17].
-
-Still one challenge in QA was the relationship to requirements engineering. The
-QA functions were struggling if there was problems such as ambiguity or
-ineffective breakdown of large requirements [P48]. In a waterfall model an
-extended time was reserved for QA at the end of projects, which allowed to
-retroactively resolve ambiguities in requirements [P5]. However, ambiguity
-became an impediment to QA when development worked in short increments and
-there was no extended period for stabilization [P5].
-
-\subsubsection{Integrating other functions than development to agile}
-
-Agile methods are typically considered as a set of activities closely related to
-software development. Cases describe how development teams were the first
-entities that were considered when introducing agile [P2, P17, P31, P38].
-However, development operates in a larger context and will need to interface
-with other organizational functions.
-Challenges were faced when exposing other parts of the organization to the agile
-model, and functions that were further away from development were resistant to
-change [P2, P18, P31]. A tension was growing between development and broader
-roles in the organization [P1]. The full benefits of the transformation would
-not be attained unless the entire organization was set to work along the same
-paradigm [P18, P42].
-
-Primary sources listed various functions beyond development that are involved in
-bringing products to the market. Functions interfacing with development and
-affected by the agile transformation are presented in figure
-\ref{fig:challenges_other_functions}.
-Marketing was a particularly frequently mentioned group [P5, P17, P28, P31, P38,
-P45, P48]. Other groups involved in the whole of product delivery were sales
-[P17, P31], infrastructure and operations [P10, P45, P46, P50], user experience
-and design [P9, P10, P28], documentation [P28], legal [P10], security [P10],
-customer services [P48], and finance [P48].
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{graphics/challenges_other_functions.pdf}
-    \caption{Functions affected by the agile transformation}
-    \label{fig:challenges_other_functions}
-  \end{center}
-\end{figure}
-
-Agile affects timings in the software delivery life cycle, what caused two types
-of problems. Firstly, the iterativeness of agile models was problematic when
-interfacing with other functions. Iterativeness changed the pace of delivery
-[P17], and forced design to focus on a shorter scope [P15]. Various groups
-feeding and supporting development had earlier committed once to a long term
-plan, but in the new arrangement they needed to adjust to incremental delivery
-[P48].
-Especially user experience (UX) and interaction designers did not welcome an
-incremental model [P6], and were struggling with maintaining the big picture in
-design while working in iterations [P9, P10].
-The infrastructure and operations teams were falling behind due to the increased
-delivery speed of development teams, and they were forced to update their way of
-working [P46, P50].
-
-Federoff and Courage [P15] elaborate on the user experience team's problems
-with the short time horizons of Scrum. Previously the UX team had supplied
-development with designs within a long release cycle, but Scrum required
-completing features within weeks. The problem was that that the UX work required
-a holistic view instead of an incremental one, and the design process did not
-fit well into the time frame of a sprint. The UX team was overloaded and
-discontent, but the situation was fixed by refining the schedule for interactions
-with development teams. [P15, P28]
-
-The second type of problems was that some release activities would inevitably
-have a long lead time but required committing to a set of functionalities in
-good time. The agile model's emphasis on short time horizon and flexibility in
-prioritization did not mix well with such activities.
-For instance, it would take time to prepare printed marketing material and press
-releases, but they needed to accurately present the upcoming features of the
-product [P38].
-
-Maples [P31] describes that marketing needed three months for preparing the
-product launch, but the agile process allowed development to keep changing the
-content of the product. As a result marketing was struggling with creating
-material and campaigns without knowing the exact product features. Further, the
-processes of acquiring export clearances and licensing authority required that
-the the feature set of the product was known well on beforehand.
-As a result of these requests development felt that their process was being
-slowed down. The issues were solved by agreeing on that marketing could at any
-time request a release in 90 days, and development would be able to commit to a
-fixed feature set to be delivered [P31].
-
-% Problems with HR
-
-Problems in adopting agile were even showing in functions outside the software
-delivery life cycle. In several cases human resources (HR) was a support function
-mentioned to work against the agile adoption. In order to gain full results of
-the transformation the entire organization should to be aligned, including HR
-[P38, P42, P49].
-Especially the rewarding practices set by HR were seen as problematic for the
-agile model. Rewarding was tied to personal performance, which was acting
-against team centric thinking and the agile model in general [P3, P6, P49].
-One case reported even on a practice of tracking failures down to individuals
-[P40].
-
-\subsubsection{Other challenges in transformation}
-
-There were still some other challenges in the transformation that do not fit in
-the categories presented above. Two particular problems are worth mentioning.
-Firstly, there were certain factors that pulled people back to the old ways of
-working, and secondly, excessive enthusiasm in agile had negative effects.
-
-% Falling back to old model
-
-In several cases challenges in the transformation resulted in people falling
-back to the old model of working. In some cases it was only a temporary struggle
-while learning agile practices, but in other cases the old model displaced
-agile.
-Development work has to go on during the transformation but there will be new
-things to learn for the team. Stress caused by the combination of schedule
-pressure and much change at once will pull people back to the old way of working
-[P11, P18, P38].
-Schatz and Abdelshafi tell how even subtle trouble may risk the transformation,
-as people will always look for reasons to revert to familiar behavior [P44].
-Teams without adequate training were struggling with applying agile practices
-correctly, and the challenge the new practices posed made people to abandon them
-and return to the ways they know [P6, P49].
-
-In some cases agile was seen as an overhead, and was therefore abandoned. For
-instance, as new practices were being introduced there was a decrease
-performance, and when teams realized that the benefits are not immediate they
-started reverting to the old model [P21].
-Evans [P12] describes that new senior appointments changed the management to be
-less favorable towards agile, and a waterfall development model started to
-return.
-
-% Agile zealots
-
-A phenomena that troubled some organizations was over-enthusiasm towards agile
-methods. In several cases there was mentionings about change leaders or teams
-becoming agile zealots. For instance, Evans [P12] describes how some members of
-the agile community were becoming too evangelic, which made people take sides
-for or against agile development. In another case team members started out with
-an overly eager attitude, but when agile did not deliver improvements the
-interest faded and started to go back to the old process [P20]. Attempting to
-implement agile too strictly may cause conflicts, and especially in a large
-organization the change leaders need to maintain a collaborative attitude
-towards various groups in the surrounding organization [P6]. Introducing agile
-methods does not guarantee success, and therefore they should not be followed
-blindly [P4].
-
-% When a release schedule with fixed deadlines that business required was combined
-% to agile managers started to behave authoritatively and the self organization of
-% agile started to break down [P31]. % (Skeptics used this as a sign of that agile can not work)
-
-% This??
-% % People do not speak their minds
-% In one case a passive-aggressive organization culture was described as a
-% challenge. Ranganath describes how people did not speak their minds in meetings,
-% thus creating a false sense of agreement when talking about how the organization
-% should develop. Because of this resistance and concerns were left growing, and
-% acted against achieving a successful transformation. To to avoid these issues
-% change leaders worked with team members on an individual level to win confidence 
-% and make their voices heard [P40].
-
-\clearpage
-
-\subsection{Success factors in transformation}
-
-The analyzed primary studies brought up several factors that contributed in
-success in the transformation. As stated by Research Question 4, we studied what
-factors were reported to facilitate the transformation process. We present
-factors that the primary studies particularly highlighted as being important in
-advancing the transformation process.
-
-\begin{figure}[!b]
-  \begin{center}
-    \includegraphics[width=1\textwidth]{graphics/success_summary.pdf}
-    \caption{Overview of success factors in agile transformations}
-    \label{fig:success_summary}
-  \end{center}
-\end{figure}
-
-Figure \ref{fig:success_summary} summarizes success factors evident in the
-primary studies. The success factors are organized into ten categories within
-three themes, which are elaborated in this section. The first theme relates to
-management and leadership in the transformation. The second theme presents
-success factors that relate to finding a suitable agile model for the
-organization. The last theme includes success factors that are realized on the
-level of development teams.
-
-The discovered success factors did not exhibit strong differences in
-significance in respect to each other. Three categories of success factors can
-be considered of higher importance: management support, training and coaching,
-and choosing the agile model carefully. The remaining success factors should be
-considered having even significance.
-
-% * What is our standpoint on recommendations that experience reports give, but
-%   which were not necessarily tried in the organization?
-% 
-% * Some success factors are just calculated decisions that were made because
-%   the organization saw them as the best way to proceed. The question whether
-%   the decision really were critical for success is not proven. 
-% 
-% * Some success factors are compiled from ``what we learned'' or ``advice for
-%   others'' lists. We considered whether the advice was something the
-%   organization had tried in practice.
-% 
-% * For instance, implementing continuous integration and automated builds was not
-%   explicitly stated as a factor contributing to the success of the
-%   transformation, although it is unquestionable that such systems are critical
-%   for maintaining quality and enabling frequent releases.
-
-\subsubsection{Management support}
-
-The most prominent and clear success factor was having support from management.
-When the leadership of the organization is committed and supportive change is
-much more likely to happen. Table \ref{table:success_management} lists
-success factors related to management support.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.41\textwidth}
-                     >{\raggedright\arraybackslash}p{0.51\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Have management support  &
-                P8, P11, P16, P18, P27, P31, P33, P36,
-                P43, P44, P46, P47, P49  \\
-        Visible management support encourages employees  &
-                P9, P11, P24, P40  \\
-        Educate management on agile methods  &
-                P6, P8, P9, P11, P16, P29, P35, P46, P47  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Success factors relating to management support for agile}
-    \label{table:success_management}
-\end{table}
-
-% Management support in general
-
-Numerous cases made it clear that management support is an absolute necessity
-[P8, P18, P27, P31, P33, P36, P43, P47, P49].
-This was reflected in statements such as ``Adopting agile, or implementing any
-significant change, requires an executive’s sincere support.'' [P44],
-``Executive commitment was crucial to implementing massive change.'' [P16], and
-``Having upper management engaged, supportive and visible is critical for
-wholesale organization involvement with Scrum.'' [P11].
-
-Managers were seen to be in a key role in making the change stick, as they had
-the authority and power to remove impediments [P18]. Seffernick [P46] describes
-how a number of people attempted to explain away the applicability of agile
-methods, but the objections were overruled by the director's commitment to make
-agile work. Management support was similarly needed when tight release schedules
-had to be flexed in order to give room for the adoption process [P16, P27].
-Favorable management decisions were also critical when additional resources were
-allocated to training and coaching [P47].
-
-Visible involvement of management was also reported to motivate and encourage
-employees to adopt the new model [P40]. For instance, the CTO organized training
-sessions personally [P24], and the engineering VP frequently visited sprint
-demos [P11]. When the corporate level support for the agile initiative was
-showing teams adopted agile methods even spontaneously [P9].
-
-% Educate management
-
-In order to gain support for agile several cases highlighted the need to educate
-management [P8, P47].
-Without education managers might make harmful interventions in the process, or
-just stick to the old model [P16, P35, P46].
-Managers not understanding the principles of agile were feeling left out with
-the introduction of self organizing teams, which some times resulted in
-backlashes [P6]. Providing proper training was reported to correct the
-situation, and even create strong agile supporters in management [P6, P9].
-Training also cleared misconceptions that management had [P29], and helped
-create a consistent implementation of the agile model across the organization
-[P11].
-
-There were a few suggestions on how to involve management better. Foremost, a
-successful experience will likely rise the interest of management [P36].
-Mencke [P33] suggests giving executives tasks related to the agile process,
-so that they have better visibility of the new way of working.
-
-\subsubsection{Commitment to change}
-
-To make a large organization change the commitment to change must be made clear.
-People should be made to understand that there is no return to the old model.
-If and when the changes uncover problems they must not be allowed to set back
-the transformation. Table \ref{table:success_commitment} lists sources which
-present commitment to change as a success factor.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Make it understood that the change is non-negotiable  &
-                P11, P22, P48, P49  \\
-        Commitment to change  &
-                P8, P10, P43  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Commitment to change as a success factor}
-    \label{table:success_commitment}
-\end{table}
-
-% Make it understood that there is no return to the old model
-
-Articulating urgency at start will help getting the change under way [P49].
-Feedback of changes should certainly be listened to, but in the end it must be
-made clear that the old model can not be returned to [P22].
-For a case where the organization was changed all at once it was reported that
-the magnitude of change helped created the impression that the change was non
-negotiable [P11]. Firstly, as it was evident that a large change was to happen
-people felt encouragement and permission to move on from the old way, and
-secondly, the urgency eliminated wasteful discussions on whether Scrum was a
-good model or not [P11].
-When the organization culture is ingrown he change vision must constantly be
-reminded of, and each step towards the new model consider a victory [P48].
-
-% Make change necessary
-
-Transformation can be facilitated by creating mechanisms that force change.
-Spayd [P49] describes that senior management pushed hard on a mandate to have
-deliveries every 90 days. The mandate made change necessary, and while the
-strong pressure did not promote agile practices in every case, the drive for
-change was perceived good in general [P49].
-
-% Commitment (?)
-
-A large change will inevitably require strong commitment, but problems in the
-transformation can put the commitment to test. The agile model is introduced
-because of problems in the old model, and therefore there will be organizational
-issues uncovered during the transformation [P10]. People must not be demoralized
-when facing challenges, and the determination to change must be maintained [P8,
-P10]. It is not the agile model that has flaws, but it is merely exposing
-problems in the organization [P10].
-
-Ryan and Scudiere [P43] describe that management commitment was showing as
-strong focus on certain agile practices that had been chosen as non negotiable,
-and constant assessment that the practices work. The expectations were clearly
-communicated to the teams, and constant assessment made it clear that change is
-desired [P43].
-A strong commitment from management will assure teams that the change is the
-right thing [P11].
-
-% * Create a sense of urgency
-% * Find a common cause to motivate the change
-% [P11]
-% Find a rallying point.  You might not have a recession as a spur to receptivity,
-% but a common, critical cause to motivate change will make it easier to explain
-% how Scrum can contribute to the success  of that cause.
-
-\subsubsection{Leadership and alignment}
-
-The importance of leadership in change was described in many sources. Individual
-persons as well as task forces and communities acted as change leaders. A factor
-related to leadership was creating organizational alignment.
-Table \ref{table:success_changeleaders} summarizes factors related to leadership
-in change which benefit the transformation.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Have change leaders  &
-                P3, P4, P11, P16, P18, P31, P33, P42  \\
-        Have an agile community  &
-                P3, P12, P41, P42, P47  \\
-        Align the organization  &
-                P8, P18, P31, P42, P43, P46  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Success factors relating to leadership in change}
-    \label{table:success_changeleaders}
-\end{table}
-
-% Change leaders
-
-Organizing a large number of people requires some management and leadership.
-Similarly, transforming the way of working of a large group requires coordination
-and leadership [P31, P42]. In addition to the leadership provided by coaches,
-specific change leaders were also mentioned. Cases indicated the importance of
-having spokespersons for the change [P3, P4, P31]. Cowan [P11] describes how one
-person was strongly driving the transformation, and made an indisputable
-contribution for transforming the organization. Also Maples [P31] describes how
-the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
-described that the change was lead by a competent roll-out team, which had
-representatives from all parts of the organization [P16, P33]. Finally, the
-responsibility of line and project managers to act as change leaders was
-highlighted [P18].
-
-% Have an agile community
-
-The formation and influence of agile communities was reported to have a
-significant impact on transformation.
-The emergence of an agile community was reported as a success factor in the
-transformation [P3]. A community was also seen as indispensable as it has power
-to overcome impediments that would be too large for individuals to affect [P47].
-The agile communities achieved to raise awareness of agile methods in the
-organization [P41, P42, P47], and spawned eager followers who spread the word
-even further [P12].
-
-% Align the organization
-
-A factor in achieving change was creating alignment towards the common goal of
-introducing new development methods. It was seen as an important factor that all
-levels in the organization speak the same language and accept the change [P18,
-P31]. Alignment was built by promoting success stories and gathering lists of
-problems to tackle [P42, P46]. In one case complete alignment between teams and
-within management was considered as necessary to use the agile model in a large
-context [P43]. Another case highlighted creating and applying a structured
-roll-out process to uniformly introduce agile to a large number of teams [P8].
-
-\subsubsection{Use of piloting}
-
-The agile model was usually tried out in the organization before rolling it out
-on a general level. Beyond being a common practice, the use of pilot programs
-was also attributed as a success factor. Table \ref{table:success_pilots} lists
-sources discussing piloting as a success factor.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Have a pilot project at start  &
-                P3, P9, P14, P17, P31, P36, P38, P39, P47  \\
-        Gather insights from pilot projects  &
-                P8, P22, P27, P40, P45  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Piloting the agile model as a success factor}
-    \label{table:success_pilots}
-\end{table}
-
-% Have a successful pilot first
-
-Having a pilot project was reported as a significant success factor in several
-cases [P3, P9, P38].
-The pilot project gave proof that the agile model would be suitable for the
-organization and the general acceptance of agile was increased [P3, P17].
-The pilot also cleared disbeliefs of the suitability of agile for large
-organizations, and created acceptance for both development and management [P9,
-P31, P36].
-
-Pilots were especially important for gaining management acceptance. Cases
-reported that management gave approval for agile methods only after seeing
-successful pilots [P14, P47]. Prokhorenko [P39] also describes that seeing the
-example of successful pilots will make managers from other departments eager to
-try the new method, and thus help to spread the use of it.
-
-% Pilot provides information on how to do the rest of the transformation
-
-A particular benefit in piloting was that it provided knowledge on how to fit
-the agile method to the particular organization.
-A pilot project enabled the organization to get feedback on how teams and
-managers are best introduced to the new methods [P8, P22].
-Organizations met challenges in the pilot projects. However, the pilot projects
-served as valuable learning experiments providing insight in how to mitigate
-problems when the rest of the organization is transforming [P27, P40, P45].
-
-\subsubsection{Choosing the agile model to fit the organization}
-
-The choice of agile practices did affect the success of the transformation. It
-was common that cases described that it was beneficial to customize agile
-practices to fit the organization specific environment. As agile models do not
-prescribe practices strictly there was opportunities to make choices in the
-implementation. A few recommendations were given for choosing the agile model,
-such as using the same practices for all teams, mapping to the old model, and
-keeping practices simple. Table \ref{table:success_choosemodel} lists viewpoints
-relating to the choice of the agile model which were reported to provide
-benefits.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Customize the agile model to the needs of the organization  &
-                P8, P10, P11, P29, P31, P40, P41, P43, P45, P49, P51  \\
-        Conformity to a single agile model  &
-                P8, P11, P12, P16, P32, P43  \\
-        Mapping to the old model eases adaptation  &
-                P14, P29, P44, P46, P48  \\
-        Keep the new model simple  &
-                P5, P10, P16, P40  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Recommendations for choosing the agile model}
-    \label{table:success_choosemodel}
-\end{table}
-
-% Customizing the agile model
-
-Customizing the agile model and practices was often seen as an necessary step in
-the agile implementation. As each organization will have its own challenges in
-adopting agile, it should be carefully considered which organization specific
-areas to focus on when choosing the agile practices to implement [P8, P43].
-Lewis and Nehrer [P29] recommended to choose the agile model according to the
-current corporate model in order to avoid interference.
-
-Cases reported that customization was part of a successful agile implementation.
-For instance, teams were let to customize their practices individually, to fit
-the needs of each team [P41].
-Spayd [P49] writes that teams who modified agile practices to suite the large and
-distributed environment ended up doing better than teams that did not.
-% Some intermediate point?
-To allow teams to become innovative and perform well the agile model should be
-customized in a pragmatical way instead of following a strict textbook
-interpretation [P10]
-In order to draw the most out of agile teams should innovate and find on
-practices that really work for their case [P40, P41]
-
-Especially in a large organization it is not feasible to use the same process
-for all projects [P49]. An individual application of the agile process is needed
-for different types of development, depending on for instance the type of
-software being developed or the project size [P45].
-Applying agile at scale will require to deviate from some of the typical
-recommendations [P31]. However, it is essential to remember the agile principles
-when doing customizations, and watch out for customization that contradict with
-the principles [P31].
-
-The customization of the agile model was also seen as an evolutionary process.
-Cowan [P11] suggested that in a big-bang transformation it might be necessary to
-selectively compromise some parts of the agile implementation so that the core
-practices can be set in place.
-The agile transformation is a constantly ongoing process, and it is recommended
-to regularly challenge teams to refine the agile implementation to reflect the
-current needs of the organization [P41, P43].
-
-% Conformity in the agile model
-
-Some studies reported that conformity was considered in implementing the new
-model. A consistence common vocabulary was seen to benefit the organization and
-support the change [P8, P16, P43]. Other benefits in using a single model were
-possibility to compare work between teams, it was easier for people to relocate,
-and progress was predictable for stakeholders [P43]. Consistency and common
-understanding was created in discipline meetings and public events, or by peer
-coaching [P11, P12, P32].
-
-% Mapping agile to the old model
-
-Mapping the agile model to match the old model was seen as necessary in a few
-cases. Although a general mapping to the old model may not be good [P44, P46],
-some cases needed to preserve high level management practices [P14, P29, P48].
-By allowing management to remain unchanged it was possible to introduced agile
-step-wise on team level, and helped in getting management buy-in for the process
-[P14, P29, P48].
-In one case Scrum was considered as a wrapper for existing practices, and could
-therefore be easily fit in the organization [P44]. The agile methodology was
-considered complementing the existing culture, which eased management buy-in for
-the new method [P44].
-% Similarities and differences of agile were documented, and Scrum terminology was
-% explained in terms of the old model. Mapping of management information was done
-% so that for instance the information earlier found in Engineering Gantt charts
-% was found in the Sprint Backlog. [P29]
-
-% Keep things simple
-
-One advice given by few cases was to keep the organization and process simple
-[P10, P16]. Beavers [P5] describes that attempting to operate in a complex and
-global setting complicated even simple agile practices. The organizational chart
-was simplified, which was welcomed by both employees and managers [P5]. Rather
-than focusing on detail in process, communication practices, and tools the focus
-should be on engaging the teams [P40].
-
-\subsubsection{Considering the Product Owner role and requirements management}
-
-A particularly difficult part to implement efficiently was requirements
-management work. The key role in handling requirements is the Product Owner.
-Many cases described that careful attention to the requirements management roles
-was necessary for a successful transformation.
-Table \ref{table:success_requirements} lists references that discuss careful
-implementation of the product owner role and requirements management as success
-factors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.46\textwidth}
-                     >{\raggedright\arraybackslash}p{0.46\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Pay attention to the Product Owner role  &
-                P12, P14, P16, P24, P30, P33, P39, P46  \\
-        Pay attention to requirements management  &
-                P5, P11, P17, P33, P48  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Careful implementation of requirements management seen as a
-             success factor}
-    \label{table:success_requirements}
-\end{table}
-
-% Emphasize role of Product Owners
-
-A particular role that seemed to make much difference in the agile introduction
-was the Product Owner. Many cases reported on problems if the Product Owner did
-not perform adequately, and it was seen as critical to have a dedicated person
-in that role [P12, P14, P30, P39].
-Success or problems emerged respectively depending on how well the Product
-Owners were performing. In one case a well implemented Product Owner role was
-understood to be a key success factor when using agile methods [P16, P33].
-It was reported that when the Product Owner role was implemented correctly the
-team performed better and the work products were correct [P24]. Projects started
-off better when the Product Owners were engaged early on [P46]. Some Product
-Owners were even described to become agile enthusiasts when they learned how
-business and technology can collaborate in the agile model [P46].
-
-Several cases endorsed training and coaching for the Product Owner role [P16,
-P39, P46]. Product Owners should receive training on agile principles, backlog
-management, user story breakdown, and agile planning [P16]. Also tool support
-for backlog management and two way communication between Product Owners and
-teams should be worked on [P39].
-
-% Improve requirements handling
-
-Some cases had reported on challenges in requirements management, especially
-highlighting difficulty in managing the gap between high level requirements and
-stories handled by teams. It was stated that it is important that the
-requirements are concise and small enough for the teams to handle [P5, P11].
-It was recommended to invest in teaching skills in breaking down and writing
-user stories [P11, P33, P48]. Gat [P17] describes how the gap between
-development and requirements management was bridged by having a dedicated
-Requirements Architect to help with requirements breakdown.
-
-\subsubsection{Training and coaching}
-
-Agile methods come as new ways of working to many. In order to get a good start
-in the transformation the ideas of the new methods need to be explained
-thoroughly. Training sessions are good starting points, but teams should be
-accompanied by coaching for a longer time to ensure that the new practices are
-learned and applied correctly in practice. Table \ref{table:success_training}
-lists references that highlight training and coaching as success factors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.46\textwidth}
-                     >{\raggedright\arraybackslash}p{0.46\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Providing training on agile methods  &
-                P3, P6, P11, P14, P16, P30, P37, P46 \\
-        Coaching teams as they learn by doing  &
-                P3, P6, P8, P9 P16, P17, P18, P29,
-                P30, P33, P42, P44, P47, P48  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Training and coaching presented as success factors}
-    \label{table:success_training}
-\end{table}
-
-Training was seen as a strong factor helping the organization to transform.
-Several studies stated that training improved the chance of succeeding in the
-transformation [P3, P16, P30]. It was even highlighted that the change would have
-failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
-and again how training could make the difference between success and failure for
-a team''. Training also helped people to become more positively inclined towards
-the new model [P16, P37], and at best training made people enthusiastic to change
-[P3, P46]. A few cases reported that there was some disinclination to change,
-and training would have helped to overcome [P11, P14].
-
-Agile methods avoid prescribing exact ways of working but rather emphasize a
-mindset of adapting to the situation. It is difficult to explain by theory how
-such a mindset should be applied, and the agile practices are best learned by
-doing. Coaching teams while they apply the agile methods in practice was seen as
-an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
-that coaching was essential for success in transformation [P8, P18, P29, P33].
-Also, without coaching teams can do damage to the agile transformation if
-techniques are applied improperly [P9].
-
-There were a number of statements of the benefits of coaching. The coach
-position allowed to watch for and correct problems when they arose [P16, P42].
-Coaches also helped to draw attention away from focus on tools and towards
-understanding the principles of agile [P42, P48].
-There were also benefits in using both internal and external coaches.
-External coaches were able to have an objective view of the organization [P16,
-P44], but internal coaches were more accessible and knew the specifics of the
-organization [P3, P6].
-
-\subsubsection{Communication and transparency in the change}
-
-Communication and transparency was an important success factor brought up in the
-primary studies. The ongoing change should be communicated intensively to
-everyone. Transparency and inclusiveness were success factors related to
-communication.
-Table \ref{table:success_communication} presents success factors relating to
-communicating the change.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Communicate the change intensively  &
-                P16, P22, P33, P40, P41, P43, P48  \\
-        Enable transparency in the change  &
-                P6, P11, P16, P42, P43  \\
-        Engage people broadly in the organization  &
-                P4, P6, P16, P22, P31 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Success factors relating to communicating the change}
-    \label{table:success_communication}
-\end{table}
-
-Intensive communication was emphasized in a number of studies. It is important
-to reach a maximum number of people throughout the organization, as without
-communication the a new model will not take root [P16, P43].
-It was recommended that use of the new model is made highly visible on many
-communication channels and even over-communicated [P16, P33, P41]. Mencke [P33]
-summarizes the viewpoint on over-communication: ``Even if you think that your
-teams understand a new method or process, repeat your communication to be
-sure.'' Workshops, coaching sessions, online discussions, and newsletters are
-examples on different communication formats used [P41]. Another approach in
-communicating the change was that managers explained and encouraged change in an
-extensive series of one-to-one discussions [P22].
-
-Also communicating the goals of the transformation was highlighted in a some
-studies.
-Having a clear message of expectations helped to remove confusion and make
-people understand the purpose of the transformation [P43, P48]. The motivation
-to change can be increased by having a simple proposal on how to reach the
-desired outcomes [P40].
-McDowell and Dourambeis [P32] describe how the organization launched a series of
-communications events especially designed to emphasize the reasons behind the
-agile practices.
-
-Enabling transparency during the transformation was another communication
-related theme that was brought up, and even highlighted as a critical factor for
-success [P11, P16]. Fry and Greene [P16] underline the importance of
-transparency and sharing of information: ``This bias to sharing information with
-everyone was critical in our ability to adapt on a daily basis to ensure our
-success.'' Transparency was achieved by showing both successes and challenges
-[P6], actively reaching out for feedback [P6], using project management tools to
-display project status publicly [P11], by rearranging physical spaces [P42], and
-by holding the meetings of the roll-out team in public [P16].
-By sharing experiences and status of the transformation the organization was
-aligned and everyone was moving in the same direction [P42, P43].
-
-% The inertia and fear of sharing information must be overcome [P33].
-
-% Sharing everything with everyone was key to success [P16].
-
-% Organizations made also an effort to communicate the status of the
-% transformation [P16, P43].
-
-% Because of the volatility of software development there must be continuous and
-% fluent communication between managers, POs, and teams [P42].
-
-% Engage people broadly
-
-Many cases highlighted the importance of engaging people broadly in the
-organization.
-One of the lessons learned presented was that it is important to involve all
-stakeholders to gain acceptance in the transformation [P4, P31]. The transition
-team made an effort to engage people by both inviting to give feedback online
-and holding an extensive number of feedback meetings [P6, P16, P22] Being
-inclusive towards everyone was seen as one of the key success factors in the
-transformation [P16]. Inclusiveness will encourage people to participate and use
-the new model visibly [P16].
-
-\subsubsection{Enable autonomy for development teams}
-
-A key principle in agile is to allow development teams decide for themselves how
-they execute their tasks. To make the agile model work well teams need to
-self-organize, and be allowed to make decisions on the pace of development.
-When teams can perform their tasks autonomically the transformation towards
-agile will progress.
-Table \ref{table:success_autonomy} lists success factors relating to autonomy of
-development teams.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
-                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Allow teams to self-organize  &
-                P5, P16, P18, P41  \\
-        Empower teams to decide over speed and quality  &
-                P27, P30, P34, P45  \\
-        No top-down mandate  &
-                P3, P6, P41  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Success factors relating to enabling autonomy for development teams}
-    \label{table:success_autonomy}
-\end{table}
-
-% Empowering teams
-
-The agile principle of giving teams the power to decide over themselves was seen
-as an important factor in advancing the transformation.
-In some cases management initially attempted to prescribe how the new practices
-should be implemented [P5, P18, P41]. However, it was learned that only when
-full control was given to teams the agile methods could be properly established
-[P5, P16, P41].
-Allowing self-organization will create commitment to the change and motivation
-to continued use [P16, P18]. It will allow teams to take ownership of the
-development model and voluntarily improve it even further [P41].
-
-The acceptance of agile methods was easy when teams gained authority to decide
-on development speed and quality [P45]. Favoring empowerment over prescribing
-the new practices was also reported to improve performance [P27]. Giving teams
-authority to decide over work items increased productivity and morale [P30,
-P34].
-
-Roche and Vasquez-McCall [P41] describe how prescribing the agile methods to use
-lead only so-far in the transformation. The effectiveness of teaching and
-communicating the transformation started to decline. To enable the
-transformation to proceed further, an organization-wide challenge was created
-where teams would independently develop and showcase the best agile practices.
-As a result the ownership of the methods was transferred to teams, and the new
-model gained a secure foothold. [P41]
-
-% No top-down mandate
-
-An interesting success factor was the absence of a top-down mandate. Atlas [P3]
-describes that the use of Scrum was spreading because teams were both allowed to
-use and enabled to learn the method. The transformation itself was agile when
-people felt empowerment in making the decision to adopt. The incentive to change
-was created when teams learned about agile methods, and perceived that a change
-would be beneficial [P3].
-The absence of mandate granted genuine support for the model on the grass roots
-level, which was a necessity for the process to work [P6].
-It was also thought that proceeding with a top-down mandate would have made
-teams conform to a single process [P41]. This would have been sub-optimal, as it
-was important that each individual team and project tailored their practices
-[P41].
-
-\subsubsection{Choosing the right people}
-
-The transformation did not entirely rely on choices in technology and
-management, but also different people factors were given as basis of success.
-The choice of people to initially involve in the transformation was reported to
-make a difference. Interest or previous experience of agile methods were
-valuable characteristics to look for.
-Table \ref{table:success_people} lists factors choosing people that can benefit
-the transformation.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ ll }
-        \toprule
-        Success factors  &  Described in \\
-        \midrule
-        Start with people who are receptive for agile methods  &
-                P6, P29, P34  \\
-        Use people with previous agile experience  &
-                P6, P10, P37  \\
-        Have change leaders without baggage of the past  &
-                P11, P16  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Success factors relating to choosing the right people}
-    \label{table:success_people}
-\end{table}
-
-Choosing people with a disposition towards agile methods was seen as a
-requirement for the change to take the right track. Teams are staffed usually
-based on technical competences, but considering personality aspects alike was
-emphasized for agile teams [P29, P34]. The personality of people was seen as a
-key aspect for achieving change [P6]. There was a need for collaborative and
-understanding persons who are prepared to discard preconceptions and willing to
-try new untested approaches [P6, P29].
-% Also, having previous agile experience did not directly imply suitability to
-% work in a large scale agile environment [P34].
-% In the beginning of the transformation team members who did not find themselves
-% comfortable with the new model were exchanged for better suited persons [P29].
-
-% People with previous agile expertise
-
-A few cases mentioned the importance people with previous agile experience had
-at the start of the transformation [P6]. For instance, it helped the product
-management office to implement agile over the entire enterprise when the
-staffing of was updated to include people with agile experience [P10]. Staffing
-teams partially with developers familiar with agile helped the rest of the team
-get a good hold of agile development [P37].
-
-% People without baggage of the past
-
-A few cases discussed the benefit of having change leaders without baggage of
-the past. Cowan [P11] describes how the newly hired director was able to
-circumvent territorial battles that existed from before, and therefore focus
-fully on getting the agile model implemented. In another case external coaches
-were described to better spot places for correction in the agile model, and
-their advice was received better because they were considered impartial when
-being external to the organization [P16].
-% Old employees were also willing to try something new, because the organization
-% had gone through a time of uncertainty and restructuring [P11].
-
-% S: Have social events (get-togethers, etc.)
-% 
-% * Management team event --> Investment
-% * The transformation management team met in person --> Personal bonding
-% * A common understanding of the principles --> Consistent implementation
-% * Excursion and training
-% [P11]
-% 
-% * It was seen beneficial that the coach team had an event
-% * The team had been dispersed
-% * Road trip
-% [P23]
-% 
-% * Various social activities are given as example of improve team bonding
-% * Ways to build trust --> Trust is needed
-% * Team outings
-% [P27]
-% 
-% * Learn the difference between functions and formal roles
-% * Teams participated in learning games
-% * Learning games created understanding for team dynamics in the agile environment
-% [P40]
-% 
-% * Team competition-based events
-% * Short events in the middle of the work week
-% * Teams got excited when they were able to compete and demonstrate their own best practices
-% * Teams got encouragement to tailor practices
-% [P41]
-
-\subsubsection{Other success factors}
-
-A few success factors which did not fit in the above categories stood out. These
-success factors are creating a positive experience in the beginning, arranging
-social events, emphasizing principles over mechanics, and the previous
-organizational culture being like agile.
-
-% Having a positive experience
-
-It is clear that any positive results will facilitate an organizational
-transformation. Although a recipe to fabricate positive experiences can not be
-given, many cases described how they had succeeded. Benefield [P6] describes how
-a survey was done to objectively measure satisfaction in agile, and the positive
-results helped teams make the decision to change their way of working. Benefield
-describes further how management was requesting proof on better performance as a
-response to requests to increase coaching and training budget. After conducting
-inquiries the coaching team was able to present an estimate of a 30\% increase
-in performance, which made management highly convinced [P6]. Another case
-describes how the corporate agile awareness and teaching event was designed to
-be fun and engaging, which made people take more enthusiasm in applying the agile
-practices [P32]. The move towards agile is assisted by making any benefits
-publicly visible and celebrating even the small victories [P40, P42].
-
-The dynamics of positive results were explained in further ways. Several cases
-highlighted that the agile transformation was spread effectively through
-positive word-of-mouth [P6, P40]. When good results were shown by a team it
-created interest in others, and enthusiasm to try the new model will spread
-[P39, P46]. Some companies were using agile and waterfall methods side by side.
-This setting made it comparison possible, bringing up the benefits of the agile
-model [P19, P36].
-
-% Have social events
-
-Social events were reported to benefit the transformation. A few cases even
-described a the transformation being driven by a series of events where people
-received information and had the possibility to participate in shaping the new
-model of working [P32, P41]. Various social activities were presented as
-valuable tools for improving bonding within teams [P11, P27, P40]. The importance of
-creating personal bonding was also highlighted for change leading entities such
-as management and coaching teams [P11, P23].
-
-% Emphasize principles over mechanics
-
-In a number of studies it was described that the principles of agile should be
-emphasized over practices and simple mechanics [P46, P48]. When people
-understand the agile values they will also understand why the change is being
-done and feel motivated [P1, P16, P42].
-Some cases described that inexperienced coaches and Scrum masters made mistakes
-in focusing too much on the implementation of practices [P23, P30].
-
-% Previous culture is like agile
-
-A few organizations had already some structures in place that were similar to
-agile, which made the transformation easier.
-For instance, a previous organizational model based on small and autonomous
-teams strongly aided the adoption [P3].
-Another case described the original on-demand delivery model a natural fit for
-agile, and the transformation was presented as a return to core values instead
-of a remodeling [P16].
-
-\clearpage
-
-\section{Discussion}
-\label{sec:discussion}
-
-Grounded theory may not be able to provide a single generally applicable model
-for transformation, at least not in a literature review made based on the
-current agile transformation publications. This is because of two things.
-First, the organizations seem to differ too much from each other. Secondly, the
-published studies focus too much on isolated phenomena within the larger
-transformation.
-It would require too big of an interpretation to combine all the various cases
-into one theoretical model. We have chosen to present transformations using an
-elementary two-axial classification of transformations, and listings of
-individual phenomena that typically occur in a transformation.
-
-* Vertailu kirjallisuuteen
-
-* Some success factors and challenges relate
-
-* Some success factors and challenges are ``symmetrical'', so that lack of some
-  factor is a challenge, but availability of it provides success
-
-* Success factors and challenges may be complements, and may be presented as
-  complements
-
-* What is special in large scale?
-
-* Customization of practices was seen as necessary, as large scale has demands
-  over the original agile ideas
-
-* What is similar in organizations of all scales?
-
-* Transformation and scaling up are connected concepts. Although this research
-  focuses on transformation process only, it can not be completely separated
-  from discussion on scaling up, or on large-scale agile implementation in
-  general.
-
-* Transformation of an organization to use agile can be seen as transforming
-  also the agile method to suit the organization.
-  --> A large-scale agile model may require tailoring and trial over time
-
-* It can be argued that any development effort with hundreds of teams will be
-  difficult to coordinate
-
-* Even though most of the primary studies are experience reports and subject to
-  bias, this study can draw valuable conclusions. --> If a matter is being
-  talked about it must be of some significance.
-
-* Have flexibility for different teams in the agile model (agile found
-  hard [P41]) vs. conformity to a single model.
-
-* The factors that are listed are the ones to discuss -- there may be other
-  factors affecting transformation, but this study provides evidence that these
-  are the most important ones. We included only challenges that are discussed in
-  more than studies, which proves that some filtering has been done.
-
-* System development life cycle ?? -- There are inevitably other areas than
-  development of increments, even in an agile process
-
-* The primary studies were quite openly pro-agile -- how large a factor of bias
-  is this?
-
-* How do transformations fit within Kotter's 8 step process?
-
-* A qualitative interpretation requires much text. Therefore the results got
-  long. See ``Meta-ethnography''.
-
-* It is difficult to distinguish between team and organization. In most cases
-  when talking about teams it refers to a handful of people familiar to each
-  other. However, in some cases it was evident that an entire department was
-  referred to as a team.
-
-* Projects and organizations. How to distinguish between a separate adoption in
-  the beginning of each project, and an organization wide adoption.
-
-% ---------------------------------------------------------------------
-\section{Limitations and future research}
-\label{sec:conclusion}
-
-* Compare to change leadership research, and organizational research. Which
-  challenges are typical to any organization, which are specific to software
-  development organizations.
-
-* Compare how recommendations in large-scale application models of agile relate
-  to the challenges and success factors reported in the transformation process.
-
-% ---------------------------------------------------------------------
-% -------------- BIBLIOGRAPHY -----------------------------------------
-% ---------------------------------------------------------------------
-
-\bibliographystyle{elsarticle-harv}
-%\bibliographystylegen{plain}
-
-\bibliography{sources_background,sources_primary,sources_excluded}
+\input{content.tex}
 

### 2014-08-09 journal/content.tex
Preambles for article

@@
-% ---------------------------------------------------------------------
-% -------------- PREAMBLE ---------------------------------------------
-% ---------------------------------------------------------------------
-
-%\documentclass[12pt,a4paper,english,oneside]{article}
-\documentclass[preprint,authoryear,12pt]{elsarticle}
-
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{calc}
-\usepackage{natbib}
-\usepackage{url}
-%\usepackage{listings}
-\usepackage{hyphenat}
-
-\usepackage{supertabular,array}
-\usepackage{lipsum}
-\usepackage{booktabs}
-\usepackage{graphicx}
-
-% Strike through
-\usepackage[normalem]{ulem}
-
-% Punctuation for references
-\bibpunct{[}{]}{;}{n}{,}{,}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% Multiple bibliographies
-%\usepackage{multibib}
-%\newcites{gen}{References}
-%\usepackage{bibentry}
-
-% ---------------------------------------------------------------------
-% -------------- DOCUMENT ---------------------------------------------
-% ---------------------------------------------------------------------
-
-\journal{Information and Software Technology}
-
-\begin{document}
-
-% Extra definitions?
-\selectlanguage{english}
-
-\begin{frontmatter}
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for the associated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for the associated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for the associated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%%
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{<author name>}
-%% \address[label1]{<address>}
-%% \address[label2]{<address>}
-
-\title{Adopting agile software development in the large organizations:
-       A systematic literature review}
-
-\author{Kim Dikert}
-\ead{kim-karol.dikert@aalto.fi}
-
-\author{Maria Paasivaara}
-\ead{maria.paasivaara@aalto.fi}
-
-\author{Casper Lassenius}
-\ead{casper.lassenius@aalto.fi}
-
-\address{Aalto University, School of Science, Department of Computer Science and Engineering}
-
-% ---------------------------------------------------------------------
-% ------------------------- ABSTRACT ----------------------------------
-% ---------------------------------------------------------------------
-
-\begin{abstract}
-
-Agile methods have become an appealing alternative for companies striving to
-improve their performance, but the methods are originally designed for small and
-individual teams. This creates unique challenges when introducing agile to large
-organizations where development teams can not be autonomous. We conducted a
-systematic literature review including publications on 40 industry cases,
-describing the process of taking agile methods into use in large organizations.
-
-The primary motive for agile adoption is the need to improve performance. For a
-successful transformation the entire organization must be aligned and all
-organizational units must participate.
-Creating an agile community in the organization will help the change to stick.
-Typical problems are misunderstanding of the agile model, and difficulties
-coordinating multiple agile teams.
-
-\end{abstract}
-
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-agile \sep transformation \sep large-scale \sep adopting agile
-
-\end{keyword}
-
-\end{frontmatter}
-
-% ---------------------------------------------------------------------
-% --------------------------- TEXT ------------------------------------
-% ---------------------------------------------------------------------
-
 \section{Introduction}
@@
 \bibliography{sources_background,sources_primary,sources_excluded}
-
-\end{document}

### 2014-08-10 journal/content.tex
Bibliography settings
Changed section labels for thesis style

@@
-\section{Introduction}
+\chapter{Introduction}
 
@@
 
-\section{Background}
+\chapter{Background}
 \label{sec:background}
 
-\subsection{Agile software development}
+\section{Agile software development}
 
@@
 
-\subsection{Large scale agile development}
+\section{Large scale agile development}
 
@@
 
-\subsection{Adopting agile methods in large organizations}
+\section{Adopting agile methods in large organizations}
 
@@
 
-% \subsection{Challenges in adopting agile methods}
+% \section{Challenges in adopting agile methods}
 %
 
-% \subsection{Success factors in adopting agile methods}
+% \section{Success factors in adopting agile methods}
 %
 
-\subsection{Summary of things that may need to be explained}
+\section{Summary of things that may need to be explained}
 
@@
 
-\section{Research method}
+\chapter{Research method}
 \label{sec:method}
@@
 
-\subsection{Research questions}
+\section{Research questions}
 
@@
 
-\subsection{Definition of large scale}
+\section{Definition of large scale}
 \label{sec:largescale}
@@
 
-\subsection{Research process}
+\section{Research process}
 
@@
 
-\subsubsection{Inclusion criteria}
+\subsection{Inclusion criteria}
 \label{sec:inclusioncriteria}
@@
 
-\subsubsection{Preliminary searches}
+\subsection{Preliminary searches}
 
@@
 
-\subsubsection{Identification of primary studies}
+\subsection{Identification of primary studies}
 
@@
 
-\subsubsection{Study selection}
+\subsection{Study selection}
 
@@
 
-\subsubsection{Handling of duplicate reports on a single case}
+\subsection{Handling of duplicate reports on a single case}
 
@@
 
-\subsubsection{Study quality assessment}
+\subsection{Study quality assessment}
 
@@
 
-\subsubsection{Coding of primary studies}
+\subsection{Coding of primary studies}
 \label{sec:coding}
@@
 
-\subsubsection{Validation of coding}
+\subsection{Validation of coding}
 
@@
 
-\subsubsection{Synthesis of primary studies based on coding}
+\subsection{Synthesis of primary studies based on coding}
 
@@
 
-\subsection{Threats to validity and limitations}
+\section{Threats to validity and limitations}
 
@@
 
-\section{Results}
+\chapter{Results}
 \label{sec:results}
@@
 
-\subsection{Overview of primary studies}
+\section{Overview of primary studies}
 
@@
 
-\subsection{Agile methods used}
+\section{Agile methods used}
 
@@
 
-\subsection{Reasons to change}
+\section{Reasons to change}
 
@@
 
-\subsubsection{Business demands speed and flexibility}
+\subsection{Business demands speed and flexibility}
 
@@
 
-\subsubsection{Management issues}
+\subsection{Management issues}
 
@@
 
-\subsubsection{Process problems}
+\subsection{Process problems}
 
@@
 
-\subsubsection{Other reasons to change}
+\subsection{Other reasons to change}
 
@@
 
-\subsection{Transformation process}
+\section{Transformation process}
 
@@
 
-\subsubsection{The initial state of the organization}
+\subsection{The initial state of the organization}
 
@@
 
-\subsubsection{Modes of transformation}
+\subsection{Modes of transformation}
 
@@
 
-\subsubsection{How change was lead}
+\subsection{How change was lead}
 
@@
 
-\subsubsection{Training, coaching, and awareness}
+\subsection{Training, coaching, and awareness}
 
@@
 
-\subsubsection{Piloting}
+\subsection{Piloting}
 
@@
 
-\subsubsection{Investing in new structures in transformation}
+\subsection{Investing in new structures in transformation}
 
@@
 
-\subsection{Challenges in transformation}
+\section{Challenges in transformation}
 
@@
 
-\subsubsection{Change resistance}
+\subsection{Change resistance}
 
@@
 
-\subsubsection{Lack of investment in transformation}
+\subsection{Lack of investment in transformation}
 
@@
 
-\subsubsection{Agile is difficult to implement}
+\subsection{Agile is difficult to implement}
 
@@
 
-\subsubsection{Coordinating work between many teams}
+\subsection{Coordinating work between many teams}
 
@@
 
-\subsubsection{Issues specific to large organizations}
+\subsection{Issues specific to large organizations}
 
@@
 
-\subsubsection{Integrating requirements management and quality assurance}
+\subsection{Integrating requirements management and quality assurance}
 
@@
 
-\subsubsection{Integrating other functions than development to agile}
+\subsection{Integrating other functions than development to agile}
 
@@
 
-\subsubsection{Other challenges in transformation}
+\subsection{Other challenges in transformation}
 
@@
 
-\subsection{Success factors in transformation}
+\section{Success factors in transformation}
 
@@
 
-\subsubsection{Management support}
+\subsection{Management support}
 
@@
 
-\subsubsection{Commitment to change}
+\subsection{Commitment to change}
 
@@
 
-\subsubsection{Leadership and alignment}
+\subsection{Leadership and alignment}
 
@@
 
-\subsubsection{Use of piloting}
+\subsection{Use of piloting}
 
@@
 
-\subsubsection{Choosing the agile model to fit the organization}
+\subsection{Choosing the agile model to fit the organization}
 
@@
 
-\subsubsection{Considering the Product Owner role and requirements management}
+\subsection{Considering the Product Owner role and requirements management}
 
@@
 
-\subsubsection{Training and coaching}
+\subsection{Training and coaching}
 
@@
 
-\subsubsection{Communication and transparency in the change}
+\subsection{Communication and transparency in the change}
 
@@
 
-\subsubsection{Enable autonomy for development teams}
+\subsection{Enable autonomy for development teams}
 
@@
 
-\subsubsection{Choosing the right people}
+\subsection{Choosing the right people}
 
@@
 
-\subsubsection{Other success factors}
+\subsection{Other success factors}
 
@@
 
-\section{Discussion}
+\chapter{Discussion}
 \label{sec:discussion}
@@
 % ---------------------------------------------------------------------
-\section{Limitations and future research}
+\chapter{Limitations and future research}
 \label{sec:conclusion}
@@
 
-\bibliographystyle{elsarticle-harv}
-%\bibliographystylegen{plain}
-
 \bibliography{sources_background,sources_primary,sources_excluded}

### 2014-08-10 journal/content.tex
Layout updates for thesis style
Moved validity and limitations to the end, organized conclusion chapter

@@
 
-\begin{table}
+\begin{table}[b]
     \centering
@@
 %   result text)
-
-\section{Threats to validity and limitations}
-
-To assess the validity of this research internal validity and external validity
-needs to be considered. In this section we present the identified threats to
-validity and the strategies used to mitigate them. We also highlight two
-specific limitations in the research: weakness in keyword search, and the
-strength of evidence of primary sources.
-
-The threat to internal validity is subjective bias, which affects two parts of
-the research. First, the selection of the primary studies may have been
-distorted by interpreting the inclusion criteria falsely. This risk was
-mitigated at start by using three researchers in designing the inclusion
-criteria. When the inclusion criteria was subsequently applied the initial
-filtering was performed independently by two researchers, and unclear cases were
-resolved by case-by-case discussions between two or three researchers.
-
-The second part of the research that may have been affected by subjective bias
-was the elicitation of results by coding and analysis. Our tools for mitigating
-this threat were limited, as the work stages in question are particularly
-laborious. For resource constraints the process could not be duplicated. We
-assessed the bias of the coding step using an experiment, and prevented
-subjective bias in analysis by making the results as traceable as possible. The
-coding experiment revealed that two researchers can find roughly 2 of 3 themes
-in an experience report identically. Traceability of results was achieved by
-supplying references to each claim presented in the results.
-
-This study had a minor threat to external validity. The systematic literature
-review approach includes steps to identify all studies that may be relevant in
-the study area. As long as the systematic approach is adhered to, including
-systematic identification of research, threats against external validity are
-minimized. External validity was also confirmed by saturation of the primary
-studies. Saturation was evident, as when the majority of the primary studies had
-been coded almost all quotations for the remaining studies could be assigned to
-existing codes.
-
-A particular problem in this systematic literature review was the limitations of
-Boolean keyword searches in online databases. A keyword search can not identify
-the facet of \emph{large scale}, and also \emph{transformation} is difficult to
-capture with keywords. Because of this we assume that it is likely that a some
-studies have been missed by the keyword search, potentially causing bias. To
-eliminate the potential bias we would suggest manually searching the entire
-archive of \emph{Agile Conference} and \emph{Agile Processes in Software
-Engineering and Extreme Programming} for relevant studies. These two conferences
-provided more than two thirds of the primary sources.
-
-Another particular problem was that few of the primary studies presented a
-research method. Most of the studies were experience reports where the author
-was a member of the organization discussed. Because of this we must assume that
-the primary studies are biased to some extent. Due to this risk of bias we chose
-not to make quantitative interpretations in the results, and use only
-qualitative analysis. We conclude that in the current state of research it is
-necessary to include studies with varying strength of evidence in order to
-relevantly aggregate evidence on large scale transformations.
 
@@
 
-\chapter{Discussion}
+\chapter{Discussion and conclusions}
 \label{sec:discussion}
+
+In this chapter we discuss the results and provide a conclusion on this work.
+First the results are discussed, including highlighting of most important
+findings and comparison to previous work. Then we present limitations and
+thereats to the validity of this work. This chapter ends with a concluding
+overview of the work.
+
+\section{Discussion}
+
+* ...
 
@@
 
+\section{Threats to validity and limitations}
+
+To assess the validity of this research internal validity and external validity
+needs to be considered. In this section we present the identified threats to
+validity and the strategies used to mitigate them. We also highlight two
+specific limitations in the research: weakness in keyword search, and the
+strength of evidence of primary sources.
+
+The threat to internal validity is subjective bias, which affects two parts of
+the research. First, the selection of the primary studies may have been
+distorted by interpreting the inclusion criteria falsely. This risk was
+mitigated at start by using three researchers in designing the inclusion
+criteria. When the inclusion criteria was subsequently applied the initial
+filtering was performed independently by two researchers, and unclear cases were
+resolved by case-by-case discussions between two or three researchers.
+
+The second part of the research that may have been affected by subjective bias
+was the elicitation of results by coding and analysis. Our tools for mitigating
+this threat were limited, as the work stages in question are particularly
+laborious. For resource constraints the process could not be duplicated. We
+assessed the bias of the coding step using an experiment, and prevented
+subjective bias in analysis by making the results as traceable as possible. The
+coding experiment revealed that two researchers can find roughly 2 of 3 themes
+in an experience report identically. Traceability of results was achieved by
+supplying references to each claim presented in the results.
+
+This study had a minor threat to external validity. The systematic literature
+review approach includes steps to identify all studies that may be relevant in
+the study area. As long as the systematic approach is adhered to, including
+systematic identification of research, threats against external validity are
+minimized. External validity was also confirmed by saturation of the primary
+studies. Saturation was evident, as when the majority of the primary studies had
+been coded almost all quotations for the remaining studies could be assigned to
+existing codes.
+
+A particular problem in this systematic literature review was the limitations of
+Boolean keyword searches in online databases. A keyword search can not identify
+the facet of \emph{large scale}, and also \emph{transformation} is difficult to
+capture with keywords. Because of this we assume that it is likely that a some
+studies have been missed by the keyword search, potentially causing bias. To
+eliminate the potential bias we would suggest manually searching the entire
+archive of \emph{Agile Conference} and \emph{Agile Processes in Software
+Engineering and Extreme Programming} for relevant studies. These two conferences
+provided more than two thirds of the primary sources.
+
+Another particular problem was that few of the primary studies presented a
+research method. Most of the studies were experience reports where the author
+was a member of the organization discussed. Because of this we must assume that
+the primary studies are biased to some extent. Due to this risk of bias we chose
+not to make quantitative interpretations in the results, and use only
+qualitative analysis. We conclude that in the current state of research it is
+necessary to include studies with varying strength of evidence in order to
+relevantly aggregate evidence on large scale transformations.
+
+\section{Conclusion}
+
+A concluding overview.
+
+* Recap of the work
+
+* Recap of mosti important results
+
+(* Future research?)
+
 % ---------------------------------------------------------------------
-\chapter{Limitations and future research}
-\label{sec:conclusion}
-
-* Compare to change leadership research, and organizational research. Which
-  challenges are typical to any organization, which are specific to software
-  development organizations.
-
-* Compare how recommendations in large-scale application models of agile relate
-  to the challenges and success factors reported in the transformation process.
+% \chapter{Limitations and future research}
+% \label{sec:conclusion}
+% 
+% * Compare to change leadership research, and organizational research. Which
+%   challenges are typical to any organization, which are specific to software
+%   development organizations.
+% 
+% * Compare how recommendations in large-scale application models of agile relate
+%   to the challenges and success factors reported in the transformation process.
 

### 2014-08-10 journal/content.tex
Commented out notes

@@
 
-\section{Large scale agile development}
-
-Laanti 2008:
-[2] Kähkönen Tuomo, “Agile Methods for large organizations – building communities of practice”.
-Proc. Agile Development Conf. (ADC), 2004, pp. 2-10.
-[7] Eckstein Jutta, “Agile Software Development in the large: Diving into the deep”,
-Dorset House Publishing, 2004. ISBN 0-932633-57-9.
-[8] Leffingwell Dean, “Scaling Software Agility: Best Practices for Large Enterprises”,
-Addison-Wesley Professional, 2007. ISBN-10: 0321458192.
-[12] Highsmith J., Agile for the Enterprise, “From Agile Teams to Agile Organizations”,
-Cutter, 2005. http://www.cutter.com/program/ fulltext/reports/2005/01/index.html .
-[13] Schwaber Ken, “Enterprise and Scrum”, Microsoft Press 2007, ISBN-10: 0735623376.
-
-Vlaanderen 2012:
-22. Sidky, A.S.: A Structured Approach to Adopting Agile Practices: The Agile Adop-
-tion Framework (June 2007)
-
-Cockburn 2003:
-R. Crocker, Large-Scale Agile Software Development, Addison-Wesley, 2003.
+% \section{Large scale agile development}
+% 
+% Laanti 2008:
+% [2] Kähkönen Tuomo, “Agile Methods for large organizations – building communities of practice”.
+% Proc. Agile Development Conf. (ADC), 2004, pp. 2-10.
+% [7] Eckstein Jutta, “Agile Software Development in the large: Diving into the deep”,
+% Dorset House Publishing, 2004. ISBN 0-932633-57-9.
+% [8] Leffingwell Dean, “Scaling Software Agility: Best Practices for Large Enterprises”,
+% Addison-Wesley Professional, 2007. ISBN-10: 0321458192.
+% [12] Highsmith J., Agile for the Enterprise, “From Agile Teams to Agile Organizations”,
+% Cutter, 2005. http://www.cutter.com/program/ fulltext/reports/2005/01/index.html .
+% [13] Schwaber Ken, “Enterprise and Scrum”, Microsoft Press 2007, ISBN-10: 0735623376.
+% 
+% Vlaanderen 2012:
+% 22. Sidky, A.S.: A Structured Approach to Adopting Agile Practices: The Agile Adop-
+% tion Framework (June 2007)
+% 
+% Cockburn 2003:
+% R. Crocker, Large-Scale Agile Software Development, Addison-Wesley, 2003.
 
@@
 
-\section{Summary of things that may need to be explained}
-
-* Scrum
-
-* XP
-
-* Definition of sprint
-
-* Definition of product owner
-
-* Self organization and independence of teams -- This is a question relating how
-  to organize management.
-
-* Scrum master must not command and control -- Scrum master as manager??
-
-* People centricity -- agile is not for squeezing efficiency -- agile is not for
-  fewer career opportunities in a flatter organization.
-
-* Team centricity -- agile focuses on team rather than individuals
-
-* Agile focuses on short time horizons ???
-
-* Agile is a mindset -- it can not be studied from a book -- need of coaching
-
-* What are the agile principles? What is being talked about when ``agile
-  principles'' is referred to?
-
-* What is coaching? Why is it done?
-
-* Principle of build often? Principle of nightly builds or CI? Is this from XP?
-  Is this an agile principle?
-
-* Kotter: Having a ``guiding team'' for change
+% \section{Summary of things that may need to be explained}
+% 
+% * Scrum
+% 
+% * XP
+% 
+% * Definition of sprint
+% 
+% * Definition of product owner
+% 
+% * Self organization and independence of teams -- This is a question relating how
+%   to organize management. 
+% 
+% * Scrum master must not command and control -- Scrum master as manager??
+% 
+% * People centricity -- agile is not for squeezing efficiency -- agile is not for
+%   fewer career opportunities in a flatter organization.
+% 
+% * Team centricity -- agile focuses on team rather than individuals
+% 
+% * Agile focuses on short time horizons ???
+% 
+% * Agile is a mindset -- it can not be studied from a book -- need of coaching
+% 
+% * What are the agile principles? What is being talked about when ``agile
+%   principles'' is referred to?
+% 
+% * What is coaching? Why is it done?
+% 
+% * Principle of build often? Principle of nightly builds or CI? Is this from XP?
+%   Is this an agile principle?
+% 
+% * Kotter: Having a ``guiding team'' for change
 

### 2014-08-10 journal/content.tex
Style review

@@
     \end{tabular}
-    \caption{Context information}
+    \caption{Context information specified to be gathered from primary studies.}
     \label{table:contextualcodes}
@@
     \end{tabular}
-    \caption{Code families}
+    \caption{Code families defined before start of coding. Examples on
+    potential codes to be inducted in parentheses.}
     \label{table:codefamilies}
@@
     \end{tabular}
-    \caption{Number of codes and quotations per code family}
+    \caption{Number of codes and quotations per code family, as the final result
+    of coding the primary studies}
     \label{table:codecount}
@@
 wiki-based website to maintain and distribute information on agile practices
-[P14, P16, P45]. Hanly et al. [P23] describes how a n organization specific
+[P14, P16, P45]. Hanly et al. [P23] describes how an organization specific
 cookbook for agile implementation was created. The cookbook resolved
@@
 
-\begin{table}[h]
+\begin{table}[b]
     \centering
@@
 
-\begin{table}[b]
+\begin{table}[t]
     \centering

### 2014-08-10 journal/content.tex
Layout and spell check

@@
 
-\begin{table}[b]
+We proceeded with constructing a search string based on the facets presented in
+section \ref{sec:inclusioncriteria}. However, preliminary searches had showed
+that picking keywords with good precision was difficult. Especially the facet
+\emph{large scale} was difficult to represent with precise words. Also the facet
+of \emph{empiricality} could not sensibly be represented by keywords. We decided
+to include only the facets \emph{agile software development} and
+\emph{organizational transformation} in the keyword search, with the consequence
+of increasing the manual filtering effort in the subsequent steps.
+
+Having a substantial part of the matches in non-relevant areas such as agile
+manufacturing, we included only articles including the term ``software'' or
+articles published in relevant conferences. Instead of engineering the search
+terms any further, which proved to result in excluding some relevant articles,
+we chose to rely more on manual filtering. The resulting facets and keywords are
+listed in Table \ref{table:searchterms}.
+
+The database keyword search matched 1875 unique papers. The results of the
+search were gathered for the selection step using reference management software.
+
+\begin{table}[h]
     \centering
@@
 \end{table}
-
-We proceeded with constructing a search string based on the facets presented in
-section \ref{sec:inclusioncriteria}. However, preliminary searches had showed
-that picking keywords with good precision was difficult. Especially the facet
-\emph{large scale} was difficult to represent with precise words. Also the facet
-of \emph{empiricality} could not sensibly be represented by keywords. We decided
-to include only the facets \emph{agile software development} and
-\emph{organizational transformation} in the keyword search, with the consequence
-of increasing the manual filtering effort in the subsequent steps.
-
-Having a substantial part of the matches in non-relevant areas such as agile
-manufacturing, we included only articles including the term ``software'' or
-articles published in relevant conferences. Instead of engineering the search
-terms any further, which proved to result in excluding some relevant articles,
-we chose to rely more on manual filtering. The resulting facets and keywords are
-listed in Table \ref{table:searchterms}.
 
@@
 
-The database keyword search matched 1875 unique papers. The results of the
-search were gathered for the selection step using reference management software.
-
 \subsection{Study selection}
@@
 		Distinguishable practices that are used or established during transformation.
-		Factors that are presented as neutral in terms of successes and challenges
-		in the primary studies.
+		Factors presented as neutral relating to successes and challenges.
 		(coaching, piloting, continuous integration, communities of practice) \\
@@
 
-\begin{table}[t]
+\begin{table}
     \centering
@@
 
+\clearpage
+
 \subsection{Validation of coding}
@@
 As an accurate description had been created for each code, the initially defined
-categories could be refined. Some codes were recategorized, and the definitions
+categories could be refined. Some codes were re-categorized, and the definitions
 of a few categories were revisited, until a final ordering was reached. The
@@
 % What public events
-Organizations typically strived to increase the agile awareness by introducing
+Organizations were typically striving to increase the agile awareness by introducing
 series of events for promoting agile, ranging from public seminars to brown bag
@@
 [P12, P45].
-Companies strived to create communities of practice [P42] or special interest
+Companies were striving to create communities of practice [P42] or special interest
 groups [P41] in order to reinforce continuous learning in the organization.
@@
 Some organizations started their agile journey in a state where people were
-overcommitted. It was later realized that when people are overloaded they will
+over-committed. It was later realized that when people are overloaded they will
 not be able to changing their behavior and learn new ways of working [P37, P42].
@@
 However, even when there was preparation for extra work people with specific
-expertise became overloaded. After time it was realized that overcommitment must
+expertise became overloaded. After time it was realized that over-commitment must
 be reduced by allowing to shift excess work for later time [P11].
@@
 findings and comparison to previous work. Then we present limitations and
-thereats to the validity of this work. This chapter ends with a concluding
+threats to the validity of this work. This chapter ends with a concluding
 overview of the work.
@@
 
-* Recap of mosti important results
+* Recap of most important results
 

### 2014-08-10 journal/content.tex
Large scale definition as part of background

@@
 \label{sec:background}
+
+In this chapter we present existing background literature for this work. This
+chapter begins with an overview of agile software development. The second
+part presents an overview of existing studies on introducing agile in large
+organizations. The last part of this chapter defines large scale in the context
+of this study.
 
@@
 
-% ---------------------------------------------------------------------
-\clearpage
-
-\chapter{Research method}
-\label{sec:method}
-
-The goal of the research was to aggregate evidence on how large organizations
-transform into an agile way of working. The research was conducted as a
-systematic literature review based on empirical studies.
-
-The research process is presented starting from the research questions. After
-that we present the definition of large scale which this research builds upon.
-The bulk of this section concentrates on presenting the steps of the research
-process, starting from the definition of inclusion criteria, continuing with
-describing how the primary studies were selected, and finally presenting the
-method for analyzing and synthesizing the primary studies. The section concludes
-with a discussion on limitations of the method.
-
-\section{Research questions}
-
-The following research questions were set to guide the research.
-
-\begin{itemize}
-
-\item
-RQ1: Why do large software development organizations initiate an agile
-transformation?
-
-\item
-RQ2: How do large-scale agile transformations usually proceed?
-
-\item
-RQ3: What are typical challenges in a large-scale agile transformation?
-
-\item
-RQ4: What factors facilitate adopting of agile methods?
-
-\end{itemize}
-
 \section{Definition of large scale}
@@
 %     required numbers. Small companies in growth are therefore excluded.
+
+% ---------------------------------------------------------------------
+\clearpage
+
+\chapter{Research method}
+\label{sec:method}
+
+The goal of the research was to aggregate evidence on how large organizations
+transform into an agile way of working. The research was conducted as a
+systematic literature review based on empirical studies.
+
+The research process is presented starting with the research questions.
+%After that we present the definition of large scale which this research builds upon.
+The bulk of this section concentrates on presenting the steps of the research
+process, starting from the definition of inclusion criteria, continuing with
+describing how the primary studies were selected, and finally presenting the
+method for analyzing and synthesizing the primary studies.
+%The section concludes with a discussion on limitations of the method.
+
+\section{Research questions}
+
+The following research questions were set to guide the research.
+
+\begin{itemize}
+
+\item
+RQ1: Why do large software development organizations initiate an agile
+transformation?
+
+\item
+RQ2: How do large-scale agile transformations usually proceed?
+
+\item
+RQ3: What are typical challenges in a large-scale agile transformation?
+
+\item
+RQ4: What factors facilitate adopting of agile methods?
+
+\end{itemize}
 

### 2014-08-10 journal/content.tex
Bigger image
T: External consultants and investments separate subjects

@@
   \begin{center}
-    \includegraphics{graphics/research_process.pdf}
+    \includegraphics[width=1\textwidth]{graphics/research_process.pdf}
     \caption{Outline of the research process}
@@
     \end{tabular}
-    \caption{Summary of activities aiming to advance the transformation}
+    \caption{Sources discussing viewpoints on piloting}
     \label{table:transformation_piloting}
@@
 
+\subsection{Use of external consultants}
+
+It was very typical that external help was brought in to help with the
+transformation.
+Primary sources presenting different ways external consultants were employed in
+the transformation are listed in table \ref{table:transformation_consultants}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
+                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
+        \toprule
+        Sources discussing use of external consultants  &
+                P5-P9, P16-P21, P23 P25, P27, P29-P31,
+                P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
+        \bottomrule
+    \end{tabular}
+    \caption{Sources discussing use of external consultants}
+    \label{table:transformation_consultants}
+\end{table}
+
+% External consultants
+
+External consultants were employed especially at the start of the
+transformation. At the time it was realized that the way of working should be
+changed a consultant was often hired to assess the current state and give a
+recommendation on how to proceed [P16, P20, P41, P43, P41, P44, P51].
+In the beginning external consultants were used to ignite interest in agile
+methods, and provide initial education [P7, P9, P27, P30, P43].
+It was recommended to bring in external help as early as possible, when the
+internal agile expertise is lacking and before transformation challenges appear
+[P6, P16, P30].
+
+When the transformation was getting on its way, agile consultants were typically
+leading training sessions, arranging agile workshops, and coaching teams.
+In some cases consultants helped specifically with scaling up the use of agile
+beyond single teams [P31, P34].
+Consultants were used mostly in the first stages of the transformation.
+The need for external consultants reduced over time, and their tasks were
+transferred to internal positions [P29, P47].
+It was described that as the agile journey progressed, external consultants were
+invited every now and then to bring in fresh ideas to the organization [P6].
+
+External consultants were regarded to bring in critical assistance in
+transforming the organization [P6, P16, P17]. Qualities in external consultants
+that organizations looked for was leveraging their experience to bring more
+speed to development and agile adoption [P17, P27, P34, P35, P40], and to bring
+in agile skill when only little existed in-house [P27, P47]. It was considered
+beneficial having experts with previous experience of transforming organizations
+[P16, P27].
+Still one important aspect in using consultants was that people were more
+comfortable when the advice was given by impartial persons external to the
+organization [P16, P29, P44].
+
+% In one case the teams were sceptical of having external consultants embedded
+% in temas [P27].
+
 \subsection{Investing in new structures in transformation}
@@
 tangible investments. These investments were focused on better enabling agile
-software development. They also highlight the concerns that organizations had when
-implementing an agile model. Typical investments were bringing in external help,
-new management tools, improvements in testing environments, and updating of the
-physical working spaces. Table \ref{table:transformation_investments} summarizes
-common types of investments.
+software development. They also highlight the concerns that organizations had
+when implementing an agile model. Typical investments were new management tools,
+improvements in testing environments, and updating of the physical working
+spaces. Table \ref{table:transformation_investments} summarizes common types of
+investments.
 
@@
         \midrule
-        Using external consultants  &
-                P5-P9, P16-P21, P23 P25, P27, P29-P31,
-                P34, P35, P37-P41, P44, P46, P47, P50, P51  \\  % 29
         New project management tools for agile  &
@@
 \end{table}
-
-% External consultants
-
-It was very typical that external help was brought in to help with the
-transformation.
-External consultants were employed especially at the start of the
-transformation. At the time it was realized that the way of working should be
-changed a consultant was often hired to assess the current state and give a
-recommendation on how to proceed [P16, P20, P41, P43, P41, P44, P51].
-In the beginning external consultants were used to ignite interest in agile
-methods, and provide initial education [P7, P9, P27, P30, P43].
-It was recommended to bring in external help as early as possible, when the
-internal agile expertise is lacking and before transformation challenges appear
-[P6, P16, P30].
-
-When the transformation was getting on its way, agile consultants were typically
-leading training sessions, arranging agile workshops, and coaching teams.
-In some cases consultants helped specifically with scaling up the use of agile
-beyond single teams [P31, P34].
-Consultants were used mostly in the first stages of the transformation.
-The need for external consultants reduced over time, and their tasks were
-transferred to internal positions [P29, P47].
-It was described that as the agile journey progressed, external consultants were
-invited every now and then to bring in fresh ideas to the organization [P6].
-
-External consultants were regarded to bring in critical assistance in
-transforming the organization [P6, P16, P17]. Qualities in external consultants
-that organizations looked for was leveraging their experience to bring more
-speed to development and agile adoption [P17, P27, P34, P35, P40], and to bring
-in agile skill when only little existed in-house [P27, P47]. It was considered
-beneficial having experts with previous experience of transforming organizations
-[P16, P27].
-Still one important aspect in using consultants was that people were more
-comfortable when the advice was given by impartial persons external to the
-organization [P16, P29, P44].
-
-% In one case the teams were sceptical of having external consultants embedded
-% in temas [P27].
 

### 2014-08-10 journal/content.tex
Layout fix
Conclusion section

@@
 
-\begin{table}[t]
+\begin{table}[!b]
     \centering
@@
 
-A concluding overview.
-
-* Recap of the work
-
-* Recap of most important results
-
-(* Future research?)
+The goal of this work was to review and analyze empirical studies of large
+organizations taking agile software development methods into use. By applying
+the method of systematic literature review we analyzed the findings of 52
+studies describing the agile transformation in 40 different organizations. As a
+result we presented qualitative findings describing typical reasons to change,
+characteristics of transformations, challenges, and success factors.
+
+The primary motive to change was a demand to faster and more flexibly respond to
+business needs. Other reasons included the desire to mitigate known problems in
+the organization.
+Transformation processes could be characterized along two dimensions,
+one distinguishing between top-down and bottom-up leadership, and the other
+being the transformation velocity. Typical characteristics in transformations
+were use of pilot projects, bringing in external consultants, and investing in
+training.
+
+Typical challenges included change resistance and lack of commitment and
+investment in transformation. Some challenges related particularly to large
+scale, such as implementing effective coordination between teams, and friction
+caused by the fact that a large organization can not transform all at once.
+The most important success factors were management support, use of training and
+coaching and carefully customizing the agile model to fit the organization.
+
+Where to from here?
+Adopting agile methods in large organizations is closely related to generic
+discussion on organizational change. A possible direction for future research
+would be to compare the agile transformations to research on change in knowledge
+organizations from the viewpoint of organizational research. The organizational
+research domain is significantly older than research on agile development, and
+would therefore likely provide additional insight in agile transformations.
+Also, agile methods and organizations using them mature over time. A research
+topic related to transformation is the follow-up of the evolving agile process.
+Research seems to be scarce on the evolving of agile methods after introduction
+in large organizations.
 

### 2014-08-10 journal/content.tex
Missing comma

@@
 The most important success factors were management support, use of training and
-coaching and carefully customizing the agile model to fit the organization.
+coaching, and carefully customizing the agile model to fit the organization.
 

### 2014-08-10 journal/content.tex
Discussion

@@
 
-* ...
-
-Grounded theory may not be able to provide a single generally applicable model
-for transformation, at least not in a literature review made based on the
+Grounded theory may not be able to provide a generally applicable models for
+agile transformation, at least not in a literature review made based on the
 current agile transformation publications. This is because of two things.
@@
 It would require too big of an interpretation to combine all the various cases
-into one theoretical model. We have chosen to present transformations using an
-elementary two-axial classification of transformations, and listings of
-individual phenomena that typically occur in a transformation.
+into one theoretical model. For these reasons we chose to present
+transformations using an elementary two-axial classification of transformations,
+and listings of individual phenomena that typically occur in a transformation.
+
+% Transformation and scaling up
+
+Transformation and scaling up are related concepts. Although this research
+focuses on transformation process only, it can not be completely separated from
+discussion on scaling up, or on large-scale agile implementation in general.
+Examples on where scaling up can be seen as the transformation itself are
+organizations which start with a successful pilot project, after which agile
+methods are introduced step wise to the rest of the organization.
+
+% ------------------------------
+
+% Comparison with literature
+
+\citet{Misra2010} states that an agile adoption requires a change in mindset.
+--> What evidence on this was there?
+
+The higher level of coordination required in a large organization compared to
+ideal small agile organizations was showing.
 
 * Vertailu kirjallisuuteen
+
+* What is similar in organizations of all scales?
+
+* What is special in large scale?
+
+* Customization of practices was seen as necessary, as large scale has demands
+  over the original agile ideas
+
+* It can be argued that any development effort with hundreds of teams will be
+  difficult to coordinate
+
+% Related challenges and success factors
 
@@
 
-* What is special in large scale?
-
-* Customization of practices was seen as necessary, as large scale has demands
-  over the original agile ideas
-
-* What is similar in organizations of all scales?
-
-* Transformation and scaling up are connected concepts. Although this research
-  focuses on transformation process only, it can not be completely separated
-  from discussion on scaling up, or on large-scale agile implementation in
-  general.
+* Have flexibility for different teams in the agile model (agile found
+  hard [P41]) vs. conformity to a single model.
+
+% ?
 
@@
 
-* It can be argued that any development effort with hundreds of teams will be
-  difficult to coordinate
+% Defence against criticism for weak evidence in primary studies
+
+* The primary studies were quite openly pro-agile -- how large a factor of bias
+  is this?
 
@@
   talked about it must be of some significance.
-
-* Have flexibility for different teams in the agile model (agile found
-  hard [P41]) vs. conformity to a single model.
 
@@
 
-* System development life cycle ?? -- There are inevitably other areas than
-  development of increments, even in an agile process
-
-* The primary studies were quite openly pro-agile -- how large a factor of bias
-  is this?
-
-* How do transformations fit within Kotter's 8 step process?
-
-* A qualitative interpretation requires much text. Therefore the results got
-  long. See ``Meta-ethnography''.
+% Different organizations
 
@@
   the beginning of each project, and an organization wide adoption.
+
+% Other ?
+
+* System development life cycle ?? -- There are inevitably other areas than
+  development of increments, even in an agile process
+
+* How do transformations fit within Kotter's 8 step process?
+
+* A qualitative interpretation requires much text. Therefore the results got
+  long. See ``Meta-ethnography''.
 

### 2014-08-11 journal/content.tex
Introduction

@@
 
---> Introductory paragraph
+Software organizations are constantly challenged to function more efficiently,
+most often due to the fast evolving software marketplace. Agile methods have
+been developed to improve efficiency of software development, but are originally
+designed for use in projects of limited size. The benefits of agile are pursued
+by large organizations
+
+Despite the
 
@@
    base.
---> There is interest in large scale [ref XP13 workshop] \cite{Dingsoyr2013a}
-
---> Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
+
+--> Why is large-scale software development different from smaller scale?
+
+A number of studies on large scale use of agile exist. However, the studies
+present either single cases, do not provide a research method, or does not
+provide insights on the adoption aspect of large scale agile development.
+
+* There is interest in large scale [ref XP13 workshop] \cite{Dingsoyr2013a}
+
+* Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
     Practitioners," IEEE Software, pp. 8-9, 2010.
 
---> Why is large-scale software development different from smaller scale?
-
---> This literature review is intended for researchers as a theoretical
-   and proven foundation for the real problems that organizations face in
-   transformations.
+This literature review is intended for researchers as a theoretical and proven
+foundation for the real problems that organizations face in transformations.
 This is also intended to be directly usable by practitioners, to serve as a
-   base of general advice with strong motivation (although every organization
-   has unique needs).
+base of general advice with strong motivation.
+%(although every organization has unique needs)
 
@@
 
-Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
-Perplexed: Addison-Wesley, 2003.
-Boehm, B., "Get ready for agile methods, with care," IEEE Computer, vol. 35,
-pp. 64 - 69, 2002.
---> Size requires more discipline?
+% Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
+% Perplexed: Addison-Wesley, 2003.
+% Boehm, B., "Get ready for agile methods, with care," IEEE Computer, vol. 35,
+% pp. 64 - 69, 2002.
+% --> Size requires more discipline?
+
+This work is organized as follows. In the second chapter we present agile
+methods in general, and previous research on adopting agile methods on large
+scale. The third chapter discussed the research method used in this work.
+The following chapter presents the findings. The final chapter discusses the
+results and concludes this work.
 

### 2014-08-11 journal/content.tex
Fixed citation tags

@@
 
-* There is interest in large scale [ref XP13 workshop] \cite{Dingsoyr2013a}
+* There is interest in large scale [ref XP13 workshop] \citep{Dingsoyr2013a}
 
@@
 in agile development.
-\cite{Highsmith2001,Cockburn2001}
+\citep{Highsmith2001,Cockburn2001}
 
@@
 that accommodating change may be a factor in both success and failure
-\cite{Boehm2002}. It has been shown that agile methods have improved
+\citep{Boehm2002}. It has been shown that agile methods have improved
 satisfaction of both customers and developers, but on the other hand there is
 evidence that agile methods may not suite well for large undertakings
-\cite{Dyba2009}. As a solution \citet{Boehm2002} has suggested that each
+\citep{Dyba2009}. As a solution \citet{Boehm2002} has suggested that each
 organization should find its own balance of agile and plan driven methods.
@@
 Two of the most popular agile methods at the time being are Extreme Programming
-(XP) and Scrum \cite{Hamed2013}. Scrum is a method which focuses on the project
-management viewpoint of agile development \cite{Schwaber2002}. The Scrum method
+(XP) and Scrum \citep{Hamed2013}. Scrum is a method which focuses on the project
+management viewpoint of agile development \citep{Schwaber2002}. The Scrum method
 focuses on time-boxing, continuous tracking of project progress, and customer
@@
 are time boxes of 1 to 4 weeks, aimed for producing a finished increment of
-working software \cite{Schwaber2002}. The XP development method is a collection
-of practices for enabling efficient incremental development \cite{Beck1999}.
+working software \citep{Schwaber2002}. The XP development method is a collection
+of practices for enabling efficient incremental development \citep{Beck1999}.
 Some of the key practices of XP are continuous feedback on development, test
 driven development, close customer involvement, pair programming, and continuous
-integration (CI) \cite{Beck1999}.
+integration (CI) \citep{Beck1999}.
 
@@
 Core principles of lean software development are optimizing the whole, removal
-of waste, building quality in, and continuous learning. \cite{Poppendieck2012}
+of waste, building quality in, and continuous learning. \citep{Poppendieck2012}
 
@@
 Introducing agile methods in large organizations is more difficult than it is in
-small organizations \cite{Dyba2008}. The difficulty is partly related to size
+small organizations \citep{Dyba2008}. The difficulty is partly related to size
 bringing higher organizational inertia which slows down organizational change
-\cite{Livermore2008}. Agile development is not founded on the use of individual
+\citep{Livermore2008}. Agile development is not founded on the use of individual
 tools or practices, but rather on a holistic way of thinking. Adopting agile
-often requires change of the entire organizational culture \cite{Misra2010}.
+often requires change of the entire organizational culture \citep{Misra2010}.
 
@@
 organizations have more dependencies between projects and teams. This increases
-the need for formal documentation and thus reduces agility \cite{Lindvall2004}.
+the need for formal documentation and thus reduces agility \citep{Lindvall2004}.
 In addition to inter-team coordination development teams must interact with
@@
 HR may demand individuals to have strictly specified roles in projects
-\cite{Boehm2005}, or a change control board may inhibit the use of continuous
-integration or refactoring \cite{Lindvall2004}. All units affected by the agile
+\citep{Boehm2005}, or a change control board may inhibit the use of continuous
+integration or refactoring \citep{Lindvall2004}. All units affected by the agile
 transformation need to be informed and consulted, and the agile process must be
-adjusted according to their needs \cite{Lindvall2004, Cohn2003, Boehm2005}.
+adjusted according to their needs \citep{Lindvall2004, Cohn2003, Boehm2005}.
 
@@
 challenge is that management must move away from life-cycle models and towards
-iterative and feature centric models \cite{Nerur2005}. The focus must be shifted
-from a large scale scope to shorter term project planning \cite{Misra2010}, as
+iterative and feature centric models \citep{Nerur2005}. The focus must be shifted
+from a large scale scope to shorter term project planning \citep{Misra2010}, as
 agile methods emphasize that planning is only meaningful for the near future
-\cite{Cohn2003}. However, the lack of planning can be a concern as business and
+\citep{Cohn2003}. However, the lack of planning can be a concern as business and
 customer relationships often build on long term roadmapping. Enabling operation
 with shorter term planning requires educating stakeholders and reviewing
-contracting practices \cite{Boehm2005}.
+contracting practices \citep{Boehm2005}.
 
@@
 The definition of large scale was recently discussed in a workshop at XP2013
-\cite{Dingsoyr2013a}. Proceeding the workshop, Dingsøyr et al.
-compiled a description of the definition of large scale \cite{Dingsoyr2013b}.
+\citep{Dingsoyr2013a}. Proceeding the workshop, Dingsøyr et al.
+compiled a description of the definition of large scale \citep{Dingsoyr2013b}.
 The most common view in the XP2013 workshop was that the involvement of many
-people is the main factor in large scale \cite{Dingsoyr2013b}.
+people is the main factor in large scale \citep{Dingsoyr2013b}.
 
 Dingsøyr et al. conducted a brief literature search to identify what definitions
-on large scale are used \cite{Dingsoyr2013b}. The results showed that large
+on large scale are used \citep{Dingsoyr2013b}. The results showed that large
 scale had been regarded in terms of size in persons or teams, project budget,
-code base size, and project duration. As examples from the results, Paasivaara
-et al. \cite{Paasivaara2008} referred to an organization of 40 people and 7
-teams. Berger and Beynon-Davies \cite{Berger2009} present the project cost of
-over £10 million as a factor of large scale, but also highlight a team size of
-over 50 people. A product with a code base size of over 5 million lines of code
-was considered to represent a large scale development effort by Petersen and
-Wohlin \cite{Petersen2010}. A project time of 2 years was considered as a factor
-of large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a project
-scope of 60-80 features.
+code base size, and project duration. As examples from the results,
+\citet{Paasivaara2008} referred to an organization of 40 people and 7 teams.
+\citet{Berger2009} present the project cost of over £10 million as a factor of
+large scale, but also highlight a team size of over 50 people. A product with a
+code base size of over 5 million lines of code was considered to represent a
+large scale development effort by \citet{Petersen2010}. A project time of 2
+years was considered as a factor of large scale by \citet{Bjarnason2011}, as
+well as a project scope of 60-80 features.
 
@@
 of these definitions referred to the number of people involved.
-In early work on agile Fowler \cite{Fowler2000} regards the Crystal methodology
-being suitable for up to 50 people, thus setting a boundary beyond which agile
-methods would be expected to need special considerations. Also Williams and
-Cockburn \cite{Williams2003} present 50 people as a limit which practitioners
-and researchers has seen as the largest organization suitable for agile.
-Elshamy and Elssamadisy \cite{Elshamy2006} considered a development project
-large if it had a staff between 50 and 100 people, including all project
-personnel. Finally, the largest numbers were presented by Moore and Spens
-\cite{Moore2008} who define large scale in the scope of their research as 300
-people across 3 sites.
+In early work on agile \citet{Fowler2000} regards the Crystal methodology being
+suitable for up to 50 people, thus setting a boundary beyond which agile methods
+would be expected to need special considerations. Also \citet{Williams2003}
+present 50 people as a limit which practitioners and researchers has seen as the
+largest organization suitable for agile.
+\citet{Elshamy2006} considered a development project large if it had a staff
+between 50 and 100 people, including all project personnel. Finally, the largest
+numbers were presented by \citet{Moore2008} who define large scale in the scope
+of their research as 300 people across 3 sites.
 
@@
 
-The research was conducted as an application of Kitchenham's
-\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
-primary studies was done using first a keyword based database search to identify
+The research was conducted as an application of the guidelines for systematic
+literature review presented by \citet{Kitchenham2007}. Selection of primary
+studies was done using first a keyword based database search to identify
 potentially relevant sources, and then manually filtering the search result. The
@@
 organizational transformation was aimed to introduce agile methods, which
-excluded other development methods than agile \cite{Sagesser2013}, and use of
-agile methods in other contexts than software development \cite{Hodgkins2007}.
+excluded other development methods than agile \citep{Sagesser2013}, and use of
+agile methods in other contexts than software development \citep{Hodgkins2007}.
 
@@
 how the transformations happen. Examples on excluded topics are comparing the
-original and agile development methods \cite{Petersen2010}, discussing use of
+original and agile development methods \citep{Petersen2010}, discussing use of
 agile in a large organization but not describing how the new methods were
-introduced \cite{Mishra2011}, and merely presenting agile tools in large-scale
-use \cite{Kim2012}. These sources were excluded as they did not provide insights
+introduced \citep{Mishra2011}, and merely presenting agile tools in large-scale
+use \citep{Kim2012}. These sources were excluded as they did not provide insights
 on organizational transformation.
@@
 However, we excluded cases where an initially small organization was scaled up
-\cite{Maranzato2012}, and discussions focusing on processes or tools without
-describing organizational change \cite{Lyon2008}.
+\citep{Maranzato2012}, and discussions focusing on processes or tools without
+describing organizational change \citep{Lyon2008}.
 % Also [Greening 2010] on process centered viewpoint without transformation
@@
 for large scale. In some cases the source presented very vague indicators on
-size. For instance, we deemed the case presented by Cloke \cite{Cloke2007} to be
+size. For instance, we deemed the case presented by \citet{Cloke2007} to be
 included as there was indications of large-scale considerations, although the
-size remained unclear. The case presented by Miller and Haddad \cite{Miller2012}
-was deemed to be excluded as there was no indications whatsoever on size. A
-borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
-no indication of issues indicating large size, although there were several
-development teams.
+size remained unclear. The case presented by \citet{Miller2012} was deemed to be
+excluded as there was no indications whatsoever on size. A borderline exclude
+was the case presented by \citet{Tudor2006} as there was no indication of issues
+indicating large size, although there were several development teams.
 To complicate matters, some papers talked about ``the team'' in singular when
-referring to the organization \cite{Hodgkins2007}, making it clearly nontrivial
+referring to the organization \citep{Hodgkins2007}, making it clearly nontrivial
 to judge whether an organization was large based on the choice of words of the
@@
 Further examples on exclusion by the \emph{large scale} facet were cases with
-large organizations but only a single team adopting agile \cite{Fulgham2011}.
+large organizations but only a single team adopting agile \citep{Fulgham2011}.
 We considered cases of single teams (although in large organizations) unrelated
@@
 Also piloting cases that resulted only in single teams using agile were excluded
-\cite{Scott2008}. Finally, cases where the organization was growing to large
+\citep{Scott2008}. Finally, cases where the organization was growing to large
 scale, but did not meet the size criteria at the start of the transformation,
@@
 organization focused typically on different aspects. For example, one paper
-would highlight the viewpoint of developers \cite{Fry2007}, and another would
+would highlight the viewpoint of developers \citep{Fry2007}, and another would
 consider the transformation from user experience designers' point of view
-\cite{Federoff2009}.
+\citep{Federoff2009}.
 
@@
 same case to be included. Instead of using the most complete paper as suggested
-by Kitchenham \cite{Kitchenham2007}, we combined the results presented in each
+by \citet{Kitchenham2007}, we combined the results presented in each
 paper and considered the case as a single unit. Keeping in mind the potential
@@
 
-The primary studies were coded using the Atlas TI qualitative data analysis software.
-An integrated deductive and inductive approach was chosen for coding the primary
-studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
-and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
-list of codes was established for contextual information, as presented in Table
+The primary studies were coded using the Atlas TI qualitative data analysis
+software. An integrated deductive and inductive approach was chosen for coding
+the primary studies \citep{Cruzes2011a}. The coding was designed to have a
+contextual part and a findings part as also \citet{Cruzes2011a} suggests. A list
+of codes was established for contextual information, as presented in Table
 \ref{table:contextualcodes}. Codes related to the research questions were chosen
@@
 than using an a priori approach was to avoid the researchers' previous
-assumptions of the research area to affect the choice of codes
+assumptions of the research area to affect the choice of codes.
 
@@
 
-To give direction for code induction seven different code families were
+To give direction for code induction, seven different code families were
 established on beforehand, as presented in Table \ref{table:codefamilies}. A
@@
 Coding was done by only one researcher due to resource constraints and volume of
-source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
+source material. As recommended by \citet{Kitchenham2007} we conducted
 an experiment with the goal to assess the consistency of coding that two
@@
 process driven waterfall'' [P23]. Korhonen [P25] referred to the model presented
-by Royce \cite{Royce1970} when describing the initial state.
+by \citet{Royce1970} when describing the initial state.
 
@@
 
-Transformation can be facilitated by creating mechanisms that force change.
+Transformation can be facilitated by setting up mechanisms that force change.
 Spayd [P49] describes that senior management pushed hard on a mandate to have

### 2014-08-11 journal/content.tex
Yet one reference on "large scale"
Introduction: Relevance of this study

@@
 
-Software organizations are constantly challenged to function more efficiently,
-most often due to the fast evolving software marketplace. Agile methods have
+Software organizations are constantly challenged to function more efficiently.
+In the fast evolving software marketplace  . Agile methods have
 been developed to improve efficiency of software development, but are originally
-designed for use in projects of limited size. The benefits of agile are pursued
-by large organizations
-
-Despite the
-
--- Who will be interested of this?
+designed for use in projects of limited size.
+
+The benefits of agile are pursued by large organizations
+
+Agile methods are more difficult to implement in large organizations \citep{Dyba2008}.
+
+It is an interesting topic, but agile in large scale is still an evolving
+research topic.
 
@@
 
-A number of studies on large scale use of agile exist. However, the studies
-present either single cases, do not provide a research method, or does not
-provide insights on the adoption aspect of large scale agile development.
-
-* There is interest in large scale [ref XP13 workshop] \citep{Dingsoyr2013a}
-
-* Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
-    Practitioners," IEEE Software, pp. 8-9, 2010.
-
-This literature review is intended for researchers as a theoretical and proven
-foundation for the real problems that organizations face in transformations.
-This is also intended to be directly usable by practitioners, to serve as a
-base of general advice with strong motivation.
-%(although every organization has unique needs)
-
-Dingsoyr:
-The nature of software development, being innovative work that is only partially
-compatible with programmed coordination, requires a strong emphasis on personal
-communication [Kraut, R. E. and Streeter, L. A., "Coordination in software
-development," Communications of the ACM, vol. 38, pp. 69-81, 1995]
+% A number of studies on large scale use of agile exist. However, the studies
+% present either single cases, do not provide a research method, or does not
+% provide insights on the adoption aspect of large scale agile development.
+
+This literature review is targeted for both practitioners and researchers. For
+practitioners this work provides a reference of cases describing introduction
+of agile methods, which can be valuable in planning or reflecting on a agile
+adoption process. The systematic review approach can provide practitioners an
+objective view on how agile adoptions typically progress.
+
+Applying agile in large scale has been noted as a current research trend. In
+2010 \citet{Freudenberg2010} conducted a survey at the XP conference,
+highlighting ``agile and large projects'' as the one of the hottest research
+topics. The trend was continued in the 2013 XP conference, where a research
+trends workshop focused specifically on large scale agile development. The
+viewpoint on introduction of agile methods was highlighted in the workshop
+\citep{Dingsoyr2013a}.
+
+Research on agile software development is accumulating and maturing, and has
+provided ground for performing systematic literature reviews
+\citep{Dyba2008,Jalali2012,Senapathi2013,Kaisti2013}. Still, the area of large
+scale agile development has not yet been studied through secondary studies. In
+the mean time numerous studies discussing agile adoption have been published.
+This setting provides an opportunity to aggregate the findings of primary
+studies, and thereby fill a gap in research on large scale agile adoption.
+
+This work is organized as follows. In the second chapter we present agile
+methods in general, and previous research on adopting agile methods on large
+scale. The third chapter discussed the research method used in this work.
+The following chapter presents the findings. The final chapter discusses the
+results and concludes this work.
+
+% Dingsoyr:
+% The nature of software development, being innovative work that is only partially
+% compatible with programmed coordination, requires a strong emphasis on personal
+% communication [Kraut, R. E. and Streeter, L. A., "Coordination in software
+% development," Communications of the ACM, vol. 38, pp. 69-81, 1995]
 
@@
 % --> Size requires more discipline?
-
-This work is organized as follows. In the second chapter we present agile
-methods in general, and previous research on adopting agile methods on large
-scale. The third chapter discussed the research method used in this work.
-The following chapter presents the findings. The final chapter discusses the
-results and concludes this work.
 
@@
 largest organization suitable for agile.
-\citet{Elshamy2006} considered a development project large if it had a staff
-between 50 and 100 people, including all project personnel. Finally, the largest
-numbers were presented by \citet{Moore2008} who define large scale in the scope
-of their research as 300 people across 3 sites.
+\citet{Koehnemann2009} referred to agile projects including up to 50 people as
+small. \citet{Elshamy2006} considered a development project large if it had a
+staff between 50 and 100 people, including all project personnel. Finally, the
+largest numbers were presented by \citet{Moore2008} who define large scale in
+the scope of their research as 300 people across 3 sites.
 

### 2014-08-11 journal/content.tex
Comment on a very close background match
Fixed mistakes with study count numbers

@@
 contracting practices \citep{Boehm2005}.
+
+% Rohunen 2010:
+% Approaches to Agile Adoption in Large Settings: A Comparison of the Results
+% from a Literature Analysis and an Industrial Inventory
 
@@
 the studies was an interview article. The publication forums of the primary
-studies were distributed so that 57 sources were conference proceedings, 4
+studies were distributed so that 47 sources were conference proceedings, 4
 sources were journal articles, and one source was a technical report.
@@
 the method of systematic literature review we analyzed the findings of 52
-studies describing the agile transformation in 40 different organizations. As a
+studies describing the agile transformation in 42 different organizations. As a
 result we presented qualitative findings describing typical reasons to change,

### 2014-08-12 journal/content.tex
Fixing layout
Introduction

@@
 Software organizations are constantly challenged to function more efficiently.
-In the fast evolving software marketplace  . Agile methods have
-been developed to improve efficiency of software development, but are originally
-designed for use in projects of limited size.
-
-The benefits of agile are pursued by large organizations
-
-Agile methods are more difficult to implement in large organizations \citep{Dyba2008}.
-
-It is an interesting topic, but agile in large scale is still an evolving
-research topic.
-
--- What can be learned from this?
-
---> Motivation for research
--- There are no systematic reviews for this
--- A collection of experience reports can be used for creating a theoretical
-   base.
-
---> Why is large-scale software development different from smaller scale?
-
-% A number of studies on large scale use of agile exist. However, the studies
-% present either single cases, do not provide a research method, or does not
-% provide insights on the adoption aspect of large scale agile development.
-
-This literature review is targeted for both practitioners and researchers. For
-practitioners this work provides a reference of cases describing introduction
-of agile methods, which can be valuable in planning or reflecting on a agile
-adoption process. The systematic review approach can provide practitioners an
-objective view on how agile adoptions typically progress.
-
-Applying agile in large scale has been noted as a current research trend. In
-2010 \citet{Freudenberg2010} conducted a survey at the XP conference,
+In the fast evolving software marketplace companies must seek new ways to
+respond to the competition. Agile methods have proven to improve efficiency and
+quality \citep{Livermore2008}, which makes them attractive options when
+considering ways for improving the development process. Agile methods give
+advantage by enabling shorter lead times and stronger focus on customer needs
+\citep{Petersen2010}.
+
+However, agile methods have originally been designed for use in projects of
+limited size \citep{Boehm2005}. It has proven that it may be problematic
+to introduce agile practices in large organizations \citep{Dyba2009}.
+Nevertheless, the benefits of agile are pursued by large organizations.
+% It would also be sociatelly desirable to effectivize large software projects,
+% because large projects often have societal influence.
+
+% --> Why is large-scale software development different from smaller scale?
+% Agile methods are more difficult to implement in large organizations
+% \citep{Dyba2008}.
+Compared to small organizations ideal for agile development, large organizations
+are characterized by the need of additional coordination. A particular problem
+in applying agile to large software organizations is how to handle inter-team
+coordination. Large organizations also tend to have additional organizational
+units that development must interface with, such as human resources.
+Still further, large organizations may cause users and other stakeholders become
+distant from development teams.
+% The full set of requirements of a large system could be too cumbersome for
+% individual teams to manage.
+\citet{Boehm2003b} have suggested that sizable organizations should not attempt
+to become fully agile.
+
+This thesis presents a study on how large software organizations transform their
+way of working towards agile methods. The focus of this study is to explain the
+process of transformation, but we have decided to leave out discussion on
+measuring the success of the transformation. The study has been performed as a
+systematic literature review, drawing the conclusions from previously published
+research. Using the systematic literature review method we aim to provide an
+objective and broad view on the topic.
+
+This study is targeted for both practitioners and researchers. For practitioners
+this we provide a reference of cases describing introduction of agile methods,
+which can be valuable in planning or reflecting on a agile adoption process. The
+systematic literature review approach can provide practitioners an objective
+view on how agile adoptions typically progress.
+
+Applying agile in large scale has been noted as a current research trend. 
+\citep{Freudenberg2010} conducted a survey at the 2010 XP conference,
 highlighting ``agile and large projects'' as the one of the hottest research
@@
 results and concludes this work.
+
+% A number of studies on large scale use of agile exist. However, the studies
+% present either single cases, do not provide a research method, or does not
+% provide insights on the adoption aspect of large scale agile development.
 
@@
 
-\begin{table}[t]
-    \centering
-    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
-                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
-        \toprule
-        Code family          &  Description
-        \\
-        \midrule
-
-        RQ1: Reason to change &
-        Reasons to start the transformation (demand for faster delivery). \\
-
-		RQ2: Transformation process &
-		Statements that describe the transformation process (top-down, big bang,
-		step wise). \\
-
-		RQ3: Challenges &
-		Statements that present challenges in the transformation (change resistance).
-		\\
-
-		RQ3: Success factors &
-		Statements that present success factors in the transformation (management
-		support).
-		\\
-
-		Investing in change  &
-		Factors that present how the organization is investing in the
-		transformation (training, consultants, tools). \\
-
-		Practices &
-		Distinguishable practices that are used or established during transformation.
-		Factors presented as neutral relating to successes and challenges.
-		(coaching, piloting, continuous integration, communities of practice) \\
-
-		Contextual &
-		Contextual codes defined in Table \ref{table:contextualcodes}. \\
-
-        \bottomrule
-    \end{tabular}
-    \caption{Code families defined before start of coding. Examples on
-    potential codes to be inducted in parentheses.}
-    \label{table:codefamilies}
-\end{table}
-
 The inductive coding was done according to the guidelines described below.
@@
 
-The total number of codes and quotations created in the coding process is
-presented in Table \ref{table:codecount}. The total number of quotations is less
-than the sum of quotations in the categories because a single quotation may have
-multiple codes.
+\begin{table}
+    \centering
+    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
+                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
+        \toprule
+        Code family          &  Description
+        \\
+        \midrule
+
+        RQ1: Reason to change &
+        Reasons to start the transformation (demand for faster delivery). \\
+
+        RQ2: Transformation process &
+        Statements that describe the transformation process (top-down, big bang,
+        step wise). \\
+
+        RQ3: Challenges &
+        Statements that present challenges in the transformation (change resistance).
+        \\
+
+        RQ3: Success factors &
+        Statements that present success factors in the transformation (management
+        support).
+        \\
+
+        Investing in change  &
+        Factors that present how the organization is investing in the
+        transformation (training, consultants, tools). \\
+
+        Practices &
+        Distinguishable practices that are used or established during transformation.
+        Factors presented as neutral relating to successes and challenges.
+        (coaching, piloting, continuous integration, communities of practice) \\
+
+        Contextual &
+        Contextual codes defined in Table \ref{table:contextualcodes}. \\
+
+        \bottomrule
+    \end{tabular}
+    \caption{Code families defined before start of coding. Examples on
+    potential codes to be inducted in parentheses.}
+    \label{table:codefamilies}
+\end{table}
 
@@
 \clearpage
+
+The total number of codes and quotations created in the coding process is
+presented in Table \ref{table:codecount}. The total number of quotations is less
+than the sum of quotations in the categories because a single quotation may have
+multiple codes.
 

### 2014-08-12 journal/content.tex
Improvements on introduction

@@
 However, agile methods have originally been designed for use in projects of
-limited size \citep{Boehm2005}. It has proven that it may be problematic
-to introduce agile practices in large organizations \citep{Dyba2009}.
-Nevertheless, the benefits of agile are pursued by large organizations.
+limited size \citep{Boehm2005}. It has proven that agile methods are more
+difficult to implement in large organizations \citep{Dyba2009}.
+Still, regardless of the challenges that large scale agile adoption poses, the
+benefits of agile are pursued by large organizations.
 % It would also be sociatelly desirable to effectivize large software projects,
@@
 
-% --> Why is large-scale software development different from smaller scale?
 % Agile methods are more difficult to implement in large organizations
@@
 distant from development teams.
+Regardless of these known problems there is an industry trend of large scale
+agile adoptions. An additional challenge is the lack of research on how to
+introduce agile methods to large organizations.
 % The full set of requirements of a large system could be too cumbersome for
 % individual teams to manage.
-\citet{Boehm2003b} have suggested that sizable organizations should not attempt
-to become fully agile.
+% \citet{Boehm2003b} have suggested that sizable organizations should not
+% attempt to become fully agile.
 
@@
 process of transformation, but we have decided to leave out discussion on
-measuring the success of the transformation. The study has been performed as a
+measuring the success of the transformation. The study has been implemented as a
 systematic literature review, drawing the conclusions from previously published
@@
 
-Applying agile in large scale has been noted as a current research trend. 
-\citep{Freudenberg2010} conducted a survey at the 2010 XP conference,
+In research large scale agile development is still a new area, although
+it is gaining interest.
+\citet{Freudenberg2010} conducted a survey at the 2010 XP conference,
 highlighting ``agile and large projects'' as the one of the hottest research
@@
 trends workshop focused specifically on large scale agile development. The
-viewpoint on introduction of agile methods was highlighted in the workshop
-\citep{Dingsoyr2013a}.
+viewpoint on introduction of agile methods was one of the highlighted themes in
+the workshop \citep{Dingsoyr2013a}.
 
@@
 scale agile development has not yet been studied through secondary studies. In
-the mean time numerous studies discussing agile adoption have been published.
+the mean time numerous reports discussing agile adoption have been published.
 This setting provides an opportunity to aggregate the findings of primary

### 2014-08-12 journal/content.tex
Work on conclusion

@@
 
+% Defence against criticism for weak evidence in primary studies
+
+* The primary studies were quite openly pro-agile -- how large a factor of bias
+  is this?
+
+* Even though most of the primary studies are experience reports and subject to
+  bias, this study can draw valuable conclusions. --> If a matter is being
+  talked about it must be of some significance.
+
+* The factors that are listed are the ones to discuss -- there may be other
+  factors affecting transformation, but this study provides evidence that these
+  are the most important ones. We included only challenges that are discussed in
+  more than studies, which proves that some filtering has been done.
+
 % Transformation and scaling up
@@
 
-* Some success factors and challenges relate
-
-* Some success factors and challenges are ``symmetrical'', so that lack of some
-  factor is a challenge, but availability of it provides success
+It is evident in the results that some success factors and challenges relate.
+Some success factors and challenges are complementary, so that lack of some
+factor is a challenge, but availability of it provides success.
+A particular complementary relationship could be observed in use of training
 
@@
 
-% Defence against criticism for weak evidence in primary studies
-
-* The primary studies were quite openly pro-agile -- how large a factor of bias
-  is this?
-
-* Even though most of the primary studies are experience reports and subject to
-  bias, this study can draw valuable conclusions. --> If a matter is being
-  talked about it must be of some significance.
-
-* The factors that are listed are the ones to discuss -- there may be other
-  factors affecting transformation, but this study provides evidence that these
-  are the most important ones. We included only challenges that are discussed in
-  more than studies, which proves that some filtering has been done.
-
 % Different organizations
-
-* It is difficult to distinguish between team and organization. In most cases
-  when talking about teams it refers to a handful of people familiar to each
-  other. However, in some cases it was evident that an entire department was
-  referred to as a team.
-
-* Projects and organizations. How to distinguish between a separate adoption in
-  the beginning of each project, and an organization wide adoption.
+% 
+% * A recommendation for further research
+%
+% * It is difficult to distinguish between team and organization. In most cases
+%   when talking about teams it refers to a handful of people familiar to each
+%   other. However, in some cases it was evident that an entire department was
+%   referred to as a team.
+% 
+% * Projects and organizations. How to distinguish between a separate adoption in
+%   the beginning of each project, and an organization wide adoption.
 
 % Other ?
-
-* System development life cycle ?? -- There are inevitably other areas than
-  development of increments, even in an agile process
-
-* How do transformations fit within Kotter's 8 step process?
-
-* A qualitative interpretation requires much text. Therefore the results got
-  long. See ``Meta-ethnography''.
+% 
+% * System development life cycle ?? -- There are inevitably other areas than
+%   development of increments, even in an agile process
+% 
+% * How do transformations fit within Kotter's 8 step process?
+% 
+% * A qualitative interpretation requires much text. Therefore the results got
+%   long. See ``Meta-ethnography''.
 

### 2014-08-13 journal/content.tex
Work on discussion

@@
 % Defence against criticism for weak evidence in primary studies
-
-* The primary studies were quite openly pro-agile -- how large a factor of bias
-  is this?
-
-* Even though most of the primary studies are experience reports and subject to
-  bias, this study can draw valuable conclusions. --> If a matter is being
-  talked about it must be of some significance.
-
-* The factors that are listed are the ones to discuss -- there may be other
-  factors affecting transformation, but this study provides evidence that these
-  are the most important ones. We included only challenges that are discussed in
-  more than studies, which proves that some filtering has been done.
-
-% Transformation and scaling up
+\subsection{Possible weakness of evidence due to primary study quality}
+
+The strength of evidence of this research may be claimed to be weakened because
+of the fact that primary studies were not ranked based on study quality. Primary
+studies without proper research methods may bring bias to our results. For
+instance, many primary studies were openly pro-agile, without giving a solid
+motivation for the standpoint. Although most of the primary studies are
+subjective experience reports, we still think this study can draw accurate and
+valuable conclusions.
+
+Even if the primary studies do not include rigorous research methods, their
+authors have chosen to bring up matters that they consider important. It can be
+argued that if a matter is being talked about it must be of some significance.
+We believe that the viewpoints on agile transformation that have been elicitated
+in this research are the ones to discuss.
+In order to strengthen the evidence, we included only viewpoints that are
+discussed in several studies.
+There may be other factors affecting transformation, but this study provides
+evidence that these are the most important ones.
+
+\subsection{Similarity of transformation and scaling up}
 

### 2014-08-13 journal/content.tex
Discussion

@@
 In this chapter we discuss the results and provide a conclusion on this work.
-First the results are discussed, including highlighting of most important
-findings and comparison to previous work. Then we present limitations and
+First the results are discussed, including highlighting some findings and
+comparison to previous work. Then we present an appraisal of limitations and
 threats to the validity of this work. This chapter ends with a concluding
@@
 
-Grounded theory may not be able to provide a generally applicable models for
-agile transformation, at least not in a literature review made based on the
-current agile transformation publications. This is because of two things.
-First, the organizations seem to differ too much from each other. Secondly, the
-published studies focus too much on isolated phenomena within the larger
-transformation.
-It would require too big of an interpretation to combine all the various cases
-into one theoretical model. For these reasons we chose to present
-transformations using an elementary two-axial classification of transformations,
-and listings of individual phenomena that typically occur in a transformation.
+The discussion of the results is divided into several sections, each bringing up
+a particular viewpoint. The first part of this section focuses on interpretation
+of results and comparison to the background literature. The remainder of this
+section discusses problems that were encountered during the research process.
+
+\subsection{Related challenges and success factors}
+
+It is evident in the results that some success factors and challenges relate.
+Some success factors and challenges are complementary, so that lack of some
+factor is a challenge, but availability of it provides success.
+A particular complementary relationship was apparent in use of training
+and coaching. The lack of training was seen as a challenge, and similarly
+investments in training were reported as a success factor.
+
+Another complementary relationship could be observed in discussion on
+requirements management. Implementing effective requirements breakdown and
+long term requirements planning were seen as challenges. The complementary
+success factor was to pay close attention on how the Product Owner role 
+was implemented.
+
+Yet another complementary relationship was observed in that implementing the
+agile model was found to be difficult, but choosing the the agile methods
+carefully was seen as a success factor. Implementation difficulties were due to
+misconceptions on agile and lack of guidance on how to proceed during the agile
+transformation. Complementing these challenges primary studies emphasized
+using care in choosing the agile model to pursue.
+
+% * Have flexibility for different teams in the agile model (agile found
+%   hard [P41]) vs. conformity to a single model.
+
+The significance of the complementary challenges and success factors can be
+interpreted in a few ways. The simplest interpretation is that success factors
+in transformation that have complementing challenges have a higher significance.
+This means that investment in training and coaching should be considered as a
+significant success factor. Similarly, paying attention to requirements
+management and especially the Product Owner role can be considered as
+significant success factors. The relationship between implementation challenges
+with the agile model and the success factors related to careful choice of the
+model may not indicate greater significance. Instead the connection between
+these challenges and success factors can be considered to indicate that they
+simply should be studied together.
+
+\subsection{Comparing results with existing literature}
+
+Implementing an agile model is more about a change in mindset than application
+of a set of new practices \citep{Misra2010}. Although this is an important point
+in discussion on agile transformations, it was discussed quite little in the
+primary studies. However, the use of coaching was a particular practice that was
+aimed towards changing the mindset of people. A particular viewpoint that was
+brought up when presenting coaching as a success factor was the importance of
+how coaching could create a change in mindset. Coaching was attributed as a way
+to teach by practice how to apply an agile mindset in development.
+
+\subsection{Similarity of transformation and scaling up}
+
+A particular complication in this research was distinguishing between the
+concepts of transformation and scaling up. Distinguishing between these concepts
+was an important factor due to the systematic nature of this research. Although
+this research focuses on transformation process only, it can not be completely
+separated from discussion on scaling up, or on large-scale agile implementation
+in general.
+Examples on where scaling up can be seen as the transformation itself are
+organizations which start with a successful pilot project, after which agile
+methods are introduced step wise to the rest of the organization.
 
@@
 argued that if a matter is being talked about it must be of some significance.
-We believe that the viewpoints on agile transformation that have been elicitated
+We believe that the viewpoints on agile transformation that have been elicited
 in this research are the ones to discuss.
@@
 
-\subsection{Similarity of transformation and scaling up}
-
-Transformation and scaling up are related concepts. Although this research
-focuses on transformation process only, it can not be completely separated from
-discussion on scaling up, or on large-scale agile implementation in general.
-Examples on where scaling up can be seen as the transformation itself are
-organizations which start with a successful pilot project, after which agile
-methods are introduced step wise to the rest of the organization.
-
-% ------------------------------
-
-% Comparison with literature
-
-\citet{Misra2010} states that an agile adoption requires a change in mindset.
---> What evidence on this was there?
-
-The higher level of coordination required in a large organization compared to
-ideal small agile organizations was showing.
-
-* Vertailu kirjallisuuteen
-
-* What is similar in organizations of all scales?
-
-* What is special in large scale?
-
-* Customization of practices was seen as necessary, as large scale has demands
-  over the original agile ideas
-
-* It can be argued that any development effort with hundreds of teams will be
-  difficult to coordinate
-
-% Related challenges and success factors
-
-It is evident in the results that some success factors and challenges relate.
-Some success factors and challenges are complementary, so that lack of some
-factor is a challenge, but availability of it provides success.
-A particular complementary relationship could be observed in use of training
-
-* Success factors and challenges may be complements, and may be presented as
-  complements
-
-* Have flexibility for different teams in the agile model (agile found
-  hard [P41]) vs. conformity to a single model.
-
-% ?
-
-* Transformation of an organization to use agile can be seen as transforming
-  also the agile method to suit the organization.
-  --> A large-scale agile model may require tailoring and trial over time
+\subsection{Problems using grounded theory for analyzing distinctive organizations}
+
+Grounded theory may not be able to provide a generally applicable models for
+agile transformation, at least not in a literature review made based on the
+current agile transformation publications. This is because of two things.
+First, the organizations seem to differ too much from each other. Secondly, the
+published studies focus too much on isolated phenomena within the larger
+transformation.
+In our case it would require too big of an interpretation to combine all the
+various cases into a theoretical model. For these reasons we chose to present
+transformations using an elementary two-axial classification of transformations,
+listings of individual phenomena, and other typical factors in transforming
+organizations.
+
+% * What is similar in organizations of all scales?
+% 
+% * What is special in large scale?
+% 
+% The higher level of coordination required in a large organization compared to
+% ideal small agile organizations was showing.
+% 
+% * It can be argued that any development effort with hundreds of teams will be
+%   difficult to coordinate
+% 
+% * Customization of practices was seen as necessary, as large scale has demands
+%   over the original agile ideas
+
+% ???
+% 
+% * Transformation of an organization to use agile can be seen as transforming
+%   also the agile method to suit the organization.
+%   --> A large-scale agile model may require tailoring and trial over time
 

### 2014-08-13 journal/content.tex
Discussion

@@
 \subsection{Integrating other functions than development to agile}
+\label{section_other_functions}
 
@@
 
-\subsection{Comparing results with existing literature}
+\subsection{Comparison with existing literature}
 
@@
 to teach by practice how to apply an agile mindset in development.
+
+A challenge realting particularly to large scale is that development teams need
+to interface with various other organizational functions \citep{Lindvall2004}.
+Particular function mentioned are other development teams and requirements
+management. \citet{Boehm2005} and also \citet{Cohn2003} mentions human resources
+as an additional group.
+Our research reflects very closely on the challenges development has in
+interfacing with these groups. Our research identified still a number of other
+groups that were not mentioned in existing literature (see section
+\ref{section_other_functions}).
+
+\citet{Cohn2003} brings up a number of challenges that our research also
+discovered.
+
+As a conclusion on the comparison between this study and previous literature it
+can be said that there is a very high level of correspondence. However, the
+existing publications discussing large scale agile adoption provides very little
+grounds for the claims presented. Our research can be considered to groundedly
+confirm the claims presented in existing literature.
+
+The majority of transformation related topics presented in existing literature
+is found in this research. However, this does not go the other way, as each of
+the sources discussed as background is missing some area that this study brings
+up.
+It should also be noted that the thoroughness of discussion on various topics
+differ between this study and the background literature. Therefore each study
+will give a different impression on the relative importance of the topics
+discussed. In this research the number of sources referring to a particular
+topic may be used as an estimate of relative importance.
 

### 2014-08-13 journal/content.tex
Finishing up on discussion

@@
 \subsection{Agile is difficult to implement}
+\label{section_difficult_to_implement}
 
@@
 
-A challenge realting particularly to large scale is that development teams need
+A challenge relating particularly to large scale is that development teams need
 to interface with various other organizational functions \citep{Lindvall2004}.
@@
 \citet{Cohn2003} brings up a number of challenges that our research also
-discovered.
+discovered, including change resistance, misunderstanding the agile model, and
+overzealous teams. Although the topics are the same, the discussion of
+\citeauthor{Cohn2003} differ from the results of our research. For instance,
+they present the challenge of misunderstanding agile development merely as fear
+of micromanagement, whereas our research provides a rich set of examples (see
+section \ref{section_difficult_to_implement}).
+The discussion on overzealous teams also seems to get disproportionate attention
+compared to the groundedness observed in our research.
 
@@
 existing publications discussing large scale agile adoption provides very little
-grounds for the claims presented. Our research can be considered to groundedly
-confirm the claims presented in existing literature.
+grounds for the claims presented. Our research can be considered to confirm the
+claims presented in existing literature in a grounded way.
 
@@
 up.
-It should also be noted that the thoroughness of discussion on various topics
+It should also be noted that the extensiveness of discussion on various topics
 differ between this study and the background literature. Therefore each study
@@
 discussed. In this research the number of sources referring to a particular
-topic may be used as an estimate of relative importance.
+topic may be used as a grounded estimate on relative importance.
 
@@
 
+\newpage
 \section{Conclusion}

### 2014-08-13 journal/content.tex
Bias towards successful transformations

@@
 % Defence against criticism for weak evidence in primary studies
-\subsection{Possible weakness of evidence due to primary study quality}
+\subsection{Possible weakness of evidence and bias due to primary study quality}
 
@@
 instance, many primary studies were openly pro-agile, without giving a solid
-motivation for the standpoint. Although most of the primary studies are
-subjective experience reports, we still think this study can draw accurate and
-valuable conclusions.
+motivation for the standpoint. 
+Although most of the primary studies are subjective experience reports, we still
+think this study can draw accurate and valuable conclusions.
+
+A tendency to publish only positive results is another particular problem for
+this literature review. In the collection of primary studies it can be observed
+that authors typically considered the transformation being successful.
+It can be argued that this apparent tendency to publish only cases that have
+experienced success will bias the results.
+However, even if the transformations were considered successful in general, most
+studies brought up different challenges that had been experienced during the
+transformation.
 

### 2014-08-13 journal/content.tex
Statements on contribution

@@
 eliminate the potential bias we would suggest manually searching the entire
-archive of \emph{Agile Conference} and \emph{Agile Processes in Software
-Engineering and Extreme Programming} for relevant studies. These two conferences
-provided more than two thirds of the primary sources.
+archive of \emph{Agile Conference} and \emph{International conference on agile
+software development} (also known as XP conference) for relevant studies.
+% XP = Agile Processes in Software Engineering and Extreme Programming
+These two conferences provided more than two thirds of the primary sources.
 
@@
 
-\newpage
 \section{Conclusion}
+
+This work makes a unique contribution on a topical area in research on agile
+software development. In the past decade there has been numerous publications
+discussing use of agile in large scale organizations. However, the process of
+adopting agile in large organizations has not been studied extensively before.
+Numerous experience reports and case studies discussing large agile adoptions
+exist, but there is a lack of aggregating work providing an overview of this
+subject area.
+This work presents a unique compilation and summarizing analysis of existing
+publications on large scale agile adoptions.
+
+The systematic literature review method creates solid evidence that can act as
+background for further research.
+For practitioners the results provide an extensive collection of viewpoints to
+consider in an adoption process, and the primary sources act as a listing of
+cases to compare against.
 
@@
 the method of systematic literature review we analyzed the findings of 52
-studies describing the agile transformation in 42 different organizations. As a
+studies describing 42 different organizations adopting agile methods. As a
 result we presented qualitative findings describing typical reasons to change,

### 2014-08-13 journal/content.tex
Using Times New Roman as font; layout fixes
Fixed difficult sentences

@@
 
-\begin{table}[b]
+\begin{table}
     \centering
@@
 
-\begin{table}[t]
+\begin{table}[h]
     \centering
@@
 
-\begin{table}[!b]
+\begin{table}
     \centering
@@
 Another complementary relationship could be observed in discussion on
-requirements management. Implementing effective requirements breakdown and
-long term requirements planning were seen as challenges. The complementary
-success factor was to pay close attention on how the Product Owner role 
-was implemented.
+requirements management. Reported transformation challenges included
+difficulties in implementing effective requirements breakdown and long term
+requirements planning. The complementary success factor was to pay close
+attention to the Product Owner role. A notable part of discussion on how to
+successfully implement requirements management highlighted the role of the
+Product Owner.
 
@@
 section \ref{section_difficult_to_implement}).
-The discussion on overzealous teams also seems to get disproportionate attention
-compared to the groundedness observed in our research.
+\citeauthor{Cohn2003} also seem to give disproportionate attention on the
+discussion on overzealous teams, compared to the groundedness of the topic
+in our results.
 

### 2014-08-14 journal/content.tex
Corrections

@@
 
-The goal of this work was to review and analyze empirical studies of large
-organizations taking agile software development methods into use. By applying
-the method of systematic literature review we analyzed the findings of 52
-studies describing 42 different organizations adopting agile methods. As a
-result we presented qualitative findings describing typical reasons to change,
-characteristics of transformations, challenges, and success factors.
+The goal of this work was to review and analyze empirical studies and experience
+reports presenting large organizations taking agile software development methods
+into use. By applying the method of systematic literature review we analyzed the
+findings of 52 papers describing 42 different organizations adopting agile
+methods. As a result we presented qualitative findings describing typical
+reasons to change, characteristics of transformations, challenges, and success
+factors.
 
